### 2020-05-10 paper_PLOS/paper_PLOS_draft.Rmd
first draft

---
title: Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Paenungulates
author:
  - name: Juan Manuel Vazquez
    email: juanvazquez@uchicago.edu
    affiliation: University of Chicago
    corresponding: juanvazquez@uchicago.edu
  - name: Vincent J Lynch
    email: vjlynch@buffalo.edu
    affiliation: University at Buffalo, SUNY
address:
  - code: University of Chicago
    address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
  - code: University at Buffalo, SUNY
    address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
bibliography: rbhb_paenungulates.bib
output: 
  workflowr::wflow_html: default
  rticles::plos_article: 
    includes:
      in_header: preamble.tex
  word_document: default
csl: plos.csl
editor_options:
  chunk_output_type: console
---

<!--_Text based on plos sample manuscript, see [http://journals.plos.org/ploscompbiol/s/latex](http://journals.plos.org/ploscompbiol/s/latex)_-->

```{r Libraries and functions, include = F}
knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = T, # fig.width=5, 
                      fig.retina = T)
set.seed(1234)
suppressMessages(suppressWarnings(require(data.table)))
suppressMessages(suppressWarnings(require(tidyverse)))
options(readr.num_columns = 0)
suppressMessages(suppressWarnings(require(ape)))
suppressMessages(suppressWarnings(require(geiger)))
suppressMessages(suppressWarnings(require(nlme)))
suppressMessages(suppressWarnings(require(treeio)))
suppressMessages(suppressWarnings(require(tidytree)))
suppressMessages(suppressWarnings(require(ggtree)))
suppressMessages(suppressWarnings(require(ggimage)))
suppressMessages(suppressWarnings(require(ggpubr)))
suppressMessages(suppressWarnings(require(ggsci)))
suppressMessages(suppressWarnings(require(plotly)))
suppressMessages(suppressWarnings(require(ggplotify)))
suppressMessages(suppressWarnings(require(UpSetR)))
suppressMessages(suppressWarnings(require(ggrepel)))
suppressMessages(suppressWarnings(require(magick)))
suppressMessages(suppressWarnings(require(ggrepel)))
suppressMessages(suppressWarnings(require(gghighlight)))
suppressMessages(suppressWarnings(require(broom.mixed)))
suppressMessages(suppressWarnings(require(purrr)))
suppressMessages(suppressWarnings(require(stargazer)))
source("../code/generalFunctions.R")
```
```{r Data Files, include = F, cache=T}
UP000005640 <- read_tsv("../data/input/UP000005640.withheader.tsv")
gene_list <- read_lines("../output/recBlastDBPrep/AvA_geneList.txt")
gene_list_filtered <- read_lines("../output/recBlastDBPrep/AvA_geneList_filtered.txt")
genomeTable <- read_csv("../output/other/genomeTable.csv")

bsize.tree <- read.beast("../output/AncSizeRec/AnnotatedTreeForFigure.tree") %>% 
  as_tibble %>% 
  mutate(sqrt.rate = branch.length) %>% 
  as.treedata
# st.sqrt <- read.tree("../data/stableTraits/eutheria.sqchange.tree") 
# st.sqrt$node.label <- st.sqrt$node.label %>% str_remove("node")
# st.sqrt %<>% 
#   as_tibble %>% 
#   mutate(sqrt.rate = branch.length) %>% 
#   as.treedata
st.nexus <- read.nexus("../data/stableTraits/eutheria.nexus") %>%
  as_tibble() %>%
  rename(real_label=label) %>%
  mutate(
    name = str_extract(real_label, "[A-Za-z][A-Za-z0-9*.]+'?$"),
    label = str_extract(real_label, "^(\\d)+(?=\\.[:alpha:])|^'(\\d)+(?=\\.[:alpha:])|^(\\d)+$") %>% str_remove("'")
  )
st.nexus[is.na(st.nexus$label),]$label <- st.nexus[is.na(st.nexus$label),]$name
st.nexus %<>% as.treedata()
st.other <- full_join(
  read_tsv("../data/stableTraits/eutheria.ancstates") %>% mutate(Parameter = str_remove(Parameter, "^n(?=\\d)")) %>% rename(label= Parameter), 
  read_tsv("../data/stableTraits/eutheria.brlens") %>% mutate(label=Branch %>% str_split("->") %>% sapply("[",1) %>% str_remove("^n(?=\\d)")), 
  by="label"
  ) %>% 
  mutate(node = Branch %>% str_split("->") %>% sapply("[",1) %>% str_remove("^n(?=\\d)"), parent = Branch %>% str_split("->") %>% sapply("[",2) %>% str_remove("^n(?=\\d)"))
st.other[st.other$label=="0",]$node <- "0"
class(st.other) <- c(class(st.other), "tbl_tree")
st.other %<>% select(parent, node, label, everything()) %>% filter(!is.na(parent)) %>% arrange(parent, node) %>% as.treedata()
st.part <- merge_tree(st.nexus, st.other)
st.full <- merge_tree(st.part, bsize.tree)

ml.treedata <- read.beast("../output/geneLists/maxLikelihood_model_MK+FQ+I+G4-dataType_MORPH-asrMin_0.8/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata_RBB_filtered_dyn/atlantogenata_RBB_filtered_dyn.nexus")

geneCopyTable <- read_csv("../output/geneCopyTable/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata-GeneCopyTable_RBB_filtered_long.csv")
ORA.all <- lineage.pathway.fdr.table("../output/ORA") %>% 
  mutate(
    EnrichRatio = as.numeric(EnrichRatio), FDR=as.numeric(FDR), Size=as.numeric(Size),
    GeneList = Genes %>% str_replace_all(";", "\n") %>% str_trunc(100, "center", ellipsis = "...\n"),
    Class = class.pathway(Pathway)
         ) 
loxAf4.gene_transcript <- fread(
  "../output/RBHBID-to-StringTieTranscriptID/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/loxAfr4_RBB_ID-TranscriptID.tsv", 
  sep = "\t",
  header=F,
  col.names = c("Locus", "details")
  )[, 
    geneID := tstrsplit(details, ";", fixed=TRUE),
    by=Locus][,
      .(geneID = str_extract(geneID, '(?<=gene_id ")[A-Za-z0-9.]+(?=")')),
    by=Locus] %>% 
  unique

triManLat2.gene_transcript <- fread(
  "../output/RBHBID-to-StringTieTranscriptID/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/triManLat2_RBB_ID-TranscriptID.tsv", 
  sep = "\t",
  header=F,
  col.names = c("Locus", "details")
  )[, 
    geneID := tstrsplit(details, ";", fixed=TRUE),
    by=Locus][,
      .(geneID = str_extract(geneID, '(?<=gene_id ")[A-Za-z0-9.]+(?=")')),
    by=Locus] %>% 
  unique

loxAfr4.transcripts <- getTPM.dt("loxAfr4")
triManLat2.transcripts <- getTPM.dt("triManLat2")

genomeTable.atlantogenata <- genomeTable %>% filter(label %in% ml.treedata@phylo$tip.label)

CEGMA.qc <- read_csv("../output/CEGMA/CEGMA_QC.csv") %>%
    separate_rows(Value, sep = ",") %>% 
  mutate(
    Genome = as.factor(Genome), 
    Stat = as.factor(Stat), 
    Value = Value %>% str_trim(side="both") %>% str_split(" ") %>% sapply(., "[", 1),
    Number = Value %>% str_extract("\\d+(\\.\\d+)?$") %>% as.numeric(),
    Type = Value %>% str_extract("^[A-Za-z\\.]+") %>% str_c(" (", ., ")") %>% if_else(is.na(.), "", .),
    NewStat = str_c(Stat, Type)
    ) %>% 
  select(-Value, -Type, -Stat, Genome, Stat=NewStat, Value=Number)
CEGMA.qc[is.na(CEGMA.qc$Value),]$Value <- 0
CEGMA.qc %<>% pivot_wider(id_cols=Genome, names_from = Stat, values_from=Value)
CEGMA.completeness <- read_csv("../output/CEGMA/CEGMA_Completeness.csv", comment = "#", col_names = c("Genome","Group","#Prots","%Completeness","#Total","Average","%Ortho","Missing Proteins")) %>% 
  filter(!is.na(Average))

anage <- read_tsv("../data/AnAge/anage_build14.txt", col_types = list("Weaning (days)"=col_number(), "Weaning weight (g)"=col_number(), "References" = col_character())) %>% 
  select(Genus, Species,"Maximum longevity (yrs)", "Adult weight (g)") %>% 
  unite("label", Genus, Species, sep=".") %>% 
  mutate(lnSize.anage = sapply(
    `Adult weight (g)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  ),
  Lifespan = sapply(
    `Maximum longevity (yrs)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  )
  ) %>% 
  select(label, lnSize.anage, Lifespan)
```
```{r Query List, include=F}
UP000005640.full <- UP000005640 %>% 
  filter(!is.na(Name))
UP000005640.genenames <- unique(UP000005640.full$Name)
query.binTable <- list(AvA = gene_list, UP000005640 = UP000005640.genenames) %>% 
  binTableFromList() %>% 
  left_join(., UP000005640, by=c("Gene"="Name")) %>% 
  mutate(
    "C.ORF" = str_detect(Gene, "^C[0-9XY]{1,2}orf\\d+"),  # No C-Orfs
    "LOC" = str_detect(Gene, "^LOC\\d+"),   # No uncharacterized loci
    "LINC" = str_detect(Gene, "^LINC\\d+"),   # No Long "Non-Coding" RNA
    "HLA" = str_detect(Gene, "^HLA|^HCP5|^CD74"), # No HLAs
    "Histone" = str_detect(Gene, "^H[1-4]|^HIST\\d"), # No Histones
    "ZincFinger" = str_detect(Gene, "^ZNF\\d+"),   # No Zinc Fingers
    "Riboprotein" = str_detect(Gene, "^RP[LS]\\d+"),   # No Riboproteins
    "OlfactoryReceptor" = str_detect(Gene, "^OR\\d+"),   # No Olfactory Receptors
    "Uncharacterized" = str_detect(header, "[uU]ncharacterized"),  # No Uncharacterized proteins
    "Putative" = str_detect(header, "[pP]utative"),  # No putative proteins
    "Viral" = str_detect(header, "[pP]olyprotein|ERVK|HERV"),  # No viral proteins
    "Fragment" = str_detect(header, "[Ff]ragment")  # No fragment proteins
  )
```
```{r Body Size Data: BSize only, include=F}
bsize.innerNodes <- tribble(
  ~"Node",                                                     ~"label",
  MRCA(bsize.tree, "Loxodonta.africana", "Loxodonta.cyclotis"), "Loxodontini",
  MRCA(bsize.tree, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis"), "Loxodona",
  MRCA(bsize.tree, "Loxodonta.africana", "Elephas.maximus"), "Elephantidae",
  MRCA(bsize.tree, "Elephas.maximus", "Mammuthus.primigenius"), "Elephantina",
  MRCA(bsize.tree, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
  MRCA(bsize.tree, "Loxodonta.africana", "Mammut.americanum"), "Proboscidae",
  MRCA(bsize.tree, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
  MRCA(bsize.tree, "Loxodonta.africana", "Procavia.capensis"), "Paenungulata",
  MRCA(bsize.tree, "Loxodonta.africana", "Orycteropus.afer"), "Pseudoungulata",
  MRCA(bsize.tree, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
  MRCA(bsize.tree, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosorcida",
  MRCA(bsize.tree, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
  MRCA(bsize.tree, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), "Xenarthra",
  MRCA(bsize.tree, "Dasypus.novemcinctus", "Loxodonta.africana"), "Atlantogenata"
)
bsize.tree %<>% as_tibble()

bsize.tree[which(bsize.tree$node %in% bsize.innerNodes$Node),]$label <- sapply(bsize.tree[which(bsize.tree$node %in% bsize.innerNodes$Node),]$node, function(n, tb=bsize.innerNodes){tb %>% filter(Node==n) %>% pull(label) %>% unique})

bsize.ancAtlantogenata <- bsize.innerNodes %>% filter(label=="Atlantogenata") %>% pull(Node)

bsize.tree <- full_join(bsize.tree %>% as.treedata(), genomeTable, by = "label")

bsize.Atlantogenata <- bsize.tree %>% 
  # as.treedata() %>% 
  tree_subset(., node=bsize.ancAtlantogenata, levels_back=0)

bsize.Atlantogenata.table <- bsize.Atlantogenata  %>%
  as_tibble()

size.scale <- bsize.tree %>% as_tibble %>% pull(lnSize) %>% unique %>% as.numeric() %>% sort
size.scale.min <- size.scale[1]
size.scale.max <- size.scale[length(size.scale)]

bsize.Atlantogenata.nodes <- list(
  "Loxodonta.africana"= bsize.Atlantogenata.table %>% filter(label=="Loxodonta africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= bsize.Atlantogenata.table %>% filter(label=="Trichechus manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= bsize.Atlantogenata.table %>% filter(label=="Procavia capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= bsize.Atlantogenata.table %>% filter(label=="Orycteropus afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= bsize.Atlantogenata.table %>% filter(label=="Choloepus hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = bsize.Atlantogenata.table %>% filter(label=="Dasypus novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = bsize.Atlantogenata.table %>% filter(label=="Chrysochloris asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = bsize.Atlantogenata.table %>% filter(label=="Echinops telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = bsize.Atlantogenata.table %>% filter(label=="Elephantulus edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = bsize.Atlantogenata.table %>% filter(label=="Loxodonta cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = bsize.Atlantogenata.table %>% filter(label=="Palaeoloxodon antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = bsize.Atlantogenata.table %>% filter(label=="Elephas maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = bsize.Atlantogenata.table %>% filter(label=="Mammut americanum") %>% pull(node) %>% unique,
  "Loxodontini" = bsize.Atlantogenata.table %>% filter(label=="Loxodontini") %>% pull(node) %>% unique,
  "Loxodona" = bsize.Atlantogenata.table %>% filter(label=="Loxodona") %>% pull(node) %>% unique,
  "Elephantidae" = bsize.Atlantogenata.table %>% filter(label=="Elephantidae") %>% pull(node) %>% unique,
  "Elephantina" = bsize.Atlantogenata.table %>% filter(label=="Elephantina") %>% pull(node) %>% unique,
  "Mammuthus" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
  "Proboscidae" = bsize.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(node) %>% unique,
  "Tethytheria" = bsize.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
  "Paenungulata" = bsize.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(node) %>% unique,
  "Pseudoungulata" = bsize.Atlantogenata.table %>% filter(label=="Pseudoungulata") %>% pull(node) %>% unique,
  "Afrotheria" = bsize.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
  "Afrosorcida" = bsize.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(node) %>% unique,
  "Macroscelidae" = bsize.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
  "Xenarthra" = bsize.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(node) %>% unique,
  "Atlantogenata" = bsize.Atlantogenata.table %>% filter(label=="Atlantogenata") %>% pull(node) %>% unique
)

bsize.atlantogenata.name.node.list <- names(bsize.Atlantogenata.nodes) %>% set_names(bsize.Atlantogenata.nodes)

bsize.Atlantogenata.table %<>% mutate(label = sapply(seq_along(node), function(x, l=bsize.atlantogenata.name.node.list, Label=label, Node=as.character(node)){ifelse(Node[x] %in% names(l), l[[Node[x]]], Label[x])}))

species.highlight <- c("Loxodonta.africana", "Trichechus.manatus", "Procavia.capensis", "Orycteropus.afer", "Dasypus.novemcinctus", "Choloepus.hoffmanni", "Chrysochloris.asiatica", "Echinops.telfairi", "Elephantulus.edwardii") %>% 
  set_names(.,.)

bsize.Atlantogenata.table %<>% mutate(highlight = sapply(seq_along(label), function(x, l=species.highlight, Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)}))

bsize.Atlantogenata <- bsize.Atlantogenata.table %>% as.treedata()
```

```{r Body Size Data: All, include=F}
# Tree
st.innerNodes <- tribble(
  ~"Node",                                                     ~"label",
  MRCA(st.full, "Loxodonta.africana", "Loxodonta.cyclotis"), "Loxodontini",
  MRCA(st.full, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis"), "Loxodona",
  MRCA(st.full, "Loxodonta.africana", "Elephas.maximus"), "Elephantidae",
  MRCA(st.full, "Elephas.maximus", "Mammuthus.primigenius"), "Elephantina",
  MRCA(st.full, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
  MRCA(st.full, "Loxodonta.africana", "Mammut.americanum"), "Proboscidae",
  MRCA(st.full, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
  MRCA(st.full, "Loxodonta.africana", "Procavia.capensis"), "Paenungulata",
  MRCA(st.full, "Loxodonta.africana", "Orycteropus.afer"), "Pseudoungulata",
  MRCA(st.full, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
  MRCA(st.full, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosorcida",
  MRCA(st.full, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
  MRCA(st.full, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), "Xenarthra",
  MRCA(st.full, "Dasypus.novemcinctus", "Loxodonta.africana"), "Atlantogenata"
)
st.full %<>% 
  as_tibble() %>% 
  select(-label.y, label=label.x)

st.full[which(st.full$node %in% st.innerNodes$Node),]$label <- sapply(st.full[which(st.full$node %in% st.innerNodes$Node),]$node, function(n, tb=st.innerNodes){tb %>% filter(Node==n) %>% pull(label) %>% unique})
# A bit of manual labor
st.full[which(st.full$label %in% as.character(1859:1876)),]$label <- NA

st.ancAtlantogenata <- st.innerNodes %>% filter(label=="Atlantogenata") %>% pull(Node)

st.tree <- full_join(st.full %>% as.treedata(), genomeTable, by = "label") %>% as_tibble() %>% filter(!is.na(node)) %>% as.treedata()

st.Atlantogenata <- st.tree %>% 
  # as.treedata() %>% 
  tree_subset(., node=st.ancAtlantogenata, levels_back=0) %>%
  as_tibble() %>% 
  select(parent, node, branch.length, label, Rate, Orig_len, Brownian, Median, `95%CI_Low`, `95%CI_High`, sqrt.rate, lnSize, Genome, Species.strict, `Common Name`, Species, Spc.short, BestGenome, phyloPicUID) %>%
  mutate(branch.length=Orig_len) %>% 
  mutate_at(vars(-label, -Genome, -Species.strict, -`Common Name`, -Species, -Spc.short, -BestGenome, -phyloPicUID), . %>% as.numeric %>% round(2)) %>% 
  mutate(label=label %>% str_replace_all("\\.", " ")) %>% 
  as.treedata()

# st.tree %>% as_tibble() %>% filter(node==ancNode.Atlantogenata)

tree.Atlantogenata <- st.Atlantogenata

tree.Atlantogenata.table <- tree.Atlantogenata %>% as_tibble

tree.Atlantogenata.nodes <- list(
  "Loxodonta.africana"= tree.Atlantogenata.table %>% filter(label=="Loxodonta africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= tree.Atlantogenata.table %>% filter(label=="Trichechus manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= tree.Atlantogenata.table %>% filter(label=="Procavia capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= tree.Atlantogenata.table %>% filter(label=="Orycteropus afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= tree.Atlantogenata.table %>% filter(label=="Choloepus hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = tree.Atlantogenata.table %>% filter(label=="Dasypus novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = tree.Atlantogenata.table %>% filter(label=="Chrysochloris asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = tree.Atlantogenata.table %>% filter(label=="Echinops telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = tree.Atlantogenata.table %>% filter(label=="Elephantulus edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = tree.Atlantogenata.table %>% filter(label=="Loxodonta cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = tree.Atlantogenata.table %>% filter(label=="Palaeoloxodon antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = tree.Atlantogenata.table %>% filter(label=="Elephas maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = tree.Atlantogenata.table %>% filter(label=="Mammuthus columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = tree.Atlantogenata.table %>% filter(label=="Mammuthus primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = tree.Atlantogenata.table %>% filter(label=="Mammut americanum") %>% pull(node) %>% unique,
  "Loxodontini" = tree.Atlantogenata.table %>% filter(label=="Loxodontini") %>% pull(node) %>% unique,
  "Loxodona" = tree.Atlantogenata.table %>% filter(label=="Loxodona") %>% pull(node) %>% unique,
  "Elephantidae" = tree.Atlantogenata.table %>% filter(label=="Elephantidae") %>% pull(node) %>% unique,
  "Elephantina" = tree.Atlantogenata.table %>% filter(label=="Elephantina") %>% pull(node) %>% unique,
  "Mammuthus" = tree.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
  "Proboscidae" = tree.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(node) %>% unique,
  "Tethytheria" = tree.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
  "Paenungulata" = tree.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(node) %>% unique,
  "Pseudoungulata" = tree.Atlantogenata.table %>% filter(label=="Pseudoungulata") %>% pull(node) %>% unique,
  "Afrotheria" = tree.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
  "Afrosorcida" = tree.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(node) %>% unique,
  "Macroscelidae" = tree.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
  "Xenarthra" = tree.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(node) %>% unique,
  "Atlantogenata" = tree.Atlantogenata.table %>% filter(label=="Atlantogenata") %>% pull(node) %>% unique
)

atlantogenata.name.node.list <- names(tree.Atlantogenata.nodes) %>% set_names(tree.Atlantogenata.nodes)

tree.Atlantogenata.table %<>% mutate(label = sapply(seq_along(node), function(x, l=atlantogenata.name.node.list, Label=label, Node=as.character(node)){ifelse(Node[x] %in% names(l), l[[Node[x]]], Label[x])}))

species.highlight <- c("Loxodonta africana", "Trichechus manatus", "Procavia capensis", "Orycteropus afer", "Dasypus novemcinctus", "Choloepus hoffmanni", "Chrysochloris asiatica", "Echinops telfairi", "Elephantulus edwardii") %>% 
  set_names(.,.)

tree.Atlantogenata.table %<>% mutate(highlight = sapply(seq_along(label), function(x, l=species.highlight, Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)}))

tree.Atlantogenata <- tree.Atlantogenata.table %>% as.treedata()
```
```{r Body Size Table}
st.core <- st.Atlantogenata %>% as_tibble %>% filter(label %in% c(genomeTable.atlantogenata$label%>% str_replace_all("\\.", " "), st.innerNodes$label)) %>% pull(label)

t.bodysize.atlantogenata.withNodes <- st.Atlantogenata %>% 
  as_tibble() %>% 
  filter(!is.na(label)) %>%
  select(
    `Ancestor/Species`=label, 
    `Estimated Body Size (log(g))`=Median, 
    `95% CI (Low)` = `95%CI_Low`, 
    `95% CI (High)` = `95%CI_High`,  
    `Rate (sqrt)`=sqrt.rate
  ) %>%
  mutate_at(vars(-`Ancestor/Species`, ), . %>% as.numeric %>% round(2)) %>% 
  knitr::kable(
    ., 
    longtable = TRUE, 
    booktabs = TRUE,
    caption = 'A table generated by the longtable package.'
  )

p.bodysize.atlantogenata <- st.Atlantogenata %>% 
  as_tibble() %>% 
  filter(label %in% st.core) %>% 
  mutate(label = label %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidae", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosorcida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata"))) %>% 
  select(-parent, -node, -branch.length) %>%
  ggplot(
    aes(
      y=as.factor(label),
      x=Median, 
      xmin=`95%CI_Low`,
      xmax=`95%CI_High`,
      )
  ) +
  geom_errorbarh() +
  geom_point(color="red") +
  labs(
    x="Body size (95%CI)",
    y="Species/Ancestor"
  ) + 
  theme_pubclean()

# p.tree.bodysize.atlantogenata <- st.Atlantogenata %>% 
#   ggtree()+
#   geom_tiplab()+
#   geom_nodelab()

# p.atlantogenata.zoom <-
#   st.Atlantogenata %>% 
#   ggtree() + 
#   geom_tiplab() + 
#   geom_nodelab(geom="label", nudge_x=-2, size=3) + 
#   xlim(-2, 105)

```
```{r Plot Body Size, include=F}
p.tree <- bsize.tree %>% 
  ggtree(aes(color=lnSize), continuous=T) +
  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))")) +
  geom_highlight(node = bsize.ancAtlantogenata, fill="#EFD28D") +
  xlim(-100, NA) +
  theme_tree2() +
  theme(legend.position="right") + 
  xlab("Rate of Change, Body Size")

p.tree.atlantogenata.circ <- bsize.Atlantogenata %>% 
  ggtree(
    aes(color=lnSize), 
    layout="circular",
    continuous=T
    ) + 
  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))")) #+
  #theme(legend.position="bottom") + 
  # geom_tiplab2(aes(label=Spc.short), hjust = -.1, col="black", size=2)#+ theme(plot.margin=grid::unit(c(-20,-20,-20,-20), "mm"))
  #geom_tiplab(aes(label=Spc.short), hjust = -.1, col="black", size=2)

get.size.color <- function(n){
  viridis::viridis(n=length(size.scale), )[size.scale[n]]
}

p.tree.atlantogenata <-  
  bsize.Atlantogenata %>% 
  ggtree(
    aes(color=lnSize),
    continuous=T,
    size = 1.5
  ) + 
  geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic", 
              show_guide=F, align = F, linesize = 0, offset= 100
              ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidae, label="Proboscidae", 
                  fontsize = 4, align = T, offset = 50, angle=0, barsize = 1, 
                  offset.text = 20,
                  color = bsize.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(lnSize) %>% get.size.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="Tethytheria", 
                  fontsize = 4, align = T, offset = 25, angle=0, barsize = 1, 
                  offset.text = 20, color = bsize.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(lnSize) %>% get.size.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="Paenungulata", 
                  fontsize = 4, align = T, angle=0, barsize = 1, offset.text = 20,
                  color = bsize.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(lnSize) %>% get.size.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosorcida, label="Afrosorcida",
                  fontsize = 4, align = T, angle=0, barsize = 1,
                  color=bsize.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(lnSize) %>% get.size.color
                  ) +
    geom_cladelabel(node=bsize.Atlantogenata.nodes$Macroscelidae, 
                    label="Macroscelidae", 
                    fontsize = 4, align = T, angle=0, barsize = 1,
                    color=bsize.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(lnSize) %>% get.size.color
                    ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Xenarthra, label="Xenarthra", 
                  fontsize = 4, align = T, angle=0, barsize = 1,
                  color=bsize.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(lnSize) %>% get.size.color
                  ) +
  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), 
                        guide = guide_colourbar(title="Body size (ln(g))"))+
  xlim(0,1500) + 
  theme_tree2() + 
  theme(
    legend.position="right",
    panel.background = element_rect(fill = "lightgrey")
  ) + 
  xlab("Rate of Change, Body Size")

```
```{r Gene Copy Trees, include=F}
ml.tree <- ml.treedata@phylo 
ml.treedata.table <- ml.treedata %>% 
  as_tibble() %>% 
  mutate_at(
    vars(-parent, -node, -branch.length, -label),
    . %>% as.character %>% if_else(is.na(.), "?", .) %>% if_else(.=="32", "32+", .)
  )

ml.tree %<>% full_join(., genomeTable.atlantogenata, by="label")
ml.treedata.table <- ml.treedata %>% 
  as_tibble()

bodysize.tree.atlantogenata <- tree.Atlantogenata.table %>% 
  filter(label %in% ml.treedata.table$label) %>% 
  select(label, lnSize)

ml.tree.nodes <- list(
  "Loxodonta.africana"= ml.treedata.table %>% filter(label=="Loxodonta.africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= ml.treedata.table %>% filter(label=="Trichechus.manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= ml.treedata.table %>% filter(label=="Procavia.capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= ml.treedata.table %>% filter(label=="Orycteropus.afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= ml.treedata.table %>% filter(label=="Choloepus.hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = ml.treedata.table %>% filter(label=="Dasypus.novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = ml.treedata.table %>% filter(label=="Chrysochloris.asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = ml.treedata.table %>% filter(label=="Echinops.telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = ml.treedata.table %>% filter(label=="Elephantulus.edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = ml.treedata.table %>% filter(label=="Loxodonta.cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = ml.treedata.table %>% filter(label=="Palaeoloxodon.antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = ml.treedata.table %>% filter(label=="Elephas.maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = ml.treedata.table %>% filter(label=="Mammuthus.columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = ml.treedata.table %>% filter(label=="Mammuthus.primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = ml.treedata.table %>% filter(label=="Mammut.americanum") %>% pull(node) %>% unique,
  "Afrotheria" =  ml.treedata %>% MRCA("Loxodonta.africana", "Chrysochloris.asiatica"),
  "Afroinsectivora" =  ml.treedata %>% MRCA("Elephantulus.edwardii", "Chrysochloris.asiatica"),
  "Afrosoricida" =  ml.treedata %>% MRCA("Echinops.telfairi", "Chrysochloris.asiatica"),
  "Pseudoungulata" =  ml.treedata %>% MRCA("Loxodonta.africana", "Orycteropus.afer"),
  "Paenungulata" =  ml.treedata %>% MRCA("Loxodonta.africana", "Procavia.capensis"),
  "Tethytheria" =  ml.treedata %>% MRCA("Loxodonta.africana", "Trichechus.manatus"),
  "Proboscidea" =  ml.treedata %>% MRCA("Loxodonta.africana", "Mammut.americanum"),
  "Elephantidae" =  ml.treedata %>% MRCA("Loxodonta.africana", "Elephas.maximus"),
  "Loxodontini" =  ml.treedata %>% MRCA("Loxodonta.africana", "Loxodonta.cyclotis"),
  "Loxodona" =  ml.treedata %>% MRCA("Loxodonta.cyclotis", "Palaeoloxodon.antiquus"),
  "Elephantina" =  ml.treedata %>% MRCA("Elephas.maximus", "Mammuthus.columbi"),
  "Mammuthus" =  ml.treedata %>% MRCA("Mammuthus.columbi", "Mammuthus.primigenius"),
  "Root" = ml.treedata %>% MRCA("Loxodonta.africana", "Dasypus.novemcinctus")
)

comparisons <- list(
  c("Proboscidea", "Elephantidae"),
  c("Elephantidae", "Elephantina"),
  c("Elephantina", "Elephas.maximus"),
  c("Elephantina", "Mammuthus"),
  c("Mammuthus", "Mammuthus.columbi"),
  c("Mammuthus", "Mammuthus.primigenius"),
  c("Elephantidae", "Loxodontini"),
  c("Loxodontini", "Loxodonta.africana"),
  c("Loxodontini", "Loxodona"),
  c("Loxodona", "Loxodonta.cyclotis"),
  c("Loxodona", "Palaeoloxodon.antiquus"),
  c("Proboscidea", "Mammut.americanum"),
  c("Tethytheria", "Proboscidea"),
  c("Tethytheria", "Trichechus.manatus"),
  c("Paenungulata", "Tethytheria"),
  c("Paenungulata", "Procavia.capensis"),
  c("Pseudoungulata", "Paenungulata"),
  c("Pseudoungulata", "Orycteropus.afer"),
  c("Afrotheria", "Pseudoungulata"),
  c("Afrotheria", "Afroinsectivora"),
  c("Afroinsectivora", "Elephantulus.edwardii"),
  c("Afroinsectivora", "Afrosoricida"),
  c("Afrosoricida", "Echinops.telfairi"),
  c("Afrosoricida", "Chrysochloris.asiatica"),
  c("Root", "Afrotheria"),
  c("Root", "Choloepus.hoffmanni"),
  c("Root", "Dasypus.novemcinctus")
) %>% 
  set_names(., sapply(.,"[[", 2)) %>% 
  sapply(.,. %>% paste0(., collapse="-to-") %>% paste0("increased-", ., collapse = ""))

ml.tree.increasing.geneLists <- comparisons %>% 
  lapply(
    .,
    function(f){
      dir(
        path = "../output/geneLists/maxLikelihood_model_MK+FQ+I+G4-dataType_MORPH-asrMin_0.8/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata_RBB_filtered_dyn/", 
        pattern=paste0(f, ".*"), 
        full.names = T
        ) %>% 
        read_lines() %>% 
        paste0(collapse = ",")
      }
  )

ml.tree.increasing.geneCounts <- ml.tree.increasing.geneLists %>% 
  lapply(., . %>% str_split(",") %>% unlist %>% stringi::stri_remove_empty(.) %>% length) %>% 
  as_tibble() %>% 
  pivot_longer(everything(), names_to = c("label"), values_to = "N.Genes")

ml.tree.increasing.geneCounts <- full_join(ml.tree, ml.tree.increasing.geneCounts, by="label")

ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts %>% 
  as_tibble %>% 
  full_join(., bodysize.tree.atlantogenata, by="label") %>%
  mutate(delta.GeneCN=log(branch.length)) %>% 
  as.treedata()

ml.tree.increasing.geneCounts.withSize@phylo$node.label[ml.tree.increasing.geneCounts@phylo$node.label == "Xenarthra"] <- NA

not.me <- function(attr1, attr2, threshold=5){
  sapply(
    seq_along(attr1),
    function(n, a1=attr1, a2=attr2, t=threshold){
      ifelse(a1[n] >= t, a2[n], NA)
    }
  )
}
yes.me <- function(attr1, attr2, threshold=5){
  sapply(
    seq_along(attr1),
    function(n, a1=attr1, a2=attr2, t=threshold){
      ifelse(a1[n] < t, a2[n], NA)
    }
  )
}
# ml.tree.increasing.geneCounts.withSize %>% as.tibble %>% print(n=1351)
```

```{r Plot Gene Copy Trees, include=F}
p.ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts.withSize %>% 
  ggtree(
    .,
    branch.length = "none",
    #continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=2
    ) +
  scale_color_viridis_c(guide=F) +
  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold"))) +
  geom_tiplab(aes(label=Species.strict %>% str_replace(" ","\n")), offset = 0.4, color="black", size=3)+
  geom_tiplab(aes(label=not.me(lnSize, N.Genes)), geom="label",  offset=-0.5, color="black", size=3)+ 
  geom_tiplab(aes(label=yes.me(lnSize, N.Genes)), geom="label",  offset=-0.5, color="grey", size=3)+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="black", size=3)+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="grey", size=3)+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, label)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="black", size=3)+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="grey", size=3)+ 
  geom_nodelab2(aes(subset=label=="Root", x=branch), geom="label",
               size=3,
               color="grey")+
  theme_tree() + 
  xlim(0,13) +
  theme(legend.position = "bottom")+
  xlab("Rate of Change, Body Size")
# p.ml.tree.increasing.geneCounts.withSize

p.ml.tree.increasing.geneCounts.withSize.withBL <- ml.tree.increasing.geneCounts.withSize %>% 
  ggtree(
    .,
    # branch.length = "none",
    #continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=2
    ) +
  scale_color_viridis_c(guide=F) +
  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold"))) +
  geom_tiplab(aes(label=Species.strict), 
              offset = 0.02, size=3,
              color="black")+
  geom_tiplab(aes(label=not.me(lnSize, N.Genes)), geom="label",  
              offset=0, size=3,
              color="black")+ 
  geom_tiplab(aes(label=yes.me(lnSize, N.Genes)), geom="label",  
              offset=0.0, size=3,
              color="grey")+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes)), geom="label", 
               nudge_y = 0.2, nudge_x = 0, size=3, 
               color="black")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes)), geom="label", 
               nudge_y = -0.2, nudge_x = 0, size=3, 
               color="grey")+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, label)), geom="label", 
               nudge_y = 0.2, nudge_x = -0.031, size=3, 
               color="black")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label)), geom="label", 
               nudge_y = -0.2, nudge_x = -0.031, size=3, 
               color="grey")+ 
  geom_nodelab2(aes(subset=label=="Root", x=branch), geom="label", 
               size=3, 
               color="grey")+ 
  theme_tree2() + 
  xlim(-0.01,0.45) +
  theme(legend.position = "bottom") +
  xlab("Rate of Change, Body Size")
# p.ml.tree.increasing.geneCounts.withSize.withBL

```

)

# Introduction

One of the major constraints on the evolution of large body sizes in animals is an increased risk of developing cancer. If all cells in all organisms have a similar risk of malignant transformation and equivalent cancer suppression mechanisms, organism with many cells should have a higher prevalence of cancer than organisms with fewer cells.  Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, human cancer incidence increases with increasing adult height [@Green2011; @Nunney:20181c2] and cancer incidence is positively correlated with body size in dogs [@Dobson2013; @Beuchat2015]. There is no correlation, however, between body size and cancer risk between species. This lack of correlation is often referred to as ‘Peto’s Paradox’ [@CaulinAndMaley2011; @Leroi2003; @Peto1975]. While it is clear that a resolution to Peto’s Paradox must involve the evolution of enhanced cancer protection alongside increases in body size and lifespan, the specific genetic, molecular, and cellular mechanisms that underlie this resistance have proven elusive. [c.f. @Ashur-Fabian2004; @Seluanov2008; @Gorbunova2012; @Tian2013; @Sulak2016].

Among the challenges for discovering how animals evolved enhanced cancer protection mechanisms is identifying lineages in which large bodied species are nested within species with small body sizes. Afrotherian mammals are generally small-bodied, similarly to the prediced common ancestor of Eutherian mammals. For example, maximum adult weights are ~70g in golden moles, ~120g in tenrecs, ~170g in elephant shrews, ~3kg in hyraxes, and 60kg in aardvarks [@HAGR]. However, while these extant species are relatively small, the fossil evidence demonstrates that their ancestral lineages reached enormous sizes. For example, while extant hyraxes are relatively small, the extinct Titanohyrax is estimated to have weighted up to ~1300kg [@Schwartz1995]. The largest members of Afrotheria, too, are dwarfed by the size of their recent ancestors: extant cows manatees are large bodied (~322-480kg) but are relatively small compared to the extinct Stellar’s sea cow which is estimated to have weight 8000-10000kg [@Scheffer1972]. Similarly African (4,800kg) and Asian elephants (3,200kg) are the largest living elephant species, but are dwarfed by the truly gigantic extinct Proboscideans such as Deinotherium (~132,000kg), Mammut borsoni (110,000kg), and the Asian straight-tusked elephant (~220,000kg), the largest known land mammal [@Larramendi:20151c2]. Remarkably these large-bodied Afrotherian lineages are nested within small bodied species (Fig. 1) [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], indicating that gigantism independently evolved in hyraxes, sea cows, and elephants (Paenungulates). Thus, Paenungulates are an excellent model system in which to explore the mechanisms that underlie the evolution of large body sizes and augmented cancer resistance.

Although many mechanisms can potentially resolve Peto’s paradox, the most parsimonious route to enhanced cancer resistance is likely through an increased copy number of tumor suppressors. Such an example has been seen in the case of candidate genes such as _TP53_ and _LIF_ [@Abegglen:JAMA2015; @Sulak2016; @Vazquez2018] as well as in studies involving a limited set of candidate genes [@Caulin2015; @Doherty2016]. As these studies focus on _a priori_ gene sets, however, it remains unknown whether this is a general, genome-wide trend in Afrotherian genomes; and whether such a general trend is associated with the recent increases in body size – and therefore expected cancer risk – in these species.

Here, we trace the evolution of body mass and gene copy number variation in Afrotherians in order to investigate whether gene duplications are enriched in large, long-lived species for genes involved in known tumor suppression pathways. Our estimates of the evolution of body mass, similarly to previous studies [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], show that large body masses evolved in a step-wise manner, with major increases in body mass in the Pseudoungulata (17kg), Paenungulata (25kg), Tethytheria (296kg), and Proboscidea (4,100kg) stem-lineages. Furthermore, we see that the ancestral body size increases in Hydracoidia and Sirenia were independent events. To study the evolution of gene copy number, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene duplications in Afrotherian genomes, and used parsimony to infer the lineages in which those duplications occurred. We found gene duplications in lineages with increased body mass were enriched in functions related to tumor suppression, including regulation of the cell cycle, DNA damage repair, and regulation of apoptosis. These data suggest that duplication of tumor suppressors played a role in the evolution of large, long-lived in Afrotherians.

# Methods

## Ancestral Body Size Reconstruction

We built a time-calibrated supertree of Eutherian mammals by combining the time-calibrated molecular phylogeny of Bininda-Emonds _et al._ [-@Bininda-Emonds2008] with the time-calibrated total evidence Afrotherian phylogeny from Puttick and Thomas [@PuttickAnddThomas2015]. While the Bininda-Emonds _et al._ [-@Bininda-Emonds2008] phylogeny includes 1,679 species, only 34 are Afrotherian, and no fossil data are included. The inclusion of fossil data from extinct species is essential to ensure that ancestral state reconstructions of body mass are not biased by only including extant species. This can lead to inaccurate reconstructions, for example, if lineages convergently evolved large body masses from a small bodied ancestor. In contrast, the total evidence Afrotherian phylogeny of Puttick and Thomas [-@PuttickAndThomas2015] includes 77 extant species and fossil data from 39 extinct species. Therefore we replaced the Afrotherian clade in the Bininda-Emonds _et al._ [-@Bininda-Emonds2008] phylogeny with the Afrotherian phylogeny of Puttick and Thomas [-@PuttickAndThomas2015] using Mesquite. Next, we jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of the Brownian motion model that relaxes assumptions of neutrality and gradualism by considering increments to evolving characters to be drawn from a heavy-tailed stable distribution (the "Stable Model") [@ElliotAndMooers2014]. The stable model allows for occasional large jumps in traits and has previously been shown to out-perform other models of body mass evolution, including standard Brownian motion models, Ornstein–Uhlenbeck models, early burst maximum likelihood models, and heterogeneous multi-rate models [@ElliotAndMooers2014].

## Identification of Duplicate Genes

_Reciprocal Best-Hit BLAT:_ We developed a reciprocal best hit BLAT (RBHB) pipeline to quickly identify homologs and estimate gene copy numbers (__Figure 1A__). The Reciprocal Best Hit (RBH) search strategy is conceptually straightforward: 1) Given a gene of interest $G_A$ in a query genome $A$, one searches a target genome $B$ for all possible matches to $G_A$; 2) For each of these hits, one then performs the reciprocal search in the original query genome to identify the highest-scoring hit; 3) A hit in genome $B$ is defined as a homolog of gene $G_A$ if and only if the original gene $G_A$ is the top reciprocal search hit in genome $A$. We selected BLAT [@blat] as our algorithm of choice, as this algorithm is sensitive to highly simliar (>90% identity) sequences, thus identifying the highest-confidence homologs while minimizing many-to-one mapping problems when searching for multiple genes. RBH performs similar to other more complex methods of orthology prediction, and is particularly good at identifying incomplete genes that may be fragmented in low quality/poor assembled regions of the genome [@AltenhoffAndDessimoz2009; @SalichosAndRokas2011].

_Effective Copy Number By Coverage:_ In lower-quality genomes, many genes are fragmented across multiple scaffolds, which results in BLAT calling multiple hits when in reality there is only one gene. To compensate for this, we came up with a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we see each nucleotides of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (Supplementary Figure 1). This allows us to correct for genes that have been fragmented across incomplete genomes, without underestimating the copy number of proteins that lack domains found in the human sequence.

_RecSearch Pipeline:_ We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome. For the query sequences in our search, we used the hg38 Proteome provided by UniProt [Accession #UP000005640; @uniprot], which is a comprehensive set of protein sequences curated from a combination of predicted and validated protein sequences generated by the UniProt Consortium. In order to refine our search, we omitted protein sequences originating from long, noncoding RNA loci (e.g. LINC genes); poorly-studied genes from predicted open reading frames (C-ORFs); and sequences with highly repetitive sequences such as zinc fingers, protocadherins, and transposon-containing genes, as these were prone to high levels of false positive hits. After filtering out problematic protein queries, we then used our pipeline (Figure 1A) to search for all copies of our `r length(UP000005640.genenames)` query genes in publicly available Afrotherian genomes, including African savannah elephant (_Loxodonta africana_: loxAfr3, loxAfr4, loxAfrC), African forest elephant (_Loxodonta cyclotis_: loxCycF), Asian Elephant (_Elephas maximus_: eleMaxD), Woolly Mammoth (_Mammuthus primigenius_: mamPriV), Colombian mammoth (_Mammuthus columbi_: mamColU), American mastodon (_Mammut americanum_: mamAmeI), Rock Hyrax (_Procavia capensis_: proCap1, proCap2, proCap2_HiC), West Indian Manatee (_Trichechus manatus latirostris_: triManLat1, triManLat1_HiC), Aardvark (_Orycteropus afer_: oryAfe1, oryAfe1_HiC), Lesser Hedgehog Tenrec (_Echinops telfairi_: echTel2), Nine-banded armadillo (_Dasypus novemcinctus_: dasNov3), Hoffman's two-toed sloth (_Choloepus hoffmannii_: choHof1, choHof2, choHof2_HiC), Cape golden mole (_Chrysochloris asiatica_: chrAsi1), and Cape elephant shrew (_Elephantulus edwardii_: eleEdw1). For many of these species, we covered multiple assemblies in order to test the effects of assembly size and quality on our hits.

_Duplication gene inclusion criteria:_ In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call. The resulting gene-level copy number table was then combined with the maximum ECNC values observed for each gene in order to call gene duplications. We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes, therefore partial gene duplications were included provided they passed additional inclusion criteria. The ECNC cut off of 1.5 was selected empirically, as this value minimized the number of false positives seen in a test set of genes and genomes. The results of our initial search are summarized in Figure 1B. Overall, we identified [MEDIAN] genes across all species, or [%HITS/QUERIES] of our starting query genes.

_Duplicate gene exclusion criteria:_ We excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (`r query.binTable %>% filter(C.ORF) %>% length`), LOC (`r query.binTable %>% filter(LOC) %>% length`), HLA genes (`r query.binTable %>% filter(HLA) %>% length`), replication dependent histones (`r query.binTable %>% filter(Histone) %>% length`), odorant receptors (`r query.binTable %>% filter(OlfactoryReceptor) %>% length`), ribosomal proteins (`r query.binTable %>% filter(Riboprotein) %>% length`), zinc finger transcription factors (`r query.binTable %>% filter(ZincFinger) %>% length`), viral and repetitive-element-associated proteins (`r query.binTable %>% filter(Viral) %>% length`) and any protein described as either "Uncharacterized," "Putative," or "Fragment" by UniProt in UP000005640 (`r query.binTable %>% filter(Fragment | Uncharacterized | Putative) %>% length`).

## Evidence for Functionality of Identified Genes

To validate and filter out RBHB results, we intersected our results with either gene prediction or transcriptomic evidence as a proxy for functionality.

__Transcriptome Assembly:__ For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated _de novo_ transcriptomes using publically-available RNA-sequencing data from NCBI SRA. We mapped reads to all genomes available for each species, and assembled transcripts using HISAT2 and StringTie, respectively [@HISAT2; @StringTie; @Tuxedo]. RNA-sequencing data was not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax, Aardvark, or the Lesser Hedgehog Tenrec.

__Gene Prediction:__ We obtained tracks for genes predicted using GenScan for all the genomes available via UCSC Genome Browser: African savannah elephant (loxAfr3), Rock Hyrax (proCap1), West Indian Manatee (triManLat1), Aardvark (oryAfe1), Lesser Hedgehog Tenrec (echTel2), Nine-banded armadillo (dasNov3), Hoffman's Two-Toed Sloth (choHof1), Cape golden mole (chrAsi1), and Cape Elephant Shrew (eleEdw1); gene prediction tracks for higher-quality assemblies were not available.

__Evidenced Duplicate Criteria:__ We intersected our records of duplicate hits identified in each genome with the gene prediction tracks and/or transcriptome assemblies using `bedtools` [@bedtools]. When multiple lines of evidence for functionality were present for a genome, we used the union of all intersections as the final output for evidenced duplicates. When analyzing the highest-quality assemblies available for each species, if a species had neither gene prediction tracks nor RNA-seq data for the highest-quality genome available, we conservatively included all hits for the genome in the final set of evidenced duplicates.

## Reconstruction of Ancestral Copy Numbers

We implemented a maximum likelihood method for determining the ancestral copy numbers of genes in _Atlantogenata_ using IQ-Tree. For this analysis, we used an unrooted subset of our prior species tree, including only the aforementioned _Atlantogenata_ species. We generated PHYLIP files containing the copy number of each gene in the highest quality genome for each species, encoding genes on a scale from 1-31+ copies as 1-9, A-V; and encoding a gene's copy number as uncetain ("?") when we did not identify it in the genome. We used the included tree-searching and model-testing functionality in IQ-Tree to determine the most likely topology for the species tree, and to obtain the most likely model for copy number changes in the genome. We defined the ancestral state of a node if it had greater than an 80% posterior probability.

## Pathway Enrichment Analysis

To determine which pathways were associated with duplicated genes in each species and lineage, we used WEBGESTALT to perform overrepresentation analysis (ORA) of the duplicated gene lists relative to our initial query gene list. For the database of pathways used in the analysis, we used Reactome [@Reactome], Wikipathways and Wikipathways_cancer [@Wikipathways], and KEGG [@KEGG].

# Results

```{r Genes in TSG Pathways in ORA}
# ORA.all %<>% mutate(Class = class.pathway(Pathway))
ORA.TSG.list <- ORA.all %>% filter(Direction=="increased") %>% 
  split(., .$Node) %>% 
  lapply(
    ., . %>% 
  filter(str_detect(Pathway, "APC|[Cc]yclin|[Cc]ell [Cc]ycle|Ubiq|NER|ROS|DNA|G1|[Mm]eiotic|Checkpoint|[Ss]tress|ptosis|[Cc]ancer|[Dd]eath|[Pp]53|[Cc][Dd][CcKk]|AKT|Ras|[Rr]epair|ATM")) %>%
  separate_rows(Genes, sep=";") %>% 
  group_by(Pathway) %>% 
  summarize_all(. %>% unique %>% toString) %>% 
  ungroup %>% 
    select(Node, RBHB, Pathway, Genes, FDR.Threshold) %>% 
    distinct %>% 
    mutate(
      N.Genes = Genes %>% str_split(", ") %>% sapply(length),
      FDR.Threshold = FDR.Threshold %>% str_split(", ") %>% sapply(. %>% as.numeric %>% min)
        )
  )

ORA.TSG.list.all <- ORA.all%>% 
  split(., .$Node) %>% 
  lapply(
    ., . %>% 
  filter(str_detect(Pathway, "APC|[Cc]yclin|[Cc]ell [Cc]ycle|Ubiq|NER|ROS|DNA|G1|[Mm]eiotic|Checkpoint|[Ss]tress|ptosis|[Cc]ancer|[Dd]eath|[Pp]53|[Cc][Dd][CcKk]|AKT|Ras|[Rr]epair|ATM")) %>%
  separate_rows(Genes, sep=";") %>% 
  group_by(Pathway) %>% 
  summarize_all(. %>% unique %>% toString) %>% 
  ungroup %>% 
    select(Node, RBHB, Pathway, Genes, FDR.Threshold) %>% 
    distinct %>% 
    mutate(
      N.Genes = Genes %>% str_split(", ") %>% sapply(length),
      FDR.Threshold = FDR.Threshold %>% str_split(", ") %>% sapply(. %>% as.numeric %>% min)
        )
  )
```

```{r LoxAfr All TSG Gene Transcripts}
loxAfr.duplicated <- geneCopyTable %>% 
  filter(Genome == "loxAfr4", ECNC >=1.5, Copy>=2) %>% 
  pull(Gene) %>% 
  unique

genes.loxAfr.TSG.any <- ORA.TSG.list.all$Loxodonta_africana %>% 
  separate_rows(Genes, sep=", ") %>% 
  group_by(Genes) %>% 
  summarize_all(. %>% unique %>% toString) %>% 
  ungroup %>% 
  rename(Gene = Genes) %>% 
  filter(Gene %in% loxAfr.duplicated)

ml.treedata.loxAfrHighlight.TSG.any <- genes.loxAfr.TSG.any %>% 
  pull(Gene) %>% 
  unique %>% 
  set_names(.,.) %>% 
  lapply(., function(gene){
    ml.treedata.table %>% 
      select(parent, node, branch.length, label, all_of(gene)) %>%
      rename("Count"=all_of(gene))%>% 
      as.treedata()
    })

ORA.loxAfr.table.TSG.any <- geneCopyTable %>% 
  filter(str_detect(Genome, "loxAfr|triMan")) %>% 
  mutate(ECNC = str_c("(", ECNC, ")")) %>% 
  unite("CopyECNC", Copy, ECNC, sep=" ") %>% 
  spread(key="Genome", value=c("CopyECNC")) %>% 
  right_join(., genes.loxAfr.TSG.any, by="Gene")

loxAfr4.geneTPM <- merge(
  loxAf4.gene_transcript,
  loxAfr4.transcripts,
  by="geneID"
  )[, Gene :=  str_remove(Locus, "_\\d+$")][]

setkey(loxAfr4.geneTPM, Gene)
setcolorder(loxAfr4.geneTPM)

loxAfr4.geneTPM.ORA.TSG.any <- merge(
  loxAfr4.geneTPM,
  ORA.loxAfr.table.TSG.any,
  by="Gene"
)
```

```{r LoxAfr Increased Gene Transcripts}
genes.loxAfr.TSG.increased <- ORA.TSG.list$Loxodonta_africana %>% 
  separate_rows(Genes, sep=", ") %>% 
  group_by(Genes) %>% 
  summarize_all(. %>% unique %>% toString) %>% 
  ungroup %>% 
  rename(Gene = Genes)

ml.treedata.loxAfrHighlight.TSG.increased <- genes.loxAfr.TSG.increased %>% 
  pull(Gene) %>% 
  unique %>% 
  set_names(.,.) %>% 
  lapply(., function(gene){
    ml.treedata.table %>% 
      select(parent, node, branch.length, label, all_of(gene)) %>%
      rename("Count"=all_of(gene))%>% 
      as.treedata()
    })

ORA.loxAfr.table.TSG.increased <- geneCopyTable %>% 
  filter(str_detect(Genome, "loxAfr|triMan")) %>% 
  mutate(ECNC = str_c("(", ECNC, ")")) %>% 
  unite("CopyECNC", Copy, ECNC, sep=" ") %>% 
  spread(key="Genome", value=c("CopyECNC")) %>% 
  right_join(., genes.loxAfr.TSG.increased, by="Gene")

loxAfr4.geneTPM.ORA.TSG.increased <- merge(
  loxAfr4.geneTPM,
  ORA.loxAfr.table.TSG.increased,
  by="Gene"
)

```
```{r LoxAfr Other Gene Transcripts}
genesToHighlight.plus <- c(
   names(ml.treedata.loxAfrHighlight.TSG.any),
  "RBX1", #Cell cycle, ubiqu., Probo
  "LIN9", # Cell cycle, Probo
  "MND1", # Cell Cycle, Elephantidea,
  "E2F2", #Cell cycle, Tethytheria
  "MAX", # Cell Cycle, Tethytheria
  "SYCP3", #Cell cycle, unique to LoxAfr

  "CDK1", # APC and Cell Cycle, Atlantogenata
  "RNF168", # ATM, cell cycle, Paen or Pseudo
  "RBBP8", # ATM Signaling, Cell Cycle, DSB repair, Tethy
  "MAD2L1", # APC, Tethytheria
  "CSNK1A1", #APC? Tethytheria
  "RPA2", # ATR, cell cycle, and stress response, Probo
  "PSMB3", # APC/C, ubiq, Probo
  "BUB3", # APC, 2 copies until 3 in LoxAfr

  "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata
  "STK11", #REgulation of TP53, Paenungulata
  "COX5A", #TP53 transcriptional regulation, PRobodiscea
  "BRD7", # TP53 regulation, Tethytheria
  "TP53", #TP53 duplicates in Tethytheria
  "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria
  "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants
  "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri
  "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants
  "COX20", # TP53 Regulates Metabolic Genes, LoxAfr

  "SOD1",  # Oxidative Stress, Probodiscea
  "SCMH1", # DNA damage response and OS-induced senescence, Probo
  "RNF2", # OSISenecense, Tethytheria

  "ORC3", # Replication Stress, Loxondontini, Elephas

  "RACK1", # Death Receptor Signaling, Tethy or Paen
  "PCBP2", # Ferroptosis, Loxodontini, Trichechus
  "NSMAF", # Death Receptor Signaling, Tethytheria
  "PCBP1", # Ferroptosis, Probodiscea
  "SERBP1", # Apoptosis, Probodiscea
  "CASP12", # Apoptosis, Probodiscea
  "NET1", # Apoptosis in cancers, Elephantidae
  "CSF2RB", #Apoptosis, Elephantidae

  "RASA1", # RAS signaling Pathway, Paen
  "RAB5B", # RAS signaling, Paen
  "NRAS", #Literally just RAS, Tethytheria
  "GSK3A", #RAS, Tethy

  "POLR2K", # TC-NER, Elephantidae

  "HSBP1", # Heat Shock, Elephantidae
  "ERCC8", #Ubiquitin mediated proteolysis, Elephantidea
  "DTX3L", # Ubiquitination & Proteasome degradation, Loxodontini
  "CTTN", # Proteoglycans in Cancer, Loxodontini

  "DNAJC2", # Heat Shock, Probo
  "DET1", # Ubiq-mediate proteolysis, Probo

  "MAPK3", # Many cancers, Afrotheria
  "GABARAP", #Senescence and autophagy in cancer, Paen or Tethy
  "HNRNPK", #LncRNA involvement in Wnt and Cancer, Probodiscea
  "CACYBP", # Gastric Cancer Network 2, Probodiscea
  "CHPT1", # Choline Metabolism in Cancer, Probodiscea
  "PYCR2", # Metabolic reprogramming in colon cancer, Elephantidae
  "PLPP2", # Choline Metabolism in Cancer, Loxodontini, Elephas, Trichechus
  "PLAC8", # Gastric Cancer Network 2, continuously increases towards Loxodona

  "LIMK1", #G13 Signaling, Probo

  "EEF1A1", # Vinnie wanted to highlight this
  "EIF4E", #Probodiscea
  "EIF4A1" # Loxodontini
)

ml.treedata.loxAfrHighlight.plus <- genesToHighlight.plus %>% 
  set_names(.,.) %>% 
  lapply(., function(gene){
    ml.treedata.table %>% 
      select(parent, node, branch.length, label, all_of(gene)) %>% 
      rename("Count"=all_of(gene))%>% 
      as.treedata()
    })

loxAfr4.geneTPM.genesToHighlight.plus <- merge(
  loxAfr4.geneTPM,
  geneCopyTable %>% 
    filter(str_detect(Genome, "loxAfr|triMan")) %>% 
    mutate(ECNC = str_c("(", ECNC, ")")) %>% 
    unite("CopyECNC", Copy, ECNC, sep=" ") %>% 
    spread(key="Genome", value=c("CopyECNC")),
  by="Gene"
)

# loxAfr4.geneTPM.duplicated %>% inner_join(., ORA.loxAfr.all.table, by="Gene") %>% arrange(Gene)
```

```{r LoxAfr ORA Gene Transcripts}

p.ml.treedata.loxAfrHighlight.TSG.any <- ml.treedata.loxAfrHighlight.TSG.any %>% 
  names %>% 
  set_names(.,.) %>% 
  lapply(
    .,
    function(n, l.tr=ml.treedata.loxAfrHighlight.TSG.any){
      ggtree(l.tr[[n]]) +
        geom_tiplab(aes(label=label), offset = 0.01) +
        geom_label(aes(label=Count)) + 
        theme_tree2() + 
        labs(x= "Rate of Change", caption=n)
      }
  ) # %>%
  # ggarrange(plotlist = ., ncol = 2, nrow = length(.)/2, labels = names(.))

p.ml.treedata.loxAfrHighlight.TSG.increased <- ml.treedata.loxAfrHighlight.TSG.increased %>% 
  names %>% 
  set_names(.,.) %>% 
  lapply(
    .,
    function(n, l.tr=ml.treedata.loxAfrHighlight.TSG.increased){
      ggtree(l.tr[[n]]) +
        geom_tiplab(aes(label=label), offset = 0.01) +
        geom_label(aes(label=Count)) + 
        theme_tree2() + 
        labs(x= "Rate of Change", caption=n)
      }
  )

p.ml.treedata.loxAfrHighlight.plus <- ml.treedata.loxAfrHighlight.plus %>% 
  names %>% 
  set_names(.,.) %>% 
  lapply(
    .,
    function(n, l.tr=ml.treedata.loxAfrHighlight.plus){
      ggtree(l.tr[[n]]) +
        geom_tiplab(aes(label=label), offset = 0.01) +
        geom_label(aes(label=Count)) + 
        theme_tree2() + 
        labs(x= "Rate of Change", caption=n)+
        labs_pubr() + 
        theme(
          axis.text.y = element_blank()
        )
      }
  )

```

```{r expressed loxAfr duplicates}
loxAfr4.expressedDups <- loxAfr4.geneTPM %>% 
  group_by(Gene) %>% 
  summarize(n = n_distinct(Locus)) %>% 
  filter(n>1) %>% 
  pull(Gene)

loxAfr4.geneTPM.ORA.expressedDups <- loxAfr4.geneTPM.ORA.TSG.any %>% 
  filter(Gene %in% loxAfr4.expressedDups)
```

```{r Plot: Expressed LoxAfr Duplicates ORA}
plot.coolgenes <- loxAfr4.geneTPM.ORA.expressedDups %>% 
  filter(Gene %in% genesToHighlight.plus) %>% 
  mutate(Reactome.Pathway.Prime = Pathway %>% str_split(", ") %>% sapply("[", 1),
         id = str_remove(Locus, ".*_") %>% factor(., levels=sort(as.numeric(unique(.))))) %>% 
  group_by(Locus) %>% 
  summarize(TPM=max(TPM), Reactome.Pathway.Prime=unique(Reactome.Pathway.Prime), id=unique(id), Gene = unique(Gene)) %>% 
    ggplot(
      aes(
        x=id, 
        y=TPM, 
        color=Reactome.Pathway.Prime
        )
      ) +
    geom_point(
      size=3
    )+
    geom_segment(
      aes(
        x=id,
        xend=id,
        y=0,
        yend=TPM
      )
    )+
    scale_y_sqrt()+
    #scale_color_brewer(palette = "Dark2", name="Reactome Pathway")+
    guides(color=guide_legend(nrow = 3))+
    facet_grid(cols = vars(Gene), scales = "free_x") + 
    theme_pubclean()+
    labs(
      x="Copy",
      y="TPM (sqrt)"
    ) + 
  theme(legend.position="bottom")

```

```{r ORA.dotplots}
p.loxAfr.ORA.dotplot <- ORA.all %>% 
    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome") %>% 
    graph.ORA.dotplot(., plotly=F)

p.loxAfr.ORA.dotplot.increased <- ORA.all %>% 
    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome", Direction=="increased") %>% 
    graph.ORA.dotplot(., plotly=T)
```

## Step-wise evolution of large, long-lived Afrotherians

```{r Figure 1, fig.cap="Figure 1: Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata. __A)__ Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch. Atlantogenata is highlighted at the bottom. __B)__ Zoom-in of (A) on Atlantogenata. Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size"}
f1a <- p.tree
f1b <- p.tree.atlantogenata
f1 <- ggarrange(
  f1a, f1b, 
  nrow=1, ncol = 2, 
  labels="AUTO", 
  common.legend = T, legend = "bottom")
f1
```
```{r Table 1}
t.bodysize.atlantogenata.withNodes
```

To trace the evolutionary history of body mass and lifespan in Afrotherians, we built a time-calibrated supertree of Eutherian mammals combining 1,679 species from Bininda-Emonds et al [@Bininda-Emonds2008] with a total evidence Afrotherian phylogeny including 77 extant and fossil data from 39 extinct species [@PuttickAndThomas2015]. Fossil data from extinct species were included to ensure that ancestral state reconstructions of body mass in Afrotherians were not biased by only including extant species, which can lead to inaccurate reconstructions, for example, if lineages multiple lineages evolved large body masses from a small bodied ancestor. We jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of a Brownian model of character evolution, which allows for occasional large jumps in traits (stable model) and out performs standard Brownian motion and Ornstein?Uhlenbeck models of character evolution [@ElliotAndMooers2014].

Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a 65× increase in body mass in the stem-lineage of _Pseudoungulata_ (17kg), a 1.5× increase in body mass in the stem-lineage of _Paenungulata_ (25kg), a 12× increase in body mass in the stem-lineage of _Tehthytheria_ (296kg), and a 14× increase in body mass in the stem-lineage of _Proboscidea_ (4,100kg; Figure 1). The ancestral _Hyracoidea_ was inferred to be relatively small (2.86-15.71kg), and rate accelerations were coincident with independent body mass increases in large hyraxes such as _Titanohyrax andrewsi_ (67× increase in body mass). While the body mass of the ancestral _Sirenian_ was inferred to be large (61-656kg), a rate acceleration occurred coincident with a 10× body mass increase in Stellar's sea cow. Rate accelerations also occurred coincident with 36× body mass reduction in the stem-lineage of the dwarf elephants _Elephas_ (_Palaeoloxodon_) _falconeri_ and _Palaeoloxodon cypriotes_. These data suggest that gigantism in _Afrotherians_ evolved step-wise, from small to medium bodies in the _Pseudoungulata_ stem-lineage, medium to large bodies in the _Tehthytherian_ stem-lineage and extinct hyraxes, and from large to exceptionally large bodies independently in the _Proboscidean_ stem-lineage and Stellar's sea cow (Figure 1).

## Pervasive duplication of tumor suppressors during the origins of large bodied Afrotheirans
```{r ECNC DIAGRAMS, include=F}
ecnc.df1 <- tibble(
  hit = as.character(1) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0),
  x2=c(100),
  y1=c(0),
  y2 = c(1)
)
ecnc.df2 <- tibble(
  hit = as.character(1:2) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0, 0),
  x2=c(100, 100),
  y1=c(0, 1),
  y2 = c(1, 2)
)
ecnc.df3 <- tibble(
  hit = as.character(1:3) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0,25,75),
  x2=c(25,75,100),
  y1=c(0, 0, 0),
  y2 = c(1, 1, 1)
)
ecnc.df4 <- tibble(
  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0,25,75, 0),
  x2=c(25,75,100, 100),
  y1=c(0, 0, 0, 1),
  y2 = c(1, 1, 1, 2)
)

ecnc.df5 <- tibble(
  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0,0,0, 0),
  x2=c(100,15,15, 15),
  y1=c(0, 1, 2, 3),
  y2 = c(1, 2, 3, 4)
)

ecnc.df6 <- tibble(
  hit = as.character(c(1,2,1,3, 3)) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  label = factor(c("Hit 1", "Hit 2", "Unidentified", "Unidentified", "Hit 3")),
  x1=c(0, 0, 75, 75, 50),
  x2=c(75, 50, 100, 100, 75),
  y1=c(0, 1, 0, 1, 1),
  y2 = c(1, 2, 1, 2, 2),
  transparent = c(0,0,1,1,0) %>% as.factor
)

ecnc.onegene.onehit <- 
  ecnc.df1 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: One Contiguous Gene, One Hit"
  )

ecnc.twogene.twohit <- 
  ecnc.df2 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: Two Contiguous Genes, Two Hits"
  )

ecnc.onegene.threehits <-
  ecnc.df3 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: One Fragmented Gene, Multiple Hits"
  )

ecnc.twogene.fourhits <-
  ecnc.df4 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"#,
    #title = "ECNC: Two Genes, Multiple Hits"
  )

ecnc.onegene.manypartialhits <-
  ecnc.df5 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1.45, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
    # title = "ECNC: One Real Gene, Multiple Hits on Conserved Domain"
  )

ecnc.twogene.partialhits <-
  ecnc.df6 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit,
      alpha = transparent,
      linetype = transparent
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=label), color="black", inherit.aes = F)+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_alpha_manual(values = c("0"=1,"1"=0.5), guide=F) + 
  scale_linetype_manual(values = c("0"="solid","1"="dotted"), guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: Two Real Genes, Unidentified Tails"
  )

```
```{r Correlations between genome quality scores}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}
reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

CEGMA <- full_join(CEGMA.qc, CEGMA.completeness, by="Genome")
CEGMA.completeonly <- CEGMA %>% 
  filter(Group == "Complete_all") %>% 
  rename(
    Completeness = `%Completeness`,
    totalLength = `Total length (nt)`,
    meanSequenceLength = `Mean sequence length (nt)`,
    L50 = `L50 sequence count`,
    N50 = `N50 sequence length (nt)`,
    Total = `#Total`, 
    pc.Ortho = `%Ortho`
  )

GCT.CEGMA.all <- geneCopyTable %>% 
  mutate(
    ECNC = ECNC %>% sapply(
        .,
        . %>% if_else(is.na(.), 0, .) %>% 
      sum(., ifelse(.>round(.), 0.01, 0)) %>% 
      round()
    ),
    Copy = Copy %>% sapply(. %>% if_else(is.na(.), 0, .)),
    CN = if_else(ECNC<Copy, ECNC, Copy)
  )%>% 
  inner_join(., CEGMA, by="Genome") %>% 
  mutate_if(is.character, as.factor)%>% 
  select(-Group)

GCT.CEGMA.completeOnly <- geneCopyTable %>% 
  mutate(
    ECNC = ECNC %>% sapply(
        .,
        . %>% if_else(is.na(.), 0, .) %>% 
      sum(., ifelse(.>round(.), 0.01, 0)) %>% 
      round()
    ),
    Copy = Copy %>% sapply(. %>% if_else(is.na(.), 0, .)),
    CN = if_else(ECNC<Copy, ECNC, Copy)
  )%>% 
  inner_join(., CEGMA.completeonly, by="Genome") %>% 
  mutate_if(is.character, as.factor) %>% 
  select(-Group)

cor.GCT.CEGMA.completeOnly <- GCT.CEGMA.completeOnly %>% 
  mutate_if(is.factor, as.numeric) %>% 
  as.matrix %>%
  cor %>%
  reorder_cormat %>%
  get_upper_tri %>% 
  melt(., na.rm=T) 
  # as.data.frame %>%
  # rownames_to_column(var = 'var1') %>%
  # gather(var2, value, -var1) %>% 
  # mutate(var_order = paste(var1, var2) %>%
  #          strsplit(split = ' ') %>%
  #          map_chr(
  #            ~ sort(.x) %>%
  #              paste(collapse = ' ')
  #          )
  # ) %>%
  # mutate(cnt = 1) %>%
  # group_by(var_order) %>%
  # mutate(cumsum = cumsum(cnt)) %>%
  # filter(cumsum != 2) %>%
  # ungroup %>%
  # select(-var_order, -cnt, -cumsum)

cor.GCT.CEGMA.completeOnly %>% filter(Var2 %in% c("ECNC", "Copy", "CN"), !Var1 %in% c("ECNC", "Copy", "CN"))
p.cor.GCT.CEGMA.completeOnly <- cor.GCT.CEGMA.completeOnly %>% 
  # filter(value>0) %>% 
  ggplot(
    aes(x=Var2, y=Var1, fill = value)
    )+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()+ 
# geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))

# Best independent vars:
#   Genome, N50, Mean Sequence Length, Completeness, #Total, %Ortho, Average
# Note: Genome highly correlated with "sum length seq", which is HC w/ "N50"
# Nothing correlates with Gene
# 
```
```{r Phylogenetic Correlations}
CEGMA.completeonly.wGenome <- inner_join(
  GCT.CEGMA.completeOnly, 
  genomeTable.atlantogenata %>% select(label, BestGenome), 
  by=c("Genome"="BestGenome")
) %>% 
  select(label, everything())
missing.bsize.atl.CEGMA.completeonly <- bsize.Atlantogenata %>% 
  as.tibble %>% 
  filter(!label %in% CEGMA.completeonly.wGenome$label) %>% pull(label)
bsize.atl <- bsize.Atlantogenata@phylo %>% 
  drop.tip(missing.bsize.atl.CEGMA.completeonly)

bm<-corBrownian(1, bsize.atl)
pl<-corPagel(1, bsize.atl)

# CEGMA.completeonly.wGenome %<>% filter(Gene == "TP53")
# rown <- CEGMA.completeonly.wGenome$label %>% unique
# CEGMA.completeonly.wGenome %<>% mutate_if(is.character, . %>% as.factor %>%  as.numeric)%>% 
#   as.data.frame
# rownames(CEGMA.completeonly.wGenome) <- rown
```

```{r Phylogenetic Correlations: model building}
warning("TODO: Collapse RBHB Genes into Total, Duplicate, and Median CN")
warning("TODO: CEGMA duplicate: multiply CEGMA %Ortho times #Total to get a Median CN")
warning("TODO: Rename CEGMA things to CEGMA.[thing]")
warning("TODO: I have a linear model for each gene regressing genome quality stats with gene copy numbers in a PGLS; figure out why this is useful and how to plot it.")

runGLS.bm <- function(df, pcor = bm){
  # print(class(df))
  # print(head(df))
  bla <- gls(
  CN~Genome + N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average, 
  data = df, 
  correlation=pcor
  ) 
  # print(bla)
  # print(class(bla))
  return(bla)
}
tidyGLS <- function(df){
  print(names(df))
  print(class(df$results))
  a <- tidy(df$results, conf.int=T)
  print(a)
  return(a)
}

gls.CN.bm <- CEGMA.completeonly.wGenome %>% 
  # filter(Gene %in% c("TP53", "LIF")) %>%
  nest(-Gene) %>% 
  mutate(
    data = map(
      data,
      . %>%
        mutate_if(is.character, . %>% as.factor) %>% 
        mutate(label=as.character(label)) %>% 
        mutate_if(is.factor, as.numeric) %>% 
        as.data.frame %>% 
        column_to_rownames("label")
    ),
    model = map(
      data, 
      possibly(
        ~runGLS.bm(.),
        otherwise = NA_real_
      ),
    ),
    tidied = map(
      model,
      possibly(
        .%>% tidy(., conf.int=T),
        otherwise = NA_real_
      )
    )
  )

gls.CN.bm.perGene.tidy <- gls.CN.bm %>% 
  unnest(tidied) #%>%
  # select(-model)

# gls.atl.1<-CEGMA.completeonly.wGenome %>% 
#   mutate_if(is.character, . %>% as.factor) %>% 
#   mutate(label=as.character(label)) %>% 
#   mutate_if(is.factor, as.numeric) %>% 
#   group_by(Gene) %>% 
#   do(model = . %>% 
#       as.data.frame %>% 
#       column_to_rownames("label") %>% 
#        gls(
#   CN~Genome + N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average, 
#   data = ., 
#   correlation=bm)
#   ) %>% tidy(model)
# gls.atl.2<-gls(
#   CN~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average,
#   data=CEGMA.completeonly.wGenome %>% filter(Gene == "TP53") %>%
#       mutate_if(is.character, . %>% as.factor) %>% 
#       mutate(label=as.character(label)) %>% 
#       mutate_if(is.factor, as.numeric) %>% 
#       as.data.frame %>% 
#       column_to_rownames("label"), 
#   correlation=pl)
# gls.atl.3<-gls(
#   Copy~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average,
#   data=CEGMA.completeonly.wGenome %>% filter(Gene == "TP53") %>%
#       mutate_if(is.character, . %>% as.factor) %>% 
#       mutate(label=as.character(label)) %>% 
#       mutate_if(is.factor, as.numeric) %>% 
#       as.data.frame %>% 
#       column_to_rownames("label"), 
#   correlation=bm)
# gls.atl.4<-gls(
#   Copy~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average,
#   data=CEGMA.completeonly.wGenome %>% filter(Gene == "TP53") %>%
#       mutate_if(is.character, . %>% as.factor) %>% 
#       mutate(label=as.character(label)) %>% 
#       mutate_if(is.factor, as.numeric) %>% 
#       as.data.frame %>% 
#       column_to_rownames("label"), 
#   correlation=pl)
# gls.atl.4 %>% summary
# AIC(gls.atl.3, gls.atl.1,gls.atl.2,gls.atl.4)
```

```{r LM Duplicates~Genome quality metrics}
m1 <- lm(data = GCT.CEGMA.completeOnly, formula = Copy~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average)
```

```{r Figure 2C1, fig.cap="Figure 2 (Candidate 1): A Reciprocal Best-Hit BLAT pipeline for identifying gene copy number in other genomes. A) A graphic summary of the reciprocal best-hit strategy. B) Estimated Copy Number by Coverage. C) "}
f2a <- image_read("../output/figs/RBHBDiagram.png") %>% image_ggplot()
f2b <- ecnc.twogene.fourhits
f2c <- p.ml.tree.increasing.geneCounts.withSize
f2ab <- ggarrange(f2a, f2b, nrow=2, ncol=1, labels = LETTERS[1:2])
f2.1 <- ggarrange(f2ab, f2c, nrow=1, ncol=2, widths = c(1, 2), labels=c(NA, "C"))
f2.1
```
```{r Figure 2C2, fig.cap="Figure 2 (Candidate 2): Gene Duplicates in _Atlantogenata_. A) A graphic summary of the reciprocal best-hit strategy. B) Estimated Copy Number by Coverage. C)"}
f2a <- p.ml.tree.increasing.geneCounts.withSize
f2b <- p.bodysize.atlantogenata
f2.2 <- ggarrange(f2a, f2b, nrow=2, ncol=1, labels = "AUTO")
f2.2
```
```{r Table 2: ML model summary}
warning("Table 2")
```

Enhanced cancer suppression may have evolved through many mechanisms; among the most parsimonious is an increase in the copy number of genes with tumor suppressor functions. Previous studies focusing on candidate gene studies, for example, have identified increased copy number of the tumor suppressors _TP53_ and _LIF_ in elephants [@Abegglen2015; @Caulin2015; @Doherty2016; @Sulak2016; @Vazquez2018]. Therefore, in order to test whether this was a pervasive phenomena genome-wide in _Afrotherians_, we used a Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in _Afrotherian_ and _Atlantogenatan_ genomes (Fig. 2A). Because RBHB-like approaches can over-estimate copy number when genes are fragmented  or incorrectly assembled across multiple scaffolds, we also inferred copy number using a complementary method that quantifies the ratio between observed and expected gene coverage per nucleotide (ECNC) (Supplementary Figure 1). By only including nucleotides from the query sequence that were observed in the target genome, we also correct for partial hits where some or all of the homologs of a gene have diverged from the human homolog.

Because our sequence database included various protein transcripts for each gene, in order to obtain gene-level copy number information and elimate any many-to-one mappings of hits, we labeled each exon of every reciprocal best hit (RBH) with the gene corresponding to the query transcript and merged all overlapping exons; next, we eliminated any many-to-one exons that resulted from the previous step. Finally, we reassembled the gene loci based on the original transcript starts and ends, and the collapsed exon data, obtaining the full sequence of each RBH locus. Genes were considered to be duplicated if its copy number via RBHB was greater than or equal to 2, and the maximum ECNC among all transcripts prior to filtering was greater than or equal to 1.50. This cutoff of ECNC was selected to account for truncated gene duplications, which have been shown to be functional in various examples [@CITE examples of this].

To reconcile the Atlantogenatan phylogeny with duplication events, we used maximum likelihood to reconstruct likely ancestral copy numbers for each gene at each node in the phylogeny. To define the copy number of a gene, we conservatively used the lesser value between the RBHB hit count, and the ECNC value rounded to the nearest whole number. In order to perform

Next, in order to select genes and duplicates which were likely functional, we omitted any hits that were not supported by either the gene prediction method GenScan, or by at least one transcript assembled from publically-available RNA-seq data.

We describe the number of genes that increased in each lineage in _Atlantogenata_ in Figure 2. Among the genes that increased in copy number in the elephant lineage are TP53 and LIF, as previously described. Furthermore, we identify `r warning("TODO: Finish this section")`

## Retroduplication of EEF1A1 in the stem-lineage of Tehthytheria

```{r Fig3, fig.cap="Retroduplication of Eukaryotic translation Elongation Factor 1 Alpha 1 (EEF1A1). A) Gene Copy Number of EEF1A1 in _Atlantogenata_. B) Expression of EEF1A1 copies in the West Indian Manatee and African Elephant"}
f3a <- p.ml.treedata.loxAfrHighlight.plus$EEF1A1
# f3b <- 
# f3c <- 
f3 <- ggarrange(
  f3a,
  # f3b, f3c,
  # nrows=1, ncol=3
  labels="AUTO"
)
f3
warning("Fig 3: EEF1A1")
```

Previous studies have identified genes whose transgenic overexpression delays aging related phenotypes and extends lifespan. Overexpression of the eukaryotic translation elongation factor 1 alpha 1 (EEF1A1) gene, for example, extends lifespan in transgenic Drosophila. Like most mammalian genomes, we found that Afrotherian genomes encoded numerous EEF1A1 retrogenes (Fig. 3A). While the majority of these retrogenes have premature stop codons and insertions/deletions characteristic of pseudogenes, at least 13 Tehthytherian-specific EEF1A1 retrogenes have the potential to encode functional proteins (Fig. 3B). Unfortunately it is difficult to determine if these retrogenes are transcribed because they have relatively high sequence similarity (83-88%), suggesting evidence of transcription using RNA-Seq data is likely an artifact. Indeed, nearly all of the EEF1A1 retrogene transcripts assembled by StringTie were confined to their gene body and UTRs and thus lacked unique sequences. The EEF1A1RTG13 transcript, however, initiates within an upstream MLT1B retrovirus-like MaLR transposable element and includes unique 5?- and 3?-UTR sequences, indicating at least one EEF1A1 retrogene is transcribed (Fig. 3C).

## Duplications that occured recently in Probodiscea are enriched for tumor suppressor pathways

In order to infer the functional consequences of these gene duplications, we tested if duplicate genes were enriched in specific pathways relative to our initial query set of genes. We used `r warning("TODO: Finish this section")`

## Concerted duplication of TP53 and TP53-related genes towards _Probodiscea_

```{r Fig4 }
warning("Fig4")
```

## Step-wise reduction of intrinsic cancer risk in large, long-lived Afrotherians

```{r Lifespan Regression Atlantogenata}
tree.size.lifespan.Atlantogenata <- left_join(
  bsize.Atlantogenata %>% as_tibble %>% select(parent, node, branch.length, label, lnSize), 
  anage %>% select(-lnSize.anage), 
  by="label"
  ) %>% 
  as.treedata()

atl.noLife <- tree.size.lifespan.Atlantogenata %>%
  as_tibble() %>%
  filter(is.na(Lifespan)) %>% 
  pull(label)

atl.withLife <- tree.size.lifespan.Atlantogenata %>% 
  as_tibble() %>%
  filter(!is.na(Lifespan)) %>% 
  pull(label)

tree.size.lifespan.Atlantogenata.withLife <-
  tree.size.lifespan.Atlantogenata %>%
  drop.tip(atl.noLife) %>% 
  as.phylo()
dat.size.lifespan.Atlantogenata.withLife <-
  tree.size.lifespan.Atlantogenata %>% 
  as_tibble %>%
  filter(label %in% atl.withLife) %>% as.data.frame()

rownames(dat.size.lifespan.Atlantogenata.withLife) <- dat.size.lifespan.Atlantogenata.withLife$label

size.lifespan.Atlantogenata.lm <- gls(
  Lifespan ~ lnSize,
  data=dat.size.lifespan.Atlantogenata.withLife,
  correlation = corPagel(1, tree.size.lifespan.Atlantogenata.withLife)
)

tree.size.lifespan.Atlantogenata.noLife <-
  tree.size.lifespan.Atlantogenata %>%
  drop.tip(atl.withLife) %>% 
  as.phylo()
dat.size.lifespan.Atlantogenata.noLife <-
  tree.size.lifespan.Atlantogenata %>% 
  as_tibble %>%
  filter(label %in% atl.noLife)%>% 
  select(-Lifespan) %>% 
  mutate(
    label=sapply(
      seq_along(label),
      function(n, l=label, nd=node){
        ifelse(is.na(l[n]), nd[n], l[n])
      }
    )
  ) %>% 
  as.data.frame()

rownames(dat.size.lifespan.Atlantogenata.noLife) <- dat.size.lifespan.Atlantogenata.noLife$label

dat.size.lifespan.Atlantogenata <- predict(
  object = size.lifespan.Atlantogenata.lm, 
  newdata = dat.size.lifespan.Atlantogenata.noLife,
  interval="confidence"
    ) %>%
  tibble(Lifespan = ., label=rownames(dat.size.lifespan.Atlantogenata.noLife)) %>%
  full_join(., dat.size.lifespan.Atlantogenata.noLife, by="label") %>%
  rbind(., dat.size.lifespan.Atlantogenata.withLife) %>% 
  rename(lnLifespan = Lifespan) %>% 
  mutate(Lifespan = exp(lnLifespan))

tree.size.lifespan.Atlantogenata.full <- full_join(tree.size.lifespan.Atlantogenata, dat.size.lifespan.Atlantogenata, by='node') %>% as_tibble %>% select(parent, node, branch.length=branch.length.x, -label.y, label=label.x, -lnSize.y, lnSize=lnSize.x, -Lifespan.x, Lifespan=Lifespan.y) %>% as.treedata()

# tree.size.lifespan.Atlantogenata.full %>% as.tibble %>% print(n=12351235123)
```
```{r Estimated Cancer Risk Per Node}
cancerSucceptibility.Atlantogenata <- tree.size.lifespan.Atlantogenata.full %>% 
  as_tibble %>% 
  mutate(K1 = (Lifespan)^6 * exp(lnSize),
         lnK1 = 6*log(Lifespan) + lnSize)
table.cancerSucceptibility.Atlantogenata <- cancerSucceptibility.Atlantogenata %>% 
  # arrange(parent) %>% 
  mutate(
  K2 = parent %>% 
    sapply(
     .,
     function(n, d=cancerSucceptibility.Atlantogenata){
       d %>% filter(node==n) %>% pull(K1) %>% unique
     }
    ),
  CancerSucceptabilityChange = K2/K1,
  log2CancerSucceptabilityChange = log2(K2/K1)
)
tree.cancerSucceptibility.Atlantogenata <- table.cancerSucceptibility.Atlantogenata %>% 
  as.treedata()
```

```{r Fig 5, fig.cap=""}
warning("TODO: Caption Fig 5")
quiet <- function(x) { 
  sink(tempfile()) 
  on.exit(sink()) 
  invisible(force(x)) 
} 
scale.log2CancerSucceptabilityChange <- table.cancerSucceptibility.Atlantogenata %>% 
  pull(log2CancerSucceptabilityChange) %>% 
  unique %>% 
  as.numeric() %>% 
  sort %>% 
  set_names(.,.)

scale.log2CancerSucceptabilityChange.min <- min(scale.log2CancerSucceptabilityChange)
scale.log2CancerSucceptabilityChange.max <- scale.log2CancerSucceptabilityChange[length(scale.log2CancerSucceptabilityChange)]

get.log2CancerSucceptabilityChange.color <- function(n){
  scale = viridis::viridis(n=length(scale.log2CancerSucceptabilityChange)) %>% 
    set_names(., as.character(scale.log2CancerSucceptabilityChange))
  scale[as.character(n)]
}

t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
  size.lifespan.Atlantogenata.lm,
  title = "ln(Lifespan) & ln(Body Size) Regression",
  style = "all"
))

f5a <- tree.cancerSucceptibility.Atlantogenata %>% 
  ggtree(
    aes(color=as.numeric(log2CancerSucceptabilityChange)),
    continuous=T,
    size = 1.5
  ) + 
  # geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic", 
  #             show_guide=F, align = F, linesize = 0, offset= 100
  #             ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidae, label="Proboscidae", 
                  fontsize = 4, align = T, offset = 50, angle=0, barsize = 1, 
                  offset.text = 20,
                  color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Proboscidae") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="Tethytheria", 
                  fontsize = 4, align = T, offset = 25, angle=0, barsize = 1, 
                  offset.text = 20, color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="Paenungulata", 
                  fontsize = 4, align = T, angle=0, barsize = 1, offset.text = 20,
                  color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Paenungulata") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosorcida, label="Afrosorcida",
                  fontsize = 4, align = T, angle=0, barsize = 1,
                  color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Afrosorcida") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                  ) +
    geom_cladelabel(node=bsize.Atlantogenata.nodes$Macroscelidae, 
                    label="Macroscelidae", 
                    fontsize = 4, align = T, angle=0, barsize = 1,
                    color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Macroscelidae") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                    ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Xenarthra, label="Xenarthra", 
                  fontsize = 4, align = T, angle=0, barsize = 1,
                  color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                  ) +
  scale_color_viridis_c(limits=c(scale.log2CancerSucceptabilityChange.min, scale.log2CancerSucceptabilityChange.max), 
                        guide = guide_colourbar(title="Log2 Cancer Succeptibility Change"))+
  xlim(0,1500) + 
  theme_tree2() + 
  theme(
    legend.position="right",
    panel.background = element_rect(fill = "lightgrey")
  ) + 
  xlab("Rate of Change, Body Size")
# f5b

f5 <- ggarrange(
  f5a, 
  #nrow=2, ncol=1,
  labels="AUTO"
  )

cat(t.size.lifespan.Atlantogenata.lm, sep = "\n")
f5
```

`r warning("TODO: internally add ")`
The dramatic increase in body mass and lifespan in some Afrotherian lineages implies those lineages evolved reduced cancer risk. To infer the magnitude of these reductions we estimated differences in cancer risk between small bodied, short-lived species and large bodied, long-lived species as well as for reconstructed ancestral Afrotherians. Following [@Peto2015b] we estimate the intrinsic cancer risk as the product of risk associated with body mass and lifespan. Differences in cancer susceptibility due to body mass differences between species can be approximated simply as the fold difference in body mass ($D$) between species [@Peto2015]. The risk of developing cancer also increases in proportion to the sixth power of age and is approximated by the formula $Ct^6$, in which the proportionality constant C that determines susceptibility to cancer induction is multiplied by the sixth power of the age in years, $t$ [@ArmitageAndDoll1954; @Nordling1953; @Peto2015]. Thus we can estimate differences in the intrinsic cancer risk between species as $D(Ct^6)$. We inferred that susceptibility to cancer induction ($C$) was reduced 3.40x10^3-fold in the ancestor of Pseudoungulata, 7.20×10^3-fold in the ancestor of Paenungulata, 8.81x10^5-fold in the ancestor of Tehthytheria, 1.46x10^8-fold in the ancestor of Proboscidea, and 1.28x10^6-fold in the ancestor of Sirenia relative to the ancestral Afrotherian (Table 3). What biological mechanisms underlie the evolution of thousand- to hundred million-fold reductions in cancer susceptibility during the origins of Afrotherians, which are essential for large body size and long lifespan to evolve

# Discussion

`r warning("TODO: Discuss Body Size & implications of acordioning of Body Size")`
`r warning("TODO: Discuss Copy Numbers, specifically how rapidly some gene families expanded, and the issue of genome quality")`
`r warning("TODO: Discuss effect of genome quality with CNs")`
`r warning("I placed this here, but its half-discussion, half-intro, so edit!")`
Candidate gene studies, for example, have identified functional duplicates of the tumor suppressors TP53  and LIF in elephants. In a larger candidate gene study, Caulin et al. characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans. However, the probability of developing cancer is similar for small, short-lived mammals such as mice and for large, long-lived mammals such as elephants.

In stark contrast, genome-wide studies of unusually large or long-lived species such as the bowhead whale (Keane et al., 2015), Myotid bats (Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al., 2011), and blind mole rat (Fang et al., 2014) did not find an over representation of tumor suppressors among duplicate genes.

A genomic analysis of genetic changes associated with the evolution of enhanced cancer resistance in the elephant lineage has yet to be performed. Thus it is not clear if the duplication of TP53 and LIF reflects a general pattern of tumor suppressor duplication in the elephant lineage, unlike other lineages that resolved Peto?s paradox, or from the kinds of ascertainment biases common in candidate gene studies.

`r warning("TODO: Discuss ORAs + cancer risk")`

# Acknowledgements
I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for his indispensible support.

# Conflicts of Interest
The Authors have no conflicts of interest.

#

# SUPPLEMENTARY FIGURES
<!-- ```{r SFig 1, fig.cap="As expected, there was a correlation between the number of hits identified in a species with the evolutionary distance relative to humans"} -->
<!--  -->
<!-- ``` -->
`r warning("TODO: I might have made more figures. Rename this after verifying that all figures are included somewhere in the paper.")`

```{r Supplementary Figure 1, fig.cap="Supplementary Figure 1: Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs. __A)__ A single, contiguous gene homolog in a target genome with 100% query length coverage has an ECNC of 1.0. __B)__ Two contiguous gene homologs, each with 100% query length coverage have an ECNC of 2.0. __C)__ A single gene homolog, split across multiple scaffolds and contigs in a fragmented target genome; BLAT identifies each fragment as a single hit. Per nucleotide of query sequence, there is only one corresponding nucleotide over all the hits, thus the ECNC is 1.0. __D)__ Two gene homologs, one fragmented and one contiguous. 100% of nucleotides in the query sequence are represented between all hits; however, every nucleotide in the query has two matching nucleotides in the target genome, thus the ECNC is 2.0. __E)__ One true gene homolog in the target genome, plus multiple hits of a conserved domain that span 20% of the query sequence. While 100% of the query sequence is represented in total, 20% of the nucleotides have 4 hits. Thus, the ECNC for this gene is 1.45. __F)__ Two real gene homologs; one hit is contiguous, one hit is fragmented in two, and the tail end of both sequences was not identified by BLAT due to sequence divergence. Only 75% of the query sequence was covered in total between the hits, but for that 75%, each nucleotide has two hits. As such, ECNC is equal to 2.0 for this gene."}
sf1 <- list(
    ecnc.onegene.onehit, #A
    ecnc.twogene.twohit, #B
    ecnc.onegene.threehits, #C 
    ecnc.twogene.fourhits, #D
    ecnc.onegene.manypartialhits, #E
    ecnc.twogene.partialhits #F
    ) %>% 
  lapply(
    function(p){
      p+theme_pubclean()
    }
  ) %>% 
  ggarrange(
    plotlist = .,
    ncol=3,
    nrow=2,
    labels = "AUTO"
    )
#A:"ECNC: One Contiguous Gene, One Hit"
#B: "ECNC: Two Contiguous Genes, Two Hits"
#C: "ECNC: One Fragmented Gene, Multiple Hits"
#D: "ECNC: Two Genes, Multiple Hits"
#E: "ECNC: One Real Gene, Multiple Hits on Conserved Domain"
#F: "ECNC: Two Real Genes, Unidentified Tails"
sf1
```

```{r, Supplementary Figure 2, fig.cap="Supplementary Figure 2: Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number"}
p.ml.tree.increasing.geneCounts.withSize.withBL
```

```{r, Supplementary Figure 3, fig.cap="Supplementary Figure 3: Full version of RecBlat strategy, but low quality"}
warning("TODO: Make HQ version of this figure")
p.bodysize.atlantogenata

```

```{r, Supplementary Figure 4, fig.cap="Supplementary Figure 4: Dot and Line plot for Body Sizes and such"}
p.bodysize.atlantogenata
```

```{r Supplementary Figure 5, fig.cap="Supplementary Figure 5: All the Gene Copy Trees for interesting genes duplicated in LoxAfr4 (RIP any trees if this gets printed)"}
p.ml.treedata.loxAfrHighlight.plus 
```

```{r Supplementary Figure 6, fig.cap="Supplementary Figure 6: Expression data for interesting genes duplicated in LoxAfr4"}
plot.coolgenes
```

```{r Supplementary Figure 7, fig.cap="Supplementary Figure 7: Correlation matrix heat map for genome quality metrics, ECNC, RBHB Copy, and Estimated Copy Number (lesser between ECNC and RBHB)"}
p.cor.GCT.CEGMA.completeOnly
```

# References {#references .unnumbered}

### 2020-05-10 paper_PLOS/paper_PLOS_draft.tex
first draft

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{University of Chicago,University at Buffalo, SUNY}
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Pervasive duplication of tumor suppressor genes preceded parallel
evolution of large bodied Paenungulates} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Manuel Vazquez\textsuperscript{\getAff{University of Chicago}}\textsuperscript{*},
Vincent J Lynch\textsuperscript{\getAff{University at Buffalo, SUNY}}\\
\bigskip
\textbf{\getAff{University of Chicago}}Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637\\
\textbf{\getAff{University at Buffalo, SUNY}}Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260\\
\bigskip
* Corresponding author: juanvazquez@uchicago.edu\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
)

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

One of the major constraints on the evolution of large body sizes in
animals is an increased risk of developing cancer. If all cells in all
organisms have a similar risk of malignant transformation and equivalent
cancer suppression mechanisms, organism with many cells should have a
higher prevalence of cancer than organisms with fewer cells. Consistent
with this expectation there is a strong positive correlation between
body size and cancer incidence within species, for example, human cancer
incidence increases with increasing adult height {[}1,2{]} and cancer
incidence is positively correlated with body size in dogs
{[}{\textbf{???}},3{]}. There is no correlation, however, between body
size and cancer risk between species. This lack of correlation is often
referred to as `Peto's Paradox' {[}4--6{]}. While it is clear that a
resolution to Peto's Paradox must involve the evolution of enhanced
cancer protection alongside increases in body size and lifespan, the
specific genetic, molecular, and cellular mechanisms that underlie this
resistance have proven elusive. {[}7--11{]}.

Among the challenges for discovering how animals evolved enhanced cancer
protection mechanisms is identifying lineages in which large bodied
species are nested within species with small body sizes. Afrotherian
mammals are generally small-bodied, similarly to the prediced common
ancestor of Eutherian mammals. For example, maximum adult weights are
\textasciitilde{}70g in golden moles, \textasciitilde{}120g in tenrecs,
\textasciitilde{}170g in elephant shrews, \textasciitilde{}3kg in
hyraxes, and 60kg in aardvarks {[}12{]}. However, while these extant
species are relatively small, the fossil evidence demonstrates that
their ancestral lineages reached enormous sizes. For example, while
extant hyraxes are relatively small, the extinct Titanohyrax is
estimated to have weighted up to \textasciitilde{}1300kg {[}13{]}. The
largest members of Afrotheria, too, are dwarfed by the size of their
recent ancestors: extant cows manatees are large bodied
(\textasciitilde{}322-480kg) but are relatively small compared to the
extinct Stellar's sea cow which is estimated to have weight 8000-10000kg
{[}14{]}. Similarly African (4,800kg) and Asian elephants (3,200kg) are
the largest living elephant species, but are dwarfed by the truly
gigantic extinct Proboscideans such as Deinotherium
(\textasciitilde{}132,000kg), Mammut borsoni (110,000kg), and the Asian
straight-tusked elephant (\textasciitilde{}220,000kg), the largest known
land mammal {[}15{]}. Remarkably these large-bodied Afrotherian lineages
are nested within small bodied species (Fig. 1) {[}16--19{]}, indicating
that gigantism independently evolved in hyraxes, sea cows, and elephants
(Paenungulates). Thus, Paenungulates are an excellent model system in
which to explore the mechanisms that underlie the evolution of large
body sizes and augmented cancer resistance.

Although many mechanisms can potentially resolve Peto's paradox, the
most parsimonious route to enhanced cancer resistance is likely through
an increased copy number of tumor suppressors. Such an example has been
seen in the case of candidate genes such as \emph{TP53} and \emph{LIF}
{[}11,20,21{]} as well as in studies involving a limited set of
candidate genes {[}22,23{]}. As these studies focus on \emph{a priori}
gene sets, however, it remains unknown whether this is a general,
genome-wide trend in Afrotherian genomes; and whether such a general
trend is associated with the recent increases in body size -- and
therefore expected cancer risk -- in these species.

Here, we trace the evolution of body mass and gene copy number variation
in Afrotherians in order to investigate whether gene duplications are
enriched in large, long-lived species for genes involved in known tumor
suppression pathways. Our estimates of the evolution of body mass,
similarly to previous studies {[}16--19{]}, show that large body masses
evolved in a step-wise manner, with major increases in body mass in the
Pseudoungulata (17kg), Paenungulata (25kg), Tethytheria (296kg), and
Proboscidea (4,100kg) stem-lineages. Furthermore, we see that the
ancestral body size increases in Hydracoidia and Sirenia were
independent events. To study the evolution of gene copy number, we used
a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene
duplications in Afrotherian genomes, and used parsimony to infer the
lineages in which those duplications occurred. We found gene
duplications in lineages with increased body mass were enriched in
functions related to tumor suppression, including regulation of the cell
cycle, DNA damage repair, and regulation of apoptosis. These data
suggest that duplication of tumor suppressors played a role in the
evolution of large, long-lived in Afrotherians.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{ancestral-body-size-reconstruction}{%
\subsection{Ancestral Body Size
Reconstruction}\label{ancestral-body-size-reconstruction}}

We built a time-calibrated supertree of Eutherian mammals by combining
the time-calibrated molecular phylogeny of Bininda-Emonds \emph{et al.}
{[}24{]} with the time-calibrated total evidence Afrotherian phylogeny
from Puttick and Thomas {[}{\textbf{???}}{]}. While the Bininda-Emonds
\emph{et al.} {[}24{]} phylogeny includes 1,679 species, only 34 are
Afrotherian, and no fossil data are included. The inclusion of fossil
data from extinct species is essential to ensure that ancestral state
reconstructions of body mass are not biased by only including extant
species. This can lead to inaccurate reconstructions, for example, if
lineages convergently evolved large body masses from a small bodied
ancestor. In contrast, the total evidence Afrotherian phylogeny of
Puttick and Thomas {[}19{]} includes 77 extant species and fossil data
from 39 extinct species. Therefore we replaced the Afrotherian clade in
the Bininda-Emonds \emph{et al.} {[}24{]} phylogeny with the Afrotherian
phylogeny of Puttick and Thomas {[}19{]} using Mesquite. Next, we
jointly estimated rates of body mass evolution and reconstructed
ancestral states using a generalization of the Brownian motion model
that relaxes assumptions of neutrality and gradualism by considering
increments to evolving characters to be drawn from a heavy-tailed stable
distribution (the ``Stable Model'') {[}25{]}. The stable model allows
for occasional large jumps in traits and has previously been shown to
out-perform other models of body mass evolution, including standard
Brownian motion models, Ornstein--Uhlenbeck models, early burst maximum
likelihood models, and heterogeneous multi-rate models {[}25{]}.

\hypertarget{identification-of-duplicate-genes}{%
\subsection{Identification of Duplicate
Genes}\label{identification-of-duplicate-genes}}

\emph{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit BLAT
(RBHB) pipeline to quickly identify homologs and estimate gene copy
numbers (\textbf{Figure 1A}). The Reciprocal Best Hit (RBH) search
strategy is conceptually straightforward: 1) Given a gene of interest GA
in a query genome A, one searches a target genome B for all possible
matches to GA; 2) For each of these hits, one then performs the
reciprocal search in the original query genome to identify the
highest-scoring hit; 3) A hit in genome B is defined as a homolog of
gene GA if and only if the original gene GA is the top reciprocal search
hit in genome A. We selected BLAT {[}26{]} as our algorithm of choice,
as this algorithm is sensitive to highly simliar (\textgreater{}90\%
identity) sequences, thus identifying the highest-confidence homologs
while minimizing many-to-one mapping problems when searching for
multiple genes. RBH performs similar to other more complex methods of
orthology prediction, and is particularly good at identifying incomplete
genes that may be fragmented in low quality/poor assembled regions of
the genome {[}{\textbf{???}},27{]}.

\emph{Effective Copy Number By Coverage:} In lower-quality genomes, many
genes are fragmented across multiple scaffolds, which results in BLAT
calling multiple hits when in reality there is only one gene. To
compensate for this, we came up with a novel statistic, Estimated Copy
Number by Coverage (ECNC), which averages the number of times we see
each nucleotides of a query sequence in a target genome over the total
number of nucleotides of the query sequence found overall in each target
genome (\textbf{Figure 1B}; Supplementary Figure 1). This allows us to
correct for genes that have been fragmented across incomplete genomes,
without underestimating the copy number of proteins that lack domains
found in the human sequence.

\emph{RecSearch Pipeline:} We created a custom Python pipeline for
automating RBHB searches between a single reference genome and multiple
target genomes using a list of query sequences from the reference
genome. For the query sequences in our search, we used the hg38 Proteome
provided by UniProt {[}28{]}, which is a comprehensive set of protein
sequences curated from a combination of predicted and validated protein
sequences generated by the UniProt Consortium. In order to refine our
search, we omitted protein sequences originating from long, noncoding
RNA loci (e.g.~LINC genes); poorly-studied genes from predicted open
reading frames (C-ORFs); and sequences with highly repetitive sequences
such as zinc fingers, protocadherins, and transposon-containing genes,
as these were prone to high levels of false positive hits. After
filtering out problematic protein queries, we then used our pipeline
(Figure 1A) to search for all copies of our 20456 query genes in
publicly available Afrotherian genomes, including African savannah
elephant (\emph{Loxodonta africana}: loxAfr3, loxAfr4, loxAfrC), African
forest elephant (\emph{Loxodonta cyclotis}: loxCycF), Asian Elephant
(\emph{Elephas maximus}: eleMaxD), Woolly Mammoth (\emph{Mammuthus
primigenius}: mamPriV), Colombian mammoth (\emph{Mammuthus columbi}:
mamColU), American mastodon (\emph{Mammut americanum}: mamAmeI), Rock
Hyrax (\emph{Procavia capensis}: proCap1, proCap2, proCap2\_HiC), West
Indian Manatee (\emph{Trichechus manatus latirostris}: triManLat1,
triManLat1\_HiC), Aardvark (\emph{Orycteropus afer}: oryAfe1,
oryAfe1\_HiC), Lesser Hedgehog Tenrec (\emph{Echinops telfairi}:
echTel2), Nine-banded armadillo (\emph{Dasypus novemcinctus}: dasNov3),
Hoffman's two-toed sloth (\emph{Choloepus hoffmannii}: choHof1, choHof2,
choHof2\_HiC), Cape golden mole (\emph{Chrysochloris asiatica}:
chrAsi1), and Cape elephant shrew (\emph{Elephantulus edwardii}:
eleEdw1). For many of these species, we covered multiple assemblies in
order to test the effects of assembly size and quality on our hits.

\emph{Duplication gene inclusion criteria:} In order to condense
transcript-level hits into single gene loci, and to resolve many-to-one
genome mappings, we removed exons where transcripts from different genes
overlapped, and merged overlapping transcripts of the same gene into a
single gene locus call. The resulting gene-level copy number table was
then combined with the maximum ECNC values observed for each gene in
order to call gene duplications. We called a gene duplicated if its copy
number was two or more, and if the maximum ECNC value of all the gene
transcripts searched was 1.5 or greater; previous studies have shown
that incomplete duplications can encode functional genes, therefore
partial gene duplications were included provided they passed additional
inclusion criteria. The ECNC cut off of 1.5 was selected empirically, as
this value minimized the number of false positives seen in a test set of
genes and genomes. The results of our initial search are summarized in
Figure 1B. Overall, we identified {[}MEDIAN{]} genes across all species,
or {[}\%HITS/QUERIES{]} of our starting query genes. As expected, there
was a correlation between the number of hits identified in a species
with the evolutionary distance relative to humans {[}Sup Figure 1{]}.

\emph{Duplicate gene exclusion criteria:} We excluded genes from
downstream analyses for which assignment of homology was uncertain,
including uncharacterized ORFs (17), LOC (17), HLA genes (17),
replication dependent histones (17), odorant receptors (17), ribosomal
proteins (17), zinc finger transcription factors (17), viral and
repetitive-element-associated proteins (17) and any protein described as
either ``Uncharacterized,'' ``Putative,'' or ``Fragment'' by UniProt in
UP000005640 (17).

\hypertarget{evidence-for-functionality-of-identified-genes}{%
\subsection{Evidence for Functionality of Identified
Genes}\label{evidence-for-functionality-of-identified-genes}}

To validate and filter out RBHB results, we intersected our results with
either gene prediction or transcriptomic evidence as a proxy for
functionality.

\textbf{Transcriptome Assembly:} For the African Savana Elephant, Asian
Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated
\emph{de novo} transcriptomes using publically-available RNA-sequencing
data from NCBI SRA. We mapped reads to all genomes available for each
species, and assembled transcripts using HISAT2 and StringTie,
respectively {[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}.
RNA-sequencing data was not available for Cape Golden Mole, Cape
Elephant Shrew, Rock Hyrax, Aardvark, or the Lesser Hedgehog Tenrec.

\textbf{Gene Prediction:} We obtained tracks for genes predicted using
GenScan for all the genomes available via UCSC Genome Browser: African
savannah elephant (loxAfr3), Rock Hyrax (proCap1), West Indian Manatee
(triManLat1), Aardvark (oryAfe1), Lesser Hedgehog Tenrec (echTel2),
Nine-banded armadillo (dasNov3), Hoffman's Two-Toed Sloth (choHof1),
Cape golden mole (chrAsi1), and Cape Elephant Shrew (eleEdw1); gene
prediction tracks for higher-quality assemblies were not available.

\textbf{Evidenced Duplicate Criteria:} We intersected our records of
duplicate hits identified in each genome with the gene prediction tracks
and/or transcriptome assemblies using \texttt{bedtools}
{[}{\textbf{???}}{]}. When multiple lines of evidence for functionality
were present for a genome, we used the union of all intersections as the
final output for evidenced duplicates. When analyzing the
highest-quality assemblies available for each species, if a species had
neither gene prediction tracks nor RNA-seq data for the highest-quality
genome available, we conservatively included all hits for the genome in
the final set of evidenced duplicates.

\hypertarget{reconstruction-of-ancestral-copy-numbers}{%
\subsection{Reconstruction of Ancestral Copy
Numbers}\label{reconstruction-of-ancestral-copy-numbers}}

We implemented a maximum likelihood method for determining the ancestral
copy numbers of genes in \emph{Atlantogenata} using IQ-Tree. For this
analysis, we used an unrooted subset of our prior species tree,
including only the aforementioned \emph{Atlantogenata} species. We
generated PHYLIP files containing the copy number of each gene in the
highest quality genome for each species, encoding genes on a scale from
1-31+ copies as 1-9, A-V; and encoding a gene's copy number as uncetain
(``?'') when we did not identify it in the genome. We used the included
tree-searching and model-testing functionality in IQ-Tree to determine
the most likely topology for the species tree, and to obtain the most
likely model for copy number changes in the genome. We defined the
ancestral state of a node if it had greater than an 80\% posterior
probability.

\hypertarget{pathway-enrichment-analysis}{%
\subsection{Pathway Enrichment
Analysis}\label{pathway-enrichment-analysis}}

To determine which pathways were associated with duplicated genes in
each species and lineage, we used WEBGESTALT to perform
overrepresentation analysis (ORA) of the duplicated gene lists relative
to our initial query gene list. For the database of pathways used in the
analysis, we used Reactome {[}{\textbf{???}}{]}, Wikipathways and
Wikipathways\_cancer {[}{\textbf{???}}{]}, and KEGG
{[}{\textbf{???}}{]}.

\hypertarget{results}{%
\section{Results}\label{results}}

\begin{verbatim}
## Warning: Tried to calculate with group_by(), but the calculation failed.
## Falling back to ungrouped filter operation...

## Warning: Tried to calculate with group_by(), but the calculation failed.
## Falling back to ungrouped filter operation...
\end{verbatim}

\begin{verbatim}
## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
##   If you'd like to see this geom implemented,
##   Please open an issue with your example code at
##   https://github.com/ropensci/plotly/issues

## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
##   If you'd like to see this geom implemented,
##   Please open an issue with your example code at
##   https://github.com/ropensci/plotly/issues

## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
##   If you'd like to see this geom implemented,
##   Please open an issue with your example code at
##   https://github.com/ropensci/plotly/issues
\end{verbatim}

\hypertarget{step-wise-evolution-of-large-long-lived-afrotherians}{%
\subsection{Step-wise evolution of large, long-lived
Afrotherians}\label{step-wise-evolution-of-large-long-lived-afrotherians}}

\begin{verbatim}
## Warning: Removed 120 rows containing missing values (geom_image).
\end{verbatim}

\begin{longtable}[]{@{}lrrrr@{}}
\caption{A table generated by the longtable package.}\tabularnewline
\toprule
Ancestor/Species & Estimated Body Size (log(g)) & 95\% CI (Low) & 95\%
CI (High) & Rate (sqrt)\tabularnewline
\midrule
\endfirsthead
\toprule
Ancestor/Species & Estimated Body Size (log(g)) & 95\% CI (Low) & 95\%
CI (High) & Rate (sqrt)\tabularnewline
\midrule
\endhead
Cryptochloris wintoni & 3.13 & 3.13 & 3.13 & 5.78\tabularnewline
Amblysomus marleyi & 3.53 & 3.53 & 3.53 & 3.79\tabularnewline
Elephantulus revoili & 3.48 & 3.48 & 3.48 & 1.10\tabularnewline
Titanohyrax andrewsi & 12.97 & 12.97 & 12.97 & 0.07\tabularnewline
Titanohyrax ultimus & 14.08 & 14.08 & 14.08 & 34.61\tabularnewline
Megalohyrax sp nov & 12.52 & 12.52 & 12.52 & 7.21\tabularnewline
Elephas maximus asurus & 15.66 & 15.66 & 15.66 & 0.34\tabularnewline
Protenrec tricuspis & 1.14 & 1.14 & 1.14 & 69.75\tabularnewline
Microgale parvula & 1.16 & 1.16 & 1.16 & 33.46\tabularnewline
Microgale pusilla & 1.25 & 1.25 & 1.25 & 34.31\tabularnewline
Geogale aurita & 1.90 & 1.90 & 1.90 & 40.07\tabularnewline
Microgale longicaudata & 2.09 & 2.09 & 2.09 & 0.77\tabularnewline
Microgale brevicaudata & 2.19 & 2.19 & 2.19 & 0.60\tabularnewline
Microgale jobihely & 2.30 & 2.30 & 2.30 & 1.07\tabularnewline
Microgale principula & 2.32 & 2.32 & 2.32 & 0.17\tabularnewline
Dilambdogale gheerbranti & 2.38 & 2.38 & 2.38 & 2.21\tabularnewline
Microgale taiva & 2.47 & 2.47 & 2.47 & 0.13\tabularnewline
Microgale cowani & 2.62 & 2.62 & 2.62 & 0.57\tabularnewline
Eremitalpa granti & 3.14 & 3.14 & 3.14 & 9.65\tabularnewline
Calcochloris obtusirostris & 3.27 & 3.27 & 3.27 & 13.38\tabularnewline
Neamblysomus julianae & 3.33 & 3.33 & 3.33 & 5.72\tabularnewline
Chlorotalpa duthieae & 3.38 & 3.38 & 3.38 & 0.32\tabularnewline
Chlorotalpa sclateri & 3.54 & 3.54 & 3.54 & 0.09\tabularnewline
Macroscelides proboscideus & 3.64 & 3.64 & 3.64 & 14.17\tabularnewline
Chrysochloris stuhlmanni & 3.74 & 3.74 & 3.74 & 0.33\tabularnewline
Oryzorictes hova & 3.79 & 3.79 & 3.79 & 22.77\tabularnewline
Elephantulus myurus & 3.81 & 3.81 & 3.81 & 0.95\tabularnewline
Elephantulus brachyrhynchus & 3.81 & 3.81 & 3.81 & 0.93\tabularnewline
Elephantulus rozeti & 3.81 & 3.81 & 3.81 & 10.51\tabularnewline
Elephantulus fuscus & 3.82 & 3.82 & 3.82 & 0.68\tabularnewline
Elephantulus intufi & 3.82 & 3.82 & 3.82 & 1.15\tabularnewline
Microgale talazaci & 3.88 & 3.88 & 3.88 & 61.40\tabularnewline
Chrysochloris asiatica & 3.89 & 3.89 & 3.89 & 3.34\tabularnewline
Elephantulus edwardii & 3.90 & 3.90 & 3.90 & 0.24\tabularnewline
Carpitalpa arendsi & 3.94 & 3.94 & 3.94 & 0.45\tabularnewline
Amblysomus corriae & 3.94 & 3.94 & 3.94 & 0.98\tabularnewline
Amblysomus hottentotus & 3.98 & 3.98 & 3.98 & 0.02\tabularnewline
Elephantulus fuscipes & 4.04 & 4.04 & 4.04 & 1.93\tabularnewline
Elephantulus rufescens & 4.05 & 4.05 & 4.05 & 0.12\tabularnewline
Neamblysomus gunningi & 4.09 & 4.09 & 4.09 & 3.26\tabularnewline
Elephantulus rupestris & 4.12 & 4.12 & 4.12 & 0.32\tabularnewline
Amblysomus septentrionalis & 4.23 & 4.23 & 4.23 & 0.52\tabularnewline
Chambius kasserinensis & 4.27 & 4.27 & 4.27 & 11.84\tabularnewline
Amblysomus robustus & 4.33 & 4.33 & 4.33 & 1.38\tabularnewline
Micropotamogale lamottei & 4.36 & 4.36 & 4.36 & 2.82\tabularnewline
Echinops telfairi & 4.47 & 4.47 & 4.47 & 7.75\tabularnewline
Limnogale mergulus & 4.52 & 4.52 & 4.52 & 121.95\tabularnewline
Hemicentetes semispinosus & 4.75 & 4.75 & 4.75 & 4.68\tabularnewline
Chrysospalax villosus & 4.77 & 4.77 & 4.77 & 0.13\tabularnewline
Petrodromus tetradactylus & 5.29 & 5.29 & 5.29 & 24.61\tabularnewline
Herodotius pattersoni & 5.50 & 5.50 & 5.50 & 11.64\tabularnewline
Setifer setosus & 5.61 & 5.61 & 5.61 & 12.52\tabularnewline
Rhynchocyon cirnei & 5.86 & 5.86 & 5.86 & 3.30\tabularnewline
Metoldobotes sp nov & 5.93 & 5.93 & 5.93 & 15.94\tabularnewline
Chrysospalax trevelyani & 6.13 & 6.13 & 6.13 & 62.84\tabularnewline
Rhynchocyon petersi & 6.15 & 6.15 & 6.15 & 2.13\tabularnewline
Rhynchocyon chrysopygus & 6.28 & 6.28 & 6.28 & 0.40\tabularnewline
Potamogale velox & 6.49 & 6.49 & 6.49 & 103.04\tabularnewline
Rhynchocyon udzungwensis & 6.57 & 6.57 & 6.57 & 4.33\tabularnewline
Tenrec ecaudatus & 6.75 & 6.75 & 6.75 & 79.50\tabularnewline
Dasypus sabanicola & 7.05 & 7.05 & 7.05 & 12.18\tabularnewline
Tolypeutes matacus & 7.11 & 7.11 & 7.11 & 15.96\tabularnewline
Dasypus septemcinctus & 7.30 & 7.30 & 7.30 & 4.44\tabularnewline
Zaedyus pichiy & 7.31 & 7.31 & 7.31 & 5.54\tabularnewline
Dasypus hybridus & 7.31 & 7.31 & 7.31 & 4.05\tabularnewline
Chaetophractus villosus & 7.61 & 7.61 & 7.61 & 0.42\tabularnewline
Chaetophractus nationi & 7.67 & 7.67 & 7.67 & 0.09\tabularnewline
Heterohyrax brucei & 7.78 & 7.78 & 7.78 & 1.64\tabularnewline
Cabassous centralis & 7.92 & 7.92 & 7.92 & 0.25\tabularnewline
Seggeurius amourensis & 7.98 & 7.98 & 7.98 & 2.82\tabularnewline
Procavia capensis & 8.01 & 8.01 & 8.01 & 0.00\tabularnewline
Dendrohyrax dorsalis & 8.06 & 8.06 & 8.06 & 1.86\tabularnewline
Microhyrax lavocati & 8.13 & 8.13 & 8.13 & 0.73\tabularnewline
Bradypus tridactylus & 8.23 & 8.23 & 8.23 & 0.48\tabularnewline
Bradypus torquatus & 8.27 & 8.27 & 8.27 & 0.03\tabularnewline
Dasypus novemcinctus & 8.37 & 8.37 & 8.37 & 14.73\tabularnewline
Euphractus sexcinctus & 8.43 & 8.43 & 8.43 & 14.99\tabularnewline
Choloepus hoffmanni & 8.47 & 8.47 & 8.47 & 0.32\tabularnewline
Bradypus variegatus & 8.49 & 8.49 & 8.49 & 0.51\tabularnewline
Tamandua tetradactyla & 8.52 & 8.52 & 8.52 & 10.44\tabularnewline
Cyclopes didactylus & 8.53 & 8.53 & 8.53 & 2.15\tabularnewline
Choloepus didactylus & 8.71 & 8.71 & 8.71 & 0.64\tabularnewline
Thyrohyrax meyeri & 8.78 & 8.78 & 8.78 & 3.55\tabularnewline
Saghatherium bowni & 9.13 & 9.13 & 9.13 & 15.85\tabularnewline
Dasypus kappleri & 9.23 & 9.23 & 9.23 & 74.13\tabularnewline
Thyrohyrax domorictus & 9.30 & 9.30 & 9.30 & 1.15\tabularnewline
Dimaitherium patnaiki & 9.57 & 9.57 & 9.57 & 18.23\tabularnewline
Phosphatherium escuilliei & 9.62 & 9.62 & 9.62 & 326.23\tabularnewline
Saghatherium antiquum & 9.73 & 9.73 & 9.73 & 2.90\tabularnewline
Thyrohyrax litholagus & 10.01 & 10.01 & 10.01 & 28.58\tabularnewline
Myrmecophaga tridactyla & 10.26 & 10.26 & 10.26 & 41.03\tabularnewline
Myorycteropus africanus & 10.27 & 10.27 & 10.27 & 0.57\tabularnewline
Selenohyrax chatrathi & 10.73 & 10.73 & 10.73 & 14.99\tabularnewline
Priodontes maximus & 10.82 & 10.82 & 10.82 & 268.43\tabularnewline
Orycteropus afer & 10.87 & 10.87 & 10.87 & 6.59\tabularnewline
Antilohyrax pectidens & 10.93 & 10.93 & 10.93 & 13.69\tabularnewline
Bunohyrax fajumensis & 11.32 & 11.32 & 11.32 & 1.45\tabularnewline
Afrohyrax championi & 11.32 & 11.32 & 11.32 & 0.19\tabularnewline
Geniohyus mirus & 11.33 & 11.33 & 11.33 & 5.44\tabularnewline
Prorastomus sirenoides & 11.49 & 11.49 & 11.49 & 13.61\tabularnewline
Elephas antiquus falconeri & 11.51 & 11.51 & 11.51 & 6.12\tabularnewline
Pachyhyrax crassidentatus & 11.81 & 11.81 & 11.81 & 2.29\tabularnewline
Megalohyrax eocaenus & 11.95 & 11.95 & 11.95 & 0.24\tabularnewline
Elephas cypriotes & 12.21 & 12.21 & 12.21 & 1.90\tabularnewline
Bunohyrax major & 12.36 & 12.36 & 12.36 & 11.39\tabularnewline
Titanohyrax angustidens & 12.48 & 12.48 & 12.48 & 0.04\tabularnewline
Daouitherium rebouli & 12.80 & 12.80 & 12.80 & 0.74\tabularnewline
Arcanotherium savagei & 12.89 & 12.89 & 12.89 & 7.29\tabularnewline
Dugong dugon & 12.92 & 12.92 & 12.92 & 5.85\tabularnewline
Trichechus senegalensis & 13.03 & 13.03 & 13.03 & 0.57\tabularnewline
Trichechus inunguis & 13.08 & 13.08 & 13.08 & 0.69\tabularnewline
Protosiren smithae & 13.20 & 13.20 & 13.20 & 33.69\tabularnewline
Numidotherium koholense & 13.23 & 13.23 & 13.23 & 2.29\tabularnewline
Omanitherium dhofarensis & 13.35 & 13.35 & 13.35 & 0.03\tabularnewline
Trichechus manatus & 13.44 & 13.44 & 13.44 & 1.39\tabularnewline
Moeritherium spp & 13.82 & 13.82 & 13.82 & 5.71\tabularnewline
Phiomia spp & 13.89 & 13.89 & 13.89 & 3.64\tabularnewline
Elephas maximus & 15.02 & 15.02 & 15.02 & 5.81\tabularnewline
Barytherium spp & 15.20 & 15.20 & 15.20 & 73.58\tabularnewline
Mammuthus primigenius & 15.27 & 15.27 & 15.27 & 2.17\tabularnewline
Mammut borsoni & 16.49 & 16.49 & 16.49 & 15.33\tabularnewline
Mammuthus trogontherii & 16.38 & 16.38 & 16.38 & 16.00\tabularnewline
Loxodonta africana & 15.35 & 15.35 & 15.35 & 1.28\tabularnewline
Loxodonta cyclotis & 15.37 & 15.37 & 15.37 & 3.72\tabularnewline
Palaeoloxodon antiquus & 16.14 & 16.14 & 16.14 & 0.01\tabularnewline
Palaeoloxodon namadicus & 16.81 & 16.81 & 16.81 & 12.81\tabularnewline
Mammut americanum & 15.61 & 15.61 & 15.61 & 0.95\tabularnewline
Mammuthus columbi & 15.71 & 15.71 & 15.71 & 0.91\tabularnewline
Hydrodamalis gigas & 15.72 & 15.72 & 15.72 & 172.52\tabularnewline
Atlantogenata & 5.55 & 4.06 & 7.95 & 0.03\tabularnewline
Afrotheria & 5.55 & 4.05 & 7.96 & 0.00\tabularnewline
Afrosorcida & 4.35 & 2.58 & 6.13 & 44.49\tabularnewline
Macroscelidae & 5.27 & 3.98 & 6.85 & 2.49\tabularnewline
Pseudoungulata & 9.76 & 5.21 & 12.78 & 545.83\tabularnewline
Paenungulata & 10.13 & 7.24 & 13.02 & 4.42\tabularnewline
Tethytheria & 12.60 & 10.25 & 13.81 & 187.47\tabularnewline
Proboscidae & 15.23 & 14.22 & 16.24 & 30.28\tabularnewline
Elephantidae & 15.49 & 14.89 & 16.10 & 2.21\tabularnewline
Elephantina & 15.51 & 15.08 & 15.96 & 0.01\tabularnewline
Mammuthus & 15.54 & 15.24 & 15.85 & 0.47\tabularnewline
Loxodontini & 15.55 & 15.02 & 16.11 & 0.11\tabularnewline
Loxodona & 15.72 & 15.16 & 16.30 & 0.86\tabularnewline
Xenarthra & 7.57 & 5.96 & 9.18 & 124.94\tabularnewline
\bottomrule
\end{longtable}

To trace the evolutionary history of body mass and lifespan in
Afrotherians, we built a time-calibrated supertree of Eutherian mammals
combining 1,679 species from Bininda-Emonds et al {[}24{]} with a total
evidence Afrotherian phylogeny including 77 extant and fossil data from
39 extinct species {[}19{]}. Fossil data from extinct species were
included to ensure that ancestral state reconstructions of body mass in
Afrotherians were not biased by only including extant species, which can
lead to inaccurate reconstructions, for example, if lineages multiple
lineages evolved large body masses from a small bodied ancestor. We
jointly estimated rates of body mass evolution and reconstructed
ancestral states using a generalization of a Brownian model of character
evolution, which allows for occasional large jumps in traits (stable
model) and out performs standard Brownian motion and Ornstein?Uhlenbeck
models of character evolution {[}25{]}.

Similar to previous studies of Afrotherian body size {[}19,25{]}, we
found that the body mass of the Afrotherian ancestor was inferred to be
small (0.26kg, 95\% CI: 0.31-3.01kg) and that substantial accelerations
in the rate of body mass evolution occurred coincident with a 65×
increase in body mass in the stem-lineage of \emph{Pseudoungulata}
(17kg), a 1.5× increase in body mass in the stem-lineage of
\emph{Paenungulata} (25kg), a 12× increase in body mass in the
stem-lineage of \emph{Tehthytheria} (296kg), and a 14× increase in body
mass in the stem-lineage of \emph{Proboscidea} (4,100kg; Figure 1). The
ancestral \emph{Hyracoidea} was inferred to be relatively small
(2.86-15.71kg), and rate accelerations were coincident with independent
body mass increases in large hyraxes such as \emph{Titanohyrax andrewsi}
(67× increase in body mass). While the body mass of the ancestral
\emph{Sirenian} was inferred to be large (61-656kg), a rate acceleration
occurred coincident with a 10× body mass increase in Stellar's sea cow.
Rate accelerations also occurred coincident with 36× body mass reduction
in the stem-lineage of the dwarf elephants \emph{Elephas}
(\emph{Palaeoloxodon}) \emph{falconeri} and \emph{Palaeoloxodon
cypriotes}. These data suggest that gigantism in \emph{Afrotherians}
evolved step-wise, from small to medium bodies in the
\emph{Pseudoungulata} stem-lineage, medium to large bodies in the
\emph{Tehthytherian} stem-lineage and extinct hyraxes, and from large to
exceptionally large bodies independently in the \emph{Proboscidean}
stem-lineage and Stellar's sea cow (Figure 1).

\hypertarget{pervasive-duplication-of-tumor-suppressors-during-the-origins-of-large-bodied-afrotheirans}{%
\subsection{Pervasive duplication of tumor suppressors during the
origins of large bodied
Afrotheirans}\label{pervasive-duplication-of-tumor-suppressors-during-the-origins-of-large-bodied-afrotheirans}}

\begin{verbatim}
## Warning: Column `Genome` joining factor and character vector, coercing into
## character vector
\end{verbatim}

\begin{verbatim}
## Warning in melt(., na.rm = T): The melt generic in data.table has been passed a
## matrix and will attempt to redirect to the relevant reshape2 method; please note
## that reshape2 is deprecated, and this redirection is now deprecated as well.
## To continue using melt methods from reshape2 while both libraries are attached,
## e.g. melt.list, you can prepend the namespace like reshape2::melt(.). In the
## next version, this warning will become an error.
\end{verbatim}

\begin{verbatim}
##                                        Var1 Var2        value
## 1                                      Gene Copy  0.011177676
## 2                                    Genome Copy -0.104365173
## 3       Sum length of sequences > 10M  (nt) Copy -0.074960529
## 4                     Longest sequence (nt) Copy -0.067231077
## 5                  N50 sequence length (nt) Copy -0.050741921
## 6                  Base composition (%) (A) Copy -0.001502967
## 7                  Base composition (%) (T) Copy -0.001379093
## 8                  Base composition (%) (G) Copy -0.043271156
## 9                  Base composition (%) (C) Copy -0.042268177
## 10             Base composition (%) (Other) Copy -0.050228064
## 11 # of sequences containing non-ACGTN (nt) Copy -0.050621596
## 12                Mean sequence length (nt) Copy -0.074983744
## 13              Median sequence length (nt) Copy -0.074897040
## 14                         Missing Proteins Copy -0.079692396
## 15                                   #Prots Copy -0.101300994
## 16                             Completeness Copy -0.101321644
## 17                 Base composition (%) (N) Copy  0.019768158
## 18                              totalLength Copy  0.067454870
## 19       Sum length of sequences > 1M  (nt) Copy  0.051354838
## 20               # of sequences > 100K (nt) Copy  0.056049593
## 21                                      L50 Copy  0.082175611
## 22               # of sequences >   1M (nt) Copy  0.074461957
## 23               # of sequences >  10M (nt) Copy  0.075679451
## 24                   Shortest sequence (nt) Copy -0.013216232
## 25                           GC-content (%) Copy -0.064591581
## 26                           # of sequences Copy  0.076630197
## 27               # of sequences >   1K (nt) Copy  0.047474518
## 28               # of sequences >  10K (nt) Copy  0.045583001
## 29                                   #Total Copy  0.116391876
## 30                                  Average Copy  0.124341009
## 31                                   %Ortho Copy  0.120616153
## 32                                     Gene ECNC  0.011137324
## 33                                   Genome ECNC -0.090522003
## 34      Sum length of sequences > 10M  (nt) ECNC -0.061380983
## 35                    Longest sequence (nt) ECNC -0.053145168
## 36                 N50 sequence length (nt) ECNC -0.040007282
## 37                 Base composition (%) (A) ECNC  0.004626621
## 38                 Base composition (%) (T) ECNC  0.004862588
## 39                 Base composition (%) (G) ECNC -0.037787699
## 40                 Base composition (%) (C) ECNC -0.037045693
## 41             Base composition (%) (Other) ECNC -0.040274294
## 42 # of sequences containing non-ACGTN (nt) ECNC -0.040587311
## 43                Mean sequence length (nt) ECNC -0.061569667
## 44              Median sequence length (nt) ECNC -0.060928944
## 45                         Missing Proteins ECNC -0.062209031
## 46                                   #Prots ECNC -0.084046941
## 47                             Completeness ECNC -0.084067608
## 48                 Base composition (%) (N) ECNC  0.013486259
## 49                              totalLength ECNC  0.061175951
## 50       Sum length of sequences > 1M  (nt) ECNC  0.046387082
## 51               # of sequences > 100K (nt) ECNC  0.049326449
## 52                                      L50 ECNC  0.070430706
## 53               # of sequences >   1M (nt) ECNC  0.063725329
## 54               # of sequences >  10M (nt) ECNC  0.062895861
## 55                   Shortest sequence (nt) ECNC -0.012113794
## 56                           GC-content (%) ECNC -0.066249117
## 57                           # of sequences ECNC  0.070289641
## 58               # of sequences >   1K (nt) ECNC  0.042151677
## 59               # of sequences >  10K (nt) ECNC  0.041222984
## 60                                   #Total ECNC  0.102797584
## 61                                  Average ECNC  0.108543367
## 62                                   %Ortho ECNC  0.104394790
## 63                                     Gene   CN  0.013089940
## 64                                   Genome   CN -0.092501845
## 65      Sum length of sequences > 10M  (nt)   CN -0.061610123
## 66                    Longest sequence (nt)   CN -0.052591263
## 67                 N50 sequence length (nt)   CN -0.038770715
## 68                 Base composition (%) (A)   CN  0.006139187
## 69                 Base composition (%) (T)   CN  0.006354445
## 70                 Base composition (%) (G)   CN -0.036741673
## 71                 Base composition (%) (C)   CN -0.035971303
## 72             Base composition (%) (Other)   CN -0.041544748
## 73 # of sequences containing non-ACGTN (nt)   CN -0.041868348
## 74                Mean sequence length (nt)   CN -0.063786387
## 75              Median sequence length (nt)   CN -0.063541494
## 76                         Missing Proteins   CN -0.063703014
## 77                                   #Prots   CN -0.085286259
## 78                             Completeness   CN -0.085306203
## 79                 Base composition (%) (N)   CN  0.012122309
## 80                              totalLength   CN  0.060342667
## 81       Sum length of sequences > 1M  (nt)   CN  0.044889905
## 82               # of sequences > 100K (nt)   CN  0.047874628
## 83                                      L50   CN  0.069838784
## 84               # of sequences >   1M (nt)   CN  0.062839574
## 85               # of sequences >  10M (nt)   CN  0.062215472
## 86                   Shortest sequence (nt)   CN -0.013168670
## 87                           GC-content (%)   CN -0.066875760
## 88                           # of sequences   CN  0.073563144
## 89               # of sequences >   1K (nt)   CN  0.044231140
## 90               # of sequences >  10K (nt)   CN  0.042436480
## 91                                   #Total   CN  0.105113732
## 92                                  Average   CN  0.110764419
## 93                                   %Ortho   CN  0.106456398
\end{verbatim}

\includegraphics{paper_PLOS_draft_files/figure-latex/Correlations between genome quality scores-1.pdf}

\begin{verbatim}
## Warning: Column `Genome`/`BestGenome` joining factor and character vector,
## coercing into character vector
\end{verbatim}

\begin{verbatim}
## Warning: `as.tibble()` is deprecated as of tibble 2.0.0.
## Please use `as_tibble()` instead.
## The signature and semantics have changed, see `?as_tibble`.
## This warning is displayed once every 8 hours.
## Call `lifecycle::last_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{verbatim}
## Warning: Removed 1 rows containing missing values (geom_text).
\end{verbatim}

\begin{figure}
\centering
\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2-1.pdf}
\caption{Figure 2: A Reciprocal Best-Hit BLAT pipeline for identifying
gene copy number in other genomes. A) A graphic summary of the
reciprocal best-hit strategy. B) Estimated Copy Number by Coverage. C)}
\end{figure}

\begin{verbatim}
## Warning: Table 2
\end{verbatim}

Enhanced cancer suppression may have evolved through many mechanisms;
among the most parsimonious is an increase in the copy number of genes
with tumor suppressor functions. Previous studies focusing on candidate
gene studies, for example, have identified increased copy number of the
tumor suppressors \emph{TP53} and \emph{LIF} in elephants
{[}{\textbf{???}},11,21--23{]}. Therefore, in order to test whether this
was a pervasive phenomena genome-wide in \emph{Afrotherians}, we used a
Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in
\emph{Afrotherian} and \emph{Atlantogenatan} genomes (Fig. 2A). Because
RBHB-like approaches can over-estimate copy number when genes are
fragmented or incorrectly assembled across multiple scaffolds, we also
inferred copy number using a complementary method that quantifies the
ratio between observed and expected gene coverage per nucleotide (ECNC)
(Fig. 2B; Sup. Fig. 2A-D). By only including nucleotides from the query
sequence that were observed in the target genome, we also correct for
partial hits where some or all of the homologs of a gene have diverged
from the human homolog.

Because our sequence database included various protein transcripts for
each gene, in order to obtain gene-level copy number information and
elimate any many-to-one mappings of hits, we labeled each exon of every
reciprocal best hit (RBH) with the gene corresponding to the query
transcript and merged all overlapping exons; next, we eliminated any
many-to-one exons that resulted from the previous step. Finally, we
reassembled the gene loci based on the original transcript starts and
ends, and the collapsed exon data, obtaining the full sequence of each
RBH locus. Genes were considered to be duplicated if its copy number via
RBHB was greater than or equal to 2, and the maximum ECNC among all
transcripts prior to filtering was greater than or equal to 1.50. This
cutoff of ECNC was selected to account for truncated gene duplications,
which have been shown to be functional in various examples
{[}{\textbf{???}} examples of this{]}.

To reconcile the Atlantogenatan phylogeny with duplication events, we
used maximum likelihood to reconstruct likely ancestral copy numbers for
each gene at each node in the phylogeny. To define the copy number of a
gene, we conservatively used the lesser value between the RBHB hit
count, and the ECNC value rounded to the nearest whole number. In order
to perform

Next, in order to select genes and duplicates which were likely
functional, we omitted any hits that were not supported by either the
gene prediction method GenScan, or by at least one transcript assembled
from publically-available RNA-seq data.

We describe the number of genes that increased in each lineage in
\emph{Atlantogenata} in Figure 2. Among the genes that increased in copy
number in the elephant lineage are TP53 and LIF, as previously
described. Furthermore, we identify

\hypertarget{duplications-that-occured-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}{%
\subsection{Duplications that occured recently in Probodiscea are
enriched for tumor suppressor
pathways}\label{duplications-that-occured-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}}

In order to infer the functional consequences of these gene
duplications, we tested if duplicate genes were enriched in specific
pathways relative to our initial query set of genes. We used

\hypertarget{retroduplication-of-eef1a1-in-the-stem-lineage-of-tehthytheria}{%
\subsection{Retroduplication of EEF1A1 in the stem-lineage of
Tehthytheria}\label{retroduplication-of-eef1a1-in-the-stem-lineage-of-tehthytheria}}

\begin{verbatim}
## Warning: Fig 3: EEF1A1
\end{verbatim}

Previous studies have identified genes whose transgenic overexpression
delays aging related phenotypes and extends lifespan. Overexpression of
the eukaryotic translation elongation factor 1 alpha 1 (EEF1A1) gene,
for example, extends lifespan in transgenic Drosophila. Like most
mammalian genomes, we found that Afrotherian genomes encoded numerous
EEF1A1 retrogenes (Fig. XA). While the majority of these retrogenes have
premature stop codons and insertions/deletions characteristic of
pseudogenes, at least 13 Tehthytherian-specific EEF1A1 retrogenes have
the potential to encode functional proteins (Fig. XB). Unfortunately it
is difficult to determine if these retrogenes are transcribed because
they have relatively high sequence similarity (83-88\%), suggesting
evidence of transcription using RNA-Seq data is likely an artifact.
Indeed, nearly all of the EEF1A1 retrogene transcripts assembled by
StringTie were confined to their gene body and UTRs and thus lacked
unique sequences. The EEF1A1RTG13 transcript, however, initiates within
an upstream MLT1B retrovirus-like MaLR transposable element and includes
unique 5?- and 3?-UTR sequences, indicating at least one EEF1A1
retrogene is transcribed (Fig. XC). EEF1A1

Candidate gene studies, for example, have identified functional
duplicates of the tumor suppressors TP53 and LIF in elephants. In a
larger candidate gene study, Caulin et al.~characterized the copy number
of 830 known tumor-suppressor genes across 36 mammals and identified 382
putative duplicates, including duplicates in species with large body
sizes and long life-spans. however, the probability of developing cancer
is similar for small, short-lived mammals such as mice and for large,
long-lived mammals such as elephants.

In stark contrast, genome-wide studies of unusually large or long-lived
species such as the bowhead whale (Keane et al., 2015), Myotid bats
(Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al.,
2011), and blind mole rat (Fang et al., 2014) did not find an over
representation of tumor suppressors among duplicate genes.

A genomic analysis of genetic changes associated with the evolution of
enhanced cancer resistance in the elephant lineage has yet to be
performed. Thus it is not clear if the duplication of TP53 and LIF
reflects a general pattern of tumor suppressor duplication in the
elephant lineage, unlike other lineages that resolved Peto?s paradox, or
from the kinds of ascertainment biases common in candidate gene studies.

\hypertarget{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}{%
\subsection{\texorpdfstring{Concerted duplication of TP53 and
TP53-related genes towards
\emph{Probodiscea}}{Concerted duplication of TP53 and TP53-related genes towards Probodiscea}}\label{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}}

\begin{verbatim}
## Warning: Fig4
\end{verbatim}

\hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
\subsection{Step-wise reduction of intrinsic cancer risk in large,
long-lived
Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}

\begin{verbatim}
## # A tibble: 253 x 6
##     parent  node branch.length label                       lnSize Lifespan
##      <int> <int>         <dbl> <chr>                        <dbl>    <dbl>
##   1    135     1     40.1      Geogale.aurita                   2    1.72 
##   2    140     2    122.       Limnogale.mergulus               5    2.30 
##   3    140     3      0.135    Microgale.taiva                  2    1.72 
##   4    139     4     33.5      Microgale.parvula                1    1.52 
##   5    138     5      0.595    Microgale.brevicaudata           2    1.72 
##   6    142     6      0.571    Microgale.cowani                 3    1.91 
##   7    142     7      1.07     Microgale.jobihely               2    1.72 
##   8    145     8      0.772    Microgale.longicaudata           2    1.72 
##   9    145     9      0.173    Microgale.principula             2    1.72 
##  10    144    10     34.3      Microgale.pusilla                1    1.52 
##  11    143    11     61.4      Microgale.talazaci               4    1.76 
##  12    136    12     22.8      Oryzorictes.hova                 4    2.11 
##  13    147    13      7.75     Echinops.telfairi                4    2.94 
##  14    147    14     12.5      Setifer.setosus                  6    2.65 
##  15    148    15      4.68     Hemicentetes.semispinosus        5    0.993
##  16    148    16     79.5      Tenrec.ecaudatus                 7    2.16 
##  17    150    17      2.82     Micropotamogale.lamottei         4    2.11 
##  18    150    18    103.       Potamogale.velox                 6    2.50 
##  19    151    19     69.8      Protenrec.tricuspis              1    1.52 
##  20    151    20      2.21     Dilambdogale.gheerbranti         2    1.72 
##  21    157    21      0.454    Carpitalpa.arendsi               4    2.11 
##  22    158    22      3.26     Neamblysomus.gunningi            4    2.11 
##  23    158    23      5.72     Neamblysomus.julianae            3    1.91 
##  24    162    24      0.976    Amblysomus.corriae               4    2.11 
##  25    162    25      1.38     Amblysomus.robustus              4    2.11 
##  26    161    26      0.524    Amblysomus.septentrionalis       4    2.11 
##  27    160    27      0.0152   Amblysomus.hottentotus           4    2.11 
##  28    159    28      3.79     Amblysomus.marleyi               4    2.11 
##  29    163    29      0.321    Chlorotalpa.duthieae             3    1.91 
##  30    163    30      0.0874   Chlorotalpa.sclateri             4    2.11 
##  31    165    31     62.8      Chrysospalax.trevelyani          6    2.50 
##  32    165    32      0.130    Chrysospalax.villosus            5    2.30 
##  33    164    33     13.4      Calcochloris.obtusirostris       3    1.91 
##  34    153    34      9.65     Eremitalpa.granti                3    1.91 
##  35    167    35      3.34     Chrysochloris.asiatica           4    2.11 
##  36    167    36      5.78     Cryptochloris.wintoni            3    1.91 
##  37    166    37      0.333    Chrysochloris.stuhlmanni         4    2.11 
##  38    175    38      0.926    Elephantulus.brachyrhynchus      4    2.11 
##  39    175    39      0.118    Elephantulus.rufescens           4    2.07 
##  40    176    40      1.15     Elephantulus.intufi              4    2.23 
##  41    176    41      0.319    Elephantulus.rupestris           4    2.11 
##  42    177    42      0.237    Elephantulus.edwardii            4    2.11 
##  43    177    43      0.953    Elephantulus.myurus              4    2.11 
##  44    179    44     10.5      Elephantulus.rozeti              4    1.97 
##  45    179    45     24.6      Petrodromus.tetradactylus        5    2.30 
##  46    178    46     14.2      Macroscelides.proboscideus       4    2.16 
##  47    181    47      1.93     Elephantulus.fuscipes            4    2.11 
##  48    182    48      0.678    Elephantulus.fuscus              4    2.11 
##  49    182    49      1.10     Elephantulus.revoili             3    1.91 
##  50    184    50      0.403    Rhynchocyon.chrysopygus          6    2.40 
##  51    185    51      3.30     Rhynchocyon.cirnei               6    2.50 
##  52    185    52      4.33     Rhynchocyon.udzungwensis         7    2.69 
##  53    183    53      2.13     Rhynchocyon.petersi              6    2.50 
##  54    170    54     15.9      Metoldobotes.sp..nov             6    2.50 
##  55    186    55     11.8      Chambius.kasserinensis           4    2.11 
##  56    186    56     11.6      Herodotius.pattersoni            5    2.30 
##  57    188    57      6.59     Orycteropus.afer                11    3.39 
##  58    188    58      0.569    Myorycteropus.africanus         10    3.27 
##  59    194    59      1.86     Dendrohyrax.dorsalis             8    2.88 
##  60    195    60      1.64     Heterohyrax.brucei               8    2.88 
##  61    195    61      0.000068 Procavia.capensis                8    2.69 
##  62    193    62      1.15     Thyrohyrax.domorictus            9    3.08 
##  63    192    63      3.55     Thyrohyrax.meyeri                9    3.08 
##  64    198    64     28.6      Thyrohyrax.litholagus           10    3.27 
##  65    201    65      2.29     Pachyhyrax.crassidentatus       12    3.66 
##  66    201    66      1.45     Bunohyrax.fajumensis            11    3.47 
##  67    202    67      5.44     Geniohyus.mirus                 11    3.47 
##  68    202    68     11.4      Bunohyrax.major                 12    3.66 
##  69    203    69      0.235    Megalohyrax.eocaenus            12    3.66 
##  70    203    70      7.21     Megalohyrax.sp..nov             13    3.86 
##  71    207    71      0.0699   Titanohyrax.andrewsi            13    3.86 
##  72    207    72     34.6      Titanohyrax.ultimus             14    4.05 
##  73    206    73      0.0353   Titanohyrax.angustidens         12    3.66 
##  74    205    74     13.7      Antilohyrax.pectidens           11    3.47 
##  75    204    75      0.194    Afrohyrax.championi             11    3.47 
##  76    209    76      2.90     Saghatherium.antiquum           10    3.27 
##  77    209    77     15.0      Selenohyrax.chatrathi           11    3.47 
##  78    208    78     15.8      Saghatherium.bowni               9    3.08 
##  79    211    79      2.82     Seggeurius.amourensis            8    2.88 
##  80    211    80      0.733    Microhyrax.lavocati              8    2.88 
##  81    210    81     18.2      Dimaitherium.patnaiki           10    3.27 
##  82    220    82      5.81     Elephas.maximus                 15    4.18 
##  83    222    83      6.12     Elephas.antiquus.falconeri      12    3.66 
##  84    222    84      1.90     Elephas.cypriotes               12    3.66 
##  85    221    85      0.340    Elephas.maximus.asurus          16    4.44 
##  86    223    86     16.0      Mammuthus.trogontherii          16    4.44 
##  87    224    87      0.910    Mammuthus.columbi               16    4.44 
##  88    224    88      2.17     Mammuthus.primigenius           15    4.24 
##  89    225    89      1.28     Loxodonta.africana              15    4.17 
##  90    227    90      0.0149   Palaeoloxodon.antiquus          16    4.44 
##  91    227    91     12.8      Palaeoloxodon.namadicus         17    4.63 
##  92    226    92      3.72     Loxodonta.cyclotis              15    4.24 
##  93    228    93     15.3      Mammut.borsoni                  16    4.44 
##  94    228    94      0.951    Mammut.americanum               16    4.44 
##  95    216    95      3.64     Phiomia.spp                     14    4.05 
##  96    230    96      0.0280   Omanitherium.dhofarensis        13    3.86 
##  97    230    97      7.29     Arcanotherium.savagei           13    3.86 
##  98    229    98     73.6      Barytherium.spp                 15    4.24 
##  99    214    99      5.71     Moeritherium.spp                14    4.05 
## 100    231   100    326.       Phosphatherium.escuilliei       10    3.27 
## 101    232   101      2.29     Numidotherium.koholense         13    3.86 
## 102    232   102      0.742    Daouitherium.rebouli            13    3.86 
## 103    235   103      5.85     Dugong.dugon                    13    4.29 
## 104    235   104    173.       Hydrodamalis.gigas              16    4.44 
## 105    237   105      0.688    Trichechus.inunguis             13    3.69 
## 106    237   106      1.39     Trichechus.manatus              13    4.23 
## 107    236   107      0.574    Trichechus.senegalensis         13    3.86 
## 108    238   108     13.6      Prorastomus.sirenoides          11    3.47 
## 109    238   109     33.7      Protosiren.smithae              13    3.86 
## 110    241   110      0.640    Choloepus.didactylus             9    3.61 
## 111    241   111      0.316    Choloepus.hoffmanni              8    3.71 
## 112    242   112      0.0297   Bradypus.torquatus               8    2.88 
## 113    243   113      0.483    Bradypus.tridactylus             8    2.88 
## 114    243   114      0.512    Bradypus.variegatus              8    2.88 
## 115    245   115     10.4      Tamandua.tetradactyla            9    2.94 
## 116    245   116     41.0      Myrmecophaga.tridactyla         10    3.43 
## 117    244   117      2.15     Cyclopes.didactylus              9    3.08 
## 118    249   118     15.0      Euphractus.sexcinctus            8    3.10 
## 119    249   119      5.54     Zaedyus.pichiy                   7    2.53 
## 120    250   120      0.0899   Chaetophractus.nationi           8    2.88 
## 121    250   121      0.418    Chaetophractus.villosus          8    3.23 
## 122    251   122    268.       Priodontes.maximus              11    2.71 
## 123    252   123     16.0      Tolypeutes.matacus               7    3.61 
## 124    252   124      0.248    Cabassous.centralis              8    2.88 
## 125    253   125      4.05     Dasypus.hybridus                 7    2.69 
## 126    253   126     74.1      Dasypus.kappleri                 9    3.08 
## 127    253   127     14.7      Dasypus.novemcinctus             8    3.10 
## 128    253   128     12.2      Dasypus.sabanicola               7    2.69 
## 129    253   129      4.44     Dasypus.septemcinctus            7    2.82 
## 130    130   130     NA        Atlantogenata                    2    1.72 
## 131    130   131      0.000982 Afrotheria                       6    2.50 
## 132    131   132     44.5      Afrosorcida                      4    2.11 
## 133    132   133     11.8      <NA>                             4    2.11 
## 134    133   134      0.420    <NA>                             4    2.11 
## 135    134   135     10.1      <NA>                             3    1.91 
## 136    135   136      0.373    <NA>                             3    1.91 
## 137    136   137      2.96     <NA>                             3    1.91 
## 138    137   138      2.69     <NA>                             3    1.91 
## 139    138   139      0.450    <NA>                             2    1.72 
## 140    139   140      3.36     <NA>                             3    1.91 
## 141    137   141      0.153    <NA>                             3    1.91 
## 142    141   142      0.116    <NA>                             2    1.72 
## 143    141   143      0.222    <NA>                             2    1.72 
## 144    143   144      0.771    <NA>                             2    1.72 
## 145    144   145      0.109    <NA>                             3    1.91 
## 146    134   146     38.5      <NA>                             5    2.30 
## 147    146   147      1.81     <NA>                             3    1.91 
## 148    146   148      5.26     <NA>                             3    1.91 
## 149    133   149      4.28     <NA>                             3    1.91 
## 150    149   150     52.3      <NA>                             5    2.30 
## 151    149   151     15.6      <NA>                             3    1.91 
## 152    132   152     12.0      <NA>                             4    2.11 
## 153    152   153      0.0295   <NA>                             4    2.11 
## 154    153   154      0.173    <NA>                             3    1.91 
## 155    154   155      0.0141   <NA>                             3    1.91 
## 156    155   156      0.107    <NA>                             4    2.11 
## 157    156   157      0.00544  <NA>                             3    1.91 
## 158    157   158      0.102    <NA>                             3    1.91 
## 159    156   159      0.150    <NA>                             3    1.91 
## 160    159   160      0.483    <NA>                             3    1.91 
## 161    160   161      0.321    <NA>                             3    1.91 
## 162    161   162      0.00743  <NA>                             4    2.11 
## 163    155   163      2.12     <NA>                             3    1.91 
## 164    154   164      0.774    <NA>                             4    2.11 
## 165    164   165     18.6      <NA>                             5    2.30 
## 166    152   166      0.260    <NA>                             3    1.91 
## 167    166   167      0.155    <NA>                             3    1.91 
## 168    131   168      0.00016  <NA>                             6    2.50 
## 169    168   169      2.49     Macroscelidae                    5    2.30 
## 170    169   170      0.105    <NA>                             5    2.30 
## 171    170   171      4.23     <NA>                             5    2.30 
## 172    171   172      6.20     <NA>                             3    1.91 
## 173    172   173      0.977    <NA>                             3    1.91 
## 174    173   174      0.448    <NA>                             4    2.11 
## 175    174   175      0.349    <NA>                             4    2.11 
## 176    174   176      0.176    <NA>                             4    2.11 
## 177    173   177      1.62     <NA>                             4    2.11 
## 178    172   178      0.155    <NA>                             3    1.91 
## 179    178   179      0.180    <NA>                             4    2.11 
## 180    171   180      0.100    <NA>                             5    2.30 
## 181    180   181     11.2      <NA>                             3    1.91 
## 182    181   182     12.0      <NA>                             3    1.91 
## 183    180   183     30.6      <NA>                             3    1.91 
## 184    183   184      2.34     <NA>                             3    1.91 
## 185    184   185      0.0165   <NA>                             3    1.91 
## 186    169   186      4.53     <NA>                             5    2.30 
## 187    168   187    546.       Pseudoungulata                  10    3.27 
## 188    187   188     12.9      <NA>                            10    3.27 
## 189    187   189      4.42     Paenungulata                    10    3.27 
## 190    189   190      6.90     <NA>                             5    2.30 
## 191    190   191      1.47     <NA>                            10    3.27 
## 192    191   192     17.9      <NA>                             9    3.08 
## 193    192   193      0.00758  <NA>                             9    3.08 
## 194    193   194     19.5      <NA>                             5    2.30 
## 195    194   195      2.67     <NA>                             5    2.30 
## 196    191   196      5.61     <NA>                             3    1.91 
## 197    196   197      8.46     <NA>                             3    1.91 
## 198    197   198      0.602    <NA>                            11    3.47 
## 199    198   199      8.81     <NA>                            12    3.66 
## 200    199   200      0.347    <NA>                             3    1.91 
## 201    200   201      0.177    <NA>                            12    3.66 
## 202    200   202      0.584    <NA>                             3    1.91 
## 203    199   203      8.90     <NA>                            12    3.66 
## 204    197   204      9.93     <NA>                            11    3.47 
## 205    204   205      1.23     <NA>                            12    3.66 
## 206    205   206     22.2      <NA>                             3    1.91 
## 207    206   207     10.2      <NA>                            13    3.86 
## 208    196   208      6.63     <NA>                             3    1.91 
## 209    208   209      1.16     <NA>                            10    3.27 
## 210    190   210     22.7      <NA>                             9    3.08 
## 211    210   211      8.32     <NA>                             8    2.88 
## 212    189   212    187.       Tethytheria                     13    3.86 
## 213    212   213      3.40     <NA>                             3    1.91 
## 214    213   214      6.37     <NA>                            13    3.86 
## 215    214   215      1.15     <NA>                            14    4.05 
## 216    215   216     13.4      <NA>                             3    1.91 
## 217    216   217     30.3      Proboscidae                      3    1.91 
## 218    217   218      2.21     Elephantidae                    15    4.24 
## 219    218   219      0.0116   Elephantina                     16    4.44 
## 220    219   220      0.128    <NA>                            15    4.24 
## 221    220   221      0.369    <NA>                             3    1.91 
## 222    221   222    400.       <NA>                             3    1.91 
## 223    219   223      0.645    <NA>                             3    1.91 
## 224    223   224      0.467    Mammuthus                       16    4.44 
## 225    218   225      0.108    Loxodontini                     16    4.44 
## 226    225   226      0.862    Loxodona                        16    4.44 
## 227    226   227      6.09     <NA>                             3    1.91 
## 228    217   228      9.63     <NA>                             3    1.91 
## 229    215   229      0.197    <NA>                             3    1.91 
## 230    229   230      2.47     <NA>                             3    1.91 
## 231    213   231      0.128    <NA>                            13    3.86 
## 232    231   232      0.268    <NA>                             3    1.91 
## 233    212   233      0.309    <NA>                            12    3.66 
## 234    233   234     16.4      <NA>                             3    1.91 
## 235    234   235      0.530    <NA>                             3    1.91 
## 236    234   236      0.134    <NA>                             3    1.91 
## 237    236   237      0.145    <NA>                             3    1.91 
## 238    233   238      3.59     <NA>                             3    1.91 
## 239    130   239    125.       Xenarthra                       11    3.47 
## 240    239   240     14.6      <NA>                             3    1.91 
## 241    240   241      3.04     <NA>                             3    1.91 
## 242    240   242      0.0660   <NA>                            11    3.47 
## 243    242   243      0.112    <NA>                            11    3.47 
## 244    239   244     15.1      <NA>                             3    1.91 
## 245    244   245     21.7      <NA>                             9    3.08 
## 246    239   246      0.958    <NA>                            10    3.27 
## 247    246   247      0.0602   <NA>                             3    1.91 
## 248    247   248      0.0914   <NA>                             3    1.91 
## 249    248   249      0.000579 <NA>                             8    2.88 
## 250    248   250      0.000908 <NA>                            11    3.47 
## 251    247   251      0.217    <NA>                             3    1.91 
## 252    251   252      0.0523   <NA>                            11    3.47 
## 253    246   253      0.140    <NA>                             8    2.88
\end{verbatim}

\begin{verbatim}
## Warning: Fig5
\end{verbatim}

The dramatic increase in body mass and lifespan in some Afrotherian
lineages implies those lineages evolved reduced cancer risk. To infer
the magnitude of these reductions we estimated differences in cancer
risk between small bodied, short-lived species and large bodied,
long-lived species as well as for reconstructed ancestral Afrotherians.
Following {[}{\textbf{???}}{]} we estimate the intrinsic cancer risk as
the product of risk associated with body mass and lifespan. Differences
in cancer susceptibility due to body mass differences between species
can be approximated simply as the fold difference in body mass (\(D\))
between species {[}{\textbf{???}}{]}. The risk of developing cancer also
increases in proportion to the sixth power of age and is approximated by
the formula \(Ct^6\), in which the proportionality constant C that
determines susceptibility to cancer induction is multiplied by the sixth
power of the age in years, \(t\)
{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. Thus we can estimate
differences in the intrinsic cancer risk between species as \(D(Ct^6)\).
We inferred that susceptibility to cancer induction (\(C\)) was reduced
3.40x10\^{}3-fold in the ancestor of Pseudoungulata, 7.20×10\^{}3-fold
in the ancestor of Paenungulata, 8.81x10\^{}5-fold in the ancestor of
Tehthytheria, 1.46x10\^{}8-fold in the ancestor of Proboscidea, and
1.28x10\^{}6-fold in the ancestor of Sirenia relative to the ancestral
Afrotherian (Table 3). What biological mechanisms underlie the evolution
of thousand- to hundred million-fold reductions in cancer susceptibility
during the origins of Afrotherians, which are essential for large body
size and long lifespan to evolve

\hypertarget{other-figures}{%
\section{OTHER FIGURES}\label{other-figures}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Green2011}{}%
1. Green J, Cairns BJ, Casabonne D, Wright FL, Reeves G, Beral V, et al.
Height and cancer incidence in the Million Women Study: prospective
cohort, and meta-analysis of prospective studies of height and total
cancer risk. The Lancet Oncology. 2011;12: 785--794.
doi:\href{https://doi.org/10.1016/s1470-2045(11)70154-1}{10.1016/s1470-2045(11)70154-1}

\leavevmode\hypertarget{ref-Nunney:20181c2}{}%
2. Nunney L. Size matters: height, cell number and a person's risk of
cancer. Proc R Soc B. 2018;285: 20181743.
doi:\href{https://doi.org/10.1098/rspb.2018.1743}{10.1098/rspb.2018.1743}

\leavevmode\hypertarget{ref-Dobson2013}{}%
3. Dobson JM. Breed-predispositions to cancer in pedigree dogs. ISRN
veterinary science. 2013;2013: 941275.
doi:\href{https://doi.org/10.1155/2013/941275}{10.1155/2013/941275}

\leavevmode\hypertarget{ref-CaulinAndMaley2011}{}%
4. Caulin AF, Maley CC. Peto's Paradox: evolution's prescription for
cancer prevention. Trends in ecology \& evolution. 2011;26: 175--82.
doi:\href{https://doi.org/10.1016/j.tree.2011.01.002}{10.1016/j.tree.2011.01.002}

\leavevmode\hypertarget{ref-Leroi2003}{}%
5. Leroi AM, Koufopanou V, Burt A. Cancer selection. Nature Reviews
Cancer. 2003;3: 226--231.
doi:\href{https://doi.org/10.1038/nrc1016}{10.1038/nrc1016}

\leavevmode\hypertarget{ref-Peto1975}{}%
6. Peto R, Roe F, Lee P, Levy L, Clack J. Cancer and ageing in mice and
men. British Journal of Cancer. 1975;32: 411--426.
doi:\href{https://doi.org/10.1038/bjc.1975.242}{10.1038/bjc.1975.242}

\leavevmode\hypertarget{ref-Ashur-Fabian2004}{}%
7. Ashur-Fabian O, Avivi A, Trakhtenbrot L, Adamsky K, Cohen M, Kajakaro
G, et al. Evolution of p53 in hypoxia-stressed Spalax mimics human tumor
mutation. Proceedings of the National Academy of Sciences. 2004;101:
12236--12241.
doi:\href{https://doi.org/10.1073/pnas.0404998101}{10.1073/pnas.0404998101}

\leavevmode\hypertarget{ref-Seluanov2008}{}%
8. Seluanov A, Hine C, Bozzella M, Hall A, Sasahara THC, Ribeiro AACM,
et al. Distinct tumor suppressor mechanisms evolve in rodent species
that differ in size and lifespan. Aging cell. 2008;7: 813--23.
doi:\href{https://doi.org/10.1111/j.1474-9726.2008.00431.x}{10.1111/j.1474-9726.2008.00431.x}

\leavevmode\hypertarget{ref-Gorbunova2012}{}%
9. Gorbunova V, Hine C, Tian X, Ablaeva J, Gudkov AV, Nevo E, et al.
Cancer resistance in the blind mole rat is mediated by concerted
necrotic cell death mechanism. Proceedings of the National Academy of
Sciences of the United States of America. 2012;109: 19392--6.
doi:\href{https://doi.org/10.1073/pnas.1217211109}{10.1073/pnas.1217211109}

\leavevmode\hypertarget{ref-Tian2013}{}%
10. Tian X, Azpurua J, Hine C, Vaidya A, Myakishev-Rempel M, Ablaeva J,
et al. High molecular weight hyaluronan mediates the cancer resistance
of the naked mole-rat. 2013;499.
doi:\href{https://doi.org/10.1038/nature12234}{10.1038/nature12234}

\leavevmode\hypertarget{ref-Sulak2016}{}%
11. Sulak M, Fong L, Mika K, Chigurupati S, Yon L, Mongan NP, et al.
TP53 copy number expansion is associated with the evolution of increased
body size and an enhanced DNA damage response in elephants. eLife.
2016;5: e11994.
doi:\href{https://doi.org/10.7554/elife.11994}{10.7554/elife.11994}

\leavevmode\hypertarget{ref-HAGR}{}%
12. Tacutu R, Craig T, Budovsky A, Wuttke D, Lehmann G, Taranukha D, et
al. Human Ageing Genomic Resources: Integrated databases and tools for
the biology and genetics of ageing. Nucleic Acids Research. 2013;41:
D1027--D1033.
doi:\href{https://doi.org/10.1093/nar/gks1155}{10.1093/nar/gks1155}

\leavevmode\hypertarget{ref-Schwartz1995}{}%
13. Schwartz GT, Rasmussen DT, Smith RJ. Body-Size Diversity and
Community Structure of Fossil Hyracoids. Journal of Mammalogy. 1995;76:
1088--1099. doi:\href{https://doi.org/10.2307/1382601}{10.2307/1382601}

\leavevmode\hypertarget{ref-Scheffer1972}{}%
14. Scheffer VB. The Weight of the Steller Sea Cow. Journal of
Mammalogy. 1972;53: 912--914.
doi:\href{https://doi.org/10.2307/1379236}{10.2307/1379236}

\leavevmode\hypertarget{ref-Larramendi:20151c2}{}%
15. Larramendi A. Shoulder Height, Body Mass, and Shape of
Proboscideans. Acta Palaeontologica Polonica. 2015;61.
doi:\href{https://doi.org/10.4202/app.00136.2014}{10.4202/app.00136.2014}

\leavevmode\hypertarget{ref-OLeary2013a}{}%
16. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
NP, et al. The placental mammal ancestor and the post-K-Pg radiation of
placentals. Science (New York, NY). 2013;339: 662--7.
doi:\href{https://doi.org/10.1126/science.1229237}{10.1126/science.1229237}

\leavevmode\hypertarget{ref-Springer2013}{}%
17. Springer MS, Meredith RW, Teeling EC, Murphy WJ. Technical comment
on "The placental mammal ancestor and the post-K-Pg radiation of
placentals". Science (New York, NY). 2013;341: 613.
doi:\href{https://doi.org/10.1126/science.1238025}{10.1126/science.1238025}

\leavevmode\hypertarget{ref-OLeary2013b}{}%
18. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
NP, et al. Response to comment on "The placental mammal ancestor and the
post-K-Pg radiation of placentals". Science (New York, NY). 2013;341:
613.
doi:\href{https://doi.org/10.1126/science.1238162}{10.1126/science.1238162}

\leavevmode\hypertarget{ref-PuttickAndThomas2015}{}%
19. Puttick MN, Thomas GH. Fossils and living taxa agree on patterns of
body mass evolution: a case study with Afrotheria. Proceedings
Biological sciences / The Royal Society. 2015;282: 20152023.
doi:\href{https://doi.org/10.1098/rspb.2015.2023}{10.1098/rspb.2015.2023}

\leavevmode\hypertarget{ref-Abegglen:JAMA2015}{}%
20. Abegglen LM, Caulin AF, Chan A, Lee K, Robinson R, Campbell MS, et
al. Potential Mechanisms for Cancer Resistance in Elephants and
Comparative Cellular Response to DNA Damage in Humans. JAMA. 2015;314:
1850--1860.
doi:\href{https://doi.org/10.1001/jama.2015.13134}{10.1001/jama.2015.13134}

\leavevmode\hypertarget{ref-Vazquez2018}{}%
21. Vazquez JM, Sulak M, Chigurupati S, Lynch VJ. A Zombie LIF Gene in
Elephants Is Upregulated by TP53 to Induce Apoptosis in Response to DNA
Damage. Cell Reports. 2018;24: 1765--1776.
doi:\href{https://doi.org/10.1016/j.celrep.2018.07.042}{10.1016/j.celrep.2018.07.042}

\leavevmode\hypertarget{ref-Caulin2015}{}%
22. Caulin AF, Graham TA, Wang L-S, Maley CC. Solutions to Peto's
paradox revealed by mathematical modelling and cross-species cancer gene
analysis. Philosophical transactions of the Royal Society of London
Series B, Biological sciences. 2015;370: 20140222.
doi:\href{https://doi.org/10.1098/rstb.2014.0222}{10.1098/rstb.2014.0222}

\leavevmode\hypertarget{ref-Doherty2016}{}%
23. Doherty A, Magalhães J de. Has gene duplication impacted the
evolution of Eutherian longevity? Aging Cell. 2016;15: 978--980.
doi:\href{https://doi.org/10.1111/acel.12503}{10.1111/acel.12503}

\leavevmode\hypertarget{ref-Bininda-Emonds2008}{}%
24. Bininda-Emonds ORP, Cardillo M, Jones KE, MacPhee RDE, Beck RMD,
Grenyer R, et al. Erratum: The delayed rise of present-day mammals.
Nature. 2008;456: 274--274.
doi:\href{https://doi.org/10.1038/nature07347}{10.1038/nature07347}

\leavevmode\hypertarget{ref-ElliotAndMooers2014}{}%
25. Elliot MG, Mooers AØ. Inferring ancestral states without assuming
neutrality or gradualism using a stable model of continuous character
evolution. BMC evolutionary biology. 2014;14: 226.
doi:\href{https://doi.org/10.1186/s12862-014-0226-8}{10.1186/s12862-014-0226-8}

\leavevmode\hypertarget{ref-blat}{}%
26. Kent JW. BLAT---The BLAST-Like Alignment Tool. Genome Research.
2002;12: 656--664.
doi:\href{https://doi.org/10.1101/gr.229202}{10.1101/gr.229202}

\leavevmode\hypertarget{ref-AltenhoffAndDessimoz2009}{}%
27. Altenhoff AM, Dessimoz C. Phylogenetic and functional assessment of
orthologs inference projects and methods. PLoS computational biology.
2009;5: e1000262.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000262}{10.1371/journal.pcbi.1000262}

\leavevmode\hypertarget{ref-uniprot}{}%
28. Consortium TU. UniProt: the universal protein knowledgebase. Nucleic
Acids Research. 2017;45: D158--D169.
doi:\href{https://doi.org/10.1093/nar/gkw1099}{10.1093/nar/gkw1099}

\nolinenumbers

\end{document}

### 2020-05-14 paper_PLOS/paper_PLOS_draft.Rmd
additional figures, cleaned major errors, added completion warnings

@@
 output: 
-  workflowr::wflow_html: default
   rticles::plos_article: 
@@
       in_header: preamble.tex
+  workflowr::wflow_html: default
   word_document: default
@@
 ```{r Libraries and functions, include = F}
-knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = T, # fig.width=5, 
+knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=5, 
                       fig.retina = T)
+
+if(basename(getwd()) != "paper_PLOS") setwd("paper_PLOS")
+
 set.seed(1234)
@@
 
+stabletraits.progress <- read_tsv("../data/stableTraits/eutheria.progress")
+
+time.tree <- read.tree("../data/stableTraits/eutheria.tree")
+time.tree.atlantogenata <- tree_subset(time.tree, MRCA(time.tree, "Elephas.maximus", "Dasypus.novemcinctus"), levels_back = 0)
+
+#################################################################################
+
 bsize.tree <- read.beast("../output/AncSizeRec/AnnotatedTreeForFigure.tree") %>% 
@@
 
+#################################################################################
+
 ml.treedata <- read.beast("../output/geneLists/maxLikelihood_model_MK+FQ+I+G4-dataType_MORPH-asrMin_0.8/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata_RBB_filtered_dyn/atlantogenata_RBB_filtered_dyn.nexus")
 
+#################################################################################
+
 geneCopyTable <- read_csv("../output/geneCopyTable/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata-GeneCopyTable_RBB_filtered_long.csv")
+
+#################################################################################
+
+ORA.all.f <- "../output/extra/ORA_all.csv"
+if (file.exists(ORA.all.f)){
+  ORA.all <- read_csv(ORA.all.f)
+} else{
   ORA.all <- lineage.pathway.fdr.table("../output/ORA") %>% 
@@
     )
+  dir.create("../output/extra", showWarnings = F)
+  ORA.all %>% write_csv(path=ORA.all.f)
+}
+
+#################################################################################
+
 loxAf4.gene_transcript <- fread(
@@
 
+#################################################################################
+
+loxAfr4.t.f <- "../output/extra/loxafr4_transcripts.csv"
+if (file.exists(loxAfr4.t.f)){
+  loxAfr4.transcripts <- fread(loxAfr4.t.f)
+} else{
   loxAfr4.transcripts <- getTPM.dt("loxAfr4")
+  dir.create(dirname(loxAfr4.t.f), showWarnings = F)
+  loxAfr4.transcripts %>% fwrite(file = loxAfr4.t.f)
+}
+
+triManLat2.t.f <- "../output/extra/triManLat2_transcripts.csv"
+if (file.exists(triManLat2.t.f)){
+  triManLat2.transcripts <- fread(triManLat2.t.f)
+} else{
   triManLat2.transcripts <- getTPM.dt("triManLat2")
+  dir.create(dirname(triManLat2.t.f), showWarnings = F)
+  triManLat2.transcripts %>% fwrite(file = triManLat2.t.f)
+}
 
@@
 CEGMA.qc %<>% pivot_wider(id_cols=Genome, names_from = Stat, values_from=Value)
-CEGMA.completeness <- read_csv("../output/CEGMA/CEGMA_Completeness.csv", comment = "#", col_names = c("Genome","Group","#Prots","%Completeness","#Total","Average","%Ortho","Missing Proteins")) %>% 
-  filter(!is.na(Average))
+CEGMA.completeness <- read_csv("../output/CEGMA/CEGMA_Completeness_clean.csv")
 
@@
 size.scale.max <- size.scale[length(size.scale)]
-
+get.size.color <- function(n, scale = size.scale){
+  scale.colors = viridis::viridis(n=length(scale)) %>% 
+    set_names(., as.character(scale))
+  scale.colors[as.character(n)]
+}
 bsize.Atlantogenata.nodes <- list(
@@
 
-get.size.color <- function(n){
-  viridis::viridis(n=length(size.scale), )[size.scale[n]]
-}
-
+warning("TODO: de-comment geom_tiplab in p.tree.atlantogenata! Commented currently for offline usage!")
 p.tree.atlantogenata <-  
@@
   ) + 
-  geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic", 
-              show_guide=F, align = F, linesize = 0, offset= 100
-              ) +
+  # geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic", 
+  #             show_guide=F, align = F, linesize = 0, offset= 100
+  #             ) +
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidae, label="Proboscidae", 
@@
 
-not.me <- function(attr1, attr2, threshold=5){
-  sapply(
-    seq_along(attr1),
-    function(n, a1=attr1, a2=attr2, t=threshold){
-      ifelse(a1[n] >= t, a2[n], NA)
-    }
-  )
-}
-yes.me <- function(attr1, attr2, threshold=5){
-  sapply(
-    seq_along(attr1),
-    function(n, a1=attr1, a2=attr2, t=threshold){
-      ifelse(a1[n] < t, a2[n], NA)
-    }
-  )
-}
-# ml.tree.increasing.geneCounts.withSize %>% as.tibble %>% print(n=1351)
+# ml.tree.increasing.geneCounts.withSize %>% as_tibble %>% print(n=1351)
 ```
@@
 To compensate for this, we came up with a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we see each nucleotides of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (Supplementary Figure 1).
-This allows us to correct for genes that have been fragmented across incomplete genomes, without underestimating the copy number of proteins that lack domains found in the human sequence.
+This allows us to correct for genes that have been fragmented across incomplete genomes, while also taking into account missing sequences from the human query in the target genome.
+Mathematically, this can be written as:
+$$ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}$$
+where $n$ is a given nucleotide in the query, $l$ is the total length of the query, $C_n$ is the number of instances that $n$ is present within a reciprocal best hit, and $bool(C_n)$ is 1 if $C_n > 0$ or 0 if $C_n = 0$.
 
@@
 
+__Orthogonal Genome Assessment using CEGMA__ In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome.
+CEGMA was run using the default settings of [`r warning("TODO: add CEGMA settings")`], and the mammalian-specific core gene sets.
+
 ## Evidence for Functionality of Identified Genes
@@
 
-To determine which pathways were associated with duplicated genes in each species and lineage, we used WEBGESTALT to perform overrepresentation analysis (ORA) of the duplicated gene lists relative to our initial query gene list.
-For the database of pathways used in the analysis, we used Reactome [@Reactome], Wikipathways and Wikipathways_cancer [@Wikipathways], and KEGG [@KEGG].
+To determine which pathways were associated with duplicated genes in each species and lineage, we used WEBGESTALT to perform overrepresentation analysis (ORA) of the duplicated gene lists relative to our initial query gene list [@webgestalt2019].
+For the database of pathways used in the analysis, we used Reactome [@Reactome], Wikipathways, and Wikipathways_cancer [@Wikipathways], and KEGG [@KEGG].
+For the ORA, we used FDR for determining significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and FDR=0.5.
+
+## Lifespan Phylogenetic Least-Square Regression and Calculating Estimated Cancer Risk Thoughout Atlantogenata
+
+In order to determine the cancer risk at each node, we used a simplified multistage cancer risk model for body size and lifespan [@ArmitageAndDol; @Peto2015; @Peto2015b]. We defined the
 
@@
 ```
-
 ```{r LoxAfr All TSG Gene Transcripts}
@@
 ```
-
 ```{r LoxAfr Increased Gene Transcripts}
@@
 ```{r LoxAfr Other Gene Transcripts}
-genesToHighlight.plus <- c(
-   names(ml.treedata.loxAfrHighlight.TSG.any),
+genesToHighlight.plus <- list(
+  "CellCycle" = c(
     "RBX1", #Cell cycle, ubiqu., Probo
@@
     "MAX", # Cell Cycle, Tethytheria
-  "SYCP3", #Cell cycle, unique to LoxAfr
-
+    "SYCP3" #Cell cycle, unique to LoxAfr
+  ),
+  "APC" = c(
     "CDK1", # APC and Cell Cycle, Atlantogenata
@@
     "PSMB3", # APC/C, ubiq, Probo
-  "BUB3", # APC, 2 copies until 3 in LoxAfr
-
+    "BUB3" # APC, 2 copies until 3 in LoxAfr
+  ),
+  "TP53" = c(
     "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata
@@
     "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants
-  "COX20", # TP53 Regulates Metabolic Genes, LoxAfr
-
+    "COX20" # TP53 Regulates Metabolic Genes, LoxAfr
+  ),
+  "Stress" = c(
     "SOD1",  # Oxidative Stress, Probodiscea
@@
     "RNF2", # OSISenecense, Tethytheria
-
     "ORC3", # Replication Stress, Loxondontini, Elephas
-
+    "POLR2K", # TC-NER, Elephantidae
+    "HSBP1", # Heat Shock, Elephantidae
+    "ERCC8", #Ubiquitin mediated proteolysis, Elephantidea
+    "DTX3L", # Ubiquitination & Proteasome degradation, Loxodontini
+    "DNAJC2", # Heat Shock, Probo
+    "DET1" # Ubiq-mediate proteolysis, Probo
+  ),
+  "Death" = c(
     "RACK1", # Death Receptor Signaling, Tethy or Paen
@@
     "NET1", # Apoptosis in cancers, Elephantidae
-  "CSF2RB", #Apoptosis, Elephantidae
-
+    "CSF2RB" #Apoptosis, Elephantidae
+  ),
+  "RAS" = c(
     "RASA1", # RAS signaling Pathway, Paen
@@
     "NRAS", #Literally just RAS, Tethytheria
-  "GSK3A", #RAS, Tethy
-
-  "POLR2K", # TC-NER, Elephantidae
-
-  "HSBP1", # Heat Shock, Elephantidae
-  "ERCC8", #Ubiquitin mediated proteolysis, Elephantidea
-  "DTX3L", # Ubiquitination & Proteasome degradation, Loxodontini
-  "CTTN", # Proteoglycans in Cancer, Loxodontini
-
-  "DNAJC2", # Heat Shock, Probo
-  "DET1", # Ubiq-mediate proteolysis, Probo
-
+    "GSK3A" #RAS, Tethy
+  ),
+  "CancerRelatedGenes" = c(
     "MAPK3", # Many cancers, Afrotheria
@@
     "HNRNPK", #LncRNA involvement in Wnt and Cancer, Probodiscea
+    "CTTN", # Proteoglycans in Cancer, Loxodontini
     "CACYBP", # Gastric Cancer Network 2, Probodiscea
@@
     "PLPP2", # Choline Metabolism in Cancer, Loxodontini, Elephas, Trichechus
-  "PLAC8", # Gastric Cancer Network 2, continuously increases towards Loxodona
-
-  "LIMK1", #G13 Signaling, Probo
-
+    "PLAC8" # Gastric Cancer Network 2, continuously increases towards Loxodona
+  ),
+  "EEF.EIF" = c(
     "EEF1A1", # Vinnie wanted to highlight this
@@
   )
+)
 
 ml.treedata.loxAfrHighlight.plus <- genesToHighlight.plus %>% 
+  unlist %>% 
   set_names(.,.) %>% 
@@
 ```
-
 ```{r LoxAfr ORA Gene Transcripts}
@@
 ```
-
-```{r expressed loxAfr duplicates}
+```{r Expressed LoxAfr Duplicates}
 loxAfr4.expressedDups <- loxAfr4.geneTPM %>% 
@@
 ```
-
 ```{r Plot: Expressed LoxAfr Duplicates ORA}
@@
 
-```{r ORA.dotplots}
-p.loxAfr.ORA.dotplot <- ORA.all %>% 
-    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome") %>% 
-    graph.ORA.dotplot(., plotly=F)
-
-p.loxAfr.ORA.dotplot.increased <- ORA.all %>% 
-    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome", Direction=="increased") %>% 
-    graph.ORA.dotplot(., plotly=T)
-```
-
-## Step-wise evolution of large, long-lived Afrotherians
+## Body size frequently and independently expands and contracts throughout _Atlantogenata_
 
@@
 f1b <- p.tree.atlantogenata
+
 f1 <- ggarrange(
@@
 ```
-```{r Table 1}
+```{r Table 1, fig.cap="Table 1: Body Size and Confidence Intervals in _Atlantogenata_ estimated using StableTraits."}
 t.bodysize.atlantogenata.withNodes
@@
 ```
-```{r Correlations between genome quality scores}
-# Get upper triangle of the correlation matrix
-get_upper_tri <- function(cormat){
-  cormat[lower.tri(cormat)]<- NA
-  return(cormat)
-}
-reorder_cormat <- function(cormat){
-# Use correlation between variables as distance
-dd <- as.dist((1-cormat)/2)
-hc <- hclust(dd)
-cormat <-cormat[hc$order, hc$order]
-}
-
-CEGMA <- full_join(CEGMA.qc, CEGMA.completeness, by="Genome")
-CEGMA.completeonly <- CEGMA %>% 
-  filter(Group == "Complete_all") %>% 
+```{r Correlations between genome quality scores (1)}
+CEGMA.completeness.clean <-
+  CEGMA.completeness %>%
+  filter(str_detect(Group, "_all")) %>%
+  select(-`Missing Proteins`) %>%
   rename(
-    Completeness = `%Completeness`,
-    totalLength = `Total length (nt)`,
-    meanSequenceLength = `Mean sequence length (nt)`,
-    L50 = `L50 sequence count`,
-    N50 = `N50 sequence length (nt)`,
-    Total = `#Total`, 
+    n.Prots = `#Prots`,
+    pc.Completeness = `%Completeness`,
+    n.Total = `#Total`,
     pc.Ortho = `%Ortho`
-  )
-
-GCT.CEGMA.all <- geneCopyTable %>% 
+  ) %>%
+  reshape2::melt(., id = c("Genome", "Group")) %>%
+  mutate(Group = Group %>% str_remove("_all")) %>%
+  unite("variable", variable, Group) %>%
+  pivot_wider(id_cols = Genome, names_from = variable, values_from = value)
+
+CEGMA <- full_join(CEGMA.qc, CEGMA.completeness.clean, by="Genome")
+
+geneCopyTable.summary <-
+  geneCopyTable %>%
   mutate(
@@
     CN = if_else(ECNC<Copy, ECNC, Copy)
-  )%>% 
+  ) %>%
+  group_by(Genome) %>%
+  summarise(
+    # median.ECNC = median(ECNC),
+    mean.ECNC = mean(ECNC),
+    sd.ECNC = sd(ECNC),
+    # median.Copy = median(Copy),
+    mean.Copy = mean(Copy),
+    sd.Copy = sd(Copy),
+    # median.CN = median(CN),
+    mean.CN = mean(CN),
+    sd.CN = sd(CN),
+  )
+
+GCT.CEGMA <- geneCopyTable.summary %>%
   inner_join(., CEGMA, by="Genome") %>%
-  mutate_if(is.character, as.factor)%>% 
-  select(-Group)
-
-GCT.CEGMA.completeOnly <- geneCopyTable %>% 
-  mutate(
-    ECNC = ECNC %>% sapply(
-        .,
-        . %>% if_else(is.na(.), 0, .) %>% 
-      sum(., ifelse(.>round(.), 0.01, 0)) %>% 
-      round()
-    ),
-    Copy = Copy %>% sapply(. %>% if_else(is.na(.), 0, .)),
-    CN = if_else(ECNC<Copy, ECNC, Copy)
-  )%>% 
-  inner_join(., CEGMA.completeonly, by="Genome") %>% 
-  mutate_if(is.character, as.factor) %>% 
-  select(-Group)
-
-cor.GCT.CEGMA.completeOnly <- GCT.CEGMA.completeOnly %>% 
+  mutate_if(is.character, as.factor)
+
+cor.GCT.CEGMA <- GCT.CEGMA %>%
   mutate_if(is.factor, as.numeric) %>%
@@
   reorder_cormat %>%
-  get_upper_tri %>% 
-  melt(., na.rm=T) 
-  # as.data.frame %>%
-  # rownames_to_column(var = 'var1') %>%
-  # gather(var2, value, -var1) %>% 
-  # mutate(var_order = paste(var1, var2) %>%
-  #          strsplit(split = ' ') %>%
-  #          map_chr(
-  #            ~ sort(.x) %>%
-  #              paste(collapse = ' ')
-  #          )
-  # ) %>%
-  # mutate(cnt = 1) %>%
-  # group_by(var_order) %>%
-  # mutate(cumsum = cumsum(cnt)) %>%
-  # filter(cumsum != 2) %>%
-  # ungroup %>%
-  # select(-var_order, -cnt, -cumsum)
-
-cor.GCT.CEGMA.completeOnly %>% filter(Var2 %in% c("ECNC", "Copy", "CN"), !Var1 %in% c("ECNC", "Copy", "CN"))
-p.cor.GCT.CEGMA.completeOnly <- cor.GCT.CEGMA.completeOnly %>% 
+  # get_upper_tri %>%
+  get_lower_tri %>%
+  reshape2::melt(., na.rm=T)
+
+p.cor.GCT.CEGMA <- cor.GCT.CEGMA %>%
   # filter(value>0) %>%
   ggplot(
-    aes(x=Var2, y=Var1, fill = value)
+    # aes(x=Var2, y=Var1, fill = value)
+    aes(x=Var1, y=Var2, fill = value)
   )+
   geom_tile(color = "white")+
- scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
-   midpoint = 0, limit = c(-1,1), space = "Lab", 
-    name="Pearson\nCorrelation") +
-  theme_minimal()+ # minimal theme
- theme(axis.text.x = element_text(angle = 45, vjust = 1, 
-    size = 12, hjust = 1))+
+  # scale_fill_gradient2(low = "blue", high = "red", mid = "white",
+  #   midpoint = 0, limit = c(-1,1), space = "Lab",
+  #    name="Pearson\nCorrelation") +
+  scale_fill_viridis_c() +
+  theme_minimal()+
   coord_fixed()+
-# geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
+  # coord_flip() +
+  # scale_x_discrete(position = "top") +
+  scale_x_discrete(#limits = rev(levels(cor.GCT.CEGMA$Var1)), 
+    position = "top") +
+  scale_y_discrete(position = "right") +
+  # scale_y_discrete(limits = rev(levels(cor.GCT.CEGMA$Var1)), position = "right") +
   theme(
+    # axis.text.x = element_text(
+    #   angle = -45, vjust = 0,
+    #   size = 9, hjust = 0
+    #   ),
+    axis.text.x = element_text(
+      angle = -330, vjust = 0,
+      size = 9, hjust = 0
+    ),
     axis.title.x = element_blank(),
@@
     axis.ticks = element_blank(),
-  legend.justification = c(1, 0),
-  legend.position = c(0.6, 0.7),
+    legend.justification = c(1, 1),
+    legend.position = c(0.4,0.7), #c(0.6, 0.7),
     legend.direction = "horizontal")+
-  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
+  guides(fill = guide_colorbar(title="Pearson\nCorrelation",
+                               barwidth = 7, barheight = 1,
                                title.position = "top", title.hjust = 0.5))
-
-# Best independent vars:
-#   Genome, N50, Mean Sequence Length, Completeness, #Total, %Ortho, Average
-# Note: Genome highly correlated with "sum length seq", which is HC w/ "N50"
-# Nothing correlates with Gene
-# 
 ```
 ```{r Phylogenetic Correlations}
-CEGMA.completeonly.wGenome <- inner_join(
-  GCT.CEGMA.completeOnly, 
-  genomeTable.atlantogenata %>% select(label, BestGenome), 
-  by=c("Genome"="BestGenome")
-) %>% 
-  select(label, everything())
-missing.bsize.atl.CEGMA.completeonly <- bsize.Atlantogenata %>% 
-  as.tibble %>% 
-  filter(!label %in% CEGMA.completeonly.wGenome$label) %>% pull(label)
-bsize.atl <- bsize.Atlantogenata@phylo %>% 
-  drop.tip(missing.bsize.atl.CEGMA.completeonly)
-
-bm<-corBrownian(1, bsize.atl)
-pl<-corPagel(1, bsize.atl)
-
-# CEGMA.completeonly.wGenome %<>% filter(Gene == "TP53")
-# rown <- CEGMA.completeonly.wGenome$label %>% unique
-# CEGMA.completeonly.wGenome %<>% mutate_if(is.character, . %>% as.factor %>%  as.numeric)%>% 
+
+to_drop <-tree.Atlantogenata %>% as_tibble %>% filter(!BestGenome %in% (GCT.CEGMA %>% pull(Genome))) %>% pull(label)
+
+tree.GCT.CEGMA <- full_join(
+  tree.Atlantogenata %>% 
+    drop.tip(to_drop) %>% 
+    as_tibble %>% 
+    mutate(Genome = BestGenome),
+  GCT.CEGMA, 
+  by="Genome"
+) %>% 
+  as.treedata()
+GCT.CEGMA.wGenome <- tree.GCT.CEGMA %>% as_tibble
+
+# GCT.CEGMA.wGenome %<>% filter(Gene == "TP53")
+# rown <- GCT.CEGMA.wGenome$label %>% unique
+# GCT.CEGMA.wGenome %<>% mutate_if(is.character, . %>% as.factor %>%  as.numeric)%>%
 #   as.data.frame
-# rownames(CEGMA.completeonly.wGenome) <- rown
-```
-
-```{r Phylogenetic Correlations: model building}
-warning("TODO: Collapse RBHB Genes into Total, Duplicate, and Median CN")
-warning("TODO: CEGMA duplicate: multiply CEGMA %Ortho times #Total to get a Median CN")
-warning("TODO: Rename CEGMA things to CEGMA.[thing]")
-warning("TODO: I have a linear model for each gene regressing genome quality stats with gene copy numbers in a PGLS; figure out why this is useful and how to plot it.")
-
-runGLS.bm <- function(df, pcor = bm){
-  # print(class(df))
-  # print(head(df))
-  bla <- gls(
-  CN~Genome + N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average, 
-  data = df, 
-  correlation=pcor
-  ) 
-  # print(bla)
-  # print(class(bla))
-  return(bla)
-}
-tidyGLS <- function(df){
-  print(names(df))
-  print(class(df$results))
-  a <- tidy(df$results, conf.int=T)
-  print(a)
-  return(a)
-}
-
-gls.CN.bm <- CEGMA.completeonly.wGenome %>% 
-  # filter(Gene %in% c("TP53", "LIF")) %>%
-  nest(-Gene) %>% 
-  mutate(
-    data = map(
-      data,
-      . %>%
+# rownames(GCT.CEGMA.wGenome) <- rown
+```
+```{r Phylogenetic Correlations: model building (1), warning=T, eval=F}
+training.table.tree.GCT.CEGMA <- tree.GCT.CEGMA %>%
+  as_tibble() %>%
+  filter(!is.na(mean.CN)) %>%
   mutate_if(is.character, . %>% as.factor) %>%
-        mutate(label=as.character(label)) %>% 
+  mutate(label=as.character(label) %>% str_replace_all("\\.", " ")) %>%
   mutate_if(is.factor, as.numeric) %>%
@@
   column_to_rownames("label")
-    ),
-    model = map(
-      data, 
-      possibly(
-        ~runGLS.bm(.),
-        otherwise = NA_real_
-      ),
-    ),
-    tidied = map(
-      model,
-      possibly(
-        .%>% tidy(., conf.int=T),
-        otherwise = NA_real_
-      )
-    )
-  )
-
-gls.CN.bm.perGene.tidy <- gls.CN.bm %>% 
-  unnest(tidied) #%>%
-  # select(-model)
-
-# gls.atl.1<-CEGMA.completeonly.wGenome %>% 
-#   mutate_if(is.character, . %>% as.factor) %>% 
-#   mutate(label=as.character(label)) %>% 
-#   mutate_if(is.factor, as.numeric) %>% 
-#   group_by(Gene) %>% 
-#   do(model = . %>% 
-#       as.data.frame %>% 
-#       column_to_rownames("label") %>% 
-#        gls(
-#   CN~Genome + N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average, 
-#   data = ., 
-#   correlation=bm)
-#   ) %>% tidy(model)
-# gls.atl.2<-gls(
-#   CN~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average,
-#   data=CEGMA.completeonly.wGenome %>% filter(Gene == "TP53") %>%
-#       mutate_if(is.character, . %>% as.factor) %>% 
-#       mutate(label=as.character(label)) %>% 
-#       mutate_if(is.factor, as.numeric) %>% 
-#       as.data.frame %>% 
-#       column_to_rownames("label"), 
-#   correlation=pl)
-# gls.atl.3<-gls(
-#   Copy~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average,
-#   data=CEGMA.completeonly.wGenome %>% filter(Gene == "TP53") %>%
-#       mutate_if(is.character, . %>% as.factor) %>% 
-#       mutate(label=as.character(label)) %>% 
-#       mutate_if(is.factor, as.numeric) %>% 
-#       as.data.frame %>% 
-#       column_to_rownames("label"), 
-#   correlation=bm)
-# gls.atl.4<-gls(
-#   Copy~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average,
-#   data=CEGMA.completeonly.wGenome %>% filter(Gene == "TP53") %>%
-#       mutate_if(is.character, . %>% as.factor) %>% 
-#       mutate(label=as.character(label)) %>% 
-#       mutate_if(is.factor, as.numeric) %>% 
-#       as.data.frame %>% 
-#       column_to_rownames("label"), 
-#   correlation=pl)
-# gls.atl.4 %>% summary
-# AIC(gls.atl.3, gls.atl.1,gls.atl.2,gls.atl.4)
-```
-
-```{r LM Duplicates~Genome quality metrics}
-m1 <- lm(data = GCT.CEGMA.completeOnly, formula = Copy~Genome+ N50 + meanSequenceLength + Completeness + Total+ pc.Ortho + Average)
+
+names(training.table.tree.GCT.CEGMA) %<>% make.names(unique = T) %>% str_replace_all("\\.+", ".")
+
+phyloGCT.CEGMA <- tree.GCT.CEGMA@phylo %>% as.phylo()
+
+bm.gct <-corBrownian(1, phyloGCT.CEGMA)
+pl.gct <-corPagel(1, phyloGCT.CEGMA)
+
+models.vars <- training.table.tree.GCT.CEGMA %>% 
+  select(
+    # -mean.Copy, 
+    # -mean.ECNC, 
+    # -mean.CN,
+    # -sd.ECNC, 
+    # -sd.Copy, 
+    # -sd.CN,
+    -X.of.sequences.containing.non.ACGTN.nt., -Base.composition.Other., -parent, -node, 
+    -Brownian, -Median, -X95.CI_Low, -X95.CI_High, -branch.length, -Species, 
+    -Species.strict, -Spc.short, -BestGenome, -highlight, -Base.composition.C.,
+    -Base.composition.A., -Base.composition.T., -Base.composition.G., 
+    -phyloPicUID, -Rate,-Orig_len, -Common.Name, -Longest.sequence.nt., 
+    -Shortest.sequence.nt., -Mean.sequence.length.nt., -Median.sequence.length.nt., 
+    -starts_with("X.of.sequences"), -starts_with("Sum.length")
+  )
+
+models.mean.ECNC <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.ECNC ~ ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.CN <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.CN ~ ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.Copy <-names(models.vars) %>% 
+    str_remove("mean.*|sd.*") %>% 
+    stringi::stri_omit_empty() %>% 
+    paste0("mean.Copy ~ ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+aic.models.mean.ECNC <- AIC(models.mean.ECNC[[1]], models.mean.ECNC[[2]], models.mean.ECNC[[3]], models.mean.ECNC[[4]], models.mean.ECNC[[5]], models.mean.ECNC[[6]], models.mean.ECNC[[7]], models.mean.ECNC[[8]], models.mean.ECNC[[9]], models.mean.ECNC[[10]], models.mean.ECNC[[11]], models.mean.ECNC[[12]], models.mean.ECNC[[13]], models.mean.ECNC[[14]],
+    models.mean.ECNC[[15]], models.mean.ECNC[[16]], models.mean.ECNC[[17]], models.mean.ECNC[[18]]#,models.mean.ECNC[[19]], models.mean.ECNC[[20]], models.mean.ECNC[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.ECNC)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.Copy <- AIC(models.mean.Copy[[1]], models.mean.Copy[[2]], models.mean.Copy[[3]], models.mean.Copy[[4]], models.mean.Copy[[5]], models.mean.Copy[[6]], models.mean.Copy[[7]], models.mean.Copy[[8]], models.mean.Copy[[9]], models.mean.Copy[[10]], models.mean.Copy[[11]], models.mean.Copy[[12]], models.mean.Copy[[13]], models.mean.Copy[[14]],
+    models.mean.Copy[[15]], models.mean.Copy[[16]], models.mean.Copy[[17]], models.mean.Copy[[18]]#,models.mean.Copy[[19]], models.mean.Copy[[20]], models.mean.Copy[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.Copy)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.CN <- AIC(models.mean.CN[[1]], models.mean.CN[[2]], models.mean.CN[[3]], models.mean.CN[[4]], models.mean.CN[[5]], models.mean.CN[[6]], models.mean.CN[[7]], models.mean.CN[[8]], models.mean.CN[[9]], models.mean.CN[[10]], models.mean.CN[[11]], models.mean.CN[[12]], models.mean.CN[[13]], models.mean.CN[[14]],
+    models.mean.CN[[15]], models.mean.CN[[16]], models.mean.CN[[17]], models.mean.CN[[18]]#,models.mean.CN[[19]], models.mean.CN[[20]], models.mean.CN[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.CN)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+```
+```{r Phylogenetic Correlations: model building (1 + lnsize), warning=T, eval=F}
+models.mean.ECNC.lnSize <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*|lnSize") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.ECNC ~ lnSize + ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.CN.lnSize <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*|lnSize") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.CN ~ lnSize + ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.Copy.lnSize <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*|lnSize") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.Copy ~ lnSize + ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+aic.models.mean.ECNC.lnSize <- AIC(models.mean.ECNC.lnSize[[1]], models.mean.ECNC.lnSize[[2]], models.mean.ECNC.lnSize[[3]], models.mean.ECNC.lnSize[[4]], models.mean.ECNC.lnSize[[5]], models.mean.ECNC.lnSize[[6]], models.mean.ECNC.lnSize[[7]], models.mean.ECNC.lnSize[[8]], models.mean.ECNC.lnSize[[9]], models.mean.ECNC.lnSize[[10]], models.mean.ECNC.lnSize[[11]], models.mean.ECNC.lnSize[[12]], models.mean.ECNC.lnSize[[13]], models.mean.ECNC.lnSize[[14]],
+    models.mean.ECNC.lnSize[[15]], models.mean.ECNC.lnSize[[16]], models.mean.ECNC.lnSize[[17]]#, models.mean.ECNC.lnSize[[18]],models.mean.ECNC.lnSize[[19]], models.mean.ECNC.lnSize[[20]]#, models.mean.ECNC.lnSize[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.ECNC.lnSize)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.Copy.lnSize <- AIC(models.mean.Copy.lnSize[[1]], models.mean.Copy.lnSize[[2]], models.mean.Copy.lnSize[[3]], models.mean.Copy.lnSize[[4]], models.mean.Copy.lnSize[[5]], models.mean.Copy.lnSize[[6]], models.mean.Copy.lnSize[[7]], models.mean.Copy.lnSize[[8]], models.mean.Copy.lnSize[[9]], models.mean.Copy.lnSize[[10]], models.mean.Copy.lnSize[[11]], models.mean.Copy.lnSize[[12]], models.mean.Copy.lnSize[[13]], models.mean.Copy.lnSize[[14]],
+    models.mean.Copy.lnSize[[15]], models.mean.Copy.lnSize[[16]], models.mean.Copy.lnSize[[17]]#, models.mean.Copy.lnSize[[18]],models.mean.Copy.lnSize[[19]], models.mean.Copy.lnSize[[20]]#, models.mean.Copy.lnSize[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.Copy.lnSize)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.CN.lnSize <- AIC(models.mean.CN.lnSize[[1]], models.mean.CN.lnSize[[2]], models.mean.CN.lnSize[[3]], models.mean.CN.lnSize[[4]], models.mean.CN.lnSize[[5]], models.mean.CN.lnSize[[6]], models.mean.CN.lnSize[[7]], models.mean.CN.lnSize[[8]], models.mean.CN.lnSize[[9]], models.mean.CN.lnSize[[10]], models.mean.CN.lnSize[[11]], models.mean.CN.lnSize[[12]], models.mean.CN.lnSize[[13]], models.mean.CN.lnSize[[14]],
+    models.mean.CN.lnSize[[15]], models.mean.CN.lnSize[[16]], models.mean.CN.lnSize[[17]]#, models.mean.CN.lnSize[[18]],models.mean.CN.lnSize[[19]], models.mean.CN.lnSize[[20]]#, models.mean.CN.lnSize[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.CN.lnSize)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+
 ```
@@
 ```
-```{r Table 2: ML model summary}
-warning("Table 2")
+```{r PREP Table 2: PGLS of Copy Number ~ Genomic Stats, eval=F}
+# if(knitr::is_latex_output()){
+t.GCT.CEGMA.gls <- quiet(stargazer(
+  models.mean.Copy,models.mean.ECNC, models.mean.CN,
+  type="latex",
+  title = "Gene Copy ~ various genomic variables",
+  style = "all"
+)  %>% 
+  c("\\newgeometry{margin=1cm}\n\\begin{landscape}", ., "\\end{landscape}\n\\restoregeometry") %>% 
+  paste0(collapse="\n")
+  )
+# } else if(knitr::is_html_output()){
+#   t.GCT.CEGMA.gls <- quiet(stargazer(
+#     c(models.mean.CN %>% unname, models.mean.Copy%>% unname, models.mean.ECNC%>% unname),
+#     type="html",
+#     title = "Gene Copy ~ various genomic variables",
+#     style = "all"
+#   ))
+# } else{
+#   t.GCT.CEGMA.gls <- quiet(stargazer(
+#     c(models.mean.CN %>% unname, models.mean.Copy%>% unname, models.mean.ECNC%>% unname),
+#     type="text",
+#     title = "Gene Copy ~ various genomic variables",
+#     style = "all"
+#   ))
+# }
+```
+```{r Table 2: ML model summary, warning=T}
+warning("TODO: Table 2 is done, but Stargazer has some weird bug stopping it from running...")
 ```
@@
 
+<!--
 ## Retroduplication of EEF1A1 in the stem-lineage of Tehthytheria
 
-```{r Fig3, fig.cap="Retroduplication of Eukaryotic translation Elongation Factor 1 Alpha 1 (EEF1A1).
-A) Gene Copy Number of EEF1A1 in _Atlantogenata_.
-B) Expression of EEF1A1 copies in the West Indian Manatee and African Elephant"}
+```{r Fig3 prep}
 f3a <- p.ml.treedata.loxAfrHighlight.plus$EEF1A1
@@
 )
+```
+
+```{r Fig3, fig.cap="Retroduplication of Eukaryotic translation Elongation Factor 1 Alpha 1 (EEF1A1).
+A) Gene Copy Number of EEF1A1 in _Atlantogenata_.
+B) Expression of EEF1A1 copies in the West Indian Manatee and African Elephant", warning=T}
 f3
@@
 3C).
+-->
 
 ## Duplications that occured recently in Probodiscea are enriched for tumor suppressor pathways
+
+```{r ORA.dotplots}
+p.loxAfr.ORA.dotplot <- ORA.all %>% 
+    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome") %>% 
+    graph.ORA.dotplot(., plotly=knitr::is_html_output())
+
+p.loxAfr.ORA.dotplot.increased <- ORA.all %>% 
+    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome", Direction=="increased") %>% 
+    graph.ORA.dotplot(., plotly=knitr::is_html_output())
+```
+
+```{r Fig 3, fig.pos="hold", fig.cap="Fig 3"}
+f3a <- p.loxAfr.ORA.dotplot
+f3b <- p.loxAfr.ORA.dotplot.increased
+
+f3 <- ggarrange(f3a, f3b, nrow=1, ncol=2, common.legend = T, legend = "bottom", label="AUTO")
+# f3
+f3a
+f3b
+```
 
@@
 
-```{r Fig4 }
-warning("Fig4")
+```{r Fig4 prep}
+
+    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata
+    # "STK11", #REgulation of TP53, Paenungulata
+    # "COX5A", #TP53 transcriptional regulation, PRobodiscea
+    # "BRD7", # TP53 regulation, Tethytheria
+    # "TP53", #TP53 duplicates in Tethytheria
+    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria
+    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants
+    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri
+    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants
+    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr
+
+genesToHighlight.plus.TP53 <- genesToHighlight.plus$TP53
+
+loxAfr4.geneTPM.ORA.expressedDups.TP53 <- loxAfr4.geneTPM.ORA.expressedDups %>% 
+  filter(Gene %in% genesToHighlight.plus.TP53) %>% 
+  mutate(
+    Reactome.Pathway.Prime = Pathway %>% str_split(", ") %>% sapply(. %>% str_extract(., ".*[Tt]?[Pp]53.*")%>% stringi::stri_remove_empty_na() %>% paste(collapse =", ")),
+    id = str_remove(Locus, ".*_") %>% factor(., levels=sort(as.numeric(unique(.)))),
+    Gene = Gene %>% factor(., levels=genesToHighlight.plus$TP53)
+    ) %>% 
+  group_by(Locus) %>% 
+  summarize(Gene = unique(Gene), id=unique(id), TPM=max(TPM), Reactome.Pathway.Prime=unique(Reactome.Pathway.Prime)) %>% 
+  arrange(Gene, id)
+
+# TP53.genes.plots <- genesToHighlight.plus.TP53 %>% 
+#   set_names(.,.) %>% 
+#   lapply(
+#     .,
+#     function(gene){
+#       list(
+#         "GeneCopyTree"=ml.treedata.loxAfrHighlight.plus[[gene]],
+#         "GeneCopyExpression"= 
+#       )
+#     }
+#   )
+
+p.tp53.genes.expression <- loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% 
+  # filter(Gene == gene) %>% 
+  ggplot(
+    aes(
+      x=id, 
+      y=TPM, 
+      color=Reactome.Pathway.Prime
+    )
+  ) +
+  geom_point(
+    size=3
+  )+
+  geom_segment(
+    aes(
+      x=id,
+      xend=id,
+      y=0,
+      yend=TPM
+    )
+  )+
+  scale_y_sqrt()+
+  #scale_color_brewer(palette = "Dark2", name="Reactome Pathway")+
+  guides(color=guide_legend(nrow = 3))+
+  facet_grid(cols = vars(Gene), scales = "free_x") + 
+  theme_pubclean()+
+  labs(
+    x="Copy",
+    y="TPM (sqrt)"
+  ) + 
+  theme(legend.position="bottom")
+warning("TODO: 1. Color each facet a different color")
+warning("TODO: 2. Make a tree with stars at each node where one of these genes duplicated, color coded by genes")
+```
+
+```{r Fig4, warning=T}
+warning("TODO: Make Fig4")
+
 ```
@@
   data=dat.size.lifespan.Atlantogenata.withLife,
-  correlation = corPagel(1, tree.size.lifespan.Atlantogenata.withLife)
+  correlation = corBrownian(1, tree.size.lifespan.Atlantogenata.withLife)
 )
@@
 
-# tree.size.lifespan.Atlantogenata.full %>% as.tibble %>% print(n=12351235123)
+# For cancer resistance:
+to_drop <-tree.size.lifespan.Atlantogenata.full %>% as_tibble %>% filter(!label %in% genomeTable.atlantogenata$label) %>% pull(label)
+tree.size.lifespan.Atlantogenata.genomes <- tree.size.lifespan.Atlantogenata.full %>% 
+  drop.tip(to_drop) %>% 
+  as_tibble() %>% 
+  mutate(label = label %>% sapply(function(l){ifelse(is.na(l), "Afroinsectivora", l)})) %>% 
+  as.treedata()
+
 ```
 ```{r Estimated Cancer Risk Per Node}
-cancerSucceptibility.Atlantogenata <- tree.size.lifespan.Atlantogenata.full %>% 
+cancerSucceptibility.Atlantogenata.genomes <-  tree.size.lifespan.Atlantogenata.genomes %>% 
   as_tibble %>% 
@@
          lnK1 = 6*log(Lifespan) + lnSize)
-table.cancerSucceptibility.Atlantogenata <- cancerSucceptibility.Atlantogenata %>% 
+
+table.cancerSucceptibility.Atlantogenata.genomes <- tree.size.lifespan.Atlantogenata.genomes %>% 
+  as_tibble%>% 
+  mutate(
+    Ancestor = label %>% 
+    sapply(
+        .,
+        function(n, d=cancerSucceptibility.Atlantogenata.genomes){
+            if(is.na(n)){return(NA)}
+            #print(n)
+            #print(length(n))
+            p.node <- d %>% filter(label==n) %>% pull(parent) %>% unique
+            d %>% filter(node == p.node) %>% pull(label) %>% unique %>% return
+            }
+
+    ) %>% str_replace_all("\\.", " ") %>% str_replace("Root", "Atlantogenata"),
+    K1 = (Lifespan)^6 * exp(lnSize),
+    lnK1 = 6*log(Lifespan) + lnSize,
+    K2 = Ancestor %>% 
+    sapply(
+     .,
+     function(n, d=cancerSucceptibility.Atlantogenata.genomes){
+       d %>% filter(label==n) %>% pull(K1) %>% unique
+     }
+    ),
+  lnK2 = Ancestor %>% 
+    sapply(
+     .,
+     function(n, d=cancerSucceptibility.Atlantogenata.genomes){
+       d %>% filter(label==n) %>% pull(lnK1) %>% unique
+     }
+    ),
+  CancerSucceptabilityChange = K1/K2,
+  log2CancerSucceptabilityChange = log2(K1)-log2(K2)
+)
+tree.cancerSucceptibility.Atlantogenata.genomes <- table.cancerSucceptibility.Atlantogenata.genomes %>% 
+  full_join(., genomeTable, by = "label") %>% 
+  mutate(highlight = sapply(seq_along(label), function(x, l=species.highlight, Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)})) %>% 
+  as.treedata() %>% 
+  drop.tip(tree.Atlantogenata %>% as_tibble %>% filter(!BestGenome %in% (GCT.CEGMA %>% pull(Genome))) %>% pull(label))
+##########################################################################################
+cancerSucceptibility.Atlantogenata.full <- tree.size.lifespan.Atlantogenata.full %>% 
+  as_tibble %>% 
+  mutate(K1 = (Lifespan)^6 * exp(lnSize),
+         lnK1 = 6*log(Lifespan) + lnSize)
+table.cancerSucceptibility.Atlantogenata <- cancerSucceptibility.Atlantogenata.full %>% 
   # arrange(parent) %>% 
   mutate(
+    Ancestor = label %>% 
+    sapply(
+        .,
+        function(n, d=cancerSucceptibility.Atlantogenata.full){
+            if(is.na(n)){return(NA)}
+            #print(n)
+            #print(length(n))
+            p.node <- d %>% filter(label==n) %>% pull(parent) %>% unique
+            d %>% filter(node == p.node) %>% pull(label) %>% unique %>% return
+            }
+
+    ) %>% str_replace_all("\\.", " ") %>% str_replace("Root", "Atlantogenata"),
     K2 = parent %>% 
@@
      .,
-     function(n, d=cancerSucceptibility.Atlantogenata){
+     function(n, d=cancerSucceptibility.Atlantogenata.full){
        d %>% filter(node==n) %>% pull(K1) %>% unique
@@
     ),
-  CancerSucceptabilityChange = K2/K1,
-  log2CancerSucceptabilityChange = log2(K2/K1)
+  lnK2 = parent %>% 
+    sapply(
+     .,
+     function(n, d=cancerSucceptibility.Atlantogenata.full){
+       d %>% filter(node==n) %>% pull(lnK1) %>% unique
+     }
+    ),
+  CancerSucceptabilityChange = K1/K2,
+  log2CancerSucceptabilityChange = log2(K1)-log2(K2)
 )
 tree.cancerSucceptibility.Atlantogenata <- table.cancerSucceptibility.Atlantogenata %>% 
+  full_join(., genomeTable.atlantogenata, by = "label") %>% 
+  mutate(highlight = sapply(seq_along(label), function(x, l=str_replace(species.highlight, " ", ".") %>% set_names(.,.), Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)})) %>%
   as.treedata()
 ```
-
-```{r Fig 5, fig.cap=""}
-warning("TODO: Caption Fig 5")
-quiet <- function(x) { 
-  sink(tempfile()) 
-  on.exit(sink()) 
-  invisible(force(x)) 
-} 
+```{r Fig 5 prep: Scales}
 scale.log2CancerSucceptabilityChange <- table.cancerSucceptibility.Atlantogenata %>% 
@@
   set_names(.,.)
-
 scale.log2CancerSucceptabilityChange.min <- min(scale.log2CancerSucceptabilityChange)
@@
 
-get.log2CancerSucceptabilityChange.color <- function(n){
-  scale = viridis::viridis(n=length(scale.log2CancerSucceptabilityChange)) %>% 
-    set_names(., as.character(scale.log2CancerSucceptabilityChange))
-  scale[as.character(n)]
+get.log2CancerSucceptabilityChange.color <- function(n, scale = scale.log2CancerSucceptabilityChange){
+  scale.colors = viridis::viridis(n=length(scale)) %>% 
+    set_names(., as.character(scale))
+  scale.colors[as.character(n)]
 }
-
+cancerSucceptibility.Atlantogenata.nodes <- table.cancerSucceptibility.Atlantogenata.genomes$node %>% as.list() %>%
+set_names(., table.cancerSucceptibility.Atlantogenata.genomes$label)
+```
+
+```{r Fig 5 prep: Table 3}
+if(knitr::is_latex_output()){
+# size.lifespan.Atlantogenata.lm
   t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
     size.lifespan.Atlantogenata.lm,
-  title = "ln(Lifespan) & ln(Body Size) Regression",
+    type="latex",
+    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
     style = "all"
-))
-
-f5a <- tree.cancerSucceptibility.Atlantogenata %>% 
+  )) %>% paste0(collapse="\n")
+} else if(knitr::is_html_output()){
+  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
+    size.lifespan.Atlantogenata.lm,
+    type="html",
+    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
+    style = "all"
+  )) %>% paste0(collapse="\n")
+} else{
+  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
+    size.lifespan.Atlantogenata.lm,
+    type="text",
+    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
+    style = "all"
+  )) %>% paste0(collapse="\n")
+  t.size.lifespan.Atlantogenata.lm %>% cat
+}
+```
+
+```{r Fig 5 prep}
+f5a <- tree.cancerSucceptibility.Atlantogenata  %>% as_tibble %>% mutate(CancerSucceptabilityChange = as.numeric(CancerSucceptabilityChange)) %>% as.treedata %>% 
+  ggtree(
+    aes(color=log2CancerSucceptabilityChange),
+    continuous=T,
+    size = 1.5, 
+    yscale="log2CancerSucceptability"
+  ) + 
+  # geom_tiplab(aes(image=highlight, size=I(0.075)), geom="phylopic", color="black"
+  #             ) +
+  scale_color_viridis_c(limits=c(scale.log2CancerSucceptabilityChange.min, scale.log2CancerSucceptabilityChange.max), 
+                        guide = guide_colourbar(title="Log2 Cancer Succeptibility Change"))+
+  theme_minimal() +
+  theme(
+    legend.position="bottom",
+    panel.background = element_rect(fill = "darkgrey")
+  ) + 
+  xlab("Rate of Change, Body Size") +
+    ylab("Log2 Cancer Succeptibility Change")
+
+#############################################################################################
+#############################################################################################
+
+f5a2 <- tree.cancerSucceptibility.Atlantogenata  %>% 
   ggtree(
@@
   xlab("Rate of Change, Body Size")
-# f5b
+# f5a
 
@@
 )
-
-cat(t.size.lifespan.Atlantogenata.lm, sep = "\n")
+```
+
+```{r Fig 5 prep: Table 4}
+t.cancersucc.atl <- table.cancerSucceptibility.Atlantogenata.genomes %>% 
+  # filter(label %in% ml.treedata.table$label) %>%  
+  # select(-parent, -node, -branch.length) %>%
+  select(
+    Node = label, 
+    lnSize,
+    `Est. Lifespan`=Lifespan,
+    `Est. Cancer Susceptibility (K1)` = K1,
+    `Est. Ancestral K2` = K2,
+    `Change in K (K2/K1)` = CancerSucceptabilityChange,
+    `log2 Fold Change` = log2CancerSucceptabilityChange
+  ) %>%
+  mutate_at(
+    vars(`Est. Cancer Susceptibility (K1)`, `Est. Ancestral K2`), 
+    . %>% formatC(., format = "e", digits = 2)
+            ) %>%
+  mutate(
+    `Est. Lifespan` = `Est. Lifespan` %>% round(2), 
+    `Change in K (K2/K1)` = `Change in K (K2/K1)` %>% round(2), 
+    `log2 Fold Change` = `log2 Fold Change` %>% round(2)
+  ) %>% 
+  mutate(
+    Ancestor = Node %>% 
+    sapply(
+        .,
+        function(n, d=table.cancerSucceptibility.Atlantogenata.genomes){
+            p.node <- d %>% filter(label==n) %>% pull(parent) %>% unique
+            d %>% filter(node == p.node) %>% pull(label) %>% unique %>% return
+        }
+    ) %>% str_replace_all("\\.", " ") %>% str_replace("Root", "Atlantogenata") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora")),
+        Node = Node %>% str_replace_all("\\.", " ") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosorcida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora"))
+  ) %>%
+    arrange(Node) %>% 
+  knitr::kable(
+    ., 
+    longtable = TRUE, 
+    booktabs = TRUE,
+    caption = 'A table generated by the longtable package.'
+  )
+```
+```{r Fig 5, fig.cap="", warning=T}
+warning("TODO: Caption Fig 5")
+warning("TODO: de-comment geom_tiplab in f5a! Commented currently for offline usage!")
 f5
 ```
-
-`r warning("TODO: internally add ")`
+```{r Table 3, fig.cap="Table 3: Model Statistics for PGLS of Cancer Succeptibility"}
+t.size.lifespan.Atlantogenata.lm %>% cat
+```
+```{r Table 4, fig.cap="Table 4: Estimated Cancer Succeptibilty for nodes in Atlantogenata"}
+t.cancersucc.atl
+```
+
 The dramatic increase in body mass and lifespan in some Afrotherian lineages implies those lineages evolved reduced cancer risk.
@@
 Following [@Peto2015b] we estimate the intrinsic cancer risk as the product of risk associated with body mass and lifespan.
-Differences in cancer susceptibility due to body mass differences between species can be approximated simply as the fold difference in body mass ($D$) between species [@Peto2015].
+Differences in cancer susceptibility $K$ due to body mass differences between species can be approximated simply as the fold difference in body mass ($D$) between species [@Peto2015].
 The risk of developing cancer also increases in proportion to the sixth power of age and is approximated by the formula $Ct^6$, in which the proportionality constant C that determines susceptibility to cancer induction is multiplied by the sixth power of the age in years, $t$ [@ArmitageAndDoll1954; @Nordling1953; @Peto2015].
-Thus we can estimate differences in the intrinsic cancer risk between species as $D(Ct^6)$.
-We inferred that susceptibility to cancer induction ($C$) was reduced 3.40x10^3-fold in the ancestor of Pseudoungulata, 7.20×10^3-fold in the ancestor of Paenungulata, 8.81x10^5-fold in the ancestor of Tehthytheria, 1.46x10^8-fold in the ancestor of Proboscidea, and 1.28x10^6-fold in the ancestor of Sirenia relative to the ancestral Afrotherian (Table 3).
-What biological mechanisms underlie the evolution of thousand- to hundred million-fold reductions in cancer susceptibility during the origins of Afrotherians, which are essential for large body size and long lifespan to evolve
+Thus we can estimate the intrinsic cancer risk for a species as $K \approx Dt^6$.
+
+In order to estimate the intrinsic cancer risk of a species, we first obtained estimates for lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon$ (Figure `r warning("make lifespan figure")`).
+With this information in hand, we calculated $K_{1}$ at all nodes, and then estimated the fold change in cancer succeptibility between an ancestral node and a given node as $\frac{K_{2}}{K_{1}}$ (__Table 4__).
+
+As shown in __Table 4__, cancer succeptibility skyrocketed at the initial divergence of Atlantogenata, followed by a generally upwards trend.
+At the common ancestor of Afrotheria there is an inital `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrotheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer risk.
+In parallel to Afrotheria, cancer succeptibility increases `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold in Xenarthra.
+However, cancer risk slowly deflates as size decreases as one moves along the tree towards extant species, such as in Hoffman's Two Toed Sloth (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Choloepus hoffmanii") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change) and in the Nine-banded Armadillo (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Dasypus novemcinctus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change).
+
+Within Afrotheria, cancer succeptibility drops in Afrosorcida as species shrink (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrosorcida") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, then stagnates for the Cape Golden Mole) - but then rises `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Echinops telfairi") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold towards the Lesser Hedgehog Tenrec.
+In parallel, Afroinsectivora does not increase in cancer succeptibility, and decreases once more at the Cape Elephant Shrew (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantulus edwardii") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
+The emergence of Pseudoungulata sees the next big leap in cancer succeptibility with a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Pseudoungulata") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase.
+The Aardvark further increases `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Orycteropus afer") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, while we don't observe an increase at the common ancestor of Paenungulates.
+While the Rock Hyrax decreases in cancer succeptibility as expected (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Procavia capensis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold), Tethytheria sees a sharp increase in cancer risk (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
+Within Tethytheria, the Manatee's cancer risk increases once more `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Trichechus manatus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, while Proboscidae's cancer risk drops precipitously along with its body size (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Proboscidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
+Yet, within Proboscidae we see the biggest increases: right off the bat, we see that the cancer succeptibility of  Elephantidae and the American Mastodon skyrocket by `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammut americanum") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively.
+Both Elephantina and Loxodontini in Elephantidae have a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantina") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer succeptibility.
+Within Elephantina, cancer susceptibility stays stable at Mammuthus and in the Colombian Mammoth, and slightly decreases in the Wooly Mammoth (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammuthus primigenius") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
+The three extant elephants - Asian Elephant in Elephantina, the African Savana Elephant in Loxodontini, and the African Forest Elephant in Loxodona, meanwhile, have parallel and similar decreases in both size and cancer succeptibility (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephas maximus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-, `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Loxodonta africana") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-, and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Loxodonta cyclotis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively).
+Neither the common ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further changes in cancer succeptibility.
 
@@
 
-`r warning("TODO: Discuss Body Size & implications of acordioning of Body Size")`
-`r warning("TODO: Discuss Copy Numbers, specifically how rapidly some gene families expanded, and the issue of genome quality")`
-`r warning("TODO: Discuss effect of genome quality with CNs")`
+What biological mechanisms underlie the evolution of thousand- to hundred million- increases in cancer susceptibility during the origins of Afrotherians, which are essential for large body size and long lifespan to evolve
+
+`r warning("TODO: Body Size & implications of accordioning of Body Size")`
+`r warning("TODO: Next, discuss how the giant initial increases in cancer risk, and the lack of fully commensurate changes even in nodes that decrease in size, may explain why we see TSG duplicates conserved even in smaller species")`
+`r warning("TODO: Lifespan regression, and the possible impact of lacking other data. Key: We have no other data for most species, and would have to estimate them too.")`
+`r warning("TODO: Copy Numbers, specifically how rapidly some gene families expanded, and the issue of genome quality")`
+`r warning("TODO: Effect of genome quality with CNs.\nCHANGE: On second thought, after fixing the model, it doesn't seem any one metric affects CN too much, except for Genome (duh).
+Even Genome though has a much smaller impact than CEGMA stats, although those too don't explain things fully.
+Haven't included yet older genome models though, so this is TBD for now.")`
+`r warning("TODO: ")`
 `r warning("I placed this here, but its half-discussion, half-intro, so edit!")`
@@
 # Conflicts of Interest
-The Authors have no conflicts of interest.
-
-#
+The Authors have no conflicts of interest to report
+
+# Funding Source
+We would like to thank the Department of Human Genetics at the University of Chicago for supporting this project.
 
@@
 
-```{r, Supplementary Figure 3, fig.cap="Supplementary Figure 3: Full version of RecBlat strategy, but low quality"}
+```{r, Supplementary Figure 3, fig.cap="Supplementary Figure 3: Full version of RecBlat strategy, but low quality", warning=T}
 warning("TODO: Make HQ version of this figure")
-p.bodysize.atlantogenata
-
+f2.1
 ```
@@
 ```{r Supplementary Figure 5, fig.cap="Supplementary Figure 5: All the Gene Copy Trees for interesting genes duplicated in LoxAfr4 (RIP any trees if this gets printed)"}
-p.ml.treedata.loxAfrHighlight.plus 
+p.ml.treedata.loxAfrHighlight.plus[1:10]
+warning("Don't print the full list, seriously, its like 600+ pages")
 ```
@@
 ```{r Supplementary Figure 6, fig.cap="Supplementary Figure 6: Expression data for interesting genes duplicated in LoxAfr4"}
-plot.coolgenes
+# plot.coolgenes
 ```
@@
 ```{r Supplementary Figure 7, fig.cap="Supplementary Figure 7: Correlation matrix heat map for genome quality metrics, ECNC, RBHB Copy, and Estimated Copy Number (lesser between ECNC and RBHB)"}
-p.cor.GCT.CEGMA.completeOnly
+p.cor.GCT.CEGMA
+```
+
+```{r Supplementary Figure 8, fig.cap="Supplementary Figure 8: The time-calibrated Eutherian tree used as input for StableTraits", warning=T}
+# p.time.time.tree.Eutheria <-
+#   time.tree %>% 
+#   as_tibble %>% 
+#   mutate(time = branch.length) %>% 
+#   select(-parent, -node, -branch.length) %>% 
+#   full_join(., bsize.tree %>% as_tibble, by="label") %>% 
+#   as.treedata %>% 
+#   ggtree(
+#     branch.length = "time",
+#     aes(color = lnSize),
+#     layout="circular"
+#   ) %>% 
+#   ggtree::revts() + 
+#     geom_highlight(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), fill="goldenrod") + geom_cladelabel(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), label="Atlantogenata", align = F, offset = 2, offset.text = 2, geom = "label", fill="goldenrod") +
+#   geom_highlight(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill="aquamarine") + geom_cladelabel(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill = "aquamarine", label="Chiroptera", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label") +
+#     geom_highlight(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1") + geom_cladelabel(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1", label="Rodentia", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label") +
+#   geom_highlight(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue") + geom_cladelabel(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue", label="Primata", align = F, offset = 2, offset.text = 3, hjust = 0, geom = "label") +
+#   geom_highlight(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1") + geom_cladelabel(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1", label="Cetacea", align = F, offset = 2, offset.text = 2, hjust = 0, geom = "label") +
+#   theme_tree()
+warning("TODO: this causes a segfault, find out why!")
+```
+
+# Supplementary data
+
+```{r Supplementary Table 1, fig.cap="Supplementary Table 1: Parameters from StableTraits", warning=T}
+warning("TODO: Read in Eutheria.progress and present it to the bold.")
 ```

### 2020-05-14 paper_PLOS/paper_PLOS_draft.tex
additional figures, cleaned major errors, added completion warnings

@@
 \usepackage{booktabs}
+\usepackage{geometry}
+\usepackage{pdflscape}
 
@@
 numbers (\textbf{Figure 1A}). The Reciprocal Best Hit (RBH) search
-strategy is conceptually straightforward: 1) Given a gene of interest GA
-in a query genome A, one searches a target genome B for all possible
-matches to GA; 2) For each of these hits, one then performs the
-reciprocal search in the original query genome to identify the
-highest-scoring hit; 3) A hit in genome B is defined as a homolog of
-gene GA if and only if the original gene GA is the top reciprocal search
-hit in genome A. We selected BLAT {[}26{]} as our algorithm of choice,
-as this algorithm is sensitive to highly simliar (\textgreater{}90\%
-identity) sequences, thus identifying the highest-confidence homologs
-while minimizing many-to-one mapping problems when searching for
-multiple genes. RBH performs similar to other more complex methods of
-orthology prediction, and is particularly good at identifying incomplete
-genes that may be fragmented in low quality/poor assembled regions of
-the genome {[}{\textbf{???}},27{]}.
+strategy is conceptually straightforward: 1) Given a gene of interest
+\(G_A\) in a query genome \(A\), one searches a target genome \(B\) for
+all possible matches to \(G_A\); 2) For each of these hits, one then
+performs the reciprocal search in the original query genome to identify
+the highest-scoring hit; 3) A hit in genome \(B\) is defined as a
+homolog of gene \(G_A\) if and only if the original gene \(G_A\) is the
+top reciprocal search hit in genome \(A\). We selected BLAT {[}26{]} as
+our algorithm of choice, as this algorithm is sensitive to highly
+simliar (\textgreater{}90\% identity) sequences, thus identifying the
+highest-confidence homologs while minimizing many-to-one mapping
+problems when searching for multiple genes. RBH performs similar to
+other more complex methods of orthology prediction, and is particularly
+good at identifying incomplete genes that may be fragmented in low
+quality/poor assembled regions of the genome {[}{\textbf{???}},27{]}.
 
@@
 number of nucleotides of the query sequence found overall in each target
-genome (\textbf{Figure 1B}; Supplementary Figure 1). This allows us to
-correct for genes that have been fragmented across incomplete genomes,
-without underestimating the copy number of proteins that lack domains
-found in the human sequence.
+genome (Supplementary Figure 1). This allows us to correct for genes
+that have been fragmented across incomplete genomes, while also taking
+into account missing sequences from the human query in the target
+genome. Mathematically, this can be written as:\\
+\[ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}\] where
+\(n\) is a given nucleotide in the query, \(l\) is the total length of
+the query, \(C_n\) is the number of instances that \(n\) is present
+within a reciprocal best hit, and \(bool(C_n)\) is 1 if \(C_n > 0\) or 0
+if \(C_n = 0\).
 
@@
 Figure 1B. Overall, we identified {[}MEDIAN{]} genes across all species,
-or {[}\%HITS/QUERIES{]} of our starting query genes. As expected, there
-was a correlation between the number of hits identified in a species
-with the evolutionary distance relative to humans {[}Sup Figure 1{]}.
+or {[}\%HITS/QUERIES{]} of our starting query genes.
 
@@
 UP000005640 (17).
+
+\textbf{Orthogonal Genome Assessment using CEGMA} In order to determine
+the effect of genome quality on our results, we used the gVolante
+webserver and CEGMA to assess the quality and completeness of the
+genome. CEGMA was run using the default settings of {[}{]}, and the
+mammalian-specific core gene sets.
 
@@
 overrepresentation analysis (ORA) of the duplicated gene lists relative
-to our initial query gene list. For the database of pathways used in the
-analysis, we used Reactome {[}{\textbf{???}}{]}, Wikipathways and
-Wikipathways\_cancer {[}{\textbf{???}}{]}, and KEGG
-{[}{\textbf{???}}{]}.
+to our initial query gene list {[}{\textbf{???}}{]}. For the database of
+pathways used in the analysis, we used Reactome {[}{\textbf{???}}{]},
+Wikipathways, and Wikipathways\_cancer {[}{\textbf{???}}{]}, and KEGG
+{[}{\textbf{???}}{]}. For the ORA, we used FDR for determining
+significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and
+FDR=0.5.
+
+\hypertarget{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}{%
+\subsection{Lifespan Phylogenetic Least-Square Regression and
+Calculating Estimated Cancer Risk Thoughout
+Atlantogenata}\label{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}}
+
+In order to determine the cancer risk at each node, we used a simplified
+multistage cancer risk model for body size and lifespan
+{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. We defined the
 
@@
 
-\begin{verbatim}
-## Warning: Tried to calculate with group_by(), but the calculation failed.
-## Falling back to ungrouped filter operation...
-
-## Warning: Tried to calculate with group_by(), but the calculation failed.
-## Falling back to ungrouped filter operation...
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
-##   If you'd like to see this geom implemented,
-##   Please open an issue with your example code at
-##   https://github.com/ropensci/plotly/issues
-
-## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
-##   If you'd like to see this geom implemented,
-##   Please open an issue with your example code at
-##   https://github.com/ropensci/plotly/issues
-
-## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
-##   If you'd like to see this geom implemented,
-##   Please open an issue with your example code at
-##   https://github.com/ropensci/plotly/issues
-\end{verbatim}
-
-\hypertarget{step-wise-evolution-of-large-long-lived-afrotherians}{%
-\subsection{Step-wise evolution of large, long-lived
-Afrotherians}\label{step-wise-evolution-of-large-long-lived-afrotherians}}
-
-\begin{verbatim}
-## Warning: Removed 120 rows containing missing values (geom_image).
-\end{verbatim}
+\hypertarget{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}{%
+\subsection{\texorpdfstring{Body size frequently and independently
+expands and contracts throughout
+\emph{Atlantogenata}}{Body size frequently and independently expands and contracts throughout Atlantogenata}}\label{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 1-1.pdf}
+\caption{Figure 1: Body sizes rapidly and frequently expand in
+Eutherians, especially in Atlantogenata. \textbf{A)} Tree of Eutherian
+species, colored by ln(Body Size) and with branch lengths set to the
+rate of change in body sizes, normalized by the square root of the root
+branch. Atlantogenata is highlighted at the bottom. \textbf{B)} Zoom-in
+of (A) on Atlantogenata. Silhuetes for the African Elephant, West Indian
+Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole,
+Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their
+extant body sizes, while clade labels are colored based on the common
+ancestor's estimated body size}
+\end{figure}
 
@@
 
-\begin{verbatim}
-## Warning: Column `Genome` joining factor and character vector, coercing into
-## character vector
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in melt(., na.rm = T): The melt generic in data.table has been passed a
-## matrix and will attempt to redirect to the relevant reshape2 method; please note
-## that reshape2 is deprecated, and this redirection is now deprecated as well.
-## To continue using melt methods from reshape2 while both libraries are attached,
-## e.g. melt.list, you can prepend the namespace like reshape2::melt(.). In the
-## next version, this warning will become an error.
-\end{verbatim}
-
-\begin{verbatim}
-##                                        Var1 Var2        value
-## 1                                      Gene Copy  0.011177676
-## 2                                    Genome Copy -0.104365173
-## 3       Sum length of sequences > 10M  (nt) Copy -0.074960529
-## 4                     Longest sequence (nt) Copy -0.067231077
-## 5                  N50 sequence length (nt) Copy -0.050741921
-## 6                  Base composition (%) (A) Copy -0.001502967
-## 7                  Base composition (%) (T) Copy -0.001379093
-## 8                  Base composition (%) (G) Copy -0.043271156
-## 9                  Base composition (%) (C) Copy -0.042268177
-## 10             Base composition (%) (Other) Copy -0.050228064
-## 11 # of sequences containing non-ACGTN (nt) Copy -0.050621596
-## 12                Mean sequence length (nt) Copy -0.074983744
-## 13              Median sequence length (nt) Copy -0.074897040
-## 14                         Missing Proteins Copy -0.079692396
-## 15                                   #Prots Copy -0.101300994
-## 16                             Completeness Copy -0.101321644
-## 17                 Base composition (%) (N) Copy  0.019768158
-## 18                              totalLength Copy  0.067454870
-## 19       Sum length of sequences > 1M  (nt) Copy  0.051354838
-## 20               # of sequences > 100K (nt) Copy  0.056049593
-## 21                                      L50 Copy  0.082175611
-## 22               # of sequences >   1M (nt) Copy  0.074461957
-## 23               # of sequences >  10M (nt) Copy  0.075679451
-## 24                   Shortest sequence (nt) Copy -0.013216232
-## 25                           GC-content (%) Copy -0.064591581
-## 26                           # of sequences Copy  0.076630197
-## 27               # of sequences >   1K (nt) Copy  0.047474518
-## 28               # of sequences >  10K (nt) Copy  0.045583001
-## 29                                   #Total Copy  0.116391876
-## 30                                  Average Copy  0.124341009
-## 31                                   %Ortho Copy  0.120616153
-## 32                                     Gene ECNC  0.011137324
-## 33                                   Genome ECNC -0.090522003
-## 34      Sum length of sequences > 10M  (nt) ECNC -0.061380983
-## 35                    Longest sequence (nt) ECNC -0.053145168
-## 36                 N50 sequence length (nt) ECNC -0.040007282
-## 37                 Base composition (%) (A) ECNC  0.004626621
-## 38                 Base composition (%) (T) ECNC  0.004862588
-## 39                 Base composition (%) (G) ECNC -0.037787699
-## 40                 Base composition (%) (C) ECNC -0.037045693
-## 41             Base composition (%) (Other) ECNC -0.040274294
-## 42 # of sequences containing non-ACGTN (nt) ECNC -0.040587311
-## 43                Mean sequence length (nt) ECNC -0.061569667
-## 44              Median sequence length (nt) ECNC -0.060928944
-## 45                         Missing Proteins ECNC -0.062209031
-## 46                                   #Prots ECNC -0.084046941
-## 47                             Completeness ECNC -0.084067608
-## 48                 Base composition (%) (N) ECNC  0.013486259
-## 49                              totalLength ECNC  0.061175951
-## 50       Sum length of sequences > 1M  (nt) ECNC  0.046387082
-## 51               # of sequences > 100K (nt) ECNC  0.049326449
-## 52                                      L50 ECNC  0.070430706
-## 53               # of sequences >   1M (nt) ECNC  0.063725329
-## 54               # of sequences >  10M (nt) ECNC  0.062895861
-## 55                   Shortest sequence (nt) ECNC -0.012113794
-## 56                           GC-content (%) ECNC -0.066249117
-## 57                           # of sequences ECNC  0.070289641
-## 58               # of sequences >   1K (nt) ECNC  0.042151677
-## 59               # of sequences >  10K (nt) ECNC  0.041222984
-## 60                                   #Total ECNC  0.102797584
-## 61                                  Average ECNC  0.108543367
-## 62                                   %Ortho ECNC  0.104394790
-## 63                                     Gene   CN  0.013089940
-## 64                                   Genome   CN -0.092501845
-## 65      Sum length of sequences > 10M  (nt)   CN -0.061610123
-## 66                    Longest sequence (nt)   CN -0.052591263
-## 67                 N50 sequence length (nt)   CN -0.038770715
-## 68                 Base composition (%) (A)   CN  0.006139187
-## 69                 Base composition (%) (T)   CN  0.006354445
-## 70                 Base composition (%) (G)   CN -0.036741673
-## 71                 Base composition (%) (C)   CN -0.035971303
-## 72             Base composition (%) (Other)   CN -0.041544748
-## 73 # of sequences containing non-ACGTN (nt)   CN -0.041868348
-## 74                Mean sequence length (nt)   CN -0.063786387
-## 75              Median sequence length (nt)   CN -0.063541494
-## 76                         Missing Proteins   CN -0.063703014
-## 77                                   #Prots   CN -0.085286259
-## 78                             Completeness   CN -0.085306203
-## 79                 Base composition (%) (N)   CN  0.012122309
-## 80                              totalLength   CN  0.060342667
-## 81       Sum length of sequences > 1M  (nt)   CN  0.044889905
-## 82               # of sequences > 100K (nt)   CN  0.047874628
-## 83                                      L50   CN  0.069838784
-## 84               # of sequences >   1M (nt)   CN  0.062839574
-## 85               # of sequences >  10M (nt)   CN  0.062215472
-## 86                   Shortest sequence (nt)   CN -0.013168670
-## 87                           GC-content (%)   CN -0.066875760
-## 88                           # of sequences   CN  0.073563144
-## 89               # of sequences >   1K (nt)   CN  0.044231140
-## 90               # of sequences >  10K (nt)   CN  0.042436480
-## 91                                   #Total   CN  0.105113732
-## 92                                  Average   CN  0.110764419
-## 93                                   %Ortho   CN  0.106456398
-\end{verbatim}
-
-\includegraphics{paper_PLOS_draft_files/figure-latex/Correlations between genome quality scores-1.pdf}
-
-\begin{verbatim}
-## Warning: Column `Genome`/`BestGenome` joining factor and character vector,
-## coercing into character vector
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: `as.tibble()` is deprecated as of tibble 2.0.0.
-## Please use `as_tibble()` instead.
-## The signature and semantics have changed, see `?as_tibble`.
-## This warning is displayed once every 8 hours.
-## Call `lifecycle::last_warnings()` to see where this warning was generated.
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 1 rows containing missing values (geom_text).
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2-1.pdf}
-\caption{Figure 2: A Reciprocal Best-Hit BLAT pipeline for identifying
-gene copy number in other genomes. A) A graphic summary of the
-reciprocal best-hit strategy. B) Estimated Copy Number by Coverage. C)}
-\end{figure}
-
-\begin{verbatim}
-## Warning: Table 2
+\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2C1-1.pdf}
+\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2C2-1.pdf}
+
+\begin{verbatim}
+## Warning: TODO: Table 2 is done, but Stargazer has some weird bug stopping it
+## from running...
 \end{verbatim}
@@
 ratio between observed and expected gene coverage per nucleotide (ECNC)
-(Fig. 2B; Sup. Fig. 2A-D). By only including nucleotides from the query
+(Supplementary Figure 1). By only including nucleotides from the query
 sequence that were observed in the target genome, we also correct for
@@
 
+\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 3-1.pdf}
+\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 3-2.pdf}
+
 In order to infer the functional consequences of these gene
@@
 pathways relative to our initial query set of genes. We used
-
-\hypertarget{retroduplication-of-eef1a1-in-the-stem-lineage-of-tehthytheria}{%
-\subsection{Retroduplication of EEF1A1 in the stem-lineage of
-Tehthytheria}\label{retroduplication-of-eef1a1-in-the-stem-lineage-of-tehthytheria}}
-
-\begin{verbatim}
-## Warning: Fig 3: EEF1A1
-\end{verbatim}
-
-Previous studies have identified genes whose transgenic overexpression
-delays aging related phenotypes and extends lifespan. Overexpression of
-the eukaryotic translation elongation factor 1 alpha 1 (EEF1A1) gene,
-for example, extends lifespan in transgenic Drosophila. Like most
-mammalian genomes, we found that Afrotherian genomes encoded numerous
-EEF1A1 retrogenes (Fig. XA). While the majority of these retrogenes have
-premature stop codons and insertions/deletions characteristic of
-pseudogenes, at least 13 Tehthytherian-specific EEF1A1 retrogenes have
-the potential to encode functional proteins (Fig. XB). Unfortunately it
-is difficult to determine if these retrogenes are transcribed because
-they have relatively high sequence similarity (83-88\%), suggesting
-evidence of transcription using RNA-Seq data is likely an artifact.
-Indeed, nearly all of the EEF1A1 retrogene transcripts assembled by
-StringTie were confined to their gene body and UTRs and thus lacked
-unique sequences. The EEF1A1RTG13 transcript, however, initiates within
-an upstream MLT1B retrovirus-like MaLR transposable element and includes
-unique 5?- and 3?-UTR sequences, indicating at least one EEF1A1
-retrogene is transcribed (Fig. XC). EEF1A1
-
-Candidate gene studies, for example, have identified functional
-duplicates of the tumor suppressors TP53 and LIF in elephants. In a
-larger candidate gene study, Caulin et al.~characterized the copy number
-of 830 known tumor-suppressor genes across 36 mammals and identified 382
-putative duplicates, including duplicates in species with large body
-sizes and long life-spans. however, the probability of developing cancer
-is similar for small, short-lived mammals such as mice and for large,
-long-lived mammals such as elephants.
-
-In stark contrast, genome-wide studies of unusually large or long-lived
-species such as the bowhead whale (Keane et al., 2015), Myotid bats
-(Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al.,
-2011), and blind mole rat (Fang et al., 2014) did not find an over
-representation of tumor suppressors among duplicate genes.
-
-A genomic analysis of genetic changes associated with the evolution of
-enhanced cancer resistance in the elephant lineage has yet to be
-performed. Thus it is not clear if the duplication of TP53 and LIF
-reflects a general pattern of tumor suppressor duplication in the
-elephant lineage, unlike other lineages that resolved Peto?s paradox, or
-from the kinds of ascertainment biases common in candidate gene studies.
 
@@
 \begin{verbatim}
-## Warning: Fig4
+## Warning: TODO: Make Fig4
 \end{verbatim}
@@
 \begin{verbatim}
-## # A tibble: 253 x 6
-##     parent  node branch.length label                       lnSize Lifespan
-##      <int> <int>         <dbl> <chr>                        <dbl>    <dbl>
-##   1    135     1     40.1      Geogale.aurita                   2    1.72 
-##   2    140     2    122.       Limnogale.mergulus               5    2.30 
-##   3    140     3      0.135    Microgale.taiva                  2    1.72 
-##   4    139     4     33.5      Microgale.parvula                1    1.52 
-##   5    138     5      0.595    Microgale.brevicaudata           2    1.72 
-##   6    142     6      0.571    Microgale.cowani                 3    1.91 
-##   7    142     7      1.07     Microgale.jobihely               2    1.72 
-##   8    145     8      0.772    Microgale.longicaudata           2    1.72 
-##   9    145     9      0.173    Microgale.principula             2    1.72 
-##  10    144    10     34.3      Microgale.pusilla                1    1.52 
-##  11    143    11     61.4      Microgale.talazaci               4    1.76 
-##  12    136    12     22.8      Oryzorictes.hova                 4    2.11 
-##  13    147    13      7.75     Echinops.telfairi                4    2.94 
-##  14    147    14     12.5      Setifer.setosus                  6    2.65 
-##  15    148    15      4.68     Hemicentetes.semispinosus        5    0.993
-##  16    148    16     79.5      Tenrec.ecaudatus                 7    2.16 
-##  17    150    17      2.82     Micropotamogale.lamottei         4    2.11 
-##  18    150    18    103.       Potamogale.velox                 6    2.50 
-##  19    151    19     69.8      Protenrec.tricuspis              1    1.52 
-##  20    151    20      2.21     Dilambdogale.gheerbranti         2    1.72 
-##  21    157    21      0.454    Carpitalpa.arendsi               4    2.11 
-##  22    158    22      3.26     Neamblysomus.gunningi            4    2.11 
-##  23    158    23      5.72     Neamblysomus.julianae            3    1.91 
-##  24    162    24      0.976    Amblysomus.corriae               4    2.11 
-##  25    162    25      1.38     Amblysomus.robustus              4    2.11 
-##  26    161    26      0.524    Amblysomus.septentrionalis       4    2.11 
-##  27    160    27      0.0152   Amblysomus.hottentotus           4    2.11 
-##  28    159    28      3.79     Amblysomus.marleyi               4    2.11 
-##  29    163    29      0.321    Chlorotalpa.duthieae             3    1.91 
-##  30    163    30      0.0874   Chlorotalpa.sclateri             4    2.11 
-##  31    165    31     62.8      Chrysospalax.trevelyani          6    2.50 
-##  32    165    32      0.130    Chrysospalax.villosus            5    2.30 
-##  33    164    33     13.4      Calcochloris.obtusirostris       3    1.91 
-##  34    153    34      9.65     Eremitalpa.granti                3    1.91 
-##  35    167    35      3.34     Chrysochloris.asiatica           4    2.11 
-##  36    167    36      5.78     Cryptochloris.wintoni            3    1.91 
-##  37    166    37      0.333    Chrysochloris.stuhlmanni         4    2.11 
-##  38    175    38      0.926    Elephantulus.brachyrhynchus      4    2.11 
-##  39    175    39      0.118    Elephantulus.rufescens           4    2.07 
-##  40    176    40      1.15     Elephantulus.intufi              4    2.23 
-##  41    176    41      0.319    Elephantulus.rupestris           4    2.11 
-##  42    177    42      0.237    Elephantulus.edwardii            4    2.11 
-##  43    177    43      0.953    Elephantulus.myurus              4    2.11 
-##  44    179    44     10.5      Elephantulus.rozeti              4    1.97 
-##  45    179    45     24.6      Petrodromus.tetradactylus        5    2.30 
-##  46    178    46     14.2      Macroscelides.proboscideus       4    2.16 
-##  47    181    47      1.93     Elephantulus.fuscipes            4    2.11 
-##  48    182    48      0.678    Elephantulus.fuscus              4    2.11 
-##  49    182    49      1.10     Elephantulus.revoili             3    1.91 
-##  50    184    50      0.403    Rhynchocyon.chrysopygus          6    2.40 
-##  51    185    51      3.30     Rhynchocyon.cirnei               6    2.50 
-##  52    185    52      4.33     Rhynchocyon.udzungwensis         7    2.69 
-##  53    183    53      2.13     Rhynchocyon.petersi              6    2.50 
-##  54    170    54     15.9      Metoldobotes.sp..nov             6    2.50 
-##  55    186    55     11.8      Chambius.kasserinensis           4    2.11 
-##  56    186    56     11.6      Herodotius.pattersoni            5    2.30 
-##  57    188    57      6.59     Orycteropus.afer                11    3.39 
-##  58    188    58      0.569    Myorycteropus.africanus         10    3.27 
-##  59    194    59      1.86     Dendrohyrax.dorsalis             8    2.88 
-##  60    195    60      1.64     Heterohyrax.brucei               8    2.88 
-##  61    195    61      0.000068 Procavia.capensis                8    2.69 
-##  62    193    62      1.15     Thyrohyrax.domorictus            9    3.08 
-##  63    192    63      3.55     Thyrohyrax.meyeri                9    3.08 
-##  64    198    64     28.6      Thyrohyrax.litholagus           10    3.27 
-##  65    201    65      2.29     Pachyhyrax.crassidentatus       12    3.66 
-##  66    201    66      1.45     Bunohyrax.fajumensis            11    3.47 
-##  67    202    67      5.44     Geniohyus.mirus                 11    3.47 
-##  68    202    68     11.4      Bunohyrax.major                 12    3.66 
-##  69    203    69      0.235    Megalohyrax.eocaenus            12    3.66 
-##  70    203    70      7.21     Megalohyrax.sp..nov             13    3.86 
-##  71    207    71      0.0699   Titanohyrax.andrewsi            13    3.86 
-##  72    207    72     34.6      Titanohyrax.ultimus             14    4.05 
-##  73    206    73      0.0353   Titanohyrax.angustidens         12    3.66 
-##  74    205    74     13.7      Antilohyrax.pectidens           11    3.47 
-##  75    204    75      0.194    Afrohyrax.championi             11    3.47 
-##  76    209    76      2.90     Saghatherium.antiquum           10    3.27 
-##  77    209    77     15.0      Selenohyrax.chatrathi           11    3.47 
-##  78    208    78     15.8      Saghatherium.bowni               9    3.08 
-##  79    211    79      2.82     Seggeurius.amourensis            8    2.88 
-##  80    211    80      0.733    Microhyrax.lavocati              8    2.88 
-##  81    210    81     18.2      Dimaitherium.patnaiki           10    3.27 
-##  82    220    82      5.81     Elephas.maximus                 15    4.18 
-##  83    222    83      6.12     Elephas.antiquus.falconeri      12    3.66 
-##  84    222    84      1.90     Elephas.cypriotes               12    3.66 
-##  85    221    85      0.340    Elephas.maximus.asurus          16    4.44 
-##  86    223    86     16.0      Mammuthus.trogontherii          16    4.44 
-##  87    224    87      0.910    Mammuthus.columbi               16    4.44 
-##  88    224    88      2.17     Mammuthus.primigenius           15    4.24 
-##  89    225    89      1.28     Loxodonta.africana              15    4.17 
-##  90    227    90      0.0149   Palaeoloxodon.antiquus          16    4.44 
-##  91    227    91     12.8      Palaeoloxodon.namadicus         17    4.63 
-##  92    226    92      3.72     Loxodonta.cyclotis              15    4.24 
-##  93    228    93     15.3      Mammut.borsoni                  16    4.44 
-##  94    228    94      0.951    Mammut.americanum               16    4.44 
-##  95    216    95      3.64     Phiomia.spp                     14    4.05 
-##  96    230    96      0.0280   Omanitherium.dhofarensis        13    3.86 
-##  97    230    97      7.29     Arcanotherium.savagei           13    3.86 
-##  98    229    98     73.6      Barytherium.spp                 15    4.24 
-##  99    214    99      5.71     Moeritherium.spp                14    4.05 
-## 100    231   100    326.       Phosphatherium.escuilliei       10    3.27 
-## 101    232   101      2.29     Numidotherium.koholense         13    3.86 
-## 102    232   102      0.742    Daouitherium.rebouli            13    3.86 
-## 103    235   103      5.85     Dugong.dugon                    13    4.29 
-## 104    235   104    173.       Hydrodamalis.gigas              16    4.44 
-## 105    237   105      0.688    Trichechus.inunguis             13    3.69 
-## 106    237   106      1.39     Trichechus.manatus              13    4.23 
-## 107    236   107      0.574    Trichechus.senegalensis         13    3.86 
-## 108    238   108     13.6      Prorastomus.sirenoides          11    3.47 
-## 109    238   109     33.7      Protosiren.smithae              13    3.86 
-## 110    241   110      0.640    Choloepus.didactylus             9    3.61 
-## 111    241   111      0.316    Choloepus.hoffmanni              8    3.71 
-## 112    242   112      0.0297   Bradypus.torquatus               8    2.88 
-## 113    243   113      0.483    Bradypus.tridactylus             8    2.88 
-## 114    243   114      0.512    Bradypus.variegatus              8    2.88 
-## 115    245   115     10.4      Tamandua.tetradactyla            9    2.94 
-## 116    245   116     41.0      Myrmecophaga.tridactyla         10    3.43 
-## 117    244   117      2.15     Cyclopes.didactylus              9    3.08 
-## 118    249   118     15.0      Euphractus.sexcinctus            8    3.10 
-## 119    249   119      5.54     Zaedyus.pichiy                   7    2.53 
-## 120    250   120      0.0899   Chaetophractus.nationi           8    2.88 
-## 121    250   121      0.418    Chaetophractus.villosus          8    3.23 
-## 122    251   122    268.       Priodontes.maximus              11    2.71 
-## 123    252   123     16.0      Tolypeutes.matacus               7    3.61 
-## 124    252   124      0.248    Cabassous.centralis              8    2.88 
-## 125    253   125      4.05     Dasypus.hybridus                 7    2.69 
-## 126    253   126     74.1      Dasypus.kappleri                 9    3.08 
-## 127    253   127     14.7      Dasypus.novemcinctus             8    3.10 
-## 128    253   128     12.2      Dasypus.sabanicola               7    2.69 
-## 129    253   129      4.44     Dasypus.septemcinctus            7    2.82 
-## 130    130   130     NA        Atlantogenata                    2    1.72 
-## 131    130   131      0.000982 Afrotheria                       6    2.50 
-## 132    131   132     44.5      Afrosorcida                      4    2.11 
-## 133    132   133     11.8      <NA>                             4    2.11 
-## 134    133   134      0.420    <NA>                             4    2.11 
-## 135    134   135     10.1      <NA>                             3    1.91 
-## 136    135   136      0.373    <NA>                             3    1.91 
-## 137    136   137      2.96     <NA>                             3    1.91 
-## 138    137   138      2.69     <NA>                             3    1.91 
-## 139    138   139      0.450    <NA>                             2    1.72 
-## 140    139   140      3.36     <NA>                             3    1.91 
-## 141    137   141      0.153    <NA>                             3    1.91 
-## 142    141   142      0.116    <NA>                             2    1.72 
-## 143    141   143      0.222    <NA>                             2    1.72 
-## 144    143   144      0.771    <NA>                             2    1.72 
-## 145    144   145      0.109    <NA>                             3    1.91 
-## 146    134   146     38.5      <NA>                             5    2.30 
-## 147    146   147      1.81     <NA>                             3    1.91 
-## 148    146   148      5.26     <NA>                             3    1.91 
-## 149    133   149      4.28     <NA>                             3    1.91 
-## 150    149   150     52.3      <NA>                             5    2.30 
-## 151    149   151     15.6      <NA>                             3    1.91 
-## 152    132   152     12.0      <NA>                             4    2.11 
-## 153    152   153      0.0295   <NA>                             4    2.11 
-## 154    153   154      0.173    <NA>                             3    1.91 
-## 155    154   155      0.0141   <NA>                             3    1.91 
-## 156    155   156      0.107    <NA>                             4    2.11 
-## 157    156   157      0.00544  <NA>                             3    1.91 
-## 158    157   158      0.102    <NA>                             3    1.91 
-## 159    156   159      0.150    <NA>                             3    1.91 
-## 160    159   160      0.483    <NA>                             3    1.91 
-## 161    160   161      0.321    <NA>                             3    1.91 
-## 162    161   162      0.00743  <NA>                             4    2.11 
-## 163    155   163      2.12     <NA>                             3    1.91 
-## 164    154   164      0.774    <NA>                             4    2.11 
-## 165    164   165     18.6      <NA>                             5    2.30 
-## 166    152   166      0.260    <NA>                             3    1.91 
-## 167    166   167      0.155    <NA>                             3    1.91 
-## 168    131   168      0.00016  <NA>                             6    2.50 
-## 169    168   169      2.49     Macroscelidae                    5    2.30 
-## 170    169   170      0.105    <NA>                             5    2.30 
-## 171    170   171      4.23     <NA>                             5    2.30 
-## 172    171   172      6.20     <NA>                             3    1.91 
-## 173    172   173      0.977    <NA>                             3    1.91 
-## 174    173   174      0.448    <NA>                             4    2.11 
-## 175    174   175      0.349    <NA>                             4    2.11 
-## 176    174   176      0.176    <NA>                             4    2.11 
-## 177    173   177      1.62     <NA>                             4    2.11 
-## 178    172   178      0.155    <NA>                             3    1.91 
-## 179    178   179      0.180    <NA>                             4    2.11 
-## 180    171   180      0.100    <NA>                             5    2.30 
-## 181    180   181     11.2      <NA>                             3    1.91 
-## 182    181   182     12.0      <NA>                             3    1.91 
-## 183    180   183     30.6      <NA>                             3    1.91 
-## 184    183   184      2.34     <NA>                             3    1.91 
-## 185    184   185      0.0165   <NA>                             3    1.91 
-## 186    169   186      4.53     <NA>                             5    2.30 
-## 187    168   187    546.       Pseudoungulata                  10    3.27 
-## 188    187   188     12.9      <NA>                            10    3.27 
-## 189    187   189      4.42     Paenungulata                    10    3.27 
-## 190    189   190      6.90     <NA>                             5    2.30 
-## 191    190   191      1.47     <NA>                            10    3.27 
-## 192    191   192     17.9      <NA>                             9    3.08 
-## 193    192   193      0.00758  <NA>                             9    3.08 
-## 194    193   194     19.5      <NA>                             5    2.30 
-## 195    194   195      2.67     <NA>                             5    2.30 
-## 196    191   196      5.61     <NA>                             3    1.91 
-## 197    196   197      8.46     <NA>                             3    1.91 
-## 198    197   198      0.602    <NA>                            11    3.47 
-## 199    198   199      8.81     <NA>                            12    3.66 
-## 200    199   200      0.347    <NA>                             3    1.91 
-## 201    200   201      0.177    <NA>                            12    3.66 
-## 202    200   202      0.584    <NA>                             3    1.91 
-## 203    199   203      8.90     <NA>                            12    3.66 
-## 204    197   204      9.93     <NA>                            11    3.47 
-## 205    204   205      1.23     <NA>                            12    3.66 
-## 206    205   206     22.2      <NA>                             3    1.91 
-## 207    206   207     10.2      <NA>                            13    3.86 
-## 208    196   208      6.63     <NA>                             3    1.91 
-## 209    208   209      1.16     <NA>                            10    3.27 
-## 210    190   210     22.7      <NA>                             9    3.08 
-## 211    210   211      8.32     <NA>                             8    2.88 
-## 212    189   212    187.       Tethytheria                     13    3.86 
-## 213    212   213      3.40     <NA>                             3    1.91 
-## 214    213   214      6.37     <NA>                            13    3.86 
-## 215    214   215      1.15     <NA>                            14    4.05 
-## 216    215   216     13.4      <NA>                             3    1.91 
-## 217    216   217     30.3      Proboscidae                      3    1.91 
-## 218    217   218      2.21     Elephantidae                    15    4.24 
-## 219    218   219      0.0116   Elephantina                     16    4.44 
-## 220    219   220      0.128    <NA>                            15    4.24 
-## 221    220   221      0.369    <NA>                             3    1.91 
-## 222    221   222    400.       <NA>                             3    1.91 
-## 223    219   223      0.645    <NA>                             3    1.91 
-## 224    223   224      0.467    Mammuthus                       16    4.44 
-## 225    218   225      0.108    Loxodontini                     16    4.44 
-## 226    225   226      0.862    Loxodona                        16    4.44 
-## 227    226   227      6.09     <NA>                             3    1.91 
-## 228    217   228      9.63     <NA>                             3    1.91 
-## 229    215   229      0.197    <NA>                             3    1.91 
-## 230    229   230      2.47     <NA>                             3    1.91 
-## 231    213   231      0.128    <NA>                            13    3.86 
-## 232    231   232      0.268    <NA>                             3    1.91 
-## 233    212   233      0.309    <NA>                            12    3.66 
-## 234    233   234     16.4      <NA>                             3    1.91 
-## 235    234   235      0.530    <NA>                             3    1.91 
-## 236    234   236      0.134    <NA>                             3    1.91 
-## 237    236   237      0.145    <NA>                             3    1.91 
-## 238    233   238      3.59     <NA>                             3    1.91 
-## 239    130   239    125.       Xenarthra                       11    3.47 
-## 240    239   240     14.6      <NA>                             3    1.91 
-## 241    240   241      3.04     <NA>                             3    1.91 
-## 242    240   242      0.0660   <NA>                            11    3.47 
-## 243    242   243      0.112    <NA>                            11    3.47 
-## 244    239   244     15.1      <NA>                             3    1.91 
-## 245    244   245     21.7      <NA>                             9    3.08 
-## 246    239   246      0.958    <NA>                            10    3.27 
-## 247    246   247      0.0602   <NA>                             3    1.91 
-## 248    247   248      0.0914   <NA>                             3    1.91 
-## 249    248   249      0.000579 <NA>                             8    2.88 
-## 250    248   250      0.000908 <NA>                            11    3.47 
-## 251    247   251      0.217    <NA>                             3    1.91 
-## 252    251   252      0.0523   <NA>                            11    3.47 
-## 253    246   253      0.140    <NA>                             8    2.88
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Fig5
-\end{verbatim}
+## Warning: TODO: Caption Fig 5
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: TODO: de-comment geom_tiplab in f5a! Commented currently for offline
+## usage!
+\end{verbatim}
+
+\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 5-1.pdf}
+
+\begin{verbatim}
+## 
+## % Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
+## % Date and time: Wed, May 13, 2020 - 20:08:58
+## \begin{table}[!htbp] \centering 
+##   \caption{Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression} 
+##   \label{} 
+## \begin{tabular}{@{\extracolsep{5pt}}lc} 
+## \\[-1.8ex]\hline 
+## \hline \\[-1.8ex] 
+##  & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
+## \cline{2-2} 
+## \\[-1.8ex] & Lifespan \\ 
+## \hline \\[-1.8ex] 
+##  lnSize & 0.100 \\ 
+##   & (0.121) \\ 
+##   & t = 0.826 \\ 
+##   & p = 0.409 \\ 
+##   Constant & 1.943 \\ 
+##   & (1.385) \\ 
+##   & t = 1.403 \\ 
+##   & p = 0.161 \\ 
+##  \hline \\[-1.8ex] 
+## Observations & 28 \\ 
+## Log Likelihood & $-$39.636 \\ 
+## Akaike Inf. Crit. & 85.273 \\ 
+## Bayesian Inf. Crit. & 89.047 \\ 
+## \hline 
+## \hline \\[-1.8ex] 
+## \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+## \end{tabular} 
+## \end{table}
+\end{verbatim}
+
+\begin{longtable}[]{@{}lrrllrrl@{}}
+\caption{A table generated by the longtable package.}\tabularnewline
+\toprule
+Node & lnSize & Est. Lifespan & Est. Cancer Susceptibility (K1) & Est.
+Ancestral K2 & Change in K (K2/K1) & log2 Fold Change &
+Ancestor\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Node & lnSize & Est. Lifespan & Est. Cancer Susceptibility (K1) & Est.
+Ancestral K2 & Change in K (K2/K1) & log2 Fold Change &
+Ancestor\tabularnewline
+\midrule
+\endhead
+Loxodontini & 16 & 34.38 & 1.47e+16 & 2.97e+15 & 4.940000e+00 & 2.31 &
+Elephantidae\tabularnewline
+Loxodonta africana & 15 & 65.00 & 2.47e+17 & 1.47e+16 & 1.681000e+01 &
+4.07 & Loxodontini\tabularnewline
+Loxodona & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 & 0.00 &
+Loxodontini\tabularnewline
+Loxodonta cyclotis & 15 & 31.12 & 2.97e+15 & 1.47e+16 & 2.000000e-01 &
+-2.31 & Loxodona\tabularnewline
+Palaeoloxodon antiquus & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00
+& 0.00 & Loxodona\tabularnewline
+Elephantidae & 15 & 31.12 & 2.97e+15 & 1.40e+07 & 2.125534e+08 & 27.66 &
+NA\tabularnewline
+Elephantina & 16 & 34.38 & 1.47e+16 & 2.97e+15 & 4.940000e+00 & 2.31 &
+Elephantidae\tabularnewline
+Elephas maximus & 15 & 65.50 & 2.58e+17 & 1.47e+16 & 1.760000e+01 & 4.14
+& Elephantina\tabularnewline
+Mammuthus & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 & 0.00 &
+Elephantina\tabularnewline
+Mammuthus primigenius & 15 & 31.12 & 2.97e+15 & 1.47e+16 & 2.000000e-01
+& -2.31 & Mammuthus\tabularnewline
+Mammuthus columbi & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 &
+0.00 & Mammuthus\tabularnewline
+Mammut americanum & 16 & 34.38 & 1.47e+16 & 1.40e+07 & 1.050567e+09 &
+29.97 & NA\tabularnewline
+Tethytheria & 13 & 25.49 & 1.21e+14 & 1.01e+12 & 1.207400e+02 & 6.92 &
+Paenungulata\tabularnewline
+Trichechus manatus & 13 & 69.00 & 4.77e+16 & 1.21e+14 & 3.929800e+02 &
+8.62 & Tethytheria\tabularnewline
+Paenungulata & 10 & 18.91 & 1.01e+12 & 1.01e+12 & 1.000000e+00 & 0.00 &
+Pseudoungulata\tabularnewline
+Procavia capensis & 8 & 14.80 & 3.13e+10 & 1.01e+12 & 3.000000e-02 &
+-5.01 & Paenungulata\tabularnewline
+Pseudoungulata & 10 & 18.91 & 1.01e+12 & 1.69e+09 & 5.967900e+02 & 9.22
+& Afroinsectivora\tabularnewline
+Orycteropus afer & 11 & 29.80 & 4.19e+13 & 1.01e+12 & 4.167000e+01 &
+5.38 & Pseudoungulata\tabularnewline
+Elephantulus edwardii & 4 & 10.40 & 6.90e+07 & 1.69e+09 & 4.000000e-02 &
+-4.61 & Afroinsectivora\tabularnewline
+Afrosorcida & 4 & 10.40 & 6.90e+07 & 1.69e+09 & 4.000000e-02 & -4.61 &
+Afrotheria\tabularnewline
+Chrysochloris asiatica & 4 & 10.40 & 6.90e+07 & 6.90e+07 & 1.000000e+00
+& 0.00 & NA\tabularnewline
+Echinops telfairi & 4 & 19.00 & 2.57e+09 & 6.90e+07 & 3.722000e+01 &
+5.22 & NA\tabularnewline
+Afrotheria & 6 & 12.69 & 1.69e+09 & 2.83e+06 & 5.967900e+02 & 9.22 &
+Atlantogenata\tabularnewline
+Xenarthra & 11 & 20.89 & 4.97e+12 & 2.83e+06 & 1.760358e+06 & 20.75 &
+Atlantogenata\tabularnewline
+Dasypus novemcinctus & 8 & 22.30 & 3.67e+11 & 4.97e+12 & 7.000000e-02 &
+-3.76 & Xenarthra\tabularnewline
+Choloepus hoffmanni & 8 & 41.00 & 1.42e+13 & 4.97e+12 & 2.850000e+00 &
+1.51 & Xenarthra\tabularnewline
+Atlantogenata & 2 & 8.52 & 2.83e+06 & 2.83e+06 & 1.000000e+00 & 0.00 &
+Atlantogenata\tabularnewline
+Afroinsectivora & 6 & 12.69 & 1.69e+09 & 1.69e+09 & 1.000000e+00 & 0.00
+& Afrotheria\tabularnewline
+NA & 3 & 9.41 & 1.40e+07 & 1.21e+14 & 0.000000e+00 & -23.05 &
+Tethytheria\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 the product of risk associated with body mass and lifespan. Differences
-in cancer susceptibility due to body mass differences between species
-can be approximated simply as the fold difference in body mass (\(D\))
-between species {[}{\textbf{???}}{]}. The risk of developing cancer also
-increases in proportion to the sixth power of age and is approximated by
-the formula \(Ct^6\), in which the proportionality constant C that
-determines susceptibility to cancer induction is multiplied by the sixth
-power of the age in years, \(t\)
+in cancer susceptibility \(K\) due to body mass differences between
+species can be approximated simply as the fold difference in body mass
+(\(D\)) between species {[}{\textbf{???}}{]}. The risk of developing
+cancer also increases in proportion to the sixth power of age and is
+approximated by the formula \(Ct^6\), in which the proportionality
+constant C that determines susceptibility to cancer induction is
+multiplied by the sixth power of the age in years, \(t\)
 {[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. Thus we can estimate
-differences in the intrinsic cancer risk between species as \(D(Ct^6)\).
-We inferred that susceptibility to cancer induction (\(C\)) was reduced
-3.40x10\^{}3-fold in the ancestor of Pseudoungulata, 7.20×10\^{}3-fold
-in the ancestor of Paenungulata, 8.81x10\^{}5-fold in the ancestor of
-Tehthytheria, 1.46x10\^{}8-fold in the ancestor of Proboscidea, and
-1.28x10\^{}6-fold in the ancestor of Sirenia relative to the ancestral
-Afrotherian (Table 3). What biological mechanisms underlie the evolution
-of thousand- to hundred million-fold reductions in cancer susceptibility
-during the origins of Afrotherians, which are essential for large body
-size and long lifespan to evolve
-
-\hypertarget{other-figures}{%
-\section{OTHER FIGURES}\label{other-figures}}
+the intrinsic cancer risk for a species as \(K \approx Dt^6\).
+
+In order to estimate the intrinsic cancer risk of a species, we first
+obtained estimates for lifespans at ancestral nodes using PGLS and the
+model
+\(ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon\)
+(Figure ). With this information in hand, we calculated \(K_{1}\) at all
+nodes, and then estimated the fold change in cancer succeptibility
+between an ancestral node and a given node as \(\frac{K_{2}}{K_{1}}\)
+(\textbf{Table 4}).
+
+As shown in \textbf{Table 4}, cancer succeptibility skyrocketed at the
+initial divergence of Atlantogenata, followed by a generally upwards
+trend. At the common ancestor of Afrotheria there is an inital 9.22-fold
+increase in cancer risk. In parallel to Afrotheria, cancer
+succeptibility increases 20.75-fold in Xenarthra. However, cancer risk
+slowly deflates as size decreases as one moves along the tree towards
+extant species, such as in Hoffman's Two Toed Sloth (-fold change) and
+in the Nine-banded Armadillo (-fold change).
+
+Within Afrotheria, cancer succeptibility drops in Afrosorcida as species
+shrink (-4.61-fold, then stagnates for the Cape Golden Mole) - but then
+rises -fold towards the Lesser Hedgehog Tenrec. In parallel,
+Afroinsectivora does not increase in cancer succeptibility, and
+decreases once more at the Cape Elephant Shrew (-fold). The emergence of
+Pseudoungulata sees the next big leap in cancer succeptibility with a
+9.22-fold increase. The Aardvark further increases -fold, while we don't
+observe an increase at the common ancestor of Paenungulates. While the
+Rock Hyrax decreases in cancer succeptibility as expected (-fold),
+Tethytheria sees a sharp increase in cancer risk (6.92-fold). Within
+Tethytheria, the Manatee's cancer risk increases once more -fold, while
+Proboscidae's cancer risk drops precipitously along with its body size
+(-23.05-fold). Yet, within Proboscidae we see the biggest increases:
+right off the bat, we see that the cancer succeptibility of Elephantidae
+and the American Mastodon skyrocket by 27.66-fold and -fold,
+respectively. Both Elephantina and Loxodontini in Elephantidae have a
+2.31-fold increase in cancer succeptibility. Within Elephantina, cancer
+susceptibility stays stable at Mammuthus and in the Colombian Mammoth,
+and slightly decreases in the Wooly Mammoth (-fold). The three extant
+elephants - Asian Elephant in Elephantina, the African Savana Elephant
+in Loxodontini, and the African Forest Elephant in Loxodona, meanwhile,
+have parallel and similar decreases in both size and cancer
+succeptibility (-, -, and -fold, respectively). Neither the common
+ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further
+changes in cancer succeptibility.
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+What biological mechanisms underlie the evolution of thousand- to
+hundred million- increases in cancer susceptibility during the origins
+of Afrotherians, which are essential for large body size and long
+lifespan to evolve
+
+Candidate gene studies, for example, have identified functional
+duplicates of the tumor suppressors TP53 and LIF in elephants. In a
+larger candidate gene study, Caulin et al.~characterized the copy number
+of 830 known tumor-suppressor genes across 36 mammals and identified 382
+putative duplicates, including duplicates in species with large body
+sizes and long life-spans. However, the probability of developing cancer
+is similar for small, short-lived mammals such as mice and for large,
+long-lived mammals such as elephants.
+
+In stark contrast, genome-wide studies of unusually large or long-lived
+species such as the bowhead whale (Keane et al., 2015), Myotid bats
+(Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al.,
+2011), and blind mole rat (Fang et al., 2014) did not find an over
+representation of tumor suppressors among duplicate genes.
+
+A genomic analysis of genetic changes associated with the evolution of
+enhanced cancer resistance in the elephant lineage has yet to be
+performed. Thus it is not clear if the duplication of TP53 and LIF
+reflects a general pattern of tumor suppressor duplication in the
+elephant lineage, unlike other lineages that resolved Peto?s paradox, or
+from the kinds of ascertainment biases common in candidate gene studies.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for
+his indispensible support.
+
+\hypertarget{conflicts-of-interest}{%
+\section{Conflicts of Interest}\label{conflicts-of-interest}}
+
+The Authors have no conflicts of interest to report
+
+\hypertarget{funding-source}{%
+\section{Funding Source}\label{funding-source}}
+
+We would like to thank the Department of Human Genetics at the
+University of Chicago for supporting this project.
+
+\hypertarget{supplementary-figures}{%
+\section{SUPPLEMENTARY FIGURES}\label{supplementary-figures}}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 1-1.pdf}
+\caption{Supplementary Figure 1: Estimated Copy Number by Coverage
+(ECNC) consolidates fragmented genes while accounting for missing
+domains in homologs. \textbf{A)} A single, contiguous gene homolog in a
+target genome with 100\% query length coverage has an ECNC of 1.0.
+\textbf{B)} Two contiguous gene homologs, each with 100\% query length
+coverage have an ECNC of 2.0. \textbf{C)} A single gene homolog, split
+across multiple scaffolds and contigs in a fragmented target genome;
+BLAT identifies each fragment as a single hit. Per nucleotide of query
+sequence, there is only one corresponding nucleotide over all the hits,
+thus the ECNC is 1.0. \textbf{D)} Two gene homologs, one fragmented and
+one contiguous. 100\% of nucleotides in the query sequence are
+represented between all hits; however, every nucleotide in the query has
+two matching nucleotides in the target genome, thus the ECNC is 2.0.
+\textbf{E)} One true gene homolog in the target genome, plus multiple
+hits of a conserved domain that span 20\% of the query sequence. While
+100\% of the query sequence is represented in total, 20\% of the
+nucleotides have 4 hits. Thus, the ECNC for this gene is 1.45.
+\textbf{F)} Two real gene homologs; one hit is contiguous, one hit is
+fragmented in two, and the tail end of both sequences was not identified
+by BLAT due to sequence divergence. Only 75\% of the query sequence was
+covered in total between the hits, but for that 75\%, each nucleotide
+has two hits. As such, ECNC is equal to 2.0 for this gene.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 2-1.pdf}
+\caption{Supplementary Figure 2: Gene copy increases polarized along
+Atlantogenata, colored by ln(Body Size), with branch lengths equal to
+the change in gene copy number}
+\end{figure}
+
+\begin{verbatim}
+## Warning: TODO: Make HQ version of this figure
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 1 rows containing missing values (geom_text).
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 3-1.pdf}
+\caption{Supplementary Figure 3: Full version of RecBlat strategy, but
+low quality}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 4-1.pdf}
+\caption{Supplementary Figure 4: Dot and Line plot for Body Sizes and
+such}
+\end{figure}
+
+\begin{verbatim}
+## $RBX1
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-1.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $LIN9
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-2.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $MND1
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-3.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $E2F2
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-4.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $MAX
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-5.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $SYCP3
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-6.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $CDK1
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-7.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $RNF168
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-8.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $RBBP8
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-9.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{verbatim}
+## 
+## $MAD2L1
+\end{verbatim}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-10.pdf}
+\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
+genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 7-1.pdf}
+\caption{Supplementary Figure 7: Correlation matrix heat map for genome
+quality metrics, ECNC, RBHB Copy, and Estimated Copy Number (lesser
+between ECNC and RBHB)}
+\end{figure}
+
+\begin{verbatim}
+## Warning: TODO: this causes a segfault, find out why!
+\end{verbatim}
+
+\hypertarget{supplementary-data}{%
+\section{Supplementary data}\label{supplementary-data}}
+
+\begin{verbatim}
+## Warning: TODO: Read in Eutheria.progress and present it to the bold.
+\end{verbatim}
 

### 2020-05-14 code/generalFunctions.R

@@ -656,7 +656,7 @@ graph.ORA.dotplot <- function(ORA.df, plotly=F){
       # text=GeneList, 
       color=Class
     )) + 
-    geom_point(position = position_jitter(width = .02, height=.02, seed = NULL)) +
+    geom_point() + #position = position_jitter(width = .02, height=.02, seed = NULL)) +
     # guides(col=guide_legend())+
     # facet_wrap(~Database) + 
     gghighlight(Class != "Other", keep_scales = T, calculate_per_facet = T, use_direct_label = F, label_key = NULL)+
@@ -664,8 +664,10 @@ graph.ORA.dotplot <- function(ORA.df, plotly=F){
     scale_x_log10(guide=guide_axis(title="FDR (log)")) +
     scale_y_log10(guide=guide_axis(title="Enrichment Ratio (log)")) +
     #geom_text_repel(data=. %>% filter(FDR<=1e-4|Class %in% c("TP53 Pathway", "Senescence", "Cell Death Pathways")), nudge_y = 0.5, show.legend = F, force = 15, size=3) +
-    geom_text_repel(data=. %>% filter(FDR<=1e-4), nudge_y = 0.5, show.legend = F, force = 15, size=3) +
-    geom_text_repel(data=. %>% filter(EnrichRatio>=7), nudge_y = -0.2, force = 15, show.legend=F, size=3) + 
+    geom_text_repel(data=. %>% filter(FDR<=1e-4), nudge_y = 0.5,
+                    show.legend = F, force = 15, size=3)+#, position = position_jitter(width = .02, height=.02, seed = NULL)) +
+    geom_text_repel(data=. %>% filter(EnrichRatio>=7), nudge_y = -0.2, 
+                    force = 15, show.legend=F, size=3)+#, position = position_jitter(width = .02, height=.02, seed = NULL)) +
     # geom_text_repel(data=. %>% filter(Class %in% c("TP53 Pathway", "Senescence", "Cell Death Pathways")), nudge_y = 0.5, force = 10, show.legend=F) + 
     scale_color_d3(palette="category20") +
     theme_pubclean() + 
@@ -674,3 +676,64 @@ graph.ORA.dotplot <- function(ORA.df, plotly=F){
   if(plotly) {p %<>% ggplotly}
   return(p)
 }
+
+
+get_upper_tri <- function(cormat){
[…]

### 2020-05-20 paper_PLOS/paper_PLOS_draft.Rmd
first draft

@@
 ---
-title: Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Paenungulates
+title: Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans
 author:
@@
 abstract: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+  Cancer is an intrinsic disease of multicellular organisms.
+Within a species, the size of an animal, - correlated with the individual's number of cells - and its lifespan - correlated with increasing cellular damage over time - positively correlate with the risk any individual has to form tumors.
+Between species, however, we do not observe any correlation between size, lifespan, and cancer, a phenomenon that referred to as _Peto's Paradox_.
+Elephants are a particularly interesting member of this class of paradoxical animals, since they are a set of large species deeply nested in a clade of smaller species, indicating a recent gain of size.
+Recent work has identified several individual cases of gene duplicates contributing to the increased cancer resistance of elephants, which suggests that duplication of tumor suppressor genes may play a more general role in mediating Peto's Paradox by increasing cancer resistance in large, long-lived species.
+  By using a Reciprocal Best-Hit BLAT search approach, we investigated copy numbers of all protein-coding genes in _Atlantogenatan_ genomes to see if there is any correlation between the copy number of duplicates and changes body size along the phylogenetic tree.
+From an initial set of 18,011 protein-coding genes in hg38, we identified a median of 13,880 genes in _Atlantogenatan_ genomes, of which a median of 940 genes are duplicated.
+We find that, just as body size fluctuates throughout _Atlantogenata_, genes involved in tumor suppressor pathways are also duplicated throughout the phylogenetic tree.
+Extant species of elephants, however, show active transcription of both canonical and duplicated copies of tummor suppressors that duplicated prior to and during their sudden increase in body size, suggesting that the duplication of tumor suppressor genes facilitates the evolution of increased body size by compensating for the increased cancer risk.
 
@@
 output: 
+  workflowr::wflow_html: default
   rticles::plos_article: 
@@
       in_header: preamble.tex
-  workflowr::wflow_html: default
   word_document: default
@@
 knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=5, 
-                      fig.retina = T)
+                      fig.retina = T, fig.align = "left")
 
@@
 
-anage <- read_tsv("../data/AnAge/anage_build14.txt", col_types = list("Weaning (days)"=col_number(), "Weaning weight (g)"=col_number(), "References" = col_character())) %>% 
-  select(Genus, Species,"Maximum longevity (yrs)", "Adult weight (g)") %>% 
-  unite("label", Genus, Species, sep=".") %>% 
+anage_full <- read_tsv("../data/AnAge/anage_build14.txt", col_types = list("Weaning (days)"=col_number(), "Weaning weight (g)"=col_number(), "References" = col_character())) %>% unite("label", Genus, Species, sep=".")
+anage <- anage_full %>% 
+  select(label,Family, "Maximum longevity (yrs)", "Adult weight (g)") %>% 
   mutate(lnSize.anage = sapply(
@@
   ) %>% 
-  select(label, lnSize.anage, Lifespan)
+  select(label, Family, lnSize.anage, Lifespan)
 ```
@@
   MRCA(bsize.tree, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
-  MRCA(bsize.tree, "Loxodonta.africana", "Mammut.americanum"), "Proboscidae",
+  MRCA(bsize.tree, "Loxodonta.africana", "Mammut.americanum"), "Proboscidea",
   MRCA(bsize.tree, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
@@
   MRCA(bsize.tree, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
-  MRCA(bsize.tree, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosorcida",
+  MRCA(bsize.tree, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosoricida",
   MRCA(bsize.tree, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
@@
   "Mammuthus" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
-  "Proboscidae" = bsize.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(node) %>% unique,
+  "Proboscidea" = bsize.Atlantogenata.table %>% filter(label=="Proboscidea") %>% pull(node) %>% unique,
   "Tethytheria" = bsize.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
@@
   "Afrotheria" = bsize.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
-  "Afrosorcida" = bsize.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(node) %>% unique,
+  "Afrosoricida" = bsize.Atlantogenata.table %>% filter(label=="Afrosoricida") %>% pull(node) %>% unique,
   "Macroscelidae" = bsize.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
@@
 ```
-
 ```{r Body Size Data: All, include=F}
@@
   MRCA(st.full, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
-  MRCA(st.full, "Loxodonta.africana", "Mammut.americanum"), "Proboscidae",
+  MRCA(st.full, "Loxodonta.africana", "Mammut.americanum"), "Proboscidea",
   MRCA(st.full, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
@@
   MRCA(st.full, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
-  MRCA(st.full, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosorcida",
+  MRCA(st.full, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosoricida",
   MRCA(st.full, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
@@
   "Mammuthus" = tree.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
-  "Proboscidae" = tree.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(node) %>% unique,
+  "Proboscidea" = tree.Atlantogenata.table %>% filter(label=="Proboscidea") %>% pull(node) %>% unique,
   "Tethytheria" = tree.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
@@
   "Afrotheria" = tree.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
-  "Afrosorcida" = tree.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(node) %>% unique,
+  "Afrosoricida" = tree.Atlantogenata.table %>% filter(label=="Afrosoricida") %>% pull(node) %>% unique,
   "Macroscelidae" = tree.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
@@
 
-tree.Atlantogenata.table %<>% mutate(highlight = sapply(seq_along(label), function(x, l=species.highlight, Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)}))
+tree.Atlantogenata.table %<>% mutate(
+  highlight = sapply(seq_along(label), function(x, l=species.highlight, Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)}),
+  size.change = node %>% sapply(., branch.size.ratio, tree.data=tree.Atlantogenata.table)
+  )
 
@@
   select(
-    `Ancestor/Species`=label, 
-    `Estimated Body Size (log(g))`=Median, 
+    `Node`=label, 
+    `Size (log(g))`=Median, 
     `95% CI (Low)` = `95%CI_Low`, 
@@
   ) %>%
-  mutate_at(vars(-`Ancestor/Species`, ), . %>% as.numeric %>% round(2)) %>% 
+  mutate_at(vars(-`Node`, ), . %>% as.numeric %>% round(2)) %>% 
   knitr::kable(
@@
   filter(label %in% st.core) %>% 
-  mutate(label = label %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidae", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosorcida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata"))) %>%
+  mutate(label = label %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata"))) %>%
   select(-parent, -node, -branch.length) %>%
@@
       xmin=`95%CI_Low`,
-      xmax=`95%CI_High`,
-      )
-  ) +
-  geom_errorbarh() +
-  geom_point(color="red") +
+      xmax=`95%CI_High`
+      )
+  ) +
+  geom_pointrange(size=1) +
+  geom_point(aes(color=Median), size=3) +
+  geom_point(aes(x=`95%CI_Low`, color=`95%CI_Low`), size=5, stroke=5, pch="[") +
+  geom_point(aes(x=`95%CI_High`, color=`95%CI_High`), size=5, stroke=5, pch="]") +
+  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))")) +
   labs(
@@
   ) + 
-  theme_pubclean()
+  theme_pubclean()+
+  labs_pubr()
 
@@
   theme(legend.position="right") + 
-  xlab("Rate of Change, Body Size")
+  xlab("Rate of Change, Body Size") + 
+  theme(
+    legend.position="right",
+    # panel.background = element_rect(fill = "lightgrey"),
+    text = element_text(face = "plain", 
+                        colour = "black", lineheight = 0.9, 
+                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                        debug = FALSE), 
+    axis.text.x = element_text(size = rel(0.86), 
+                               colour = "black", face = "bold"),
+    axis.title = element_text(size = rel(1), 
+                              colour = "black", face = "bold"),
+    plot.title = element_text(size = rel(1), 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.title = element_text(size = rel(0.7), 
+                                face = "bold", colour = "black"), 
+    legend.text = element_text(size = rel(0.7), 
+                               face = "plain", colour = "black")
+  )
 
@@
     ) + 
-  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))")) #+
-  #theme(legend.position="bottom") + 
-  # geom_tiplab2(aes(label=Spc.short), hjust = -.1, col="black", size=2)#+ theme(plot.margin=grid::unit(c(-20,-20,-20,-20), "mm"))
-  #geom_tiplab(aes(label=Spc.short), hjust = -.1, col="black", size=2)
-
-warning("TODO: de-comment geom_tiplab in p.tree.atlantogenata! Commented currently for offline usage!")
+  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))"))
+
+# warning("TODO: de-comment geom_tiplab in p.tree.atlantogenata! Commented currently for offline usage!")
 p.tree.atlantogenata <-  
@@
   ) + 
-  # geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic", 
-  #             show_guide=F, align = F, linesize = 0, offset= 100
-  #             ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidae, label="Proboscidae", 
+  geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic",
+              show_guide=F, align = F, linesize = 0, offset= 100
+              ) +
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidea, label="Proboscidea", 
                   fontsize = 4, align = T, offset = 50, angle=0, barsize = 1, 
                   offset.text = 20,
-                  color = bsize.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(lnSize) %>% get.size.color
+                  color = bsize.Atlantogenata.table %>% filter(label=="Proboscidea") %>% pull(lnSize) %>% get.size.color
                   ) +
@@
                   ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosorcida, label="Afrosorcida",
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosoricida, label="Afrosoricida",
                   fontsize = 4, align = T, angle=0, barsize = 1,
-                  color=bsize.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(lnSize) %>% get.size.color
+                  color=bsize.Atlantogenata.table %>% filter(label=="Afrosoricida") %>% pull(lnSize) %>% get.size.color
                   ) +
@@
   theme_tree2() +
+  xlab("Rate of Change, (Body Size, sqrt)") + 
   theme(
     legend.position="right",
-    panel.background = element_rect(fill = "lightgrey")
-  ) + 
-  xlab("Rate of Change, Body Size")
+    # panel.background = element_rect(fill = "lightgrey"),
+    text = element_text(face = "plain", 
+                        colour = "black", lineheight = 0.9, 
+                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                        debug = FALSE), 
+    axis.text.x = element_text(size = rel(0.86), 
+                               colour = "black", face = "bold"),
+    axis.title = element_text(size = rel(1), 
+                              colour = "black", face = "bold"),
+    plot.title = element_text(size = rel(1), 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.title = element_text(size = rel(0.7), 
+                                face = "bold", colour = "black"), 
+    legend.text = element_text(size = rel(0.7), 
+                               face = "plain", colour = "black")
+  )
 
@@
 p.ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts.withSize %>%
+  as_tibble %>% mutate(Spc.short = Spc.short %>% str_c('bold("', ., '")')) %>% 
+  as.treedata() %>% 
   ggtree(
@@
   scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold"))) +
-  geom_tiplab(aes(label=Species.strict %>% str_replace(" ","\n")), offset = 0.4, color="black", size=3)+
+  geom_tiplab(aes(label=Spc.short), offset = 0.4, color="black", size=4, parse=T)+
   geom_tiplab(aes(label=not.me(lnSize, N.Genes)), geom="label",  offset=-0.5, color="black", size=3)+ 
@@
 
-_Reciprocal Best-Hit BLAT:_ We developed a reciprocal best hit BLAT (RBHB) pipeline to quickly identify homologs and estimate gene copy numbers (__Figure 1A__).
+__Reciprocal Best-Hit BLAT:__ We developed a reciprocal best hit BLAT (RBHB) pipeline to quickly identify homologs and estimate gene copy numbers (__Figure 1A__).
 The Reciprocal Best Hit (RBH) search strategy is conceptually straightforward: 1) Given a gene of interest $G_A$ in a query genome $A$, one searches a target genome $B$ for all possible matches to $G_A$; 2) For each of these hits, one then performs the reciprocal search in the original query genome to identify the highest-scoring hit; 3) A hit in genome $B$ is defined as a homolog of gene $G_A$ if and only if the original gene $G_A$ is the top reciprocal search hit in genome $A$.
@@
 
-_Effective Copy Number By Coverage:_ In lower-quality genomes, many genes are fragmented across multiple scaffolds, which results in BLAT calling multiple hits when in reality there is only one gene.
+__Effective Copy Number By Coverage:__ In lower-quality genomes, many genes are fragmented across multiple scaffolds, which results in BLAT calling multiple hits when in reality there is only one gene.
 To compensate for this, we came up with a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we see each nucleotides of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (Supplementary Figure 1).
@@
 where $n$ is a given nucleotide in the query, $l$ is the total length of the query, $C_n$ is the number of instances that $n$ is present within a reciprocal best hit, and $bool(C_n)$ is 1 if $C_n > 0$ or 0 if $C_n = 0$.
-
-_RecSearch Pipeline:_ We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome.
+```{r RBHB Stats}
+n.TranscriptsSearched <- query.binTable %>% filter(!C.ORF, !LOC, !LINC, !HLA, !Histone, !ZincFinger,!Riboprotein, !OlfactoryReceptor, !Uncharacterized, !Putative, !Viral, !Fragment) %>% pull(ID) %>% unique %>% length()
+n.GenesSearched <- query.binTable %>% filter(!C.ORF, !LOC, !LINC, !HLA, !Histone, !ZincFinger,!Riboprotein, !OlfactoryReceptor, !Uncharacterized, !Putative, !Viral, !Fragment) %>% pull(Gene) %>% unique %>% length()
+genesPerGenome <- geneCopyTable %>% filter(Copy>=1, !is.na(ECNC)) %>% group_by(Genome) %>% summarise(
+  Hits=n(), pc.Found=Hits/n.GenesSearched, 
+  mean.ECNC = mean(ECNC), q1.ECNC = quantile(ECNC, 1/4), median.ECNC = median(ECNC), q3.ECNC = quantile(ECNC, 3/4), max.ECNC = max(ECNC), sd.ECNC = sd(ECNC), 
+  mean.Copy = mean(Copy), q1.Copy = quantile(Copy, 1/4), median.Copy = median(Copy), q3.Copy = quantile(Copy, 3/4), max.Copy = max(Copy), sd.Copy = sd(Copy), 
+  mean.ECNCPerHit = mean(ECNC/Copy), q1.ECNCPerHit = quantile(ECNC/Copy, 1/4), median.ECNCPerHit = median(ECNC/Copy), q3.ECNCPerHit = quantile(ECNC/Copy, 3/4), max.ECNCPerHit = max(ECNC/Copy), min.ECNCPerHit = min(ECNC/Copy), sd.ECNCPerHit = sd(ECNC/Copy))
+duplicatedGenesPerGenome <- geneCopyTable %>% filter(Copy>=1, ECNC>=1.5) %>% group_by(Genome) %>% summarise(Duplicated=n())
+
+genomeStatsRBHB <- full_join(genesPerGenome, duplicatedGenesPerGenome, by="Genome") %>% mutate(pc.Hits.Duplicated = Duplicated %>% divide_by(Hits))
+
+median.genesPerGenome <- genesPerGenome %>% pull(Hits) %>% median
+median.DuplicatedGenesPerGenome <- duplicatedGenesPerGenome %>% pull(Duplicated) %>% median
+```
+__RecSearch Pipeline__ We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome.
 For the query sequences in our search, we used the hg38 Proteome provided by UniProt [Accession #UP000005640; @uniprot], which is a comprehensive set of protein sequences curated from a combination of predicted and validated protein sequences generated by the UniProt Consortium.
@@
 LINC genes); poorly-studied genes from predicted open reading frames (C-ORFs); and sequences with highly repetitive sequences such as zinc fingers, protocadherins, and transposon-containing genes, as these were prone to high levels of false positive hits.
-After filtering out problematic protein queries, we then used our pipeline (Figure 1A) to search for all copies of our `r length(UP000005640.genenames)` query genes in publicly available Afrotherian genomes, including African savannah elephant (_Loxodonta africana_: loxAfr3, loxAfr4, loxAfrC), African forest elephant (_Loxodonta cyclotis_: loxCycF), Asian Elephant (_Elephas maximus_: eleMaxD), Woolly Mammoth (_Mammuthus primigenius_: mamPriV), Colombian mammoth (_Mammuthus columbi_: mamColU), American mastodon (_Mammut americanum_: mamAmeI), Rock Hyrax (_Procavia capensis_: proCap1, proCap2, proCap2_HiC), West Indian Manatee (_Trichechus manatus latirostris_: triManLat1, triManLat1_HiC), Aardvark (_Orycteropus afer_: oryAfe1, oryAfe1_HiC), Lesser Hedgehog Tenrec (_Echinops telfairi_: echTel2), Nine-banded armadillo (_Dasypus novemcinctus_: dasNov3), Hoffman's two-toed sloth (_Choloepus hoffmannii_: choHof1, choHof2, choHof2_HiC), Cape golden mole (_Chrysochloris asiatica_: chrAsi1), and Cape elephant shrew (_Elephantulus edwardii_: eleEdw1).
+  After filtering out problematic protein queries, we then used our pipeline (Figure 1A) to search for all copies of our `r n.GenesSearched` query genes in publicly available Afrotherian genomes, including African savannah elephant (_Loxodonta africana_: loxAfr3, loxAfr4, loxAfrC), African forest elephant (_Loxodonta cyclotis_: loxCycF), Asian Elephant (_Elephas maximus_: eleMaxD), Woolly Mammoth (_Mammuthus primigenius_: mamPriV), Colombian mammoth (_Mammuthus columbi_: mamColU), American mastodon (_Mammut americanum_: mamAmeI), Rock Hyrax (_Procavia capensis_: proCap1, proCap2, proCap2_HiC), West Indian Manatee (_Trichechus manatus latirostris_: triManLat1, triManLat1_HiC), Aardvark (_Orycteropus afer_: oryAfe1, oryAfe1_HiC), Lesser Hedgehog Tenrec (_Echinops telfairi_: echTel2), Nine-banded armadillo (_Dasypus novemcinctus_: dasNov3), Hoffman's two-toed sloth (_Choloepus hoffmannii_: choHof1, choHof2, choHof2_HiC), Cape golden mole (_Chrysochloris asiatica_: chrAsi1), and Cape elephant shrew (_Elephantulus edwardii_: eleEdw1).
 For many of these species, we covered multiple assemblies in order to test the effects of assembly size and quality on our hits.
 
-_Duplication gene inclusion criteria:_ In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call.
+__Query protein criteria__ To assemble our query list, we first removed all unnamed genes from UP000005640.
+Next, we excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (`r query.binTable %>% filter(C.ORF) %>% nrow`), LOC (`r query.binTable %>% filter(LOC) %>% nrow`), HLA genes (`r query.binTable %>% filter(HLA) %>% nrow`), replication dependent histones (`r query.binTable %>% filter(Histone) %>% nrow`), odorant receptors (`r query.binTable %>% filter(OlfactoryReceptor) %>% nrow`), ribosomal proteins (`r query.binTable %>% filter(Riboprotein) %>% nrow`), zinc finger transcription factors (`r query.binTable %>% filter(ZincFinger) %>% nrow`), viral and repetitive-element-associated proteins (`r query.binTable %>% filter(Viral) %>% nrow`) and any protein described as either "Uncharacterized," "Putative," or "Fragment" by UniProt in UP000005640 (`r query.binTable %>% filter(Fragment | Uncharacterized | Putative) %>% nrow`), leaving us with a final set of `r n.TranscriptsSearched` query protein sequences, corresponding to `r n.GenesSearched` genes.
+
+__Duplication gene inclusion criteria__ In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call.
 The resulting gene-level copy number table was then combined with the maximum ECNC values observed for each gene in order to call gene duplications.
@@
 The ECNC cut off of 1.5 was selected empirically, as this value minimized the number of false positives seen in a test set of genes and genomes.
-The results of our initial search are summarized in Figure 1B.
-Overall, we identified [MEDIAN] genes across all species, or [%HITS/QUERIES] of our starting query genes.
-
-_Duplicate gene exclusion criteria:_ We excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (`r query.binTable %>% filter(C.ORF) %>% length`), LOC (`r query.binTable %>% filter(LOC) %>% length`), HLA genes (`r query.binTable %>% filter(HLA) %>% length`), replication dependent histones (`r query.binTable %>% filter(Histone) %>% length`), odorant receptors (`r query.binTable %>% filter(OlfactoryReceptor) %>% length`), ribosomal proteins (`r query.binTable %>% filter(Riboprotein) %>% length`), zinc finger transcription factors (`r query.binTable %>% filter(ZincFinger) %>% length`), viral and repetitive-element-associated proteins (`r query.binTable %>% filter(Viral) %>% length`) and any protein described as either "Uncharacterized," "Putative," or "Fragment" by UniProt in UP000005640 (`r query.binTable %>% filter(Fragment | Uncharacterized | Putative) %>% length`).
-
-__Orthogonal Genome Assessment using CEGMA__ In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome.
-CEGMA was run using the default settings of [`r warning("TODO: add CEGMA settings")`], and the mammalian-specific core gene sets.
+The results of our initial search are summarized in Figure 1B`r warning('TODO: Check if this is still the case')`.
+Overall, we identified `r median.genesPerGenome` genes across all species, or `r median.genesPerGenome %>% divide_by(n.GenesSearched) %>% multiply_by(100) %>% round(digit=1) %>% paste0(.,"%")` of our starting query genes.
+
+__Genome Quality Assessment using CEGMA__ In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome [@gVolante; @CEGMA].
+CEGMA was run using the default settings for mammals ("Cut-off length for sequence statistics and composition" = 1;"CEGMA max intron length" = 100000; "CEGMA gene flanks" = 10000, "Selected reference gene set" = CVG).
+For each genome, we generated a correlation matrix using the aforementioned genome quality scores, and either the mean Copy Number or mean ECNC for all hits in the genome.
 
@@
 
-In order to determine the cancer risk at each node, we used a simplified multistage cancer risk model for body size and lifespan [@ArmitageAndDol; @Peto2015; @Peto2015b]. We defined the
+In order to determine the cancer risk $K$ at each node, we first needed to calcultate ancestral lifespans at each node.
+To do so, we used a Phylogenetic Generalized Least-Square Regression (PGLS) [@Felsenstein1985; @MartinsAndHansen1997] to calculate estimated ancestral lifespans across _Atlantogenata_ using our estimates for body size at each node.
+
+Next, we used a simplified multistage cancer risk model for body size $D$ and lifespan $t$: $K \approx Dt^6$ [@ArmitageAndDoll; @Peto2015; @Peto2015b].
+To determine the change in $K$ between nodes, we obtained the ratio between the cancer risk $K_{1}$ at any given node, and the cancer risk $K_{2}$ at its ancestral node, using the equation $\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}$.
+Finally, to simplify comparisons, we calculated the fold change cancer risk between a node and its ancestor as $\log_{2}(\frac{K_{2}}{K_{1}})$.
 
@@
 f1b <- p.tree.atlantogenata
-
-f1 <- ggarrange(
+f1c <- p.bodysize.atlantogenata
+# tree.Atlantogenata %>% tree_subset(., MRCA(.,"Loxodonta.africana", "Elephantulus.edwardii"), levels_back=3) %>% ggtree(aes(color=lnSize)) + geom_tiplab() +scale_color_viridis_c() +
+#   geom_point2(aes(subset=(node==MRCA(tree.Atlantogenata, "Loxodonta.africana", MRCA(tree.Atlantogenata, "Loxodonta.africana", "Mammut.americanum"))), color=lnSize))# Paenungulates, simple view
+#   # geom_point2(aes(subset=(node== MRCA(tree.Atlantogenata, "Loxodonta.africana", "Echinops.telfairi")), color=lnSize))# Paenungulates, simple view
+f1ab <- ggarrange(
   f1a, f1b, 
@@
   labels="AUTO", 
-  common.legend = T, legend = "bottom")
+  legend = "none",
+  widths = c(1,2)
+  )
+f1 <-
+  ggarrange(
+  ggplotify::as.ggplot(f1ab),
+  f1c,
+  nrow=2, ncol=1,
+  labels = c("", "C")
+)
 f1
@@
 Fossil data from extinct species were included to ensure that ancestral state reconstructions of body mass in Afrotherians were not biased by only including extant species, which can lead to inaccurate reconstructions, for example, if lineages multiple lineages evolved large body masses from a small bodied ancestor.
-We jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of a Brownian model of character evolution, which allows for occasional large jumps in traits (stable model) and out performs standard Brownian motion and Ornstein?Uhlenbeck models of character evolution [@ElliotAndMooers2014].
-
-Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a 65× increase in body mass in the stem-lineage of _Pseudoungulata_ (17kg), a 1.5× increase in body mass in the stem-lineage of _Paenungulata_ (25kg), a 12× increase in body mass in the stem-lineage of _Tehthytheria_ (296kg), and a 14× increase in body mass in the stem-lineage of _Proboscidea_ (4,100kg; Figure 1).
-The ancestral _Hyracoidea_ was inferred to be relatively small (2.86-15.71kg), and rate accelerations were coincident with independent body mass increases in large hyraxes such as _Titanohyrax andrewsi_ (67× increase in body mass).
-While the body mass of the ancestral _Sirenian_ was inferred to be large (61-656kg), a rate acceleration occurred coincident with a 10× body mass increase in Stellar's sea cow.
-Rate accelerations also occurred coincident with 36× body mass reduction in the stem-lineage of the dwarf elephants _Elephas_ (_Palaeoloxodon_) _falconeri_ and _Palaeoloxodon cypriotes_.
+We jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of a Brownian model of character evolution, which allows for occasional large jumps in traits (stable model) and out performs standard Brownian motion and Ornstein-Uhlenbeck models of character evolution [@ElliotAndMooers2014].
+
+Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Orycteropus.afer")` in body mass in the stem-lineage of _Pseudoungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Orycteropus.afer")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Procavia.capensis")` in body mass in the stem-lineage of _Paenungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Procavia.capensis")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Trichechus.manatus")` in body mass in the stem-lineage of _Tehthytheria_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Trichechus.manatus")`), and a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Mammut.americanum")` __(Vinnie: "14× increase", note I tried to divide the MRCA of LoxAfer with Phosphatherium, Moeritherium, Omanitherium, Phiomia, and Mammut by their respective ancestral nodes, but never got anything that was a 14x increase...)__ in body mass in the stem-lineage of _Proboscidea_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Mammut.americanum")`) __(Vinnie:"(4,100kg)"; I get 4114kg; can't find an MRCA with that value)__ (Figure 1B,C).
+The ancestral _Hyracoidea_ was inferred to be relatively small `r text.bodysize.ancNode.byMRCA.CI("Microhyrax.lavocati", "Procavia.capensis")` __(Vinnie wrote (2.86-15.71kg); I get an upper range of 118.18kg)__, and rate accelerations were coincident with independent body mass increases in large hyraxes such as _Titanohyrax andrewsi_ `r text.bodysize.byLabel("Titanohyrax.andrewsi")`; `r text.ratio.nodeByMRCA.ancNode(nodeA="Titanohyrax.andrewsi","Orycteropus.afer")`.
+While the body mass of the ancestral _Sirenian_ was inferred to be large `r text.bodysize.ancNode.byMRCA.CI("Trichechus.manatus", "Prorastomus.sirenoides")` _(Vinnie: (61-656kg); I get an upper range of 956kg)__, a rate acceleration occurred coincident with a `r text.ratio.node.ancnode.atLabel("Hydrodamalis.gigas")` in body mass in Stellar's sea cow.
+Rate accelerations also occurred coincident with `r text.ratio.nodeByMRCA.ancNode("Elephas.antiquus falconeri", "Elephas.cypriotes")` in body mass in the stem-lineage of the dwarf elephants _Elephas (Palaeoloxodon) antiquus falconeri_ and _Elephas cypriotes_.
 These data suggest that gigantism in _Afrotherians_ evolved step-wise, from small to medium bodies in the _Pseudoungulata_ stem-lineage, medium to large bodies in the _Tehthytherian_ stem-lineage and extinct hyraxes, and from large to exceptionally large bodies independently in the _Proboscidean_ stem-lineage and Stellar's sea cow (Figure 1).
@@
 
-geneCopyTable.summary <-
-  geneCopyTable %>%
-  mutate(
-    ECNC = ECNC %>% sapply(
-      .,
-      . %>% if_else(is.na(.), 0, .) %>%
-        sum(., ifelse(.>round(.), 0.01, 0)) %>%
-        round()
-    ),
-    Copy = Copy %>% sapply(. %>% if_else(is.na(.), 0, .)),
-    CN = if_else(ECNC<Copy, ECNC, Copy)
-  ) %>%
-  group_by(Genome) %>%
-  summarise(
-    # median.ECNC = median(ECNC),
-    mean.ECNC = mean(ECNC),
-    sd.ECNC = sd(ECNC),
-    # median.Copy = median(Copy),
-    mean.Copy = mean(Copy),
-    sd.Copy = sd(Copy),
-    # median.CN = median(CN),
-    mean.CN = mean(CN),
-    sd.CN = sd(CN),
-  )
-
-GCT.CEGMA <- geneCopyTable.summary %>%
+GCT.CEGMA <- genomeStatsRBHB %>%
+  select(-starts_with("q"), -starts_with("Median")) %>% 
   inner_join(., CEGMA, by="Genome") %>%
@@
 
-p.cor.GCT.CEGMA <- cor.GCT.CEGMA %>%
-  # filter(value>0) %>%
+p.cor.GCT.CEGMA <-
+  cor.GCT.CEGMA %>%
   ggplot(
-    # aes(x=Var2, y=Var1, fill = value)
     aes(x=Var1, y=Var2, fill = value)
@@
   geom_tile(color = "white")+
-  # scale_fill_gradient2(low = "blue", high = "red", mid = "white",
-  #   midpoint = 0, limit = c(-1,1), space = "Lab",
-  #    name="Pearson\nCorrelation") +
-  scale_fill_viridis_c() +
   theme_minimal()+
   coord_fixed()+
-  # coord_flip() +
-  # scale_x_discrete(position = "top") +
-  scale_x_discrete(#limits = rev(levels(cor.GCT.CEGMA$Var1)), 
-    position = "top") +
+  scale_x_discrete(
+    position = "top"
+    ) +
   scale_y_discrete(position = "right") +
-  # scale_y_discrete(limits = rev(levels(cor.GCT.CEGMA$Var1)), position = "right") +
+  scale_fill_viridis_c(breaks = c(-1,-0.5, 0, 0.5, 1),labels = c(-1,-0.5, 0, 0.5, 1)) +
   theme(
-    # axis.text.x = element_text(
-    #   angle = -45, vjust = 0,
-    #   size = 9, hjust = 0
-    #   ),
     axis.text.x = element_text(
@@
     legend.position = c(0.4,0.7), #c(0.6, 0.7),
-    legend.direction = "horizontal")+
-  guides(fill = guide_colorbar(title="Pearson\nCorrelation",
-                               barwidth = 7, barheight = 1,
-                               title.position = "top", title.hjust = 0.5))
+    legend.direction = "horizontal",
+    legend.text = element_text(size=rel(1)),
+    legend.title =  element_text(size=rel(1.2))
+    ) +
+  guides(
+    fill = guide_colorbar(
+      title="Pearson\nCorrelation",
+      barwidth = 9, barheight = 3,
+      title.position = "top", title.hjust = 0.5,
+      draw.llim = T
+    )
+  )
 ```
@@
 
-```{r Figure 2C1, fig.cap="Figure 2 (Candidate 1): A Reciprocal Best-Hit BLAT pipeline for identifying gene copy number in other genomes.
-A) A graphic summary of the reciprocal best-hit strategy.
-B) Estimated Copy Number by Coverage.
-C) "}
-f2a <- image_read("../output/figs/RBHBDiagram.png") %>% image_ggplot()
-f2b <- ecnc.twogene.fourhits
-f2c <- p.ml.tree.increasing.geneCounts.withSize
-f2ab <- ggarrange(f2a, f2b, nrow=2, ncol=1, labels = LETTERS[1:2])
-f2.1 <- ggarrange(f2ab, f2c, nrow=1, ncol=2, widths = c(1, 2), labels=c(NA, "C"))
-f2.1
-```
-```{r Figure 2C2, fig.cap="Figure 2 (Candidate 2): Gene Duplicates in _Atlantogenata_. A) A graphic summary of the reciprocal best-hit strategy. B) Estimated Copy Number by Coverage. C)"}
-f2a <- p.ml.tree.increasing.geneCounts.withSize
-f2b <- p.bodysize.atlantogenata
+```{r Genome Table Prep, warning=T}
+warning("TODO: Label each genome with citation number")
+genome.table.forFig <- genomeTable.atlantogenata %>% 
+  left_join(anage, by="label") %>% 
+  left_join(tree.Atlantogenata.table, by="label") %>% 
+  mutate(
+    Median = exp(Median) %>% formatC(format = "g", digits = 2),
+    Lifespan = Lifespan %>% exp %>% formatC(format = "g", digits = 2) %>% as.character() %>% str_replace_na(replacement = "(Unknown)")
+  ) %>% 
+  select(
+    `Common Name` = `Common Name.x`,
+    Species = Species.strict.x,
+    # Clade = Family, 
+    `Size (g)` = Median,
+    # `Lifespan (y)`= Lifespan, 
+    Genome = BestGenome.x
+  ) %>% 
+  left_join(
+    genomeStatsRBHB %>% 
+      select(Genome, `#Hits`=Hits, `% Genes Found`=pc.Found, `#Duplicated` = Duplicated,`% Hits Duplicated` = pc.Hits.Duplicated, `Max Copy#`=max.Copy, `Max ECNC`=max.ECNC, `Mean ECNC/Hit`=mean.ECNCPerHit),
+    by="Genome"
+  ) %>% 
+  mutate(
+    `Common Name` = str_to_title(`Common Name`),
+    `% Genes Found` = `% Genes Found` %>% as.percentage,
+    `% Hits Duplicated` = `% Hits Duplicated` %>% as.percentage
+  )
+
+# #clades
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Elephant$|Mammoth$"),]$Clade <- "Paenungulata, Probodiscea"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Mastodon"),]$Clade <- "Paenungulata, Probodiscea"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "^Cape|Tenrec"),]$Clade <- "Macroscelididae/Afroinsectivora"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Manatee"),]$Clade <- "Paenungulata, Tethytheria"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Hyrax"),]$Clade <- "Paenungulata"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Aardvark"),]$Clade <- "Pseudoungulata"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Sloth|Armadillo"),]$Clade <- "Xenarthra"
+# 
+# #Lifespan caveats
+# genome.table.forFig[which(genome.table.forFig$Species == "Elephantulus edwardii"),]$`Lifespan (y)` = "~5.7*"
+# genome.table.forFig[which(genome.table.forFig$Species == "Loxodonta cyclotis"),]$`Lifespan (y)` = paste0(genome.table.forFig[which(genome.table.forFig$Species == "Loxodonta africana"),]$`Lifespan (y)`, "\u2020")
+```
+```{r Figure 2 Prep: Genome Table}
+ggt.GenomeTable.withNote <- genome.table.forFig %>% 
+  select(Species, `Common Name`, starts_with("Size"), starts_with("#"), starts_with("%"), starts_with("Mean")) %>% 
+  arrange(Species) %>% 
+  ggtexttable(
+    theme = ttheme(base_style="minimal", base_size=8)
+  ) #%>% annotate_figure(bottom = "* Observation derived from a single captive individual; real lifespan unknown.\n\u2020 Lifespan estimate derived from Loxodonta africana before the taxonomic species split.")
+```
+
+```{r Figure 2, fig.cap="Gene duplications occur readily throughout Atlantogenata", fig.height=7, fig.width=5}
+f2a <- ggt.GenomeTable.withNote
+f2b <- p.ml.tree.increasing.geneCounts.withSize
 f2.2 <- ggarrange(f2a, f2b, nrow=2, ncol=1, labels = "AUTO")
@@
 ```
-```{r PREP Table 2: PGLS of Copy Number ~ Genomic Stats, eval=F}
+```{r Prep Table 2: PGLS of Copy Number ~ Genomic Stats, eval=F}
 # if(knitr::is_latex_output()){
-t.GCT.CEGMA.gls <- quiet(stargazer(
-  models.mean.Copy,models.mean.ECNC, models.mean.CN,
-  type="latex",
-  title = "Gene Copy ~ various genomic variables",
-  style = "all"
-)  %>% 
-  c("\\newgeometry{margin=1cm}\n\\begin{landscape}", ., "\\end{landscape}\n\\restoregeometry") %>% 
-  paste0(collapse="\n")
-  )
+# t.GCT.CEGMA.gls <- quiet(stargazer(
+#   models.mean.Copy,models.mean.ECNC, models.mean.CN,
+#   type="latex",
+#   title = "Gene Copy ~ various genomic variables",
+#   style = "all"
+# )  %>% 
+#   c("\\newgeometry{margin=1cm}\n\\begin{landscape}", ., "\\end{landscape}\n\\restoregeometry") %>% 
+#   paste0(collapse="\n")
+#   )
 # } else if(knitr::is_html_output()){
@@
 ```
-```{r Table 2: ML model summary, warning=T}
-warning("TODO: Table 2 is done, but Stargazer has some weird bug stopping it from running...")
-```
 
@@
 Therefore, in order to test whether this was a pervasive phenomena genome-wide in _Afrotherians_, we used a Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in _Afrotherian_ and _Atlantogenatan_ genomes (Fig.
-2A).
+2A, Supplementary Figure 1).
 Because RBHB-like approaches can over-estimate copy number when genes are fragmented  or incorrectly assembled across multiple scaffolds, we also inferred copy number using a complementary method that quantifies the ratio between observed and expected gene coverage per nucleotide (ECNC) (Supplementary Figure 1).
@@
 
-To reconcile the Atlantogenatan phylogeny with duplication events, we used maximum likelihood to reconstruct likely ancestral copy numbers for each gene at each node in the phylogeny.
-To define the copy number of a gene, we conservatively used the lesser value between the RBHB hit count, and the ECNC value rounded to the nearest whole number.
-In order to perform
+Once we applied these criteria to the results, we obtained the results shown in Figure 2A.
+Our approach positively identified an average of `r genomeStatsRBHB$pc.found %>% mean %>% as.percentage` of the queries we searched.
+We observed that the percentage of duplicated genes in non-Pseudoungulatan genomes was significantly higher: while these genomes had duplicates percentages ranging anywhere from `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`, outgroup species' duplication rates ranged from `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`.
+
+```{r Figure 3: Correlation}
+p.cor.GCT.CEGMA
+```
+
+Given the preliminary status of the genomes for _Chrysochloris asiatica_ and _Elephantulus edwardii_; and the fact that more revisions have been performed for Paenungulatan genomes; we hypothesized that genome quality may be confounding the results we've observed.
+We used `CEGMA` and the `gVolante` server [@CEGMA; @gVolante] to assess the quality of all our genomes, to see if any of these metrics correlated with our metrics of gene counts and copy number.
+As shown in Figure 3, mean ECNC, mean Copy Number, and mean CN (the lesser of Copy Number and ECNC per gene) correlate moderately strongly with genomic quality metrics related to the length and assembly quality, like LD50 and the number of scaffolds and contigs with a length above either 100K or 1M, supporting our hypothesis that increases in copy number correlate .
+
+In order to determine when in the phylogeny our identified duplications occured, we used maximum likelihood to reconstruct ancestral gene copy numbers across the phylogeny.
+We encoded the copy number of each gene as a discrete trait ranging from 0 ("?") to 31+, and then ran the ancestral gene copy number reconstruction using the phylogenetics suite `iqtree`, which rapidly tests and runs the best evolutionary model for a given trait of interest.
+Importantly, `iqtree` can be set to only output states where the likelihood of the state is greater than a certain threshold.
+Setting the threshold to 0.8, we obtained the results shown in
 
@@
 
-<!--
-## Retroduplication of EEF1A1 in the stem-lineage of Tehthytheria
-
-```{r Fig3 prep}
-f3a <- p.ml.treedata.loxAfrHighlight.plus$EEF1A1
-# f3b <- 
-# f3c <- 
-f3 <- ggarrange(
-  f3a,
-  # f3b, f3c,
-  # nrows=1, ncol=3
-  labels="AUTO"
-)
-```
-
-```{r Fig3, fig.cap="Retroduplication of Eukaryotic translation Elongation Factor 1 Alpha 1 (EEF1A1).
-A) Gene Copy Number of EEF1A1 in _Atlantogenata_.
-B) Expression of EEF1A1 copies in the West Indian Manatee and African Elephant", warning=T}
-f3
-warning("Fig 3: EEF1A1")
-```
-
-Previous studies have identified genes whose transgenic overexpression delays aging related phenotypes and extends lifespan.
-Overexpression of the eukaryotic translation elongation factor 1 alpha 1 (EEF1A1) gene, for example, extends lifespan in transgenic Drosophila.
-Like most mammalian genomes, we found that Afrotherian genomes encoded numerous EEF1A1 retrogenes (Fig.
-3A).
-While the majority of these retrogenes have premature stop codons and insertions/deletions characteristic of pseudogenes, at least 13 Tehthytherian-specific EEF1A1 retrogenes have the potential to encode functional proteins (Fig.
-3B).
-Unfortunately it is difficult to determine if these retrogenes are transcribed because they have relatively high sequence similarity (83-88%), suggesting evidence of transcription using RNA-Seq data is likely an artifact.
-Indeed, nearly all of the EEF1A1 retrogene transcripts assembled by StringTie were confined to their gene body and UTRs and thus lacked unique sequences.
-The EEF1A1RTG13 transcript, however, initiates within an upstream MLT1B retrovirus-like MaLR transposable element and includes unique 5?- and 3?-UTR sequences, indicating at least one EEF1A1 retrogene is transcribed (Fig.
-3C).
--->
-
 ## Duplications that occured recently in Probodiscea are enriched for tumor suppressor pathways
@@
 p.loxAfr.ORA.dotplot <- ORA.all %>% 
-    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome") %>% 
+    filter(Node == "Loxodonta_africana", FDR.Threshold <= 0.3, Database == "Reactome", RBHB == "RBB") %>% 
     graph.ORA.dotplot(., plotly=knitr::is_html_output())
@@
 p.loxAfr.ORA.dotplot.increased <- ORA.all %>% 
-    filter(Node == "Loxodonta_africana", FDR.Threshold %in% c("0.1", "0.2", "0.3"), Database == "Reactome", Direction=="increased") %>% 
+    filter(Node == "Loxodonta_africana", FDR.Threshold <= 0.3, Database == "Reactome", RBHB == "RBB", Direction=="increased") %>% 
     graph.ORA.dotplot(., plotly=knitr::is_html_output())
-```
-
-```{r Fig 3, fig.pos="hold", fig.cap="Fig 3"}
-f3a <- p.loxAfr.ORA.dotplot
-f3b <- p.loxAfr.ORA.dotplot.increased
-
-f3 <- ggarrange(f3a, f3b, nrow=1, ncol=2, common.legend = T, legend = "bottom", label="AUTO")
-# f3
-f3a
-f3b
-```
-
-In order to infer the functional consequences of these gene duplications, we tested if duplicate genes were enriched in specific pathways relative to our initial query set of genes.
-We used `r warning("TODO: Finish this section")`
+# p.Proboscidea.ORA.dotplot <- ORA.all %>% 
+#   filter(Direction=="increased", Node == "Proboscidea", Database == "Reactome", RBHB == "RBB") %>% 
+#   graph.ORA.dotplot(., plotly=knitr::is_html_output())
+```
+
+```{r Branches which experienced an increase in size}
+
+ORA.numbers <- ORA.all  %>% filter(RBHB=="RBB") %>% group_by(Direction, Ancestor, Node, FDR.Threshold, Database) %>% summarise(N.Pathways=n_distinct(Pathway)) %>% ungroup %>% pivot_wider(names_from = FDR.Threshold, values_from = N.Pathways, names_prefix = "Pathways at FDR<") %>% mutate_if(is.numeric, .
+%>% ifelse(is.na(.), 0,.))
+
+ORA.numbers.ReactomeOnly <- ORA.all %>% filter(Database=="Reactome", RBHB=="RBB") %>% group_by(Direction, Ancestor, Node, FDR.Threshold) %>% summarise(N.Pathways=n_distinct(Pathway)) %>% ungroup %>% pivot_wider(names_from = FDR.Threshold, values_from = N.Pathways, names_prefix = "Pathways at FDR<") %>% mutate_if(is.numeric, .
+%>% ifelse(is.na(.), 0,.))
+
+noEnrichmentIncreasing <- setdiff(ORA.numbers %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers %>% filter(Direction == "increased") %>%  pull(Node)) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+noEnrichmentIncreasing.Reactome <- setdiff(ORA.numbers.ReactomeOnly %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers.ReactomeOnly %>% filter(Direction == "increased") %>%  pull(Node)) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+
+noEnrichment.any <- setdiff(ml.treedata.table$label %>% str_replace_all("\\.", "_"), union(ORA.numbers %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers %>% filter(Direction == "increased") %>%  pull(Node))) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+noEnrichment.any.Reactome <- setdiff(ml.treedata.table$label %>% str_replace_all("\\.", "_"), union(ORA.numbers.ReactomeOnly %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers.ReactomeOnly %>% filter(Direction == "increased") %>%  pull(Node))) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+
+p.ORA <- ORA.numbers.ReactomeOnly %>% 
+  filter(Direction == "increased") %>% 
+  select(-Direction) %>% 
+  ggtexttable(
+    theme = ttheme(base_style="minimal", base_size=8)
+  )
+
+```
+
+```{r Fig 3, fig.cap="Fig 3"}
+f3a <- p.ORA
+f3b <- p.loxAfr.ORA.dotplot
+f3c <- p.loxAfr.ORA.dotplot.increased
+
+f3bc <- ggarrange(f3b, f3c, nrow=1, ncol=2, common.legend = T, legend = "bottom", labels=c("B", "C"))
+f3 <- ggarrange(f3a, f3bc, nrow=2, ncol=1, labels=c("A",""))
+
+f3
+```
+`r warning("This figure needs work. Start by changing that table to a bar graph, for one....")`
+
+As previously stated, our inital hypothesis was that genes duplicated in lineages that experienced a growth in size would have an increased number of gene duplications that were enriched for membership in tumor suppressor pathways.
+Thus, we used WebGestalt and its Overrepresentation Analysis (ORA) functionality to determine what pathways were enriched in our duplicated gene sets in each branch relative to our initial query set.
+For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways_cancer databases using WebGestalt ORA.
+Going through the tree, at no FDR<0.5 is there any significant pathway representation for genes that increased in the branches leading to `r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`; furthermore, there are no significant pathway enrichments at FDR<0.5 for copy-number-stable genes for `r noEnrichment.any.Reactome[1:length(noEnrichment.any.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichment.any.Reactome[length(noEnrichment.any.Reactome)]`.
+Note that because _Xenarthra_ was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.
+
+For the other branches, the number of pathways that came up as significantly enriched at each FDR is shown in Figure 3A (Supplementary Figure 3).
+For the species with high duplication rates and lower-quality, highly-fragmented genomes, such as with _Chrysochloris asiatica_ (`r genomeStatsRBHB %>% filter(Genome == "chrAsi1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits) and _Elephantulus edwardii_ (`r genomeStatsRBHB %>% filter(Genome == "eleEdw1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits), it is unsurprising that there is an equivalently astronomical number of pathway enrichments present.
+In the case of these two species, their many pathway enrichments also span an incredible range of processes at every level of biology; this, in combination with the astronomically high number of copy numbers identified for these genes, further suggests a need for improvement and refinement in these genomes.
+In the cell cycle pathways called as significant in the genomes of these two species plus _Orycteropus afer_ and _Echinops telfairi_, we see that the member duplicated genes are from the same sets families, such as the APC subunit family; the proteosome subunit families; and the protein phosphatase 2 family, among others.
+It is highly possible that these results reflect true expansions of these gene families, especially in the higher-quality _Orycteropus afer_ genome; however, as discussed, it is also possible that it simply reflects artifcatual duplication events, and so require futher study.
+
+The pathway enrichments for the stable - yet duplicated - copy number, and increased copy number sets of genes for the African Elephant are shown in Figure 3B-C.
+In support of our original hypothesis, among the few enriched pathways in the African Elephant, we see that that two tumor suppression pathways - APC Complex-related pathways, and "TP53 Regulates Metabolic Genes"- appear not only in the case of stable genes, but also in the set of newly duplicated genes.
+The other pathways we see in the set of recently-duplicated genes include "Functionalization of Compounds" and its daughter pathway "Xenobiotics".
+Genes in these pathways serve to add functional groups to lipophylic compounds which would otherwise not be reactive in the cell, and are types of metabolic pathways.
+In the stable set we see enrichment of pathways such as "Neuronal Systems" and "Axon Guidance," which fit in well with what is known about elephant biology and evolution [@Cite probably].
+Overall in elephants, we see enrichments within duplicated genes for pathways involved in what makes an elephant an elephant - including tumor suppressor pathways.
 
@@
 ```{r Fig4 prep}
-
-    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata
-    # "STK11", #REgulation of TP53, Paenungulata
-    # "COX5A", #TP53 transcriptional regulation, PRobodiscea
-    # "BRD7", # TP53 regulation, Tethytheria
+loxAfr4.geneTPM.genesToHighlight.plus
+    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
+    # "STK11", #REgulation of TP53, Paenungulata                  #STK11 is a TSG
+    # "COX5A", #TP53 transcriptional regulation, PRobodiscea      #CytoC OXidase
+    # "BRD7", # TP53 regulation, Tethytheria                      #Required for p53-dependent oncogene-induced senescence
     # "TP53", #TP53 duplicates in Tethytheria
-    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria
-    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants
-    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri
-    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants
-    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr
+    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria   #E3 Ligase, induces apoptosis
+    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants  #Induces apoptosis, duh
+    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri          #Decomposes H2O2
+    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants   #Antiapoptotic?
+    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr            #CytoC OXidase
 
@@
   summarize(Gene = unique(Gene), id=unique(id), TPM=max(TPM), Reactome.Pathway.Prime=unique(Reactome.Pathway.Prime)) %>% 
-  arrange(Gene, id)
-
-# TP53.genes.plots <- genesToHighlight.plus.TP53 %>% 
-#   set_names(.,.) %>% 
-#   lapply(
-#     .,
-#     function(gene){
-#       list(
-#         "GeneCopyTree"=ml.treedata.loxAfrHighlight.plus[[gene]],
-#         "GeneCopyExpression"= 
-#       )
-#     }
-#   )
+  arrange(Gene, id) %>% 
+  mutate(Class = ifelse(
+    str_detect(Reactome.Pathway.Prime, "[Rr]egulation of [Tt][Pp]53")& Gene != "TP53",
+    "Regulator of TP53",
+    ifelse(
+      str_detect(Gene, "COX")& Gene != "TP53",
+      "Cytochrome C Oxidase Component",
+      ifelse(
+        str_detect(Gene, "CASP")& Gene != "TP53",
+      "Apoptosis",
+      "TP53 Pathway, Other"
+      )
+    )
+
+  )
+           )
+
+treedata.tp53.table <- ml.treedata.table %>% 
+  full_join(., bodysize.tree.atlantogenata, by="label") %>%
+  select(parent, node, branch.length, label, lnSize, all_of(loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% pull(Gene) %>% unique)) %>% 
+  mutate_at(vars(loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% pull(Gene) %>% unique), . %>% sapply(. %>% ifelse(is.na(.), 1, .))) %>% 
+  gene.increase.tree()
+
+TP53.gene.color <- RColorBrewer::brewer.pal(length(loxAfr4.geneTPM.ORA.expressedDups.TP53$Gene %>% unique), "Set1") %>% 
+  set_names(loxAfr4.geneTPM.ORA.expressedDups.TP53$Gene %>% unique)
+
+# TODO: Make the recursion make NA values to the parent node value, unlike the manual fix above.
+# TODO: Arrange the le
+p.treedata.tp53.table <- treedata.tp53.table %>% 
+  mutate(label=str_replace(label, "\\.", " ")) %>% 
+  as.treedata() %>% 
+  ggtree(
+    .,
+    branch.length = "none",
+    #continuous=T,
+    aes(
+    ),
+    size=2
+  ) +  
+  geom_point2(aes(x=branch, subset=(GTF2F1>0), color="GTF2F1"), size=5, position = position_nudge(x=-0.2, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(STK11>0), color="STK11"), size=5, position = position_nudge(x=0, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(COX5A>0), color="COX5A"), size=5, position = position_nudge(x=0.2, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(BRD7>0), color="BRD7"), size=5, position = position_nudge(x=-0.2, y=0)) +
+  geom_point2(aes(x=branch, subset=(TP53>0), color="TP53"), size=5, position = position_nudge(x=0, y=0)) +
+  geom_point2(aes(x=branch, subset=(CASP9>0), color="CASP9"), size=5, position = position_nudge(x=0.2, y=0)) +
+  geom_point2(aes(x=branch, subset=(PRDX1>0), color="PRDX1"), size=5, position = position_nudge(x=-0.2, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(LAMTOR5>0), color="LAMTOR5"), size=5, position = position_nudge(x=0, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(COX20>0), color="COX20"), size=5, position = position_nudge(x=0.2, y=0.2)) +
+  geom_tiplab() +
+  scale_color_manual(values = TP53.gene.color) + 
+  guides(color=guide_legend(title="Gene"))+
+  xlim(0,12) +
+  theme_tree2() + 
+  theme(legend.position = "bottom")
 
 p.tp53.genes.expression <- loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% 
-  # filter(Gene == gene) %>% 
   ggplot(
@@
       y=TPM, 
-      color=Reactome.Pathway.Prime
+      color=Class
     )
@@
   ) + 
-  theme(legend.position="bottom")
-warning("TODO: 1. Color each facet a different color")
-warning("TODO: 2. Make a tree with stars at each node where one of these genes duplicated, color coded by genes")
+  theme(legend.position="bottom")+
+  labs_pubr()
 ```
@@
 ```{r Fig4, warning=T}
-warning("TODO: Make Fig4")
-
-```
+f4a <- p.treedata.tp53.table
+f4b <- p.tp53.genes.expression
+fig4 <- ggarrange(f4a, f4b, nrow=1, ncol=2, labels="AUTO")
+
+fig4
+    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
+    # "STK11", #REgulation of TP53, Paenungulata                  #STK11 is a TSG
+    # "COX5A", #TP53 transcriptional regulation, PRobodiscea      #CytoC OXidase
+    # "BRD7", # TP53 regulation, Tethytheria                      #Required for p53-dependent oncogene-induced senescence
+    # "TP53", #TP53 duplicates in Tethytheria
+    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria   #E3 Ligase, induces apoptosis
+    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants  #Induces apoptosis, duh
+    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri          #Decomposes H2O2
+    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants   #Antiapoptotic?
+    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr            #CytoC OXidase
+```
+
+Prior studies looking at the duplication of TP53 in the African Elephant motivated further study for the enrichment of genes involved in TP53-related metabolic pathway.
+We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, and used publically-available RNA-seq data to see which genes are actively expressess in living elephants.
+Excitingly, we found that the initial duplication of TP53 in Tethytheria, where body size expanded, was presceded by the duplication of GTF2F1 and STK11 in _Paenungulata_; and was coincided by the duplication of BRD7.
+[TALK ABOUT HOW THESE TWO MAY HAVE FACILITATED THE DUPLICATION OF TP53]
+GTF2F1 is a general transcription factor [@CITE and talk more about this]
+STK11 is a tumor suppressor gene [@CITE and talk more]
+
+The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either alongside TP53 in the case of SIAH1, or subsequently in _Probodiscea_, _Elephantidae_, and in modern elephants.
+(Figure 4A).
+These genes are all expressed in RNA-seq data, suggesting that they encode functional genes in modern elephants (Fig 4B).
 
@@
   CancerSucceptabilityChange = K1/K2,
-  log2CancerSucceptabilityChange = log2(K1)-log2(K2)
+  log2CancerSucceptabilityChange = log2(K1)-log2(K2),
+  label = label %>% str_replace_all("\\.", " ")
 )
@@
   CancerSucceptabilityChange = K1/K2,
-  log2CancerSucceptabilityChange = log2(K1)-log2(K2)
+  log2CancerSucceptabilityChange = log2(K1)-log2(K2),
+  label = label %>% str_replace_all("\\.", " ")
 )
 tree.cancerSucceptibility.Atlantogenata <- table.cancerSucceptibility.Atlantogenata %>% 
-  full_join(., genomeTable.atlantogenata, by = "label") %>% 
+  full_join(., genomeTable.atlantogenata %>% mutate(label=str_replace_all(label, "\\.", " ")), by = "label") %>% 
   mutate(highlight = sapply(seq_along(label), function(x, l=str_replace(species.highlight, " ", ".") %>% set_names(.,.), Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)})) %>%
@@
 ```{r Fig 5 prep}
-f5a <- tree.cancerSucceptibility.Atlantogenata  %>% as_tibble %>% mutate(CancerSucceptabilityChange = as.numeric(CancerSucceptabilityChange)) %>% as.treedata %>% 
+f5a1 <- tree.cancerSucceptibility.Atlantogenata  %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(CancerSucceptabilityChange = as.numeric(CancerSucceptabilityChange)) %>% as.treedata %>% 
   ggtree(
@@
     size = 1.5, 
-    yscale="log2CancerSucceptability"
+    yscale="log2CancerSucceptabilityChange"
   ) + 
@@
 
-f5a2 <- tree.cancerSucceptibility.Atlantogenata  %>% 
+f5a <- tree.cancerSucceptibility.Atlantogenata %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% as.treedata %>% 
   ggtree(
@@
   #             ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidae, label="Proboscidae", 
-                  fontsize = 4, align = T, offset = 50, angle=0, barsize = 1, 
-                  offset.text = 20,
-                  color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Proboscidae") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidea, label="Proboscidea", 
+                  fontsize = 4, align = T, offset = 10, angle=0, barsize = 1, 
+                  offset.text = 0,
+                  color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Proboscidea") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                   ) +
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="Tethytheria", 
-                  fontsize = 4, align = T, offset = 25, angle=0, barsize = 1, 
-                  offset.text = 20, color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
+                  fontsize = 4, align = T, offset = 5, angle=0, barsize = 1, 
+                  offset.text = 5, color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                   ) +
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="Paenungulata", 
-                  fontsize = 4, align = T, angle=0, barsize = 1, offset.text = 20,
+                  fontsize = 4, align = T, angle=0, barsize = 1, offset.text = 10, 
                   color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Paenungulata") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                   ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosorcida, label="Afrosorcida",
-                  fontsize = 4, align = T, angle=0, barsize = 1,
-                  color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Afrosorcida") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosoricida, label="Afrosoricida",
+                  fontsize = 4, align = T, angle=0, barsize = 1,offset.text = 10,
+                  color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Afrosoricida") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                   ) +
@@
                     label="Macroscelidae", 
-                    fontsize = 4, align = T, angle=0, barsize = 1,
+                    fontsize = 4, align = T, angle=0, barsize = 1,offset.text = 10,
                     color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Macroscelidae") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
@@
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Xenarthra, label="Xenarthra", 
-                  fontsize = 4, align = T, angle=0, barsize = 1,
+                  fontsize = 4, align = T, angle=0, barsize = 1,offset.text = 10,
                   color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
@@
                         guide = guide_colourbar(title="Log2 Cancer Succeptibility Change"))+
-  xlim(0,1500) + 
+  xlim(0,250) +
   theme_tree2() + 
@@
   ) + 
-  xlab("Rate of Change, Body Size")
+  xlab("|Log2CancerSucceptabilityChange|")
 # f5a
@@
     ) %>% str_replace_all("\\.", " ") %>% str_replace("Root", "Atlantogenata") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora")),
-        Node = Node %>% str_replace_all("\\.", " ") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosorcida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora"))
+        Node = Node %>% str_replace_all("\\.", " ") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora"))
   ) %>%
@@
 With this information in hand, we calculated $K_{1}$ at all nodes, and then estimated the fold change in cancer succeptibility between an ancestral node and a given node as $\frac{K_{2}}{K_{1}}$ (__Table 4__).
+As body size and phylogeny are the strongest predictors of lifespan, we find that this regression is sufficiently robust without further metabolic or other covariates for our purposes herein.
 
@@
 In parallel to Afrotheria, cancer succeptibility increases `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold in Xenarthra.
-However, cancer risk slowly deflates as size decreases as one moves along the tree towards extant species, such as in Hoffman's Two Toed Sloth (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Choloepus hoffmanii") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change) and in the Nine-banded Armadillo (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Dasypus novemcinctus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change).
-
-Within Afrotheria, cancer succeptibility drops in Afrosorcida as species shrink (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrosorcida") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, then stagnates for the Cape Golden Mole) - but then rises `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Echinops telfairi") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold towards the Lesser Hedgehog Tenrec.
+However, cancer risk slowly deflates as size decreases as one moves along the tree towards extant species, such as in Hoffman's Two Toed Sloth (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Choloepus hoffmanni") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change) and in the Nine-banded Armadillo (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Dasypus novemcinctus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change).
+
+Within Afrotheria, cancer succeptibility drops in Afrosoricida as species shrink (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrosoricida") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, then stagnates for the Cape Golden Mole) - but then rises `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Echinops telfairi") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold towards the Lesser Hedgehog Tenrec.
 In parallel, Afroinsectivora does not increase in cancer succeptibility, and decreases once more at the Cape Elephant Shrew (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantulus edwardii") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
@@
 While the Rock Hyrax decreases in cancer succeptibility as expected (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Procavia capensis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold), Tethytheria sees a sharp increase in cancer risk (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
-Within Tethytheria, the Manatee's cancer risk increases once more `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Trichechus manatus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, while Proboscidae's cancer risk drops precipitously along with its body size (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Proboscidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
-Yet, within Proboscidae we see the biggest increases: right off the bat, we see that the cancer succeptibility of  Elephantidae and the American Mastodon skyrocket by `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammut americanum") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively.
+Within Tethytheria, the Manatee's cancer risk increases once more `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Trichechus manatus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, while Proboscidea's cancer risk drops precipitously along with its body size (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Proboscidea") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
+Yet, within Proboscidea we see the biggest increases: right off the bat, we see that the cancer succeptibility of  Elephantidae and the American Mastodon skyrocket by `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammut americanum") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively.
 Both Elephantina and Loxodontini in Elephantidae have a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantina") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer succeptibility.
@@
 
-What biological mechanisms underlie the evolution of thousand- to hundred million- increases in cancer susceptibility during the origins of Afrotherians, which are essential for large body size and long lifespan to evolve
-
-`r warning("TODO: Body Size & implications of accordioning of Body Size")`
-`r warning("TODO: Next, discuss how the giant initial increases in cancer risk, and the lack of fully commensurate changes even in nodes that decrease in size, may explain why we see TSG duplicates conserved even in smaller species")`
-`r warning("TODO: Lifespan regression, and the possible impact of lacking other data. Key: We have no other data for most species, and would have to estimate them too.")`
-`r warning("TODO: Copy Numbers, specifically how rapidly some gene families expanded, and the issue of genome quality")`
-`r warning("TODO: Effect of genome quality with CNs.\nCHANGE: On second thought, after fixing the model, it doesn't seem any one metric affects CN too much, except for Genome (duh).
-Even Genome though has a much smaller impact than CEGMA stats, although those too don't explain things fully.
-Haven't included yet older genome models though, so this is TBD for now.")`
-`r warning("TODO: ")`
-`r warning("I placed this here, but its half-discussion, half-intro, so edit!")`
-Candidate gene studies, for example, have identified functional duplicates of the tumor suppressors TP53  and LIF in elephants.
-In a larger candidate gene study, Caulin et al.
-characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans.
-However, the probability of developing cancer is similar for small, short-lived mammals such as mice and for large, long-lived mammals such as elephants.
-
-In stark contrast, genome-wide studies of unusually large or long-lived species such as the bowhead whale (Keane et al., 2015), Myotid bats (Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al., 2011), and blind mole rat (Fang et al., 2014) did not find an over representation of tumor suppressors among duplicate genes.
-
-A genomic analysis of genetic changes associated with the evolution of enhanced cancer resistance in the elephant lineage has yet to be performed.
-Thus it is not clear if the duplication of TP53 and LIF reflects a general pattern of tumor suppressor duplication in the elephant lineage, unlike other lineages that resolved Peto?s paradox, or from the kinds of ascertainment biases common in candidate gene studies.
-
-`r warning("TODO: Discuss ORAs + cancer risk")`
+With our results, we have demonstrated that gigantism in _Atlantogenata_ was not limited to extant elephants, but rather occured at various points in the evolution of the clade; however, the hundred-fold to hundred-million-fold increases in cancer risk that is associated with these increases however poses an inate challenge in the evolution and persistance of this trait.
+The environmental selective pressures on body size have long been the fascination of evolutionary biologists, and the influence of climate, predation, geography, and ecological niche on body size have been well established.
+[Cooper & Purvis 2010, MORE CITATIONS].
+Indeed, there is a general trend, known as Cope's Rule, for body sizes of species to increase over time [Citation for Cope's Rule].
+However, for all the research on body size that has been done thus far, the mechanisms that enable a release from the negative pressure on body size exerted on cancer risk has proven more elusive.
+
+Furthermore, we show that tumor suppressor duplications are enriched not only in large, extant species, but also in large common ancestors, and that these duplications evolved throughout the tree, rather than in concert.
+The ancestral body sizes of many of the subclades in _Atlantogenata_ were estimated to be large, and the estimated cancer risk increases - even for the small clades like _Afroinsectivora_ may explain why these TSG gene duplications occured early on, and persisted even in species that we did not expect to have high risks of cancer.
+However, some of our results also provide interesting evidence for a paradigm of TSG duplicaitons being a pervasive phenomenon, rather than a specific mechansim that occurs after the evolution of gigantism.
+At the common ancestor of _Proboscidea_, which was small relative to both its ancestors and its decendants, we see the emergence of various TSG duplications, which may have enabled the stratospheric increase in body size of modern elephants.
+We also identified many TSG duplication events in smaller species and lineages such as in _Chrysochloris asiatica_ and _Elephantulus edwardii_, although these may have been the result of low-quality genomes.
+
+The impacts and takeaways of this study are limited by the quality and quantity of _Atlantogenatan_ genomes that were available, and our available knowledge of the livespans and cancer risk of the extant species.
+For many of our species, studies in captivity are limited, and the species are relatively understudied from a longitudinal perspective, such as with the Cape Elephant Shrew and Cape Golden Mole.
+Furthermore, while there is recent interest in resequencing and improving the quality of these assemblies, at the time of this writing there is still quite a ways to go in order to have genomes of a suffiently rigorous quality to make stronger inferences about gene copy number expansions and contractions (which were not considered in this study for this reason).
+
+The lack of a stronger signal from tumor suppressor duplications is likely a result of the strong effect size on both cancer risk and organismal toxicity that a TSG duplication would provide.
+The duplication of a tumor suppressor in many cases is associated with mild toxicity, although it greatly varies given the context and TSG in questionl; however, a single TSG duplication can also provide significant protection against cancer.
+For example, the overexpression of TP53 in mice, while protective of cancer, is associated with progeria and early death; however, if an additional copy of TP53 is introduced with its regulatory elements intact, the mice are healthy and experience normaging, while also demonstrating an enhanced response to cellular stress and lower rates of cancer.
+[@CITE the mouse TP53 papers].
+In light of this, it is fascinating that our results in the elephant lineage suggest that the duplication of TP53 regulators preceded the retroduplication and expansion of TP53, as this likely would have lowered the toxicity of the intial duplication and thus enabled it to occur.
+Given a suffienct selective pressure on increasing body size, it stands to reason that events like this could alleviate the negative pleiotropy of TSG duplications sufficiently to enable their persistence and allow for subsequent refinment over evolutionary time.
+
+By combining a phylogenetic study on body size in addition to a survey of copy number across nearly all protein-coding genes, we provide a comprehensive look at the question of the role of cancer risk and body size in _Atlantogenata_ that may provide broader insight to other mammalian species.
+Our study was initially motivated by the identification of functional duplicates of tumor suppressors, such as TP53 and LIF in elephants [@Abbeglen2015; @Sulak2016; @Vazquez2018].
+Further in support of our results, a larger candidate gene study by Caulin et al.
+[@CITE] characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans.
+However, while candidate gene studies are useful, by their very design they are biased in determining larger patterns of evolution of traits.
+Without addressing these questions with a genome-wide approach, any and all insights will be inevitably limited to a fraction of the whole story.
+
+Our results suggest that the pervasive duplication of tumor suppressors may help enable the evolution of larger body sizes by lowering the cancer risk of species, either prior to or in lockstep with increasing body size.
+however, this is unlikely to be the only genetic mechanism at play in this scenario.
+In genome-wide studies of unusually large or long-lived species such as the bowhead whale (Keane et al., 2015), Myotid bats (Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al., 2011), and blind mole rat (Fang et al., 2014), there were cases of overrepresentation of TSGs among duplicate genes that were outshadowed by the identification of strong signatures of positive selection at TSGs.
+While the evolution of regulatory and coding elements of both TSGs and other non-cannonical tumor suppressor genes have been shown to be important for mediating the cancer risk of long-lived species, there has been no attention given to the possiblity of TSG duplications providing a relaxation of possible negative pleitropy that could result from these traits.
+It has been well-established in the literature that genes duplication events allow for evolutionary drift in one of the copies, which frequently results in neofunctionalization or specialization of the two copies [@ now actually support this].
+While this is beyond the scope of our study, a promising future direction of this work would include an evolutionary analysis of duplicated genes relative to each other to see if this has already occured between the pairs of genes we have identified.
 
@@
 warning("TODO: Make HQ version of this figure")
-f2.1
+image_read("../output/figs/RBHBDiagram.png") %>% image_ggplot()
 ```
@@
 ```{r Supplementary Figure 8, fig.cap="Supplementary Figure 8: The time-calibrated Eutherian tree used as input for StableTraits", warning=T}
-# p.time.time.tree.Eutheria <-
-#   time.tree %>% 
+p.time.tree.Eutheria <-
+  time.tree %>%
   # as_tibble %>%
@@
   # as.treedata %>%
-#   ggtree(
+  ggtree(
     # branch.length = "time",
     # aes(color = lnSize),
-#     layout="circular"
-#   ) %>% 
-#   ggtree::revts() + 
-#     geom_highlight(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), fill="goldenrod") + geom_cladelabel(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), label="Atlantogenata", align = F, offset = 2, offset.text = 2, geom = "label", fill="goldenrod") +
-#   geom_highlight(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill="aquamarine") + geom_cladelabel(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill = "aquamarine", label="Chiroptera", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label") +
-#     geom_highlight(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1") + geom_cladelabel(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1", label="Rodentia", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label") +
-#   geom_highlight(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue") + geom_cladelabel(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue", label="Primata", align = F, offset = 2, offset.text = 3, hjust = 0, geom = "label") +
-#   geom_highlight(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1") + geom_cladelabel(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1", label="Cetacea", align = F, offset = 2, offset.text = 2, hjust = 0, geom = "label") +
-#   theme_tree()
-warning("TODO: this causes a segfault, find out why!")
+    layout="circular"
+  ) %>%
+  ggtree::revts() +
+    geom_highlight(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), fill="goldenrod") + geom_cladelabel(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), label="Atlantogenata", align = F, offset = 2, offset.text = 2, geom = "label", fill="goldenrod") +
+  geom_highlight(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill="aquamarine") + geom_cladelabel(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill = "aquamarine", label="Chiroptera", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label") +
+    geom_highlight(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1") + geom_cladelabel(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1", label="Rodentia", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label") +
+  geom_highlight(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue") + geom_cladelabel(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue", label="Primata", align = F, offset = 2, offset.text = 3, hjust = 0, geom = "label") +
+  geom_highlight(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1") + geom_cladelabel(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1", label="Cetacea", align = F, offset = 2, offset.text = 2, hjust = 0, geom = "label") +
+  theme_tree()
+
+time.tree.atlantogenata <- time.tree %>%
+  tree_subset(., MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), levels_back = 0)
+
+p.time.tree.Atlantogenata <-
+  time.tree.atlantogenata %>% 
+  # as_tibble %>%
+  # mutate(time = branch.length) %>%
+  # select(-parent, -node, -branch.length) %>%
+  # full_join(., bsize.tree %>% as_tibble, by="label") %>%
+  # as.treedata %>%
+  ggtree(
+    # branch.length = "time",
+    # aes(color = lnSize),
+    layout="circular"
+  ) %>%
+  ggtree::revts() +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Loxodonta.africana", "Procavia.capensis"), fill="goldenrod") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Loxodonta.africana", "Procavia.capensis"), label="Paenungulata", align = F, offset = 2, offset.text = 10, geom = "label", fill="goldenrod", fontsize = 8) +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Elephantulus.edwardii", "Chambius.kasserinensis"), fill="firebrick1") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Elephantulus.edwardii", "Chambius.kasserinensis"), fill="firebrick1", label="Macroscelidae", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label", fontsize = 8) +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Chrysochloris.asiatica", "Echinops.telfairi"), fill="deepskyblue") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Chrysochloris.asiatica", "Echinops.telfairi"), fill="deepskyblue", label="Afrosoricida", align = F, offset = 2, offset.text = 15, hjust = 0, geom = "label", fontsize = 8) +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), fill="darkseagreen1") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), fill="darkseagreen1", label="Xenarthra", align = F, offset = 2, offset.text = 7, hjust = 0, geom = "label", fontsize = 8) +
+  geom_point2(aes(subset=(label %in% c("Loxodonta.africana", "Trichechus.manatus", "Procavia.capensis"))), size=4, pch = 16, fill="black")+
+    geom_point2(aes(subset=(label %in% c("Loxodonta.africana", "Trichechus.manatus", "Procavia.capensis"))), size=2, pch = 16, color="red")+
+  theme_tree()
+# p.time.tree.Atlantogenata2 <-
+  time.tree.atlantogenata %>% 
+  # as_tibble %>%
+  # mutate(time = branch.length) %>%
+  # select(-parent, -node, -branch.length) %>%
+  # full_join(., bsize.tree %>% as_tibble, by="label") %>%
+  # as.treedata %>%
+  ggtree(
+    # branch.length = "time",
+    # aes(color = lnSize),
+    layout="circular"
+  ) %>%
+  ggtree::revts() +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Loxodonta.africana", "Procavia.capensis"), fill="goldenrod") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Loxodonta.africana", "Procavia.capensis"), label="Paenungulata", align = F, offset = 2, offset.text = 10, geom = "label", fill="goldenrod", fontsize = 8) +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Elephantulus.edwardii", "Chambius.kasserinensis"), fill="firebrick1") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Elephantulus.edwardii", "Chambius.kasserinensis"), fill="firebrick1", label="Macroscelidae", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label", fontsize = 8) +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Chrysochloris.asiatica", "Echinops.telfairi"), fill="deepskyblue") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Chrysochloris.asiatica", "Echinops.telfairi"), fill="deepskyblue", label="Afrosoricida", align = F, offset = 2, offset.text = 15, hjust = 0, geom = "label", fontsize = 8) +
+  geom_highlight(node = MRCA(time.tree.atlantogenata, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), fill="darkseagreen1") + 
+    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), fill="darkseagreen1", label="Xenarthra", align = F, offset = 2, offset.text = 7, hjust = 0, geom = "label", fontsize = 8) +
+  geom_point2(aes(subset=(label %in% c("Choloepus.hoffmanni", "Chrysochloris.asiatica", "Dasypus.novemcinctus", "Echinops.telfairi", "Elephantulus.edwardii", "Elephas.maximus", "Loxodonta.africana", "Loxodonta.cyclotis", "Mammut.americanum", "Mammuthus.columbi", "Mammuthus.primigenius", "Orycteropus.afer", "Palaeoloxodon.antiquus", "Procavia.capensis", "Trichechus.manatus"))), size=6, pch = 16, fill="black")+
+    geom_point2(aes(subset=(label %in% c("Choloepus.hoffmanni", "Chrysochloris.asiatica", "Dasypus.novemcinctus", "Echinops.telfairi", "Elephantulus.edwardii", "Elephas.maximus", "Loxodonta.africana", "Loxodonta.cyclotis", "Mammut.americanum", "Mammuthus.columbi", "Mammuthus.primigenius", "Orycteropus.afer", "Palaeoloxodon.antiquus", "Procavia.capensis", "Trichechus.manatus"))), size=3, pch = 16, color="white")+
+  theme_tree()
+# warning("TODO: this causes a segfault, find out why!")
 ```

### 2020-05-20 paper_PLOS/paper_PLOS_draft.tex
first draft

@@
 \textbf\newline{Pervasive duplication of tumor suppressor genes preceded parallel
-evolution of large bodied Paenungulates} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+evolution of large bodied Atlantogenatans} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
-porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
-dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
-velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
-finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
-diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
-ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
-pulvinar, volutpat risus.
+Cancer is an intrinsic disease of multicellular organisms. Within a
+species, the size of an animal, - correlated with the individual's
+number of cells - and its lifespan - correlated with increasing cellular
+damage over time - positively correlate with the risk any individual has
+to form tumors. Between species, however, we do not observe any
+correlation between size, lifespan, and cancer, a phenomenon that
+referred to as \emph{Peto's Paradox}. Elephants are a particularly
+interesting member of this class of paradoxical animals, since they are
+a set of large species deeply nested in a clade of smaller species,
+indicating a recent gain of size. Recent work has identified several
+individual cases of gene duplicates contributing to the increased cancer
+resistance of elephants, which suggests that duplication of tumor
+suppressor genes may play a more general role in mediating Peto's
+Paradox by increasing cancer resistance in large, long-lived species. By
+using a Reciprocal Best-Hit BLAT search approach, we investigated copy
+numbers of all protein-coding genes in \emph{Atlantogenatan} genomes to
+see if there is any correlation between the copy number of duplicates
+and changes body size along the phylogenetic tree. From an initial set
+of 18,011 protein-coding genes in hg38, we identified a median of 13,880
+genes in \emph{Atlantogenatan} genomes, of which a median of 940 genes
+are duplicated. We find that, just as body size fluctuates throughout
+\emph{Atlantogenata}, genes involved in tumor suppressor pathways are
+also duplicated throughout the phylogenetic tree. Extant species of
+elephants, however, show active transcription of both canonical and
+duplicated copies of tummor suppressors that duplicated prior to and
+during their sudden increase in body size, suggesting that the
+duplication of tumor suppressor genes facilitates the evolution of
+increased body size by compensating for the increased cancer risk.
 
@@
 
-\emph{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit BLAT
-(RBHB) pipeline to quickly identify homologs and estimate gene copy
+\textbf{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit
+BLAT (RBHB) pipeline to quickly identify homologs and estimate gene copy
 numbers (\textbf{Figure 1A}). The Reciprocal Best Hit (RBH) search
@@
 
-\emph{Effective Copy Number By Coverage:} In lower-quality genomes, many
-genes are fragmented across multiple scaffolds, which results in BLAT
-calling multiple hits when in reality there is only one gene. To
+\textbf{Effective Copy Number By Coverage:} In lower-quality genomes,
+many genes are fragmented across multiple scaffolds, which results in
+BLAT calling multiple hits when in reality there is only one gene. To
 compensate for this, we came up with a novel statistic, Estimated Copy
@@
 
-\emph{RecSearch Pipeline:} We created a custom Python pipeline for
+\textbf{RecSearch Pipeline} We created a custom Python pipeline for
 automating RBHB searches between a single reference genome and multiple
@@
 filtering out problematic protein queries, we then used our pipeline
-(Figure 1A) to search for all copies of our 20456 query genes in
+(Figure 1A) to search for all copies of our 18011 query genes in
 publicly available Afrotherian genomes, including African savannah
@@
 
-\emph{Duplication gene inclusion criteria:} In order to condense
+\textbf{Query protein criteria} To assemble our query list, we first
+removed all unnamed genes from UP000005640. Next, we excluded genes from
+downstream analyses for which assignment of homology was uncertain,
+including uncharacterized ORFs (991), LOC (63), HLA genes (402),
+replication dependent histones (72), odorant receptors (499), ribosomal
+proteins (410), zinc finger transcription factors (1983), viral and
+repetitive-element-associated proteins (82) and any protein described as
+either ``Uncharacterized,'' ``Putative,'' or ``Fragment'' by UniProt in
+UP000005640 (30724), leaving us with a final set of 37582 query protein
+sequences, corresponding to 18011 genes.
+
+\textbf{Duplication gene inclusion criteria} In order to condense
 transcript-level hits into single gene loci, and to resolve many-to-one
@@
 genes and genomes. The results of our initial search are summarized in
-Figure 1B. Overall, we identified {[}MEDIAN{]} genes across all species,
-or {[}\%HITS/QUERIES{]} of our starting query genes.
-
-\emph{Duplicate gene exclusion criteria:} We excluded genes from
-downstream analyses for which assignment of homology was uncertain,
-including uncharacterized ORFs (17), LOC (17), HLA genes (17),
-replication dependent histones (17), odorant receptors (17), ribosomal
-proteins (17), zinc finger transcription factors (17), viral and
-repetitive-element-associated proteins (17) and any protein described as
-either ``Uncharacterized,'' ``Putative,'' or ``Fragment'' by UniProt in
-UP000005640 (17).
-
-\textbf{Orthogonal Genome Assessment using CEGMA} In order to determine
-the effect of genome quality on our results, we used the gVolante
-webserver and CEGMA to assess the quality and completeness of the
-genome. CEGMA was run using the default settings of {[}{]}, and the
-mammalian-specific core gene sets.
+Figure 1B. Overall, we identified 13880 genes across all species, or
+77.1\% of our starting query genes.
+
+\textbf{Genome Quality Assessment using CEGMA} In order to determine the
+effect of genome quality on our results, we used the gVolante webserver
+and CEGMA to assess the quality and completeness of the genome
+{[}{\textbf{???}},{\textbf{??}}{]}. CEGMA was run using the default
+settings for mammals (``Cut-off length for sequence statistics and
+composition'' = 1;``CEGMA max intron length'' = 100000; ``CEGMA gene
+flanks'' = 10000, ``Selected reference gene set'' = CVG). For each
+genome, we generated a correlation matrix using the aforementioned
+genome quality scores, and either the mean Copy Number or mean ECNC for
+all hits in the genome.
 
@@
 
-In order to determine the cancer risk at each node, we used a simplified
-multistage cancer risk model for body size and lifespan
-{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. We defined the
+In order to determine the cancer risk \(K\) at each node, we first
+needed to calcultate ancestral lifespans at each node. To do so, we used
+a Phylogenetic Generalized Least-Square Regression (PGLS)
+{[}{\textbf{???}},{\textbf{??}}{]} to calculate estimated ancestral
+lifespans across \emph{Atlantogenata} using our estimates for body size
+at each node.
+
+Next, we used a simplified multistage cancer risk model for body size
+\(D\) and lifespan \(t\): \(K \approx Dt^6\)
+{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. To determine the
+change in \(K\) between nodes, we obtained the ratio between the cancer
+risk \(K_{1}\) at any given node, and the cancer risk \(K_{2}\) at its
+ancestral node, using the equation
+\(\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}\).
+Finally, to simplify comparisons, we calculated the fold change cancer
+risk between a node and its ancestor as
+\(\log_{2}(\frac{K_{2}}{K_{1}})\).
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 1-1.pdf}
-\caption{Figure 1: Body sizes rapidly and frequently expand in
-Eutherians, especially in Atlantogenata. \textbf{A)} Tree of Eutherian
-species, colored by ln(Body Size) and with branch lengths set to the
-rate of change in body sizes, normalized by the square root of the root
-branch. Atlantogenata is highlighted at the bottom. \textbf{B)} Zoom-in
-of (A) on Atlantogenata. Silhuetes for the African Elephant, West Indian
-Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole,
-Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their
-extant body sizes, while clade labels are colored based on the common
-ancestor's estimated body size}
-\end{figure}
+\textbackslash{}begin\{figure\}
+
+\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 1-1}
+\hfill{}
+
+\textbackslash{}caption\{Figure 1: Body sizes rapidly and frequently
+expand in Eutherians, especially in Atlantogenata. \textbf{A)} Tree of
+Eutherian species, colored by ln(Body Size) and with branch lengths set
+to the rate of change in body sizes, normalized by the square root of
+the root branch. Atlantogenata is highlighted at the bottom. \textbf{B)}
+Zoom-in of (A) on Atlantogenata. Silhuetes for the African Elephant,
+West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape
+Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are
+colored by their extant body sizes, while clade labels are colored based
+on the common ancestor's estimated body size\}\label{fig:Figure 1}
+\textbackslash{}end\{figure\}
 
@@
 \toprule
-Ancestor/Species & Estimated Body Size (log(g)) & 95\% CI (Low) & 95\%
-CI (High) & Rate (sqrt)\tabularnewline
+Node & Size (log(g)) & 95\% CI (Low) & 95\% CI (High) & Rate
+(sqrt)\tabularnewline
 \midrule
@@
 evolution, which allows for occasional large jumps in traits (stable
-model) and out performs standard Brownian motion and Ornstein?Uhlenbeck
+model) and out performs standard Brownian motion and Ornstein-Uhlenbeck
 models of character evolution {[}25{]}.
@@
 small (0.26kg, 95\% CI: 0.31-3.01kg) and that substantial accelerations
-in the rate of body mass evolution occurred coincident with a 65×
-increase in body mass in the stem-lineage of \emph{Pseudoungulata}
-(17kg), a 1.5× increase in body mass in the stem-lineage of
-\emph{Paenungulata} (25kg), a 12× increase in body mass in the
-stem-lineage of \emph{Tehthytheria} (296kg), and a 14× increase in body
-mass in the stem-lineage of \emph{Proboscidea} (4,100kg; Figure 1). The
-ancestral \emph{Hyracoidea} was inferred to be relatively small
-(2.86-15.71kg), and rate accelerations were coincident with independent
-body mass increases in large hyraxes such as \emph{Titanohyrax andrewsi}
-(67× increase in body mass). While the body mass of the ancestral
-\emph{Sirenian} was inferred to be large (61-656kg), a rate acceleration
-occurred coincident with a 10× body mass increase in Stellar's sea cow.
-Rate accelerations also occurred coincident with 36× body mass reduction
-in the stem-lineage of the dwarf elephants \emph{Elephas}
+in the rate of body mass evolution occurred coincident with a 2.77x
+increase \textbf{(Vinnie wrote 65× increase")} in body mass in the
+stem-lineage of \emph{Pseudoungulata} (9.76kg) \textbf{(``Vinnie wrote
+(17kg)'')}, a 1.57x increase \textbf{(``Vinnie wrote 1.5x increase'')}
+in body mass in the stem-lineage of \emph{Paenungulata} (10.13kg)
+\textbf{(``Vinnie wrote (25kg)'')}, a 1.52x increase \textbf{(``Vinnie
+wrote 12× increase'')} in body mass in the stem-lineage of
+\emph{Tehthytheria} (12.6kg) \textbf{(``Vinnie wrote (296kg)'')}, and a
+1.22x increase \textbf{(``Vinnie wrote 14× increase'')} in body mass in
+the stem-lineage of \emph{Proboscidea} (15.23kg) \textbf{(``Vinnie wrote
+(4,100kg)'')} (Figure 1). The ancestral \emph{Hyracoidea} was inferred
+to be relatively small 7.96kg-11.68kg \textbf{(Vinnie wrote
+(2.86-15.71kg))}, and rate accelerations were coincident with
+independent body mass increases in large hyraxes such as
+\emph{Titanohyrax andrewsi} 12.97kg; 1.33x increase (67× increase in
+body mass). While the body mass of the ancestral \emph{Sirenian} was
+inferred to be large (61-656kg), a rate acceleration occurred coincident
+with a 10× body mass increase in Stellar's sea cow. Rate accelerations
+also occurred coincident with 36× body mass reduction in the
+stem-lineage of the dwarf elephants \emph{Elephas}
 (\emph{Palaeoloxodon}) \emph{falconeri} and \emph{Palaeoloxodon
@@
 \hypertarget{pervasive-duplication-of-tumor-suppressors-during-the-origins-of-large-bodied-afrotheirans}{%
-\subsection{Pervasive duplication of tumor suppressors during the
+\subsection*{Pervasive duplication of tumor suppressors during the
 origins of large bodied
 Afrotheirans}\label{pervasive-duplication-of-tumor-suppressors-during-the-origins-of-large-bodied-afrotheirans}}
-
-\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2C1-1.pdf}
-\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2C2-1.pdf}
-
-\begin{verbatim}
-## Warning: TODO: Table 2 is done, but Stargazer has some weird bug stopping it
-## from running...
-\end{verbatim}
-
-Enhanced cancer suppression may have evolved through many mechanisms;
-among the most parsimonious is an increase in the copy number of genes
-with tumor suppressor functions. Previous studies focusing on candidate
-gene studies, for example, have identified increased copy number of the
-tumor suppressors \emph{TP53} and \emph{LIF} in elephants
-{[}{\textbf{???}},11,21--23{]}. Therefore, in order to test whether this
-was a pervasive phenomena genome-wide in \emph{Afrotherians}, we used a
-Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in
-\emph{Afrotherian} and \emph{Atlantogenatan} genomes (Fig. 2A). Because
-RBHB-like approaches can over-estimate copy number when genes are
-fragmented or incorrectly assembled across multiple scaffolds, we also
-inferred copy number using a complementary method that quantifies the
-ratio between observed and expected gene coverage per nucleotide (ECNC)
-(Supplementary Figure 1). By only including nucleotides from the query
-sequence that were observed in the target genome, we also correct for
-partial hits where some or all of the homologs of a gene have diverged
-from the human homolog.
-
-Because our sequence database included various protein transcripts for
-each gene, in order to obtain gene-level copy number information and
-elimate any many-to-one mappings of hits, we labeled each exon of every
-reciprocal best hit (RBH) with the gene corresponding to the query
-transcript and merged all overlapping exons; next, we eliminated any
-many-to-one exons that resulted from the previous step. Finally, we
-reassembled the gene loci based on the original transcript starts and
-ends, and the collapsed exon data, obtaining the full sequence of each
-RBH locus. Genes were considered to be duplicated if its copy number via
-RBHB was greater than or equal to 2, and the maximum ECNC among all
-transcripts prior to filtering was greater than or equal to 1.50. This
-cutoff of ECNC was selected to account for truncated gene duplications,
-which have been shown to be functional in various examples
-{[}{\textbf{???}} examples of this{]}.
-
-To reconcile the Atlantogenatan phylogeny with duplication events, we
-used maximum likelihood to reconstruct likely ancestral copy numbers for
-each gene at each node in the phylogeny. To define the copy number of a
-gene, we conservatively used the lesser value between the RBHB hit
-count, and the ECNC value rounded to the nearest whole number. In order
-to perform
-
-Next, in order to select genes and duplicates which were likely
-functional, we omitted any hits that were not supported by either the
-gene prediction method GenScan, or by at least one transcript assembled
-from publically-available RNA-seq data.
-
-We describe the number of genes that increased in each lineage in
-\emph{Atlantogenata} in Figure 2. Among the genes that increased in copy
-number in the elephant lineage are TP53 and LIF, as previously
-described. Furthermore, we identify
-
-\hypertarget{duplications-that-occured-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}{%
-\subsection{Duplications that occured recently in Probodiscea are
-enriched for tumor suppressor
-pathways}\label{duplications-that-occured-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}}
-
-\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 3-1.pdf}
-\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 3-2.pdf}
-
-In order to infer the functional consequences of these gene
-duplications, we tested if duplicate genes were enriched in specific
-pathways relative to our initial query set of genes. We used
-
-\hypertarget{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}{%
-\subsection{\texorpdfstring{Concerted duplication of TP53 and
-TP53-related genes towards
-\emph{Probodiscea}}{Concerted duplication of TP53 and TP53-related genes towards Probodiscea}}\label{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}}
-
-\begin{verbatim}
-## Warning: TODO: Make Fig4
-\end{verbatim}
-
-\hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
-\subsection{Step-wise reduction of intrinsic cancer risk in large,
-long-lived
-Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}
-
-\begin{verbatim}
-## Warning: TODO: Caption Fig 5
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: TODO: de-comment geom_tiplab in f5a! Commented currently for offline
-## usage!
-\end{verbatim}
-
-\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 5-1.pdf}
-
-\begin{verbatim}
-## 
-## % Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
-## % Date and time: Wed, May 13, 2020 - 20:08:58
-## \begin{table}[!htbp] \centering 
-##   \caption{Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression} 
-##   \label{} 
-## \begin{tabular}{@{\extracolsep{5pt}}lc} 
-## \\[-1.8ex]\hline 
-## \hline \\[-1.8ex] 
-##  & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
-## \cline{2-2} 
-## \\[-1.8ex] & Lifespan \\ 
-## \hline \\[-1.8ex] 
-##  lnSize & 0.100 \\ 
-##   & (0.121) \\ 
-##   & t = 0.826 \\ 
-##   & p = 0.409 \\ 
-##   Constant & 1.943 \\ 
-##   & (1.385) \\ 
-##   & t = 1.403 \\ 
-##   & p = 0.161 \\ 
-##  \hline \\[-1.8ex] 
-## Observations & 28 \\ 
-## Log Likelihood & $-$39.636 \\ 
-## Akaike Inf. Crit. & 85.273 \\ 
-## Bayesian Inf. Crit. & 89.047 \\ 
-## \hline 
-## \hline \\[-1.8ex] 
-## \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-## \end{tabular} 
-## \end{table}
-\end{verbatim}
-
-\begin{longtable}[]{@{}lrrllrrl@{}}
-\caption{A table generated by the longtable package.}\tabularnewline
-\toprule
-Node & lnSize & Est. Lifespan & Est. Cancer Susceptibility (K1) & Est.
-Ancestral K2 & Change in K (K2/K1) & log2 Fold Change &
-Ancestor\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Node & lnSize & Est. Lifespan & Est. Cancer Susceptibility (K1) & Est.
-Ancestral K2 & Change in K (K2/K1) & log2 Fold Change &
-Ancestor\tabularnewline
-\midrule
-\endhead
-Loxodontini & 16 & 34.38 & 1.47e+16 & 2.97e+15 & 4.940000e+00 & 2.31 &
-Elephantidae\tabularnewline
-Loxodonta africana & 15 & 65.00 & 2.47e+17 & 1.47e+16 & 1.681000e+01 &
-4.07 & Loxodontini\tabularnewline
-Loxodona & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 & 0.00 &
-Loxodontini\tabularnewline
-Loxodonta cyclotis & 15 & 31.12 & 2.97e+15 & 1.47e+16 & 2.000000e-01 &
--2.31 & Loxodona\tabularnewline
-Palaeoloxodon antiquus & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00
-& 0.00 & Loxodona\tabularnewline
-Elephantidae & 15 & 31.12 & 2.97e+15 & 1.40e+07 & 2.125534e+08 & 27.66 &
-NA\tabularnewline
-Elephantina & 16 & 34.38 & 1.47e+16 & 2.97e+15 & 4.940000e+00 & 2.31 &
-Elephantidae\tabularnewline
-Elephas maximus & 15 & 65.50 & 2.58e+17 & 1.47e+16 & 1.760000e+01 & 4.14
-& Elephantina\tabularnewline
-Mammuthus & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 & 0.00 &
-Elephantina\tabularnewline
-Mammuthus primigenius & 15 & 31.12 & 2.97e+15 & 1.47e+16 & 2.000000e-01
-& -2.31 & Mammuthus\tabularnewline
-Mammuthus columbi & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 &
-0.00 & Mammuthus\tabularnewline
-Mammut americanum & 16 & 34.38 & 1.47e+16 & 1.40e+07 & 1.050567e+09 &
-29.97 & NA\tabularnewline
-Tethytheria & 13 & 25.49 & 1.21e+14 & 1.01e+12 & 1.207400e+02 & 6.92 &
-Paenungulata\tabularnewline
-Trichechus manatus & 13 & 69.00 & 4.77e+16 & 1.21e+14 & 3.929800e+02 &
-8.62 & Tethytheria\tabularnewline
-Paenungulata & 10 & 18.91 & 1.01e+12 & 1.01e+12 & 1.000000e+00 & 0.00 &
-Pseudoungulata\tabularnewline
-Procavia capensis & 8 & 14.80 & 3.13e+10 & 1.01e+12 & 3.000000e-02 &
--5.01 & Paenungulata\tabularnewline
-Pseudoungulata & 10 & 18.91 & 1.01e+12 & 1.69e+09 & 5.967900e+02 & 9.22
-& Afroinsectivora\tabularnewline
-Orycteropus afer & 11 & 29.80 & 4.19e+13 & 1.01e+12 & 4.167000e+01 &
-5.38 & Pseudoungulata\tabularnewline
-Elephantulus edwardii & 4 & 10.40 & 6.90e+07 & 1.69e+09 & 4.000000e-02 &
--4.61 & Afroinsectivora\tabularnewline
-Afrosorcida & 4 & 10.40 & 6.90e+07 & 1.69e+09 & 4.000000e-02 & -4.61 &
-Afrotheria\tabularnewline
-Chrysochloris asiatica & 4 & 10.40 & 6.90e+07 & 6.90e+07 & 1.000000e+00
-& 0.00 & NA\tabularnewline
-Echinops telfairi & 4 & 19.00 & 2.57e+09 & 6.90e+07 & 3.722000e+01 &
-5.22 & NA\tabularnewline
-Afrotheria & 6 & 12.69 & 1.69e+09 & 2.83e+06 & 5.967900e+02 & 9.22 &
-Atlantogenata\tabularnewline
-Xenarthra & 11 & 20.89 & 4.97e+12 & 2.83e+06 & 1.760358e+06 & 20.75 &
-Atlantogenata\tabularnewline
-Dasypus novemcinctus & 8 & 22.30 & 3.67e+11 & 4.97e+12 & 7.000000e-02 &
--3.76 & Xenarthra\tabularnewline
-Choloepus hoffmanni & 8 & 41.00 & 1.42e+13 & 4.97e+12 & 2.850000e+00 &
-1.51 & Xenarthra\tabularnewline
-Atlantogenata & 2 & 8.52 & 2.83e+06 & 2.83e+06 & 1.000000e+00 & 0.00 &
-Atlantogenata\tabularnewline
-Afroinsectivora & 6 & 12.69 & 1.69e+09 & 1.69e+09 & 1.000000e+00 & 0.00
-& Afrotheria\tabularnewline
-NA & 3 & 9.41 & 1.40e+07 & 1.21e+14 & 0.000000e+00 & -23.05 &
-Tethytheria\tabularnewline
-\bottomrule
-\end{longtable}
-
-The dramatic increase in body mass and lifespan in some Afrotherian
-lineages implies those lineages evolved reduced cancer risk. To infer
-the magnitude of these reductions we estimated differences in cancer
-risk between small bodied, short-lived species and large bodied,
-long-lived species as well as for reconstructed ancestral Afrotherians.
-Following {[}{\textbf{???}}{]} we estimate the intrinsic cancer risk as
-the product of risk associated with body mass and lifespan. Differences
-in cancer susceptibility \(K\) due to body mass differences between
-species can be approximated simply as the fold difference in body mass
-(\(D\)) between species {[}{\textbf{???}}{]}. The risk of developing
-cancer also increases in proportion to the sixth power of age and is
-approximated by the formula \(Ct^6\), in which the proportionality
-constant C that determines susceptibility to cancer induction is
-multiplied by the sixth power of the age in years, \(t\)
-{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. Thus we can estimate
-the intrinsic cancer risk for a species as \(K \approx Dt^6\).
-
-In order to estimate the intrinsic cancer risk of a species, we first
-obtained estimates for lifespans at ancestral nodes using PGLS and the
-model
-\(ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon\)
-(Figure ). With this information in hand, we calculated \(K_{1}\) at all
-nodes, and then estimated the fold change in cancer succeptibility
-between an ancestral node and a given node as \(\frac{K_{2}}{K_{1}}\)
-(\textbf{Table 4}).
-
-As shown in \textbf{Table 4}, cancer succeptibility skyrocketed at the
-initial divergence of Atlantogenata, followed by a generally upwards
-trend. At the common ancestor of Afrotheria there is an inital 9.22-fold
-increase in cancer risk. In parallel to Afrotheria, cancer
-succeptibility increases 20.75-fold in Xenarthra. However, cancer risk
-slowly deflates as size decreases as one moves along the tree towards
-extant species, such as in Hoffman's Two Toed Sloth (-fold change) and
-in the Nine-banded Armadillo (-fold change).
-
-Within Afrotheria, cancer succeptibility drops in Afrosorcida as species
-shrink (-4.61-fold, then stagnates for the Cape Golden Mole) - but then
-rises -fold towards the Lesser Hedgehog Tenrec. In parallel,
-Afroinsectivora does not increase in cancer succeptibility, and
-decreases once more at the Cape Elephant Shrew (-fold). The emergence of
-Pseudoungulata sees the next big leap in cancer succeptibility with a
-9.22-fold increase. The Aardvark further increases -fold, while we don't
-observe an increase at the common ancestor of Paenungulates. While the
-Rock Hyrax decreases in cancer succeptibility as expected (-fold),
-Tethytheria sees a sharp increase in cancer risk (6.92-fold). Within
-Tethytheria, the Manatee's cancer risk increases once more -fold, while
-Proboscidae's cancer risk drops precipitously along with its body size
-(-23.05-fold). Yet, within Proboscidae we see the biggest increases:
-right off the bat, we see that the cancer succeptibility of Elephantidae
-and the American Mastodon skyrocket by 27.66-fold and -fold,
-respectively. Both Elephantina and Loxodontini in Elephantidae have a
-2.31-fold increase in cancer succeptibility. Within Elephantina, cancer
-susceptibility stays stable at Mammuthus and in the Colombian Mammoth,
-and slightly decreases in the Wooly Mammoth (-fold). The three extant
-elephants - Asian Elephant in Elephantina, the African Savana Elephant
-in Loxodontini, and the African Forest Elephant in Loxodona, meanwhile,
-have parallel and similar decreases in both size and cancer
-succeptibility (-, -, and -fold, respectively). Neither the common
-ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further
-changes in cancer succeptibility.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-What biological mechanisms underlie the evolution of thousand- to
-hundred million- increases in cancer susceptibility during the origins
-of Afrotherians, which are essential for large body size and long
-lifespan to evolve
-
-Candidate gene studies, for example, have identified functional
-duplicates of the tumor suppressors TP53 and LIF in elephants. In a
-larger candidate gene study, Caulin et al.~characterized the copy number
-of 830 known tumor-suppressor genes across 36 mammals and identified 382
-putative duplicates, including duplicates in species with large body
-sizes and long life-spans. However, the probability of developing cancer
-is similar for small, short-lived mammals such as mice and for large,
-long-lived mammals such as elephants.
-
-In stark contrast, genome-wide studies of unusually large or long-lived
-species such as the bowhead whale (Keane et al., 2015), Myotid bats
-(Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al.,
-2011), and blind mole rat (Fang et al., 2014) did not find an over
-representation of tumor suppressors among duplicate genes.
-
-A genomic analysis of genetic changes associated with the evolution of
-enhanced cancer resistance in the elephant lineage has yet to be
-performed. Thus it is not clear if the duplication of TP53 and LIF
-reflects a general pattern of tumor suppressor duplication in the
-elephant lineage, unlike other lineages that resolved Peto?s paradox, or
-from the kinds of ascertainment biases common in candidate gene studies.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for
-his indispensible support.
-
-\hypertarget{conflicts-of-interest}{%
-\section{Conflicts of Interest}\label{conflicts-of-interest}}
-
-The Authors have no conflicts of interest to report
-
-\hypertarget{funding-source}{%
-\section{Funding Source}\label{funding-source}}
-
-We would like to thank the Department of Human Genetics at the
-University of Chicago for supporting this project.
-
-\hypertarget{supplementary-figures}{%
-\section{SUPPLEMENTARY FIGURES}\label{supplementary-figures}}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 1-1.pdf}
-\caption{Supplementary Figure 1: Estimated Copy Number by Coverage
-(ECNC) consolidates fragmented genes while accounting for missing
-domains in homologs. \textbf{A)} A single, contiguous gene homolog in a
-target genome with 100\% query length coverage has an ECNC of 1.0.
-\textbf{B)} Two contiguous gene homologs, each with 100\% query length
-coverage have an ECNC of 2.0. \textbf{C)} A single gene homolog, split
-across multiple scaffolds and contigs in a fragmented target genome;
-BLAT identifies each fragment as a single hit. Per nucleotide of query
-sequence, there is only one corresponding nucleotide over all the hits,
-thus the ECNC is 1.0. \textbf{D)} Two gene homologs, one fragmented and
-one contiguous. 100\% of nucleotides in the query sequence are
-represented between all hits; however, every nucleotide in the query has
-two matching nucleotides in the target genome, thus the ECNC is 2.0.
-\textbf{E)} One true gene homolog in the target genome, plus multiple
-hits of a conserved domain that span 20\% of the query sequence. While
-100\% of the query sequence is represented in total, 20\% of the
-nucleotides have 4 hits. Thus, the ECNC for this gene is 1.45.
-\textbf{F)} Two real gene homologs; one hit is contiguous, one hit is
-fragmented in two, and the tail end of both sequences was not identified
-by BLAT due to sequence divergence. Only 75\% of the query sequence was
-covered in total between the hits, but for that 75\%, each nucleotide
-has two hits. As such, ECNC is equal to 2.0 for this gene.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 2-1.pdf}
-\caption{Supplementary Figure 2: Gene copy increases polarized along
-Atlantogenata, colored by ln(Body Size), with branch lengths equal to
-the change in gene copy number}
-\end{figure}
-
-\begin{verbatim}
-## Warning: TODO: Make HQ version of this figure
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 1 rows containing missing values (geom_text).
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 3-1.pdf}
-\caption{Supplementary Figure 3: Full version of RecBlat strategy, but
-low quality}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 4-1.pdf}
-\caption{Supplementary Figure 4: Dot and Line plot for Body Sizes and
-such}
-\end{figure}
-
-\begin{verbatim}
-## $RBX1
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-1.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $LIN9
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-2.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $MND1
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-3.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $E2F2
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-4.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $MAX
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-5.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $SYCP3
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-6.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $CDK1
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-7.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $RNF168
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-8.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $RBBP8
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-9.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{verbatim}
-## 
-## $MAD2L1
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 5-10.pdf}
-\caption{Supplementary Figure 5: All the Gene Copy Trees for interesting
-genes duplicated in LoxAfr4 (RIP any trees if this gets printed)}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{paper_PLOS_draft_files/figure-latex/Supplementary Figure 7-1.pdf}
-\caption{Supplementary Figure 7: Correlation matrix heat map for genome
-quality metrics, ECNC, RBHB Copy, and Estimated Copy Number (lesser
-between ECNC and RBHB)}
-\end{figure}
-
-\begin{verbatim}
-## Warning: TODO: this causes a segfault, find out why!
-\end{verbatim}
-
-\hypertarget{supplementary-data}{%
-\section{Supplementary data}\label{supplementary-data}}
-
-\begin{verbatim}
-## Warning: TODO: Read in Eutheria.progress and present it to the bold.
-\end{verbatim}
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\addcontentsline{toc}{subsection}{Pervasive duplication of tumor
+suppressors during the origins of large bodied Afrotheirans}
 

### 2020-05-20 code/generalFunctions.R

@@ -616,7 +616,7 @@ class.pathway <- function(Pathway){
                               str_detect(P, "Pathways in Cancer|\\b[Rr][Aa][Ss]\\b|\\bRb\\b|Retinoblastoma|\\bPTEN\\b|Tumor suppressor|PI3"),
                               "Other Cancer-Related Pathways",
                               ifelse(
-                                str_detect(P, "[Ss]tress|[Hh]ypoxia|Oxidative damage|UPR") %>% str_detect("[Ff]luid", negate=T),
+                                str_detect(P, "(?!Fluid)(\\b[Ss]tress\\b|\\b[Hh]ypoxia\\b|Oxidative damage|\\bUPR\\b)"),
                                 "Cell Stress Response",
                                 ifelse(
                                   str_detect(P, "TOR"),
@@ -627,7 +627,11 @@ class.pathway <- function(Pathway){
                                     ifelse(
                                       str_detect(P, "[Oo]ncogen"),
                                       "Oncogenic Pathway",
-                                      "Other"
+                                      ifelse(
+                                        str_detect(P, "Insulin|IGF"),
+                                        "Insulin- and IGF-Related Pathways",
+                                        "Other"
+                                      )
                                       )
                                   )
                                 )
@@ -737,3 +741,237 @@ quiet <- function(x) {
   on.exit(sink()) 
   invisible(force(x)) 
 } 
+
+text.ratio.nodeByMRCA.ancNode <- function(nodeA, outNode, tree_table=tree.Atlantogenata.table, nodeid=T){
+  tree_table %<>% mutate(label=str_replace(label, " ", "."), Median = Median %>% exp)
+  if (nodeid){
+    nodeA = nodeid(tree_table, nodeA)
+    outNode = nodeid(tree_table, outNode)
+  }
+  real.node = tree_table %>% filter(node == MRCA(tree_table, nodeA, outNode) %>% pull(node)) %>% pull(node)
[…]

### 2020-05-28 paper_PLOS/paper_PLOS_draft.Rmd
Figure 1 corrected

@@
 ---
-title: Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans
+title: Pervasive duplication of tumor suppressor genes preceded parallel evolution
+  of large bodied Atlantogenatans
 author:
 - name: Juan Manuel Vazquez
-    email: juanvazquez@uchicago.edu
-    affiliation: University of Chicago
-    corresponding: juanvazquez@uchicago.edu
-  - name: Vincent J Lynch
-    email: vjlynch@buffalo.edu
-    affiliation: University at Buffalo, SUNY
+  affiliation: "University of Chicago"
+- affiliation: "SUNY Buffalo"
+  name: Vincent J Lynch
 address:
-  - code: University of Chicago
-    address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
-  - code: University at Buffalo, SUNY
-    address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
-abstract: |
-  Cancer is an intrinsic disease of multicellular organisms.
-Within a species, the size of an animal, - correlated with the individual's number of cells - and its lifespan - correlated with increasing cellular damage over time - positively correlate with the risk any individual has to form tumors.
-Between species, however, we do not observe any correlation between size, lifespan, and cancer, a phenomenon that referred to as _Peto's Paradox_.
-Elephants are a particularly interesting member of this class of paradoxical animals, since they are a set of large species deeply nested in a clade of smaller species, indicating a recent gain of size.
-Recent work has identified several individual cases of gene duplicates contributing to the increased cancer resistance of elephants, which suggests that duplication of tumor suppressor genes may play a more general role in mediating Peto's Paradox by increasing cancer resistance in large, long-lived species.
-  By using a Reciprocal Best-Hit BLAT search approach, we investigated copy numbers of all protein-coding genes in _Atlantogenatan_ genomes to see if there is any correlation between the copy number of duplicates and changes body size along the phylogenetic tree.
-From an initial set of 18,011 protein-coding genes in hg38, we identified a median of 13,880 genes in _Atlantogenatan_ genomes, of which a median of 940 genes are duplicated.
-We find that, just as body size fluctuates throughout _Atlantogenata_, genes involved in tumor suppressor pathways are also duplicated throughout the phylogenetic tree.
-Extant species of elephants, however, show active transcription of both canonical and duplicated copies of tummor suppressors that duplicated prior to and during their sudden increase in body size, suggesting that the duplication of tumor suppressor genes facilitates the evolution of increased body size by compensating for the increased cancer risk.
-
-author_summary: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
-bibliography: rbhb_paenungulates.bib
+- address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
+  code: "University of Chicago"
+- address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
+  code: "SUNY Buffalo"
 output:
-  workflowr::wflow_html: default
-  rticles::plos_article: 
+  rticles::elsevier_article:
     includes:
       in_header: preamble.tex
+  pdf_document: default
   word_document: default
+  workflowr::wflow_html: default
 csl: plos.csl
@@
   chunk_output_type: console
+bibliography: rbhb_paenungulates.bib
+abstract: "Cancer is an intrinsic disease of multicellular organisms. Within a species,
+  the size of an animal, - correlated with the individual's number of cells - and
+  its lifespan - correlated with increasing cellular damage over time - positively
+  correlate with the risk any individual has to form tumors. Between species, however,
+  we do not observe any correlation between size, lifespan, and cancer, a phenomenon
+  that referred to as _Peto's Paradox_. Elephants are a particularly interesting member
+  of this class of paradoxical animals, since they are a set of large species deeply
+  nested in a clade of smaller species, indicating a recent gain of size. Recent work
+  has identified several individual cases of gene duplicates contributing to the increased
+  cancer resistance of elephants, which suggests that duplication of tumor suppressor
+  genes may play a more general role in mediating Peto's Paradox by increasing cancer
+  resistance in large, long-lived species. \nBy using a Reciprocal Best-Hit BLAT search
+  approach, we investigated copy numbers of all protein-coding genes in _Atlantogenatan_
+  genomes to see if there is any correlation between the copy number of duplicates
+  and changes body size along the phylogenetic tree. From an initial set of 18,011
+  protein-coding genes in hg38, we identified a median of 13,880 genes in _Atlantogenatan_
+  genomes, of which a median of 940 genes are duplicated. We find that, just as body
+  size fluctuates throughout _Atlantogenata_, genes involved in tumor suppressor pathways
+  are also duplicated throughout the phylogenetic tree. Extant species of elephants,
+  however, show active transcription of both canonical and duplicated copies of tummor
+  suppressors that duplicated prior to and during their sudden increase in body size,
+  suggesting that the duplication of tumor suppressor genes facilitates the evolution
+  of increased body size by compensating for the increased cancer risk. \n"
+
 ---
@@
     ```{r Libraries and functions, include = F}
-knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=5, 
-                      fig.retina = T, fig.align = "left")
+knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=6, 
+                      fig.retina = 2, fig.align = "left", dpi = 300, dev="Cairo_pdf")
 
@@
 ```
-```{r Data Files, include = F, cache=T}
+```{r Data Files, include = F}
 UP000005640 <- read_tsv("../data/input/UP000005640.withheader.tsv")
@@
   ) +
-  geom_pointrange(size=1) +
-  geom_point(aes(color=Median), size=3) +
-  geom_point(aes(x=`95%CI_Low`, color=`95%CI_Low`), size=5, stroke=5, pch="[") +
-  geom_point(aes(x=`95%CI_High`, color=`95%CI_High`), size=5, stroke=5, pch="]") +
-  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))")) +
+  geom_pointrange(size=0.5) +
+  geom_point(aes(color=Median), size=1) +
+  geom_point(aes(x=`95%CI_Low`, color=`95%CI_Low`), size=1.5, stroke=2, pch="[") +
+  geom_point(aes(x=`95%CI_High`, color=`95%CI_High`), size=1.5, stroke=2, pch="]") +
+  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), 
+                        guide = guide_colorbar(
+                          title="Body size (ln(g))",
+                          title.theme =  element_text(size = 6, face = "bold", colour = "black"),
+                          title.position = "top",
+                          title.hjust = 0.5
+                          )
+                        ) +
   labs(
     x="Body size (95%CI)",
-    y="Species/Ancestor"
+    y="Node"
   ) + 
   theme_pubclean()+
-  labs_pubr()
+  labs_pubr() + 
+  theme(
+    text = element_text(face = "plain", 
+                        colour = "black", lineheight = 0.9, 
+                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                        debug = FALSE), 
+    axis.text = element_text(size = 4, 
+                               colour = "black", face = "bold"),
+    axis.title = element_text(size = 6, 
+                              colour = "black", face = "bold"),
+    plot.title = element_text(size = 6, 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.text = element_text(size = 4, 
+                               face = "plain", colour = "black"),
+    legend.position = "bottom"
+  )
 
@@
                         debug = FALSE), 
-    axis.text.x = element_text(size = rel(0.86), 
+    axis.text.x = element_text(size = 4, 
                                colour = "black", face = "bold"),
-    axis.title = element_text(size = rel(1), 
+    axis.title = element_text(size = 6, 
                               colour = "black", face = "bold"),
-    plot.title = element_text(size = rel(1), 
+    plot.title = element_text(size = 6, 
                               colour = "black", lineheight = 1, face = "bold"), 
-    legend.title = element_text(size = rel(0.7), 
+    legend.title = element_text(size = 6, 
                                 face = "bold", colour = "black"), 
-    legend.text = element_text(size = rel(0.7), 
+    legend.text = element_text(size = 6, 
                                face = "plain", colour = "black")
-  )
-
-p.tree.atlantogenata.circ <- bsize.Atlantogenata %>% 
-  ggtree(
-    aes(color=lnSize), 
-    layout="circular",
-    continuous=T
-    ) + 
-  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))"))
-
-# warning("TODO: de-comment geom_tiplab in p.tree.atlantogenata! Commented currently for offline usage!")
+  ) + 
+  guides(color=guide_none())
+
 p.tree.atlantogenata <-  
@@
     aes(color=lnSize),
-    continuous=T,
-    size = 1.5
+    continuous=T
   ) + 
@@
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidea, label="Proboscidea", 
-                  fontsize = 4, align = T, offset = 50, angle=0, barsize = 1, 
+                  fontsize = 2, align = T, offset = 50, angle=0, barsize = 1,
                   offset.text = 20,
@@
                   ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="Tethytheria", 
-                  fontsize = 4, align = T, offset = 25, angle=0, barsize = 1, 
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="\nTethytheria", 
+                  fontsize = 2, align = T, offset = 25, angle=0, barsize = 1, 
                   offset.text = 20, color = bsize.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(lnSize) %>% get.size.color
                   ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="Paenungulata", 
-                  fontsize = 4, align = T, angle=0, barsize = 1, offset.text = 20,
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="\n\nPaenungulata", 
+                  fontsize = 2, align = T, angle=0, barsize = 1, offset.text = 20,
                   color = bsize.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(lnSize) %>% get.size.color
@@
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosoricida, label="Afrosoricida",
-                  fontsize = 4, align = T, angle=0, barsize = 1,
+                  fontsize = 2, align = T, angle=0, barsize = 1,
                   color=bsize.Atlantogenata.table %>% filter(label=="Afrosoricida") %>% pull(lnSize) %>% get.size.color
@@
                     label="Macroscelidae", 
-                    fontsize = 4, align = T, angle=0, barsize = 1,
+                    fontsize = 2, align = T, angle=0, barsize = 1,
                     color=bsize.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(lnSize) %>% get.size.color
@@
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Xenarthra, label="Xenarthra", 
-                  fontsize = 4, align = T, angle=0, barsize = 1,
+                  fontsize = 2, align = T, angle=0, barsize = 1,
                   color=bsize.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(lnSize) %>% get.size.color
@@
                         guide = guide_colourbar(title="Body size (ln(g))"))+
-  xlim(0,1500) + 
+  xlim(0,1750) + 
   theme_tree2() +
@@
                         debug = FALSE), 
-    axis.text.x = element_text(size = rel(0.86), 
+    axis.text.x = element_text(size = 4, 
                                colour = "black", face = "bold"),
-    axis.title = element_text(size = rel(1), 
+    axis.title = element_text(size = 6, 
                               colour = "black", face = "bold"),
-    plot.title = element_text(size = rel(1), 
+    plot.title = element_text(size = 6, 
                               colour = "black", lineheight = 1, face = "bold"), 
-    legend.title = element_text(size = rel(0.7), 
+    legend.title = element_text(size = 6, 
                                 face = "bold", colour = "black"), 
-    legend.text = element_text(size = rel(0.7), 
+    legend.text = element_text(size = 6, 
                                face = "plain", colour = "black")
-  )
+  ) + 
+  guides(color=guide_none())
 
@@
 # p.ml.tree.increasing.geneCounts.withSize.withBL
-
 ```
@@
 If all cells in all organisms have a similar risk of malignant transformation and equivalent cancer suppression mechanisms, organism with many cells should have a higher prevalence of cancer than organisms with fewer cells.
-Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, human cancer incidence increases with increasing adult height [@Green2011; @Nunney:20181c2] and cancer incidence is positively correlated with body size in dogs [@Dobson2013; @Beuchat2015].
+Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, human cancer incidence increases with increasing adult height [@Green2011; @Nunney:20181c2] and cancer incidence is positively correlated with body size in dogs [@Dobson2013; @Dorn1968].
 There is no correlation, however, between body size and cancer risk between species.
-This lack of correlation is often referred to as ‘Peto’s Paradox’ [@CaulinAndMaley2011; @Leroi2003; @Peto1975].
+This lack of correlation is often referred to as ‘Peto’s Paradox’ [@Caulin2011; @Leroi2003; @Peto1975].
 While it is clear that a resolution to Peto’s Paradox must involve the evolution of enhanced cancer protection alongside increases in body size and lifespan, the specific genetic, molecular, and cellular mechanisms that underlie this resistance have proven elusive.
 [c.f.
-@Ashur-Fabian2004; @Seluanov2008; @Gorbunova2012; @Tian2013; @Sulak2016].
+@Ashur-Fabian2004; @Seluanov2008; @Gorbunova2012; @Tian2013; @Sulak2015].
 
@@
 Similarly African (4,800kg) and Asian elephants (3,200kg) are the largest living elephant species, but are dwarfed by the truly gigantic extinct Proboscideans such as Deinotherium (~132,000kg), Mammut borsoni (110,000kg), and the Asian straight-tusked elephant (~220,000kg), the largest known land mammal [@Larramendi:20151c2].
-Remarkably these large-bodied Afrotherian lineages are nested within small bodied species (Fig.
+These large-bodied Afrotherian lineages are nested within small bodied species (Fig.
 1) [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], indicating that gigantism independently evolved in hyraxes, sea cows, and elephants (Paenungulates).
@@
 
-Although many mechanisms can potentially resolve Peto’s paradox, the most parsimonious route to enhanced cancer resistance is likely through an increased copy number of tumor suppressors.
-Such an example has been seen in the case of candidate genes such as _TP53_ and _LIF_ [@Abegglen:JAMA2015; @Sulak2016; @Vazquez2018] as well as in studies involving a limited set of candidate genes [@Caulin2015; @Doherty2016].
+Although many mechanisms can potentially resolve Peto’s paradox, among the most parsimonious routes to enhanced cancer resistance is through an increased copy number of tumor suppressors.
+Such an example has been seen in the case of candidate genes such as _TP53_ and _LIF_ [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018] as well as in studies involving a limited set of candidate genes [@Caulin2015; @Doherty2016].
 As these studies focus on _a priori_ gene sets, however, it remains unknown whether this is a general, genome-wide trend in Afrotherian genomes; and whether such a general trend is associated with the recent increases in body size – and therefore expected cancer risk – in these species.
@@
 Furthermore, we see that the ancestral body size increases in Hydracoidia and Sirenia were independent events.
-To study the evolution of gene copy number, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene duplications in Afrotherian genomes, and used parsimony to infer the lineages in which those duplications occurred.
+To study the evolution of gene copy number, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene duplications in Afrotherian genomes, and used maximum likelihood (treating copy number as a discrete trait) to infer the lineages in which those duplications occurred.
 We found gene duplications in lineages with increased body mass were enriched in functions related to tumor suppression, including regulation of the cell cycle, DNA damage repair, and regulation of apoptosis.
@@
 
-We built a time-calibrated supertree of Eutherian mammals by combining the time-calibrated molecular phylogeny of Bininda-Emonds _et al._ [-@Bininda-Emonds2008] with the time-calibrated total evidence Afrotherian phylogeny from Puttick and Thomas [@PuttickAnddThomas2015].
+We built a time-calibrated supertree of Eutherian mammals by combining the time-calibrated molecular phylogeny of Bininda-Emonds _et al._ [-@Bininda-Emonds2008] with the time-calibrated total evidence Afrotherian phylogeny from Puttick and Thomas [@PuttickAndThomas2015].
 While the Bininda-Emonds _et al._ [-@Bininda-Emonds2008] phylogeny includes 1,679 species, only 34 are Afrotherian, and no fossil data are included.
@@
 
-__Reciprocal Best-Hit BLAT:__ We developed a reciprocal best hit BLAT (RBHB) pipeline to quickly identify homologs and estimate gene copy numbers (__Figure 1A__).
+__Reciprocal Best-Hit BLAT:__ We developed a reciprocal best hit BLAT (RBHB) pipeline to identify putative homologs and estimate gene copy numbers across species (__Figure 1A__).
 The Reciprocal Best Hit (RBH) search strategy is conceptually straightforward: 1) Given a gene of interest $G_A$ in a query genome $A$, one searches a target genome $B$ for all possible matches to $G_A$; 2) For each of these hits, one then performs the reciprocal search in the original query genome to identify the highest-scoring hit; 3) A hit in genome $B$ is defined as a homolog of gene $G_A$ if and only if the original gene $G_A$ is the top reciprocal search hit in genome $A$.
@@
 ```
-__RecSearch Pipeline__ We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome.
+__RecSearch Pipeline:__ We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome.
 For the query sequences in our search, we used the hg38 Proteome provided by UniProt [Accession #UP000005640; @uniprot], which is a comprehensive set of protein sequences curated from a combination of predicted and validated protein sequences generated by the UniProt Consortium.
@@
 LINC genes); poorly-studied genes from predicted open reading frames (C-ORFs); and sequences with highly repetitive sequences such as zinc fingers, protocadherins, and transposon-containing genes, as these were prone to high levels of false positive hits.
-  After filtering out problematic protein queries, we then used our pipeline (Figure 1A) to search for all copies of our `r n.GenesSearched` query genes in publicly available Afrotherian genomes, including African savannah elephant (_Loxodonta africana_: loxAfr3, loxAfr4, loxAfrC), African forest elephant (_Loxodonta cyclotis_: loxCycF), Asian Elephant (_Elephas maximus_: eleMaxD), Woolly Mammoth (_Mammuthus primigenius_: mamPriV), Colombian mammoth (_Mammuthus columbi_: mamColU), American mastodon (_Mammut americanum_: mamAmeI), Rock Hyrax (_Procavia capensis_: proCap1, proCap2, proCap2_HiC), West Indian Manatee (_Trichechus manatus latirostris_: triManLat1, triManLat1_HiC), Aardvark (_Orycteropus afer_: oryAfe1, oryAfe1_HiC), Lesser Hedgehog Tenrec (_Echinops telfairi_: echTel2), Nine-banded armadillo (_Dasypus novemcinctus_: dasNov3), Hoffman's two-toed sloth (_Choloepus hoffmannii_: choHof1, choHof2, choHof2_HiC), Cape golden mole (_Chrysochloris asiatica_: chrAsi1), and Cape elephant shrew (_Elephantulus edwardii_: eleEdw1).
+
+After filtering out problematic protein queries (see "Query gene inclusion criteria"), we then used our pipeline (Figure 1A) to search for all copies of our `r n.GenesSearched` query genes in publicly available Afrotherian genomes, including African savannah elephant (_Loxodonta africana_: loxAfr3, loxAfr4, loxAfrC), African forest elephant (_Loxodonta cyclotis_: loxCycF), Asian Elephant (_Elephas maximus_: eleMaxD), Woolly Mammoth (_Mammuthus primigenius_: mamPriV), Colombian mammoth (_Mammuthus columbi_: mamColU), American mastodon (_Mammut americanum_: mamAmeI), Rock Hyrax (_Procavia capensis_: proCap1, proCap2, proCap2_HiC), West Indian Manatee (_Trichechus manatus latirostris_: triManLat1, triManLat1_HiC), Aardvark (_Orycteropus afer_: oryAfe1, oryAfe1_HiC), Lesser Hedgehog Tenrec (_Echinops telfairi_: echTel2), Nine-banded armadillo (_Dasypus novemcinctus_: dasNov3), Hoffman's two-toed sloth (_Choloepus hoffmannii_: choHof1, choHof2, choHof2_HiC), Cape golden mole (_Chrysochloris asiatica_: chrAsi1), and Cape elephant shrew (_Elephantulus edwardii_: eleEdw1).
 For many of these species, we covered multiple assemblies in order to test the effects of assembly size and quality on our hits.
 
-__Query protein criteria__ To assemble our query list, we first removed all unnamed genes from UP000005640.
+__Query gene inclusion criteria:__ To assemble our query list, we first removed all unnamed genes from UP000005640.
 Next, we excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (`r query.binTable %>% filter(C.ORF) %>% nrow`), LOC (`r query.binTable %>% filter(LOC) %>% nrow`), HLA genes (`r query.binTable %>% filter(HLA) %>% nrow`), replication dependent histones (`r query.binTable %>% filter(Histone) %>% nrow`), odorant receptors (`r query.binTable %>% filter(OlfactoryReceptor) %>% nrow`), ribosomal proteins (`r query.binTable %>% filter(Riboprotein) %>% nrow`), zinc finger transcription factors (`r query.binTable %>% filter(ZincFinger) %>% nrow`), viral and repetitive-element-associated proteins (`r query.binTable %>% filter(Viral) %>% nrow`) and any protein described as either "Uncharacterized," "Putative," or "Fragment" by UniProt in UP000005640 (`r query.binTable %>% filter(Fragment | Uncharacterized | Putative) %>% nrow`), leaving us with a final set of `r n.TranscriptsSearched` query protein sequences, corresponding to `r n.GenesSearched` genes.
 
-__Duplication gene inclusion criteria__ In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call.
+__Duplication gene inclusion criteria:__ In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call.
 The resulting gene-level copy number table was then combined with the maximum ECNC values observed for each gene in order to call gene duplications.
-We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes, therefore partial gene duplications were included provided they passed additional inclusion criteria.
+We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes [@Sulak2015; @Vazquez2018], therefore partial gene duplications were included provided they passed additional inclusion criteria.
 The ECNC cut off of 1.5 was selected empirically, as this value minimized the number of false positives seen in a test set of genes and genomes.
@@
 
-__Genome Quality Assessment using CEGMA__ In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome [@gVolante; @CEGMA].
+__Genome Quality Assessment using CEGMA:__ In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome [@gVolante; @CEGMA].
 CEGMA was run using the default settings for mammals ("Cut-off length for sequence statistics and composition" = 1;"CEGMA max intron length" = 100000; "CEGMA gene flanks" = 10000, "Selected reference gene set" = CVG).
@@
 
-To validate and filter out RBHB results, we intersected our results with either gene prediction or transcriptomic evidence as a proxy for functionality.
-
-__Transcriptome Assembly:__ For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated _de novo_ transcriptomes using publically-available RNA-sequencing data from NCBI SRA.
-We mapped reads to all genomes available for each species, and assembled transcripts using HISAT2 and StringTie, respectively [@HISAT2; @StringTie; @Tuxedo].
+To validate and filter out duplicate gene calls, we intersected our results with either gene prediction or transcriptomic evidence as a proxy for functionality.
+
+__Transcriptome Assembly:__ For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated _de novo_ transcriptomes using publically-available RNA-sequencing data from NCBI SRA (Supplementary Table 1).
+We mapped reads to all genomes available for each species, and assembled transcripts using HISAT2 and StringTie, respectively [@HISAT; @StringTie; @Tuxedo].
 RNA-sequencing data was not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax, Aardvark, or the Lesser Hedgehog Tenrec.
@@
 
-__Evidenced Duplicate Criteria:__ We intersected our records of duplicate hits identified in each genome with the gene prediction tracks and/or transcriptome assemblies using `bedtools` [@bedtools].
+__Evidenced Duplicate Criteria:__ We intersected our records of duplicate hits identified in each genome with the gene prediction tracks and/or transcriptome assemblies using `bedtools`.
 When multiple lines of evidence for functionality were present for a genome, we used the union of all intersections as the final output for evidenced duplicates.
@@
 We implemented a maximum likelihood method for determining the ancestral copy numbers of genes in _Atlantogenata_ using IQ-Tree.
-For this analysis, we used an unrooted subset of our prior species tree, including only the aforementioned _Atlantogenata_ species.
+For this analysis, we used an unrooted species tree including only the aforementioned _Atlantogenata_ species.
 We generated PHYLIP files containing the copy number of each gene in the highest quality genome for each species, encoding genes on a scale from 1-31+ copies as 1-9, A-V; and encoding a gene's copy number as uncetain ("?") when we did not identify it in the genome.
 We used the included tree-searching and model-testing functionality in IQ-Tree to determine the most likely topology for the species tree, and to obtain the most likely model for copy number changes in the genome.
+The most likely model for the evolution of copy number was inferred to be a Jukes-Cantor type model for morphological data with equal state frequencies, allowing for a proportion of invariable sites, and using a discrete Gamma model with 4 rate categories.
+("MK+FQ+I+G4").
 We defined the ancestral state of a node if it had greater than an 80% posterior probability.
@@
 
-To determine which pathways were associated with duplicated genes in each species and lineage, we used WEBGESTALT to perform overrepresentation analysis (ORA) of the duplicated gene lists relative to our initial query gene list [@webgestalt2019].
-For the database of pathways used in the analysis, we used Reactome [@Reactome], Wikipathways, and Wikipathways_cancer [@Wikipathways], and KEGG [@KEGG].
+To determine which pathways were associated with duplicated genes in each species and lineage, we used WEBGESTALT to perform overrepresentation analysis (ORA) of the duplicated gene lists relative to our initial query gene list [@WebGestalt2019].
+For the database of pathways used in the analysis, we used Reactome [@Reactome], Wikipathways, [@Wikipathways], and KEGG [@KEGG].
 For the ORA, we used FDR for determining significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and FDR=0.5.
@@
 
-Next, we used a simplified multistage cancer risk model for body size $D$ and lifespan $t$: $K \approx Dt^6$ [@ArmitageAndDoll; @Peto2015; @Peto2015b].
+Next, we used a simplified multistage cancer risk model for body size $D$ and lifespan $t$: $K \approx Dt^6$ [@Armitage:19851c2; @Armitage:20041c2; @Peto:20151c2; @Peto1975].
 To determine the change in $K$ between nodes, we obtained the ratio between the cancer risk $K_{1}$ at any given node, and the cancer risk $K_{2}$ at its ancestral node, using the equation $\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}$.
@@
 
-```{r Figure 1, fig.cap="Figure 1: Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
-__A)__ Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch.
+```{r Figure1, fig.cap="Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
+**A)** Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch.
 Atlantogenata is highlighted at the bottom.
-__B)__ Zoom-in of (A) on Atlantogenata.
-Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size"}
+**B)** Zoom-in of (**A**) on Atlantogenata.
+Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size.
+**C)** Confidence interval plot for representative species and ancestral nodes.", out.width="6in", fig.asp=1}
 f1a <- p.tree
@@
   labels="AUTO",
-  legend = "none",
-  widths = c(1,2)
+  widths = c(1,2),
+  font.label = list(size=6, face="bold")
   )
@@
   nrow=2, ncol=1,
-  labels = c("", "C")
+  labels = c("", "C"),
+  font.label = list(size=6, face="bold")
 )
 f1
-```
-```{r Table 1, fig.cap="Table 1: Body Size and Confidence Intervals in _Atlantogenata_ estimated using StableTraits."}
-t.bodysize.atlantogenata.withNodes
 ```
@@
 
-Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Orycteropus.afer")` in body mass in the stem-lineage of _Pseudoungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Orycteropus.afer")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Procavia.capensis")` in body mass in the stem-lineage of _Paenungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Procavia.capensis")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Trichechus.manatus")` in body mass in the stem-lineage of _Tehthytheria_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Trichechus.manatus")`), and a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Mammut.americanum")` __(Vinnie: "14× increase", note I tried to divide the MRCA of LoxAfer with Phosphatherium, Moeritherium, Omanitherium, Phiomia, and Mammut by their respective ancestral nodes, but never got anything that was a 14x increase...)__ in body mass in the stem-lineage of _Proboscidea_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Mammut.americanum")`) __(Vinnie:"(4,100kg)"; I get 4114kg; can't find an MRCA with that value)__ (Figure 1B,C).
-The ancestral _Hyracoidea_ was inferred to be relatively small `r text.bodysize.ancNode.byMRCA.CI("Microhyrax.lavocati", "Procavia.capensis")` __(Vinnie wrote (2.86-15.71kg); I get an upper range of 118.18kg)__, and rate accelerations were coincident with independent body mass increases in large hyraxes such as _Titanohyrax andrewsi_ `r text.bodysize.byLabel("Titanohyrax.andrewsi")`; `r text.ratio.nodeByMRCA.ancNode(nodeA="Titanohyrax.andrewsi","Orycteropus.afer")`.
-While the body mass of the ancestral _Sirenian_ was inferred to be large `r text.bodysize.ancNode.byMRCA.CI("Trichechus.manatus", "Prorastomus.sirenoides")` _(Vinnie: (61-656kg); I get an upper range of 956kg)__, a rate acceleration occurred coincident with a `r text.ratio.node.ancnode.atLabel("Hydrodamalis.gigas")` in body mass in Stellar's sea cow.
+Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Orycteropus.afer")` in body mass in the stem-lineage of _Pseudoungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Orycteropus.afer")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Procavia.capensis")` in body mass in the stem-lineage of _Paenungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Procavia.capensis")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Trichechus.manatus")` in body mass in the stem-lineage of _Tehthytheria_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Trichechus.manatus")`), and a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Mammut.americanum")` in body mass in the stem-lineage of _Proboscidea_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Mammut.americanum")`) (Figure 1B,C).
+The ancestral _Hyracoidea_ was inferred to be relatively small `r text.bodysize.ancNode.byMRCA.CI("Microhyrax.lavocati", "Procavia.capensis")`, and rate accelerations were coincident with independent body mass increases in large hyraxes such as _Titanohyrax andrewsi_ `r text.bodysize.byLabel("Titanohyrax.andrewsi")`; `r text.ratio.nodeByMRCA.ancNode(nodeA="Titanohyrax.andrewsi","Orycteropus.afer")`.
+While the body mass of the ancestral _Sirenian_ was inferred to be large `r text.bodysize.ancNode.byMRCA.CI("Trichechus.manatus", "Prorastomus.sirenoides")`, a rate acceleration occurred coincident with a `r text.ratio.node.ancnode.atLabel("Hydrodamalis.gigas")` in body mass in Stellar's sea cow.
 Rate accelerations also occurred coincident with `r text.ratio.nodeByMRCA.ancNode("Elephas.antiquus falconeri", "Elephas.cypriotes")` in body mass in the stem-lineage of the dwarf elephants _Elephas (Palaeoloxodon) antiquus falconeri_ and _Elephas cypriotes_.
@@
 
-## Pervasive duplication of tumor suppressors during the origins of large bodied Afrotheirans
-```{r ECNC DIAGRAMS, include=F}
-ecnc.df1 <- tibble(
-  hit = as.character(1) %>% paste("Hit", ., sep=" ") %>% as.factor(),
-  x1=c(0),
-  x2=c(100),
-  y1=c(0),
-  y2 = c(1)
-)
-ecnc.df2 <- tibble(
-  hit = as.character(1:2) %>% paste("Hit", ., sep=" ") %>% as.factor(),
-  x1=c(0, 0),
-  x2=c(100, 100),
-  y1=c(0, 1),
-  y2 = c(1, 2)
-)
-ecnc.df3 <- tibble(
-  hit = as.character(1:3) %>% paste("Hit", ., sep=" ") %>% as.factor(),
-  x1=c(0,25,75),
-  x2=c(25,75,100),
-  y1=c(0, 0, 0),
-  y2 = c(1, 1, 1)
-)
-ecnc.df4 <- tibble(
-  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
-  x1=c(0,25,75, 0),
-  x2=c(25,75,100, 100),
-  y1=c(0, 0, 0, 1),
-  y2 = c(1, 1, 1, 2)
-)
-
-ecnc.df5 <- tibble(
-  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
-  x1=c(0,0,0, 0),
-  x2=c(100,15,15, 15),
-  y1=c(0, 1, 2, 3),
-  y2 = c(1, 2, 3, 4)
-)
-
-ecnc.df6 <- tibble(
-  hit = as.character(c(1,2,1,3, 3)) %>% paste("Hit", ., sep=" ") %>% as.factor(),
-  label = factor(c("Hit 1", "Hit 2", "Unidentified", "Unidentified", "Hit 3")),
-  x1=c(0, 0, 75, 75, 50),
-  x2=c(75, 50, 100, 100, 75),
-  y1=c(0, 1, 0, 1, 1),
-  y2 = c(1, 2, 1, 2, 2),
-  transparent = c(0,0,1,1,0) %>% as.factor
-)
-
-ecnc.onegene.onehit <- 
-  ecnc.df1 %>% 
-  ggplot(
-    aes(
-      xmin = x1,
-      xmax = x2,
-      ymin = y1,
-      ymax = y2,
-      fill = hit
-    )
-  ) + 
-  geom_rect(color="black") + 
-  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
-  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
-  scale_fill_brewer(palette = "Dark2", guide=F) + 
-  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
-  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
-  labs(
-    x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: One Contiguous Gene, One Hit"
-  )
-
-ecnc.twogene.twohit <- 
-  ecnc.df2 %>% 
-  ggplot(
-    aes(
-      xmin = x1,
-      xmax = x2,
-      ymin = y1,
-      ymax = y2,
-      fill = hit
-    )
-  ) + 
-  geom_rect(color="black") + 
-  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
-  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
-  scale_fill_brewer(palette = "Dark2", guide=F) + 
-  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
-  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
-  labs(
-    x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: Two Contiguous Genes, Two Hits"
-  )
-
-ecnc.onegene.threehits <-
-  ecnc.df3 %>% 
-  ggplot(
-    aes(
-      xmin = x1,
-      xmax = x2,
-      ymin = y1,
-      ymax = y2,
-      fill = hit
-    )
-  ) + 
-  geom_rect(color="black") + 
-  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
-  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
-  scale_fill_brewer(palette = "Dark2", guide=F) + 
-  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
-  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
-  labs(
-    x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: One Fragmented Gene, Multiple Hits"
-  )
-
-ecnc.twogene.fourhits <-
-  ecnc.df4 %>% 
-  ggplot(
-    aes(
-      xmin = x1,
-      xmax = x2,
-      ymin = y1,
-      ymax = y2,
-      fill = hit
-    )
-  ) + 
-  geom_rect(color="black") + 
-  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
-  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
-  scale_fill_brewer(palette = "Dark2", guide=F) + 
-  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
-  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
-  labs(
-    x = "Query Length",
-    y = "Query Coverage"#,
-    #title = "ECNC: Two Genes, Multiple Hits"
-  )
-
-ecnc.onegene.manypartialhits <-
-  ecnc.df5 %>% 
-  ggplot(
-    aes(
-      xmin = x1,
-      xmax = x2,
-      ymin = y1,
-      ymax = y2,
-      fill = hit
-    )
-  ) + 
-  geom_rect(color="black") + 
-  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
-  geom_hline(yintercept = 1.45, color="red", lty="dashed", size=2) + 
-  scale_fill_brewer(palette = "Dark2", guide=F) + 
-  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
-  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
-  labs(
-    x = "Query Length",
-    y = "Query Coverage"
-    # title = "ECNC: One Real Gene, Multiple Hits on Conserved Domain"
-  )
-
-ecnc.twogene.partialhits <-
-  ecnc.df6 %>% 
-  ggplot(
-    aes(
-      xmin = x1,
-      xmax = x2,
-      ymin = y1,
-      ymax = y2,
-      fill = hit,
-      alpha = transparent,
-      linetype = transparent
-    )
-  ) + 
-  geom_rect(color="black") + 
-  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=label), color="black", inherit.aes = F)+
-  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
-  scale_fill_brewer(palette = "Dark2", guide=F) + 
-  scale_alpha_manual(values = c("0"=1,"1"=0.5), guide=F) + 
-  scale_linetype_manual(values = c("0"="solid","1"="dotted"), guide=F) + 
-  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
-  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
-  labs(
-    x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: Two Real Genes, Unidentified Tails"
-  )
-
-```
-```{r Correlations between genome quality scores (1)}
-CEGMA.completeness.clean <-
-  CEGMA.completeness %>%
-  filter(str_detect(Group, "_all")) %>%
-  select(-`Missing Proteins`) %>%
-  rename(
-    n.Prots = `#Prots`,
-    pc.Completeness = `%Completeness`,
-    n.Total = `#Total`,
-    pc.Ortho = `%Ortho`
-  ) %>%
-  reshape2::melt(., id = c("Genome", "Group")) %>%
-  mutate(Group = Group %>% str_remove("_all")) %>%
-  unite("variable", variable, Group) %>%
-  pivot_wider(id_cols = Genome, names_from = variable, values_from = value)
-
-CEGMA <- full_join(CEGMA.qc, CEGMA.completeness.clean, by="Genome")
-
-GCT.CEGMA <- genomeStatsRBHB %>%
-  select(-starts_with("q"), -starts_with("Median")) %>% 
-  inner_join(., CEGMA, by="Genome") %>%
-  mutate_if(is.character, as.factor)
-
-cor.GCT.CEGMA <- GCT.CEGMA %>%
-  mutate_if(is.factor, as.numeric) %>%
-  as.matrix %>%
-  cor %>%
-  reorder_cormat %>%
-  # get_upper_tri %>%
-  get_lower_tri %>%
-  reshape2::melt(., na.rm=T)
-
-p.cor.GCT.CEGMA <-
-  cor.GCT.CEGMA %>%
-  ggplot(
-    aes(x=Var1, y=Var2, fill = value)
-  )+
-  geom_tile(color = "white")+
-  theme_minimal()+
-  coord_fixed()+
-  scale_x_discrete(
-    position = "top"
-    ) +
-  scale_y_discrete(position = "right") +
-  scale_fill_viridis_c(breaks = c(-1,-0.5, 0, 0.5, 1),labels = c(-1,-0.5, 0, 0.5, 1)) +
-  theme(
-    axis.text.x = element_text(
-      angle = -330, vjust = 0,
-      size = 9, hjust = 0
-    ),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank(),
-    panel.grid.major = element_blank(),
-    panel.border = element_blank(),
-    panel.background = element_blank(),
-    axis.ticks = element_blank(),
-    legend.justification = c(1, 1),
-    legend.position = c(0.4,0.7), #c(0.6, 0.7),
-    legend.direction = "horizontal",
-    legend.text = element_text(size=rel(1)),
-    legend.title =  element_text(size=rel(1.2))
-    ) +
-  guides(
-    fill = guide_colorbar(
-      title="Pearson\nCorrelation",
-      barwidth = 9, barheight = 3,
-      title.position = "top", title.hjust = 0.5,
-      draw.llim = T
-    )
-  )
-```
-```{r Phylogenetic Correlations}
-
-to_drop <-tree.Atlantogenata %>% as_tibble %>% filter(!BestGenome %in% (GCT.CEGMA %>% pull(Genome))) %>% pull(label)
-
-tree.GCT.CEGMA <- full_join(
-  tree.Atlantogenata %>% 
-    drop.tip(to_drop) %>% 
-    as_tibble %>% 
-    mutate(Genome = BestGenome),
-  GCT.CEGMA, 
-  by="Genome"
-) %>% 
-  as.treedata()
-GCT.CEGMA.wGenome <- tree.GCT.CEGMA %>% as_tibble
-
-# GCT.CEGMA.wGenome %<>% filter(Gene == "TP53")
-# rown <- GCT.CEGMA.wGenome$label %>% unique
-# GCT.CEGMA.wGenome %<>% mutate_if(is.character, . %>% as.factor %>%  as.numeric)%>%
-#   as.data.frame
-# rownames(GCT.CEGMA.wGenome) <- rown
-```
-```{r Phylogenetic Correlations: model building (1), warning=T, eval=F}
-training.table.tree.GCT.CEGMA <- tree.GCT.CEGMA %>%
-  as_tibble() %>%
-  filter(!is.na(mean.CN)) %>%
-  mutate_if(is.character, . %>% as.factor) %>%
-  mutate(label=as.character(label) %>% str_replace_all("\\.", " ")) %>%
-  mutate_if(is.factor, as.numeric) %>%
-  as.data.frame %>%
-  column_to_rownames("label")
-
-names(training.table.tree.GCT.CEGMA) %<>% make.names(unique = T) %>% str_replace_all("\\.+", ".")
-
-phyloGCT.CEGMA <- tree.GCT.CEGMA@phylo %>% as.phylo()
-
-bm.gct <-corBrownian(1, phyloGCT.CEGMA)
-pl.gct <-corPagel(1, phyloGCT.CEGMA)
-
-models.vars <- training.table.tree.GCT.CEGMA %>% 
-  select(
-    # -mean.Copy, 
-    # -mean.ECNC, 
-    # -mean.CN,
-    # -sd.ECNC, 
-    # -sd.Copy, 
-    # -sd.CN,
-    -X.of.sequences.containing.non.ACGTN.nt., -Base.composition.Other., -parent, -node, 
-    -Brownian, -Median, -X95.CI_Low, -X95.CI_High, -branch.length, -Species, 
-    -Species.strict, -Spc.short, -BestGenome, -highlight, -Base.composition.C.,
-    -Base.composition.A., -Base.composition.T., -Base.composition.G., 
-    -phyloPicUID, -Rate,-Orig_len, -Common.Name, -Longest.sequence.nt., 
-    -Shortest.sequence.nt., -Mean.sequence.length.nt., -Median.sequence.length.nt., 
-    -starts_with("X.of.sequences"), -starts_with("Sum.length")
-  )
-
-models.mean.ECNC <- names(models.vars) %>% 
-  str_remove("mean.*|sd.*") %>% 
-  stringi::stri_omit_empty() %>% 
-  paste0("mean.ECNC ~ ", .) %>% 
-  set_names(., .) %>% 
-  sapply(
-    ., 
-    . %>% 
-      paste0(
-        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
-        ) %>% 
-      parse(text = .)
-    ) %>% 
-  lapply(eval)
-
-models.mean.CN <- names(models.vars) %>% 
-  str_remove("mean.*|sd.*") %>% 
-  stringi::stri_omit_empty() %>% 
-  paste0("mean.CN ~ ", .) %>% 
-  set_names(., .) %>% 
-  sapply(
-    ., 
-    . %>% 
-      paste0(
-        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
-        ) %>% 
-      parse(text = .)
-    ) %>% 
-  lapply(eval)
-
-models.mean.Copy <-names(models.vars) %>% 
-    str_remove("mean.*|sd.*") %>% 
-    stringi::stri_omit_empty() %>% 
-    paste0("mean.Copy ~ ", .) %>% 
-  set_names(., .) %>% 
-  sapply(
-    ., 
-    . %>% 
-      paste0(
-        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
-        ) %>% 
-      parse(text = .)
-    ) %>% 
-  lapply(eval)
-
-aic.models.mean.ECNC <- AIC(models.mean.ECNC[[1]], models.mean.ECNC[[2]], models.mean.ECNC[[3]], models.mean.ECNC[[4]], models.mean.ECNC[[5]], models.mean.ECNC[[6]], models.mean.ECNC[[7]], models.mean.ECNC[[8]], models.mean.ECNC[[9]], models.mean.ECNC[[10]], models.mean.ECNC[[11]], models.mean.ECNC[[12]], models.mean.ECNC[[13]], models.mean.ECNC[[14]],
-    models.mean.ECNC[[15]], models.mean.ECNC[[16]], models.mean.ECNC[[17]], models.mean.ECNC[[18]]#,models.mean.ECNC[[19]], models.mean.ECNC[[20]], models.mean.ECNC[[21]]
-    ) %>% 
-  rownames_to_column("Model") %>% 
-  mutate(Model = names(models.mean.ECNC)) %>%
-  separate(Model, into = c("Y", "X"), sep="~") %>% 
-  arrange(AIC)
-aic.models.mean.Copy <- AIC(models.mean.Copy[[1]], models.mean.Copy[[2]], models.mean.Copy[[3]], models.mean.Copy[[4]], models.mean.Copy[[5]], models.mean.Copy[[6]], models.mean.Copy[[7]], models.mean.Copy[[8]], models.mean.Copy[[9]], models.mean.Copy[[10]], models.mean.Copy[[11]], models.mean.Copy[[12]], models.mean.Copy[[13]], models.mean.Copy[[14]],
-    models.mean.Copy[[15]], models.mean.Copy[[16]], models.mean.Copy[[17]], models.mean.Copy[[18]]#,models.mean.Copy[[19]], models.mean.Copy[[20]], models.mean.Copy[[21]]
-    ) %>% 
-  rownames_to_column("Model") %>% 
-  mutate(Model = names(models.mean.Copy)) %>%
-  separate(Model, into = c("Y", "X"), sep="~") %>% 
-  arrange(AIC)
-aic.models.mean.CN <- AIC(models.mean.CN[[1]], models.mean.CN[[2]], models.mean.CN[[3]], models.mean.CN[[4]], models.mean.CN[[5]], models.mean.CN[[6]], models.mean.CN[[7]], models.mean.CN[[8]], models.mean.CN[[9]], models.mean.CN[[10]], models.mean.CN[[11]], models.mean.CN[[12]], models.mean.CN[[13]], models.mean.CN[[14]],
-    models.mean.CN[[15]], models.mean.CN[[16]], models.mean.CN[[17]], models.mean.CN[[18]]#,models.mean.CN[[19]], models.mean.CN[[20]], models.mean.CN[[21]]
-    ) %>% 
-  rownames_to_column("Model") %>% 
-  mutate(Model = names(models.mean.CN)) %>%
-  separate(Model, into = c("Y", "X"), sep="~") %>% 
-  arrange(AIC)
-```
-```{r Phylogenetic Correlations: model building (1 + lnsize), warning=T, eval=F}
-models.mean.ECNC.lnSize <- names(models.vars) %>% 
-  str_remove("mean.*|sd.*|lnSize") %>% 
-  stringi::stri_omit_empty() %>% 
-  paste0("mean.ECNC ~ lnSize + ", .) %>% 
-  set_names(., .) %>% 
-  sapply(
-    ., 
-    . %>% 
-      paste0(
-        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
-        ) %>% 
-      parse(text = .)
-    ) %>% 
-  lapply(eval)
-
-models.mean.CN.lnSize <- names(models.vars) %>% 
-  str_remove("mean.*|sd.*|lnSize") %>% 
-  stringi::stri_omit_empty() %>% 
-  paste0("mean.CN ~ lnSize + ", .) %>% 
-  set_names(., .) %>% 
-  sapply(
-    ., 
-    . %>% 
-      paste0(
-        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
-        ) %>% 
-      parse(text = .)
-    ) %>% 
-  lapply(eval)
-
-models.mean.Copy.lnSize <- names(models.vars) %>% 
-  str_remove("mean.*|sd.*|lnSize") %>% 
-  stringi::stri_omit_empty() %>% 
-  paste0("mean.Copy ~ lnSize + ", .) %>% 
-  set_names(., .) %>% 
-  sapply(
-    ., 
-    . %>% 
-      paste0(
-        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
-        ) %>% 
-      parse(text = .)
-    ) %>% 
-  lapply(eval)
-
-aic.models.mean.ECNC.lnSize <- AIC(models.mean.ECNC.lnSize[[1]], models.mean.ECNC.lnSize[[2]], models.mean.ECNC.lnSize[[3]], models.mean.ECNC.lnSize[[4]], models.mean.ECNC.lnSize[[5]], models.mean.ECNC.lnSize[[6]], models.mean.ECNC.lnSize[[7]], models.mean.ECNC.lnSize[[8]], models.mean.ECNC.lnSize[[9]], models.mean.ECNC.lnSize[[10]], models.mean.ECNC.lnSize[[11]], models.mean.ECNC.lnSize[[12]], models.mean.ECNC.lnSize[[13]], models.mean.ECNC.lnSize[[14]],
-    models.mean.ECNC.lnSize[[15]], models.mean.ECNC.lnSize[[16]], models.mean.ECNC.lnSize[[17]]#, models.mean.ECNC.lnSize[[18]],models.mean.ECNC.lnSize[[19]], models.mean.ECNC.lnSize[[20]]#, models.mean.ECNC.lnSize[[21]]
-    ) %>% 
-  rownames_to_column("Model") %>% 
-  mutate(Model = names(models.mean.ECNC.lnSize)) %>%
-  separate(Model, into = c("Y", "X"), sep="~") %>% 
-  arrange(AIC)
-aic.models.mean.Copy.lnSize <- AIC(models.mean.Copy.lnSize[[1]], models.mean.Copy.lnSize[[2]], models.mean.Copy.lnSize[[3]], models.mean.Copy.lnSize[[4]], models.mean.Copy.lnSize[[5]], models.mean.Copy.lnSize[[6]], models.mean.Copy.lnSize[[7]], models.mean.Copy.lnSize[[8]], models.mean.Copy.lnSize[[9]], models.mean.Copy.lnSize[[10]], models.mean.Copy.lnSize[[11]], models.mean.Copy.lnSize[[12]], models.mean.Copy.lnSize[[13]], models.mean.Copy.lnSize[[14]],
-    models.mean.Copy.lnSize[[15]], models.mean.Copy.lnSize[[16]], models.mean.Copy.lnSize[[17]]#, models.mean.Copy.lnSize[[18]],models.mean.Copy.lnSize[[19]], models.mean.Copy.lnSize[[20]]#, models.mean.Copy.lnSize[[21]]
-    ) %>% 
-  rownames_to_column("Model") %>% 
-  mutate(Model = names(models.mean.Copy.lnSize)) %>%
-  separate(Model, into = c("Y", "X"), sep="~") %>% 
-  arrange(AIC)
-aic.models.mean.CN.lnSize <- AIC(models.mean.CN.lnSize[[1]], models.mean.CN.lnSize[[2]], models.mean.CN.lnSize[[3]], models.mean.CN.lnSize[[4]], models.mean.CN.lnSize[[5]], models.mean.CN.lnSize[[6]], models.mean.CN.lnSize[[7]], models.mean.CN.lnSize[[8]], models.mean.CN.lnSize[[9]], models.mean.CN.lnSize[[10]], models.mean.CN.lnSize[[11]], models.mean.CN.lnSize[[12]], models.mean.CN.lnSize[[13]], models.mean.CN.lnSize[[14]],
-    models.mean.CN.lnSize[[15]], models.mean.CN.lnSize[[16]], models.mean.CN.lnSize[[17]]#, models.mean.CN.lnSize[[18]],models.mean.CN.lnSize[[19]], models.mean.CN.lnSize[[20]]#, models.mean.CN.lnSize[[21]]
-    ) %>% 
-  rownames_to_column("Model") %>% 
-  mutate(Model = names(models.mean.CN.lnSize)) %>%
-  separate(Model, into = c("Y", "X"), sep="~") %>% 
-  arrange(AIC)
-
-```
-
-```{r Genome Table Prep, warning=T}
-warning("TODO: Label each genome with citation number")
-genome.table.forFig <- genomeTable.atlantogenata %>% 
-  left_join(anage, by="label") %>% 
-  left_join(tree.Atlantogenata.table, by="label") %>% 
-  mutate(
-    Median = exp(Median) %>% formatC(format = "g", digits = 2),
-    Lifespan = Lifespan %>% exp %>% formatC(format = "g", digits = 2) %>% as.character() %>% str_replace_na(replacement = "(Unknown)")
-  ) %>% 
-  select(
-    `Common Name` = `Common Name.x`,
-    Species = Species.strict.x,
-    # Clade = Family, 
-    `Size (g)` = Median,
-    # `Lifespan (y)`= Lifespan, 
-    Genome = BestGenome.x
-  ) %>% 
-  left_join(
-    genomeStatsRBHB %>% 
-      select(Genome, `#Hits`=Hits, `% Genes Found`=pc.Found, `#Duplicated` = Duplicated,`% Hits Duplicated` = pc.Hits.Duplicated, `Max Copy#`=max.Copy, `Max ECNC`=max.ECNC, `Mean ECNC/Hit`=mean.ECNCPerHit),
-    by="Genome"
-  ) %>% 
-  mutate(
-    `Common Name` = str_to_title(`Common Name`),
-    `% Genes Found` = `% Genes Found` %>% as.percentage,
-    `% Hits Duplicated` = `% Hits Duplicated` %>% as.percentage
-  )
-
-# #clades
-# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Elephant$|Mammoth$"),]$Clade <- "Paenungulata, Probodiscea"
-# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Mastodon"),]$Clade <- "Paenungulata, Probodiscea"
-# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "^Cape|Tenrec"),]$Clade <- "Macroscelididae/Afroinsectivora"
-# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Manatee"),]$Clade <- "Paenungulata, Tethytheria"
-# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Hyrax"),]$Clade <- "Paenungulata"
-# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Aardvark"),]$Clade <- "Pseudoungulata"
-# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Sloth|Armadillo"),]$Clade <- "Xenarthra"
-# 
-# #Lifespan caveats
-# genome.table.forFig[which(genome.table.forFig$Species == "Elephantulus edwardii"),]$`Lifespan (y)` = "~5.7*"
-# genome.table.forFig[which(genome.table.forFig$Species == "Loxodonta cyclotis"),]$`Lifespan (y)` = paste0(genome.table.forFig[which(genome.table.forFig$Species == "Loxodonta africana"),]$`Lifespan (y)`, "\u2020")
-```
-```{r Figure 2 Prep: Genome Table}
-ggt.GenomeTable.withNote <- genome.table.forFig %>% 
-  select(Species, `Common Name`, starts_with("Size"), starts_with("#"), starts_with("%"), starts_with("Mean")) %>% 
-  arrange(Species) %>% 
-  ggtexttable(
-    theme = ttheme(base_style="minimal", base_size=8)
-  ) #%>% annotate_figure(bottom = "* Observation derived from a single captive individual; real lifespan unknown.\n\u2020 Lifespan estimate derived from Loxodonta africana before the taxonomic species split.")
-```
-
-```{r Figure 2, fig.cap="Gene duplications occur readily throughout Atlantogenata", fig.height=7, fig.width=5}
-f2a <- ggt.GenomeTable.withNote
-f2b <- p.ml.tree.increasing.geneCounts.withSize
-f2.2 <- ggarrange(f2a, f2b, nrow=2, ncol=1, labels = "AUTO")
-f2.2
-```
-```{r Prep Table 2: PGLS of Copy Number ~ Genomic Stats, eval=F}
-# if(knitr::is_latex_output()){
-# t.GCT.CEGMA.gls <- quiet(stargazer(
-#   models.mean.Copy,models.mean.ECNC, models.mean.CN,
-#   type="latex",
-#   title = "Gene Copy ~ various genomic variables",
-#   style = "all"
-# )  %>% 
-#   c("\\newgeometry{margin=1cm}\n\\begin{landscape}", ., "\\end{landscape}\n\\restoregeometry") %>% 
-#   paste0(collapse="\n")
-#   )
-# } else if(knitr::is_html_output()){
-#   t.GCT.CEGMA.gls <- quiet(stargazer(
-#     c(models.mean.CN %>% unname, models.mean.Copy%>% unname, models.mean.ECNC%>% unname),
-#     type="html",
-#     title = "Gene Copy ~ various genomic variables",
-#     style = "all"
-#   ))
-# } else{
-#   t.GCT.CEGMA.gls <- quiet(stargazer(
-#     c(models.mean.CN %>% unname, models.mean.Copy%>% unname, models.mean.ECNC%>% unname),
-#     type="text",
-#     title = "Gene Copy ~ various genomic variables",
-#     style = "all"
-#   ))
-# }
-```
-
-Enhanced cancer suppression may have evolved through many mechanisms; among the most parsimonious is an increase in the copy number of genes with tumor suppressor functions.
-Previous studies focusing on candidate gene studies, for example, have identified increased copy number of the tumor suppressors _TP53_ and _LIF_ in elephants [@Abegglen2015; @Caulin2015; @Doherty2016; @Sulak2016; @Vazquez2018].
-Therefore, in order to test whether this was a pervasive phenomena genome-wide in _Afrotherians_, we used a Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in _Afrotherian_ and _Atlantogenatan_ genomes (Fig.
-2A, Supplementary Figure 1).
-Because RBHB-like approaches can over-estimate copy number when genes are fragmented  or incorrectly assembled across multiple scaffolds, we also inferred copy number using a complementary method that quantifies the ratio between observed and expected gene coverage per nucleotide (ECNC) (Supplementary Figure 1).
-By only including nucleotides from the query sequence that were observed in the target genome, we also correct for partial hits where some or all of the homologs of a gene have diverged from the human homolog.
-
-Because our sequence database included various protein transcripts for each gene, in order to obtain gene-level copy number information and elimate any many-to-one mappings of hits, we labeled each exon of every reciprocal best hit (RBH) with the gene corresponding to the query transcript and merged all overlapping exons; next, we eliminated any many-to-one exons that resulted from the previous step.
-Finally, we reassembled the gene loci based on the original transcript starts and ends, and the collapsed exon data, obtaining the full sequence of each RBH locus.
-Genes were considered to be duplicated if its copy number via RBHB was greater than or equal to 2, and the maximum ECNC among all transcripts prior to filtering was greater than or equal to 1.50.
-This cutoff of ECNC was selected to account for truncated gene duplications, which have been shown to be functional in various examples [@CITE examples of this].
-
-Once we applied these criteria to the results, we obtained the results shown in Figure 2A.
-Our approach positively identified an average of `r genomeStatsRBHB$pc.found %>% mean %>% as.percentage` of the queries we searched.
-We observed that the percentage of duplicated genes in non-Pseudoungulatan genomes was significantly higher: while these genomes had duplicates percentages ranging anywhere from `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`, outgroup species' duplication rates ranged from `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`.
-
-```{r Figure 3: Correlation}
-p.cor.GCT.CEGMA
-```
-
-Given the preliminary status of the genomes for _Chrysochloris asiatica_ and _Elephantulus edwardii_; and the fact that more revisions have been performed for Paenungulatan genomes; we hypothesized that genome quality may be confounding the results we've observed.
-We used `CEGMA` and the `gVolante` server [@CEGMA; @gVolante] to assess the quality of all our genomes, to see if any of these metrics correlated with our metrics of gene counts and copy number.
-As shown in Figure 3, mean ECNC, mean Copy Number, and mean CN (the lesser of Copy Number and ECNC per gene) correlate moderately strongly with genomic quality metrics related to the length and assembly quality, like LD50 and the number of scaffolds and contigs with a length above either 100K or 1M, supporting our hypothesis that increases in copy number correlate .
-
-In order to determine when in the phylogeny our identified duplications occured, we used maximum likelihood to reconstruct ancestral gene copy numbers across the phylogeny.
-We encoded the copy number of each gene as a discrete trait ranging from 0 ("?") to 31+, and then ran the ancestral gene copy number reconstruction using the phylogenetics suite `iqtree`, which rapidly tests and runs the best evolutionary model for a given trait of interest.
-Importantly, `iqtree` can be set to only output states where the likelihood of the state is greater than a certain threshold.
-Setting the threshold to 0.8, we obtained the results shown in
-
-Next, in order to select genes and duplicates which were likely functional, we omitted any hits that were not supported by either the gene prediction method GenScan, or by at least one transcript assembled from publically-available RNA-seq data.
-
-We describe the number of genes that increased in each lineage in _Atlantogenata_ in Figure 2.
-Among the genes that increased in copy number in the elephant lineage are TP53 and LIF, as previously described.
-Furthermore, we identify `r warning("TODO: Finish this section")`
-
-## Duplications that occured recently in Probodiscea are enriched for tumor suppressor pathways
-
-```{r ORA.dotplots}
-p.loxAfr.ORA.dotplot <- ORA.all %>% 
-    filter(Node == "Loxodonta_africana", FDR.Threshold <= 0.3, Database == "Reactome", RBHB == "RBB") %>% 
-    graph.ORA.dotplot(., plotly=knitr::is_html_output())
-
-p.loxAfr.ORA.dotplot.increased <- ORA.all %>% 
-    filter(Node == "Loxodonta_africana", FDR.Threshold <= 0.3, Database == "Reactome", RBHB == "RBB", Direction=="increased") %>% 
-    graph.ORA.dotplot(., plotly=knitr::is_html_output())
-# p.Proboscidea.ORA.dotplot <- ORA.all %>% 
-#   filter(Direction=="increased", Node == "Proboscidea", Database == "Reactome", RBHB == "RBB") %>% 
-#   graph.ORA.dotplot(., plotly=knitr::is_html_output())
-```
-
-```{r Branches which experienced an increase in size}
-
-ORA.numbers <- ORA.all  %>% filter(RBHB=="RBB") %>% group_by(Direction, Ancestor, Node, FDR.Threshold, Database) %>% summarise(N.Pathways=n_distinct(Pathway)) %>% ungroup %>% pivot_wider(names_from = FDR.Threshold, values_from = N.Pathways, names_prefix = "Pathways at FDR<") %>% mutate_if(is.numeric, .
-%>% ifelse(is.na(.), 0,.))
-
-ORA.numbers.ReactomeOnly <- ORA.all %>% filter(Database=="Reactome", RBHB=="RBB") %>% group_by(Direction, Ancestor, Node, FDR.Threshold) %>% summarise(N.Pathways=n_distinct(Pathway)) %>% ungroup %>% pivot_wider(names_from = FDR.Threshold, values_from = N.Pathways, names_prefix = "Pathways at FDR<") %>% mutate_if(is.numeric, .
-%>% ifelse(is.na(.), 0,.))
-
-noEnrichmentIncreasing <- setdiff(ORA.numbers %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers %>% filter(Direction == "increased") %>%  pull(Node)) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
-noEnrichmentIncreasing.Reactome <- setdiff(ORA.numbers.ReactomeOnly %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers.ReactomeOnly %>% filter(Direction == "increased") %>%  pull(Node)) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
-
-noEnrichment.any <- setdiff(ml.treedata.table$label %>% str_replace_all("\\.", "_"), union(ORA.numbers %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers %>% filter(Direction == "increased") %>%  pull(Node))) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
-noEnrichment.any.Reactome <- setdiff(ml.treedata.table$label %>% str_replace_all("\\.", "_"), union(ORA.numbers.ReactomeOnly %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers.ReactomeOnly %>% filter(Direction == "increased") %>%  pull(Node))) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
-
-p.ORA <- ORA.numbers.ReactomeOnly %>% 
-  filter(Direction == "increased") %>% 
-  select(-Direction) %>% 
-  ggtexttable(
-    theme = ttheme(base_style="minimal", base_size=8)
-  )
-
-```
-
-```{r Fig 3, fig.cap="Fig 3"}
-f3a <- p.ORA
-f3b <- p.loxAfr.ORA.dotplot
-f3c <- p.loxAfr.ORA.dotplot.increased
-
-f3bc <- ggarrange(f3b, f3c, nrow=1, ncol=2, common.legend = T, legend = "bottom", labels=c("B", "C"))
-f3 <- ggarrange(f3a, f3bc, nrow=2, ncol=1, labels=c("A",""))
-
-f3
-```
-`r warning("This figure needs work. Start by changing that table to a bar graph, for one....")`
-
-As previously stated, our inital hypothesis was that genes duplicated in lineages that experienced a growth in size would have an increased number of gene duplications that were enriched for membership in tumor suppressor pathways.
-Thus, we used WebGestalt and its Overrepresentation Analysis (ORA) functionality to determine what pathways were enriched in our duplicated gene sets in each branch relative to our initial query set.
-For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways_cancer databases using WebGestalt ORA.
-Going through the tree, at no FDR<0.5 is there any significant pathway representation for genes that increased in the branches leading to `r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`; furthermore, there are no significant pathway enrichments at FDR<0.5 for copy-number-stable genes for `r noEnrichment.any.Reactome[1:length(noEnrichment.any.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichment.any.Reactome[length(noEnrichment.any.Reactome)]`.
-Note that because _Xenarthra_ was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.
-
-For the other branches, the number of pathways that came up as significantly enriched at each FDR is shown in Figure 3A (Supplementary Figure 3).
-For the species with high duplication rates and lower-quality, highly-fragmented genomes, such as with _Chrysochloris asiatica_ (`r genomeStatsRBHB %>% filter(Genome == "chrAsi1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits) and _Elephantulus edwardii_ (`r genomeStatsRBHB %>% filter(Genome == "eleEdw1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits), it is unsurprising that there is an equivalently astronomical number of pathway enrichments present.
-In the case of these two species, their many pathway enrichments also span an incredible range of processes at every level of biology; this, in combination with the astronomically high number of copy numbers identified for these genes, further suggests a need for improvement and refinement in these genomes.
-In the cell cycle pathways called as significant in the genomes of these two species plus _Orycteropus afer_ and _Echinops telfairi_, we see that the member duplicated genes are from the same sets families, such as the APC subunit family; the proteosome subunit families; and the protein phosphatase 2 family, among others.
-It is highly possible that these results reflect true expansions of these gene families, especially in the higher-quality _Orycteropus afer_ genome; however, as discussed, it is also possible that it simply reflects artifcatual duplication events, and so require futher study.
-
-The pathway enrichments for the stable - yet duplicated - copy number, and increased copy number sets of genes for the African Elephant are shown in Figure 3B-C.
-In support of our original hypothesis, among the few enriched pathways in the African Elephant, we see that that two tumor suppression pathways - APC Complex-related pathways, and "TP53 Regulates Metabolic Genes"- appear not only in the case of stable genes, but also in the set of newly duplicated genes.
-The other pathways we see in the set of recently-duplicated genes include "Functionalization of Compounds" and its daughter pathway "Xenobiotics".
-Genes in these pathways serve to add functional groups to lipophylic compounds which would otherwise not be reactive in the cell, and are types of metabolic pathways.
-In the stable set we see enrichment of pathways such as "Neuronal Systems" and "Axon Guidance," which fit in well with what is known about elephant biology and evolution [@Cite probably].
-Overall in elephants, we see enrichments within duplicated genes for pathways involved in what makes an elephant an elephant - including tumor suppressor pathways.
-
-## Concerted duplication of TP53 and TP53-related genes towards _Probodiscea_
-
-```{r Fig4 prep}
-loxAfr4.geneTPM.genesToHighlight.plus
-    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
-    # "STK11", #REgulation of TP53, Paenungulata                  #STK11 is a TSG
-    # "COX5A", #TP53 transcriptional regulation, PRobodiscea      #CytoC OXidase
-    # "BRD7", # TP53 regulation, Tethytheria                      #Required for p53-dependent oncogene-induced senescence
-    # "TP53", #TP53 duplicates in Tethytheria
-    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria   #E3 Ligase, induces apoptosis
-    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants  #Induces apoptosis, duh
-    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri          #Decomposes H2O2
-    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants   #Antiapoptotic?
-    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr            #CytoC OXidase
-
-genesToHighlight.plus.TP53 <- genesToHighlight.plus$TP53
-
-loxAfr4.geneTPM.ORA.expressedDups.TP53 <- loxAfr4.geneTPM.ORA.expressedDups %>% 
-  filter(Gene %in% genesToHighlight.plus.TP53) %>% 
-  mutate(
-    Reactome.Pathway.Prime = Pathway %>% str_split(", ") %>% sapply(. %>% str_extract(., ".*[Tt]?[Pp]53.*")%>% stringi::stri_remove_empty_na() %>% paste(collapse =", ")),
-    id = str_remove(Locus, ".*_") %>% factor(., levels=sort(as.numeric(unique(.)))),
-    Gene = Gene %>% factor(., levels=genesToHighlight.plus$TP53)
-    ) %>% 
-  group_by(Locus) %>% 
-  summarize(Gene = unique(Gene), id=unique(id), TPM=max(TPM), Reactome.Pathway.Prime=unique(Reactome.Pathway.Prime)) %>% 
-  arrange(Gene, id) %>% 
-  mutate(Class = ifelse(
-    str_detect(Reactome.Pathway.Prime, "[Rr]egulation of [Tt][Pp]53")& Gene != "TP53",
-    "Regulator of TP53",
-    ifelse(
-      str_detect(Gene, "COX")& Gene != "TP53",
-      "Cytochrome C Oxidase Component",
-      ifelse(
-        str_detect(Gene, "CASP")& Gene != "TP53",
-      "Apoptosis",
-      "TP53 Pathway, Other"
-      )
-    )
-
-  )
-           )
-
-treedata.tp53.table <- ml.treedata.table %>% 
-  full_join(., bodysize.tree.atlantogenata, by="label") %>%
-  select(parent, node, branch.length, label, lnSize, all_of(loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% pull(Gene) %>% unique)) %>% 
-  mutate_at(vars(loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% pull(Gene) %>% unique), . %>% sapply(. %>% ifelse(is.na(.), 1, .))) %>% 
-  gene.increase.tree()
-
-TP53.gene.color <- RColorBrewer::brewer.pal(length(loxAfr4.geneTPM.ORA.expressedDups.TP53$Gene %>% unique), "Set1") %>% 
-  set_names(loxAfr4.geneTPM.ORA.expressedDups.TP53$Gene %>% unique)
-
-# TODO: Make the recursion make NA values to the parent node value, unlike the manual fix above.
-# TODO: Arrange the le
-p.treedata.tp53.table <- treedata.tp53.table %>% 
-  mutate(label=str_replace(label, "\\.", " ")) %>% 
-  as.treedata() %>% 
-  ggtree(
-    .,
-    branch.length = "none",
-    #continuous=T,
-    aes(
-    ),
-    size=2
-  ) +  
-  geom_point2(aes(x=branch, subset=(GTF2F1>0), color="GTF2F1"), size=5, position = position_nudge(x=-0.2, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(STK11>0), color="STK11"), size=5, position = position_nudge(x=0, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(COX5A>0), color="COX5A"), size=5, position = position_nudge(x=0.2, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(BRD7>0), color="BRD7"), size=5, position = position_nudge(x=-0.2, y=0)) +
-  geom_point2(aes(x=branch, subset=(TP53>0), color="TP53"), size=5, position = position_nudge(x=0, y=0)) +
-  geom_point2(aes(x=branch, subset=(CASP9>0), color="CASP9"), size=5, position = position_nudge(x=0.2, y=0)) +
-  geom_point2(aes(x=branch, subset=(PRDX1>0), color="PRDX1"), size=5, position = position_nudge(x=-0.2, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(LAMTOR5>0), color="LAMTOR5"), size=5, position = position_nudge(x=0, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(COX20>0), color="COX20"), size=5, position = position_nudge(x=0.2, y=0.2)) +
-  geom_tiplab() +
-  scale_color_manual(values = TP53.gene.color) + 
-  guides(color=guide_legend(title="Gene"))+
-  xlim(0,12) +
-  theme_tree2() + 
-  theme(legend.position = "bottom")
-
-p.tp53.genes.expression <- loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% 
-  ggplot(
-    aes(
-      x=id, 
-      y=TPM, 
-      color=Class
-    )
-  ) +
-  geom_point(
-    size=3
-  )+
-  geom_segment(
-    aes(
-      x=id,
-      xend=id,
-      y=0,
-      yend=TPM
-    )
-  )+
-  scale_y_sqrt()+
-  #scale_color_brewer(palette = "Dark2", name="Reactome Pathway")+
-  guides(color=guide_legend(nrow = 3))+
-  facet_grid(cols = vars(Gene), scales = "free_x") + 
-  theme_pubclean()+
-  labs(
-    x="Copy",
-    y="TPM (sqrt)"
-  ) + 
-  theme(legend.position="bottom")+
-  labs_pubr()
-```
-
-```{r Fig4, warning=T}
-f4a <- p.treedata.tp53.table
-f4b <- p.tp53.genes.expression
-fig4 <- ggarrange(f4a, f4b, nrow=1, ncol=2, labels="AUTO")
-
-fig4
-    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
-    # "STK11", #REgulation of TP53, Paenungulata                  #STK11 is a TSG
-    # "COX5A", #TP53 transcriptional regulation, PRobodiscea      #CytoC OXidase
-    # "BRD7", # TP53 regulation, Tethytheria                      #Required for p53-dependent oncogene-induced senescence
-    # "TP53", #TP53 duplicates in Tethytheria
-    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria   #E3 Ligase, induces apoptosis
-    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants  #Induces apoptosis, duh
-    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri          #Decomposes H2O2
-    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants   #Antiapoptotic?
-    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr            #CytoC OXidase
-```
-
-Prior studies looking at the duplication of TP53 in the African Elephant motivated further study for the enrichment of genes involved in TP53-related metabolic pathway.
-We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, and used publically-available RNA-seq data to see which genes are actively expressess in living elephants.
-Excitingly, we found that the initial duplication of TP53 in Tethytheria, where body size expanded, was presceded by the duplication of GTF2F1 and STK11 in _Paenungulata_; and was coincided by the duplication of BRD7.
-[TALK ABOUT HOW THESE TWO MAY HAVE FACILITATED THE DUPLICATION OF TP53]
-GTF2F1 is a general transcription factor [@CITE and talk more about this]
-STK11 is a tumor suppressor gene [@CITE and talk more]
-
-The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either alongside TP53 in the case of SIAH1, or subsequently in _Probodiscea_, _Elephantidae_, and in modern elephants.
-(Figure 4A).
-These genes are all expressed in RNA-seq data, suggesting that they encode functional genes in modern elephants (Fig 4B).
+`r knitr::knit_exit()`
 
@@
   as.treedata() %>% 
-  drop.tip(tree.Atlantogenata %>% as_tibble %>% filter(!BestGenome %in% (GCT.CEGMA %>% pull(Genome))) %>% pull(label))
+  drop.tip(tree.Atlantogenata %>% as_tibble %>% filter(!BestGenome %in% (genomeStatsRBHB %>% pull(Genome))) %>% pull(label))
 ##########################################################################################
@@
 ```
-```{r Fig 5 prep: Scales}
+```{r Fig2 Prep Scales}
 scale.log2CancerSucceptabilityChange <- table.cancerSucceptibility.Atlantogenata %>% 
@@
 
-```{r Fig 5 prep: Table 3}
+```{r Fig2 Prep Table}
 if(knitr::is_latex_output()){
@@
 
-```{r Fig 5 prep}
-f5a1 <- tree.cancerSucceptibility.Atlantogenata  %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(CancerSucceptabilityChange = as.numeric(CancerSucceptabilityChange)) %>% as.treedata %>% 
-  ggtree(
-    aes(color=log2CancerSucceptabilityChange),
-    continuous=T,
-    size = 1.5, 
-    yscale="log2CancerSucceptabilityChange"
-  ) + 
-  # geom_tiplab(aes(image=highlight, size=I(0.075)), geom="phylopic", color="black"
+```{r Fig2 prep}
+# f5a1 <- tree.cancerSucceptibility.Atlantogenata  %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(CancerSucceptabilityChange = as.numeric(CancerSucceptabilityChange)) %>% as.treedata %>% 
+#   ggtree(
+#     aes(color=log2CancerSucceptabilityChange),
+#     continuous=T,
+#     size = 1.5, 
+#     yscale="log2CancerSucceptabilityChange"
 #   ) + 
-  scale_color_viridis_c(limits=c(scale.log2CancerSucceptabilityChange.min, scale.log2CancerSucceptabilityChange.max), 
-                        guide = guide_colourbar(title="Log2 Cancer Succeptibility Change"))+
-  theme_minimal() +
-  theme(
-    legend.position="bottom",
-    panel.background = element_rect(fill = "darkgrey")
-  ) + 
-  xlab("Rate of Change, Body Size") +
-    ylab("Log2 Cancer Succeptibility Change")
-
-#############################################################################################
-#############################################################################################
-
-f5a <- tree.cancerSucceptibility.Atlantogenata %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% as.treedata %>% 
+#   # geom_tiplab(aes(image=highlight, size=I(0.075)), geom="phylopic", color="black"
+#   #             ) +
+#   scale_color_viridis_c(limits=c(scale.log2CancerSucceptabilityChange.min, scale.log2CancerSucceptabilityChange.max), 
+#                         guide = guide_colourbar(title="Log2 Cancer Susceptibility Change"))+
+#   theme_minimal() +
+#   theme(
+#     legend.position="bottom",
+#     panel.background = element_rect(fill = "darkgrey")
+#   ) + 
+#   xlab("Rate of Change, Body Size") +
+#     ylab("Log2 Cancer Susceptibility Change")
+# 
+# #############################################################################################
+# #############################################################################################
+
+f2a <- tree.cancerSucceptibility.Atlantogenata %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% as.treedata %>% 
   ggtree(
@@
   ) + 
-  # geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic",
-  #             show_guide=F, align = F, linesize = 0, offset= 100
-  #             ) +
+  geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic",
+              show_guide=F, align = F, linesize = 0, offset= 100
+              ) +
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidea, label="Proboscidea", 
@@
   scale_color_viridis_c(limits=c(scale.log2CancerSucceptabilityChange.min, scale.log2CancerSucceptabilityChange.max), 
-                        guide = guide_colourbar(title="Log2 Cancer Succeptibility Change"))+
-  xlim(0,250) +
+                        guide = guide_colourbar(title="Log2 Cancer Susceptibility Change"))+
+  xlim(0,300) +
   theme_tree2() + 
@@
 
-f5 <- ggarrange(
-  f5a, 
+f2 <- ggarrange(
+  f2a, 
   #nrow=2, ncol=1,
@@
 
-```{r Fig 5 prep: Table 4}
+```{r Fig2 Prep Table 1}
 t.cancersucc.atl <- table.cancerSucceptibility.Atlantogenata.genomes %>% 
@@
     `Est. Lifespan`=Lifespan,
-    `Est. Cancer Susceptibility (K1)` = K1,
-    `Est. Ancestral K2` = K2,
-    `Change in K (K2/K1)` = CancerSucceptabilityChange,
-    `log2 Fold Change` = log2CancerSucceptabilityChange
+    `K1` = K1,
+    `K2` = K2,
+    `Change in K` = CancerSucceptabilityChange,
+    `log2 Change` = log2CancerSucceptabilityChange
   ) %>%
   mutate_at(
-    vars(`Est. Cancer Susceptibility (K1)`, `Est. Ancestral K2`), 
+    vars(`K1`, `K2`, `Change in K`), 
     . %>% formatC(., format = "e", digits = 2)
@@
     `Est. Lifespan` = `Est. Lifespan` %>% round(2), 
-    `Change in K (K2/K1)` = `Change in K (K2/K1)` %>% round(2), 
-    `log2 Fold Change` = `log2 Fold Change` %>% round(2)
+    # `Change in K` = `Change in K` %>% round(2), 
+    `log2 Change` = `log2 Change` %>% round(2)
   ) %>% 
@@
     booktabs = TRUE,
-    caption = 'A table generated by the longtable package.'
-  )
-```
-```{r Fig 5, fig.cap="", warning=T}
-warning("TODO: Caption Fig 5")
-warning("TODO: de-comment geom_tiplab in f5a! Commented currently for offline usage!")
-f5
-```
-```{r Table 3, fig.cap="Table 3: Model Statistics for PGLS of Cancer Succeptibility"}
-t.size.lifespan.Atlantogenata.lm %>% cat
-```
-```{r Table 4, fig.cap="Table 4: Estimated Cancer Succeptibilty for nodes in Atlantogenata"}
+    caption = 'Estimated Cancer Succeptibilty for nodes in Atlantogenata'
+  )
+```
+```{r Figure2, fig.cap="Cancer succecptibility across Atlantogenata"}
+f2
+```
+
+```{r Table 1}
 t.cancersucc.atl
@@
 To infer the magnitude of these reductions we estimated differences in cancer risk between small bodied, short-lived species and large bodied, long-lived species as well as for reconstructed ancestral Afrotherians.
-Following [@Peto2015b] we estimate the intrinsic cancer risk as the product of risk associated with body mass and lifespan.
-Differences in cancer susceptibility $K$ due to body mass differences between species can be approximated simply as the fold difference in body mass ($D$) between species [@Peto2015].
-The risk of developing cancer also increases in proportion to the sixth power of age and is approximated by the formula $Ct^6$, in which the proportionality constant C that determines susceptibility to cancer induction is multiplied by the sixth power of the age in years, $t$ [@ArmitageAndDoll1954; @Nordling1953; @Peto2015].
+Following [@Peto:20151c2] we estimate the intrinsic cancer risk as the product of risk associated with body mass and lifespan.
+Differences in cancer susceptibility $K$ due to body mass differences between species can be approximated simply as the fold difference in body mass ($D$) between species [@Peto:20151c2].
+The risk of developing cancer also increases in proportion to the sixth power of age and is approximated by the formula $Ct^6$, in which the proportionality constant C that determines susceptibility to cancer induction is multiplied by the sixth power of the age in years, $t$ [@ArmitageAndDoll1954; @Nordling1953; @Peto:20151c2].
 Thus we can estimate the intrinsic cancer risk for a species as $K \approx Dt^6$.
@@
 In order to estimate the intrinsic cancer risk of a species, we first obtained estimates for lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon$ (Figure `r warning("make lifespan figure")`).
-With this information in hand, we calculated $K_{1}$ at all nodes, and then estimated the fold change in cancer succeptibility between an ancestral node and a given node as $\frac{K_{2}}{K_{1}}$ (__Table 4__).
+With this information in hand, we calculated $K_{1}$ at all nodes, and then estimated the fold change in cancer susceptibility between an ancestral node and a given node as $\frac{K_{2}}{K_{1}}$ (__Table 4__).
 As body size and phylogeny are the strongest predictors of lifespan, we find that this regression is sufficiently robust without further metabolic or other covariates for our purposes herein.
 
-As shown in __Table 4__, cancer succeptibility skyrocketed at the initial divergence of Atlantogenata, followed by a generally upwards trend.
-At the common ancestor of Afrotheria there is an inital `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrotheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer risk.
-In parallel to Afrotheria, cancer succeptibility increases `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold in Xenarthra.
+As shown in __Table 4__, cancer susceptibility skyrocketed at the initial divergence of Atlantogenata, followed by a generally upwards trend.
+At the common ancestor of Afrotheria there is an initial `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrotheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer risk.
+In parallel to Afrotheria, cancer susceptibility increases `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold in Xenarthra.
 However, cancer risk slowly deflates as size decreases as one moves along the tree towards extant species, such as in Hoffman's Two Toed Sloth (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Choloepus hoffmanni") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change) and in the Nine-banded Armadillo (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Dasypus novemcinctus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change).
 
-Within Afrotheria, cancer succeptibility drops in Afrosoricida as species shrink (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrosoricida") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, then stagnates for the Cape Golden Mole) - but then rises `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Echinops telfairi") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold towards the Lesser Hedgehog Tenrec.
-In parallel, Afroinsectivora does not increase in cancer succeptibility, and decreases once more at the Cape Elephant Shrew (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantulus edwardii") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
-The emergence of Pseudoungulata sees the next big leap in cancer succeptibility with a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Pseudoungulata") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase.
+Within Afrotheria, cancer susceptibility drops in Afrosoricida as species shrink (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrosoricida") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, then stagnates for the Cape Golden Mole) - but then rises `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Echinops telfairi") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold towards the Lesser Hedgehog Tenrec.
+In parallel, Afroinsectivora does not increase in cancer susceptibility, and decreases once more at the Cape Elephant Shrew (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantulus edwardii") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
+The emergence of Pseudoungulata sees the next big leap in cancer susceptibility with a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Pseudoungulata") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase.
 The Aardvark further increases `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Orycteropus afer") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, while we don't observe an increase at the common ancestor of Paenungulates.
-While the Rock Hyrax decreases in cancer succeptibility as expected (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Procavia capensis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold), Tethytheria sees a sharp increase in cancer risk (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
+While the Rock Hyrax decreases in cancer susceptibility as expected (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Procavia capensis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold), Tethytheria sees a sharp increase in cancer risk (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
 Within Tethytheria, the Manatee's cancer risk increases once more `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Trichechus manatus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, while Proboscidea's cancer risk drops precipitously along with its body size (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Proboscidea") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
-Yet, within Proboscidea we see the biggest increases: right off the bat, we see that the cancer succeptibility of  Elephantidae and the American Mastodon skyrocket by `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammut americanum") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively.
-Both Elephantina and Loxodontini in Elephantidae have a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantina") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer succeptibility.
+Yet, within Proboscidea we see the biggest increases: right off the bat, we see that the cancer susceptibility of  Elephantidae and the American Mastodon skyrocket by `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammut americanum") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively.
+Both Elephantina and Loxodontini in Elephantidae have a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantina") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer susceptibility.
 Within Elephantina, cancer susceptibility stays stable at Mammuthus and in the Colombian Mammoth, and slightly decreases in the Wooly Mammoth (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammuthus primigenius") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
-The three extant elephants - Asian Elephant in Elephantina, the African Savana Elephant in Loxodontini, and the African Forest Elephant in Loxodona, meanwhile, have parallel and similar decreases in both size and cancer succeptibility (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephas maximus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-, `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Loxodonta africana") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-, and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Loxodonta cyclotis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively).
-Neither the common ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further changes in cancer succeptibility.
+The three extant elephants - Asian Elephant in Elephantina, the African Savana Elephant in Loxodontini, and the African Forest Elephant in Loxodona, meanwhile, have parallel and similar decreases in both size and cancer susceptibility (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephas maximus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-, `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Loxodonta africana") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-, and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Loxodonta cyclotis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, respectively).
+Neither the common ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further changes in cancer susceptibility.
+
+## Identification and evolutionary history of gene duplications
+```{r ECNC DIAGRAMS, include=F}
+ecnc.df1 <- tibble(
+  hit = as.character(1) %>% paste("Hit", ., sep=" ") %>% as.factor(),
+  x1=c(0),
+  x2=c(100),
+  y1=c(0),
+  y2 = c(1)
+)
+ecnc.df2 <- tibble(
+  hit = as.character(1:2) %>% paste("Hit", ., sep=" ") %>% as.factor(),
+  x1=c(0, 0),
+  x2=c(100, 100),
+  y1=c(0, 1),
+  y2 = c(1, 2)
+)
+ecnc.df3 <- tibble(
+  hit = as.character(1:3) %>% paste("Hit", ., sep=" ") %>% as.factor(),
+  x1=c(0,25,75),
+  x2=c(25,75,100),
+  y1=c(0, 0, 0),
+  y2 = c(1, 1, 1)
+)
+ecnc.df4 <- tibble(
+  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
+  x1=c(0,25,75, 0),
+  x2=c(25,75,100, 100),
+  y1=c(0, 0, 0, 1),
+  y2 = c(1, 1, 1, 2)
+)
+
+ecnc.df5 <- tibble(
+  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
+  x1=c(0,0,0, 0),
+  x2=c(100,15,15, 15),
+  y1=c(0, 1, 2, 3),
+  y2 = c(1, 2, 3, 4)
+)
+
+ecnc.df6 <- tibble(
+  hit = as.character(c(1,2,1,3, 3)) %>% paste("Hit", ., sep=" ") %>% as.factor(),
+  label = factor(c("Hit 1", "Hit 2", "Unidentified", "Unidentified", "Hit 3")),
+  x1=c(0, 0, 75, 75, 50),
+  x2=c(75, 50, 100, 100, 75),
+  y1=c(0, 1, 0, 1, 1),
+  y2 = c(1, 2, 1, 2, 2),
+  transparent = c(0,0,1,1,0) %>% as.factor
+)
+
+ecnc.onegene.onehit <- 
+  ecnc.df1 %>% 
+  ggplot(
+    aes(
+      xmin = x1,
+      xmax = x2,
+      ymin = y1,
+      ymax = y2,
+      fill = hit
+    )
+  ) + 
+  geom_rect(color="black") + 
+  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
+  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
+  scale_fill_brewer(palette = "Dark2", guide=F) + 
+  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
+  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
+  labs(
+    x = "Query Length",
+    y = "Query Coverage",
+    title = "ECNC: One Contiguous Gene, One Hit"
+  )
+
+ecnc.twogene.twohit <- 
+  ecnc.df2 %>% 
+  ggplot(
+    aes(
+      xmin = x1,
+      xmax = x2,
+      ymin = y1,
+      ymax = y2,
+      fill = hit
+    )
+  ) + 
+  geom_rect(color="black") + 
+  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
+  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
+  scale_fill_brewer(palette = "Dark2", guide=F) + 
+  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
+  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
+  labs(
+    x = "Query Length",
+    y = "Query Coverage",
+    title = "ECNC: Two Contiguous Genes, Two Hits"
+  )
+
+ecnc.onegene.threehits <-
+  ecnc.df3 %>% 
+  ggplot(
+    aes(
+      xmin = x1,
+      xmax = x2,
+      ymin = y1,
+      ymax = y2,
+      fill = hit
+    )
+  ) + 
+  geom_rect(color="black") + 
+  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
+  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
+  scale_fill_brewer(palette = "Dark2", guide=F) + 
+  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
+  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
+  labs(
+    x = "Query Length",
+    y = "Query Coverage",
+    title = "ECNC: One Fragmented Gene, Multiple Hits"
+  )
+
+ecnc.twogene.fourhits <-
+  ecnc.df4 %>% 
+  ggplot(
+    aes(
+      xmin = x1,
+      xmax = x2,
+      ymin = y1,
+      ymax = y2,
+      fill = hit
+    )
+  ) + 
+  geom_rect(color="black") + 
+  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
+  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
+  scale_fill_brewer(palette = "Dark2", guide=F) + 
+  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
+  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
+  labs(
+    x = "Query Length",
+    y = "Query Coverage"#,
+    #title = "ECNC: Two Genes, Multiple Hits"
+  )
+
+ecnc.onegene.manypartialhits <-
+  ecnc.df5 %>% 
+  ggplot(
+    aes(
+      xmin = x1,
+      xmax = x2,
+      ymin = y1,
+      ymax = y2,
+      fill = hit
+    )
+  ) + 
+  geom_rect(color="black") + 
+  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
+  geom_hline(yintercept = 1.45, color="red", lty="dashed", size=2) + 
+  scale_fill_brewer(palette = "Dark2", guide=F) + 
+  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
+  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
+  labs(
+    x = "Query Length",
+    y = "Query Coverage"
+    # title = "ECNC: One Real Gene, Multiple Hits on Conserved Domain"
+  )
+
+ecnc.twogene.partialhits <-
+  ecnc.df6 %>% 
+  ggplot(
+    aes(
+      xmin = x1,
+      xmax = x2,
+      ymin = y1,
+      ymax = y2,
+      fill = hit,
+      alpha = transparent,
+      linetype = transparent
+    )
+  ) + 
+  geom_rect(color="black") + 
+  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=label), color="black", inherit.aes = F)+
+  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
+  scale_fill_brewer(palette = "Dark2", guide=F) + 
+  scale_alpha_manual(values = c("0"=1,"1"=0.5), guide=F) + 
+  scale_linetype_manual(values = c("0"="solid","1"="dotted"), guide=F) + 
+  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
+  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
+  labs(
+    x = "Query Length",
+    y = "Query Coverage",
+    title = "ECNC: Two Real Genes, Unidentified Tails"
+  )
+
+```
+```{r Correlations between genome quality scores (1)}
+CEGMA.completeness.clean <-
+  CEGMA.completeness %>%
+  filter(str_detect(Group, "_all")) %>%
+  select(-`Missing Proteins`) %>%
+  rename(
+    n.Prots = `#Prots`,
+    pc.Completeness = `%Completeness`,
+    n.Total = `#Total`,
+    pc.Ortho = `%Ortho`
+  ) %>%
+  reshape2::melt(., id = c("Genome", "Group")) %>%
+  mutate(Group = Group %>% str_remove("_all")) %>%
+  unite("variable", variable, Group) %>%
+  pivot_wider(id_cols = Genome, names_from = variable, values_from = value)
+
+CEGMA <- full_join(CEGMA.qc, CEGMA.completeness.clean, by="Genome")
+
+GCT.CEGMA <- genomeStatsRBHB %>%
+  select(-starts_with("q"), -starts_with("Median")) %>% 
+  inner_join(., CEGMA, by="Genome") %>%
+  mutate_if(is.character, as.factor)
+
+cor.GCT.CEGMA <- GCT.CEGMA %>%
+  mutate_if(is.factor, as.numeric) %>%
+  as.matrix %>%
+  cor %>%
+  reorder_cormat %>%
+  # get_upper_tri %>%
+  get_lower_tri %>%
+  reshape2::melt(., na.rm=T)
+
+cor.toHighlight <- c("mean.ECNC", "mean.Copy", "L50 sequence count", "# of sequences >  10M (nt)", "# of sequences >   1M (nt)", "# of sequences >  10K (nt)", "# of sequences >   1K (nt)", "# of sequences > 100K (nt)", "# of sequences", "Sum length of sequences > 1M  (nt)", "Total length (nt)", "Base composition (%) (N)", "mean.ECNCPerHit", "n.Prots_Partial","pc.Completeness_Partial", "Median sequence length (nt)", "Mean sequence length (nt)", "Sum length of sequences > 10M  (nt)")
+
+p.cor.GCT.CEGMA <-
+  cor.GCT.CEGMA %>%
+  ggplot(
+    aes(x=Var1, y=Var2, fill = value)
+  )+
+  geom_tile(color = "white")+
+  theme_minimal()+
+  coord_fixed()+
+  scale_x_discrete(
+    position = "top"
+    ) +
+  scale_y_discrete(position = "right") +
+  scale_fill_viridis_c(breaks = c(-1,-0.5, 0, 0.5, 1),labels = c(-1,-0.5, 0, 0.5, 1)) +
+  theme(
+    axis.text.x = element_text(
+      angle = -330, vjust = 0,
+      size = highlightTextSize(cor.GCT.CEGMA$Var1, cor.toHighlight, sizes = c(6,9)), 
+      color = highlightTextColor(cor.GCT.CEGMA$Var1, cor.toHighlight, color = c("black", "red")), 
+      hjust = 0
+    ),
+    axis.text.y = element_text(
+      size = highlightTextSize(cor.GCT.CEGMA$Var2, cor.toHighlight, sizes = c(6,9)), 
+      color = highlightTextColor(cor.GCT.CEGMA$Var2, cor.toHighlight, color = c("black", "red"))
+    ),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.border = element_blank(),
+    panel.background = element_blank(),
+    axis.ticks = element_blank(),
+    legend.justification = c(1, 1),
+    legend.position = c(0.4,0.7), #c(0.6, 0.7),
+    legend.direction = "horizontal",
+    legend.text = element_text(size=9),
+    legend.title =  element_text(size=11)
+    ) +
+  guides(
+    fill = guide_colorbar(
+      title="Pearson\nCorrelation",
+      barwidth = 7, barheight = 1,
+      title.position = "top", title.hjust = 0.5,
+      draw.llim = T
+    )
+  )
+```
+```{r Phylogenetic Correlations}
+
+to_drop <-tree.Atlantogenata %>% as_tibble %>% filter(!BestGenome %in% (GCT.CEGMA %>% pull(Genome))) %>% pull(label)
+
+tree.GCT.CEGMA <- full_join(
+  tree.Atlantogenata %>% 
+    drop.tip(to_drop) %>% 
+    as_tibble %>% 
+    mutate(Genome = BestGenome),
+  GCT.CEGMA, 
+  by="Genome"
+) %>% 
+  as.treedata()
+GCT.CEGMA.wGenome <- tree.GCT.CEGMA %>% as_tibble
+
+# GCT.CEGMA.wGenome %<>% filter(Gene == "TP53")
+# rown <- GCT.CEGMA.wGenome$label %>% unique
+# GCT.CEGMA.wGenome %<>% mutate_if(is.character, . %>% as.factor %>%  as.numeric)%>%
+#   as.data.frame
+# rownames(GCT.CEGMA.wGenome) <- rown
+```
+```{r Phylogenetic Correlations: model building (1), warning=T, eval=F}
+training.table.tree.GCT.CEGMA <- tree.GCT.CEGMA %>%
+  as_tibble() %>%
+  filter(!is.na(mean.CN)) %>%
+  mutate_if(is.character, . %>% as.factor) %>%
+  mutate(label=as.character(label) %>% str_replace_all("\\.", " ")) %>%
+  mutate_if(is.factor, as.numeric) %>%
+  as.data.frame %>%
+  column_to_rownames("label")
+
+names(training.table.tree.GCT.CEGMA) %<>% make.names(unique = T) %>% str_replace_all("\\.+", ".")
+
+phyloGCT.CEGMA <- tree.GCT.CEGMA@phylo %>% as.phylo()
+
+bm.gct <-corBrownian(1, phyloGCT.CEGMA)
+pl.gct <-corPagel(1, phyloGCT.CEGMA)
+
+models.vars <- training.table.tree.GCT.CEGMA %>% 
+  select(
+    # -mean.Copy, 
+    # -mean.ECNC, 
+    # -mean.CN,
+    # -sd.ECNC, 
+    # -sd.Copy, 
+    # -sd.CN,
+    -X.of.sequences.containing.non.ACGTN.nt., -Base.composition.Other., -parent, -node, 
+    -Brownian, -Median, -X95.CI_Low, -X95.CI_High, -branch.length, -Species, 
+    -Species.strict, -Spc.short, -BestGenome, -highlight, -Base.composition.C.,
+    -Base.composition.A., -Base.composition.T., -Base.composition.G., 
+    -phyloPicUID, -Rate,-Orig_len, -Common.Name, -Longest.sequence.nt., 
+    -Shortest.sequence.nt., -Mean.sequence.length.nt., -Median.sequence.length.nt., 
+    -starts_with("X.of.sequences"), -starts_with("Sum.length")
+  )
+
+models.mean.ECNC <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.ECNC ~ ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.CN <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.CN ~ ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.Copy <-names(models.vars) %>% 
+    str_remove("mean.*|sd.*") %>% 
+    stringi::stri_omit_empty() %>% 
+    paste0("mean.Copy ~ ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+aic.models.mean.ECNC <- AIC(models.mean.ECNC[[1]], models.mean.ECNC[[2]], models.mean.ECNC[[3]], models.mean.ECNC[[4]], models.mean.ECNC[[5]], models.mean.ECNC[[6]], models.mean.ECNC[[7]], models.mean.ECNC[[8]], models.mean.ECNC[[9]], models.mean.ECNC[[10]], models.mean.ECNC[[11]], models.mean.ECNC[[12]], models.mean.ECNC[[13]], models.mean.ECNC[[14]],
+    models.mean.ECNC[[15]], models.mean.ECNC[[16]], models.mean.ECNC[[17]], models.mean.ECNC[[18]]#,models.mean.ECNC[[19]], models.mean.ECNC[[20]], models.mean.ECNC[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.ECNC)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.Copy <- AIC(models.mean.Copy[[1]], models.mean.Copy[[2]], models.mean.Copy[[3]], models.mean.Copy[[4]], models.mean.Copy[[5]], models.mean.Copy[[6]], models.mean.Copy[[7]], models.mean.Copy[[8]], models.mean.Copy[[9]], models.mean.Copy[[10]], models.mean.Copy[[11]], models.mean.Copy[[12]], models.mean.Copy[[13]], models.mean.Copy[[14]],
+    models.mean.Copy[[15]], models.mean.Copy[[16]], models.mean.Copy[[17]], models.mean.Copy[[18]]#,models.mean.Copy[[19]], models.mean.Copy[[20]], models.mean.Copy[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.Copy)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.CN <- AIC(models.mean.CN[[1]], models.mean.CN[[2]], models.mean.CN[[3]], models.mean.CN[[4]], models.mean.CN[[5]], models.mean.CN[[6]], models.mean.CN[[7]], models.mean.CN[[8]], models.mean.CN[[9]], models.mean.CN[[10]], models.mean.CN[[11]], models.mean.CN[[12]], models.mean.CN[[13]], models.mean.CN[[14]],
+    models.mean.CN[[15]], models.mean.CN[[16]], models.mean.CN[[17]], models.mean.CN[[18]]#,models.mean.CN[[19]], models.mean.CN[[20]], models.mean.CN[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.CN)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+```
+```{r Phylogenetic Correlations: model building (1 + lnsize), warning=T, eval=F}
+models.mean.ECNC.lnSize <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*|lnSize") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.ECNC ~ lnSize + ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.CN.lnSize <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*|lnSize") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.CN ~ lnSize + ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+models.mean.Copy.lnSize <- names(models.vars) %>% 
+  str_remove("mean.*|sd.*|lnSize") %>% 
+  stringi::stri_omit_empty() %>% 
+  paste0("mean.Copy ~ lnSize + ", .) %>% 
+  set_names(., .) %>% 
+  sapply(
+    ., 
+    . %>% 
+      paste0(
+        "gls(", ., ", data = training.table.tree.GCT.CEGMA, correlation=bm.gct)"
+        ) %>% 
+      parse(text = .)
+    ) %>% 
+  lapply(eval)
+
+aic.models.mean.ECNC.lnSize <- AIC(models.mean.ECNC.lnSize[[1]], models.mean.ECNC.lnSize[[2]], models.mean.ECNC.lnSize[[3]], models.mean.ECNC.lnSize[[4]], models.mean.ECNC.lnSize[[5]], models.mean.ECNC.lnSize[[6]], models.mean.ECNC.lnSize[[7]], models.mean.ECNC.lnSize[[8]], models.mean.ECNC.lnSize[[9]], models.mean.ECNC.lnSize[[10]], models.mean.ECNC.lnSize[[11]], models.mean.ECNC.lnSize[[12]], models.mean.ECNC.lnSize[[13]], models.mean.ECNC.lnSize[[14]],
+    models.mean.ECNC.lnSize[[15]], models.mean.ECNC.lnSize[[16]], models.mean.ECNC.lnSize[[17]]#, models.mean.ECNC.lnSize[[18]],models.mean.ECNC.lnSize[[19]], models.mean.ECNC.lnSize[[20]]#, models.mean.ECNC.lnSize[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.ECNC.lnSize)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.Copy.lnSize <- AIC(models.mean.Copy.lnSize[[1]], models.mean.Copy.lnSize[[2]], models.mean.Copy.lnSize[[3]], models.mean.Copy.lnSize[[4]], models.mean.Copy.lnSize[[5]], models.mean.Copy.lnSize[[6]], models.mean.Copy.lnSize[[7]], models.mean.Copy.lnSize[[8]], models.mean.Copy.lnSize[[9]], models.mean.Copy.lnSize[[10]], models.mean.Copy.lnSize[[11]], models.mean.Copy.lnSize[[12]], models.mean.Copy.lnSize[[13]], models.mean.Copy.lnSize[[14]],
+    models.mean.Copy.lnSize[[15]], models.mean.Copy.lnSize[[16]], models.mean.Copy.lnSize[[17]]#, models.mean.Copy.lnSize[[18]],models.mean.Copy.lnSize[[19]], models.mean.Copy.lnSize[[20]]#, models.mean.Copy.lnSize[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.Copy.lnSize)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+aic.models.mean.CN.lnSize <- AIC(models.mean.CN.lnSize[[1]], models.mean.CN.lnSize[[2]], models.mean.CN.lnSize[[3]], models.mean.CN.lnSize[[4]], models.mean.CN.lnSize[[5]], models.mean.CN.lnSize[[6]], models.mean.CN.lnSize[[7]], models.mean.CN.lnSize[[8]], models.mean.CN.lnSize[[9]], models.mean.CN.lnSize[[10]], models.mean.CN.lnSize[[11]], models.mean.CN.lnSize[[12]], models.mean.CN.lnSize[[13]], models.mean.CN.lnSize[[14]],
+    models.mean.CN.lnSize[[15]], models.mean.CN.lnSize[[16]], models.mean.CN.lnSize[[17]]#, models.mean.CN.lnSize[[18]],models.mean.CN.lnSize[[19]], models.mean.CN.lnSize[[20]]#, models.mean.CN.lnSize[[21]]
+    ) %>% 
+  rownames_to_column("Model") %>% 
+  mutate(Model = names(models.mean.CN.lnSize)) %>%
+  separate(Model, into = c("Y", "X"), sep="~") %>% 
+  arrange(AIC)
+
+```
+```{r Genome Table Prep}
+genome.table.forFig <- genomeTable.atlantogenata %>% 
+  left_join(anage, by="label") %>% 
+  left_join(tree.Atlantogenata.table, by="label") %>% 
+  mutate(
+    Median = exp(Median) %>% formatC(format = "g", digits = 2),
+    Lifespan = Lifespan %>% exp %>% formatC(format = "g", digits = 2) %>% as.character() %>% str_replace_na(replacement = "(Unknown)")
+  ) %>% 
+  select(
+    `Common Name` = `Common Name.x`,
+    Species = Species.strict.x,
+    # Clade = Family, 
+    `Size (g)` = Median,
+    # `Lifespan (y)`= Lifespan, 
+    Genome = BestGenome.x
+  ) %>% 
+  left_join(
+    genomeStatsRBHB %>% 
+      select(Genome, `#Hits`=Hits, `% Genes Found`=pc.Found, `#Duplicated` = Duplicated,`% Hits Duplicated` = pc.Hits.Duplicated, `Max Copy#`=max.Copy, `Max ECNC`=max.ECNC, `Mean ECNC/Hit`=mean.ECNCPerHit),
+    by="Genome"
+  ) %>% 
+  mutate(
+    `Common Name` = str_to_title(`Common Name`),
+    `% Genes Found` = `% Genes Found` %>% as.percentage,
+    `% Hits Duplicated` = `% Hits Duplicated` %>% as.percentage,
+    `Mean ECNC/Hit` = round(`Mean ECNC/Hit`, 2)
+  )
+
+# #clades
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Elephant$|Mammoth$"),]$Clade <- "Paenungulata, Probodiscea"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Mastodon"),]$Clade <- "Paenungulata, Probodiscea"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "^Cape|Tenrec"),]$Clade <- "Macroscelididae/Afroinsectivora"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Manatee"),]$Clade <- "Paenungulata, Tethytheria"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Hyrax"),]$Clade <- "Paenungulata"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Aardvark"),]$Clade <- "Pseudoungulata"
+# genome.table.forFig[str_detect(genome.table.forFig$`Common Name`, "Sloth|Armadillo"),]$Clade <- "Xenarthra"
+# 
+# #Lifespan caveats
+# genome.table.forFig[which(genome.table.forFig$Species == "Elephantulus edwardii"),]$`Lifespan (y)` = "~5.7*"
+# genome.table.forFig[which(genome.table.forFig$Species == "Loxodonta cyclotis"),]$`Lifespan (y)` = paste0(genome.table.forFig[which(genome.table.forFig$Species == "Loxodonta africana"),]$`Lifespan (y)`, "\u2020")
+```
+```{r Figure3 Prep Genome Table}
+ggt.GenomeTable.withNote <- genome.table.forFig %>% 
+  select(Species, `Common Name`, starts_with("Size"), starts_with("#"), starts_with("%"), starts_with("Mean")) %>% 
+  arrange(Species)%>% 
+  knitr::kable(
+    ., 
+    longtable = TRUE, 
+    booktabs = TRUE,
+    caption = 'Summary of duplications in Atlantogenata'
+  )
+```
+```{r Table2, fig.height=8, fig.width=6}
+ggt.GenomeTable.withNote
+```
+
+```{r Figure3, fig.cap="Gene duplications occur readily throughout Atlantogenata.
+A) A table summarizing duplication events in Atlantogenata.
+B) A tree of the species in Atlantogenata with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.", fig.height=8, fig.width=6}
+
+f3 <- p.ml.tree.increasing.geneCounts.withSize
+# f3 <- ggarrange(f3a, f3b, nrow=2, ncol=1, labels = "AUTO", heights = c(1,2))
+# f3
+f3
+```
+
+Enhanced cancer suppression may have evolved through many mechanisms; among the most parsimonious is an increase in the copy number of genes with tumor suppressor functions.
+Previous studies focusing on candidate gene studies, for example, have identified increased copy number of the tumor suppressors _TP53_ and _LIF_ in elephants [@Abegglen:JAMA2015; @Caulin2015; @Doherty2016; @Sulak2015; @Vazquez2018].
+Therefore, in order to test whether this was pervasive genome-wide in _Afrotherians_, we used a Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in _Afrotherian_ and _Atlantogenatan_ genomes (Fig.
+2A, Supplementary Figure 1).
+Because RBHB-like approaches can over-estimate copy number when genes are fragmented  or incorrectly assembled across multiple scaffolds, we also inferred copy number using a complementary method that quantifies the ratio between observed and expected gene coverage per nucleotide (ECNC) (Supplementary Figure 1).
+By only including nucleotides from the query sequence that were observed in the target genome, we also correct for partial hits where some or all of the homologs of a gene have diverged from the human homolog.
+
+As our sequence database included various protein transcripts for each gene, in order to obtain gene-level copy number information and eliminate any many-to-one mappings of hits, we labeled each exon of every reciprocal best hit (RBH) with the gene corresponding to the query transcript and merged all overlapping exons; next, we eliminated any many-to-one exons that resulted from the previous step.
+Finally, we reassembled the gene loci based on the original transcript starts and ends, and the collapsed exon data, obtaining the full sequence of each RBH locus.
+Genes were considered to be duplicated if its copy number via RBHB was greater than or equal to 2, and the maximum ECNC among all transcripts prior to filtering was greater than or equal to 1.50 in order to account for partial hits that may be due to incomplete coverage.
+
+Once we applied these criteria to the results, we obtained the results shown in Figure 2A.
+Our approach positively identified an average of `r genomeStatsRBHB$pc.found %>% mean %>% as.percentage` of the queries we searched.
+We observed that the percentage of duplicated genes in non-Pseudoungulatan genomes was significantly higher: while these genomes had duplicates percentages ranging anywhere from `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`, outgroup species' duplication rates ranged from `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`.
+
+```{r Figure4, fig.cap="Correlations between genome quality metrics and ECNC metrics"}
+f4 <- p.cor.GCT.CEGMA
+f4
+```
+
+Given the preliminary status of the genomes for _Chrysochloris asiatica_ and _Elephantulus edwardii_; and the fact that more revisions have been performed for Paenungulatan genomes; we hypothesized that genome quality may be confounding the results we've observed.
+We used `CEGMA` and the `gVolante` server [@CEGMA; @gVolante] to assess the quality of all our genomes, to see if any of these metrics correlated with our metrics of gene counts and copy number.
+As shown in Figure 3, mean ECNC, mean Copy Number, and mean CN (the lesser of Copy Number and ECNC per gene) correlate moderately strongly with genomic quality metrics related to the length and assembly quality, like LD50 and the number of scaffolds and contigs with a length above either 100K or 1M, supporting our hypothesis that increases in copy number correlate .
+
+In order to determine when in the phylogeny our identified duplications occurred, we used maximum likelihood to reconstruct ancestral gene copy numbers across the phylogeny.
+We encoded the copy number of each gene as a discrete trait ranging from 0 ("?") to 31+, and then ran the ancestral gene copy number reconstruction using the phylogenetics suite `IQTREE`, which rapidly tests and runs the best evolutionary model for a given trait of interest.
+Importantly, `IQTREE` can be set to only output states where the likelihood of the state is greater than a certain threshold.
+Setting the threshold to 0.8, we obtained the results shown in Figure 2B.
+
+Gene duplications events occured readily thoughout _Atlantogenata_ (Figure 2B).
+Among the genes that increased in copy number in the elephant lineage are TP53 and LIF, as previously described; however, we find that these two genes represent a fraction of the 940 genes that are duplicated in the African Elephant overall, which accumulated over various steps through their evolution.
+While the extinct elephantids have acceptable genome quality metrics according to CEGMA, they are nonetheless missing a significant number of sequences; this may contribute to the low number of duplicated genes that occured in internal nodes.
+The number of duplicates that occur at each branch is also proportional to the density of the sampling of the clade overall, as would be expected.
+In branches, such as _Afrosoricida_, where the number of species is relatively minuscule compared to the size of the clade, we see many significantly larger numbers of duplications private to these species.
+
+## Duplications that occurred recently in Probodiscea are enriched for tumor suppressor pathways
+
+```{r ORA.dotplots}
+p.loxAfr.ORA.dotplot <- ORA.all %>% 
+    filter(Node == "Loxodonta_africana", FDR.Threshold <= 0.3, Database == "Reactome", RBHB == "RBB") %>% 
+    graph.ORA.dotplot(., plotly=knitr::is_html_output())
+
+p.loxAfr.ORA.dotplot.increased <- ORA.all %>% 
+    filter(Node == "Loxodonta_africana", FDR.Threshold <= 0.3, Database == "Reactome", RBHB == "RBB", Direction=="increased") %>% 
+    graph.ORA.dotplot(., plotly=knitr::is_html_output())
+# p.Proboscidea.ORA.dotplot <- ORA.all %>% 
+#   filter(Direction=="increased", Node == "Proboscidea", Database == "Reactome", RBHB == "RBB") %>% 
+#   graph.ORA.dotplot(., plotly=knitr::is_html_output())
+```
+
+```{r Branches which experienced an increase in size}
+
+ORA.numbers <- ORA.all  %>% filter(RBHB=="RBB") %>% group_by(Direction, Ancestor, Node, FDR.Threshold, Database) %>% summarise(N.Pathways=n_distinct(Pathway)) %>% ungroup %>% pivot_wider(names_from = FDR.Threshold, values_from = N.Pathways, names_prefix = "Pathways at FDR<") %>% mutate_if(is.numeric, .
+%>% ifelse(is.na(.), 0,.))
+
+ORA.numbers.ReactomeOnly <- ORA.all %>% filter(Database=="Reactome", RBHB=="RBB") %>% group_by(Direction, Ancestor, Node, FDR.Threshold) %>% summarise(N.Pathways=n_distinct(Pathway)) %>% ungroup %>% pivot_wider(names_from = FDR.Threshold, values_from = N.Pathways, names_prefix = "Pathways at FDR<") %>% mutate_if(is.numeric, .
+%>% ifelse(is.na(.), 0,.))
+
+noEnrichmentIncreasing <- setdiff(ORA.numbers %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers %>% filter(Direction == "increased") %>%  pull(Node)) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+noEnrichmentIncreasing.Reactome <- setdiff(ORA.numbers.ReactomeOnly %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers.ReactomeOnly %>% filter(Direction == "increased") %>%  pull(Node)) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+
+noEnrichment.any <- setdiff(ml.treedata.table$label %>% str_replace_all("\\.", "_"), union(ORA.numbers %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers %>% filter(Direction == "increased") %>%  pull(Node))) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+noEnrichment.any.Reactome <- setdiff(ml.treedata.table$label %>% str_replace_all("\\.", "_"), union(ORA.numbers.ReactomeOnly %>% filter(Direction == "stable") %>% pull(Node), ORA.numbers.ReactomeOnly %>% filter(Direction == "increased") %>%  pull(Node))) %>% str_replace_all("_", " ") %>% str_c("_", ., "_")
+
+p.ORA <- ORA.numbers.ReactomeOnly %>% 
+  filter(Direction == "increased") %>% 
+  select(-Direction) %>% 
+  knitr::kable(
+    ., 
+    longtable = TRUE, 
+    booktabs = TRUE,
+    caption = 'Number of pathways overrepresented among duplicated genes at different FDRs.'
+  )
+
+```
+
+```{r Table3}
+p.ORA
+```
+
+```{r Figure5, fig.cap="Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.", fig.width=6, fig.height=8}
+f5a <- p.loxAfr.ORA.dotplot
+f5b <- p.loxAfr.ORA.dotplot.increased
+
+f5 <- ggarrange(f5a, f5b, nrow=1, ncol=2, common.legend = T, legend = "bottom", labels="AUTO")
+
+f5
+```
+
+Our initial hypothesis was that genes which duplicated in lineages that experienced a growth in size would be enriched for membership in tumor suppressor pathways.
+Thus, we used WebGestalt and its Overrepresentation Analysis (ORA) functionality to determine what pathways were enriched in our duplicated gene sets in each branch relative to our initial query set.
+For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways_cancer databases using WebGestalt ORA.
+Going through the tree, at no FDR<0.5 is there any significant pathway representation for genes that increased in the branches leading to `r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`; furthermore, there are no significant pathway enrichments at FDR<0.5 for genes whose copy number did not change between branches (copy-number-stable) for `r noEnrichment.any.Reactome[1:length(noEnrichment.any.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichment.any.Reactome[length(noEnrichment.any.Reactome)]`.
+Note that because _Xenarthra_ was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.
+
+For the other branches, the number of pathways that came up as significantly enriched at each FDR is shown in Figure 3A (Supplementary Figure 3).
+For the species with high duplication rates and lower-quality, highly-fragmented genomes, such as with _Chrysochloris asiatica_ (`r genomeStatsRBHB %>% filter(Genome == "chrAsi1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits) and _Elephantulus edwardii_ (`r genomeStatsRBHB %>% filter(Genome == "eleEdw1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits), it is unsurprising that there is a proportionally large number of pathway enrichments.
+In the case of these two species, their many pathway enrichments also span an incredible range of processes at every level of biology; this, in combination with the high number of copy numbers identified for these genes, further suggests a need for improvement and refinement in these genomes.
+In the cell cycle pathways called as significant in the genomes of these two species plus _Orycteropus afer_ and _Echinops telfairi_, the duplicated genes included in these sets are from the same gene families, such as the APC subunit family; the proteosome subunit families; and the protein phosphatase 2 family, among others.
+It is highly possible that these results reflect true expansions of these gene families, especially in the higher-quality _Orycteropus afer_ genome; however, it is also possible that it simply reflects artificial duplication events, and so require further study.
+
+The pathway enrichments for genes whose copy number either did not change, or whose copy number increased, between _Loxodona_ and the African Elephant are shown in Figure 3B and 3C, respectively.
+Among the few enriched pathways in the African Elephant, we see that that two tumor suppression pathways - APC Complex-related pathways, and "TP53 Regulates Metabolic Genes"- appear not only in the case of stable genes, but also in the set of newly duplicated genes.
+The other pathways we see in the set of recently-duplicated genes include "Functionalization of Compounds" and its daughter pathway "Xenobiotics".
+Genes in these pathways serve to add functional groups to lipophylic compounds which would otherwise not be reactive in the cell, and are types of metabolic pathways.
+In the stable set we see enrichment of pathways such as "Neuronal Systems" and "Axon Guidance," which fit in well with what is known about elephant biology and evolution [@Goodman2009].
+Overall in elephants, we see enrichments within duplicated genes for pathways involved in what makes an elephant an elephant - including tumor suppressor pathways.
+
+## Concerted duplication of TP53 and TP53-related genes towards _Probodiscea_
+
+```{r Fig6 prep}
+# loxAfr4.geneTPM.genesToHighlight.plus
+    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
+    # "STK11", #REgulation of TP53, Paenungulata                  #STK11 is a TSG
+    # "COX5A", #TP53 transcriptional regulation, PRobodiscea      #CytoC OXidase
+    # "BRD7", # TP53 regulation, Tethytheria                      #Required for p53-dependent oncogene-induced senescence
+    # "TP53", #TP53 duplicates in Tethytheria
+    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria   #E3 Ligase, induces apoptosis
+    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants  #Induces apoptosis, duh
+    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri          #Decomposes H2O2
+    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants   #Antiapoptotic?
+    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr            #CytoC OXidase
+
+genesToHighlight.plus.TP53 <- genesToHighlight.plus$TP53
+
+loxAfr4.geneTPM.ORA.expressedDups.TP53 <- loxAfr4.geneTPM.ORA.expressedDups %>% 
+  filter(Gene %in% genesToHighlight.plus.TP53) %>% 
+  mutate(
+    Reactome.Pathway.Prime = Pathway %>% str_split(", ") %>% sapply(. %>% str_extract(., ".*[Tt]?[Pp]53.*")%>% stringi::stri_remove_empty_na() %>% paste(collapse =", ")),
+    id = str_remove(Locus, ".*_") %>% factor(., levels=sort(as.numeric(unique(.)))),
+    Gene = Gene %>% factor(., levels=genesToHighlight.plus$TP53)
+    ) %>% 
+  group_by(Locus) %>% 
+  summarize(Gene = unique(Gene), id=unique(id), TPM=max(TPM), Reactome.Pathway.Prime=unique(Reactome.Pathway.Prime)) %>% 
+  arrange(Gene, id) %>% 
+  mutate(Class = ifelse(
+    str_detect(Reactome.Pathway.Prime, "[Rr]egulation of [Tt][Pp]53")& Gene != "TP53",
+    "Regulator of TP53",
+    ifelse(
+      str_detect(Gene, "COX")& Gene != "TP53",
+      "Cytochrome C\nOxidase",
+      ifelse(
+        str_detect(Gene, "CASP")& Gene != "TP53",
+      "Apoptosis",
+      "TP53 Pathway\n(Other)"
+      )
+    )
+
+  )
+           )
+
+treedata.tp53.table <- ml.treedata.table %>% 
+  full_join(., bodysize.tree.atlantogenata, by="label") %>%
+  select(parent, node, branch.length, label, lnSize, all_of(loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% pull(Gene) %>% unique)) %>% 
+  mutate_at(vars(loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% pull(Gene) %>% unique), . %>% sapply(. %>% ifelse(is.na(.), 1, .))) %>% 
+  gene.increase.tree()
+
+TP53.gene.color <- RColorBrewer::brewer.pal(length(loxAfr4.geneTPM.ORA.expressedDups.TP53$Gene %>% unique), "Set1") %>% 
+  set_names(loxAfr4.geneTPM.ORA.expressedDups.TP53$Gene %>% unique)
+
+p.treedata.tp53.table <- treedata.tp53.table %>% 
+  mutate(label=str_replace(label, "\\.", "\n") %>% str_c("bold('", ., "')")) %>% 
+  as.treedata() %>% 
+  ggtree(
+    .,
+    branch.length = "none",
+    #continuous=T,
+    aes(
+    ),
+    size=2
+  ) +  
+  geom_point2(aes(x=branch, subset=(GTF2F1>0), color="GTF2F1"), size=3, position = position_nudge(x=-0.2, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(STK11>0), color="STK11"), size=3, position = position_nudge(x=0, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(COX5A>0), color="COX5A"), size=3, position = position_nudge(x=0.2, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(BRD7>0), color="BRD7"), size=3, position = position_nudge(x=-0.2, y=0)) +
+  geom_point2(aes(x=branch, subset=(TP53>0), color="TP53"), size=3, position = position_nudge(x=0, y=0)) +
+  geom_point2(aes(x=branch, subset=(CASP9>0), color="CASP9"), size=3, position = position_nudge(x=0.2, y=0)) +
+  geom_point2(aes(x=branch, subset=(PRDX1>0), color="PRDX1"), size=3, position = position_nudge(x=-0.2, y=-0.2)) +
+  geom_point2(aes(x=branch, subset=(LAMTOR5>0), color="LAMTOR5"), size=3, position = position_nudge(x=0, y=-0.2)) +
+  geom_point2(aes(x=branch, subset=(COX20>0), color="COX20"), size=3, position = position_nudge(x=0.2, y=-0.2)) +
+  geom_tiplab(parse=T, offset = 0.25) +
+  scale_color_manual(values = TP53.gene.color) + 
+  guides(color=guide_legend(title="Gene", title.position = "top", title.theme = element_text(size = 14, face="bold")))+
+  xlim(0,20) +
+  theme_tree()+
+  theme(
+    strip.text = element_text(size=14, face="bold"), 
+    legend.position = "bottom",
+    legend.title = element_text(size = 14, face="bold"), 
+    legend.text = element_text(size = 12),
+    axis.text.y = element_blank()
+    )
+
+p.tp53.genes.expression <- loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% 
+  ggplot(
+    aes(
+      x=id, 
+      y=TPM, 
+      color=Class
+    )
+  ) +
+  geom_point(
+    size=3
+  )+
+  geom_segment(
+    aes(
+      x=id,
+      xend=id,
+      y=0,
+      yend=TPM
+    )
+  )+
+  scale_y_sqrt()+
+  #scale_color_brewer(palette = "Dark2", name="Reactome Pathway")+
+  guides(color=guide_legend(nrow = 2, ncol=2, title= "Pathway", title.position = "top"))+
+  facet_grid(rows = vars(Gene), scales = "free_y", switch = "y") + 
+  theme_pubclean()+
+  labs(
+    x="Copy",
+    y="TPM (sqrt)"
+  ) + 
+  theme(
+    legend.position="bottom",
+    strip.placement = "outside",
+    strip.background = element_rect(fill="white"),
+    strip.text.y = element_text(face = "bold", size=14),
+    legend.title = element_text(size = 14, face="bold"), 
+    legend.text = element_text(size = 12),
+    axis.title = element_text(size=14, face="bold"),
+    axis.text.x = element_text(size=12, face="bold"),
+    axis.text.y = element_text(size=12, face = "bold")
+  )+
+  coord_flip()
+```
+
+```{r Figure6, fig.cap="", fig.width=5, fig.height=8}
+f6a <- p.treedata.tp53.table
+f6b <- p.tp53.genes.expression
+fig6 <- ggarrange(f6a, f6b, nrow=1, ncol=2, labels="AUTO", widths = c(1.5,1))
+
+fig6
+    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
+    # "STK11", #REgulation of TP53, Paenungulata                  #STK11 is a TSG
+    # "COX5A", #TP53 transcriptional regulation, PRobodiscea      #CytoC OXidase
+    # "BRD7", # TP53 regulation, Tethytheria                      #Required for p53-dependent oncogene-induced senescence
+    # "TP53", #TP53 duplicates in Tethytheria
+    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria   #E3 Ligase, induces apoptosis
+    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants  #Induces apoptosis, duh
+    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri          #Decomposes H2O2
+    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants   #Antiapoptotic?
+    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr            #CytoC OXidase
+```
+
+Prior studies looking at the duplication of TP53 in the African Elephant motivated further study for the enrichment of genes involved in TP53-related metabolic pathway.
+We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, and used publicly-available RNA-seq data to see which genes are actively expresses in living elephants.
+Excitingly, we found that the initial duplication of TP53 in Tethytheria, where body size expanded, was preceded by the duplication of GTF2F1 and STK11 in _Paenungulata_; and was coincided by the duplication of BRD7.
+These two genes are involved in regulating the transcription of TP53, and their duplication prior to that of TP53 may have facilitated its retroduplication.
+Interestingly enough, STK11 is a tumor suppressor gene in its own right, and plays additional roles in mediating tumor suppression via p21-induced senescence [@Launonen2005].
+<!-- [TALK ABOUT HOW THESE TWO MAY HAVE FACILITATED THE DUPLICATION OF TP53] -->
+<!-- GTF2F1 is a general transcription factor [@CITE and talk more about this] -->
+<!-- STK11 is a tumor suppressor gene [@CITE and talk more] -->
+
+The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either alongside TP53 in the case of SIAH1, or subsequently in _Probodiscea_, _Elephantidae_, and in modern elephants.
+(Figure 4A).
+These genes are all expressed in RNA-seq data, suggesting that they encode functional genes in modern elephants (Fig 4B).
 
@@
 
-With our results, we have demonstrated that gigantism in _Atlantogenata_ was not limited to extant elephants, but rather occured at various points in the evolution of the clade; however, the hundred-fold to hundred-million-fold increases in cancer risk that is associated with these increases however poses an inate challenge in the evolution and persistance of this trait.
+With our results, we have demonstrated that gigantism in _Atlantogenata_ was not limited to extant elephants, but rather occurred at various points in the evolution of the clade; however, the hundred-fold to hundred-million-fold increases in cancer risk that is associated with these increases however poses an innate challenge in the evolution and persistence of this trait.
 The environmental selective pressures on body size have long been the fascination of evolutionary biologists, and the influence of climate, predation, geography, and ecological niche on body size have been well established.
-[Cooper & Purvis 2010, MORE CITATIONS].
-Indeed, there is a general trend, known as Cope's Rule, for body sizes of species to increase over time [Citation for Cope's Rule].
-However, for all the research on body size that has been done thus far, the mechanisms that enable a release from the negative pressure on body size exerted on cancer risk has proven more elusive.
+[@CooperAndPurvis2010].
+Indeed, there is a general trend, known as Cope's Rule, for body sizes of species to increase over time [@HoneAndBenton2004; @HuntAndRoy2006; @Jablonski1997].
+However, for all the research on body size that has been done thus far, the mechanisms that enable a release from the negative pressure on body size exerted on cancer risk has proven more elusive [@HoneAndBenton2004; @HuntAndRoy2006; @Jablonski1997].
 
 Furthermore, we show that tumor suppressor duplications are enriched not only in large, extant species, but also in large common ancestors, and that these duplications evolved throughout the tree, rather than in concert.
-The ancestral body sizes of many of the subclades in _Atlantogenata_ were estimated to be large, and the estimated cancer risk increases - even for the small clades like _Afroinsectivora_ may explain why these TSG gene duplications occured early on, and persisted even in species that we did not expect to have high risks of cancer.
-However, some of our results also provide interesting evidence for a paradigm of TSG duplicaitons being a pervasive phenomenon, rather than a specific mechansim that occurs after the evolution of gigantism.
-At the common ancestor of _Proboscidea_, which was small relative to both its ancestors and its decendants, we see the emergence of various TSG duplications, which may have enabled the stratospheric increase in body size of modern elephants.
+The ancestral body sizes of many of the subclades in _Atlantogenata_ were estimated to be large, and the estimated cancer risk increases - even for the small clades like _Afroinsectivora_ may explain why these TSG gene duplications occurred early on, and persisted even in species that we did not expect to have high risks of cancer.
+However, some of our results also provide interesting evidence for a paradigm of TSG duplications being a pervasive phenomenon, rather than a specific mechanism that occurs after the evolution of gigantism.
+At the common ancestor of _Proboscidea_, which was small relative to both its ancestors and its descendants, we see the emergence of various TSG duplications, which may have enabled the stratospheric increase in body size of modern elephants.
 We also identified many TSG duplication events in smaller species and lineages such as in _Chrysochloris asiatica_ and _Elephantulus edwardii_, although these may have been the result of low-quality genomes.
 
-The impacts and takeaways of this study are limited by the quality and quantity of _Atlantogenatan_ genomes that were available, and our available knowledge of the livespans and cancer risk of the extant species.
+The impacts and takeaways of this study are limited by the quality and quantity of _Atlantogenatan_ genomes that were available, and our available knowledge of the lifespans and cancer risk of the extant species.
 For many of our species, studies in captivity are limited, and the species are relatively understudied from a longitudinal perspective, such as with the Cape Elephant Shrew and Cape Golden Mole.
-Furthermore, while there is recent interest in resequencing and improving the quality of these assemblies, at the time of this writing there is still quite a ways to go in order to have genomes of a suffiently rigorous quality to make stronger inferences about gene copy number expansions and contractions (which were not considered in this study for this reason).
+Furthermore, while there is recent interest in resequencing and improving the quality of these assemblies, at the time of this writing there is still quite a ways to go in order to have genomes of a sufficiently rigorous quality to make stronger inferences about gene copy number expansions and contractions (which were not considered in this study for this reason).
 
 The lack of a stronger signal from tumor suppressor duplications is likely a result of the strong effect size on both cancer risk and organismal toxicity that a TSG duplication would provide.
-The duplication of a tumor suppressor in many cases is associated with mild toxicity, although it greatly varies given the context and TSG in questionl; however, a single TSG duplication can also provide significant protection against cancer.
+The duplication of a tumor suppressor in many cases is associated with mild toxicity, although it greatly varies given the context and TSG in question; however, a single TSG duplication can also provide significant protection against cancer.
 For example, the overexpression of TP53 in mice, while protective of cancer, is associated with progeria and early death; however, if an additional copy of TP53 is introduced with its regulatory elements intact, the mice are healthy and experience normaging, while also demonstrating an enhanced response to cellular stress and lower rates of cancer.
-[@CITE the mouse TP53 papers].
-In light of this, it is fascinating that our results in the elephant lineage suggest that the duplication of TP53 regulators preceded the retroduplication and expansion of TP53, as this likely would have lowered the toxicity of the intial duplication and thus enabled it to occur.
-Given a suffienct selective pressure on increasing body size, it stands to reason that events like this could alleviate the negative pleiotropy of TSG duplications sufficiently to enable their persistence and allow for subsequent refinment over evolutionary time.
+[@Tyner:20021c2; @GarciaCao:20021c2].
+In light of this, it is fascinating that our results in the elephant lineage suggest that the duplication of TP53 regulators preceded the retroduplication and expansion of TP53, as this likely would have lowered the toxicity of the initial duplication and thus enabled it to occur.
+Given a sufficient selective pressure on increasing body size, it stands to reason that events like this could alleviate the negative pleiotropy of TSG duplications sufficiently to enable their persistence and allow for subsequent refinement over evolutionary time.
 
 By combining a phylogenetic study on body size in addition to a survey of copy number across nearly all protein-coding genes, we provide a comprehensive look at the question of the role of cancer risk and body size in _Atlantogenata_ that may provide broader insight to other mammalian species.
-Our study was initially motivated by the identification of functional duplicates of tumor suppressors, such as TP53 and LIF in elephants [@Abbeglen2015; @Sulak2016; @Vazquez2018].
+Our study was initially motivated by the identification of functional duplicates of tumor suppressors, such as TP53 and LIF in elephants [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018].
 Further in support of our results, a larger candidate gene study by Caulin et al.
-[@CITE] characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans.
+[@Caulin2015] characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans.
 However, while candidate gene studies are useful, by their very design they are biased in determining larger patterns of evolution of traits.
@@
 however, this is unlikely to be the only genetic mechanism at play in this scenario.
-In genome-wide studies of unusually large or long-lived species such as the bowhead whale (Keane et al., 2015), Myotid bats (Seim et al., 2013; Zhang et al., 2013), naked mole rat (Kim et al., 2011), and blind mole rat (Fang et al., 2014), there were cases of overrepresentation of TSGs among duplicate genes that were outshadowed by the identification of strong signatures of positive selection at TSGs.
-While the evolution of regulatory and coding elements of both TSGs and other non-cannonical tumor suppressor genes have been shown to be important for mediating the cancer risk of long-lived species, there has been no attention given to the possiblity of TSG duplications providing a relaxation of possible negative pleitropy that could result from these traits.
-It has been well-established in the literature that genes duplication events allow for evolutionary drift in one of the copies, which frequently results in neofunctionalization or specialization of the two copies [@ now actually support this].
-While this is beyond the scope of our study, a promising future direction of this work would include an evolutionary analysis of duplicated genes relative to each other to see if this has already occured between the pairs of genes we have identified.
+In genome-wide studies of unusually large or long-lived species such as the bowhead whale [@BowheadGenome], Myotid bats [@seim2013; @myoDav1_pteAle1], naked mole rat [@NakedMoleRatGenome], and blind mole rat [@BlindMoleRatGenome], there were cases of overrepresentation of TSGs among duplicate genes that were outshadowed by the identification of strong signatures of positive selection at TSGs.
+While the evolution of regulatory and coding elements of both TSGs and other non-canonical tumor suppressor genes have been shown to be important for mediating the cancer risk of long-lived species, there has been no attention given to the possibility of TSG duplications providing a relaxation of possible negative pleitropy that could result from these traits.
+It has been well-established in the literature that genes duplication events allow for evolutionary drift in one of the copies, which may result in neofunctionalization or specialization of the two copies [@Rastogi2005; @QianAndZhang2014; @Stoltzfus1999].
+While this is beyond the scope of our study, a promising future direction of this work would include an evolutionary analysis of duplicated genes relative to each other to see if this has already occurred between the pairs of genes we have identified.
 
 # Acknowledgements
-I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for his indispensible support.
+I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for his indispensable support.
 
@@
 Only 75% of the query sequence was covered in total between the hits, but for that 75%, each nucleotide has two hits.
-As such, ECNC is equal to 2.0 for this gene."}
+As such, ECNC is equal to 2.0 for this gene.", eval=F}
 sf1 <- list(
@@
 
-```{r, Supplementary Figure 2, fig.cap="Supplementary Figure 2: Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number"}
+```{r, Supplementary Figure 2, fig.cap="Supplementary Figure 2: Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number", eval=F}
 p.ml.tree.increasing.geneCounts.withSize.withBL
@@
 
-```{r, Supplementary Figure 3, fig.cap="Supplementary Figure 3: Full version of RecBlat strategy, but low quality", warning=T}
+```{r, Supplementary Figure 3, fig.cap="Supplementary Figure 3: Full version of RecBlat strategy, but low quality", warning=T, eval=F}
 warning("TODO: Make HQ version of this figure")
@@
 
-```{r, Supplementary Figure 4, fig.cap="Supplementary Figure 4: Dot and Line plot for Body Sizes and such"}
+```{r, Supplementary Figure 4, fig.cap="Supplementary Figure 4: Dot and Line plot for Body Sizes and such", eval=F}
 p.bodysize.atlantogenata
@@
 
-```{r Supplementary Figure 5, fig.cap="Supplementary Figure 5: All the Gene Copy Trees for interesting genes duplicated in LoxAfr4 (RIP any trees if this gets printed)"}
+```{r Supplementary Figure 5, fig.cap="Supplementary Figure 5: All the Gene Copy Trees for interesting genes duplicated in LoxAfr4 (RIP any trees if this gets printed)", eval=F}
 p.ml.treedata.loxAfrHighlight.plus[1:10]
@@
 
-```{r Supplementary Figure 6, fig.cap="Supplementary Figure 6: Expression data for interesting genes duplicated in LoxAfr4"}
+```{r Supplementary Figure 6, fig.cap="Supplementary Figure 6: Expression data for interesting genes duplicated in LoxAfr4", eval=F}
 # plot.coolgenes
@@
 
-```{r Supplementary Figure 7, fig.cap="Supplementary Figure 7: Correlation matrix heat map for genome quality metrics, ECNC, RBHB Copy, and Estimated Copy Number (lesser between ECNC and RBHB)"}
+```{r Supplementary Figure 7, fig.cap="Supplementary Figure 7: Correlation matrix heat map for genome quality metrics, ECNC, RBHB Copy, and Estimated Copy Number (lesser between ECNC and RBHB)", eval=F}
 p.cor.GCT.CEGMA
@@
 
-```{r Supplementary Figure 8, fig.cap="Supplementary Figure 8: The time-calibrated Eutherian tree used as input for StableTraits", warning=T}
+```{r Supplementary Figure 8, fig.cap="Supplementary Figure 8: The time-calibrated Eutherian tree used as input for StableTraits", warning=T, eval=F}
 p.time.tree.Eutheria <-
@@
 # Supplementary data
-
-```{r Supplementary Table 1, fig.cap="Supplementary Table 1: Parameters from StableTraits", warning=T}
+```{r Supplementary Table 1, eval=F}
+sra.data <- read_csv("../data/SraRunTable/SraRunTable.csv") %>% 
+  select(Organism, Run, source_name, sample_name) %>% 
+  filter(Organism %in% genomeTable.atlantogenata$Species) %>% 
+     mutate(sample_name = sample_name %>% str_remove("Nine-banded Armadillo_ ID 09-86_")) %>% 
+     mutate_all(. %>% str_replace_na("")) %>% 
+     unite("Tissue", source_name, sample_name, sep="") %>% 
+     filter(Tissue != "") %>% 
+     group_by(Organism) %>% 
+     summarize(Run = Run %>% unique %>% paste0(collapse=", "), 
+               Tissue = Tissue %>% str_to_title() %>% unique %>% paste0(collapse=", ")) %>% 
+     left_join(genomeTable.atlantogenata, by=c("Organism" = "Species")) %>% 
+  select(Organism, `Common Name`, Genome, `SRA Acc.`=Run, Tissues=Tissue) %>% 
+  mutate_all(. %>% str_wrap(width=100))
+sra.data %>% ggtexttable()
+```
+
+```{r Supplementary Table 2, fig.cap="Table 1: Body Size and Confidence Intervals in _Atlantogenata_ estimated using StableTraits.", eval=F}
+t.bodysize.atlantogenata.withNodes
+```
+
+```{r Supplementary Table 3, fig.cap="Supplementary Table 1: Parameters from StableTraits", warning=T, eval=F}
 warning("TODO: Read in Eutheria.progress and present it to the bold.")
@@
 
+```{r Table 4, fig.cap="Model Statistics for PGLS of Cancer Succeptibility", eval=F}
+t.size.lifespan.Atlantogenata.lm %>% cat
+```
+
 # References {#references .unnumbered}

### 2020-05-28 paper_PLOS/paper_PLOS_draft.tex
Figure 1 corrected

@@
-% Template for PLoS
-% Version 3.5 March 2018
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-% -- IMPORTANT NOTE
-%
-% This template contains comments intended
-% to minimize problems and delays during our production
-% process. Please follow the template instructions
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% Once your paper is accepted for publication,
-% PLEASE REMOVE ALL TRACKED CHANGES in this file
-% and leave only the final text of your manuscript.
-% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
-% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
-%
-%
-% There are no restrictions on package use within the LaTeX files except that
-% no packages listed in the template may be deleted.
-%
-% Please do not include colors or graphics in the text.
-%
-% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file.
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% For figure citations, please use "Fig" instead of "Figure".
-% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - spacing/line breaks within cells to alter layout or alignment
-% - do not nest tabular environments (no tabular environments within tabular environments)
-% - no graphics or colored text (cell background color/shading OK)
-% See http://journals.plos.org/plosone/s/tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
-%
-% For inline equations, please be sure to include all portions of an equation in the math environment.
-%
-% Do not include text that is not math in the math environment.
-%
-% Please add line breaks to long display equations when possible in order to fit size of the column.
-%
-% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
-%
-% When adding superscript or subscripts outside of brackets/braces, please group using {}.
-%
-% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% Please contact latex@plos.org with any questions.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
+\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
 
-\documentclass[10pt,letterpaper]{article}
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+\usepackage{lineno} % add
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{amsmath,amssymb}
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+%%%%%%%%%%%%%%%% end my additions to header
 
-% Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\bibliographystyle{elsarticle-harv}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans},
+            colorlinks=false,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
 
-% Use Unicode characters when possible
-\usepackage[utf8x]{inputenc}
+\setcounter{secnumdepth}{0}
+% Pandoc toggle for numbering sections (defaults to be off)
+\setcounter{secnumdepth}{0}
 
-% textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
-
-% cite package, to clean up citations in the main text. Do not remove.
-% \usepackage{cite}
-
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
-
-% line numbers
-\usepackage[right]{lineno}
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
-
-% array package and thick rules for tables
-\usepackage{array}
-
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
-
-% create \thickcline for thick horizontal lines of variable length
-\newlength\savedwidth
-\newcommand\thickcline[1]{%
-  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-  \cline{#1}%
-  \noalign{\vskip\arrayrulewidth}%
-  \noalign{\global\arrayrulewidth\savedwidth}%
-}
-
-% \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-\hline
-\noalign{\global\arrayrulewidth\savedwidth}}
-
-% Remove comment for double spacing
-%\usepackage{setspace}
-%\doublespacing
-
-% Text layout
-\raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in
-\textheight 8.75in
-
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\renewcommand{\figurename}{Fig}
-
-% Use the PLoS provided BiBTeX style
-% \bibliographystyle{plos2015}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
-%\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-%\setlength{\headheight}{27.023pt}
-%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\today}
-
-%% Include all macros below
-
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
-
+% Pandoc header
 \usepackage{longtable}
@@
 
-\usepackage{forarray}
-\usepackage{xstring}
-\newcommand{\getIndex}[2]{
-  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
-}
+\begin{document}
+\begin{frontmatter}
 
-\setcounter{secnumdepth}{0}
+  \title{Pervasive duplication of tumor suppressor genes preceded parallel
+evolution of large bodied Atlantogenatans}
+    \author[University of Chicago]{Juan Manuel Vazquez}
 
-\newcommand{\getAff}[1]{
-  \getIndex{#1}{University of Chicago,University at Buffalo, SUNY}
-}
+    \author[SUNY Buffalo]{Vincent J Lynch}
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+      \address[University of Chicago]{Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637}
+    \address[SUNY Buffalo]{Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260}
 
-\begin{document}
-\vspace*{0.2in}
-
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{Pervasive duplication of tumor suppressor genes preceded parallel
-evolution of large bodied Atlantogenatans} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
-}
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Juan Manuel Vazquez\textsuperscript{\getAff{University of Chicago}}\textsuperscript{*},
-Vincent J Lynch\textsuperscript{\getAff{University at Buffalo, SUNY}}\\
-\bigskip
-\textbf{\getAff{University of Chicago}}Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637\\
-\textbf{\getAff{University at Buffalo, SUNY}}Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260\\
-\bigskip
-* Corresponding author: juanvazquez@uchicago.edu\\
-\end{flushleft}
-% Please keep the abstract below 300 words
-\section*{Abstract}
+  \begin{abstract}
   Cancer is an intrinsic disease of multicellular organisms. Within a
@@
   increased body size by compensating for the increased cancer risk.
+  \end{abstract}
 
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step.
-% Author Summary not valid for PLOS ONE submissions.
-\section*{Author summary}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
-porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
-dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
-velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
-finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
-diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
-ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
-pulvinar, volutpat risus.
+ \end{frontmatter}
 
-\linenumbers
-
-% Use "Eq" instead of "Equation" for equation citations.
 )
@@
 incidence increases with increasing adult height {[}1,2{]} and cancer
-incidence is positively correlated with body size in dogs
-{[}{\textbf{???}},3{]}. There is no correlation, however, between body
-size and cancer risk between species. This lack of correlation is often
-referred to as `Peto's Paradox' {[}4--6{]}. While it is clear that a
-resolution to Peto's Paradox must involve the evolution of enhanced
-cancer protection alongside increases in body size and lifespan, the
-specific genetic, molecular, and cellular mechanisms that underlie this
-resistance have proven elusive. {[}7--11{]}.
+incidence is positively correlated with body size in dogs {[}3,4{]}.
+There is no correlation, however, between body size and cancer risk
+between species. This lack of correlation is often referred to as
+`Peto's Paradox' {[}5--7{]}. While it is clear that a resolution to
+Peto's Paradox must involve the evolution of enhanced cancer protection
+alongside increases in body size and lifespan, the specific genetic,
+molecular, and cellular mechanisms that underlie this resistance have
+proven elusive. {[}8--12{]}.
 
@@
 \textasciitilde{}170g in elephant shrews, \textasciitilde{}3kg in
-hyraxes, and 60kg in aardvarks {[}12{]}. However, while these extant
+hyraxes, and 60kg in aardvarks {[}13{]}. However, while these extant
 species are relatively small, the fossil evidence demonstrates that
@@
 extant hyraxes are relatively small, the extinct Titanohyrax is
-estimated to have weighted up to \textasciitilde{}1300kg {[}13{]}. The
+estimated to have weighted up to \textasciitilde{}1300kg {[}14{]}. The
 largest members of Afrotheria, too, are dwarfed by the size of their
@@
 extinct Stellar's sea cow which is estimated to have weight 8000-10000kg
-{[}14{]}. Similarly African (4,800kg) and Asian elephants (3,200kg) are
+{[}15{]}. Similarly African (4,800kg) and Asian elephants (3,200kg) are
 the largest living elephant species, but are dwarfed by the truly
@@
 straight-tusked elephant (\textasciitilde{}220,000kg), the largest known
-land mammal {[}15{]}. Remarkably these large-bodied Afrotherian lineages
-are nested within small bodied species (Fig. 1) {[}16--19{]}, indicating
-that gigantism independently evolved in hyraxes, sea cows, and elephants
+land mammal {[}16{]}. These large-bodied Afrotherian lineages are nested
+within small bodied species (Fig. 1) {[}17--20{]}, indicating that
+gigantism independently evolved in hyraxes, sea cows, and elephants
 (Paenungulates). Thus, Paenungulates are an excellent model system in
@@
 
-Although many mechanisms can potentially resolve Peto's paradox, the
-most parsimonious route to enhanced cancer resistance is likely through
-an increased copy number of tumor suppressors. Such an example has been
+Although many mechanisms can potentially resolve Peto's paradox, among
+the most parsimonious routes to enhanced cancer resistance is through an
+increased copy number of tumor suppressors. Such an example has been
 seen in the case of candidate genes such as \emph{TP53} and \emph{LIF}
-{[}11,20,21{]} as well as in studies involving a limited set of
-candidate genes {[}22,23{]}. As these studies focus on \emph{a priori}
+{[}12,21,22{]} as well as in studies involving a limited set of
+candidate genes {[}23,24{]}. As these studies focus on \emph{a priori}
 gene sets, however, it remains unknown whether this is a general,
@@
 suppression pathways. Our estimates of the evolution of body mass,
-similarly to previous studies {[}16--19{]}, show that large body masses
+similarly to previous studies {[}17--20{]}, show that large body masses
 evolved in a step-wise manner, with major increases in body mass in the
@@
 a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene
-duplications in Afrotherian genomes, and used parsimony to infer the
-lineages in which those duplications occurred. We found gene
-duplications in lineages with increased body mass were enriched in
-functions related to tumor suppression, including regulation of the cell
-cycle, DNA damage repair, and regulation of apoptosis. These data
-suggest that duplication of tumor suppressors played a role in the
-evolution of large, long-lived in Afrotherians.
+duplications in Afrotherian genomes, and used maximum likelihood
+(treating copy number as a discrete trait) to infer the lineages in
+which those duplications occurred. We found gene duplications in
+lineages with increased body mass were enriched in functions related to
+tumor suppression, including regulation of the cell cycle, DNA damage
+repair, and regulation of apoptosis. These data suggest that duplication
+of tumor suppressors played a role in the evolution of large, long-lived
+in Afrotherians.
 
@@
 the time-calibrated molecular phylogeny of Bininda-Emonds \emph{et al.}
-{[}24{]} with the time-calibrated total evidence Afrotherian phylogeny
-from Puttick and Thomas {[}{\textbf{???}}{]}. While the Bininda-Emonds
-\emph{et al.} {[}24{]} phylogeny includes 1,679 species, only 34 are
-Afrotherian, and no fossil data are included. The inclusion of fossil
-data from extinct species is essential to ensure that ancestral state
-reconstructions of body mass are not biased by only including extant
-species. This can lead to inaccurate reconstructions, for example, if
-lineages convergently evolved large body masses from a small bodied
-ancestor. In contrast, the total evidence Afrotherian phylogeny of
-Puttick and Thomas {[}19{]} includes 77 extant species and fossil data
-from 39 extinct species. Therefore we replaced the Afrotherian clade in
-the Bininda-Emonds \emph{et al.} {[}24{]} phylogeny with the Afrotherian
-phylogeny of Puttick and Thomas {[}19{]} using Mesquite. Next, we
-jointly estimated rates of body mass evolution and reconstructed
-ancestral states using a generalization of the Brownian motion model
-that relaxes assumptions of neutrality and gradualism by considering
-increments to evolving characters to be drawn from a heavy-tailed stable
-distribution (the ``Stable Model'') {[}25{]}. The stable model allows
-for occasional large jumps in traits and has previously been shown to
-out-perform other models of body mass evolution, including standard
-Brownian motion models, Ornstein--Uhlenbeck models, early burst maximum
-likelihood models, and heterogeneous multi-rate models {[}25{]}.
+{[}25{]} with the time-calibrated total evidence Afrotherian phylogeny
+from Puttick and Thomas {[}20{]}. While the Bininda-Emonds \emph{et al.}
+{[}25{]} phylogeny includes 1,679 species, only 34 are Afrotherian, and
+no fossil data are included. The inclusion of fossil data from extinct
+species is essential to ensure that ancestral state reconstructions of
+body mass are not biased by only including extant species. This can lead
+to inaccurate reconstructions, for example, if lineages convergently
+evolved large body masses from a small bodied ancestor. In contrast, the
+total evidence Afrotherian phylogeny of Puttick and Thomas {[}20{]}
+includes 77 extant species and fossil data from 39 extinct species.
+Therefore we replaced the Afrotherian clade in the Bininda-Emonds
+\emph{et al.} {[}25{]} phylogeny with the Afrotherian phylogeny of
+Puttick and Thomas {[}20{]} using Mesquite. Next, we jointly estimated
+rates of body mass evolution and reconstructed ancestral states using a
+generalization of the Brownian motion model that relaxes assumptions of
+neutrality and gradualism by considering increments to evolving
+characters to be drawn from a heavy-tailed stable distribution (the
+``Stable Model'') {[}26{]}. The stable model allows for occasional large
+jumps in traits and has previously been shown to out-perform other
+models of body mass evolution, including standard Brownian motion
+models, Ornstein--Uhlenbeck models, early burst maximum likelihood
+models, and heterogeneous multi-rate models {[}26{]}.
 
@@
 \textbf{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit
-BLAT (RBHB) pipeline to quickly identify homologs and estimate gene copy
-numbers (\textbf{Figure 1A}). The Reciprocal Best Hit (RBH) search
-strategy is conceptually straightforward: 1) Given a gene of interest
-\(G_A\) in a query genome \(A\), one searches a target genome \(B\) for
-all possible matches to \(G_A\); 2) For each of these hits, one then
-performs the reciprocal search in the original query genome to identify
-the highest-scoring hit; 3) A hit in genome \(B\) is defined as a
-homolog of gene \(G_A\) if and only if the original gene \(G_A\) is the
-top reciprocal search hit in genome \(A\). We selected BLAT {[}26{]} as
-our algorithm of choice, as this algorithm is sensitive to highly
-simliar (\textgreater{}90\% identity) sequences, thus identifying the
-highest-confidence homologs while minimizing many-to-one mapping
-problems when searching for multiple genes. RBH performs similar to
-other more complex methods of orthology prediction, and is particularly
-good at identifying incomplete genes that may be fragmented in low
-quality/poor assembled regions of the genome {[}{\textbf{???}},27{]}.
+BLAT (RBHB) pipeline to identify putative homologs and estimate gene
+copy numbers across species (\textbf{Figure 1A}). The Reciprocal Best
+Hit (RBH) search strategy is conceptually straightforward: 1) Given a
+gene of interest \(G_A\) in a query genome \(A\), one searches a target
+genome \(B\) for all possible matches to \(G_A\); 2) For each of these
+hits, one then performs the reciprocal search in the original query
+genome to identify the highest-scoring hit; 3) A hit in genome \(B\) is
+defined as a homolog of gene \(G_A\) if and only if the original gene
+\(G_A\) is the top reciprocal search hit in genome \(A\). We selected
+BLAT {[}27{]} as our algorithm of choice, as this algorithm is sensitive
+to highly simliar (\textgreater{}90\% identity) sequences, thus
+identifying the highest-confidence homologs while minimizing many-to-one
+mapping problems when searching for multiple genes. RBH performs similar
+to other more complex methods of orthology prediction, and is
+particularly good at identifying incomplete genes that may be fragmented
+in low quality/poor assembled regions of the genome {[}28,29{]}.
 
@@
 
-\textbf{RecSearch Pipeline} We created a custom Python pipeline for
+\textbf{RecSearch Pipeline:} We created a custom Python pipeline for
 automating RBHB searches between a single reference genome and multiple
@@
 genome. For the query sequences in our search, we used the hg38 Proteome
-provided by UniProt {[}28{]}, which is a comprehensive set of protein
+provided by UniProt {[}30{]}, which is a comprehensive set of protein
 sequences curated from a combination of predicted and validated protein
@@
 such as zinc fingers, protocadherins, and transposon-containing genes,
-as these were prone to high levels of false positive hits. After
-filtering out problematic protein queries, we then used our pipeline
-(Figure 1A) to search for all copies of our 18011 query genes in
-publicly available Afrotherian genomes, including African savannah
-elephant (\emph{Loxodonta africana}: loxAfr3, loxAfr4, loxAfrC), African
-forest elephant (\emph{Loxodonta cyclotis}: loxCycF), Asian Elephant
+as these were prone to high levels of false positive hits.
+
+After filtering out problematic protein queries (see ``Query gene
+inclusion criteria''), we then used our pipeline (Figure 1A) to search
+for all copies of our 18011 query genes in publicly available
+Afrotherian genomes, including African savannah elephant
+(\emph{Loxodonta africana}: loxAfr3, loxAfr4, loxAfrC), African forest
+elephant (\emph{Loxodonta cyclotis}: loxCycF), Asian Elephant
 (\emph{Elephas maximus}: eleMaxD), Woolly Mammoth (\emph{Mammuthus
@@
 
-\textbf{Query protein criteria} To assemble our query list, we first
-removed all unnamed genes from UP000005640. Next, we excluded genes from
-downstream analyses for which assignment of homology was uncertain,
-including uncharacterized ORFs (991), LOC (63), HLA genes (402),
-replication dependent histones (72), odorant receptors (499), ribosomal
-proteins (410), zinc finger transcription factors (1983), viral and
-repetitive-element-associated proteins (82) and any protein described as
-either ``Uncharacterized,'' ``Putative,'' or ``Fragment'' by UniProt in
-UP000005640 (30724), leaving us with a final set of 37582 query protein
-sequences, corresponding to 18011 genes.
+\textbf{Query gene inclusion criteria:} To assemble our query list, we
+first removed all unnamed genes from UP000005640. Next, we excluded
+genes from downstream analyses for which assignment of homology was
+uncertain, including uncharacterized ORFs (991), LOC (63), HLA genes
+(402), replication dependent histones (72), odorant receptors (499),
+ribosomal proteins (410), zinc finger transcription factors (1983),
+viral and repetitive-element-associated proteins (82) and any protein
+described as either ``Uncharacterized,'' ``Putative,'' or ``Fragment''
+by UniProt in UP000005640 (30724), leaving us with a final set of 37582
+query protein sequences, corresponding to 18011 genes.
 
-\textbf{Duplication gene inclusion criteria} In order to condense
+\textbf{Duplication gene inclusion criteria:} In order to condense
 transcript-level hits into single gene loci, and to resolve many-to-one
@@
 transcripts searched was 1.5 or greater; previous studies have shown
-that incomplete duplications can encode functional genes, therefore
-partial gene duplications were included provided they passed additional
-inclusion criteria. The ECNC cut off of 1.5 was selected empirically, as
-this value minimized the number of false positives seen in a test set of
-genes and genomes. The results of our initial search are summarized in
-Figure 1B. Overall, we identified 13880 genes across all species, or
-77.1\% of our starting query genes.
+that incomplete duplications can encode functional genes {[}12,22{]},
+therefore partial gene duplications were included provided they passed
+additional inclusion criteria. The ECNC cut off of 1.5 was selected
+empirically, as this value minimized the number of false positives seen
+in a test set of genes and genomes. The results of our initial search
+are summarized in Figure 1B. Overall, we identified 13880 genes across
+all species, or 77.1\% of our starting query genes.
 
-\textbf{Genome Quality Assessment using CEGMA} In order to determine the
-effect of genome quality on our results, we used the gVolante webserver
-and CEGMA to assess the quality and completeness of the genome
-{[}{\textbf{???}},{\textbf{??}}{]}. CEGMA was run using the default
-settings for mammals (``Cut-off length for sequence statistics and
-composition'' = 1;``CEGMA max intron length'' = 100000; ``CEGMA gene
-flanks'' = 10000, ``Selected reference gene set'' = CVG). For each
-genome, we generated a correlation matrix using the aforementioned
-genome quality scores, and either the mean Copy Number or mean ECNC for
-all hits in the genome.
+\textbf{Genome Quality Assessment using CEGMA:} In order to determine
+the effect of genome quality on our results, we used the gVolante
+webserver and CEGMA to assess the quality and completeness of the genome
+{[}31,32{]}. CEGMA was run using the default settings for mammals
+(``Cut-off length for sequence statistics and composition'' = 1;``CEGMA
+max intron length'' = 100000; ``CEGMA gene flanks'' = 10000, ``Selected
+reference gene set'' = CVG). For each genome, we generated a correlation
+matrix using the aforementioned genome quality scores, and either the
+mean Copy Number or mean ECNC for all hits in the genome.
 
@@
 
-To validate and filter out RBHB results, we intersected our results with
-either gene prediction or transcriptomic evidence as a proxy for
-functionality.
+To validate and filter out duplicate gene calls, we intersected our
+results with either gene prediction or transcriptomic evidence as a
+proxy for functionality.
 
@@
 \emph{de novo} transcriptomes using publically-available RNA-sequencing
-data from NCBI SRA. We mapped reads to all genomes available for each
-species, and assembled transcripts using HISAT2 and StringTie,
-respectively {[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}.
-RNA-sequencing data was not available for Cape Golden Mole, Cape
-Elephant Shrew, Rock Hyrax, Aardvark, or the Lesser Hedgehog Tenrec.
+data from NCBI SRA (Supplementary Table 1). We mapped reads to all
+genomes available for each species, and assembled transcripts using
+HISAT2 and StringTie, respectively {[}33--35{]}. RNA-sequencing data was
+not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax,
+Aardvark, or the Lesser Hedgehog Tenrec.
 
@@
 duplicate hits identified in each genome with the gene prediction tracks
-and/or transcriptome assemblies using \texttt{bedtools}
-{[}{\textbf{???}}{]}. When multiple lines of evidence for functionality
-were present for a genome, we used the union of all intersections as the
-final output for evidenced duplicates. When analyzing the
-highest-quality assemblies available for each species, if a species had
-neither gene prediction tracks nor RNA-seq data for the highest-quality
-genome available, we conservatively included all hits for the genome in
-the final set of evidenced duplicates.
+and/or transcriptome assemblies using \texttt{bedtools}. When multiple
+lines of evidence for functionality were present for a genome, we used
+the union of all intersections as the final output for evidenced
+duplicates. When analyzing the highest-quality assemblies available for
+each species, if a species had neither gene prediction tracks nor
+RNA-seq data for the highest-quality genome available, we conservatively
+included all hits for the genome in the final set of evidenced
+duplicates.
 
@@
 copy numbers of genes in \emph{Atlantogenata} using IQ-Tree. For this
-analysis, we used an unrooted subset of our prior species tree,
-including only the aforementioned \emph{Atlantogenata} species. We
-generated PHYLIP files containing the copy number of each gene in the
-highest quality genome for each species, encoding genes on a scale from
-1-31+ copies as 1-9, A-V; and encoding a gene's copy number as uncetain
-(``?'') when we did not identify it in the genome. We used the included
-tree-searching and model-testing functionality in IQ-Tree to determine
-the most likely topology for the species tree, and to obtain the most
-likely model for copy number changes in the genome. We defined the
-ancestral state of a node if it had greater than an 80\% posterior
-probability.
+analysis, we used an unrooted species tree including only the
+aforementioned \emph{Atlantogenata} species. We generated PHYLIP files
+containing the copy number of each gene in the highest quality genome
+for each species, encoding genes on a scale from 1-31+ copies as 1-9,
+A-V; and encoding a gene's copy number as uncetain (``?'') when we did
+not identify it in the genome. We used the included tree-searching and
+model-testing functionality in IQ-Tree to determine the most likely
+topology for the species tree, and to obtain the most likely model for
+copy number changes in the genome. The most likely model for the
+evolution of copy number was inferred to be a Jukes-Cantor type model
+for morphological data with equal state frequencies, allowing for a
+proportion of invariable sites, and using a discrete Gamma model with 4
+rate categories. (``MK+FQ+I+G4''). We defined the ancestral state of a
+node if it had greater than an 80\% posterior probability.
 
@@
 overrepresentation analysis (ORA) of the duplicated gene lists relative
-to our initial query gene list {[}{\textbf{???}}{]}. For the database of
-pathways used in the analysis, we used Reactome {[}{\textbf{???}}{]},
-Wikipathways, and Wikipathways\_cancer {[}{\textbf{???}}{]}, and KEGG
-{[}{\textbf{???}}{]}. For the ORA, we used FDR for determining
+to our initial query gene list {[}36{]}. For the database of pathways
+used in the analysis, we used Reactome {[}37{]}, Wikipathways, {[}38{]},
+and KEGG {[}39{]}. For the ORA, we used FDR for determining
 significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and
@@
 needed to calcultate ancestral lifespans at each node. To do so, we used
-a Phylogenetic Generalized Least-Square Regression (PGLS)
-{[}{\textbf{???}},{\textbf{??}}{]} to calculate estimated ancestral
-lifespans across \emph{Atlantogenata} using our estimates for body size
-at each node.
+a Phylogenetic Generalized Least-Square Regression (PGLS) {[}40,41{]} to
+calculate estimated ancestral lifespans across \emph{Atlantogenata}
+using our estimates for body size at each node.
 
 Next, we used a simplified multistage cancer risk model for body size
-\(D\) and lifespan \(t\): \(K \approx Dt^6\)
-{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. To determine the
-change in \(K\) between nodes, we obtained the ratio between the cancer
-risk \(K_{1}\) at any given node, and the cancer risk \(K_{2}\) at its
-ancestral node, using the equation
+\(D\) and lifespan \(t\): \(K \approx Dt^6\) {[}7,42--44{]}. To
+determine the change in \(K\) between nodes, we obtained the ratio
+between the cancer risk \(K_{1}\) at any given node, and the cancer risk
+\(K_{2}\) at its ancestral node, using the equation
 \(\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}\).
@@
 
-\textbackslash{}begin\{figure\}
+\begin{figure}
 
-\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 1-1}
-\hfill{}
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure1-1} \hfill{}
 
-\textbackslash{}caption\{Figure 1: Body sizes rapidly and frequently
-expand in Eutherians, especially in Atlantogenata. \textbf{A)} Tree of
-Eutherian species, colored by ln(Body Size) and with branch lengths set
-to the rate of change in body sizes, normalized by the square root of
-the root branch. Atlantogenata is highlighted at the bottom. \textbf{B)}
-Zoom-in of (A) on Atlantogenata. Silhuetes for the African Elephant,
-West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape
-Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are
-colored by their extant body sizes, while clade labels are colored based
-on the common ancestor's estimated body size\}\label{fig:Figure 1}
-\textbackslash{}end\{figure\}
-
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{A table generated by the longtable package.}\tabularnewline
-\toprule
-Node & Size (log(g)) & 95\% CI (Low) & 95\% CI (High) & Rate
-(sqrt)\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Node & Size (log(g)) & 95\% CI (Low) & 95\% CI (High) & Rate
-(sqrt)\tabularnewline
-\midrule
-\endhead
-Cryptochloris wintoni & 3.13 & 3.13 & 3.13 & 5.78\tabularnewline
-Amblysomus marleyi & 3.53 & 3.53 & 3.53 & 3.79\tabularnewline
-Elephantulus revoili & 3.48 & 3.48 & 3.48 & 1.10\tabularnewline
-Titanohyrax andrewsi & 12.97 & 12.97 & 12.97 & 0.07\tabularnewline
-Titanohyrax ultimus & 14.08 & 14.08 & 14.08 & 34.61\tabularnewline
-Megalohyrax sp nov & 12.52 & 12.52 & 12.52 & 7.21\tabularnewline
-Elephas maximus asurus & 15.66 & 15.66 & 15.66 & 0.34\tabularnewline
-Protenrec tricuspis & 1.14 & 1.14 & 1.14 & 69.75\tabularnewline
-Microgale parvula & 1.16 & 1.16 & 1.16 & 33.46\tabularnewline
-Microgale pusilla & 1.25 & 1.25 & 1.25 & 34.31\tabularnewline
-Geogale aurita & 1.90 & 1.90 & 1.90 & 40.07\tabularnewline
-Microgale longicaudata & 2.09 & 2.09 & 2.09 & 0.77\tabularnewline
-Microgale brevicaudata & 2.19 & 2.19 & 2.19 & 0.60\tabularnewline
-Microgale jobihely & 2.30 & 2.30 & 2.30 & 1.07\tabularnewline
-Microgale principula & 2.32 & 2.32 & 2.32 & 0.17\tabularnewline
-Dilambdogale gheerbranti & 2.38 & 2.38 & 2.38 & 2.21\tabularnewline
-Microgale taiva & 2.47 & 2.47 & 2.47 & 0.13\tabularnewline
-Microgale cowani & 2.62 & 2.62 & 2.62 & 0.57\tabularnewline
-Eremitalpa granti & 3.14 & 3.14 & 3.14 & 9.65\tabularnewline
-Calcochloris obtusirostris & 3.27 & 3.27 & 3.27 & 13.38\tabularnewline
-Neamblysomus julianae & 3.33 & 3.33 & 3.33 & 5.72\tabularnewline
-Chlorotalpa duthieae & 3.38 & 3.38 & 3.38 & 0.32\tabularnewline
-Chlorotalpa sclateri & 3.54 & 3.54 & 3.54 & 0.09\tabularnewline
-Macroscelides proboscideus & 3.64 & 3.64 & 3.64 & 14.17\tabularnewline
-Chrysochloris stuhlmanni & 3.74 & 3.74 & 3.74 & 0.33\tabularnewline
-Oryzorictes hova & 3.79 & 3.79 & 3.79 & 22.77\tabularnewline
-Elephantulus myurus & 3.81 & 3.81 & 3.81 & 0.95\tabularnewline
-Elephantulus brachyrhynchus & 3.81 & 3.81 & 3.81 & 0.93\tabularnewline
-Elephantulus rozeti & 3.81 & 3.81 & 3.81 & 10.51\tabularnewline
-Elephantulus fuscus & 3.82 & 3.82 & 3.82 & 0.68\tabularnewline
-Elephantulus intufi & 3.82 & 3.82 & 3.82 & 1.15\tabularnewline
-Microgale talazaci & 3.88 & 3.88 & 3.88 & 61.40\tabularnewline
-Chrysochloris asiatica & 3.89 & 3.89 & 3.89 & 3.34\tabularnewline
-Elephantulus edwardii & 3.90 & 3.90 & 3.90 & 0.24\tabularnewline
-Carpitalpa arendsi & 3.94 & 3.94 & 3.94 & 0.45\tabularnewline
-Amblysomus corriae & 3.94 & 3.94 & 3.94 & 0.98\tabularnewline
-Amblysomus hottentotus & 3.98 & 3.98 & 3.98 & 0.02\tabularnewline
-Elephantulus fuscipes & 4.04 & 4.04 & 4.04 & 1.93\tabularnewline
-Elephantulus rufescens & 4.05 & 4.05 & 4.05 & 0.12\tabularnewline
-Neamblysomus gunningi & 4.09 & 4.09 & 4.09 & 3.26\tabularnewline
-Elephantulus rupestris & 4.12 & 4.12 & 4.12 & 0.32\tabularnewline
-Amblysomus septentrionalis & 4.23 & 4.23 & 4.23 & 0.52\tabularnewline
-Chambius kasserinensis & 4.27 & 4.27 & 4.27 & 11.84\tabularnewline
-Amblysomus robustus & 4.33 & 4.33 & 4.33 & 1.38\tabularnewline
-Micropotamogale lamottei & 4.36 & 4.36 & 4.36 & 2.82\tabularnewline
-Echinops telfairi & 4.47 & 4.47 & 4.47 & 7.75\tabularnewline
-Limnogale mergulus & 4.52 & 4.52 & 4.52 & 121.95\tabularnewline
-Hemicentetes semispinosus & 4.75 & 4.75 & 4.75 & 4.68\tabularnewline
-Chrysospalax villosus & 4.77 & 4.77 & 4.77 & 0.13\tabularnewline
-Petrodromus tetradactylus & 5.29 & 5.29 & 5.29 & 24.61\tabularnewline
-Herodotius pattersoni & 5.50 & 5.50 & 5.50 & 11.64\tabularnewline
-Setifer setosus & 5.61 & 5.61 & 5.61 & 12.52\tabularnewline
-Rhynchocyon cirnei & 5.86 & 5.86 & 5.86 & 3.30\tabularnewline
-Metoldobotes sp nov & 5.93 & 5.93 & 5.93 & 15.94\tabularnewline
-Chrysospalax trevelyani & 6.13 & 6.13 & 6.13 & 62.84\tabularnewline
-Rhynchocyon petersi & 6.15 & 6.15 & 6.15 & 2.13\tabularnewline
-Rhynchocyon chrysopygus & 6.28 & 6.28 & 6.28 & 0.40\tabularnewline
-Potamogale velox & 6.49 & 6.49 & 6.49 & 103.04\tabularnewline
-Rhynchocyon udzungwensis & 6.57 & 6.57 & 6.57 & 4.33\tabularnewline
-Tenrec ecaudatus & 6.75 & 6.75 & 6.75 & 79.50\tabularnewline
-Dasypus sabanicola & 7.05 & 7.05 & 7.05 & 12.18\tabularnewline
-Tolypeutes matacus & 7.11 & 7.11 & 7.11 & 15.96\tabularnewline
-Dasypus septemcinctus & 7.30 & 7.30 & 7.30 & 4.44\tabularnewline
-Zaedyus pichiy & 7.31 & 7.31 & 7.31 & 5.54\tabularnewline
-Dasypus hybridus & 7.31 & 7.31 & 7.31 & 4.05\tabularnewline
-Chaetophractus villosus & 7.61 & 7.61 & 7.61 & 0.42\tabularnewline
-Chaetophractus nationi & 7.67 & 7.67 & 7.67 & 0.09\tabularnewline
-Heterohyrax brucei & 7.78 & 7.78 & 7.78 & 1.64\tabularnewline
-Cabassous centralis & 7.92 & 7.92 & 7.92 & 0.25\tabularnewline
-Seggeurius amourensis & 7.98 & 7.98 & 7.98 & 2.82\tabularnewline
-Procavia capensis & 8.01 & 8.01 & 8.01 & 0.00\tabularnewline
-Dendrohyrax dorsalis & 8.06 & 8.06 & 8.06 & 1.86\tabularnewline
-Microhyrax lavocati & 8.13 & 8.13 & 8.13 & 0.73\tabularnewline
-Bradypus tridactylus & 8.23 & 8.23 & 8.23 & 0.48\tabularnewline
-Bradypus torquatus & 8.27 & 8.27 & 8.27 & 0.03\tabularnewline
-Dasypus novemcinctus & 8.37 & 8.37 & 8.37 & 14.73\tabularnewline
-Euphractus sexcinctus & 8.43 & 8.43 & 8.43 & 14.99\tabularnewline
-Choloepus hoffmanni & 8.47 & 8.47 & 8.47 & 0.32\tabularnewline
-Bradypus variegatus & 8.49 & 8.49 & 8.49 & 0.51\tabularnewline
-Tamandua tetradactyla & 8.52 & 8.52 & 8.52 & 10.44\tabularnewline
-Cyclopes didactylus & 8.53 & 8.53 & 8.53 & 2.15\tabularnewline
-Choloepus didactylus & 8.71 & 8.71 & 8.71 & 0.64\tabularnewline
-Thyrohyrax meyeri & 8.78 & 8.78 & 8.78 & 3.55\tabularnewline
-Saghatherium bowni & 9.13 & 9.13 & 9.13 & 15.85\tabularnewline
-Dasypus kappleri & 9.23 & 9.23 & 9.23 & 74.13\tabularnewline
-Thyrohyrax domorictus & 9.30 & 9.30 & 9.30 & 1.15\tabularnewline
-Dimaitherium patnaiki & 9.57 & 9.57 & 9.57 & 18.23\tabularnewline
-Phosphatherium escuilliei & 9.62 & 9.62 & 9.62 & 326.23\tabularnewline
-Saghatherium antiquum & 9.73 & 9.73 & 9.73 & 2.90\tabularnewline
-Thyrohyrax litholagus & 10.01 & 10.01 & 10.01 & 28.58\tabularnewline
-Myrmecophaga tridactyla & 10.26 & 10.26 & 10.26 & 41.03\tabularnewline
-Myorycteropus africanus & 10.27 & 10.27 & 10.27 & 0.57\tabularnewline
-Selenohyrax chatrathi & 10.73 & 10.73 & 10.73 & 14.99\tabularnewline
-Priodontes maximus & 10.82 & 10.82 & 10.82 & 268.43\tabularnewline
-Orycteropus afer & 10.87 & 10.87 & 10.87 & 6.59\tabularnewline
-Antilohyrax pectidens & 10.93 & 10.93 & 10.93 & 13.69\tabularnewline
-Bunohyrax fajumensis & 11.32 & 11.32 & 11.32 & 1.45\tabularnewline
-Afrohyrax championi & 11.32 & 11.32 & 11.32 & 0.19\tabularnewline
-Geniohyus mirus & 11.33 & 11.33 & 11.33 & 5.44\tabularnewline
-Prorastomus sirenoides & 11.49 & 11.49 & 11.49 & 13.61\tabularnewline
-Elephas antiquus falconeri & 11.51 & 11.51 & 11.51 & 6.12\tabularnewline
-Pachyhyrax crassidentatus & 11.81 & 11.81 & 11.81 & 2.29\tabularnewline
-Megalohyrax eocaenus & 11.95 & 11.95 & 11.95 & 0.24\tabularnewline
-Elephas cypriotes & 12.21 & 12.21 & 12.21 & 1.90\tabularnewline
-Bunohyrax major & 12.36 & 12.36 & 12.36 & 11.39\tabularnewline
-Titanohyrax angustidens & 12.48 & 12.48 & 12.48 & 0.04\tabularnewline
-Daouitherium rebouli & 12.80 & 12.80 & 12.80 & 0.74\tabularnewline
-Arcanotherium savagei & 12.89 & 12.89 & 12.89 & 7.29\tabularnewline
-Dugong dugon & 12.92 & 12.92 & 12.92 & 5.85\tabularnewline
-Trichechus senegalensis & 13.03 & 13.03 & 13.03 & 0.57\tabularnewline
-Trichechus inunguis & 13.08 & 13.08 & 13.08 & 0.69\tabularnewline
-Protosiren smithae & 13.20 & 13.20 & 13.20 & 33.69\tabularnewline
-Numidotherium koholense & 13.23 & 13.23 & 13.23 & 2.29\tabularnewline
-Omanitherium dhofarensis & 13.35 & 13.35 & 13.35 & 0.03\tabularnewline
-Trichechus manatus & 13.44 & 13.44 & 13.44 & 1.39\tabularnewline
-Moeritherium spp & 13.82 & 13.82 & 13.82 & 5.71\tabularnewline
-Phiomia spp & 13.89 & 13.89 & 13.89 & 3.64\tabularnewline
-Elephas maximus & 15.02 & 15.02 & 15.02 & 5.81\tabularnewline
-Barytherium spp & 15.20 & 15.20 & 15.20 & 73.58\tabularnewline
-Mammuthus primigenius & 15.27 & 15.27 & 15.27 & 2.17\tabularnewline
-Mammut borsoni & 16.49 & 16.49 & 16.49 & 15.33\tabularnewline
-Mammuthus trogontherii & 16.38 & 16.38 & 16.38 & 16.00\tabularnewline
-Loxodonta africana & 15.35 & 15.35 & 15.35 & 1.28\tabularnewline
-Loxodonta cyclotis & 15.37 & 15.37 & 15.37 & 3.72\tabularnewline
-Palaeoloxodon antiquus & 16.14 & 16.14 & 16.14 & 0.01\tabularnewline
-Palaeoloxodon namadicus & 16.81 & 16.81 & 16.81 & 12.81\tabularnewline
-Mammut americanum & 15.61 & 15.61 & 15.61 & 0.95\tabularnewline
-Mammuthus columbi & 15.71 & 15.71 & 15.71 & 0.91\tabularnewline
-Hydrodamalis gigas & 15.72 & 15.72 & 15.72 & 172.52\tabularnewline
-Atlantogenata & 5.55 & 4.06 & 7.95 & 0.03\tabularnewline
-Afrotheria & 5.55 & 4.05 & 7.96 & 0.00\tabularnewline
-Afrosorcida & 4.35 & 2.58 & 6.13 & 44.49\tabularnewline
-Macroscelidae & 5.27 & 3.98 & 6.85 & 2.49\tabularnewline
-Pseudoungulata & 9.76 & 5.21 & 12.78 & 545.83\tabularnewline
-Paenungulata & 10.13 & 7.24 & 13.02 & 4.42\tabularnewline
-Tethytheria & 12.60 & 10.25 & 13.81 & 187.47\tabularnewline
-Proboscidae & 15.23 & 14.22 & 16.24 & 30.28\tabularnewline
-Elephantidae & 15.49 & 14.89 & 16.10 & 2.21\tabularnewline
-Elephantina & 15.51 & 15.08 & 15.96 & 0.01\tabularnewline
-Mammuthus & 15.54 & 15.24 & 15.85 & 0.47\tabularnewline
-Loxodontini & 15.55 & 15.02 & 16.11 & 0.11\tabularnewline
-Loxodona & 15.72 & 15.16 & 16.30 & 0.86\tabularnewline
-Xenarthra & 7.57 & 5.96 & 9.18 & 124.94\tabularnewline
-\bottomrule
-\end{longtable}
+\caption{Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
+**A)** Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch.
+Atlantogenata is highlighted at the bottom.
+**B)** Zoom-in of (**A**) on Atlantogenata.
+Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size.
+**C)** Confidence interval plot for representative species and ancestral nodes.}\label{fig:Figure1}
+\end{figure}
 
@@
 Afrotherians, we built a time-calibrated supertree of Eutherian mammals
-combining 1,679 species from Bininda-Emonds et al {[}24{]} with a total
+combining 1,679 species from Bininda-Emonds et al {[}25{]} with a total
 evidence Afrotherian phylogeny including 77 extant and fossil data from
-39 extinct species {[}19{]}. Fossil data from extinct species were
+39 extinct species {[}20{]}. Fossil data from extinct species were
 included to ensure that ancestral state reconstructions of body mass in
@@
 model) and out performs standard Brownian motion and Ornstein-Uhlenbeck
-models of character evolution {[}25{]}.
+models of character evolution {[}26{]}.
 
-Similar to previous studies of Afrotherian body size {[}19,25{]}, we
+Similar to previous studies of Afrotherian body size {[}20,26{]}, we
 found that the body mass of the Afrotherian ancestor was inferred to be
 small (0.26kg, 95\% CI: 0.31-3.01kg) and that substantial accelerations
-in the rate of body mass evolution occurred coincident with a 2.77x
-increase \textbf{(Vinnie wrote 65× increase")} in body mass in the
-stem-lineage of \emph{Pseudoungulata} (9.76kg) \textbf{(``Vinnie wrote
-(17kg)'')}, a 1.57x increase \textbf{(``Vinnie wrote 1.5x increase'')}
-in body mass in the stem-lineage of \emph{Paenungulata} (10.13kg)
-\textbf{(``Vinnie wrote (25kg)'')}, a 1.52x increase \textbf{(``Vinnie
-wrote 12× increase'')} in body mass in the stem-lineage of
-\emph{Tehthytheria} (12.6kg) \textbf{(``Vinnie wrote (296kg)'')}, and a
-1.22x increase \textbf{(``Vinnie wrote 14× increase'')} in body mass in
-the stem-lineage of \emph{Proboscidea} (15.23kg) \textbf{(``Vinnie wrote
-(4,100kg)'')} (Figure 1). The ancestral \emph{Hyracoidea} was inferred
-to be relatively small 7.96kg-11.68kg \textbf{(Vinnie wrote
-(2.86-15.71kg))}, and rate accelerations were coincident with
+in the rate of body mass evolution occurred coincident with a 67.36x
+increase in body mass in the stem-lineage of \emph{Pseudoungulata}
+(17.33kg), a 1.45x increase in body mass in the stem-lineage of
+\emph{Paenungulata} (25.08kg), a 11.82x increase in body mass in the
+stem-lineage of \emph{Tehthytheria} (296.56kg), and a 2.69x increase in
+body mass in the stem-lineage of \emph{Proboscidea} (4114.39kg) (Figure
+1B,C). The ancestral \emph{Hyracoidea} was inferred to be relatively
+small 2.86kg-118.18kg, and rate accelerations were coincident with
 independent body mass increases in large hyraxes such as
-\emph{Titanohyrax andrewsi} 12.97kg; 1.33x increase (67× increase in
-body mass). While the body mass of the ancestral \emph{Sirenian} was
-inferred to be large (61-656kg), a rate acceleration occurred coincident
-with a 10× body mass increase in Stellar's sea cow. Rate accelerations
-also occurred coincident with 36× body mass reduction in the
-stem-lineage of the dwarf elephants \emph{Elephas}
-(\emph{Palaeoloxodon}) \emph{falconeri} and \emph{Palaeoloxodon
-cypriotes}. These data suggest that gigantism in \emph{Afrotherians}
-evolved step-wise, from small to medium bodies in the
-\emph{Pseudoungulata} stem-lineage, medium to large bodies in the
+\emph{Titanohyrax andrewsi} 429.34kg; 67.36x increase. While the body
+mass of the ancestral \emph{Sirenian} was inferred to be large
+61.7kg-955.51kg, a rate acceleration occurred coincident with a 10.59x
+increase in body mass in Stellar's sea cow. Rate accelerations also
+occurred coincident with 36.6x decrease in body mass in the stem-lineage
+of the dwarf elephants \emph{Elephas (Palaeoloxodon) antiquus falconeri}
+and \emph{Elephas cypriotes}. These data suggest that gigantism in
+\emph{Afrotherians} evolved step-wise, from small to medium bodies in
+the \emph{Pseudoungulata} stem-lineage, medium to large bodies in the
 \emph{Tehthytherian} stem-lineage and extinct hyraxes, and from large to
@@
 
-\hypertarget{pervasive-duplication-of-tumor-suppressors-during-the-origins-of-large-bodied-afrotheirans}{%
-\subsection*{Pervasive duplication of tumor suppressors during the
-origins of large bodied
-Afrotheirans}\label{pervasive-duplication-of-tumor-suppressors-during-the-origins-of-large-bodied-afrotheirans}}
-\addcontentsline{toc}{subsection}{Pervasive duplication of tumor
-suppressors during the origins of large bodied Afrotheirans}
+\hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
+\subsection*{Step-wise reduction of intrinsic cancer risk in large,
+long-lived
+Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}
+\addcontentsline{toc}{subsection}{Step-wise reduction of intrinsic
+cancer risk in large, long-lived Afrotherians}
 
@@
 
-\leavevmode\hypertarget{ref-CaulinAndMaley2011}{}%
-4. Caulin AF, Maley CC. Peto's Paradox: evolution's prescription for
+\leavevmode\hypertarget{ref-Dorn1968}{}%
+4. Dorn CR, Taylor DON, Schneider R, Hibbard HH, Klauber MR. Survey of
+Animal Neoplasms in Alameda and Contra Costa Counties, California. II.
+Cancer Morbidity in Dogs and Cats From Alameda County\textless{}xref
+ref-type="fn" rid="FN2"\textgreater{}2\textless{}/xref\textgreater{}.
+JNCI: Journal of the National Cancer Institute. 1968;40: 307--318.
+doi:\href{https://doi.org/10.1093/jnci/40.2.307}{10.1093/jnci/40.2.307}
+
+\leavevmode\hypertarget{ref-Caulin2011}{}%
+5. Caulin AF, Maley CC. Peto's Paradox: evolution's prescription for
 cancer prevention. Trends in ecology \& evolution. 2011;26: 175--82.
@@
 \leavevmode\hypertarget{ref-Leroi2003}{}%
-5. Leroi AM, Koufopanou V, Burt A. Cancer selection. Nature Reviews
+6. Leroi AM, Koufopanou V, Burt A. Cancer selection. Nature Reviews
 Cancer. 2003;3: 226--231.
@@
 \leavevmode\hypertarget{ref-Peto1975}{}%
-6. Peto R, Roe F, Lee P, Levy L, Clack J. Cancer and ageing in mice and
+7. Peto R, Roe F, Lee P, Levy L, Clack J. Cancer and ageing in mice and
 men. British Journal of Cancer. 1975;32: 411--426.
@@
 \leavevmode\hypertarget{ref-Ashur-Fabian2004}{}%
-7. Ashur-Fabian O, Avivi A, Trakhtenbrot L, Adamsky K, Cohen M, Kajakaro
+8. Ashur-Fabian O, Avivi A, Trakhtenbrot L, Adamsky K, Cohen M, Kajakaro
 G, et al. Evolution of p53 in hypoxia-stressed Spalax mimics human tumor
@@
 \leavevmode\hypertarget{ref-Seluanov2008}{}%
-8. Seluanov A, Hine C, Bozzella M, Hall A, Sasahara THC, Ribeiro AACM,
+9. Seluanov A, Hine C, Bozzella M, Hall A, Sasahara THC, Ribeiro AACM,
 et al. Distinct tumor suppressor mechanisms evolve in rodent species
@@
 \leavevmode\hypertarget{ref-Gorbunova2012}{}%
-9. Gorbunova V, Hine C, Tian X, Ablaeva J, Gudkov AV, Nevo E, et al.
+10. Gorbunova V, Hine C, Tian X, Ablaeva J, Gudkov AV, Nevo E, et al.
 Cancer resistance in the blind mole rat is mediated by concerted
@@
 \leavevmode\hypertarget{ref-Tian2013}{}%
-10. Tian X, Azpurua J, Hine C, Vaidya A, Myakishev-Rempel M, Ablaeva J,
+11. Tian X, Azpurua J, Hine C, Vaidya A, Myakishev-Rempel M, Ablaeva J,
 et al. High molecular weight hyaluronan mediates the cancer resistance
@@
 
-\leavevmode\hypertarget{ref-Sulak2016}{}%
-11. Sulak M, Fong L, Mika K, Chigurupati S, Yon L, Mongan NP, et al.
+\leavevmode\hypertarget{ref-Sulak2015}{}%
+12. Sulak M, Fong L, Mika K, Chigurupati S, Yon L, Mongan NP, et al.
 TP53 copy number expansion is associated with the evolution of increased
@@
 \leavevmode\hypertarget{ref-HAGR}{}%
-12. Tacutu R, Craig T, Budovsky A, Wuttke D, Lehmann G, Taranukha D, et
+13. Tacutu R, Craig T, Budovsky A, Wuttke D, Lehmann G, Taranukha D, et
 al. Human Ageing Genomic Resources: Integrated databases and tools for
@@
 \leavevmode\hypertarget{ref-Schwartz1995}{}%
-13. Schwartz GT, Rasmussen DT, Smith RJ. Body-Size Diversity and
+14. Schwartz GT, Rasmussen DT, Smith RJ. Body-Size Diversity and
 Community Structure of Fossil Hyracoids. Journal of Mammalogy. 1995;76:
@@
 \leavevmode\hypertarget{ref-Scheffer1972}{}%
-14. Scheffer VB. The Weight of the Steller Sea Cow. Journal of
+15. Scheffer VB. The Weight of the Steller Sea Cow. Journal of
 Mammalogy. 1972;53: 912--914.
@@
 \leavevmode\hypertarget{ref-Larramendi:20151c2}{}%
-15. Larramendi A. Shoulder Height, Body Mass, and Shape of
+16. Larramendi A. Shoulder Height, Body Mass, and Shape of
 Proboscideans. Acta Palaeontologica Polonica. 2015;61.
@@
 \leavevmode\hypertarget{ref-OLeary2013a}{}%
-16. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
+17. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
 NP, et al. The placental mammal ancestor and the post-K-Pg radiation of
@@
 \leavevmode\hypertarget{ref-Springer2013}{}%
-17. Springer MS, Meredith RW, Teeling EC, Murphy WJ. Technical comment
+18. Springer MS, Meredith RW, Teeling EC, Murphy WJ. Technical comment
 on "The placental mammal ancestor and the post-K-Pg radiation of
@@
 \leavevmode\hypertarget{ref-OLeary2013b}{}%
-18. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
+19. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
 NP, et al. Response to comment on "The placental mammal ancestor and the
@@
 \leavevmode\hypertarget{ref-PuttickAndThomas2015}{}%
-19. Puttick MN, Thomas GH. Fossils and living taxa agree on patterns of
+20. Puttick MN, Thomas GH. Fossils and living taxa agree on patterns of
 body mass evolution: a case study with Afrotheria. Proceedings
@@
 \leavevmode\hypertarget{ref-Abegglen:JAMA2015}{}%
-20. Abegglen LM, Caulin AF, Chan A, Lee K, Robinson R, Campbell MS, et
+21. Abegglen LM, Caulin AF, Chan A, Lee K, Robinson R, Campbell MS, et
 al. Potential Mechanisms for Cancer Resistance in Elephants and
@@
 \leavevmode\hypertarget{ref-Vazquez2018}{}%
-21. Vazquez JM, Sulak M, Chigurupati S, Lynch VJ. A Zombie LIF Gene in
+22. Vazquez JM, Sulak M, Chigurupati S, Lynch VJ. A Zombie LIF Gene in
 Elephants Is Upregulated by TP53 to Induce Apoptosis in Response to DNA
@@
 \leavevmode\hypertarget{ref-Caulin2015}{}%
-22. Caulin AF, Graham TA, Wang L-S, Maley CC. Solutions to Peto's
+23. Caulin AF, Graham TA, Wang L-S, Maley CC. Solutions to Peto's
 paradox revealed by mathematical modelling and cross-species cancer gene
@@
 \leavevmode\hypertarget{ref-Doherty2016}{}%
-23. Doherty A, Magalhães J de. Has gene duplication impacted the
+24. Doherty A, Magalhães J de. Has gene duplication impacted the
 evolution of Eutherian longevity? Aging Cell. 2016;15: 978--980.
@@
 \leavevmode\hypertarget{ref-Bininda-Emonds2008}{}%
-24. Bininda-Emonds ORP, Cardillo M, Jones KE, MacPhee RDE, Beck RMD,
+25. Bininda-Emonds ORP, Cardillo M, Jones KE, MacPhee RDE, Beck RMD,
 Grenyer R, et al. Erratum: The delayed rise of present-day mammals.
@@
 \leavevmode\hypertarget{ref-ElliotAndMooers2014}{}%
-25. Elliot MG, Mooers AØ. Inferring ancestral states without assuming
+26. Elliot MG, Mooers AØ. Inferring ancestral states without assuming
 neutrality or gradualism using a stable model of continuous character
@@
 \leavevmode\hypertarget{ref-blat}{}%
-26. Kent JW. BLAT---The BLAST-Like Alignment Tool. Genome Research.
+27. Kent JW. BLAT?The BLAST-Like Alignment Tool. Genome Research.
 2002;12: 656--664.
@@
 \leavevmode\hypertarget{ref-AltenhoffAndDessimoz2009}{}%
-27. Altenhoff AM, Dessimoz C. Phylogenetic and functional assessment of
+28. Altenhoff AM, Dessimoz C. Phylogenetic and functional assessment of
 orthologs inference projects and methods. PLoS computational biology.
@@
 
+\leavevmode\hypertarget{ref-SalichosAndRokas2011}{}%
+29. Salichos L, Rokas A. Evaluating ortholog prediction algorithms in a
+yeast model clade. PloS one. 2011;6: e18755.
+doi:\href{https://doi.org/10.1371/journal.pone.0018755}{10.1371/journal.pone.0018755}
+
 \leavevmode\hypertarget{ref-uniprot}{}%
-28. Consortium TU. UniProt: the universal protein knowledgebase. Nucleic
+30. Consortium TU. UniProt: the universal protein knowledgebase. Nucleic
 Acids Research. 2017;45: D158--D169.
@@
 
-\nolinenumbers
+\leavevmode\hypertarget{ref-gVolante}{}%
+31. Nishimura O, Hara Y, Kuraku S. gVolante for standardizing
+completeness assessment of genome and transcriptome assemblies.
+Bioinformatics. 2017;33: 3635--3637.
+doi:\href{https://doi.org/10.1093/bioinformatics/btx445}{10.1093/bioinformatics/btx445}
+
+\leavevmode\hypertarget{ref-CEGMA}{}%
+32. Parra G, Bradnam K, Ning Z, Keane T, Korf I. Assessing the gene
+space in draft genomes. Nucleic Acids Research. 2008;37: 289--297.
+doi:\href{https://doi.org/10.1093/nar/gkn916}{10.1093/nar/gkn916}
+
+\leavevmode\hypertarget{ref-HISAT}{}%
+33. Kim D, Langmead B, Salzberg SL. HISAT: a fast spliced aligner with
+low memory requirements. Nature Methods. 2015;12: 357--360.
+doi:\href{https://doi.org/10.1038/nmeth.3317}{10.1038/nmeth.3317}
+
+\leavevmode\hypertarget{ref-StringTie}{}%
+34. Pertea M, Pertea GM, Antonescu CM, Chang T-C, Mendell JT, Salzberg
+SL. StringTie enables improved reconstruction of a transcriptome from
+RNA-seq reads. Nature Biotechnology. 2015;33: 290--295.
+doi:\href{https://doi.org/10.1038/nbt.3122}{10.1038/nbt.3122}
+
+\leavevmode\hypertarget{ref-Tuxedo}{}%
+35. Pertea M, Kim D, Pertea GM, Leek JT, Salzberg SL. Transcript-level
+expression analysis of RNA-seq experiments with HISAT, StringTie and
+Ballgown. Nature Protocols. 2016;11: 1650--1667.
+doi:\href{https://doi.org/10.1038/nprot.2016.095}{10.1038/nprot.2016.095}
+
+\leavevmode\hypertarget{ref-WebGestalt2019}{}%
+36. Liao Y, Wang J, Jaehnig EJ, Shi Z, Zhang B. WebGestalt 2019: gene
+set analysis toolkit with revamped UIs and APIs. Nucleic Acids Research.
+2019;47: W199--W205.
+doi:\href{https://doi.org/10.1093/nar/gkz401}{10.1093/nar/gkz401}
+
+\leavevmode\hypertarget{ref-Reactome}{}%
+37. Jassal B, Matthews L, Viteri G, Gong C, Lorente P, Fabregat A, et
+al. The reactome pathway knowledgebase. Nucleic acids research. 2020;48:
+D498--D503.
+doi:\href{https://doi.org/10.1093/nar/gkz1031}{10.1093/nar/gkz1031}
+
+\leavevmode\hypertarget{ref-Wikipathways}{}%
+38. Slenter DN, Kutmon M, Hanspers K, Riutta A, Windsor J, Nunes N, et
+al. WikiPathways: a multifaceted pathway database bridging metabolomics
+to other omics research. Nucleic Acids Research. 2017;46: D661--D667.
+doi:\href{https://doi.org/10.1093/nar/gkx1064}{10.1093/nar/gkx1064}
+
+\leavevmode\hypertarget{ref-KEGG}{}%
+39. Kanehisa M, Goto S. KEGG: Kyoto Encyclopedia of Genes and Genomes.
+Nucleic Acids Research. 2000;28: 27--30.
+doi:\href{https://doi.org/10.1093/nar/28.1.27}{10.1093/nar/28.1.27}
+
+\leavevmode\hypertarget{ref-Felsenstein1985}{}%
+40. Felsenstein J. Phylogenies and the Comparative Method. The American
+Naturalist. 1985;125: 1--15.
+doi:\href{https://doi.org/10.1086/284325}{10.1086/284325}
+
+\leavevmode\hypertarget{ref-MartinsAndHansen1997}{}%
+41. Martins EP, Hansen TF. Phylogenies and the Comparative Method: A
+General Approach to Incorporating Phylogenetic Information into the
+Analysis of Interspecific Data. The American Naturalist. 1997;149:
+646--667. doi:\href{https://doi.org/10.1086/286013}{10.1086/286013}
+
+\leavevmode\hypertarget{ref-Armitage:19851c2}{}%
+42. Armitage P. Multistage models of carcinogenesis. Environmental
+health perspectives. 1985;63: 195--201.
+doi:\href{https://doi.org/10.1289/ehp.8563195}{10.1289/ehp.8563195}
+
+\leavevmode\hypertarget{ref-Armitage:20041c2}{}%
+43. Armitage P, Doll R. The age distribution of cancer and a multi-stage
+theory of carcinogenesis. British Journal of Cancer. 2004;91: 6602297.
+doi:\href{https://doi.org/10.1038/sj.bjc.6602297}{10.1038/sj.bjc.6602297}
+
+\leavevmode\hypertarget{ref-Peto:20151c2}{}%
+44. Peto R. Quantitative implications of the approximate irrelevance of
+mammalian body size and lifespan to lifelong cancer risk. Phil Trans R
+Soc B. 2015;370: 20150198.
+doi:\href{https://doi.org/10.1098/rstb.2015.0198}{10.1098/rstb.2015.0198}
 

### 2020-05-28 code/generalFunctions.R

@@ -699,7 +699,7 @@ reorder_cormat <- function(cormat){
 }
 
 
-not.me <- function(attr1, attr2, threshold=5){
+not.me <- function(attr1, attr2, threshold=4){
   sapply(
     seq_along(attr1),
     function(n, a1=attr1, a2=attr2, t=threshold){
@@ -707,7 +707,7 @@ not.me <- function(attr1, attr2, threshold=5){
     }
   )
 }
-yes.me <- function(attr1, attr2, threshold=5){
+yes.me <- function(attr1, attr2, threshold=4){
   sapply(
     seq_along(attr1),
     function(n, a1=attr1, a2=attr2, t=threshold){
@@ -716,6 +716,47 @@ yes.me <- function(attr1, attr2, threshold=5){
   )
 }
 
+colorado <- function(src, boulder) {                        # From https://stackoverflow.com/a/39695859, pun is too good to change
+  if (!is.factor(src)) src <- factor(src)                   # make sure it's a factor
+  src_levels <- levels(src)                                 # retrieve the levels in their order
+  brave <- boulder %in% src_levels                          # make sure everything we want to make bold is actually in the factor levels
+  if (all(brave)) {                                         # if so
+    b_pos <- purrr::map_int(boulder, ~which(.==src_levels)) # then find out where they are
+    b_vec <- rep("plain", length(src_levels))               # make'm all plain first
+    b_vec[b_pos] <- "bold"                                  # make our targets bold
+    b_vec                                                   # return the new vector
+  } else {
+    stop("All elements of 'boulder' must be in src")
+  }
+}
+
+highlightTextSize <- function(src, boulder, sizes=c(4,8)) {
+  if (!is.factor(src)) src <- factor(src)                 
+  src_levels <- levels(src)                               
+  brave <- boulder %in% src_levels                        
+  if (all(brave)) {                                       
+    b_pos <- purrr::map_int(boulder, ~which(.==src_levels))
+    b_vec <- rep(sizes[1], length(src_levels))            
+    b_vec[b_pos] <- sizes[2]                              
+    b_vec                                                   
+  } else {
+    stop("All elements of 'boulder' must be in src")
+  }
+}
+
+highlightTextColor <- function(src, boulder, color=c("grey","black")) {
+  if (!is.factor(src)) src <- factor(src)                 
+  src_levels <- levels(src)                               
+  brave <- boulder %in% src_levels                        
[…]

### 2020-05-29 paper_PLOS/paper_PLOS_draft.Rmd
Figure 2 corrected

@@
   theme(
-    legend.position="right",
+    legend.position="bottom",
     # panel.background = element_rect(fill = "lightgrey"),
@@
 
-`r knitr::knit_exit()`
-
 ## Step-wise reduction of intrinsic cancer risk in large, long-lived Afrotherians
@@
 
-```{r Fig2 Prep Table}
+```{r Fig2 Prep Table, eval=F}
 if(knitr::is_latex_output()){
@@
     continuous=T,
-    size = 1.5
-  ) + 
-  geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic",
-              show_guide=F, align = F, linesize = 0, offset= 100
+    size = 1
+  ) + 
+  geom_tiplab(aes(image=phyloPicUID, size=I(0.1)), geom="phylopic",
+              show_guide=F, align = F, linesize = 0, offset= 5
               ) +
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidea, label="Proboscidea", 
-                  fontsize = 4, align = T, offset = 10, angle=0, barsize = 1, 
-                  offset.text = 0,
+                  fontsize = 2, align = T, offset = 45, angle=0, barsize = 1,
+                  offset.text = 5,
                   color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Proboscidea") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
                   ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="Tethytheria", 
-                  fontsize = 4, align = T, offset = 5, angle=0, barsize = 1, 
-                  offset.text = 5, color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
-                  ) +
-  geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="Paenungulata", 
-                  fontsize = 4, align = T, angle=0, barsize = 1, offset.text = 10, 
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="\nTethytheria", 
+                  fontsize = 2, align = T, offset = 40, angle=0, barsize = 1, 
+                  offset.text = 5, 
+                  color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
+                  ) +
+  geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="\n\nPaenungulata", 
+                  fontsize = 2, align = T, angle=0, barsize = 1, offset.text = 5, offset=35,
                   color = table.cancerSucceptibility.Atlantogenata %>% filter(label=="Paenungulata") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
@@
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosoricida, label="Afrosoricida",
-                  fontsize = 4, align = T, angle=0, barsize = 1,offset.text = 10,
+                  fontsize = 2, align = T, angle=0, barsize = 1, offset.text=5,offset=35,
                   color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Afrosoricida") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
@@
                     label="Macroscelidae", 
-                    fontsize = 4, align = T, angle=0, barsize = 1,offset.text = 10,
+                    fontsize = 2, align = T, angle=0, barsize = 1,offset.text=5,offset=35,
                     color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Macroscelidae") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
@@
   geom_cladelabel(node=bsize.Atlantogenata.nodes$Xenarthra, label="Xenarthra", 
-                  fontsize = 4, align = T, angle=0, barsize = 1,offset.text = 10,
+                  fontsize = 2, align = T, angle=0, barsize = 1,offset.text=5,offset=35,
                   color=table.cancerSucceptibility.Atlantogenata %>% filter(label=="Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% as.numeric %>% get.log2CancerSucceptabilityChange.color
@@
   scale_color_viridis_c(limits=c(scale.log2CancerSucceptabilityChange.min, scale.log2CancerSucceptabilityChange.max), 
-                        guide = guide_colourbar(title="Log2 Cancer Susceptibility Change"))+
-  xlim(0,300) +
+                        guide = guide_colourbar(title="Log2 Cancer Susceptibility Change", title.position = "top", title.hjust = 0.5))+
+  xlim(0,275) +
   theme_tree2()  + 
   theme(
-    legend.position="right",
-    panel.background = element_rect(fill = "lightgrey")
-  ) + 
-  xlab("|Log2CancerSucceptabilityChange|")
+    legend.position="bottom",
+    panel.background = element_rect(fill = "lightgrey"),
+    axis.text.x = element_text(size = 4, 
+                               colour = "black", face = "bold"),
+    axis.title = element_text(size = 6, 
+                              colour = "black", face = "bold"),
+    plot.title = element_text(size = 6, 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.title = element_text(size = 6, 
+                                face = "bold", colour = "black"), 
+    legend.text = element_text(size = 6, 
+                               face = "plain", colour = "black")
+  ) + 
+  xlab("| Log2 Cancer Succeptibility Change |")
 # f5a
@@
     arrange(Node) %>% 
+  select(Node, `Est. Lifespan`, K1, K2, `Change in K`, `log2 Change`) %>% 
   knitr::kable(
@@
 ```
-```{r Figure2, fig.cap="Cancer succecptibility across Atlantogenata"}
+```{r Figure2, fig.cap="Cancer succecptibility across Atlantogenata", out.width="6in", fig.asp=1}
 f2
@@
 
-```{r Table 1}
+```{r Table 2}
 t.cancersucc.atl
@@
 
-In order to estimate the intrinsic cancer risk of a species, we first obtained estimates for lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon$ (Figure `r warning("make lifespan figure")`).
+In order to estimate the intrinsic cancer risk of a species, we first obtained estimates for lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon$ (Table 2).
 With this information in hand, we calculated $K_{1}$ at all nodes, and then estimated the fold change in cancer susceptibility between an ancestral node and a given node as $\frac{K_{2}}{K_{1}}$ (__Table 4__).
@@
 
-As shown in __Table 4__, cancer susceptibility skyrocketed at the initial divergence of Atlantogenata, followed by a generally upwards trend.
+As shown in __Table 2__, cancer susceptibility skyrocketed at the initial divergence of Atlantogenata, followed by a generally upwards trend.
 At the common ancestor of Afrotheria there is an initial `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrotheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer risk.
@@
 Neither the common ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further changes in cancer susceptibility.
+
+`r knitr::knit_exit()`
 

### 2020-05-29 paper_PLOS/paper_PLOS_draft.tex
Figure 2 corrected

@@
 \bibliographystyle{elsarticle-harv}
+\usepackage{longtable}
 \ifxetex
@@
 \hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
-\subsection*{Step-wise reduction of intrinsic cancer risk in large,
+\subsection{Step-wise reduction of intrinsic cancer risk in large,
 long-lived
 Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}
-\addcontentsline{toc}{subsection}{Step-wise reduction of intrinsic
-cancer risk in large, long-lived Afrotherians}
+
+\begin{figure}
+
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure2-1} \hfill{}
+
+\caption{Cancer succecptibility across Atlantogenata}\label{fig:Figure2}
+\end{figure}
+
+\begin{longtable}[]{@{}lrlllr@{}}
+\caption{Estimated Cancer Succeptibilty for nodes in
+Atlantogenata}\tabularnewline
+\toprule
+Node & Est. Lifespan & K1 & K2 & Change in K & log2
+Change\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Node & Est. Lifespan & K1 & K2 & Change in K & log2
+Change\tabularnewline
+\midrule
+\endhead
+Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 &
+2.31\tabularnewline
+Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 &
+4.07\tabularnewline
+Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\tabularnewline
+Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 &
+-2.31\tabularnewline
+Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 &
+0.00\tabularnewline
+Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 &
+27.66\tabularnewline
+Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 &
+2.31\tabularnewline
+Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 &
+4.14\tabularnewline
+Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\tabularnewline
+Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 &
+-2.31\tabularnewline
+Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 &
+0.00\tabularnewline
+Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 &
+-23.05\tabularnewline
+Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 &
+29.97\tabularnewline
+Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 &
+6.92\tabularnewline
+Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 &
+8.62\tabularnewline
+Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 &
+0.00\tabularnewline
+Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 &
+-5.01\tabularnewline
+Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 &
+9.22\tabularnewline
+Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 &
+5.38\tabularnewline
+Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 &
+-4.61\tabularnewline
+Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 &
+-4.61\tabularnewline
+Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 &
+0.00\tabularnewline
+Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 &
+5.22\tabularnewline
+Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 &
+9.22\tabularnewline
+Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 &
+20.75\tabularnewline
+Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 &
+-3.76\tabularnewline
+Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 &
+1.51\tabularnewline
+Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 &
+0.00\tabularnewline
+Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 &
+0.00\tabularnewline
+\bottomrule
+\end{longtable}
+
+The dramatic increase in body mass and lifespan in some Afrotherian
+lineages implies those lineages evolved reduced cancer risk. To infer
+the magnitude of these reductions we estimated differences in cancer
+risk between small bodied, short-lived species and large bodied,
+long-lived species as well as for reconstructed ancestral Afrotherians.
+Following {[}44{]} we estimate the intrinsic cancer risk as the product
+of risk associated with body mass and lifespan. Differences in cancer
+susceptibility \(K\) due to body mass differences between species can be
+approximated simply as the fold difference in body mass (\(D\)) between
+species {[}44{]}. The risk of developing cancer also increases in
+proportion to the sixth power of age and is approximated by the formula
+\(Ct^6\), in which the proportionality constant C that determines
+susceptibility to cancer induction is multiplied by the sixth power of
+the age in years, \(t\) {[}44--46{]}. Thus we can estimate the intrinsic
+cancer risk for a species as \(K \approx Dt^6\).
+
+In order to estimate the intrinsic cancer risk of a species, we first
+obtained estimates for lifespans at ancestral nodes using PGLS and the
+model
+\(ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon\)
+(Table 2). With this information in hand, we calculated \(K_{1}\) at all
+nodes, and then estimated the fold change in cancer susceptibility
+between an ancestral node and a given node as \(\frac{K_{2}}{K_{1}}\)
+(\textbf{Table 4}). As body size and phylogeny are the strongest
+predictors of lifespan, we find that this regression is sufficiently
+robust without further metabolic or other covariates for our purposes
+herein.
+
+As shown in \textbf{Table 2}, cancer susceptibility skyrocketed at the
+initial divergence of Atlantogenata, followed by a generally upwards
+trend. At the common ancestor of Afrotheria there is an initial
+9.22-fold increase in cancer risk. In parallel to Afrotheria, cancer
+susceptibility increases 20.75-fold in Xenarthra. However, cancer risk
+slowly deflates as size decreases as one moves along the tree towards
+extant species, such as in Hoffman's Two Toed Sloth (1.51-fold change)
+and in the Nine-banded Armadillo (-3.76-fold change).
+
+Within Afrotheria, cancer susceptibility drops in Afrosoricida as
+species shrink (-4.61-fold, then stagnates for the Cape Golden Mole) -
+but then rises 5.22-fold towards the Lesser Hedgehog Tenrec. In
+parallel, Afroinsectivora does not increase in cancer susceptibility,
+and decreases once more at the Cape Elephant Shrew (-4.61-fold). The
+emergence of Pseudoungulata sees the next big leap in cancer
+susceptibility with a 9.22-fold increase. The Aardvark further increases
+5.38-fold, while we don't observe an increase at the common ancestor of
+Paenungulates. While the Rock Hyrax decreases in cancer susceptibility
+as expected (-5.01-fold), Tethytheria sees a sharp increase in cancer
+risk (6.92-fold). Within Tethytheria, the Manatee's cancer risk
+increases once more 8.62-fold, while Proboscidea's cancer risk drops
+precipitously along with its body size (-23.05-fold). Yet, within
+Proboscidea we see the biggest increases: right off the bat, we see that
+the cancer susceptibility of Elephantidae and the American Mastodon
+skyrocket by 27.66-fold and 29.97-fold, respectively. Both Elephantina
+and Loxodontini in Elephantidae have a 2.31-fold increase in cancer
+susceptibility. Within Elephantina, cancer susceptibility stays stable
+at Mammuthus and in the Colombian Mammoth, and slightly decreases in the
+Wooly Mammoth (-2.31-fold). The three extant elephants - Asian Elephant
+in Elephantina, the African Savana Elephant in Loxodontini, and the
+African Forest Elephant in Loxodona, meanwhile, have parallel and
+similar decreases in both size and cancer susceptibility (4.14-, 4.07-,
+and -2.31-fold, respectively). Neither the common ancestor of Loxodonta,
+nor the Straight-Tusked Mammoth see any further changes in cancer
+susceptibility.
+
+\hypertarget{identification-and-evolutionary-history-of-gene-duplications}{%
+\subsection*{Identification and evolutionary history of gene
+duplications}\label{identification-and-evolutionary-history-of-gene-duplications}}
+\addcontentsline{toc}{subsection}{Identification and evolutionary
+history of gene duplications}
 
@@
 
+\leavevmode\hypertarget{ref-ArmitageAndDoll1954}{}%
+45. Armitage P, Doll R. The Age Distribution of Cancer and a Multi-stage
+Theory of Carcinogenesis. British Journal of Cancer. 1954;8: 1--12.
+doi:\href{https://doi.org/10.1038/bjc.1954.1}{10.1038/bjc.1954.1}
+
+\leavevmode\hypertarget{ref-Nordling1953}{}%
+46. Nordling CO. A New Theory on the Cancer-inducing Mechanism. British
+Journal of Cancer. 1953;7: 68--72.
+doi:\href{https://doi.org/10.1038/bjc.1953.8}{10.1038/bjc.1953.8}
+
 \end{document}

### 2020-05-31 paper_PLOS/paper_PLOS_draft.Rmd
Added Figure, fixed figures 0-5

@@
 knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=6, 
-                      fig.retina = 2, fig.align = "left", dpi = 300, dev="Cairo_pdf")
+                      fig.retina = 2, fig.align = "left", dpi = 300)
 
@@
 suppressMessages(suppressWarnings(require(purrr)))
+suppressMessages(suppressWarnings(require(kableExtra)))
 suppressMessages(suppressWarnings(require(stargazer)))
@@
   mutate_at(vars(-`Node`, ), . %>% as.numeric %>% round(2)) %>% 
-  knitr::kable(
+  kable(
     ., 
@@
     booktabs = TRUE,
-    caption = 'A table generated by the longtable package.'
+    caption = 'All body size estimates for _Atlantogenata_.'
   )
@@
                         debug = FALSE), 
-    axis.text = element_text(size = 4, 
+    axis.text = element_text(size = 6, 
                                colour = "black", face = "bold"),
-    axis.title = element_text(size = 6, 
+    axis.title = element_text(size = 8, 
                               colour = "black", face = "bold"),
-    plot.title = element_text(size = 6, 
+    plot.title = element_text(size = 8, 
                               colour = "black", lineheight = 1, face = "bold"), 
-    legend.text = element_text(size = 4, 
+    legend.text = element_text(size = 6, 
                                face = "plain", colour = "black"),
@@
   geom_tiplab(aes(label=Spc.short), offset = 0.4, color="black", size=4, parse=T)+
-  geom_tiplab(aes(label=not.me(lnSize, N.Genes)), geom="label",  offset=-0.5, color="black", size=3)+ 
-  geom_tiplab(aes(label=yes.me(lnSize, N.Genes)), geom="label",  offset=-0.5, color="grey", size=3)+ 
-  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="black", size=3)+ 
-  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="grey", size=3)+ 
-  geom_nodelab(aes(x=branch, label=not.me(lnSize, label)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="black", size=3)+ 
-  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="grey", size=3)+ 
-  geom_nodelab2(aes(subset=label=="Root", x=branch), geom="label",
-               size=3,
+  geom_tiplab(aes(label=not.me(lnSize, N.Genes, threshold = 7)), geom="label",  offset=-0.5, color="black", size=2, fontface="bold")+ 
+  geom_tiplab(aes(label=yes.me(lnSize, N.Genes, threshold = 7)), geom="label",  offset=-0.5, color="grey", size=2)+ 
+  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes, threshold = 7)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="black", size=2, fontface="bold")+ 
+  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes, threshold = 7)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="grey", size=2, fontface="bold")+ 
+  geom_nodelab(aes(x=branch, label=not.me(lnSize, label, threshold = 7)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="black", size=2, fontface="bold")+ 
+  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label, threshold = 7)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="grey", size=2, fontface="bold")+ 
+  geom_nodelab2(aes(subset=(label=="Root"), label = "Atlantogenata", x=branch), geom="label",
+               size=2,
                color="grey")+
   theme_tree() + 
-  xlim(0,13) +
-  theme(legend.position = "bottom")+
-  xlab("Rate of Change, Body Size")
+  xlim(-0.5,13) +
+  xlab("Rate of Change, Body Size") + 
+  theme(
+    legend.position="bottom",
+    # panel.background = element_rect(fill = "lightgrey"),
+    text = element_text(face = "plain", 
+                        colour = "black", lineheight = 0.9, 
+                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                        debug = FALSE), 
+    axis.text.x = element_text(size = 4, 
+                               colour = "black", face = "bold"),
+    axis.title = element_text(size = 6, 
+                              colour = "black", face = "bold"),
+    plot.title = element_text(size = 6, 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.title = element_text(size = 6, 
+                                face = "bold", colour = "black"), 
+    legend.text = element_text(size = 6, 
+                               face = "plain", colour = "black")
+  )
 # p.ml.tree.increasing.geneCounts.withSize
@@
 
-)
+```{r Figure0Prep}
+
+groups.atlantogenata <- list(
+  # "Atlantogenata",
+  "Elephantidae" = c("Loxodonta.africana", "Loxodonta.cyclotis", "Palaeoloxodon.antiquus", "Elephas.maximus", "Mammuthus.columbi", "Mammuthus.primigenius"),
+  "Proboscidea" = c("Mammut.americanum"),
+  "Tethytheria" = c("Trichechus.manatus"),
+  "Paenungulata" = c("Procavia.capensis"),
+  "Pseudoungulata" = c("Orycteropus.afer"),
+  "Xenarthra" = c("Dasypus.novemcinctus", "Choloepus.hoffmanni"),
+  "Afrosoricida" = c("Chrysochloris.asiatica"), 
+  "Macroscelidea" = c("Elephantulus.edwardii", "Echinops.telfairi")
+)
+
+anage.genomeSpecies <- anage %>% filter(label %in% genomeTable.atlantogenata$label)
+anage.genomeSpecies <- anage.genomeSpecies %>% select(label, Lifespan) %>% 
+  full_join(tree.Atlantogenata.table %>% filter(label %in% genomeTable.atlantogenata$label) %>% select(label, Size=Median, phyloPicUID), by = "label")
+
+label.phylopic <- anage.genomeSpecies %>% select(label, uid=phyloPicUID)
+
+time.tree.atlantogenata.genomeSpecies <- time.tree.atlantogenata %>% 
+  drop.tip(time.tree.atlantogenata$tip.label[!time.tree.atlantogenata$tip.label %in% genomeTable.atlantogenata$label]) %>% 
+  full_join(label.phylopic, by="label") %>% 
+  as_tibble %>%
+  mutate(prettyLabel = label %>% str_replace("\\.", "\n")) %>% 
+  groupOTU(groups.atlantogenata) %>% #print(n=12412)
+  as.treedata()
+
+atlantogenata.tree.size.lifespan <-
+time.tree.atlantogenata.genomeSpecies %>% 
+  ggtree() %>% 
+  flip(tree_view = ., node1 =nodeid(time.tree.atlantogenata.genomeSpecies, "Loxodonta.africana"), node2 = MRCA(time.tree.atlantogenata.genomeSpecies, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis")) %>%
+  revts() +
+  geom_tiplab(aes(image=uid, size=I(0.1), color = group), geom = "phylopic") +
+  geom_tiplab(aes(label=prettyLabel), size=3, offset = 20) +
+  xlab("Divergence time (MYA)") +
+  geom_facet(
+    data = anage.genomeSpecies,
+    panel = "ln(Size)",
+    mapping = aes(x=Size),
+    geom = ggstance::geom_barh,
+    stat = "identity", 
+    fill = "aquamarine"
+  ) + 
+  geom_facet(
+    data = anage.genomeSpecies,
+    panel = "Lifespan",
+    mapping = aes(x=Lifespan),
+    geom = ggstance::geom_barh,
+    stat = "identity",
+    fill = "goldenrod"
+  ) + 
+  theme_tree2() +
+  xlim_tree(c(-90, 65)) + 
+  theme(
+    legend.position="bottom",
+    # panel.background = element_rect(fill = "lightgrey"),
+    text = element_text(face = "plain", 
+                        colour = "black", lineheight = 0.9, 
+                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                        debug = FALSE), 
+    axis.text.x = element_text(size = 4, 
+                               colour = "black", face = "bold"),
+    axis.title = element_text(size = 6, 
+                              colour = "black", face = "bold"),
+    plot.title = element_text(size = 6, 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.title = element_text(size = 8, 
+                                face = "bold", colour = "black"), 
+    legend.text = element_text(size = 6, 
+                               face = "plain", colour = "black"),
+    strip.text.y = element_text(size=8, face="bold")
+  )  + 
+  guides(color=guide_legend(title="Clade", title.position = "left"))
+
+atlantogenata.tree.size.lifespan %<>% facet_widths(c(2,1,1))
+
+```
+```{r Figure0, fig.cap="Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]",  out.width="6in", fig.asp=1.3}
+atlantogenata.tree.size.lifespan
+```
 
@@
   select(Node, `Est. Lifespan`, K1, K2, `Change in K`, `log2 Change`) %>% 
-  knitr::kable(
+  kable(
     ., 
-    longtable = TRUE, 
     booktabs = TRUE,
     caption = 'Estimated Cancer Succeptibilty for nodes in Atlantogenata'
-  )
+  ) %>% 
+  kable_styling(latex_options = c("scale_down"))
 ```
@@
 Neither the common ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further changes in cancer susceptibility.
-
-`r knitr::knit_exit()`
 
@@
 
+  # theme_pubclean()+
+  # labs_pubr() + 
+  # theme(
+  #   text = element_text(face = "plain", 
+  #                       colour = "black", lineheight = 0.9, 
+  #                       hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+  #                       debug = FALSE), 
+  #   axis.text = element_text(size = 6, 
+  #                              colour = "black", face = "bold"),
+  #   axis.title = element_text(size = 8, 
+  #                             colour = "black", face = "bold"),
+  #   plot.title = element_text(size = 8, 
+  #                             colour = "black", lineheight = 1, face = "bold"), 
+  #   legend.text = element_text(size = 6, 
+  #                              face = "plain", colour = "black"),
+  #   legend.position = "bottom"
+  # )
 p.cor.GCT.CEGMA <-
@@
   scale_fill_viridis_c(breaks = c(-1,-0.5, 0, 0.5, 1),labels = c(-1,-0.5, 0, 0.5, 1)) +
+  theme_pubclean() +
   theme(
@@
   arrange(Species)%>% 
-  knitr::kable(
+  kable(
     ., 
-    longtable = TRUE, 
     booktabs = TRUE,
     caption = 'Summary of duplications in Atlantogenata'
-  )
+  ) %>% 
+  kable_styling(latex_options = c("scale_down"))
 ```
@@
 A) A table summarizing duplication events in Atlantogenata.
-B) A tree of the species in Atlantogenata with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.", fig.height=8, fig.width=6}
+B) A tree of the species in Atlantogenata with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.", out.height="7in", out.width="6in"}
 
@@
 
-```{r Figure4, fig.cap="Correlations between genome quality metrics and ECNC metrics"}
+```{r Figure4, fig.cap="Correlations between genome quality metrics and ECNC metrics", fig.asp=1, out.width="5in", out.height="5in", fig.align="center"}
 f4 <- p.cor.GCT.CEGMA
@@
   select(-Direction) %>% 
-  knitr::kable(
+  mutate(Node = Node %>% str_replace_all("_", " ")) %>% 
+  kable(
     ., 
-    longtable = TRUE, 
     booktabs = TRUE,
     caption = 'Number of pathways overrepresented among duplicated genes at different FDRs.'
-  )
+  ) %>% 
+  kable_styling(latex_options = c("scale_down"))  # TODO: Make "Number of Pathways" a group, and make each column just "FDR<=XX"
 
@@
 
-```{r Figure5, fig.cap="Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.", fig.width=6, fig.height=8}
+```{r Figure5, fig.cap="Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.", out.width="6in", out.height="4in"}
 f5a <- p.loxAfr.ORA.dotplot
@@
 Overall in elephants, we see enrichments within duplicated genes for pathways involved in what makes an elephant an elephant - including tumor suppressor pathways.
+
+`r knitr::knit_exit()`
 
@@
 
-```{r, Supplementary Figure 4, fig.cap="Supplementary Figure 4: Dot and Line plot for Body Sizes and such", eval=F}
-p.bodysize.atlantogenata
-```
-
 ```{r Supplementary Figure 5, fig.cap="Supplementary Figure 5: All the Gene Copy Trees for interesting genes duplicated in LoxAfr4 (RIP any trees if this gets printed)", eval=F}
@@
 # plot.coolgenes
-```
-
-```{r Supplementary Figure 7, fig.cap="Supplementary Figure 7: Correlation matrix heat map for genome quality metrics, ECNC, RBHB Copy, and Estimated Copy Number (lesser between ECNC and RBHB)", eval=F}
-p.cor.GCT.CEGMA
 ```

### 2020-05-31 paper_PLOS/paper_PLOS_draft.tex
Added Figure, fixed figures 0-5

@@
 \bibliographystyle{elsarticle-harv}
-\usepackage{longtable}
 \ifxetex
@@
 \usepackage{pdflscape}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-)
+\begin{figure}
+
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure0-1} \hfill{}
+
+\caption{Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]}\label{fig:Figure0}
+\end{figure}
 
@@
 
-\begin{longtable}[]{@{}lrlllr@{}}
-\caption{Estimated Cancer Succeptibilty for nodes in
-Atlantogenata}\tabularnewline
+\begin{table}
+
+\caption{\label{tab:Fig2 Prep Table 1}Estimated Cancer Succeptibilty for nodes in Atlantogenata}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrlllr}
 \toprule
-Node & Est. Lifespan & K1 & K2 & Change in K & log2
-Change\tabularnewline
+Node & Est. Lifespan & K1 & K2 & Change in K & log2 Change\\
 \midrule
-\endfirsthead
-\toprule
-Node & Est. Lifespan & K1 & K2 & Change in K & log2
-Change\tabularnewline
-\midrule
-\endhead
-Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 &
-2.31\tabularnewline
-Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 &
-4.07\tabularnewline
-Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\tabularnewline
-Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 &
--2.31\tabularnewline
-Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 &
-0.00\tabularnewline
-Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 &
-27.66\tabularnewline
-Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 &
-2.31\tabularnewline
-Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 &
-4.14\tabularnewline
-Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\tabularnewline
-Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 &
--2.31\tabularnewline
-Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 &
-0.00\tabularnewline
-Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 &
--23.05\tabularnewline
-Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 &
-29.97\tabularnewline
-Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 &
-6.92\tabularnewline
-Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 &
-8.62\tabularnewline
-Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 &
-0.00\tabularnewline
-Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 &
--5.01\tabularnewline
-Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 &
-9.22\tabularnewline
-Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 &
-5.38\tabularnewline
-Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 &
--4.61\tabularnewline
-Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 &
--4.61\tabularnewline
-Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 &
-0.00\tabularnewline
-Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 &
-5.22\tabularnewline
-Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 &
-9.22\tabularnewline
-Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 &
-20.75\tabularnewline
-Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 &
--3.76\tabularnewline
-Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 &
-1.51\tabularnewline
-Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 &
-0.00\tabularnewline
-Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 &
-0.00\tabularnewline
+Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
+Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 & 4.07\\
+Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
+Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+\addlinespace
+Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 & 27.66\\
+Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
+Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 & 4.14\\
+Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
+\addlinespace
+Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 & -23.05\\
+Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 & 29.97\\
+Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 & 6.92\\
+Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 & 8.62\\
+\addlinespace
+Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 & 0.00\\
+Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 & -5.01\\
+Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 & 9.22\\
+Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 & 5.38\\
+Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
+\addlinespace
+Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
+Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 & 0.00\\
+Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 & 5.22\\
+Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 & 9.22\\
+Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 & 20.75\\
+\addlinespace
+Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 & -3.76\\
+Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 & 1.51\\
+Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 & 0.00\\
+Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 & 0.00\\
 \bottomrule
-\end{longtable}
+\end{tabular}}
+\end{table}
 
@@
 \hypertarget{identification-and-evolutionary-history-of-gene-duplications}{%
-\subsection*{Identification and evolutionary history of gene
+\subsection{Identification and evolutionary history of gene
 duplications}\label{identification-and-evolutionary-history-of-gene-duplications}}
-\addcontentsline{toc}{subsection}{Identification and evolutionary
-history of gene duplications}
+
+\begin{table}
+
+\caption{\label{tab:Figure3 Prep Genome Table}Summary of duplications in Atlantogenata}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lllrrllr}
+\toprule
+Species & Common Name & Size (g) & \#Hits & \#Duplicated & \% Genes Found & \% Hits Duplicated & Mean ECNC/Hit\\
+\midrule
+Choloepus hoffmanni & Hoffmans Two-Toed Sloth & 4.3e+03 & 14082 & 3204 & 78.19\% & 22.75\% & 0.98\\
+Chrysochloris asiatica & Cape Golden Mole & 49 & 13547 & 2716 & 75.22\% & 20.05\% & 0.99\\
+Dasypus novemcinctus & Nine-Banded Armadillo & 4.8e+03 & 13819 & 2605 & 76.73\% & 18.85\% & 0.98\\
+Echinops telfairi & Lesser Hedgehog Tenrec & 87 & 12903 & 1670 & 71.64\% & 12.94\% & 0.99\\
+Elephantulus edwardii & Cape Elephant Shrew & 49 & 12884 & 3048 & 71.53\% & 23.66\% & 0.99\\
+\addlinespace
+Elephas maximus & Asian Elephant & 3.3e+06 & 14073 & 907 & 78.14\% & 6.44\% & 1.00\\
+Loxodonta africana & African Savanna Elephant & 4.6e+06 & 14051 & 940 & 78.01\% & 6.69\% & 1.00\\
+Loxodonta cyclotis & African Forest Elephant & 4.7e+06 & 14065 & 900 & 78.09\% & 6.40\% & 1.00\\
+Mammut americanum & American Mastodon & 6e+06 & 13840 & 737 & 76.84\% & 5.33\% & 1.00\\
+Mammuthus columbi & Columbian Mammoth & 6.6e+06 & 13059 & 426 & 72.51\% & 3.26\% & 1.00\\
+\addlinespace
+Mammuthus primigenius & Woolly Mammoth & 4.3e+06 & 13935 & 723 & 77.37\% & 5.19\% & 1.00\\
+Orycteropus afer & Aardvark & 5.3e+04 & 13880 & 1083 & 77.06\% & 7.80\% & 0.99\\
+Palaeoloxodon antiquus & Straight Tusked Elephant & 1e+07 & 13969 & 745 & 77.56\% & 5.33\% & 1.00\\
+Procavia capensis & Rock Hyrax & 3e+03 & 13672 & 788 & 75.91\% & 5.76\% & 1.00\\
+Trichechus manatus & Manatee & 6.9e+05 & 14092 & 1046 & 78.24\% & 7.42\% & 1.00\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{figure}
+
+\includegraphics[width=6in,height=7in]{paper_PLOS_draft_files/figure-latex/Figure3-1} \hfill{}
+
+\caption{Gene duplications occur readily throughout Atlantogenata.
+A) A table summarizing duplication events in Atlantogenata.
+B) A tree of the species in Atlantogenata with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.}\label{fig:Figure3}
+\end{figure}
+
+Enhanced cancer suppression may have evolved through many mechanisms;
+among the most parsimonious is an increase in the copy number of genes
+with tumor suppressor functions. Previous studies focusing on candidate
+gene studies, for example, have identified increased copy number of the
+tumor suppressors \emph{TP53} and \emph{LIF} in elephants
+{[}12,21--24{]}. Therefore, in order to test whether this was pervasive
+genome-wide in \emph{Afrotherians}, we used a Reciprocal Best Hit BLAT
+(RBHB) approach to infer gene copy number in \emph{Afrotherian} and
+\emph{Atlantogenatan} genomes (Fig. 2A, Supplementary Figure 1). Because
+RBHB-like approaches can over-estimate copy number when genes are
+fragmented or incorrectly assembled across multiple scaffolds, we also
+inferred copy number using a complementary method that quantifies the
+ratio between observed and expected gene coverage per nucleotide (ECNC)
+(Supplementary Figure 1). By only including nucleotides from the query
+sequence that were observed in the target genome, we also correct for
+partial hits where some or all of the homologs of a gene have diverged
+from the human homolog.
+
+As our sequence database included various protein transcripts for each
+gene, in order to obtain gene-level copy number information and
+eliminate any many-to-one mappings of hits, we labeled each exon of
+every reciprocal best hit (RBH) with the gene corresponding to the query
+transcript and merged all overlapping exons; next, we eliminated any
+many-to-one exons that resulted from the previous step. Finally, we
+reassembled the gene loci based on the original transcript starts and
+ends, and the collapsed exon data, obtaining the full sequence of each
+RBH locus. Genes were considered to be duplicated if its copy number via
+RBHB was greater than or equal to 2, and the maximum ECNC among all
+transcripts prior to filtering was greater than or equal to 1.50 in
+order to account for partial hits that may be due to incomplete
+coverage.
+
+Once we applied these criteria to the results, we obtained the results
+shown in Figure 2A. Our approach positively identified an average of NA
+of the queries we searched. We observed that the percentage of
+duplicated genes in non-Pseudoungulatan genomes was significantly
+higher: while these genomes had duplicates percentages ranging anywhere
+from 3.26\% to 7.80\%, outgroup species' duplication rates ranged from
+12.94\% to 23.66\%.
+
+\begin{figure}
+
+{\centering \includegraphics[width=5in,height=5in]{paper_PLOS_draft_files/figure-latex/Figure4-1}
+
+}
+
+\caption{Correlations between genome quality metrics and ECNC metrics}\label{fig:Figure4}
+\end{figure}
+
+Given the preliminary status of the genomes for \emph{Chrysochloris
+asiatica} and \emph{Elephantulus edwardii}; and the fact that more
+revisions have been performed for Paenungulatan genomes; we hypothesized
+that genome quality may be confounding the results we've observed. We
+used \texttt{CEGMA} and the \texttt{gVolante} server {[}31,32{]} to
+assess the quality of all our genomes, to see if any of these metrics
+correlated with our metrics of gene counts and copy number. As shown in
+Figure 3, mean ECNC, mean Copy Number, and mean CN (the lesser of Copy
+Number and ECNC per gene) correlate moderately strongly with genomic
+quality metrics related to the length and assembly quality, like LD50
+and the number of scaffolds and contigs with a length above either 100K
+or 1M, supporting our hypothesis that increases in copy number correlate
+.
+
+In order to determine when in the phylogeny our identified duplications
+occurred, we used maximum likelihood to reconstruct ancestral gene copy
+numbers across the phylogeny. We encoded the copy number of each gene as
+a discrete trait ranging from 0 (``?'') to 31+, and then ran the
+ancestral gene copy number reconstruction using the phylogenetics suite
+\texttt{IQTREE}, which rapidly tests and runs the best evolutionary
+model for a given trait of interest. Importantly, \texttt{IQTREE} can be
+set to only output states where the likelihood of the state is greater
+than a certain threshold. Setting the threshold to 0.8, we obtained the
+results shown in Figure 2B.
+
+Gene duplications events occured readily thoughout \emph{Atlantogenata}
+(Figure 2B). Among the genes that increased in copy number in the
+elephant lineage are TP53 and LIF, as previously described; however, we
+find that these two genes represent a fraction of the 940 genes that are
+duplicated in the African Elephant overall, which accumulated over
+various steps through their evolution. While the extinct elephantids
+have acceptable genome quality metrics according to CEGMA, they are
+nonetheless missing a significant number of sequences; this may
+contribute to the low number of duplicated genes that occured in
+internal nodes. The number of duplicates that occur at each branch is
+also proportional to the density of the sampling of the clade overall,
+as would be expected. In branches, such as \emph{Afrosoricida}, where
+the number of species is relatively minuscule compared to the size of
+the clade, we see many significantly larger numbers of duplications
+private to these species.
+
+\hypertarget{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}{%
+\subsection{Duplications that occurred recently in Probodiscea are
+enriched for tumor suppressor
+pathways}\label{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}}
+
+\begin{table}
+
+\caption{\label{tab:Branches which experienced an increase in size}Number of pathways overrepresented among duplicated genes at different FDRs.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{llrrrr}
+\toprule
+Ancestor & Node & Pathways at FDR<0.1 & Pathways at FDR<0.2 & Pathways at FDR<0.3 & Pathways at FDR<0.5\\
+\midrule
+Afroinsectivora & Elephantulus edwardii & 252 & 37 & 30 & 87\\
+Afrosoricida & Chrysochloris asiatica & 90 & 48 & 43 & 105\\
+Afrosoricida & Echinops telfairi & 0 & 2 & 0 & 31\\
+Afrotheria & Afroinsectivora & 0 & 0 & 0 & 33\\
+Loxodontini & Loxodonta africana & 6 & 0 & 1 & 0\\
+\addlinespace
+Paenungulata & Tethytheria & 0 & 0 & 0 & 2\\
+Proboscidea & Mammut americanum & 0 & 0 & 0 & 6\\
+Pseudoungulata & Orycteropus afer & 27 & 67 & 29 & 67\\
+Tethytheria & Proboscidea & 0 & 3 & 0 & 3\\
+Tethytheria & Trichechus manatus & 4 & 0 & 0 & 2\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{figure}
+
+\includegraphics[width=6in,height=4in]{paper_PLOS_draft_files/figure-latex/Figure5-1} \hfill{}
+
+\caption{Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.}\label{fig:Figure5}
+\end{figure}
+
+Our initial hypothesis was that genes which duplicated in lineages that
+experienced a growth in size would be enriched for membership in tumor
+suppressor pathways. Thus, we used WebGestalt and its Overrepresentation
+Analysis (ORA) functionality to determine what pathways were enriched in
+our duplicated gene sets in each branch relative to our initial query
+set. For our database, we used Reactome for our primary analysis, but
+additionally used the KEGG, Panther, Wikipathways, and
+Wikipathways\_cancer databases using WebGestalt ORA. Going through the
+tree, at no FDR\textless{}0.5 is there any significant pathway
+representation for genes that increased in the branches leading to , or
+\emph{Paenungulata}; furthermore, there are no significant pathway
+enrichments at FDR\textless{}0.5 for genes whose copy number did not
+change between branches (copy-number-stable) for , or \emph{Afrotheria}.
+Note that because \emph{Xenarthra} was selected as the outgroup, it is
+not possible to polarize the changes in their gene copy numbers along
+the tree.
+
+For the other branches, the number of pathways that came up as
+significantly enriched at each FDR is shown in Figure 3A (Supplementary
+Figure 3). For the species with high duplication rates and
+lower-quality, highly-fragmented genomes, such as with
+\emph{Chrysochloris asiatica} (20.05\% duplicated hits) and
+\emph{Elephantulus edwardii} (23.66\% duplicated hits), it is
+unsurprising that there is a proportionally large number of pathway
+enrichments. In the case of these two species, their many pathway
+enrichments also span an incredible range of processes at every level of
+biology; this, in combination with the high number of copy numbers
+identified for these genes, further suggests a need for improvement and
+refinement in these genomes. In the cell cycle pathways called as
+significant in the genomes of these two species plus \emph{Orycteropus
+afer} and \emph{Echinops telfairi}, the duplicated genes included in
+these sets are from the same gene families, such as the APC subunit
+family; the proteosome subunit families; and the protein phosphatase 2
+family, among others. It is highly possible that these results reflect
+true expansions of these gene families, especially in the higher-quality
+\emph{Orycteropus afer} genome; however, it is also possible that it
+simply reflects artificial duplication events, and so require further
+study.
+
+The pathway enrichments for genes whose copy number either did not
+change, or whose copy number increased, between \emph{Loxodona} and the
+African Elephant are shown in Figure 3B and 3C, respectively. Among the
+few enriched pathways in the African Elephant, we see that that two
+tumor suppression pathways - APC Complex-related pathways, and ``TP53
+Regulates Metabolic Genes''- appear not only in the case of stable
+genes, but also in the set of newly duplicated genes. The other pathways
+we see in the set of recently-duplicated genes include
+``Functionalization of Compounds'' and its daughter pathway
+``Xenobiotics''. Genes in these pathways serve to add functional groups
+to lipophylic compounds which would otherwise not be reactive in the
+cell, and are types of metabolic pathways. In the stable set we see
+enrichment of pathways such as ``Neuronal Systems'' and ``Axon
+Guidance,'' which fit in well with what is known about elephant biology
+and evolution {[}47{]}. Overall in elephants, we see enrichments within
+duplicated genes for pathways involved in what makes an elephant an
+elephant - including tumor suppressor pathways.
+
+\hypertarget{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}{%
+\subsection*{\texorpdfstring{Concerted duplication of TP53 and
+TP53-related genes towards
+\emph{Probodiscea}}{Concerted duplication of TP53 and TP53-related genes towards Probodiscea}}\label{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}}
+\addcontentsline{toc}{subsection}{Concerted duplication of TP53 and
+TP53-related genes towards \emph{Probodiscea}}
 
@@
 
+\leavevmode\hypertarget{ref-Goodman2009}{}%
+47. Goodman M, Sterner KN, Islam M, Uddin M, Sherwood CC, Hof PR, et al.
+Phylogenomic analyses reveal convergent patterns of adaptive evolution
+in elephant and human ancestries. Proceedings of the National Academy of
+Sciences. 2009;106: 20824--20829.
+doi:\href{https://doi.org/10.1073/pnas.0911239106}{10.1073/pnas.0911239106}
+
 \end{document}

### 2020-05-31 code/generalFunctions.R

@@ -675,7 +675,22 @@ graph.ORA.dotplot <- function(ORA.df, plotly=F){
     # geom_text_repel(data=. %>% filter(Class %in% c("TP53 Pathway", "Senescence", "Cell Death Pathways")), nudge_y = 0.5, force = 10, show.legend=F) + 
     scale_color_d3(palette="category20") +
     theme_pubclean() + 
-    theme(legend.position = "bottom")+
+    labs_pubr() + 
+    theme(
+      text = element_text(face = "plain", 
+                          colour = "black", lineheight = 0.9, 
+                          hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                          debug = FALSE), 
+      axis.text = element_text(size = 6, 
+                               colour = "black", face = "bold"),
+      axis.title = element_text(size = 8, 
+                                colour = "black", face = "bold"),
+      plot.title = element_text(size = 8, 
+                                colour = "black", lineheight = 1, face = "bold"), 
+      legend.text = element_text(size = 6, 
+                                 face = "plain", colour = "black"),
+      legend.position = "bottom"
+    ) +
     guides(color=guide_legend(title="Cell Cycle Pathway", title.position = "bottom", title.hjust = 0.5), size = guide_legend(title="Size (log)", title.position = "bottom", title.hjust = 0.5)) 
   if(plotly) {p %<>% ggplotly}
   return(p)

### 2020-06-03 paper_PLOS/paper_PLOS_draft.Rmd
Refactored figures and  tables, fixed Figure 6, broke Figure 4&7.

@@
 output:
-  rticles::elsevier_article:
+  bookdown::pdf_book:
+    base_format: rticles::elsevier_article
+    number_sections: yes
     includes:
@@
 ```{r Libraries and functions, include = F}
-knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=6, 
+knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, out.width="6in",
                       fig.retina = 2, fig.align = "left", dpi = 300)
@@
 
-t.bodysize.atlantogenata.withNodes <- st.Atlantogenata %>% 
-  as_tibble() %>% 
-  filter(!is.na(label)) %>%
-  select(
-    `Node`=label, 
-    `Size (log(g))`=Median, 
-    `95% CI (Low)` = `95%CI_Low`, 
-    `95% CI (High)` = `95%CI_High`,  
-    `Rate (sqrt)`=sqrt.rate
-  ) %>%
-  mutate_at(vars(-`Node`, ), . %>% as.numeric %>% round(2)) %>% 
-  kable(
-    ., 
-    longtable = TRUE, 
-    booktabs = TRUE,
-    caption = 'All body size estimates for _Atlantogenata_.'
-  )
-
 p.bodysize.atlantogenata <- st.Atlantogenata %>% 
@@
 
-```{r Figure0Prep}
+```{r Figure1Prep}
 
@@
 ```
-```{r Figure0, fig.cap="Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]",  out.width="6in", fig.asp=1.3}
-atlantogenata.tree.size.lifespan
-```
 
@@
 
-One of the major constraints on the evolution of large body sizes in animals is an increased risk of developing cancer.
-If all cells in all organisms have a similar risk of malignant transformation and equivalent cancer suppression mechanisms, organism with many cells should have a higher prevalence of cancer than organisms with fewer cells.
-Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, human cancer incidence increases with increasing adult height [@Green2011; @Nunney:20181c2] and cancer incidence is positively correlated with body size in dogs [@Dobson2013; @Dorn1968].
-There is no correlation, however, between body size and cancer risk between species.
-This lack of correlation is often referred to as ‘Peto’s Paradox’ [@Caulin2011; @Leroi2003; @Peto1975].
-While it is clear that a resolution to Peto’s Paradox must involve the evolution of enhanced cancer protection alongside increases in body size and lifespan, the specific genetic, molecular, and cellular mechanisms that underlie this resistance have proven elusive.
-[c.f.
+Among the major constraints on the evolution of large body sizes (and long life-spans) in animals is an increased risk of developing cancer.
+If all cells in all organisms have a similar risk of malignant transformation and equivalent cancer suppression mechanisms, organism with many cells should have a higher prevalence of cancer than organisms with fewer cells, particularly because large and small animals have similar cell sizes (@Savage2007).
+Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, cancer incidence increases with increasing adult height in humans [@Green2011; @Nunney:20181c2] and in dogs [@Dobson2013; @Dorn1968].
+There is no correlation, however, between body size and cancer risk _between_ species; this lack of correlation is often referred to as ‘Peto’s Paradox’ [@Caulin2011; @Leroi2003; @Peto1975].
+The ultimate resolution to Peto’s Paradox is obvious: large bodied and long-lived species evolved enhanced cancer protection mechanisms.
+However, identifiying the specific genetic, molecular, and cellular mechanisms that underlie the evolution of augmented cancer protection has been difficult [c.f.
 @Ashur-Fabian2004; @Seluanov2008; @Gorbunova2012; @Tian2013; @Sulak2015].
@@
 Among the challenges for discovering how animals evolved enhanced cancer protection mechanisms is identifying lineages in which large bodied species are nested within species with small body sizes.
-Afrotherian mammals are generally small-bodied, similarly to the prediced common ancestor of Eutherian mammals.
+_Afrotherian_ mammals are generally small-bodied, but also include the largest extant land mammals.
 For example, maximum adult weights are ~70g in golden moles, ~120g in tenrecs, ~170g in elephant shrews, ~3kg in hyraxes, and 60kg in aardvarks [@HAGR].
-However, while these extant species are relatively small, the fossil evidence demonstrates that their ancestral lineages reached enormous sizes.
-For example, while extant hyraxes are relatively small, the extinct Titanohyrax is estimated to have weighted up to ~1300kg [@Schwartz1995].
-The largest members of Afrotheria, too, are dwarfed by the size of their recent ancestors: extant cows manatees are large bodied (~322-480kg) but are relatively small compared to the extinct Stellar’s sea cow which is estimated to have weight 8000-10000kg [@Scheffer1972].
+While extant hyraxes are relatively small, the extinct Titanohyrax is estimated to have weighted up to ~1300kg [@Schwartz1995].
+The largest members of _Afrotheria_, too, are dwarfed by the size of their recent ancestors: extant cows manatees are large bodied (~322-480kg) but are relatively small compared to the extinct Stellar’s sea cow which is estimated to have weight 8000-10000kg [@Scheffer1972].
 Similarly African (4,800kg) and Asian elephants (3,200kg) are the largest living elephant species, but are dwarfed by the truly gigantic extinct Proboscideans such as Deinotherium (~132,000kg), Mammut borsoni (110,000kg), and the Asian straight-tusked elephant (~220,000kg), the largest known land mammal [@Larramendi:20151c2].
-These large-bodied Afrotherian lineages are nested within small bodied species (Fig.
-1) [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], indicating that gigantism independently evolved in hyraxes, sea cows, and elephants (Paenungulates).
+Remarkably, these large-bodied _Afrotherian_ lineages are nested within small bodied species (__Figure \@ref(fig:Figure-intro)__) [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], indicating that gigantism independently evolved in hyraxes, sea cows, and elephants (_Paenungulata_).
 Thus, Paenungulates are an excellent model system in which to explore the mechanisms that underlie the evolution of large body sizes and augmented cancer resistance.
@@
 Although many mechanisms can potentially resolve Peto’s paradox, among the most parsimonious routes to enhanced cancer resistance is through an increased copy number of tumor suppressors.
-Such an example has been seen in the case of candidate genes such as _TP53_ and _LIF_ [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018] as well as in studies involving a limited set of candidate genes [@Caulin2015; @Doherty2016].
-As these studies focus on _a priori_ gene sets, however, it remains unknown whether this is a general, genome-wide trend in Afrotherian genomes; and whether such a general trend is associated with the recent increases in body size – and therefore expected cancer risk – in these species.
-
-Here, we trace the evolution of body mass and gene copy number variation in Afrotherians in order to investigate whether gene duplications are enriched in large, long-lived species for genes involved in known tumor suppression pathways.
-Our estimates of the evolution of body mass, similarly to previous studies [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], show that large body masses evolved in a step-wise manner, with major increases in body mass in the Pseudoungulata (17kg), Paenungulata (25kg), Tethytheria (296kg), and Proboscidea (4,100kg) stem-lineages.
-Furthermore, we see that the ancestral body size increases in Hydracoidia and Sirenia were independent events.
-To study the evolution of gene copy number, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene duplications in Afrotherian genomes, and used maximum likelihood (treating copy number as a discrete trait) to infer the lineages in which those duplications occurred.
-We found gene duplications in lineages with increased body mass were enriched in functions related to tumor suppression, including regulation of the cell cycle, DNA damage repair, and regulation of apoptosis.
-These data suggest that duplication of tumor suppressors played a role in the evolution of large, long-lived in Afrotherians.
+Indeed, candidate genes studies have found that the elephant genome encodes duplicate such as _TP53_ and _LIF_ [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018] as well as other genes with putative tumor suppressive functions [@Caulin2015; @Doherty2016].
+As these studies focus on _a priori_ gene sets, however, it remains unknown whether this is a general, genome-wide trend in _Afrotherian_ genomes; and whether such a general trend is associated with the recent increases in body size – and therefore expected cancer risk – in these species.
+
+Here, we trace the evolution of body mass and gene copy number variation in across _Afrotherian_ genomes in order to investigate whether duplication of tumors suppresor is common in large, long-lived Proboscideans.
+Our estimates of the evolution of body mass, similarly to previous studies [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], show that large body masses evolved in a step-wise manner, with major increases in body mass in the _Pseudoungulata_ (17kg), _Paenungulata_ (25kg), _Tethytheria_ (296kg), and _Proboscidea_ (4,100kg) stem-lineages.
+To explore whether duplication of tumor suppressor genes occurred coincident with the evolution of large body sizes, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) strategy to identify gene duplications, and used maximum likelihood to infer the lineages in which those duplications occurred.
+Unexpectedly, we found that duplication of tumor suppressor genes was common in all Afrotherians, both large and small.
+These data suggest that duplication of tumor suppressor genes is pervasive in Afrotherians and proceeded the evolution of species with very large body sizes.
 
@@
 Therefore we replaced the Afrotherian clade in the Bininda-Emonds _et al._ [-@Bininda-Emonds2008] phylogeny with the Afrotherian phylogeny of Puttick and Thomas [-@PuttickAndThomas2015] using Mesquite.
-Next, we jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of the Brownian motion model that relaxes assumptions of neutrality and gradualism by considering increments to evolving characters to be drawn from a heavy-tailed stable distribution (the "Stable Model") [@ElliotAndMooers2014].
+Next, we jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of the Brownian motion model that relaxes assumptions of neutrality and gradualism by considering increments to evolving characters to be drawn from a heavy-tailed stable distribution (the "Stable Model") implemented in StableTraits [@ElliotAndMooers2014].
 The stable model allows for occasional large jumps in traits and has previously been shown to out-perform other models of body mass evolution, including standard Brownian motion models, Ornstein–Uhlenbeck models, early burst maximum likelihood models, and heterogeneous multi-rate models [@ElliotAndMooers2014].
@@
 
-__Reciprocal Best-Hit BLAT:__ We developed a reciprocal best hit BLAT (RBHB) pipeline to identify putative homologs and estimate gene copy numbers across species (__Figure 1A__).
+__Reciprocal Best-Hit BLAT:__ We developed a reciprocal best hit BLAT (RBHB) pipeline to identify putative homologs and estimate gene copy number across species (\@ref(fig:Figure-RBHBStrategy)).
 The Reciprocal Best Hit (RBH) search strategy is conceptually straightforward: 1) Given a gene of interest $G_A$ in a query genome $A$, one searches a target genome $B$ for all possible matches to $G_A$; 2) For each of these hits, one then performs the reciprocal search in the original query genome to identify the highest-scoring hit; 3) A hit in genome $B$ is defined as a homolog of gene $G_A$ if and only if the original gene $G_A$ is the top reciprocal search hit in genome $A$.
@@
 __Effective Copy Number By Coverage:__ In lower-quality genomes, many genes are fragmented across multiple scaffolds, which results in BLAT calling multiple hits when in reality there is only one gene.
-To compensate for this, we came up with a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we see each nucleotides of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (Supplementary Figure 1).
+To compensate for this, we came up with a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we see each nucleotides of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (__Figure \@ref(fig:Figure-ECNC)__).
 This allows us to correct for genes that have been fragmented across incomplete genomes, while also taking into account missing sequences from the human query in the target genome.
@@
 
-After filtering out problematic protein queries (see "Query gene inclusion criteria"), we then used our pipeline (Figure 1A) to search for all copies of our `r n.GenesSearched` query genes in publicly available Afrotherian genomes, including African savannah elephant (_Loxodonta africana_: loxAfr3, loxAfr4, loxAfrC), African forest elephant (_Loxodonta cyclotis_: loxCycF), Asian Elephant (_Elephas maximus_: eleMaxD), Woolly Mammoth (_Mammuthus primigenius_: mamPriV), Colombian mammoth (_Mammuthus columbi_: mamColU), American mastodon (_Mammut americanum_: mamAmeI), Rock Hyrax (_Procavia capensis_: proCap1, proCap2, proCap2_HiC), West Indian Manatee (_Trichechus manatus latirostris_: triManLat1, triManLat1_HiC), Aardvark (_Orycteropus afer_: oryAfe1, oryAfe1_HiC), Lesser Hedgehog Tenrec (_Echinops telfairi_: echTel2), Nine-banded armadillo (_Dasypus novemcinctus_: dasNov3), Hoffman's two-toed sloth (_Choloepus hoffmannii_: choHof1, choHof2, choHof2_HiC), Cape golden mole (_Chrysochloris asiatica_: chrAsi1), and Cape elephant shrew (_Elephantulus edwardii_: eleEdw1).
-For many of these species, we covered multiple assemblies in order to test the effects of assembly size and quality on our hits.
+After filtering out problematic protein queries (see below), we then used our pipeline to search for all copies of our `r n.GenesSearched` query genes in publicly available _Afrotherian_ genomes (\@ref(tab:Table-Genomes)), including African savannah elephant (_Loxodonta africana_: loxAfr3, loxAfr4, loxAfrC), African forest elephant (_Loxodonta cyclotis_: loxCycF), Asian Elephant (_Elephas maximus_: eleMaxD), Woolly Mammoth (_Mammuthus primigenius_: mamPriV), Colombian mammoth (_Mammuthus columbi_: mamColU), American mastodon (_Mammut americanum_: mamAmeI), Rock Hyrax (_Procavia capensis_: proCap1, proCap2, proCap2_HiC), West Indian Manatee (_Trichechus manatus latirostris_: triManLat1, triManLat1_HiC), Aardvark (_Orycteropus afer_: oryAfe1, oryAfe1_HiC), Lesser Hedgehog Tenrec (_Echinops telfairi_: echTel2), Nine-banded armadillo (_Dasypus novemcinctus_: dasNov3), Hoffman's two-toed sloth (_Choloepus hoffmannii_: choHof1, choHof2, choHof2_HiC), Cape golden mole (_Chrysochloris asiatica_: chrAsi1), and Cape elephant shrew (_Elephantulus edwardii_: eleEdw1).
 
 __Query gene inclusion criteria:__ To assemble our query list, we first removed all unnamed genes from UP000005640.
-Next, we excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (`r query.binTable %>% filter(C.ORF) %>% nrow`), LOC (`r query.binTable %>% filter(LOC) %>% nrow`), HLA genes (`r query.binTable %>% filter(HLA) %>% nrow`), replication dependent histones (`r query.binTable %>% filter(Histone) %>% nrow`), odorant receptors (`r query.binTable %>% filter(OlfactoryReceptor) %>% nrow`), ribosomal proteins (`r query.binTable %>% filter(Riboprotein) %>% nrow`), zinc finger transcription factors (`r query.binTable %>% filter(ZincFinger) %>% nrow`), viral and repetitive-element-associated proteins (`r query.binTable %>% filter(Viral) %>% nrow`) and any protein described as either "Uncharacterized," "Putative," or "Fragment" by UniProt in UP000005640 (`r query.binTable %>% filter(Fragment | Uncharacterized | Putative) %>% nrow`), leaving us with a final set of `r n.TranscriptsSearched` query protein sequences, corresponding to `r n.GenesSearched` genes.
+Next, we excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (`r query.binTable %>% filter(C.ORF) %>% nrow` genes), LOC (`r query.binTable %>% filter(LOC) %>% nrow` genes), HLA genes (`r query.binTable %>% filter(HLA) %>% nrow` genes), replication dependent histones (`r query.binTable %>% filter(Histone) %>% nrow` genes), odorant receptors (`r query.binTable %>% filter(OlfactoryReceptor) %>% nrow` genes), ribosomal proteins (`r query.binTable %>% filter(Riboprotein) %>% nrow` genes), zinc finger transcription factors (`r query.binTable %>% filter(ZincFinger) %>% nrow` genes), viral and repetitive-element-associated proteins (`r query.binTable %>% filter(Viral) %>% nrow` genes) and any protein described as either "Uncharacterized," "Putative," or "Fragment" by UniProt in UP000005640 (`r query.binTable %>% filter(Fragment | Uncharacterized | Putative) %>% nrow` genes), leaving us with a final set of `r n.TranscriptsSearched` query protein isoforms, corresponding to `r n.GenesSearched` genes.
 
@@
 The resulting gene-level copy number table was then combined with the maximum ECNC values observed for each gene in order to call gene duplications.
-We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes [@Sulak2015; @Vazquez2018], therefore partial gene duplications were included provided they passed additional inclusion criteria.
+We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes [@Sulak2015; @Vazquez2018], therefore partial gene duplications were included provided they passed additional inclusion criteria (see below).
 The ECNC cut off of 1.5 was selected empirically, as this value minimized the number of false positives seen in a test set of genes and genomes.
-The results of our initial search are summarized in Figure 1B`r warning('TODO: Check if this is still the case')`.
+The results of our initial search are summarized in Figure \@ref(fig:Figure-GeneDup-Cladogram).
 Overall, we identified `r median.genesPerGenome` genes across all species, or `r median.genesPerGenome %>% divide_by(n.GenesSearched) %>% multiply_by(100) %>% round(digit=1) %>% paste0(.,"%")` of our starting query genes.
@@
 
-## Evidence for Functionality of Identified Genes
+## Evidence for Functionality of Gene Duplicates
 
@@
 
-__Transcriptome Assembly:__ For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated _de novo_ transcriptomes using publically-available RNA-sequencing data from NCBI SRA (Supplementary Table 1).
+__Transcriptome Assembly:__ For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated _de novo_ transcriptomes using publically-available RNA-sequencing data from NCBI SRA (\@ref(tab:Table-SRAData)).
 We mapped reads to all genomes available for each species, and assembled transcripts using HISAT2 and StringTie, respectively [@HISAT; @StringTie; @Tuxedo].
@@
 
-We implemented a maximum likelihood method for determining the ancestral copy numbers of genes in _Atlantogenata_ using IQ-Tree.
-For this analysis, we used an unrooted species tree including only the aforementioned _Atlantogenata_ species.
-We generated PHYLIP files containing the copy number of each gene in the highest quality genome for each species, encoding genes on a scale from 1-31+ copies as 1-9, A-V; and encoding a gene's copy number as uncetain ("?") when we did not identify it in the genome.
-We used the included tree-searching and model-testing functionality in IQ-Tree to determine the most likely topology for the species tree, and to obtain the most likely model for copy number changes in the genome.
-The most likely model for the evolution of copy number was inferred to be a Jukes-Cantor type model for morphological data with equal state frequencies, allowing for a proportion of invariable sites, and using a discrete Gamma model with 4 rate categories.
-("MK+FQ+I+G4").
-We defined the ancestral state of a node if it had greater than an 80% posterior probability.
+We encoded the copy number of each gene for each species as a discrete trait ranging from 0 (one gene copy) to 31 (for 32+ gene copies) and used IQ-TREE to select the best-fitting model of character evolution [@IQTREE; @UFBoot2; @ModelFinder; @PMSF; @Schrempf2019], which was inferred to be a Jukes-Cantor type model for morphological data (MK) with equal character state frequencies (FQ) and rate heterogeneity across sites approximated by including a class of invariable sites (I) plus a discrete Gamma model with four rate categories (G4).
+Next we inferred gene duplication and loss events with the empirical Bayesian ancestral state reconstruction (ASR) method implemented in IQ-TREE [@IQTREE; @UFBoot2; @ModelFinder; @PMSF; @Schrempf2019], the best fitting model of character evolution (MK+FQ+GR+I) [@Soubrier2012; @Yang1995], and the unrooted species tree for Atlantogenata.
+We considered ancestral state reconstructions to be reliable if they had Bayeisan Posterior Probability (BPP) ?0.80; less reliable reconstructions were excluded for pathway analyses.
 
@@
 
-To determine which pathways were associated with duplicated genes in each species and lineage, we used WEBGESTALT to perform overrepresentation analysis (ORA) of the duplicated gene lists relative to our initial query gene list [@WebGestalt2019].
-For the database of pathways used in the analysis, we used Reactome [@Reactome], Wikipathways, [@Wikipathways], and KEGG [@KEGG].
-For the ORA, we used FDR for determining significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and FDR=0.5.
-
-## Lifespan Phylogenetic Least-Square Regression and Calculating Estimated Cancer Risk Thoughout Atlantogenata
-
-In order to determine the cancer risk $K$ at each node, we first needed to calcultate ancestral lifespans at each node.
-To do so, we used a Phylogenetic Generalized Least-Square Regression (PGLS) [@Felsenstein1985; @MartinsAndHansen1997] to calculate estimated ancestral lifespans across _Atlantogenata_ using our estimates for body size at each node.
-
-Next, we used a simplified multistage cancer risk model for body size $D$ and lifespan $t$: $K \approx Dt^6$ [@Armitage:19851c2; @Armitage:20041c2; @Peto:20151c2; @Peto1975].
-To determine the change in $K$ between nodes, we obtained the ratio between the cancer risk $K_{1}$ at any given node, and the cancer risk $K_{2}$ at its ancestral node, using the equation $\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}$.
-Finally, to simplify comparisons, we calculated the fold change cancer risk between a node and its ancestor as $\log_{2}(\frac{K_{2}}{K_{1}})$.
+To determine if gene duplications were enriched in particular biological pathways, we used the WEB-based Gene SeT AnaLysis Toolkit (WEBGESTALT)[@WebGestalt2019] to perform overrepresentation analysis (ORA); pathway databases included Reactome [@Reactome], Wikipathways, [@Wikipathways], and KEGG [@KEGG].
+Gene duplicates in each lineage were used as the foreground gene set, and the initial query set was used as the background gene set.
+Statistical significance of enriched terms was assessed with a hypergeometric test, and controlled for multiple testing using the Benjamini-Hochberg false discovery rate (FDR); each analysis was run at FDR=0.1, FDR=0.2, FDR=0.3, and FDR=0.5.
+In order to determine an empirical false positive rate for term enrichment, we randomly sampled 1000-10,000 genes from our background set 1000 times, and ran the aforementioned analyses to see which terms were likely to randomly appear; there were no terms which appeared at FDR $\leq$ 0.3.
+
+## Estimating the Evolution of Cancer Risk
+
+The dramatic increase in body mass and lifespan in some Afrotherian lineages implies those lineages must have also evolved reduced cancer risk.
+To infer the magnitude of these reductions we estimated differences in intrinsic cancer risk across extant and ancestral Afrotherians.
+Following Peto [@Peto:20151c2] we estimate the intrinsic cancer risk as the product of risk associated with body mass and lifespan.
+In order to determine the intrinsic cancer risk ($K$) across species and at ancestral nodes (see below), we first needed to estimate ancestral lifespans at each node.
+We used Phylogenetic Generalized Least-Square Regression (PGLS) [@Felsenstein1985; @MartinsAndHansen1997] implemented in the R package `ape` to calculate estimated ancestral lifespans across _Atlantogenata_ using our estimates for body size at each node.
+In order to estimate the intrinsic cancer risk of a species, we first inferred lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrowninan + \beta_{2}ln(size) + \epsilon$.
+Next, we calculated $K_{1}$ at all nodes, and then estimated the fold change in cancer susceptibility between ancestral and descendant nodes (Figure \@ref(fig:Figure-CancerSuscept), Table \@ref(tab:Table-CancerSuscept)).A Next, in oder to calculate $K_{1}$ at all nodes, we used a simplified multistage cancer risk model for body size $D$ and lifespan $t$: $K \approx Dt^6$ [@Armitage:19851c2; @Armitage:20041c2; @Peto:20151c2; @Peto1975].
+The fold change in cancer risk between a node and its ancestor was then defined as $\log_{2}(\frac{K_{2}}{K_{1}})$.
 
@@
   separate_rows(Genes, sep=";") %>% 
-  group_by(Pathway) %>% 
+  # group_by(Pathway) %>% 
+    group_by(Pathway, Database) %>% 
   summarize_all(. %>% unique %>% toString) %>% 
   ungroup %>% 
-    select(Node, RBHB, Pathway, Genes, FDR.Threshold) %>% 
+    select(Node, RBHB, Database, Pathway, Genes, FDR.Threshold) %>% 
     distinct %>% 
@@
   separate_rows(Genes, sep=", ") %>% 
-  group_by(Genes) %>% 
+  group_by(Genes, Database) %>% 
   summarize_all(. %>% unique %>% toString) %>% 
@@
 
-## Body size frequently and independently expands and contracts throughout _Atlantogenata_
-
-```{r Figure1, fig.cap="Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
-**A)** Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch.
-Atlantogenata is highlighted at the bottom.
-**B)** Zoom-in of (**A**) on Atlantogenata.
-Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size.
-**C)** Confidence interval plot for representative species and ancestral nodes.", out.width="6in", fig.asp=1}
-f1a <- p.tree
-f1b <- p.tree.atlantogenata
-f1c <- p.bodysize.atlantogenata
-# tree.Atlantogenata %>% tree_subset(., MRCA(.,"Loxodonta.africana", "Elephantulus.edwardii"), levels_back=3) %>% ggtree(aes(color=lnSize)) + geom_tiplab() +scale_color_viridis_c() +
-#   geom_point2(aes(subset=(node==MRCA(tree.Atlantogenata, "Loxodonta.africana", MRCA(tree.Atlantogenata, "Loxodonta.africana", "Mammut.americanum"))), color=lnSize))# Paenungulates, simple view
-#   # geom_point2(aes(subset=(node== MRCA(tree.Atlantogenata, "Loxodonta.africana", "Echinops.telfairi")), color=lnSize))# Paenungulates, simple view
-f1ab <- ggarrange(
-  f1a, f1b, 
-  nrow=1, ncol = 2, 
-  labels="AUTO",
-  widths = c(1,2),
-  font.label = list(size=6, face="bold")
-  )
-f1 <-
-  ggarrange(
-  ggplotify::as.ggplot(f1ab),
-  f1c,
-  nrow=2, ncol=1,
-  labels = c("", "C"),
-  font.label = list(size=6, face="bold")
-)
-f1
-```
-
-To trace the evolutionary history of body mass and lifespan in Afrotherians, we built a time-calibrated supertree of Eutherian mammals combining 1,679 species from Bininda-Emonds et al [@Bininda-Emonds2008] with a total evidence Afrotherian phylogeny including 77 extant and fossil data from 39 extinct species [@PuttickAndThomas2015].
+## Step-wise evolution of body size in _Afrotherians_
+
+<!-- To trace the evolutionary history of body mass and lifespan in Afrotherians, we built a time-calibrated supertree of Eutherian mammals combining 1,679 species from Bininda-Emonds et al [@Bininda-Emonds2008] with a total evidence Afrotherian phylogeny including 77 extant and fossil data from 39 extinct species [@PuttickAndThomas2015].
 Fossil data from extinct species were included to ensure that ancestral state reconstructions of body mass in Afrotherians were not biased by only including extant species, which can lead to inaccurate reconstructions, for example, if lineages multiple lineages evolved large body masses from a small bodied ancestor.
 We jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of a Brownian model of character evolution, which allows for occasional large jumps in traits (stable model) and out performs standard Brownian motion and Ornstein-Uhlenbeck models of character evolution [@ElliotAndMooers2014].
-
-Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Orycteropus.afer")` in body mass in the stem-lineage of _Pseudoungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Orycteropus.afer")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Procavia.capensis")` in body mass in the stem-lineage of _Paenungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Procavia.capensis")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Trichechus.manatus")` in body mass in the stem-lineage of _Tehthytheria_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Trichechus.manatus")`), and a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Mammut.americanum")` in body mass in the stem-lineage of _Proboscidea_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Mammut.americanum")`) (Figure 1B,C).
-The ancestral _Hyracoidea_ was inferred to be relatively small `r text.bodysize.ancNode.byMRCA.CI("Microhyrax.lavocati", "Procavia.capensis")`, and rate accelerations were coincident with independent body mass increases in large hyraxes such as _Titanohyrax andrewsi_ `r text.bodysize.byLabel("Titanohyrax.andrewsi")`; `r text.ratio.nodeByMRCA.ancNode(nodeA="Titanohyrax.andrewsi","Orycteropus.afer")`.
-While the body mass of the ancestral _Sirenian_ was inferred to be large `r text.bodysize.ancNode.byMRCA.CI("Trichechus.manatus", "Prorastomus.sirenoides")`, a rate acceleration occurred coincident with a `r text.ratio.node.ancnode.atLabel("Hydrodamalis.gigas")` in body mass in Stellar's sea cow.
+-->
+
+Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Orycteropus.afer")` in body mass in the stem-lineage of _Pseudoungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Orycteropus.afer")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Procavia.capensis")` in body mass in the stem-lineage of _Paenungulata_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Procavia.capensis")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Trichechus.manatus")` in body mass in the stem-lineage of _Tehthytheria_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Trichechus.manatus")`), and a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Mammut.americanum")` in body mass in the stem-lineage of _Proboscidea_ (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Mammut.americanum")`) (Figure \@ref(fig:Figure-BodySize), Table \@ref(tab:Table-BodySize)).
+The ancestral _Hyracoidea_ was inferred to be relatively small (`r text.bodysize.ancNode.byMRCA.CI("Microhyrax.lavocati", "Procavia.capensis")`), and rate accelerations were coincident with independent body mass increases in large hyraxes such as _Titanohyrax andrewsi_ (`r text.bodysize.byLabel("Titanohyrax.andrewsi")`, `r text.ratio.nodeByMRCA.ancNode(nodeA="Titanohyrax.andrewsi","Orycteropus.afer")`).
+While the body mass of the ancestral _Sirenian_ was inferred to be large (`r text.bodysize.ancNode.byMRCA.CI("Trichechus.manatus", "Prorastomus.sirenoides")`), a rate acceleration occurred coincident with a `r text.ratio.node.ancnode.atLabel("Hydrodamalis.gigas")` in body mass in Stellar's sea cow.
 Rate accelerations also occurred coincident with `r text.ratio.nodeByMRCA.ancNode("Elephas.antiquus falconeri", "Elephas.cypriotes")` in body mass in the stem-lineage of the dwarf elephants _Elephas (Palaeoloxodon) antiquus falconeri_ and _Elephas cypriotes_.
-These data suggest that gigantism in _Afrotherians_ evolved step-wise, from small to medium bodies in the _Pseudoungulata_ stem-lineage, medium to large bodies in the _Tehthytherian_ stem-lineage and extinct hyraxes, and from large to exceptionally large bodies independently in the _Proboscidean_ stem-lineage and Stellar's sea cow (Figure 1).
+These data suggest that gigantism in _Afrotherians_ evolved step-wise, from small to medium bodies in the _Pseudoungulata_ stem-lineage, medium to large bodies in the _Tehthytherian_ stem-lineage and extinct hyraxes, and from large to exceptionally large bodies independently in the _Proboscidean_ stem-lineage and Stellar's sea cow (Figure \@ref(fig:Figure-BodySize), Table \@ref(tab:Table-BodySize)).
 
@@
 ```
-```{r Fig2 Prep Scales}
+```{r Fig3 Prep Scales}
 scale.log2CancerSucceptabilityChange <- table.cancerSucceptibility.Atlantogenata %>% 
@@
 
-```{r Fig2 Prep Table, eval=F}
-if(knitr::is_latex_output()){
-# size.lifespan.Atlantogenata.lm
-  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
-    size.lifespan.Atlantogenata.lm,
-    type="latex",
-    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
-    style = "all"
-  )) %>% paste0(collapse="\n")
-} else if(knitr::is_html_output()){
-  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
-    size.lifespan.Atlantogenata.lm,
-    type="html",
-    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
-    style = "all"
-  )) %>% paste0(collapse="\n")
-} else{
-  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
-    size.lifespan.Atlantogenata.lm,
-    type="text",
-    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
-    style = "all"
-  )) %>% paste0(collapse="\n")
-  t.size.lifespan.Atlantogenata.lm %>% cat
-}
-```
-
-```{r Fig2 prep}
+```{r Fig3 prep}
 # f5a1 <- tree.cancerSucceptibility.Atlantogenata  %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(CancerSucceptabilityChange = as.numeric(CancerSucceptabilityChange)) %>% as.treedata %>% 
@@
 
-f2a <- tree.cancerSucceptibility.Atlantogenata %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% as.treedata %>% 
+p.tree.cancerSucceptibility.Atlantogenata <- tree.cancerSucceptibility.Atlantogenata %>% as_tibble %>% filter(!is.na(node)) %>%  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% as.treedata %>%
   ggtree(
@@
   xlab("| Log2 Cancer Succeptibility Change |")
-# f5a
-
-f2 <- ggarrange(
-  f2a, 
-  #nrow=2, ncol=1,
-  labels="AUTO"
-)
-```
-
-```{r Fig2 Prep Table 1}
-t.cancersucc.atl <- table.cancerSucceptibility.Atlantogenata.genomes %>% 
-  # filter(label %in% ml.treedata.table$label) %>%  
-  # select(-parent, -node, -branch.length) %>%
-  select(
-    Node = label, 
-    lnSize,
-    `Est. Lifespan`=Lifespan,
-    `K1` = K1,
-    `K2` = K2,
-    `Change in K` = CancerSucceptabilityChange,
-    `log2 Change` = log2CancerSucceptabilityChange
-  ) %>%
-  mutate_at(
-    vars(`K1`, `K2`, `Change in K`), 
-    . %>% formatC(., format = "e", digits = 2)
-            ) %>%
-  mutate(
-    `Est. Lifespan` = `Est. Lifespan` %>% round(2), 
-    # `Change in K` = `Change in K` %>% round(2), 
-    `log2 Change` = `log2 Change` %>% round(2)
-  ) %>% 
-  mutate(
-    Ancestor = Node %>% 
-    sapply(
-        .,
-        function(n, d=table.cancerSucceptibility.Atlantogenata.genomes){
-            p.node <- d %>% filter(label==n) %>% pull(parent) %>% unique
-            d %>% filter(node == p.node) %>% pull(label) %>% unique %>% return
-        }
-    ) %>% str_replace_all("\\.", " ") %>% str_replace("Root", "Atlantogenata") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora")),
-        Node = Node %>% str_replace_all("\\.", " ") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora"))
-  ) %>%
-    arrange(Node) %>% 
-  select(Node, `Est. Lifespan`, K1, K2, `Change in K`, `log2 Change`) %>% 
-  kable(
-    ., 
-    booktabs = TRUE,
-    caption = 'Estimated Cancer Succeptibilty for nodes in Atlantogenata'
-  ) %>% 
-  kable_styling(latex_options = c("scale_down"))
-```
-```{r Figure2, fig.cap="Cancer succecptibility across Atlantogenata", out.width="6in", fig.asp=1}
-f2
-```
-
-```{r Table 2}
-t.cancersucc.atl
-```
-
-The dramatic increase in body mass and lifespan in some Afrotherian lineages implies those lineages evolved reduced cancer risk.
+
+```
+
+<!-- The dramatic increase in body mass and lifespan in some Afrotherian lineages implies those lineages evolved reduced cancer risk.
 To infer the magnitude of these reductions we estimated differences in cancer risk between small bodied, short-lived species and large bodied, long-lived species as well as for reconstructed ancestral Afrotherians.
@@
 Thus we can estimate the intrinsic cancer risk for a species as $K \approx Dt^6$.
-
-In order to estimate the intrinsic cancer risk of a species, we first obtained estimates for lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon$ (Table 2).
+-->
+
+<!-- In order to estimate the intrinsic cancer risk of a species, we first obtained estimates for lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon$ (Table 2).
 With this information in hand, we calculated $K_{1}$ at all nodes, and then estimated the fold change in cancer susceptibility between an ancestral node and a given node as $\frac{K_{2}}{K_{1}}$ (__Table 4__).
 As body size and phylogeny are the strongest predictors of lifespan, we find that this regression is sufficiently robust without further metabolic or other covariates for our purposes herein.
-
-As shown in __Table 2__, cancer susceptibility skyrocketed at the initial divergence of Atlantogenata, followed by a generally upwards trend.
+-->
+
+<!-- As shown in __Table 2__, cancer susceptibility skyrocketed at the initial divergence of Atlantogenata, followed by a generally upwards trend.
 At the common ancestor of Afrotheria there is an initial `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrotheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer risk.
@@
 However, cancer risk slowly deflates as size decreases as one moves along the tree towards extant species, such as in Hoffman's Two Toed Sloth (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Choloepus hoffmanni") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change) and in the Nine-banded Armadillo (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Dasypus novemcinctus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold change).
-
-Within Afrotheria, cancer susceptibility drops in Afrosoricida as species shrink (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrosoricida") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, then stagnates for the Cape Golden Mole) - but then rises `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Echinops telfairi") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold towards the Lesser Hedgehog Tenrec.
+-->
+
+<!-- Within Afrotheria, cancer susceptibility drops in Afrosoricida as species shrink (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrosoricida") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold, then stagnates for the Cape Golden Mole) - but then rises `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Echinops telfairi") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold towards the Lesser Hedgehog Tenrec.
 In parallel, Afroinsectivora does not increase in cancer susceptibility, and decreases once more at the Cape Elephant Shrew (`r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantulus edwardii") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold).
@@
 Neither the common ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further changes in cancer susceptibility.
+-->
+
+As expected, intrinsic cancer susceptibility in Afrotheria also varies with changes in body size and longevity (Figure \@ref(fig:Figure-CancerSuscept), Table \@ref(tab:Table-CancerSuscept)), with an initial `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Afrotheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold and `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in the stem-lineage of Afrotheria and Xenarthra, respectively, followed by a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Pseudoungulata") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increases in Pseudoungulata and a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Orycteropus afer") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in Aardvarks (Figure \@ref(fig:Figure-CancerSuscept), Table \@ref(tab:Table-CancerSuscept)).
+In contrast to the Paenungulate stem-lineage, there is a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer risk in Tethytheria, a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Trichechus manatus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in Manatee, and dramatic increases within Proboscidea including a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in Elephantidae and a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Mammut americanum") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold in the American Mastodon.
+Within the Elephantidae, Elephantina and Loxodontini have a `r table.cancerSucceptibility.Atlantogenata.genomes %>% filter(label == "Elephantina") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric`-fold increase in cancer susceptibility, while susceptibility is relatively stable in Mammoths.
+The three extant Proboscideans, Asian Elephant, African Savana Elephant, and the African Forest Elephant, meanwhile, have similar decreases in both size and cancer susceptibility (Figure \@ref(fig:Figure-CancerSuscept), Table \@ref(tab:Table-CancerSuscept)).
 
@@
 ```
-```{r Figure3 Prep Genome Table}
-ggt.GenomeTable.withNote <- genome.table.forFig %>% 
-  select(Species, `Common Name`, starts_with("Size"), starts_with("#"), starts_with("%"), starts_with("Mean")) %>% 
-  arrange(Species)%>% 
-  kable(
-    ., 
-    booktabs = TRUE,
-    caption = 'Summary of duplications in Atlantogenata'
-  ) %>% 
-  kable_styling(latex_options = c("scale_down"))
-```
-```{r Table2, fig.height=8, fig.width=6}
-ggt.GenomeTable.withNote
-```
-
-```{r Figure3, fig.cap="Gene duplications occur readily throughout Atlantogenata.
-A) A table summarizing duplication events in Atlantogenata.
-B) A tree of the species in Atlantogenata with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.", out.height="7in", out.width="6in"}
-
-f3 <- p.ml.tree.increasing.geneCounts.withSize
-# f3 <- ggarrange(f3a, f3b, nrow=2, ncol=1, labels = "AUTO", heights = c(1,2))
-# f3
-f3
-```
-
-Enhanced cancer suppression may have evolved through many mechanisms; among the most parsimonious is an increase in the copy number of genes with tumor suppressor functions.
+
+<!-- Enhanced cancer suppression may have evolved through many mechanisms; among the most parsimonious is an increase in the copy number of genes with tumor suppressor functions.
 Previous studies focusing on candidate gene studies, for example, have identified increased copy number of the tumor suppressors _TP53_ and _LIF_ in elephants [@Abegglen:JAMA2015; @Caulin2015; @Doherty2016; @Sulak2015; @Vazquez2018].
@@
 By only including nucleotides from the query sequence that were observed in the target genome, we also correct for partial hits where some or all of the homologs of a gene have diverged from the human homolog.
-
-As our sequence database included various protein transcripts for each gene, in order to obtain gene-level copy number information and eliminate any many-to-one mappings of hits, we labeled each exon of every reciprocal best hit (RBH) with the gene corresponding to the query transcript and merged all overlapping exons; next, we eliminated any many-to-one exons that resulted from the previous step.
+-->
+
+<!-- As our sequence database included various protein transcripts for each gene, in order to obtain gene-level copy number information and eliminate any many-to-one mappings of hits, we labeled each exon of every reciprocal best hit (RBH) with the gene corresponding to the query transcript and merged all overlapping exons; next, we eliminated any many-to-one exons that resulted from the previous step.
 Finally, we reassembled the gene loci based on the original transcript starts and ends, and the collapsed exon data, obtaining the full sequence of each RBH locus.
 Genes were considered to be duplicated if its copy number via RBHB was greater than or equal to 2, and the maximum ECNC among all transcripts prior to filtering was greater than or equal to 1.50 in order to account for partial hits that may be due to incomplete coverage.
-
-Once we applied these criteria to the results, we obtained the results shown in Figure 2A.
-Our approach positively identified an average of `r genomeStatsRBHB$pc.found %>% mean %>% as.percentage` of the queries we searched.
-We observed that the percentage of duplicated genes in non-Pseudoungulatan genomes was significantly higher: while these genomes had duplicates percentages ranging anywhere from `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`, outgroup species' duplication rates ranged from `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`.
-
-```{r Figure4, fig.cap="Correlations between genome quality metrics and ECNC metrics", fig.asp=1, out.width="5in", out.height="5in", fig.align="center"}
-f4 <- p.cor.GCT.CEGMA
-f4
-```
-
-Given the preliminary status of the genomes for _Chrysochloris asiatica_ and _Elephantulus edwardii_; and the fact that more revisions have been performed for Paenungulatan genomes; we hypothesized that genome quality may be confounding the results we've observed.
+-->
+
+<!-- We found that gene duplications were common in Atlantogenatan genomes (__Figure 4__), with number of duplications per lineage consistent with previously estimated gene duplication rates (REFS), but were particularly numerous in _Chrysochloris asiatica_,  _Elephantulus edwardii_.
+To explore whether genome quality may adversely effect our inferences, we used CEGMA and the gVolante server [@CEGMA; @gVolante] to assess the correlation between genome quality and copy number estimates.
+As shown in Supplementary Figure X, mean Copy Number, mean ECNC, and mean CN (the lesser of Copy Number and ECNC per gene) moderately or strongly correlate with genomic quality, such as LD50, the number of scaffolds, and contigs with a length above either 100K or 1M.
+-->
+
+We found that gene duplications were common in Atlantogenatan genomes (Figure \@ref(fig:Figure-GeneDup-Cladogram)), identifying an average of `r genomeStatsRBHB$pc.found %>% mean %>% as.percentage` genes, with an average of `r genomeStatsRBHB$pc.Hits.Duplicated %>% mean %>% as.percentage()` duplicated; this is in-line with other studies describing the rates of gene duplications over time.
+[@Lynch2000].
+We observed that the percentage of duplicated genes in non-_Pseudoungulatan_ genomes was significantly higher: while _Pseudoungulatan_ genomes had duplicates percentages ranging anywhere from `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(!Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`, outgroup species' duplication rates ranged from `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii"))%>% pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% min %>% as.percentage` to `r genome.table.forFig %>% filter(Species %in% c("Choloepus hoffmanni", "Chrysochloris asiatica", "Dasypus novemcinctus", "Echinops telfairi", "Elephantulus edwardii")) %>%  pull("% Hits Duplicated") %>% str_remove("%") %>% as.numeric %>% divide_by(100) %>% max %>% as.percentage`.
+To explore whether genome quality may adversely effect our inferences, we used CEGMA and the gVolante server [@CEGMA; @gVolante] to assess the correlation between genome quality and copy number estimates.
+As shown in Figure \@ref(fig:Figure-GenomeCor), mean Copy Number, mean ECNC, and mean CN (the lesser of Copy Number and ECNC per gene) moderately or strongly correlate with genomic quality, such as LD50, the number of scaffolds, and contigs with a length above either 100K or 1M.
+
+<!-- Given the preliminary status of the genomes for _Chrysochloris asiatica_ and _Elephantulus edwardii_; and the fact that more revisions have been performed for Paenungulatan genomes; we hypothesized that genome quality may be confounding the results we've observed.
 We used `CEGMA` and the `gVolante` server [@CEGMA; @gVolante] to assess the quality of all our genomes, to see if any of these metrics correlated with our metrics of gene counts and copy number.
 As shown in Figure 3, mean ECNC, mean Copy Number, and mean CN (the lesser of Copy Number and ECNC per gene) correlate moderately strongly with genomic quality metrics related to the length and assembly quality, like LD50 and the number of scaffolds and contigs with a length above either 100K or 1M, supporting our hypothesis that increases in copy number correlate .
-
-In order to determine when in the phylogeny our identified duplications occurred, we used maximum likelihood to reconstruct ancestral gene copy numbers across the phylogeny.
+-->
+
+<!-- In order to determine when in the phylogeny our identified duplications occurred, we used maximum likelihood to reconstruct ancestral gene copy numbers across the phylogeny.
 We encoded the copy number of each gene as a discrete trait ranging from 0 ("?") to 31+, and then ran the ancestral gene copy number reconstruction using the phylogenetics suite `IQTREE`, which rapidly tests and runs the best evolutionary model for a given trait of interest.
@@
 Setting the threshold to 0.8, we obtained the results shown in Figure 2B.
-
-Gene duplications events occured readily thoughout _Atlantogenata_ (Figure 2B).
+-->
+
 Among the genes that increased in copy number in the elephant lineage are TP53 and LIF, as previously described; however, we find that these two genes represent a fraction of the 940 genes that are duplicated in the African Elephant overall, which accumulated over various steps through their evolution.
@@
 ```
-
 ```{r Branches which experienced an increase in size}
@@
 
-p.ORA <- ORA.numbers.ReactomeOnly %>% 
-  filter(Direction == "increased") %>% 
-  select(-Direction) %>% 
-  mutate(Node = Node %>% str_replace_all("_", " ")) %>% 
-  kable(
-    ., 
-    booktabs = TRUE,
-    caption = 'Number of pathways overrepresented among duplicated genes at different FDRs.'
-  ) %>% 
-  kable_styling(latex_options = c("scale_down"))  # TODO: Make "Number of Pathways" a group, and make each column just "FDR<=XX"
-
-```
-
-```{r Table3}
-p.ORA
-```
-
-```{r Figure5, fig.cap="Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.", out.width="6in", out.height="4in"}
-f5a <- p.loxAfr.ORA.dotplot
-f5b <- p.loxAfr.ORA.dotplot.increased
-
-f5 <- ggarrange(f5a, f5b, nrow=1, ncol=2, common.legend = T, legend = "bottom", labels="AUTO")
-
-f5
 ```
@@
 For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways_cancer databases using WebGestalt ORA.
-Going through the tree, at no FDR<0.5 is there any significant pathway representation for genes that increased in the branches leading to `r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`; furthermore, there are no significant pathway enrichments at FDR<0.5 for genes whose copy number did not change between branches (copy-number-stable) for `r noEnrichment.any.Reactome[1:length(noEnrichment.any.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichment.any.Reactome[length(noEnrichment.any.Reactome)]`.
+Going through the tree, at no FDR $\leq$ 0.5 is there any significant pathway representation for genes that increased in the branches leading to <!--`r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`--> _Afrosoricida_, _Pseudoungulata_, _Elephantina_, _Loxodontini_, _Elephas maximus_, _Mammuthus_, _Loxodonta cyclotis_, _Palaeoloxodon antiquus_, _Loxodona_, _Mammuthus columbi_, _Mammuthus primigenius_, _Procavia capensis_, _Elephantidae_, or _Proboscidea_; furthermore, there are no significant pathway enrichments at FDR<0.5 for genes whose copy number did not change between branches (copy-number-stable) for _Afrotheria__.
 Note that because _Xenarthra_ was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.
 
-For the other branches, the number of pathways that came up as significantly enriched at each FDR is shown in Figure 3A (Supplementary Figure 3).
+For the other branches, the number of pathways that came up as significantly enriched at each FDR is shown in Table \@ref(tab:Table-OraFreqDatabase).
 For the species with high duplication rates and lower-quality, highly-fragmented genomes, such as with _Chrysochloris asiatica_ (`r genomeStatsRBHB %>% filter(Genome == "chrAsi1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits) and _Elephantulus edwardii_ (`r genomeStatsRBHB %>% filter(Genome == "eleEdw1m") %>% pull(pc.Hits.Duplicated) %>% as.percentage` duplicated hits), it is unsurprising that there is a proportionally large number of pathway enrichments.
@@
 In the cell cycle pathways called as significant in the genomes of these two species plus _Orycteropus afer_ and _Echinops telfairi_, the duplicated genes included in these sets are from the same gene families, such as the APC subunit family; the proteosome subunit families; and the protein phosphatase 2 family, among others.
-It is highly possible that these results reflect true expansions of these gene families, especially in the higher-quality _Orycteropus afer_ genome; however, it is also possible that it simply reflects artificial duplication events, and so require further study.
-
-The pathway enrichments for genes whose copy number either did not change, or whose copy number increased, between _Loxodona_ and the African Elephant are shown in Figure 3B and 3C, respectively.
+It is highly possible that these results reflect true expansions of these gene families, especially in the higher-quality _Orycteropus afer_ genome; however, it is also possible that it simply reflects artifactual duplications, and so require further study.
+
+The pathway enrichments for genes whose copy number either did not change, or whose copy number increased, between _Loxodona_ and the African Elephant are shown in Figure \@ref(fig:Figure-loxAfrORA).
 Among the few enriched pathways in the African Elephant, we see that that two tumor suppression pathways - APC Complex-related pathways, and "TP53 Regulates Metabolic Genes"- appear not only in the case of stable genes, but also in the set of newly duplicated genes.
@@
 
-`r knitr::knit_exit()`
-
 ## Concerted duplication of TP53 and TP53-related genes towards _Probodiscea_
 
-```{r Fig6 prep}
+```{r Figure7 prep}
 # loxAfr4.geneTPM.genesToHighlight.plus
@@
   filter(Gene %in% genesToHighlight.plus.TP53) %>% 
+  group_by(Database) %>% 
   mutate(
@@
     ) %>% 
-  group_by(Locus) %>% 
+    ungroup %>% 
+  group_by(Locus, Database) %>% 
   summarize(Gene = unique(Gene), id=unique(id), TPM=max(TPM), Reactome.Pathway.Prime=unique(Reactome.Pathway.Prime)) %>% 
@@
   )
+           )%>% 
+  ungroup %>% 
+  filter(Reactome.Pathway.Prime!="", Database == "Reactome")
+
+membership.matrix <- loxAfr4.geneTPM.ORA.expressedDups.TP53 %>% select(Gene, Reactome.Pathway.Prime)%>% distinct %>% pull(Reactome.Pathway.Prime) %>% str_split(", ") %>% set_names(., loxAfr4.geneTPM.ORA.expressedDups.TP53$Gene %>% unique)%>% binTableFromList() %>% data.table::transpose(make.names = "Gene", keep.names = "Gene")
+
+loxAfr4.geneTPM.ORA.expressedDups.TP53 %<>% full_join(
+  .,
+  membership.matrix %>% select(Gene, membership.matrix %>% summarise_at(vars(-Gene), sum) %>% select_if(.>1) %>% names), 
+  by="Gene"
   )
@@
     ),
-    size=2
-  ) +  
-  geom_point2(aes(x=branch, subset=(GTF2F1>0), color="GTF2F1"), size=3, position = position_nudge(x=-0.2, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(STK11>0), color="STK11"), size=3, position = position_nudge(x=0, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(COX5A>0), color="COX5A"), size=3, position = position_nudge(x=0.2, y=0.2)) +
-  geom_point2(aes(x=branch, subset=(BRD7>0), color="BRD7"), size=3, position = position_nudge(x=-0.2, y=0)) +
-  geom_point2(aes(x=branch, subset=(TP53>0), color="TP53"), size=3, position = position_nudge(x=0, y=0)) +
-  geom_point2(aes(x=branch, subset=(CASP9>0), color="CASP9"), size=3, position = position_nudge(x=0.2, y=0)) +
-  geom_point2(aes(x=branch, subset=(PRDX1>0), color="PRDX1"), size=3, position = position_nudge(x=-0.2, y=-0.2)) +
-  geom_point2(aes(x=branch, subset=(LAMTOR5>0), color="LAMTOR5"), size=3, position = position_nudge(x=0, y=-0.2)) +
-  geom_point2(aes(x=branch, subset=(COX20>0), color="COX20"), size=3, position = position_nudge(x=0.2, y=-0.2)) +
-  geom_tiplab(parse=T, offset = 0.25) +
+    size=1
+  ) +  
+  geom_point2(aes(x=branch, subset=(GTF2F1>0), color="GTF2F1"), size=1.5, position = position_nudge(x=-0.2, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(STK11>0), color="STK11"), size=1.5, position = position_nudge(x=0, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(COX5A>0), color="COX5A"), size=1.5, position = position_nudge(x=0.2, y=0.2)) +
+  geom_point2(aes(x=branch, subset=(BRD7>0), color="BRD7"), size=1.5, position = position_nudge(x=-0.2, y=0)) +
+  geom_point2(aes(x=branch, subset=(TP53>0), color="TP53"), size=1.5, position = position_nudge(x=0, y=0)) +
+  # geom_point2(aes(x=branch, subset=(CASP9>0), color="CASP9"), size=1.5, position = position_nudge(x=0.2, y=0)) +
+  geom_point2(aes(x=branch, subset=(PRDX1>0), color="PRDX1"), size=1.5, position = position_nudge(x=-0.2, y=-0.2)) +
+  geom_point2(aes(x=branch, subset=(LAMTOR5>0), color="LAMTOR5"), size=1.5, position = position_nudge(x=0, y=-0.2)) +
+  geom_point2(aes(x=branch, subset=(COX20>0), color="COX20"), size=1.5, position = position_nudge(x=0.2, y=-0.2)) +
+  geom_tiplab(parse=T, offset = 0.25, size=2) +
   scale_color_manual(values = TP53.gene.color) + 
   guides(color=guide_legend(title="Gene", title.position = "top", title.theme = element_text(size = 14, face="bold")))+
-  xlim(0,20) +
+  xlim(0,15) +
   theme_tree()+
   theme(
-    strip.text = element_text(size=14, face="bold"), 
-    legend.position = "bottom",
-    legend.title = element_text(size = 14, face="bold"), 
-    legend.text = element_text(size = 12),
-    axis.text.y = element_blank()
+    legend.position="bottom",
+    # panel.background = element_rect(fill = "lightgrey"),
+    text = element_text(face = "plain", 
+                        colour = "black", lineheight = 0.9, 
+                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                        debug = FALSE), 
+    plot.title = element_text(size = 6, 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.title = element_text(size = 8, 
+                                face = "bold", colour = "black"), 
+    legend.text = element_text(size = 6, 
+                               face = "plain", colour = "black"),
+    strip.text.y = element_text(size=8, face="bold")
   )
@@
   geom_point(
-    size=3
+    size=1.5
   )+
@@
     strip.background = element_rect(fill="white"),
-    strip.text.y = element_text(face = "bold", size=14),
-    legend.title = element_text(size = 14, face="bold"), 
-    legend.text = element_text(size = 12),
-    axis.title = element_text(size=14, face="bold"),
-    axis.text.x = element_text(size=12, face="bold"),
-    axis.text.y = element_text(size=12, face = "bold")
+    strip.text.y = element_text(face = "bold", size=8),
+    legend.title = element_text(size = 8, face="bold"), 
+    legend.text = element_text(size = 6),
+    axis.title = element_text(size=8, face="bold"),
+    axis.text.x = element_text(size=6, face="bold"),
+    axis.text.y = element_text(size=6, face = "bold")
   )+
@@
 
-```{r Figure6, fig.cap="", fig.width=5, fig.height=8}
-f6a <- p.treedata.tp53.table
-f6b <- p.tp53.genes.expression
-fig6 <- ggarrange(f6a, f6b, nrow=1, ncol=2, labels="AUTO", widths = c(1.5,1))
-
-fig6
+Prior studies looking at the duplication of TP53 in the African Elephant motivated further study for the enrichment of genes involved in TP53-related metabolic pathway.
+We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, and used publicly-available RNA-seq data to see which genes are actively expresses in living elephants.
+Excitingly, we found that the initial duplication of TP53 in Tethytheria, where body size expanded, was preceded by the duplication of GTF2F1 and STK11 in _Paenungulata_; and was coincided by the duplication of BRD7.
+These two genes are involved in regulating the transcription of TP53, and their duplication prior to that of TP53 may have facilitated its retroduplication.
+Interestingly enough, STK11 is a tumor suppressor gene in its own right, and plays additional roles in mediating tumor suppression via p21-induced senescence [@Launonen2005].
+<!-- [TALK ABOUT HOW THESE TWO MAY HAVE FACILITATED THE DUPLICATION OF TP53] -->
+<!-- GTF2F1 is a general transcription factor [@CITE and talk more about this] -->
+<!-- STK11 is a tumor suppressor gene [@CITE and talk more] -->
+
+The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either alongside TP53 in the case of SIAH1, or subsequently in _Probodiscea_, _Elephantidae_, and in modern elephants.
+(Figure \@ref(fig:Figure-loxAfrTP53)).
+These genes are all expressed in RNA-seq data, suggesting that they encode functional genes in modern elephants (Figure \@ref(fig:Figure-loxAfrTP53)).
+
+# Discussion
+
+With our results, we have demonstrated that gigantism in _Atlantogenata_ was not limited to extant elephants, but rather occurred at various points in the evolution of the clade; however, the hundred-fold to hundred-million-fold increases in cancer risk that is associated with these increases however poses an innate challenge in the evolution and persistence of this trait.
+The environmental selective pressures on body size have long been the fascination of evolutionary biologists, and the influence of climate, predation, geography, and ecological niche on body size have been well established.
+[@CooperAndPurvis2010].
+Indeed, there is a general trend, known as Cope's Rule, for body sizes of species to increase over time [@HoneAndBenton2004; @HuntAndRoy2006; @Jablonski1997].
+However, for all the research on body size that has been done thus far, the mechanisms that enable a release from the negative pressure on body size exerted on cancer risk has proven more elusive [@HoneAndBenton2004; @HuntAndRoy2006; @Jablonski1997].
+
+Furthermore, we show that tumor suppressor duplications are enriched not only in large, extant species, but also in large common ancestors, and that these duplications evolved throughout the tree, rather than in concert.
+The ancestral body sizes of many of the subclades in _Atlantogenata_ were estimated to be large, and the estimated cancer risk increases - even for the small clades like _Afroinsectivora_ may explain why these TSG gene duplications occurred early on, and persisted even in species that we did not expect to have high risks of cancer.
+However, some of our results also provide interesting evidence for a paradigm of TSG duplications being a pervasive phenomenon, rather than a specific mechanism that occurs after the evolution of gigantism.
+At the common ancestor of _Proboscidea_, which was small relative to both its ancestors and its descendants, we see the emergence of various TSG duplications, which may have enabled the stratospheric increase in body size of modern elephants.
+We also identified many TSG duplication events in smaller species and lineages such as in _Chrysochloris asiatica_ and _Elephantulus edwardii_, although these may have been the result of low-quality genomes.
+
+The impacts and takeaways of this study are limited by the quality and quantity of _Atlantogenatan_ genomes that were available, and our available knowledge of the lifespans and cancer risk of the extant species.
+For many of our species, studies in captivity are limited, and the species are relatively understudied from a longitudinal perspective, such as with the Cape Elephant Shrew and Cape Golden Mole.
+Furthermore, while there is recent interest in resequencing and improving the quality of these assemblies, at the time of this writing there is still quite a ways to go in order to have genomes of a sufficiently rigorous quality to make stronger inferences about gene copy number expansions and contractions (which were not considered in this study for this reason).
+
+The lack of a stronger signal from tumor suppressor duplications is likely a result of the strong effect size on both cancer risk and organismal toxicity that a TSG duplication would provide.
+The duplication of a tumor suppressor in many cases is associated with mild toxicity, although it greatly varies given the context and TSG in question; however, a single TSG duplication can also provide significant protection against cancer.
+For example, the overexpression of TP53 in mice, while protective of cancer, is associated with progeria and early death; however, if an additional copy of TP53 is introduced with its regulatory elements intact, the mice are healthy and experience normaging, while also demonstrating an enhanced response to cellular stress and lower rates of cancer.
+[@Tyner:20021c2; @GarciaCao:20021c2].
+In light of this, it is fascinating that our results in the elephant lineage suggest that the duplication of TP53 regulators preceded the retroduplication and expansion of TP53, as this likely would have lowered the toxicity of the initial duplication and thus enabled it to occur.
+Given a sufficient selective pressure on increasing body size, it stands to reason that events like this could alleviate the negative pleiotropy of TSG duplications sufficiently to enable their persistence and allow for subsequent refinement over evolutionary time.
+
+By combining a phylogenetic study on body size in addition to a survey of copy number across nearly all protein-coding genes, we provide a comprehensive look at the question of the role of cancer risk and body size in _Atlantogenata_ that may provide broader insight to other mammalian species.
+Our study was initially motivated by the identification of functional duplicates of tumor suppressors, such as TP53 and LIF in elephants [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018].
+Further in support of our results, a larger candidate gene study by Caulin et al.
+[@Caulin2015] characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans.
+However, while candidate gene studies are useful, by their very design they are biased in determining larger patterns of evolution of traits.
+Without addressing these questions with a genome-wide approach, any and all insights will be inevitably limited to a fraction of the whole story.
+
+Our results suggest that the pervasive duplication of tumor suppressors may help enable the evolution of larger body sizes by lowering the cancer risk of species, either prior to or in lockstep with increasing body size.
+however, this is unlikely to be the only genetic mechanism at play in this scenario.
+In genome-wide studies of unusually large or long-lived species such as the bowhead whale [@BowheadGenome], Myotid bats [@seim2013; @myoDav1_pteAle1], naked mole rat [@NakedMoleRatGenome], and blind mole rat [@BlindMoleRatGenome], there were cases of overrepresentation of TSGs among duplicate genes that were outshadowed by the identification of strong signatures of positive selection at TSGs.
+While the evolution of regulatory and coding elements of both TSGs and other non-canonical tumor suppressor genes have been shown to be important for mediating the cancer risk of long-lived species, there has been no attention given to the possibility of TSG duplications providing a relaxation of possible negative pleitropy that could result from these traits.
+It has been well-established in the literature that genes duplication events allow for evolutionary drift in one of the copies, which may result in neofunctionalization or specialization of the two copies [@Rastogi2005; @QianAndZhang2014; @Stoltzfus1999].
+While this is beyond the scope of our study, a promising future direction of this work would include an evolutionary analysis of duplicated genes relative to each other to see if this has already occurred between the pairs of genes we have identified.
+
+# Acknowledgements
+I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for his indispensable support.
+
+# Conflicts of Interest
+The Authors have no conflicts of interest to report
+
+# Funding Source
+We would like to thank the Department of Human Genetics at the University of Chicago for supporting this project.
+
+# FIGURES
+
+```{r Figure-intro, fig.cap="Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]",  out.width="6in", fig.asp=1.3}
+atlantogenata.tree.size.lifespan
+```
+
+```{r Figure-BodySize, fig.cap="Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
+**A)** Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch.
+Atlantogenata is highlighted at the bottom.
+**B)** Zoom-in of (**A**) on Atlantogenata.
+Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size.
+**C)** Confidence interval plot for representative species and ancestral nodes.", out.width="6in", fig.asp=1}
+fig.bodysize.a <- p.tree
+fig.bodysize.b <- p.tree.atlantogenata
+fig.bodysize.c <- p.bodysize.atlantogenata
+
+fig.bodysize.ab <- ggarrange(
+  fig.bodysize.a, fig.bodysize.b, 
+  nrow=1, ncol = 2, 
+  labels="AUTO",
+  widths = c(1,2),
+  font.label = list(size=6, face="bold")
+  )
+
+fig.bodysize <-
+  ggarrange(
+  ggplotify::as.ggplot(fig.bodysize.ab),
+  fig.bodysize.c,
+  nrow=2, ncol=1,
+  labels = c("", "C"),
+  font.label = list(size=6, face="bold")
+)
+fig.bodysize
+```
+
+```{r Figure-CancerSuscept, fig.cap="Cancer succecptibility across Atlantogenata.
+Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.", out.width="6in", fig.asp=1}
+p.tree.cancerSucceptibility.Atlantogenata
+```
+
+```{r Figure-GeneDup-Cladogram, fig.cap="Gene duplications occur readily throughout Atlantogenata.
+Shown here is a tree of Atlantogenatan species with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.", out.height="7in", out.width="6in"}
+p.ml.tree.increasing.geneCounts.withSize
+```
+
+```{r Figure-GenomeCor, fig.cap="Correlations between genome quality metrics and ECNC metrics.
+Gene copy number metrics, and the genome quality metrics most strongly associated with them, are highlighted in red.", fig.asp=1, out.width="6in", out.height="6in", fig.align="center"}
+p.cor.GCT.CEGMA
+```
+
+```{r Figure-loxAfrORA, fig.cap="Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.", out.width="6in", out.height="4in"}
+fig.loxAfr.ORA <- ggarrange(
+  p.loxAfr.ORA.dotplot, 
+  p.loxAfr.ORA.dotplot.increased, 
+  nrow=1, ncol=2, 
+  common.legend = T, legend = "bottom", 
+  labels="AUTO"
+  )
+
+fig.loxAfr.ORA
+```
+
+```{r Figure-loxAfrTP53, fig.cap="", out.width="6in", out.height="6in"}
+fig.TP53 <- ggarrange(
+  p.treedata.tp53.table, 
+  p.tp53.genes.expression, 
+  nrow=1, ncol=2, 
+  labels="AUTO", 
+  widths = c(1.5,1)
+  )
+fig.TP53
     # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
@@
 
-Prior studies looking at the duplication of TP53 in the African Elephant motivated further study for the enrichment of genes involved in TP53-related metabolic pathway.
-We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, and used publicly-available RNA-seq data to see which genes are actively expresses in living elephants.
-Excitingly, we found that the initial duplication of TP53 in Tethytheria, where body size expanded, was preceded by the duplication of GTF2F1 and STK11 in _Paenungulata_; and was coincided by the duplication of BRD7.
-These two genes are involved in regulating the transcription of TP53, and their duplication prior to that of TP53 may have facilitated its retroduplication.
-Interestingly enough, STK11 is a tumor suppressor gene in its own right, and plays additional roles in mediating tumor suppression via p21-induced senescence [@Launonen2005].
-<!-- [TALK ABOUT HOW THESE TWO MAY HAVE FACILITATED THE DUPLICATION OF TP53] -->
-<!-- GTF2F1 is a general transcription factor [@CITE and talk more about this] -->
-<!-- STK11 is a tumor suppressor gene [@CITE and talk more] -->
-
-The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either alongside TP53 in the case of SIAH1, or subsequently in _Probodiscea_, _Elephantidae_, and in modern elephants.
-(Figure 4A).
-These genes are all expressed in RNA-seq data, suggesting that they encode functional genes in modern elephants (Fig 4B).
-
-# Discussion
-
-With our results, we have demonstrated that gigantism in _Atlantogenata_ was not limited to extant elephants, but rather occurred at various points in the evolution of the clade; however, the hundred-fold to hundred-million-fold increases in cancer risk that is associated with these increases however poses an innate challenge in the evolution and persistence of this trait.
-The environmental selective pressures on body size have long been the fascination of evolutionary biologists, and the influence of climate, predation, geography, and ecological niche on body size have been well established.
-[@CooperAndPurvis2010].
-Indeed, there is a general trend, known as Cope's Rule, for body sizes of species to increase over time [@HoneAndBenton2004; @HuntAndRoy2006; @Jablonski1997].
-However, for all the research on body size that has been done thus far, the mechanisms that enable a release from the negative pressure on body size exerted on cancer risk has proven more elusive [@HoneAndBenton2004; @HuntAndRoy2006; @Jablonski1997].
-
-Furthermore, we show that tumor suppressor duplications are enriched not only in large, extant species, but also in large common ancestors, and that these duplications evolved throughout the tree, rather than in concert.
-The ancestral body sizes of many of the subclades in _Atlantogenata_ were estimated to be large, and the estimated cancer risk increases - even for the small clades like _Afroinsectivora_ may explain why these TSG gene duplications occurred early on, and persisted even in species that we did not expect to have high risks of cancer.
-However, some of our results also provide interesting evidence for a paradigm of TSG duplications being a pervasive phenomenon, rather than a specific mechanism that occurs after the evolution of gigantism.
-At the common ancestor of _Proboscidea_, which was small relative to both its ancestors and its descendants, we see the emergence of various TSG duplications, which may have enabled the stratospheric increase in body size of modern elephants.
-We also identified many TSG duplication events in smaller species and lineages such as in _Chrysochloris asiatica_ and _Elephantulus edwardii_, although these may have been the result of low-quality genomes.
-
-The impacts and takeaways of this study are limited by the quality and quantity of _Atlantogenatan_ genomes that were available, and our available knowledge of the lifespans and cancer risk of the extant species.
-For many of our species, studies in captivity are limited, and the species are relatively understudied from a longitudinal perspective, such as with the Cape Elephant Shrew and Cape Golden Mole.
-Furthermore, while there is recent interest in resequencing and improving the quality of these assemblies, at the time of this writing there is still quite a ways to go in order to have genomes of a sufficiently rigorous quality to make stronger inferences about gene copy number expansions and contractions (which were not considered in this study for this reason).
-
-The lack of a stronger signal from tumor suppressor duplications is likely a result of the strong effect size on both cancer risk and organismal toxicity that a TSG duplication would provide.
-The duplication of a tumor suppressor in many cases is associated with mild toxicity, although it greatly varies given the context and TSG in question; however, a single TSG duplication can also provide significant protection against cancer.
-For example, the overexpression of TP53 in mice, while protective of cancer, is associated with progeria and early death; however, if an additional copy of TP53 is introduced with its regulatory elements intact, the mice are healthy and experience normaging, while also demonstrating an enhanced response to cellular stress and lower rates of cancer.
-[@Tyner:20021c2; @GarciaCao:20021c2].
-In light of this, it is fascinating that our results in the elephant lineage suggest that the duplication of TP53 regulators preceded the retroduplication and expansion of TP53, as this likely would have lowered the toxicity of the initial duplication and thus enabled it to occur.
-Given a sufficient selective pressure on increasing body size, it stands to reason that events like this could alleviate the negative pleiotropy of TSG duplications sufficiently to enable their persistence and allow for subsequent refinement over evolutionary time.
-
-By combining a phylogenetic study on body size in addition to a survey of copy number across nearly all protein-coding genes, we provide a comprehensive look at the question of the role of cancer risk and body size in _Atlantogenata_ that may provide broader insight to other mammalian species.
-Our study was initially motivated by the identification of functional duplicates of tumor suppressors, such as TP53 and LIF in elephants [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018].
-Further in support of our results, a larger candidate gene study by Caulin et al.
-[@Caulin2015] characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans.
-However, while candidate gene studies are useful, by their very design they are biased in determining larger patterns of evolution of traits.
-Without addressing these questions with a genome-wide approach, any and all insights will be inevitably limited to a fraction of the whole story.
-
-Our results suggest that the pervasive duplication of tumor suppressors may help enable the evolution of larger body sizes by lowering the cancer risk of species, either prior to or in lockstep with increasing body size.
-however, this is unlikely to be the only genetic mechanism at play in this scenario.
-In genome-wide studies of unusually large or long-lived species such as the bowhead whale [@BowheadGenome], Myotid bats [@seim2013; @myoDav1_pteAle1], naked mole rat [@NakedMoleRatGenome], and blind mole rat [@BlindMoleRatGenome], there were cases of overrepresentation of TSGs among duplicate genes that were outshadowed by the identification of strong signatures of positive selection at TSGs.
-While the evolution of regulatory and coding elements of both TSGs and other non-canonical tumor suppressor genes have been shown to be important for mediating the cancer risk of long-lived species, there has been no attention given to the possibility of TSG duplications providing a relaxation of possible negative pleitropy that could result from these traits.
-It has been well-established in the literature that genes duplication events allow for evolutionary drift in one of the copies, which may result in neofunctionalization or specialization of the two copies [@Rastogi2005; @QianAndZhang2014; @Stoltzfus1999].
-While this is beyond the scope of our study, a promising future direction of this work would include an evolutionary analysis of duplicated genes relative to each other to see if this has already occurred between the pairs of genes we have identified.
-
-# Acknowledgements
-I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for his indispensable support.
-
-# Conflicts of Interest
-The Authors have no conflicts of interest to report
-
-# Funding Source
-We would like to thank the Department of Human Genetics at the University of Chicago for supporting this project.
+# TABLES
+
+```{r Table-CancerSuscept}
+t.cancersucc.atl <- table.cancerSucceptibility.Atlantogenata.genomes %>% 
+  # filter(label %in% ml.treedata.table$label) %>%  
+  # select(-parent, -node, -branch.length) %>%
+  select(
+    Node = label, 
+    lnSize,
+    `Est. Lifespan`=Lifespan,
+    `K1` = K1,
+    `K2` = K2,
+    `Change in K` = CancerSucceptabilityChange,
+    `log2 Change` = log2CancerSucceptabilityChange
+  ) %>%
+  mutate_at(
+    vars(`K1`, `K2`, `Change in K`), 
+    . %>% formatC(., format = "e", digits = 2)
+            ) %>%
+  mutate(
+    `Est. Lifespan` = `Est. Lifespan` %>% round(2), 
+    # `Change in K` = `Change in K` %>% round(2), 
+    `log2 Change` = `log2 Change` %>% round(2)
+  ) %>% 
+  mutate(
+    Ancestor = Node %>% 
+    sapply(
+        .,
+        function(n, d=table.cancerSucceptibility.Atlantogenata.genomes){
+            p.node <- d %>% filter(label==n) %>% pull(parent) %>% unique
+            d %>% filter(node == p.node) %>% pull(label) %>% unique %>% return
+        }
+    ) %>% str_replace_all("\\.", " ") %>% str_replace("Root", "Atlantogenata") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora")),
+        Node = Node %>% str_replace_all("\\.", " ") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora"))
+  ) %>%
+    arrange(Node) %>% 
+  select(Node, `Est. Lifespan`, K1, K2, `Change in K`, `log2 Change`) %>% 
+  kable(
+    ., 
+    booktabs = TRUE,
+    caption = 'Estimated Cancer Succeptibilty for nodes in Atlantogenata'
+  ) %>% 
+  kable_styling(latex_options = c("scale_down"))
+
+t.cancersucc.atl
+```
+
+```{r Table-GeneDup}
+ggt.GenomeTable.withNote <- genome.table.forFig %>% 
+  select(Species, `Common Name`, starts_with("Size"), starts_with("#"), starts_with("%"), starts_with("Mean")) %>% 
+  arrange(Species)%>% 
+  kable(
+    ., 
+    booktabs = TRUE,
+    caption = 'Summary of duplications in Atlantogenata'
+  ) %>% 
+  kable_styling(latex_options = c("scale_down"))
+ggt.GenomeTable.withNote
+```
+
+```{r Table-ORAFreqDatabase}
+p.ORA <- ORA.numbers.ReactomeOnly %>% 
+  filter(Direction == "increased") %>% 
+  select(-Direction) %>% 
+  mutate(Node = Node %>% str_replace_all("_", " ")) %>% 
+  kable(
+    ., 
+    booktabs = TRUE,
+    caption = 'Number of pathways overrepresented among duplicated genes at different FDRs.'
+  ) %>% 
+  kable_styling(latex_options = c("scale_down"))  # TODO: Make "Number of Pathways" a group, and make each column just "FDR<=XX"
+p.ORA
+```
+
+```{r Table-SRAData, eval=T}
+sra.data <- read_csv("../data/SraRunTable/SraRunTable.csv") %>% 
+  select(Organism, Run, source_name, sample_name) %>% 
+  filter(Organism %in% genomeTable.atlantogenata$Species) %>% 
+     mutate(sample_name = sample_name %>% str_remove("Nine-banded Armadillo_ ID 09-86_")) %>% 
+     mutate_all(. %>% str_replace_na("")) %>% 
+     unite("Tissue", source_name, sample_name, sep="") %>% 
+     filter(Tissue != "") %>% 
+     group_by(Organism) %>% 
+     summarize(Run = Run %>% unique %>% paste0(collapse=", "), 
+               Tissue = Tissue %>% str_to_title() %>% unique %>% paste0(collapse=", ")) %>% 
+     left_join(genomeTable.atlantogenata, by=c("Organism" = "Species")) %>% 
+  select(Organism, `Common Name`, Genome, `SRA Acc.`=Run, Tissues=Tissue) #%>% 
+  # mutate_all(. %>% str_wrap(width=50))
+sra.data %>% 
+  kable(
+    ., 
+    booktabs = TRUE,
+    caption = 'NCBI SRA datasets used in this study, along with key biological and genome information.'
+  ) %>% 
+  column_spec(4, width="50em") %>% 
+  column_spec(5, width="50em") %>% 
+  kable_styling(latex_options = c("scale_down"))
+```
+
+```{r Table-BodySize, eval=T}
+t.bodysize.atlantogenata.withNodes <- st.Atlantogenata %>% 
+  as_tibble() %>% 
+  filter(!is.na(label)) %>%
+  select(
+    `Node`=label, 
+    `Size (log(g))`=Median, 
+    `95% CI (Low)` = `95%CI_Low`, 
+    `95% CI (High)` = `95%CI_High`,  
+    `Rate (sqrt)`=sqrt.rate
+  ) %>%
+  mutate_at(vars(-`Node`, ), . %>% as.numeric %>% round(2)) %>% 
+  kable(
+    ., 
+    booktabs = TRUE, longtable = T,
+    caption = 'Body Size and Confidence Intervals in Atlantogenata estimated using StableTraits.'
+  ) %>% kable_styling(latex_options = "repeat_header")
+t.bodysize.atlantogenata.withNodes
+```
+
+```{r Table-PGLS-Lifespan, eval=T}
+if(knitr::is_latex_output()){
+# size.lifespan.Atlantogenata.lm
+  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
+    size.lifespan.Atlantogenata.lm,
+    type="latex",
+    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
+    style = "all"
+  )) %>% paste0(collapse="\n")
+} else if(knitr::is_html_output()){
+  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
+    size.lifespan.Atlantogenata.lm,
+    type="html",
+    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
+    style = "all"
+  )) %>% paste0(collapse="\n")
+} else{
+  t.size.lifespan.Atlantogenata.lm <- quiet(stargazer(
+    size.lifespan.Atlantogenata.lm,
+    type="text",
+    title = "Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression",
+    style = "all"
+  )) %>% paste0(collapse="\n")
+t.size.lifespan.Atlantogenata.lm %>% cat
+}
+```
+
+```{r Table-Genomes, eval=T}
+genomeTable.atlantogenata %>% 
+  mutate(Citation="NA") %>% 
+  select(Species, `Common Name`, Genomes = Genome, `Highest Quality Genome` = BestGenome, Citation) %>% 
+  kable(
+    ., 
+    booktabs = TRUE,
+    caption = 'Genomes used in this study.'
+  ) %>% 
+  kable_styling(latex_options = c("scale_down"))
+```
 
 # SUPPLEMENTARY FIGURES
-<!-- ```{r SFig 1, fig.cap="As expected, there was a correlation between the number of hits identified in a species with the evolutionary distance relative to humans"} -->
-<!--  -->
-<!-- ``` -->
-`r warning("TODO: I might have made more figures. Rename this after verifying that all figures are included somewhere in the paper.")`
-
-```{r Supplementary Figure 1, fig.cap="Supplementary Figure 1: Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
+
+```{r Figure-ECNC, fig.cap="Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
 __A)__ A single, contiguous gene homolog in a target genome with 100% query length coverage has an ECNC of 1.0.
@@
 Only 75% of the query sequence was covered in total between the hits, but for that 75%, each nucleotide has two hits.
-As such, ECNC is equal to 2.0 for this gene.", eval=F}
-sf1 <- list(
+As such, ECNC is equal to 2.0 for this gene.", eval=T}
+fig.ECNC <- list(
     ecnc.onegene.onehit, #A
@@
 #F: "ECNC: Two Real Genes, Unidentified Tails"
-sf1
-```
-
-```{r, Supplementary Figure 2, fig.cap="Supplementary Figure 2: Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number", eval=F}
+fig.ECNC
+```
+
+```{r Figure-GeneDup-BLScaled, fig.cap="Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number", eval=T}
 p.ml.tree.increasing.geneCounts.withSize.withBL
@@
 
-```{r, Supplementary Figure 3, fig.cap="Supplementary Figure 3: Full version of RecBlat strategy, but low quality", warning=T, eval=F}
+```{r Figure-RBHBStrategy, fig.cap="Full version of RecBlat strategy, but low quality", warning=T, eval=T}
 warning("TODO: Make HQ version of this figure")
-image_read("../output/figs/RBHBDiagram.png") %>% image_ggplot()
-```
-
-```{r Supplementary Figure 5, fig.cap="Supplementary Figure 5: All the Gene Copy Trees for interesting genes duplicated in LoxAfr4 (RIP any trees if this gets printed)", eval=F}
-p.ml.treedata.loxAfrHighlight.plus[1:10]
+image_read("../output/figs/RBHBDiagram.png") %>% 
+  image_rotate(180)%>% 
+  image_ggplot()
+```
+
+```{r InterestingGeneTrees, fig.cap="Supplementary Figure 5: All the Gene Copy Trees for interesting genes duplicated in LoxAfr4 (RIP any trees if this gets printed)", eval=F}
+p.ml.treedata.loxAfrHighlight.plus
 warning("Don't print the full list, seriously, its like 600+ pages")
@@
 
-```{r Supplementary Figure 6, fig.cap="Supplementary Figure 6: Expression data for interesting genes duplicated in LoxAfr4", eval=F}
-# plot.coolgenes
-```
-
-```{r Supplementary Figure 8, fig.cap="Supplementary Figure 8: The time-calibrated Eutherian tree used as input for StableTraits", warning=T, eval=F}
+```{r Figure-TimeTree-Eutheria, fig.cap="The time-calibrated Eutherian tree used as input for StableTraits, from Bininda-Emonds et al and Puttick and Thomas", warning=T, eval=T}
 p.time.tree.Eutheria <-
@@
   theme_tree()
-
+p.time.tree.Eutheria
+```
+
+```{r Figure-TimeTree-Atlantogenata, fig.cap="The time-calibrated Atlantogenatan tree used as input for StableTraits, from Puttick and Thomas", warning=T, eval=F}
 time.tree.atlantogenata <- time.tree %>%
@@
   theme_tree()
-# warning("TODO: this causes a segfault, find out why!")
-```
-
-# Supplementary data
-```{r Supplementary Table 1, eval=F}
-sra.data <- read_csv("../data/SraRunTable/SraRunTable.csv") %>% 
-  select(Organism, Run, source_name, sample_name) %>% 
-  filter(Organism %in% genomeTable.atlantogenata$Species) %>% 
-     mutate(sample_name = sample_name %>% str_remove("Nine-banded Armadillo_ ID 09-86_")) %>% 
-     mutate_all(. %>% str_replace_na("")) %>% 
-     unite("Tissue", source_name, sample_name, sep="") %>% 
-     filter(Tissue != "") %>% 
-     group_by(Organism) %>% 
-     summarize(Run = Run %>% unique %>% paste0(collapse=", "), 
-               Tissue = Tissue %>% str_to_title() %>% unique %>% paste0(collapse=", ")) %>% 
-     left_join(genomeTable.atlantogenata, by=c("Organism" = "Species")) %>% 
-  select(Organism, `Common Name`, Genome, `SRA Acc.`=Run, Tissues=Tissue) %>% 
-  mutate_all(. %>% str_wrap(width=100))
-sra.data %>% ggtexttable()
-```
-
-```{r Supplementary Table 2, fig.cap="Table 1: Body Size and Confidence Intervals in _Atlantogenata_ estimated using StableTraits.", eval=F}
-t.bodysize.atlantogenata.withNodes
-```
-
-```{r Supplementary Table 3, fig.cap="Supplementary Table 1: Parameters from StableTraits", warning=T, eval=F}
-warning("TODO: Read in Eutheria.progress and present it to the bold.")
-```
-
-```{r Table 4, fig.cap="Model Statistics for PGLS of Cancer Succeptibility", eval=F}
-t.size.lifespan.Atlantogenata.lm %>% cat
+p.time.tree.Atlantogenata
 ```

### 2020-06-03 paper_PLOS/paper_PLOS_draft.tex
Refactored figures and  tables, fixed Figure 6, broke Figure 4&7.

--- 2020-05-28 paper_PLOS/paper_PLOS_draft.tex
@@
 
-\setcounter{secnumdepth}{0}
+\setcounter{secnumdepth}{5}
 % Pandoc toggle for numbering sections (defaults to be off)
-\setcounter{secnumdepth}{0}
 
@@
 
-  \title{Pervasive duplication of tumor suppressor genes preceded parallel
-evolution of large bodied Atlantogenatans}
+  \title{Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans}
     \author[University of Chicago]{Juan Manuel Vazquez}
@@
   \begin{abstract}
-  Cancer is an intrinsic disease of multicellular organisms. Within a
-  species, the size of an animal, - correlated with the individual's
-  number of cells - and its lifespan - correlated with increasing cellular
-  damage over time - positively correlate with the risk any individual has
-  to form tumors. Between species, however, we do not observe any
-  correlation between size, lifespan, and cancer, a phenomenon that
-  referred to as \emph{Peto's Paradox}. Elephants are a particularly
-  interesting member of this class of paradoxical animals, since they are
-  a set of large species deeply nested in a clade of smaller species,
-  indicating a recent gain of size. Recent work has identified several
-  individual cases of gene duplicates contributing to the increased cancer
-  resistance of elephants, which suggests that duplication of tumor
-  suppressor genes may play a more general role in mediating Peto's
-  Paradox by increasing cancer resistance in large, long-lived species. By
-  using a Reciprocal Best-Hit BLAT search approach, we investigated copy
-  numbers of all protein-coding genes in \emph{Atlantogenatan} genomes to
-  see if there is any correlation between the copy number of duplicates
-  and changes body size along the phylogenetic tree. From an initial set
-  of 18,011 protein-coding genes in hg38, we identified a median of 13,880
-  genes in \emph{Atlantogenatan} genomes, of which a median of 940 genes
-  are duplicated. We find that, just as body size fluctuates throughout
-  \emph{Atlantogenata}, genes involved in tumor suppressor pathways are
-  also duplicated throughout the phylogenetic tree. Extant species of
-  elephants, however, show active transcription of both canonical and
-  duplicated copies of tummor suppressors that duplicated prior to and
-  during their sudden increase in body size, suggesting that the
-  duplication of tumor suppressor genes facilitates the evolution of
-  increased body size by compensating for the increased cancer risk.
+  Cancer is an intrinsic disease of multicellular organisms.
+Within a species, the size of an animal, - correlated with the individual's number of cells - and its lifespan - correlated with increasing cellular damage over time - positively correlate with the risk any individual has to form tumors.
+Between species, however, we do not observe any correlation between size, lifespan, and cancer, a phenomenon that referred to as \emph{Peto's Paradox}.
+Elephants are a particularly interesting member of this class of paradoxical animals, since they are a set of large species deeply nested in a clade of smaller species, indicating a recent gain of size.
+Recent work has identified several individual cases of gene duplicates contributing to the increased cancer resistance of elephants, which suggests that duplication of tumor suppressor genes may play a more general role in mediating Peto's Paradox by increasing cancer resistance in large, long-lived species.
+  By using a Reciprocal Best-Hit BLAT search approach, we investigated copy numbers of all protein-coding genes in \emph{Atlantogenatan} genomes to see if there is any correlation between the copy number of duplicates and changes body size along the phylogenetic tree.
+From an initial set of 18,011 protein-coding genes in hg38, we identified a median of 13,880 genes in \emph{Atlantogenatan} genomes, of which a median of 940 genes are duplicated.
+We find that, just as body size fluctuates throughout \emph{Atlantogenata}, genes involved in tumor suppressor pathways are also duplicated throughout the phylogenetic tree.
+Extant species of elephants, however, show active transcription of both canonical and duplicated copies of tummor suppressors that duplicated prior to and during their sudden increase in body size, suggesting that the duplication of tumor suppressor genes facilitates the evolution of increased body size by compensating for the increased cancer risk.
   \end{abstract}
@@
 
-)
-
 \hypertarget{introduction}{%
@@
 
-One of the major constraints on the evolution of large body sizes in
-animals is an increased risk of developing cancer. If all cells in all
-organisms have a similar risk of malignant transformation and equivalent
-cancer suppression mechanisms, organism with many cells should have a
-higher prevalence of cancer than organisms with fewer cells. Consistent
-with this expectation there is a strong positive correlation between
-body size and cancer incidence within species, for example, human cancer
-incidence increases with increasing adult height {[}1,2{]} and cancer
-incidence is positively correlated with body size in dogs {[}3,4{]}.
-There is no correlation, however, between body size and cancer risk
-between species. This lack of correlation is often referred to as
-`Peto's Paradox' {[}5--7{]}. While it is clear that a resolution to
-Peto's Paradox must involve the evolution of enhanced cancer protection
-alongside increases in body size and lifespan, the specific genetic,
-molecular, and cellular mechanisms that underlie this resistance have
-proven elusive. {[}8--12{]}.
+Among the major constraints on the evolution of large body sizes (and long life-spans) in animals is an increased risk of developing cancer.
+If all cells in all organisms have a similar risk of malignant transformation and equivalent cancer suppression mechanisms, organism with many cells should have a higher prevalence of cancer than organisms with fewer cells, particularly because large and small animals have similar cell sizes ({[}1{]}).
+Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, cancer incidence increases with increasing adult height in humans {[}2,3{]} and in dogs {[}4,5{]}.
+There is no correlation, however, between body size and cancer risk \emph{between} species; this lack of correlation is often referred to as `Peto's Paradox' {[}6--8{]}.
+The ultimate resolution to Peto's Paradox is obvious: large bodied and long-lived species evolved enhanced cancer protection mechanisms.
+However, identifiying the specific genetic, molecular, and cellular mechanisms that underlie the evolution of augmented cancer protection has been difficult {[}9--13{]}.
 
-Among the challenges for discovering how animals evolved enhanced cancer
-protection mechanisms is identifying lineages in which large bodied
-species are nested within species with small body sizes. Afrotherian
-mammals are generally small-bodied, similarly to the prediced common
-ancestor of Eutherian mammals. For example, maximum adult weights are
-\textasciitilde{}70g in golden moles, \textasciitilde{}120g in tenrecs,
-\textasciitilde{}170g in elephant shrews, \textasciitilde{}3kg in
-hyraxes, and 60kg in aardvarks {[}13{]}. However, while these extant
-species are relatively small, the fossil evidence demonstrates that
-their ancestral lineages reached enormous sizes. For example, while
-extant hyraxes are relatively small, the extinct Titanohyrax is
-estimated to have weighted up to \textasciitilde{}1300kg {[}14{]}. The
-largest members of Afrotheria, too, are dwarfed by the size of their
-recent ancestors: extant cows manatees are large bodied
-(\textasciitilde{}322-480kg) but are relatively small compared to the
-extinct Stellar's sea cow which is estimated to have weight 8000-10000kg
-{[}15{]}. Similarly African (4,800kg) and Asian elephants (3,200kg) are
-the largest living elephant species, but are dwarfed by the truly
-gigantic extinct Proboscideans such as Deinotherium
-(\textasciitilde{}132,000kg), Mammut borsoni (110,000kg), and the Asian
-straight-tusked elephant (\textasciitilde{}220,000kg), the largest known
-land mammal {[}16{]}. These large-bodied Afrotherian lineages are nested
-within small bodied species (Fig. 1) {[}17--20{]}, indicating that
-gigantism independently evolved in hyraxes, sea cows, and elephants
-(Paenungulates). Thus, Paenungulates are an excellent model system in
-which to explore the mechanisms that underlie the evolution of large
-body sizes and augmented cancer resistance.
+Among the challenges for discovering how animals evolved enhanced cancer protection mechanisms is identifying lineages in which large bodied species are nested within species with small body sizes.
+\emph{Afrotherian} mammals are generally small-bodied, but also include the largest extant land mammals.
+For example, maximum adult weights are \textasciitilde{}70g in golden moles, \textasciitilde{}120g in tenrecs, \textasciitilde{}170g in elephant shrews, \textasciitilde{}3kg in hyraxes, and 60kg in aardvarks {[}14{]}.
+While extant hyraxes are relatively small, the extinct Titanohyrax is estimated to have weighted up to \textasciitilde{}1300kg {[}15{]}.
+The largest members of \emph{Afrotheria}, too, are dwarfed by the size of their recent ancestors: extant cows manatees are large bodied (\textasciitilde{}322-480kg) but are relatively small compared to the extinct Stellar's sea cow which is estimated to have weight 8000-10000kg {[}16{]}.
+Similarly African (4,800kg) and Asian elephants (3,200kg) are the largest living elephant species, but are dwarfed by the truly gigantic extinct Proboscideans such as Deinotherium (\textasciitilde{}132,000kg), Mammut borsoni (110,000kg), and the Asian straight-tusked elephant (\textasciitilde{}220,000kg), the largest known land mammal {[}17{]}.
+Remarkably, these large-bodied \emph{Afrotherian} lineages are nested within small bodied species (\textbf{Figure \ref{fig:Figure-intro}}) {[}18--21{]}, indicating that gigantism independently evolved in hyraxes, sea cows, and elephants (\emph{Paenungulata}).
+Thus, Paenungulates are an excellent model system in which to explore the mechanisms that underlie the evolution of large body sizes and augmented cancer resistance.
 
-Although many mechanisms can potentially resolve Peto's paradox, among
-the most parsimonious routes to enhanced cancer resistance is through an
-increased copy number of tumor suppressors. Such an example has been
-seen in the case of candidate genes such as \emph{TP53} and \emph{LIF}
-{[}12,21,22{]} as well as in studies involving a limited set of
-candidate genes {[}23,24{]}. As these studies focus on \emph{a priori}
-gene sets, however, it remains unknown whether this is a general,
-genome-wide trend in Afrotherian genomes; and whether such a general
-trend is associated with the recent increases in body size -- and
-therefore expected cancer risk -- in these species.
+Although many mechanisms can potentially resolve Peto's paradox, among the most parsimonious routes to enhanced cancer resistance is through an increased copy number of tumor suppressors.
+Indeed, candidate genes studies have found that the elephant genome encodes duplicate such as \emph{TP53} and \emph{LIF} {[}13,22,23{]} as well as other genes with putative tumor suppressive functions {[}24,25{]}.
+As these studies focus on \emph{a priori} gene sets, however, it remains unknown whether this is a general, genome-wide trend in \emph{Afrotherian} genomes; and whether such a general trend is associated with the recent increases in body size -- and therefore expected cancer risk -- in these species.
 
-Here, we trace the evolution of body mass and gene copy number variation
-in Afrotherians in order to investigate whether gene duplications are
-enriched in large, long-lived species for genes involved in known tumor
-suppression pathways. Our estimates of the evolution of body mass,
-similarly to previous studies {[}17--20{]}, show that large body masses
-evolved in a step-wise manner, with major increases in body mass in the
-Pseudoungulata (17kg), Paenungulata (25kg), Tethytheria (296kg), and
-Proboscidea (4,100kg) stem-lineages. Furthermore, we see that the
-ancestral body size increases in Hydracoidia and Sirenia were
-independent events. To study the evolution of gene copy number, we used
-a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene
-duplications in Afrotherian genomes, and used maximum likelihood
-(treating copy number as a discrete trait) to infer the lineages in
-which those duplications occurred. We found gene duplications in
-lineages with increased body mass were enriched in functions related to
-tumor suppression, including regulation of the cell cycle, DNA damage
-repair, and regulation of apoptosis. These data suggest that duplication
-of tumor suppressors played a role in the evolution of large, long-lived
-in Afrotherians.
+Here, we trace the evolution of body mass and gene copy number variation in across \emph{Afrotherian} genomes in order to investigate whether duplication of tumors suppresor is common in large, long-lived Proboscideans.
+Our estimates of the evolution of body mass, similarly to previous studies {[}18--21{]}, show that large body masses evolved in a step-wise manner, with major increases in body mass in the \emph{Pseudoungulata} (17kg), \emph{Paenungulata} (25kg), \emph{Tethytheria} (296kg), and \emph{Proboscidea} (4,100kg) stem-lineages.
+To explore whether duplication of tumor suppressor genes occurred coincident with the evolution of large body sizes, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) strategy to identify gene duplications, and used maximum likelihood to infer the lineages in which those duplications occurred.
+Unexpectedly, we found that duplication of tumor suppressor genes was common in all Afrotherians, both large and small.
+These data suggest that duplication of tumor suppressor genes is pervasive in Afrotherians and proceeded the evolution of species with very large body sizes.
 
@@
 \hypertarget{ancestral-body-size-reconstruction}{%
-\subsection{Ancestral Body Size
-Reconstruction}\label{ancestral-body-size-reconstruction}}
+\subsection{Ancestral Body Size Reconstruction}\label{ancestral-body-size-reconstruction}}
 
-We built a time-calibrated supertree of Eutherian mammals by combining
-the time-calibrated molecular phylogeny of Bininda-Emonds \emph{et al.}
-{[}25{]} with the time-calibrated total evidence Afrotherian phylogeny
-from Puttick and Thomas {[}20{]}. While the Bininda-Emonds \emph{et al.}
-{[}25{]} phylogeny includes 1,679 species, only 34 are Afrotherian, and
-no fossil data are included. The inclusion of fossil data from extinct
-species is essential to ensure that ancestral state reconstructions of
-body mass are not biased by only including extant species. This can lead
-to inaccurate reconstructions, for example, if lineages convergently
-evolved large body masses from a small bodied ancestor. In contrast, the
-total evidence Afrotherian phylogeny of Puttick and Thomas {[}20{]}
-includes 77 extant species and fossil data from 39 extinct species.
-Therefore we replaced the Afrotherian clade in the Bininda-Emonds
-\emph{et al.} {[}25{]} phylogeny with the Afrotherian phylogeny of
-Puttick and Thomas {[}20{]} using Mesquite. Next, we jointly estimated
-rates of body mass evolution and reconstructed ancestral states using a
-generalization of the Brownian motion model that relaxes assumptions of
-neutrality and gradualism by considering increments to evolving
-characters to be drawn from a heavy-tailed stable distribution (the
-``Stable Model'') {[}26{]}. The stable model allows for occasional large
-jumps in traits and has previously been shown to out-perform other
-models of body mass evolution, including standard Brownian motion
-models, Ornstein--Uhlenbeck models, early burst maximum likelihood
-models, and heterogeneous multi-rate models {[}26{]}.
+We built a time-calibrated supertree of Eutherian mammals by combining the time-calibrated molecular phylogeny of Bininda-Emonds \emph{et al.} {[}26{]} with the time-calibrated total evidence Afrotherian phylogeny from Puttick and Thomas {[}21{]}.
+While the Bininda-Emonds \emph{et al.} {[}26{]} phylogeny includes 1,679 species, only 34 are Afrotherian, and no fossil data are included.
+The inclusion of fossil data from extinct species is essential to ensure that ancestral state reconstructions of body mass are not biased by only including extant species.
+This can lead to inaccurate reconstructions, for example, if lineages convergently evolved large body masses from a small bodied ancestor.
+In contrast, the total evidence Afrotherian phylogeny of Puttick and Thomas {[}21{]} includes 77 extant species and fossil data from 39 extinct species.
+Therefore we replaced the Afrotherian clade in the Bininda-Emonds \emph{et al.} {[}26{]} phylogeny with the Afrotherian phylogeny of Puttick and Thomas {[}21{]} using Mesquite.
+Next, we jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of the Brownian motion model that relaxes assumptions of neutrality and gradualism by considering increments to evolving characters to be drawn from a heavy-tailed stable distribution (the ``Stable Model'') implemented in StableTraits {[}27{]}.
+The stable model allows for occasional large jumps in traits and has previously been shown to out-perform other models of body mass evolution, including standard Brownian motion models, Ornstein--Uhlenbeck models, early burst maximum likelihood models, and heterogeneous multi-rate models {[}27{]}.
 
 \hypertarget{identification-of-duplicate-genes}{%
-\subsection{Identification of Duplicate
-Genes}\label{identification-of-duplicate-genes}}
+\subsection{Identification of Duplicate Genes}\label{identification-of-duplicate-genes}}
 
-\textbf{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit
-BLAT (RBHB) pipeline to identify putative homologs and estimate gene
-copy numbers across species (\textbf{Figure 1A}). The Reciprocal Best
-Hit (RBH) search strategy is conceptually straightforward: 1) Given a
-gene of interest \(G_A\) in a query genome \(A\), one searches a target
-genome \(B\) for all possible matches to \(G_A\); 2) For each of these
-hits, one then performs the reciprocal search in the original query
-genome to identify the highest-scoring hit; 3) A hit in genome \(B\) is
-defined as a homolog of gene \(G_A\) if and only if the original gene
-\(G_A\) is the top reciprocal search hit in genome \(A\). We selected
-BLAT {[}27{]} as our algorithm of choice, as this algorithm is sensitive
-to highly simliar (\textgreater{}90\% identity) sequences, thus
-identifying the highest-confidence homologs while minimizing many-to-one
-mapping problems when searching for multiple genes. RBH performs similar
-to other more complex methods of orthology prediction, and is
-particularly good at identifying incomplete genes that may be fragmented
-in low quality/poor assembled regions of the genome {[}28,29{]}.
+\textbf{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit BLAT (RBHB) pipeline to identify putative homologs and estimate gene copy number across species (\ref{fig:Figure-RBHBStrategy}).
+The Reciprocal Best Hit (RBH) search strategy is conceptually straightforward: 1) Given a gene of interest \(G_A\) in a query genome \(A\), one searches a target genome \(B\) for all possible matches to \(G_A\); 2) For each of these hits, one then performs the reciprocal search in the original query genome to identify the highest-scoring hit; 3) A hit in genome \(B\) is defined as a homolog of gene \(G_A\) if and only if the original gene \(G_A\) is the top reciprocal search hit in genome \(A\).
+We selected BLAT {[}28{]} as our algorithm of choice, as this algorithm is sensitive to highly simliar (\textgreater{}90\% identity) sequences, thus identifying the highest-confidence homologs while minimizing many-to-one mapping problems when searching for multiple genes.
+RBH performs similar to other more complex methods of orthology prediction, and is particularly good at identifying incomplete genes that may be fragmented in low quality/poor assembled regions of the genome {[}29,30{]}.
 
-\textbf{Effective Copy Number By Coverage:} In lower-quality genomes,
-many genes are fragmented across multiple scaffolds, which results in
-BLAT calling multiple hits when in reality there is only one gene. To
-compensate for this, we came up with a novel statistic, Estimated Copy
-Number by Coverage (ECNC), which averages the number of times we see
-each nucleotides of a query sequence in a target genome over the total
-number of nucleotides of the query sequence found overall in each target
-genome (Supplementary Figure 1). This allows us to correct for genes
-that have been fragmented across incomplete genomes, while also taking
-into account missing sequences from the human query in the target
-genome. Mathematically, this can be written as:\\
-\[ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}\] where
-\(n\) is a given nucleotide in the query, \(l\) is the total length of
-the query, \(C_n\) is the number of instances that \(n\) is present
-within a reciprocal best hit, and \(bool(C_n)\) is 1 if \(C_n > 0\) or 0
-if \(C_n = 0\).
+\textbf{Effective Copy Number By Coverage:} In lower-quality genomes, many genes are fragmented across multiple scaffolds, which results in BLAT calling multiple hits when in reality there is only one gene.
+To compensate for this, we came up with a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we see each nucleotides of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (\textbf{Figure \ref{fig:Figure-ECNC}}).
+This allows us to correct for genes that have been fragmented across incomplete genomes, while also taking into account missing sequences from the human query in the target genome.
+Mathematically, this can be written as:\\
+\[ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}\]
+where \(n\) is a given nucleotide in the query, \(l\) is the total length of the query, \(C_n\) is the number of instances that \(n\) is present within a reciprocal best hit, and \(bool(C_n)\) is 1 if \(C_n > 0\) or 0 if \(C_n = 0\).
 
-\textbf{RecSearch Pipeline:} We created a custom Python pipeline for
-automating RBHB searches between a single reference genome and multiple
-target genomes using a list of query sequences from the reference
-genome. For the query sequences in our search, we used the hg38 Proteome
-provided by UniProt {[}30{]}, which is a comprehensive set of protein
-sequences curated from a combination of predicted and validated protein
-sequences generated by the UniProt Consortium. In order to refine our
-search, we omitted protein sequences originating from long, noncoding
-RNA loci (e.g.~LINC genes); poorly-studied genes from predicted open
-reading frames (C-ORFs); and sequences with highly repetitive sequences
-such as zinc fingers, protocadherins, and transposon-containing genes,
-as these were prone to high levels of false positive hits.
+\textbf{RecSearch Pipeline:} We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome.
+For the query sequences in our search, we used the hg38 Proteome provided by UniProt {[}31{]}, which is a comprehensive set of protein sequences curated from a combination of predicted and validated protein sequences generated by the UniProt Consortium.
+In order to refine our search, we omitted protein sequences originating from long, noncoding RNA loci (e.g.~LINC genes); poorly-studied genes from predicted open reading frames (C-ORFs); and sequences with highly repetitive sequences such as zinc fingers, protocadherins, and transposon-containing genes, as these were prone to high levels of false positive hits.
 
-After filtering out problematic protein queries (see ``Query gene
-inclusion criteria''), we then used our pipeline (Figure 1A) to search
-for all copies of our 18011 query genes in publicly available
-Afrotherian genomes, including African savannah elephant
-(\emph{Loxodonta africana}: loxAfr3, loxAfr4, loxAfrC), African forest
-elephant (\emph{Loxodonta cyclotis}: loxCycF), Asian Elephant
-(\emph{Elephas maximus}: eleMaxD), Woolly Mammoth (\emph{Mammuthus
-primigenius}: mamPriV), Colombian mammoth (\emph{Mammuthus columbi}:
-mamColU), American mastodon (\emph{Mammut americanum}: mamAmeI), Rock
-Hyrax (\emph{Procavia capensis}: proCap1, proCap2, proCap2\_HiC), West
-Indian Manatee (\emph{Trichechus manatus latirostris}: triManLat1,
-triManLat1\_HiC), Aardvark (\emph{Orycteropus afer}: oryAfe1,
-oryAfe1\_HiC), Lesser Hedgehog Tenrec (\emph{Echinops telfairi}:
-echTel2), Nine-banded armadillo (\emph{Dasypus novemcinctus}: dasNov3),
-Hoffman's two-toed sloth (\emph{Choloepus hoffmannii}: choHof1, choHof2,
-choHof2\_HiC), Cape golden mole (\emph{Chrysochloris asiatica}:
-chrAsi1), and Cape elephant shrew (\emph{Elephantulus edwardii}:
-eleEdw1). For many of these species, we covered multiple assemblies in
-order to test the effects of assembly size and quality on our hits.
+After filtering out problematic protein queries (see below), we then used our pipeline to search for all copies of our 18011 query genes in publicly available \emph{Afrotherian} genomes (\ref{tab:Table-Genomes}), including African savannah elephant (\emph{Loxodonta africana}: loxAfr3, loxAfr4, loxAfrC), African forest elephant (\emph{Loxodonta cyclotis}: loxCycF), Asian Elephant (\emph{Elephas maximus}: eleMaxD), Woolly Mammoth (\emph{Mammuthus primigenius}: mamPriV), Colombian mammoth (\emph{Mammuthus columbi}: mamColU), American mastodon (\emph{Mammut americanum}: mamAmeI), Rock Hyrax (\emph{Procavia capensis}: proCap1, proCap2, proCap2\_HiC), West Indian Manatee (\emph{Trichechus manatus latirostris}: triManLat1, triManLat1\_HiC), Aardvark (\emph{Orycteropus afer}: oryAfe1, oryAfe1\_HiC), Lesser Hedgehog Tenrec (\emph{Echinops telfairi}: echTel2), Nine-banded armadillo (\emph{Dasypus novemcinctus}: dasNov3), Hoffman's two-toed sloth (\emph{Choloepus hoffmannii}: choHof1, choHof2, choHof2\_HiC), Cape golden mole (\emph{Chrysochloris asiatica}: chrAsi1), and Cape elephant shrew (\emph{Elephantulus edwardii}: eleEdw1).
 
-\textbf{Query gene inclusion criteria:} To assemble our query list, we
-first removed all unnamed genes from UP000005640. Next, we excluded
-genes from downstream analyses for which assignment of homology was
-uncertain, including uncharacterized ORFs (991), LOC (63), HLA genes
-(402), replication dependent histones (72), odorant receptors (499),
-ribosomal proteins (410), zinc finger transcription factors (1983),
-viral and repetitive-element-associated proteins (82) and any protein
-described as either ``Uncharacterized,'' ``Putative,'' or ``Fragment''
-by UniProt in UP000005640 (30724), leaving us with a final set of 37582
-query protein sequences, corresponding to 18011 genes.
+\textbf{Query gene inclusion criteria:} To assemble our query list, we first removed all unnamed genes from UP000005640.
+Next, we excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (991 genes), LOC (63 genes), HLA genes (402 genes), replication dependent histones (72 genes), odorant receptors (499 genes), ribosomal proteins (410 genes), zinc finger transcription factors (1983 genes), viral and repetitive-element-associated proteins (82 genes) and any protein described as either ``Uncharacterized,'' ``Putative,'' or ``Fragment'' by UniProt in UP000005640 (30724 genes), leaving us with a final set of 37582 query protein isoforms, corresponding to 18011 genes.
 
-\textbf{Duplication gene inclusion criteria:} In order to condense
-transcript-level hits into single gene loci, and to resolve many-to-one
-genome mappings, we removed exons where transcripts from different genes
-overlapped, and merged overlapping transcripts of the same gene into a
-single gene locus call. The resulting gene-level copy number table was
-then combined with the maximum ECNC values observed for each gene in
-order to call gene duplications. We called a gene duplicated if its copy
-number was two or more, and if the maximum ECNC value of all the gene
-transcripts searched was 1.5 or greater; previous studies have shown
-that incomplete duplications can encode functional genes {[}12,22{]},
-therefore partial gene duplications were included provided they passed
-additional inclusion criteria. The ECNC cut off of 1.5 was selected
-empirically, as this value minimized the number of false positives seen
-in a test set of genes and genomes. The results of our initial search
-are summarized in Figure 1B. Overall, we identified 13880 genes across
-all species, or 77.1\% of our starting query genes.
+\textbf{Duplication gene inclusion criteria:} In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call.
+The resulting gene-level copy number table was then combined with the maximum ECNC values observed for each gene in order to call gene duplications.
+We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes {[}13,23{]}, therefore partial gene duplications were included provided they passed additional inclusion criteria (see below).
+The ECNC cut off of 1.5 was selected empirically, as this value minimized the number of false positives seen in a test set of genes and genomes.
+The results of our initial search are summarized in Figure \ref{fig:Figure-GeneDup-Cladogram}.
+Overall, we identified 13880 genes across all species, or 77.1\% of our starting query genes.
 
-\textbf{Genome Quality Assessment using CEGMA:} In order to determine
-the effect of genome quality on our results, we used the gVolante
-webserver and CEGMA to assess the quality and completeness of the genome
-{[}31,32{]}. CEGMA was run using the default settings for mammals
-(``Cut-off length for sequence statistics and composition'' = 1;``CEGMA
-max intron length'' = 100000; ``CEGMA gene flanks'' = 10000, ``Selected
-reference gene set'' = CVG). For each genome, we generated a correlation
-matrix using the aforementioned genome quality scores, and either the
-mean Copy Number or mean ECNC for all hits in the genome.
+\textbf{Genome Quality Assessment using CEGMA:} In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome {[}32,33{]}.
+CEGMA was run using the default settings for mammals (``Cut-off length for sequence statistics and composition'' = 1;``CEGMA max intron length'' = 100000; ``CEGMA gene flanks'' = 10000, ``Selected reference gene set'' = CVG).
+For each genome, we generated a correlation matrix using the aforementioned genome quality scores, and either the mean Copy Number or mean ECNC for all hits in the genome.
 
-\hypertarget{evidence-for-functionality-of-identified-genes}{%
-\subsection{Evidence for Functionality of Identified
-Genes}\label{evidence-for-functionality-of-identified-genes}}
+\hypertarget{evidence-for-functionality-of-gene-duplicates}{%
+\subsection{Evidence for Functionality of Gene Duplicates}\label{evidence-for-functionality-of-gene-duplicates}}
 
-To validate and filter out duplicate gene calls, we intersected our
-results with either gene prediction or transcriptomic evidence as a
-proxy for functionality.
+To validate and filter out duplicate gene calls, we intersected our results with either gene prediction or transcriptomic evidence as a proxy for functionality.
 
-\textbf{Transcriptome Assembly:} For the African Savana Elephant, Asian
-Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated
-\emph{de novo} transcriptomes using publically-available RNA-sequencing
-data from NCBI SRA (Supplementary Table 1). We mapped reads to all
-genomes available for each species, and assembled transcripts using
-HISAT2 and StringTie, respectively {[}33--35{]}. RNA-sequencing data was
-not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax,
-Aardvark, or the Lesser Hedgehog Tenrec.
+\textbf{Transcriptome Assembly:} For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated \emph{de novo} transcriptomes using publically-available RNA-sequencing data from NCBI SRA (\ref{tab:Table-SRAData}).
+We mapped reads to all genomes available for each species, and assembled transcripts using HISAT2 and StringTie, respectively {[}34--36{]}.
+RNA-sequencing data was not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax, Aardvark, or the Lesser Hedgehog Tenrec.
 
-\textbf{Gene Prediction:} We obtained tracks for genes predicted using
-GenScan for all the genomes available via UCSC Genome Browser: African
-savannah elephant (loxAfr3), Rock Hyrax (proCap1), West Indian Manatee
-(triManLat1), Aardvark (oryAfe1), Lesser Hedgehog Tenrec (echTel2),
-Nine-banded armadillo (dasNov3), Hoffman's Two-Toed Sloth (choHof1),
-Cape golden mole (chrAsi1), and Cape Elephant Shrew (eleEdw1); gene
-prediction tracks for higher-quality assemblies were not available.
+\textbf{Gene Prediction:} We obtained tracks for genes predicted using GenScan for all the genomes available via UCSC Genome Browser: African savannah elephant (loxAfr3), Rock Hyrax (proCap1), West Indian Manatee (triManLat1), Aardvark (oryAfe1), Lesser Hedgehog Tenrec (echTel2), Nine-banded armadillo (dasNov3), Hoffman's Two-Toed Sloth (choHof1), Cape golden mole (chrAsi1), and Cape Elephant Shrew (eleEdw1); gene prediction tracks for higher-quality assemblies were not available.
 
-\textbf{Evidenced Duplicate Criteria:} We intersected our records of
-duplicate hits identified in each genome with the gene prediction tracks
-and/or transcriptome assemblies using \texttt{bedtools}. When multiple
-lines of evidence for functionality were present for a genome, we used
-the union of all intersections as the final output for evidenced
-duplicates. When analyzing the highest-quality assemblies available for
-each species, if a species had neither gene prediction tracks nor
-RNA-seq data for the highest-quality genome available, we conservatively
-included all hits for the genome in the final set of evidenced
-duplicates.
+\textbf{Evidenced Duplicate Criteria:} We intersected our records of duplicate hits identified in each genome with the gene prediction tracks and/or transcriptome assemblies using \texttt{bedtools}.
+When multiple lines of evidence for functionality were present for a genome, we used the union of all intersections as the final output for evidenced duplicates.
+When analyzing the highest-quality assemblies available for each species, if a species had neither gene prediction tracks nor RNA-seq data for the highest-quality genome available, we conservatively included all hits for the genome in the final set of evidenced duplicates.
 
 \hypertarget{reconstruction-of-ancestral-copy-numbers}{%
-\subsection{Reconstruction of Ancestral Copy
-Numbers}\label{reconstruction-of-ancestral-copy-numbers}}
+\subsection{Reconstruction of Ancestral Copy Numbers}\label{reconstruction-of-ancestral-copy-numbers}}
 
-We implemented a maximum likelihood method for determining the ancestral
-copy numbers of genes in \emph{Atlantogenata} using IQ-Tree. For this
-analysis, we used an unrooted species tree including only the
-aforementioned \emph{Atlantogenata} species. We generated PHYLIP files
-containing the copy number of each gene in the highest quality genome
-for each species, encoding genes on a scale from 1-31+ copies as 1-9,
-A-V; and encoding a gene's copy number as uncetain (``?'') when we did
-not identify it in the genome. We used the included tree-searching and
-model-testing functionality in IQ-Tree to determine the most likely
-topology for the species tree, and to obtain the most likely model for
-copy number changes in the genome. The most likely model for the
-evolution of copy number was inferred to be a Jukes-Cantor type model
-for morphological data with equal state frequencies, allowing for a
-proportion of invariable sites, and using a discrete Gamma model with 4
-rate categories. (``MK+FQ+I+G4''). We defined the ancestral state of a
-node if it had greater than an 80\% posterior probability.
+We encoded the copy number of each gene for each species as a discrete trait ranging from 0 (one gene copy) to 31 (for 32+ gene copies) and used IQ-TREE to select the best-fitting model of character evolution {[}37--41{]}, which was inferred to be a Jukes-Cantor type model for morphological data (MK) with equal character state frequencies (FQ) and rate heterogeneity across sites approximated by including a class of invariable sites (I) plus a discrete Gamma model with four rate categories (G4).
+Next we inferred gene duplication and loss events with the empirical Bayesian ancestral state reconstruction (ASR) method implemented in IQ-TREE {[}37--41{]}, the best fitting model of character evolution (MK+FQ+GR+I) {[}42,43{]}, and the unrooted species tree for Atlantogenata.
+We considered ancestral state reconstructions to be reliable if they had Bayeisan Posterior Probability (BPP) ?0.80; less reliable reconstructions were excluded for pathway analyses.
 
 \hypertarget{pathway-enrichment-analysis}{%
-\subsection{Pathway Enrichment
-Analysis}\label{pathway-enrichment-analysis}}
+\subsection{Pathway Enrichment Analysis}\label{pathway-enrichment-analysis}}
 
-To determine which pathways were associated with duplicated genes in
-each species and lineage, we used WEBGESTALT to perform
-overrepresentation analysis (ORA) of the duplicated gene lists relative
-to our initial query gene list {[}36{]}. For the database of pathways
-used in the analysis, we used Reactome {[}37{]}, Wikipathways, {[}38{]},
-and KEGG {[}39{]}. For the ORA, we used FDR for determining
-significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and
-FDR=0.5.
+To determine if gene duplications were enriched in particular biological pathways, we used the WEB-based Gene SeT AnaLysis Toolkit (WEBGESTALT){[}44{]} to perform overrepresentation analysis (ORA); pathway databases included Reactome {[}45{]}, Wikipathways, {[}46{]}, and KEGG {[}47{]}.
+Gene duplicates in each lineage were used as the foreground gene set, and the initial query set was used as the background gene set.
+Statistical significance of enriched terms was assessed with a hypergeometric test, and controlled for multiple testing using the Benjamini-Hochberg false discovery rate (FDR); each analysis was run at FDR=0.1, FDR=0.2, FDR=0.3, and FDR=0.5.
+In order to determine an empirical false positive rate for term enrichment, we randomly sampled 1000-10,000 genes from our background set 1000 times, and ran the aforementioned analyses to see which terms were likely to randomly appear; there were no terms which appeared at FDR \(\leq\) 0.3.
 
-\hypertarget{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}{%
-\subsection{Lifespan Phylogenetic Least-Square Regression and
-Calculating Estimated Cancer Risk Thoughout
-Atlantogenata}\label{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}}
+\hypertarget{estimating-the-evolution-of-cancer-risk}{%
+\subsection{Estimating the Evolution of Cancer Risk}\label{estimating-the-evolution-of-cancer-risk}}
 
-In order to determine the cancer risk \(K\) at each node, we first
-needed to calcultate ancestral lifespans at each node. To do so, we used
-a Phylogenetic Generalized Least-Square Regression (PGLS) {[}40,41{]} to
-calculate estimated ancestral lifespans across \emph{Atlantogenata}
-using our estimates for body size at each node.
-
-Next, we used a simplified multistage cancer risk model for body size
-\(D\) and lifespan \(t\): \(K \approx Dt^6\) {[}7,42--44{]}. To
-determine the change in \(K\) between nodes, we obtained the ratio
-between the cancer risk \(K_{1}\) at any given node, and the cancer risk
-\(K_{2}\) at its ancestral node, using the equation
-\(\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}\).
-Finally, to simplify comparisons, we calculated the fold change cancer
-risk between a node and its ancestor as
-\(\log_{2}(\frac{K_{2}}{K_{1}})\).
+The dramatic increase in body mass and lifespan in some Afrotherian lineages implies those lineages must have also evolved reduced cancer risk.
+To infer the magnitude of these reductions we estimated differences in intrinsic cancer risk across extant and ancestral Afrotherians.
+Following Peto {[}48{]} we estimate the intrinsic cancer risk as the product of risk associated with body mass and lifespan.
+In order to determine the intrinsic cancer risk (\(K\)) across species and at ancestral nodes (see below), we first needed to estimate ancestral lifespans at each node.
+We used Phylogenetic Generalized Least-Square Regression (PGLS) {[}49,50{]} implemented in the R package \texttt{ape} to calculate estimated ancestral lifespans across \emph{Atlantogenata} using our estimates for body size at each node.
+In order to estimate the intrinsic cancer risk of a species, we first inferred lifespans at ancestral nodes using PGLS and the model \(ln(lifespan) = \beta_{1}corBrowninan + \beta_{2}ln(size) + \epsilon\).
+Next, we calculated \(K_{1}\) at all nodes, and then estimated the fold change in cancer susceptibility between ancestral and descendant nodes (Figure \ref{fig:Figure-CancerSuscept}, Table \ref{tab:Table-CancerSuscept}).A Next, in oder to calculate \(K_{1}\) at all nodes, we used a simplified multistage cancer risk model for body size \(D\) and lifespan \(t\): \(K \approx Dt^6\) {[}8,48,51,52{]}.
+The fold change in cancer risk between a node and its ancestor was then defined as \(\log_{2}(\frac{K_{2}}{K_{1}})\).
 
@@
 
-\hypertarget{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}{%
-\subsection{\texorpdfstring{Body size frequently and independently
-expands and contracts throughout
-\emph{Atlantogenata}}{Body size frequently and independently expands and contracts throughout Atlantogenata}}\label{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}}
+\hypertarget{step-wise-evolution-of-body-size-in-afrotherians}{%
+\subsection{\texorpdfstring{Step-wise evolution of body size in \emph{Afrotherians}}{Step-wise evolution of body size in Afrotherians}}\label{step-wise-evolution-of-body-size-in-afrotherians}}
+
+Similar to previous studies of Afrotherian body size {[}21,27{]}, we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95\% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a 67.36x increase in body mass in the stem-lineage of \emph{Pseudoungulata} (17.33kg), a 1.45x increase in body mass in the stem-lineage of \emph{Paenungulata} (25.08kg), a 11.82x increase in body mass in the stem-lineage of \emph{Tehthytheria} (296.56kg), and a 2.69x increase in body mass in the stem-lineage of \emph{Proboscidea} (4114.39kg) (Figure \ref{fig:Figure-BodySize}, Table \ref{tab:Table-BodySize}).
+The ancestral \emph{Hyracoidea} was inferred to be relatively small (2.86kg-118.18kg), and rate accelerations were coincident with independent body mass increases in large hyraxes such as \emph{Titanohyrax andrewsi} (429.34kg, 67.36x increase).
+While the body mass of the ancestral \emph{Sirenian} was inferred to be large (61.7kg-955.51kg), a rate acceleration occurred coincident with a 10.59x increase in body mass in Stellar's sea cow.
+Rate accelerations also occurred coincident with 36.6x decrease in body mass in the stem-lineage of the dwarf elephants \emph{Elephas (Palaeoloxodon) antiquus falconeri} and \emph{Elephas cypriotes}.
+These data suggest that gigantism in \emph{Afrotherians} evolved step-wise, from small to medium bodies in the \emph{Pseudoungulata} stem-lineage, medium to large bodies in the \emph{Tehthytherian} stem-lineage and extinct hyraxes, and from large to exceptionally large bodies independently in the \emph{Proboscidean} stem-lineage and Stellar's sea cow (Figure \ref{fig:Figure-BodySize}, Table \ref{tab:Table-BodySize}).
+
+\hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
+\subsection{Step-wise reduction of intrinsic cancer risk in large, long-lived Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}
+
+As expected, intrinsic cancer susceptibility in Afrotheria also varies with changes in body size and longevity (Figure \ref{fig:Figure-CancerSuscept}, Table \ref{tab:Table-CancerSuscept}), with an initial 9.22-fold and 20.75-fold increase in the stem-lineage of Afrotheria and Xenarthra, respectively, followed by a 9.22-fold increases in Pseudoungulata and a 5.38-fold increase in Aardvarks (Figure \ref{fig:Figure-CancerSuscept}, Table \ref{tab:Table-CancerSuscept}).
+In contrast to the Paenungulate stem-lineage, there is a 6.92-fold increase in cancer risk in Tethytheria, a 8.62-fold increase in Manatee, and dramatic increases within Proboscidea including a 27.66-fold increase in Elephantidae and a 29.97-fold in the American Mastodon.
+Within the Elephantidae, Elephantina and Loxodontini have a 2.31-fold increase in cancer susceptibility, while susceptibility is relatively stable in Mammoths.
+The three extant Proboscideans, Asian Elephant, African Savana Elephant, and the African Forest Elephant, meanwhile, have similar decreases in both size and cancer susceptibility (Figure \ref{fig:Figure-CancerSuscept}, Table \ref{tab:Table-CancerSuscept}).
+
+\hypertarget{identification-and-evolutionary-history-of-gene-duplications}{%
+\subsection{Identification and evolutionary history of gene duplications}\label{identification-and-evolutionary-history-of-gene-duplications}}
+
+We found that gene duplications were common in Atlantogenatan genomes (Figure \ref{fig:Figure-GeneDup-Cladogram}), identifying an average of NA genes, with an average of 10.53\% duplicated; this is in-line with other studies describing the rates of gene duplications over time.
+{[}53{]}.
+We observed that the percentage of duplicated genes in non-\emph{Pseudoungulatan} genomes was significantly higher: while \emph{Pseudoungulatan} genomes had duplicates percentages ranging anywhere from 3.26\% to 7.80\%, outgroup species' duplication rates ranged from 12.94\% to 23.66\%.
+To explore whether genome quality may adversely effect our inferences, we used CEGMA and the gVolante server {[}32,33{]} to assess the correlation between genome quality and copy number estimates.
+As shown in Figure \ref{fig:Figure-GenomeCor}, mean Copy Number, mean ECNC, and mean CN (the lesser of Copy Number and ECNC per gene) moderately or strongly correlate with genomic quality, such as LD50, the number of scaffolds, and contigs with a length above either 100K or 1M.
+
+Among the genes that increased in copy number in the elephant lineage are TP53 and LIF, as previously described; however, we find that these two genes represent a fraction of the 940 genes that are duplicated in the African Elephant overall, which accumulated over various steps through their evolution.
+While the extinct elephantids have acceptable genome quality metrics according to CEGMA, they are nonetheless missing a significant number of sequences; this may contribute to the low number of duplicated genes that occured in internal nodes.
+The number of duplicates that occur at each branch is also proportional to the density of the sampling of the clade overall, as would be expected.
+In branches, such as \emph{Afrosoricida}, where the number of species is relatively minuscule compared to the size of the clade, we see many significantly larger numbers of duplications private to these species.
+
+\hypertarget{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}{%
+\subsection{Duplications that occurred recently in Probodiscea are enriched for tumor suppressor pathways}\label{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}}
+
+Our initial hypothesis was that genes which duplicated in lineages that experienced a growth in size would be enriched for membership in tumor suppressor pathways.
+Thus, we used WebGestalt and its Overrepresentation Analysis (ORA) functionality to determine what pathways were enriched in our duplicated gene sets in each branch relative to our initial query set.
+For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways\_cancer databases using WebGestalt ORA.
+Going through the tree, at no FDR \(\leq\) 0.5 is there any significant pathway representation for genes that increased in the branches leading to \emph{Afrosoricida}, \emph{Pseudoungulata}, \emph{Elephantina}, \emph{Loxodontini}, \emph{Elephas maximus}, \emph{Mammuthus}, \emph{Loxodonta cyclotis}, \emph{Palaeoloxodon antiquus}, \emph{Loxodona}, \emph{Mammuthus columbi}, \emph{Mammuthus primigenius}, \emph{Procavia capensis}, \emph{Elephantidae}, or \emph{Proboscidea}; furthermore, there are no significant pathway enrichments at FDR\textless{}0.5 for genes whose copy number did not change between branches (copy-number-stable) for \_Afrotheria\_\_.
+Note that because \emph{Xenarthra} was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.
+
+For the other branches, the number of pathways that came up as significantly enriched at each FDR is shown in Table \ref{tab:Table-OraFreqDatabase}.
+For the species with high duplication rates and lower-quality, highly-fragmented genomes, such as with \emph{Chrysochloris asiatica} (20.05\% duplicated hits) and \emph{Elephantulus edwardii} (23.66\% duplicated hits), it is unsurprising that there is a proportionally large number of pathway enrichments.
+In the case of these two species, their many pathway enrichments also span an incredible range of processes at every level of biology; this, in combination with the high number of copy numbers identified for these genes, further suggests a need for improvement and refinement in these genomes.
+In the cell cycle pathways called as significant in the genomes of these two species plus \emph{Orycteropus afer} and \emph{Echinops telfairi}, the duplicated genes included in these sets are from the same gene families, such as the APC subunit family; the proteosome subunit families; and the protein phosphatase 2 family, among others.
+It is highly possible that these results reflect true expansions of these gene families, especially in the higher-quality \emph{Orycteropus afer} genome; however, it is also possible that it simply reflects artifactual duplications, and so require further study.
+
+The pathway enrichments for genes whose copy number either did not change, or whose copy number increased, between \emph{Loxodona} and the African Elephant are shown in Figure \ref{fig:Figure-loxAfrORA}.
+Among the few enriched pathways in the African Elephant, we see that that two tumor suppression pathways - APC Complex-related pathways, and ``TP53 Regulates Metabolic Genes''- appear not only in the case of stable genes, but also in the set of newly duplicated genes.
+The other pathways we see in the set of recently-duplicated genes include ``Functionalization of Compounds'' and its daughter pathway ``Xenobiotics''.
+Genes in these pathways serve to add functional groups to lipophylic compounds which would otherwise not be reactive in the cell, and are types of metabolic pathways.
+In the stable set we see enrichment of pathways such as ``Neuronal Systems'' and ``Axon Guidance,'' which fit in well with what is known about elephant biology and evolution {[}54{]}.
+Overall in elephants, we see enrichments within duplicated genes for pathways involved in what makes an elephant an elephant - including tumor suppressor pathways.
+
+\hypertarget{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}{%
+\subsection{\texorpdfstring{Concerted duplication of TP53 and TP53-related genes towards \emph{Probodiscea}}{Concerted duplication of TP53 and TP53-related genes towards Probodiscea}}\label{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}}
+
+Prior studies looking at the duplication of TP53 in the African Elephant motivated further study for the enrichment of genes involved in TP53-related metabolic pathway.
+We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, and used publicly-available RNA-seq data to see which genes are actively expresses in living elephants.
+Excitingly, we found that the initial duplication of TP53 in Tethytheria, where body size expanded, was preceded by the duplication of GTF2F1 and STK11 in \emph{Paenungulata}; and was coincided by the duplication of BRD7.
+These two genes are involved in regulating the transcription of TP53, and their duplication prior to that of TP53 may have facilitated its retroduplication.
+Interestingly enough, STK11 is a tumor suppressor gene in its own right, and plays additional roles in mediating tumor suppression via p21-induced senescence {[}55{]}.
+
+The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either alongside TP53 in the case of SIAH1, or subsequently in \emph{Probodiscea}, \emph{Elephantidae}, and in modern elephants.
+(Figure \ref{fig:Figure-loxAfrTP53}).
+These genes are all expressed in RNA-seq data, suggesting that they encode functional genes in modern elephants (Figure \ref{fig:Figure-loxAfrTP53}).
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+With our results, we have demonstrated that gigantism in \emph{Atlantogenata} was not limited to extant elephants, but rather occurred at various points in the evolution of the clade; however, the hundred-fold to hundred-million-fold increases in cancer risk that is associated with these increases however poses an innate challenge in the evolution and persistence of this trait.
+The environmental selective pressures on body size have long been the fascination of evolutionary biologists, and the influence of climate, predation, geography, and ecological niche on body size have been well established.
+{[}56{]}.
+Indeed, there is a general trend, known as Cope's Rule, for body sizes of species to increase over time {[}57--59{]}.
+However, for all the research on body size that has been done thus far, the mechanisms that enable a release from the negative pressure on body size exerted on cancer risk has proven more elusive {[}57--59{]}.
+
+Furthermore, we show that tumor suppressor duplications are enriched not only in large, extant species, but also in large common ancestors, and that these duplications evolved throughout the tree, rather than in concert.
+The ancestral body sizes of many of the subclades in \emph{Atlantogenata} were estimated to be large, and the estimated cancer risk increases - even for the small clades like \emph{Afroinsectivora} may explain why these TSG gene duplications occurred early on, and persisted even in species that we did not expect to have high risks of cancer.
+However, some of our results also provide interesting evidence for a paradigm of TSG duplications being a pervasive phenomenon, rather than a specific mechanism that occurs after the evolution of gigantism.
+At the common ancestor of \emph{Proboscidea}, which was small relative to both its ancestors and its descendants, we see the emergence of various TSG duplications, which may have enabled the stratospheric increase in body size of modern elephants.
+We also identified many TSG duplication events in smaller species and lineages such as in \emph{Chrysochloris asiatica} and \emph{Elephantulus edwardii}, although these may have been the result of low-quality genomes.
+
+The impacts and takeaways of this study are limited by the quality and quantity of \emph{Atlantogenatan} genomes that were available, and our available knowledge of the lifespans and cancer risk of the extant species.
+For many of our species, studies in captivity are limited, and the species are relatively understudied from a longitudinal perspective, such as with the Cape Elephant Shrew and Cape Golden Mole.
+Furthermore, while there is recent interest in resequencing and improving the quality of these assemblies, at the time of this writing there is still quite a ways to go in order to have genomes of a sufficiently rigorous quality to make stronger inferences about gene copy number expansions and contractions (which were not considered in this study for this reason).
+
+The lack of a stronger signal from tumor suppressor duplications is likely a result of the strong effect size on both cancer risk and organismal toxicity that a TSG duplication would provide.
+The duplication of a tumor suppressor in many cases is associated with mild toxicity, although it greatly varies given the context and TSG in question; however, a single TSG duplication can also provide significant protection against cancer.
+For example, the overexpression of TP53 in mice, while protective of cancer, is associated with progeria and early death; however, if an additional copy of TP53 is introduced with its regulatory elements intact, the mice are healthy and experience normaging, while also demonstrating an enhanced response to cellular stress and lower rates of cancer.
+{[}60,61{]}.
+In light of this, it is fascinating that our results in the elephant lineage suggest that the duplication of TP53 regulators preceded the retroduplication and expansion of TP53, as this likely would have lowered the toxicity of the initial duplication and thus enabled it to occur.
+Given a sufficient selective pressure on increasing body size, it stands to reason that events like this could alleviate the negative pleiotropy of TSG duplications sufficiently to enable their persistence and allow for subsequent refinement over evolutionary time.
+
+By combining a phylogenetic study on body size in addition to a survey of copy number across nearly all protein-coding genes, we provide a comprehensive look at the question of the role of cancer risk and body size in \emph{Atlantogenata} that may provide broader insight to other mammalian species.
+Our study was initially motivated by the identification of functional duplicates of tumor suppressors, such as TP53 and LIF in elephants {[}13,22,23{]}.
+Further in support of our results, a larger candidate gene study by Caulin et al.
+{[}24{]} characterized the copy number of 830 known tumor-suppressor genes across 36 mammals and identified 382 putative duplicates, including duplicates in species with large body sizes and long life-spans.
+However, while candidate gene studies are useful, by their very design they are biased in determining larger patterns of evolution of traits.
+Without addressing these questions with a genome-wide approach, any and all insights will be inevitably limited to a fraction of the whole story.
+
+Our results suggest that the pervasive duplication of tumor suppressors may help enable the evolution of larger body sizes by lowering the cancer risk of species, either prior to or in lockstep with increasing body size.
+however, this is unlikely to be the only genetic mechanism at play in this scenario.
+In genome-wide studies of unusually large or long-lived species such as the bowhead whale {[}62{]}, Myotid bats {[}63,64{]}, naked mole rat {[}65{]}, and blind mole rat {[}66{]}, there were cases of overrepresentation of TSGs among duplicate genes that were outshadowed by the identification of strong signatures of positive selection at TSGs.
+While the evolution of regulatory and coding elements of both TSGs and other non-canonical tumor suppressor genes have been shown to be important for mediating the cancer risk of long-lived species, there has been no attention given to the possibility of TSG duplications providing a relaxation of possible negative pleitropy that could result from these traits.
+It has been well-established in the literature that genes duplication events allow for evolutionary drift in one of the copies, which may result in neofunctionalization or specialization of the two copies {[}67--69{]}.
+While this is beyond the scope of our study, a promising future direction of this work would include an evolutionary analysis of duplicated genes relative to each other to see if this has already occurred between the pairs of genes we have identified.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for his indispensable support.
+
+\hypertarget{conflicts-of-interest}{%
+\section{Conflicts of Interest}\label{conflicts-of-interest}}
+
+The Authors have no conflicts of interest to report
+
+\hypertarget{funding-source}{%
+\section{Funding Source}\label{funding-source}}
+
+We would like to thank the Department of Human Genetics at the University of Chicago for supporting this project.
+
+\hypertarget{figures}{%
+\section{FIGURES}\label{figures}}
 
@@
 
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure1-1} \hfill{}
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-intro-1} \hfill{}
+
+\caption{Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]}\label{fig:Figure-intro}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-BodySize-1} \hfill{}
 
@@
 Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size.
-**C)** Confidence interval plot for representative species and ancestral nodes.}\label{fig:Figure1}
+**C)** Confidence interval plot for representative species and ancestral nodes.}\label{fig:Figure-BodySize}
 \end{figure}
 
-To trace the evolutionary history of body mass and lifespan in
-Afrotherians, we built a time-calibrated supertree of Eutherian mammals
-combining 1,679 species from Bininda-Emonds et al {[}25{]} with a total
-evidence Afrotherian phylogeny including 77 extant and fossil data from
-39 extinct species {[}20{]}. Fossil data from extinct species were
-included to ensure that ancestral state reconstructions of body mass in
-Afrotherians were not biased by only including extant species, which can
-lead to inaccurate reconstructions, for example, if lineages multiple
-lineages evolved large body masses from a small bodied ancestor. We
-jointly estimated rates of body mass evolution and reconstructed
-ancestral states using a generalization of a Brownian model of character
-evolution, which allows for occasional large jumps in traits (stable
-model) and out performs standard Brownian motion and Ornstein-Uhlenbeck
-models of character evolution {[}26{]}.
+\begin{figure}
 
-Similar to previous studies of Afrotherian body size {[}20,26{]}, we
-found that the body mass of the Afrotherian ancestor was inferred to be
-small (0.26kg, 95\% CI: 0.31-3.01kg) and that substantial accelerations
-in the rate of body mass evolution occurred coincident with a 67.36x
-increase in body mass in the stem-lineage of \emph{Pseudoungulata}
-(17.33kg), a 1.45x increase in body mass in the stem-lineage of
-\emph{Paenungulata} (25.08kg), a 11.82x increase in body mass in the
-stem-lineage of \emph{Tehthytheria} (296.56kg), and a 2.69x increase in
-body mass in the stem-lineage of \emph{Proboscidea} (4114.39kg) (Figure
-1B,C). The ancestral \emph{Hyracoidea} was inferred to be relatively
-small 2.86kg-118.18kg, and rate accelerations were coincident with
-independent body mass increases in large hyraxes such as
-\emph{Titanohyrax andrewsi} 429.34kg; 67.36x increase. While the body
-mass of the ancestral \emph{Sirenian} was inferred to be large
-61.7kg-955.51kg, a rate acceleration occurred coincident with a 10.59x
-increase in body mass in Stellar's sea cow. Rate accelerations also
-occurred coincident with 36.6x decrease in body mass in the stem-lineage
-of the dwarf elephants \emph{Elephas (Palaeoloxodon) antiquus falconeri}
-and \emph{Elephas cypriotes}. These data suggest that gigantism in
-\emph{Afrotherians} evolved step-wise, from small to medium bodies in
-the \emph{Pseudoungulata} stem-lineage, medium to large bodies in the
-\emph{Tehthytherian} stem-lineage and extinct hyraxes, and from large to
-exceptionally large bodies independently in the \emph{Proboscidean}
-stem-lineage and Stellar's sea cow (Figure 1).
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-CancerSuscept-1} \hfill{}
 
-\hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
-\subsection*{Step-wise reduction of intrinsic cancer risk in large,
-long-lived
-Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}
-\addcontentsline{toc}{subsection}{Step-wise reduction of intrinsic
-cancer risk in large, long-lived Afrotherians}
+\caption{Cancer succecptibility across Atlantogenata.
+Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.}\label{fig:Figure-CancerSuscept}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics[width=6in,height=7in]{paper_PLOS_draft_files/figure-latex/Figure-GeneDup-Cladogram-1} \hfill{}
+
+\caption{Gene duplications occur readily throughout Atlantogenata.
+Shown here is a tree of Atlantogenatan species with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.}\label{fig:Figure-GeneDup-Cladogram}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=6in,height=6in]{paper_PLOS_draft_files/figure-latex/Figure-GenomeCor-1}
+
+}
+
+\caption{Correlations between genome quality metrics and ECNC metrics.
+Gene copy number metrics, and the genome quality metrics most strongly associated with them, are highlighted in red.}\label{fig:Figure-GenomeCor}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics[width=6in,height=4in]{paper_PLOS_draft_files/figure-latex/Figure-loxAfrORA-1} \hfill{}
+
+\caption{Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.}\label{fig:Figure-loxAfrORA}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics[width=6in,height=6in]{paper_PLOS_draft_files/figure-latex/Figure-loxAfrTP53-1} \hfill{}
+
+\end{figure}
+
+\hypertarget{tables}{%
+\section{TABLES}\label{tables}}
+
+\begin{table}
+
+\caption{\label{tab:Table-CancerSuscept}Estimated Cancer Succeptibilty for nodes in Atlantogenata}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrlllr}
+\toprule
+Node & Est. Lifespan & K1 & K2 & Change in K & log2 Change\\
+\midrule
+Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
+Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 & 4.07\\
+Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
+Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+\addlinespace
+Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 & 27.66\\
+Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
+Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 & 4.14\\
+Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
+\addlinespace
+Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 & -23.05\\
+Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 & 29.97\\
+Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 & 6.92\\
+Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 & 8.62\\
+\addlinespace
+Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 & 0.00\\
+Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 & -5.01\\
+Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 & 9.22\\
+Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 & 5.38\\
+Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
+\addlinespace
+Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
+Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 & 0.00\\
+Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 & 5.22\\
+Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 & 9.22\\
+Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 & 20.75\\
+\addlinespace
+Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 & -3.76\\
+Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 & 1.51\\
+Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 & 0.00\\
+Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 & 0.00\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-GeneDup}Summary of duplications in Atlantogenata}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lllrrllr}
+\toprule
+Species & Common Name & Size (g) & \#Hits & \#Duplicated & \% Genes Found & \% Hits Duplicated & Mean ECNC/Hit\\
+\midrule
+Choloepus hoffmanni & Hoffmans Two-Toed Sloth & 4.3e+03 & 14082 & 3204 & 78.19\% & 22.75\% & 0.98\\
+Chrysochloris asiatica & Cape Golden Mole & 49 & 13547 & 2716 & 75.22\% & 20.05\% & 0.99\\
+Dasypus novemcinctus & Nine-Banded Armadillo & 4.8e+03 & 13819 & 2605 & 76.73\% & 18.85\% & 0.98\\
+Echinops telfairi & Lesser Hedgehog Tenrec & 87 & 12903 & 1670 & 71.64\% & 12.94\% & 0.99\\
+Elephantulus edwardii & Cape Elephant Shrew & 49 & 12884 & 3048 & 71.53\% & 23.66\% & 0.99\\
+\addlinespace
+Elephas maximus & Asian Elephant & 3.3e+06 & 14073 & 907 & 78.14\% & 6.44\% & 1.00\\
+Loxodonta africana & African Savanna Elephant & 4.6e+06 & 14051 & 940 & 78.01\% & 6.69\% & 1.00\\
+Loxodonta cyclotis & African Forest Elephant & 4.7e+06 & 14065 & 900 & 78.09\% & 6.40\% & 1.00\\
+Mammut americanum & American Mastodon & 6e+06 & 13840 & 737 & 76.84\% & 5.33\% & 1.00\\
+Mammuthus columbi & Columbian Mammoth & 6.6e+06 & 13059 & 426 & 72.51\% & 3.26\% & 1.00\\
+\addlinespace
+Mammuthus primigenius & Woolly Mammoth & 4.3e+06 & 13935 & 723 & 77.37\% & 5.19\% & 1.00\\
+Orycteropus afer & Aardvark & 5.3e+04 & 13880 & 1083 & 77.06\% & 7.80\% & 0.99\\
+Palaeoloxodon antiquus & Straight Tusked Elephant & 1e+07 & 13969 & 745 & 77.56\% & 5.33\% & 1.00\\
+Procavia capensis & Rock Hyrax & 3e+03 & 13672 & 788 & 75.91\% & 5.76\% & 1.00\\
+Trichechus manatus & Manatee & 6.9e+05 & 14092 & 1046 & 78.24\% & 7.42\% & 1.00\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-ORAFreqDatabase}Number of pathways overrepresented among duplicated genes at different FDRs.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{llrrrr}
+\toprule
+Ancestor & Node & Pathways at FDR<0.1 & Pathways at FDR<0.2 & Pathways at FDR<0.3 & Pathways at FDR<0.5\\
+\midrule
+Afroinsectivora & Elephantulus edwardii & 252 & 37 & 30 & 87\\
+Afrosoricida & Chrysochloris asiatica & 90 & 48 & 43 & 105\\
+Afrosoricida & Echinops telfairi & 0 & 2 & 0 & 31\\
+Afrotheria & Afroinsectivora & 0 & 0 & 0 & 33\\
+Loxodontini & Loxodonta africana & 6 & 0 & 1 & 0\\
+\addlinespace
+Paenungulata & Tethytheria & 0 & 0 & 0 & 2\\
+Proboscidea & Mammut americanum & 0 & 0 & 0 & 6\\
+Pseudoungulata & Orycteropus afer & 27 & 67 & 29 & 67\\
+Tethytheria & Proboscidea & 0 & 3 & 0 & 3\\
+Tethytheria & Trichechus manatus & 4 & 0 & 0 & 2\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-SRAData}NCBI SRA datasets used in this study, along with key biological and genome information.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lll>{\raggedright\arraybackslash}p{50em}>{\raggedright\arraybackslash}p{50em}}
+\toprule
+Organism & Common Name & Genome & SRA Acc. & Tissues\\
+\midrule
+Dasypus novemcinctus & Nine-banded armadillo & dasNov3 & SRR494779, SRR494767, SRR494780, SRR494770, SRR309130, SRR494771, SRR4043756, SRR494776, SRR494778, SRR4043762, SRR4043755, SRR6206923, SRR4043761, SRR4043760, SRR6206913, SRR4043763, SRR494772, SRR494781, SRR494774, SRR494777, SRR494775, SRR4043754, SRR1289524, SRR4043758, SRR6206903, SRR1289523, SRR4043759, SRR3222425, SRR494768, SRR494769, SRR6206908, SRR4043757, SRR494766, SRR6206918, SRR494773 & Kidney, Spleen, Cerebellum W/ Brainstem, Rt.
+Quadricep, Mid-Stage Pregnant Endometrium, Cervix, Lung, Liver, Skeletal Muscle, Ascending Colon, Pregnant Armadillo Endometrium, Heart, Placenta\\
+Loxodonta africana & African savanna elephant & loxAfr3, loxAfrC, loxAfr4 & SRR6307198, SRR1041765, SRR6307199, SRR6307201, SRR6307196, SRR6307202, SRR6307200, SRR6307195, SRR975188, SRR6307194, SRR6307204, SRR3222430, SRR6307205, SRR975189, SRR6307197, SRR6307203 & Blood, Fibroblast, Placenta\\
+Trichechus manatus latirostris & Manatee & triMan1, triManLat2 & SRR4228542, SRR4228545, SRR4228544, SRR4228539, SRR4228541, SRR4228538, SRR4228546, SRR4228537, SRR4228540, SRR4228543, SRR4228547 & Buffy Coat\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{longtable}[t]{lrrrr}
+\caption{\label{tab:Table-BodySize}Body Size and Confidence Intervals in Atlantogenata estimated using StableTraits.}\\
+\toprule
+Node & Size (log(g)) & 95\% CI (Low) & 95\% CI (High) & Rate (sqrt)\\
+\midrule
+\endfirsthead
+\caption[]{\label{tab:Table-BodySize}Body Size and Confidence Intervals in Atlantogenata estimated using StableTraits. \textit{(continued)}}\\
+\toprule
+Node & Size (log(g)) & 95\% CI (Low) & 95\% CI (High) & Rate (sqrt)\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
+Cryptochloris wintoni & 3.13 & 3.13 & 3.13 & 5.78\\
+Amblysomus marleyi & 3.53 & 3.53 & 3.53 & 3.79\\
+Elephantulus revoili & 3.48 & 3.48 & 3.48 & 1.10\\
+Titanohyrax andrewsi & 12.97 & 12.97 & 12.97 & 0.07\\
+Titanohyrax ultimus & 14.08 & 14.08 & 14.08 & 34.61\\
+\addlinespace
+Megalohyrax sp  nov & 12.52 & 12.52 & 12.52 & 7.21\\
+Elephas maximus asurus & 15.66 & 15.66 & 15.66 & 0.34\\
+Protenrec tricuspis & 1.14 & 1.14 & 1.14 & 69.75\\
+Microgale parvula & 1.16 & 1.16 & 1.16 & 33.46\\
+Microgale pusilla & 1.25 & 1.25 & 1.25 & 34.31\\
+\addlinespace
+Geogale aurita & 1.90 & 1.90 & 1.90 & 40.07\\
+Microgale longicaudata & 2.09 & 2.09 & 2.09 & 0.77\\
+Microgale brevicaudata & 2.19 & 2.19 & 2.19 & 0.60\\
+Microgale jobihely & 2.30 & 2.30 & 2.30 & 1.07\\
+Microgale principula & 2.32 & 2.32 & 2.32 & 0.17\\
+\addlinespace
+Dilambdogale gheerbranti & 2.38 & 2.38 & 2.38 & 2.21\\
+Microgale taiva & 2.47 & 2.47 & 2.47 & 0.13\\
+Microgale cowani & 2.62 & 2.62 & 2.62 & 0.57\\
+Eremitalpa granti & 3.14 & 3.14 & 3.14 & 9.65\\
+Calcochloris obtusirostris & 3.27 & 3.27 & 3.27 & 13.38\\
+\addlinespace
+Neamblysomus julianae & 3.33 & 3.33 & 3.33 & 5.72\\
+Chlorotalpa duthieae & 3.38 & 3.38 & 3.38 & 0.32\\
+Chlorotalpa sclateri & 3.54 & 3.54 & 3.54 & 0.09\\
+Macroscelides proboscideus & 3.64 & 3.64 & 3.64 & 14.17\\
+Chrysochloris stuhlmanni & 3.74 & 3.74 & 3.74 & 0.33\\
+\addlinespace
+Oryzorictes hova & 3.79 & 3.79 & 3.79 & 22.77\\
+Elephantulus myurus & 3.81 & 3.81 & 3.81 & 0.95\\
+Elephantulus brachyrhynchus & 3.81 & 3.81 & 3.81 & 0.93\\
+Elephantulus rozeti & 3.81 & 3.81 & 3.81 & 10.51\\
+Elephantulus fuscus & 3.82 & 3.82 & 3.82 & 0.68\\
+\addlinespace
+Elephantulus intufi & 3.82 & 3.82 & 3.82 & 1.15\\
+Microgale talazaci & 3.88 & 3.88 & 3.88 & 61.40\\
+Chrysochloris asiatica & 3.89 & 3.89 & 3.89 & 3.34\\
+Elephantulus edwardii & 3.90 & 3.90 & 3.90 & 0.24\\
+Carpitalpa arendsi & 3.94 & 3.94 & 3.94 & 0.45\\
+\addlinespace
+Amblysomus corriae & 3.94 & 3.94 & 3.94 & 0.98\\
+Amblysomus hottentotus & 3.98 & 3.98 & 3.98 & 0.02\\
+Elephantulus fuscipes & 4.04 & 4.04 & 4.04 & 1.93\\
+Elephantulus rufescens & 4.05 & 4.05 & 4.05 & 0.12\\
+Neamblysomus gunningi & 4.09 & 4.09 & 4.09 & 3.26\\
+\addlinespace
+Elephantulus rupestris & 4.12 & 4.12 & 4.12 & 0.32\\
+Amblysomus septentrionalis & 4.23 & 4.23 & 4.23 & 0.52\\
+Chambius kasserinensis & 4.27 & 4.27 & 4.27 & 11.84\\
+Amblysomus robustus & 4.33 & 4.33 & 4.33 & 1.38\\
+Micropotamogale lamottei & 4.36 & 4.36 & 4.36 & 2.82\\
+\addlinespace
+Echinops telfairi & 4.47 & 4.47 & 4.47 & 7.75\\
+Limnogale mergulus & 4.52 & 4.52 & 4.52 & 121.95\\
+Hemicentetes semispinosus & 4.75 & 4.75 & 4.75 & 4.68\\
+Chrysospalax villosus & 4.77 & 4.77 & 4.77 & 0.13\\
+Petrodromus tetradactylus & 5.29 & 5.29 & 5.29 & 24.61\\
+\addlinespace
+Herodotius pattersoni & 5.50 & 5.50 & 5.50 & 11.64\\
+Setifer setosus & 5.61 & 5.61 & 5.61 & 12.52\\
+Rhynchocyon cirnei & 5.86 & 5.86 & 5.86 & 3.30\\
+Metoldobotes sp  nov & 5.93 & 5.93 & 5.93 & 15.94\\
+Chrysospalax trevelyani & 6.13 & 6.13 & 6.13 & 62.84\\
+\addlinespace
+Rhynchocyon petersi & 6.15 & 6.15 & 6.15 & 2.13\\
+Rhynchocyon chrysopygus & 6.28 & 6.28 & 6.28 & 0.40\\
+Potamogale velox & 6.49 & 6.49 & 6.49 & 103.04\\
+Rhynchocyon udzungwensis & 6.57 & 6.57 & 6.57 & 4.33\\
+Tenrec ecaudatus & 6.75 & 6.75 & 6.75 & 79.50\\
+\addlinespace
+Dasypus sabanicola & 7.05 & 7.05 & 7.05 & 12.18\\
+Tolypeutes matacus & 7.11 & 7.11 & 7.11 & 15.96\\
+Dasypus septemcinctus & 7.30 & 7.30 & 7.30 & 4.44\\
+Zaedyus pichiy & 7.31 & 7.31 & 7.31 & 5.54\\
+Dasypus hybridus & 7.31 & 7.31 & 7.31 & 4.05\\
+\addlinespace
+Chaetophractus villosus & 7.61 & 7.61 & 7.61 & 0.42\\
+Chaetophractus nationi & 7.67 & 7.67 & 7.67 & 0.09\\
+Heterohyrax brucei & 7.78 & 7.78 & 7.78 & 1.64\\
+Cabassous centralis & 7.92 & 7.92 & 7.92 & 0.25\\
+Seggeurius amourensis & 7.98 & 7.98 & 7.98 & 2.82\\
+\addlinespace
+Procavia capensis & 8.01 & 8.01 & 8.01 & 0.00\\
+Dendrohyrax dorsalis & 8.06 & 8.06 & 8.06 & 1.86\\
+Microhyrax lavocati & 8.13 & 8.13 & 8.13 & 0.73\\
+Bradypus tridactylus & 8.23 & 8.23 & 8.23 & 0.48\\
+Bradypus torquatus & 8.27 & 8.27 & 8.27 & 0.03\\
+\addlinespace
+Dasypus novemcinctus & 8.37 & 8.37 & 8.37 & 14.73\\
+Euphractus sexcinctus & 8.43 & 8.43 & 8.43 & 14.99\\
+Choloepus hoffmanni & 8.47 & 8.47 & 8.47 & 0.32\\
+Bradypus variegatus & 8.49 & 8.49 & 8.49 & 0.51\\
+Tamandua tetradactyla & 8.52 & 8.52 & 8.52 & 10.44\\
+\addlinespace
+Cyclopes didactylus & 8.53 & 8.53 & 8.53 & 2.15\\
+Choloepus didactylus & 8.71 & 8.71 & 8.71 & 0.64\\
+Thyrohyrax meyeri & 8.78 & 8.78 & 8.78 & 3.55\\
+Saghatherium bowni & 9.13 & 9.13 & 9.13 & 15.85\\
+Dasypus kappleri & 9.23 & 9.23 & 9.23 & 74.13\\
+\addlinespace
+Thyrohyrax domorictus & 9.30 & 9.30 & 9.30 & 1.15\\
+Dimaitherium patnaiki & 9.57 & 9.57 & 9.57 & 18.23\\
+Phosphatherium escuilliei & 9.62 & 9.62 & 9.62 & 326.23\\
+Saghatherium antiquum & 9.73 & 9.73 & 9.73 & 2.90\\
+Thyrohyrax litholagus & 10.01 & 10.01 & 10.01 & 28.58\\
+\addlinespace
+Myrmecophaga tridactyla & 10.26 & 10.26 & 10.26 & 41.03\\
+Myorycteropus africanus & 10.27 & 10.27 & 10.27 & 0.57\\
+Selenohyrax chatrathi & 10.73 & 10.73 & 10.73 & 14.99\\
+Priodontes maximus & 10.82 & 10.82 & 10.82 & 268.43\\
+Orycteropus afer & 10.87 & 10.87 & 10.87 & 6.59\\
+\addlinespace
+Antilohyrax pectidens & 10.93 & 10.93 & 10.93 & 13.69\\
+Bunohyrax fajumensis & 11.32 & 11.32 & 11.32 & 1.45\\
+Afrohyrax championi & 11.32 & 11.32 & 11.32 & 0.19\\
+Geniohyus mirus & 11.33 & 11.33 & 11.33 & 5.44\\
+Prorastomus sirenoides & 11.49 & 11.49 & 11.49 & 13.61\\
+\addlinespace
+Elephas antiquus falconeri & 11.51 & 11.51 & 11.51 & 6.12\\
+Pachyhyrax crassidentatus & 11.81 & 11.81 & 11.81 & 2.29\\
+Megalohyrax eocaenus & 11.95 & 11.95 & 11.95 & 0.24\\
+Elephas cypriotes & 12.21 & 12.21 & 12.21 & 1.90\\
+Bunohyrax major & 12.36 & 12.36 & 12.36 & 11.39\\
+\addlinespace
+Titanohyrax angustidens & 12.48 & 12.48 & 12.48 & 0.04\\
+Daouitherium rebouli & 12.80 & 12.80 & 12.80 & 0.74\\
+Arcanotherium savagei & 12.89 & 12.89 & 12.89 & 7.29\\
+Dugong dugon & 12.92 & 12.92 & 12.92 & 5.85\\
+Trichechus senegalensis & 13.03 & 13.03 & 13.03 & 0.57\\
+\addlinespace
+Trichechus inunguis & 13.08 & 13.08 & 13.08 & 0.69\\
+Protosiren smithae & 13.20 & 13.20 & 13.20 & 33.69\\
+Numidotherium koholense & 13.23 & 13.23 & 13.23 & 2.29\\
+Omanitherium dhofarensis & 13.35 & 13.35 & 13.35 & 0.03\\
+Trichechus manatus & 13.44 & 13.44 & 13.44 & 1.39\\
+\addlinespace
+Moeritherium spp & 13.82 & 13.82 & 13.82 & 5.71\\
+Phiomia spp & 13.89 & 13.89 & 13.89 & 3.64\\
+Elephas maximus & 15.02 & 15.02 & 15.02 & 5.81\\
+Barytherium spp & 15.20 & 15.20 & 15.20 & 73.58\\
+Mammuthus primigenius & 15.27 & 15.27 & 15.27 & 2.17\\
+\addlinespace
+Mammut borsoni & 16.49 & 16.49 & 16.49 & 15.33\\
+Mammuthus trogontherii & 16.38 & 16.38 & 16.38 & 16.00\\
+Loxodonta africana & 15.35 & 15.35 & 15.35 & 1.28\\
+Loxodonta cyclotis & 15.37 & 15.37 & 15.37 & 3.72\\
+Palaeoloxodon antiquus & 16.14 & 16.14 & 16.14 & 0.01\\
+\addlinespace
+Palaeoloxodon namadicus & 16.81 & 16.81 & 16.81 & 12.81\\
+Mammut americanum & 15.61 & 15.61 & 15.61 & 0.95\\
+Mammuthus columbi & 15.71 & 15.71 & 15.71 & 0.91\\
+Hydrodamalis gigas & 15.72 & 15.72 & 15.72 & 172.52\\
+Atlantogenata & 5.55 & 4.06 & 7.95 & 0.03\\
+\addlinespace
+Afrotheria & 5.55 & 4.05 & 7.96 & 0.00\\
+Afrosoricida & 4.35 & 2.58 & 6.13 & 44.49\\
+Macroscelidae & 5.27 & 3.98 & 6.85 & 2.49\\
+Pseudoungulata & 9.76 & 5.21 & 12.78 & 545.83\\
+Paenungulata & 10.13 & 7.24 & 13.02 & 4.42\\
+\addlinespace
+Tethytheria & 12.60 & 10.25 & 13.81 & 187.47\\
+Proboscidea & 15.23 & 14.22 & 16.24 & 30.28\\
+Elephantidae & 15.49 & 14.89 & 16.10 & 2.21\\
+Elephantina & 15.51 & 15.08 & 15.96 & 0.01\\
+Mammuthus & 15.54 & 15.24 & 15.85 & 0.47\\
+\addlinespace
+Loxodontini & 15.55 & 15.02 & 16.11 & 0.11\\
+Loxodona & 15.72 & 15.16 & 16.30 & 0.86\\
+Xenarthra & 7.57 & 5.96 & 9.18 & 124.94\\*
+\end{longtable}
+
+\begin{table}
+
+\caption{\label{tab:Table-Genomes}Genomes used in this study.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lllll}
+\toprule
+Species & Common Name & Genomes & Highest Quality Genome & Citation\\
+\midrule
+Choloepus hoffmanni & Hoffmans two-toed sloth & choHof1, choHof2, choHof-C\_hoffmanni-2.0.1\_HiC & choHof-C\_hoffmanni-2.0.1\_HiC & NA\\
+Chrysochloris asiatica & Cape golden mole & chrAsi1m & chrAsi1m & NA\\
+Dasypus novemcinctus & Nine-banded armadillo & dasNov3 & dasNov3 & NA\\
+Echinops telfairi & Lesser Hedgehog Tenrec & echTel2 & echTel2 & NA\\
+Elephantulus edwardii & Cape elephant shrew & eleEdw1m & eleEdw1m & NA\\
+\addlinespace
+Elephas maximus & Asian elephant & eleMaxD & eleMaxD & NA\\
+Loxodonta africana & African savanna elephant & loxAfr3, loxAfrC, loxAfr4 & loxAfr4 & NA\\
+Loxodonta cyclotis & African forest elephant & loxCycF & loxCycF & NA\\
+Mammut americanum & American mastodon & mamAmeI & mamAmeI & NA\\
+Mammuthus columbi & Columbian mammoth & mamColU & mamColU & NA\\
+\addlinespace
+Mammuthus primigenius & Woolly mammoth & mamPriV & mamPriV & NA\\
+Orycteropus afer & Aardvark & oryAfe1, oryAfe2 & oryAfe2 & NA\\
+Palaeoloxodon antiquus & Straight tusked elephant & palAntN & palAntN & NA\\
+Procavia capensis & Rock hyrax & proCap1, proCap2, proCap-Pcap\_2.0\_HiC & proCap-Pcap\_2.0\_HiC & NA\\
+Trichechus manatus latirostris & Manatee & triMan1, triManLat2 & triManLat2 & NA\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\hypertarget{supplementary-figures}{%
+\section{SUPPLEMENTARY FIGURES}\label{supplementary-figures}}
+
+\textbackslash{}begin\{figure\}
+
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-ECNC-1} \hfill{}
+
+\textbackslash{}caption\{Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
+\textbf{A)} A single, contiguous gene homolog in a target genome with 100\% query length coverage has an ECNC of 1.0.
+\textbf{B)} Two contiguous gene homologs, each with 100\% query length coverage have an ECNC of 2.0.
+\textbf{C)} A single gene homolog, split across multiple scaffolds and contigs in a fragmented target genome; BLAT identifies each fragment as a single hit.
+Per nucleotide of query sequence, there is only one corresponding nucleotide over all the hits, thus the ECNC is 1.0.
+\textbf{D)} Two gene homologs, one fragmented and one contiguous.
+100\% of nucleotides in the query sequence are represented between all hits; however, every nucleotide in the query has two matching nucleotides in the target genome, thus the ECNC is 2.0.
+\textbf{E)} One true gene homolog in the target genome, plus multiple hits of a conserved domain that span 20\% of the query sequence.
+While 100\% of the query sequence is represented in total, 20\% of the nucleotides have 4 hits.
+Thus, the ECNC for this gene is 1.45.
+\textbf{F)} Two real gene homologs; one hit is contiguous, one hit is fragmented in two, and the tail end of both sequences was not identified by BLAT due to sequence divergence.
+Only 75\% of the query sequence was covered in total between the hits, but for that 75\%, each nucleotide has two hits.
+As such, ECNC is equal to 2.0 for this gene.\}\label{fig:Figure-ECNC}
+\textbackslash{}end\{figure\}
+
+\begin{figure}
+
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-GeneDup-BLScaled-1} \hfill{}
+
+\caption{Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number}\label{fig:Figure-GeneDup-BLScaled}
+\end{figure}
+
+\begin{verbatim}
+## Warning: TODO: Make HQ version of this figure
+\end{verbatim}
+
+\begin{figure}
+
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-RBHBStrategy-1} \hfill{}
+
+\caption{Full version of RecBlat strategy, but low quality}\label{fig:Figure-RBHBStrategy}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-TimeTree-Eutheria-1} \hfill{}
+
+\caption{The time-calibrated Eutherian tree used as input for StableTraits, from Bininda-Emonds et al and Puttick and Thomas}\label{fig:Figure-TimeTree-Eutheria}
+\end{figure}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Savage2007}{}%
+1.
+Savage VM, Allen AP, Brown JH, Gillooly JF, Herman AB, Woodruff WH, et al.
+Scaling of number, size, and metabolic rate of cells with body size in mammals.
+Proceedings of the National Academy of Sciences.
+2007;104: 4718--4723.
+doi:\href{https://doi.org/10.1073/pnas.0611235104}{10.1073/pnas.0611235104}
+
 \leavevmode\hypertarget{ref-Green2011}{}%
-1. Green J, Cairns BJ, Casabonne D, Wright FL, Reeves G, Beral V, et al.
-Height and cancer incidence in the Million Women Study: prospective
-cohort, and meta-analysis of prospective studies of height and total
-cancer risk. The Lancet Oncology. 2011;12: 785--794.
+2.
+Green J, Cairns BJ, Casabonne D, Wright FL, Reeves G, Beral V, et al.
+Height and cancer incidence in the Million Women Study: prospective cohort, and meta-analysis of prospective studies of height and total cancer risk.
+The Lancet Oncology.
+2011;12: 785--794.
 doi:\href{https://doi.org/10.1016/s1470-2045(11)70154-1}{10.1016/s1470-2045(11)70154-1}
@@
 \leavevmode\hypertarget{ref-Nunney:20181c2}{}%
-2. Nunney L. Size matters: height, cell number and a person's risk of
-cancer. Proc R Soc B. 2018;285: 20181743.
-doi:\href{https://doi.org/10.1098/rspb.2018.1743}{10.1098/rspb.2018.1743}
+3. Nunney L. Size matters: height, cell number and a person's risk of cancer. Proc R Soc B. 2018;285: 20181743. doi:\href{https://doi.org/10.1098/rspb.2018.1743}{10.1098/rspb.2018.1743}
 
 \leavevmode\hypertarget{ref-Dobson2013}{}%
-3. Dobson JM. Breed-predispositions to cancer in pedigree dogs. ISRN
-veterinary science. 2013;2013: 941275.
-doi:\href{https://doi.org/10.1155/2013/941275}{10.1155/2013/941275}
+4. Dobson JM. Breed-predispositions to cancer in pedigree dogs. ISRN veterinary science. 2013;2013: 941275. doi:\href{https://doi.org/10.1155/2013/941275}{10.1155/2013/941275}
 
 \leavevmode\hypertarget{ref-Dorn1968}{}%
-4. Dorn CR, Taylor DON, Schneider R, Hibbard HH, Klauber MR. Survey of
-Animal Neoplasms in Alameda and Contra Costa Counties, California. II.
-Cancer Morbidity in Dogs and Cats From Alameda County\textless{}xref
-ref-type="fn" rid="FN2"\textgreater{}2\textless{}/xref\textgreater{}.
-JNCI: Journal of the National Cancer Institute. 1968;40: 307--318.
+5.
+Dorn CR, Taylor DON, Schneider R, Hibbard HH, Klauber MR.
+Survey of Animal Neoplasms in Alameda and Contra Costa Counties, California.
+II.
+Cancer Morbidity in Dogs and Cats From Alameda County\textless{}xref ref-type="fn" rid="FN2"\textgreater{}2\textless{}/xref\textgreater{}.
+JNCI: Journal of the National Cancer Institute.
+1968;40: 307--318.
 doi:\href{https://doi.org/10.1093/jnci/40.2.307}{10.1093/jnci/40.2.307}
@@
 \leavevmode\hypertarget{ref-Caulin2011}{}%
-5. Caulin AF, Maley CC. Peto's Paradox: evolution's prescription for
-cancer prevention. Trends in ecology \& evolution. 2011;26: 175--82.
+6.
+Caulin AF, Maley CC.
+Peto's Paradox: evolution's prescription for cancer prevention.
+Trends in ecology \& evolution.
+2011;26: 175--82.
 doi:\href{https://doi.org/10.1016/j.tree.2011.01.002}{10.1016/j.tree.2011.01.002}
@@
 \leavevmode\hypertarget{ref-Leroi2003}{}%
-6. Leroi AM, Koufopanou V, Burt A. Cancer selection. Nature Reviews
-Cancer. 2003;3: 226--231.
-doi:\href{https://doi.org/10.1038/nrc1016}{10.1038/nrc1016}
+7. Leroi AM, Koufopanou V, Burt A. Cancer selection. Nature Reviews Cancer. 2003;3: 226--231. doi:\href{https://doi.org/10.1038/nrc1016}{10.1038/nrc1016}
 
 \leavevmode\hypertarget{ref-Peto1975}{}%
-7. Peto R, Roe F, Lee P, Levy L, Clack J. Cancer and ageing in mice and
-men. British Journal of Cancer. 1975;32: 411--426.
-doi:\href{https://doi.org/10.1038/bjc.1975.242}{10.1038/bjc.1975.242}
+8. Peto R, Roe F, Lee P, Levy L, Clack J. Cancer and ageing in mice and men. British Journal of Cancer. 1975;32: 411--426. doi:\href{https://doi.org/10.1038/bjc.1975.242}{10.1038/bjc.1975.242}
 
 \leavevmode\hypertarget{ref-Ashur-Fabian2004}{}%
-8. Ashur-Fabian O, Avivi A, Trakhtenbrot L, Adamsky K, Cohen M, Kajakaro
-G, et al. Evolution of p53 in hypoxia-stressed Spalax mimics human tumor
-mutation. Proceedings of the National Academy of Sciences. 2004;101:
-12236--12241.
+9.
+Ashur-Fabian O, Avivi A, Trakhtenbrot L, Adamsky K, Cohen M, Kajakaro G, et al.
+Evolution of p53 in hypoxia-stressed Spalax mimics human tumor mutation.
+Proceedings of the National Academy of Sciences.
+2004;101: 12236--12241.
 doi:\href{https://doi.org/10.1073/pnas.0404998101}{10.1073/pnas.0404998101}
@@
 \leavevmode\hypertarget{ref-Seluanov2008}{}%
-9. Seluanov A, Hine C, Bozzella M, Hall A, Sasahara THC, Ribeiro AACM,
-et al. Distinct tumor suppressor mechanisms evolve in rodent species
-that differ in size and lifespan. Aging cell. 2008;7: 813--23.
+10.
+Seluanov A, Hine C, Bozzella M, Hall A, Sasahara THC, Ribeiro AACM, et al.
+Distinct tumor suppressor mechanisms evolve in rodent species that differ in size and lifespan.
+Aging cell.
+2008;7: 813--23.
 doi:\href{https://doi.org/10.1111/j.1474-9726.2008.00431.x}{10.1111/j.1474-9726.2008.00431.x}
@@
 \leavevmode\hypertarget{ref-Gorbunova2012}{}%
-10. Gorbunova V, Hine C, Tian X, Ablaeva J, Gudkov AV, Nevo E, et al.
-Cancer resistance in the blind mole rat is mediated by concerted
-necrotic cell death mechanism. Proceedings of the National Academy of
-Sciences of the United States of America. 2012;109: 19392--6.
+11.
+Gorbunova V, Hine C, Tian X, Ablaeva J, Gudkov AV, Nevo E, et al.
+Cancer resistance in the blind mole rat is mediated by concerted necrotic cell death mechanism.
+Proceedings of the National Academy of Sciences of the United States of America.
+2012;109: 19392--6.
 doi:\href{https://doi.org/10.1073/pnas.1217211109}{10.1073/pnas.1217211109}
@@
 \leavevmode\hypertarget{ref-Tian2013}{}%
-11. Tian X, Azpurua J, Hine C, Vaidya A, Myakishev-Rempel M, Ablaeva J,
-et al. High molecular weight hyaluronan mediates the cancer resistance
-of the naked mole-rat. 2013;499.
+12.
+Tian X, Azpurua J, Hine C, Vaidya A, Myakishev-Rempel M, Ablaeva J, et al.
+High molecular weight hyaluronan mediates the cancer resistance of the naked mole-rat.
+2013;499.
 doi:\href{https://doi.org/10.1038/nature12234}{10.1038/nature12234}
@@
 \leavevmode\hypertarget{ref-Sulak2015}{}%
-12. Sulak M, Fong L, Mika K, Chigurupati S, Yon L, Mongan NP, et al.
-TP53 copy number expansion is associated with the evolution of increased
-body size and an enhanced DNA damage response in elephants. eLife.
+13.
+Sulak M, Fong L, Mika K, Chigurupati S, Yon L, Mongan NP, et al.
+TP53 copy number expansion is associated with the evolution of increased body size and an enhanced DNA damage response in elephants.
+eLife.
 2016;5: e11994.
@@
 \leavevmode\hypertarget{ref-HAGR}{}%
-13. Tacutu R, Craig T, Budovsky A, Wuttke D, Lehmann G, Taranukha D, et
-al. Human Ageing Genomic Resources: Integrated databases and tools for
-the biology and genetics of ageing. Nucleic Acids Research. 2013;41:
-D1027--D1033.
+14.
+Tacutu R, Craig T, Budovsky A, Wuttke D, Lehmann G, Taranukha D, et al.
+Human Ageing Genomic Resources: Integrated databases and tools for the biology and genetics of ageing.
+Nucleic Acids Research.
+2013;41: D1027--D1033.
 doi:\href{https://doi.org/10.1093/nar/gks1155}{10.1093/nar/gks1155}
@@
 \leavevmode\hypertarget{ref-Schwartz1995}{}%
-14. Schwartz GT, Rasmussen DT, Smith RJ. Body-Size Diversity and
-Community Structure of Fossil Hyracoids. Journal of Mammalogy. 1995;76:
-1088--1099. doi:\href{https://doi.org/10.2307/1382601}{10.2307/1382601}
+15.
+Schwartz GT, Rasmussen DT, Smith RJ.
+Body-Size Diversity and Community Structure of Fossil Hyracoids.
+Journal of Mammalogy.
+1995;76: 1088--1099.
+doi:\href{https://doi.org/10.2307/1382601}{10.2307/1382601}
 
 \leavevmode\hypertarget{ref-Scheffer1972}{}%
-15. Scheffer VB. The Weight of the Steller Sea Cow. Journal of
-Mammalogy. 1972;53: 912--914.
-doi:\href{https://doi.org/10.2307/1379236}{10.2307/1379236}
+16. Scheffer VB. The Weight of the Steller Sea Cow. Journal of Mammalogy. 1972;53: 912--914. doi:\href{https://doi.org/10.2307/1379236}{10.2307/1379236}
 
 \leavevmode\hypertarget{ref-Larramendi:20151c2}{}%
-16. Larramendi A. Shoulder Height, Body Mass, and Shape of
-Proboscideans. Acta Palaeontologica Polonica. 2015;61.
-doi:\href{https://doi.org/10.4202/app.00136.2014}{10.4202/app.00136.2014}
+17. Larramendi A. Shoulder Height, Body Mass, and Shape of Proboscideans. Acta Palaeontologica Polonica. 2015;61. doi:\href{https://doi.org/10.4202/app.00136.2014}{10.4202/app.00136.2014}
 
 \leavevmode\hypertarget{ref-OLeary2013a}{}%
-17. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
-NP, et al. The placental mammal ancestor and the post-K-Pg radiation of
-placentals. Science (New York, NY). 2013;339: 662--7.
+18.
+O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini NP, et al.
+The placental mammal ancestor and the post-K-Pg radiation of placentals.
+Science (New York, NY).
+2013;339: 662--7.
 doi:\href{https://doi.org/10.1126/science.1229237}{10.1126/science.1229237}
@@
 \leavevmode\hypertarget{ref-Springer2013}{}%
-18. Springer MS, Meredith RW, Teeling EC, Murphy WJ. Technical comment
-on "The placental mammal ancestor and the post-K-Pg radiation of
-placentals". Science (New York, NY). 2013;341: 613.
+19.
+Springer MS, Meredith RW, Teeling EC, Murphy WJ.
+Technical comment on "The placental mammal ancestor and the post-K-Pg radiation of placentals".
+Science (New York, NY).
+2013;341: 613.
 doi:\href{https://doi.org/10.1126/science.1238025}{10.1126/science.1238025}
@@
 \leavevmode\hypertarget{ref-OLeary2013b}{}%
-19. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
-NP, et al. Response to comment on "The placental mammal ancestor and the
-post-K-Pg radiation of placentals". Science (New York, NY). 2013;341:
-613.
+20.
+O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini NP, et al.
+Response to comment on "The placental mammal ancestor and the post-K-Pg radiation of placentals".
+Science (New York, NY).
+2013;341: 613.
 doi:\href{https://doi.org/10.1126/science.1238162}{10.1126/science.1238162}
@@
 \leavevmode\hypertarget{ref-PuttickAndThomas2015}{}%
-20. Puttick MN, Thomas GH. Fossils and living taxa agree on patterns of
-body mass evolution: a case study with Afrotheria. Proceedings
-Biological sciences / The Royal Society. 2015;282: 20152023.
+21.
+Puttick MN, Thomas GH.
+Fossils and living taxa agree on patterns of body mass evolution: a case study with Afrotheria.
+Proceedings Biological sciences / The Royal Society.
+2015;282: 20152023.
 doi:\href{https://doi.org/10.1098/rspb.2015.2023}{10.1098/rspb.2015.2023}
@@
 \leavevmode\hypertarget{ref-Abegglen:JAMA2015}{}%
-21. Abegglen LM, Caulin AF, Chan A, Lee K, Robinson R, Campbell MS, et
-al. Potential Mechanisms for Cancer Resistance in Elephants and
-Comparative Cellular Response to DNA Damage in Humans. JAMA. 2015;314:
-1850--1860.
+22.
+Abegglen LM, Caulin AF, Chan A, Lee K, Robinson R, Campbell MS, et al.
+Potential Mechanisms for Cancer Resistance in Elephants and Comparative Cellular Response to DNA Damage in Humans.
+JAMA.
+2015;314: 1850--1860.
 doi:\href{https://doi.org/10.1001/jama.2015.13134}{10.1001/jama.2015.13134}
@@
 \leavevmode\hypertarget{ref-Vazquez2018}{}%
-22. Vazquez JM, Sulak M, Chigurupati S, Lynch VJ. A Zombie LIF Gene in
-Elephants Is Upregulated by TP53 to Induce Apoptosis in Response to DNA
-Damage. Cell Reports. 2018;24: 1765--1776.
+23.
+Vazquez JM, Sulak M, Chigurupati S, Lynch VJ.
+A Zombie LIF Gene in Elephants Is Upregulated by TP53 to Induce Apoptosis in Response to DNA Damage.
+Cell Reports.
+2018;24: 1765--1776.
 doi:\href{https://doi.org/10.1016/j.celrep.2018.07.042}{10.1016/j.celrep.2018.07.042}
@@
 \leavevmode\hypertarget{ref-Caulin2015}{}%
-23. Caulin AF, Graham TA, Wang L-S, Maley CC. Solutions to Peto's
-paradox revealed by mathematical modelling and cross-species cancer gene
-analysis. Philosophical transactions of the Royal Society of London
-Series B, Biological sciences. 2015;370: 20140222.
+24.
+Caulin AF, Graham TA, Wang L-S, Maley CC.
+Solutions to Peto's paradox revealed by mathematical modelling and cross-species cancer gene analysis.
+Philosophical transactions of the Royal Society of London Series B, Biological sciences.
+2015;370: 20140222.
 doi:\href{https://doi.org/10.1098/rstb.2014.0222}{10.1098/rstb.2014.0222}
@@
 \leavevmode\hypertarget{ref-Doherty2016}{}%
-24. Doherty A, Magalhães J de. Has gene duplication impacted the
-evolution of Eutherian longevity? Aging Cell. 2016;15: 978--980.
-doi:\href{https://doi.org/10.1111/acel.12503}{10.1111/acel.12503}
+25. Doherty A, Magalhães J de. Has gene duplication impacted the evolution of Eutherian longevity? Aging Cell. 2016;15: 978--980. doi:\href{https://doi.org/10.1111/acel.12503}{10.1111/acel.12503}
 
 \leavevmode\hypertarget{ref-Bininda-Emonds2008}{}%
-25. Bininda-Emonds ORP, Cardillo M, Jones KE, MacPhee RDE, Beck RMD,
-Grenyer R, et al. Erratum: The delayed rise of present-day mammals.
-Nature. 2008;456: 274--274.
+26.
+Bininda-Emonds ORP, Cardillo M, Jones KE, MacPhee RDE, Beck RMD, Grenyer R, et al.
+Erratum: The delayed rise of present-day mammals.
+Nature.
+2008;456: 274--274.
 doi:\href{https://doi.org/10.1038/nature07347}{10.1038/nature07347}
@@
 \leavevmode\hypertarget{ref-ElliotAndMooers2014}{}%
-26. Elliot MG, Mooers AØ. Inferring ancestral states without assuming
-neutrality or gradualism using a stable model of continuous character
-evolution. BMC evolutionary biology. 2014;14: 226.
+27.
+Elliot MG, Mooers AØ.
+Inferring ancestral states without assuming neutrality or gradualism using a stable model of continuous character evolution.
+BMC evolutionary biology.
+2014;14: 226.
 doi:\href{https://doi.org/10.1186/s12862-014-0226-8}{10.1186/s12862-014-0226-8}
@@
 \leavevmode\hypertarget{ref-blat}{}%
-27. Kent JW. BLAT?The BLAST-Like Alignment Tool. Genome Research.
-2002;12: 656--664.
-doi:\href{https://doi.org/10.1101/gr.229202}{10.1101/gr.229202}
+28. Kent JW. BLAT?The BLAST-Like Alignment Tool. Genome Research. 2002;12: 656--664. doi:\href{https://doi.org/10.1101/gr.229202}{10.1101/gr.229202}
 
 \leavevmode\hypertarget{ref-AltenhoffAndDessimoz2009}{}%
-28. Altenhoff AM, Dessimoz C. Phylogenetic and functional assessment of
-orthologs inference projects and methods. PLoS computational biology.
+29.
+Altenhoff AM, Dessimoz C.
+Phylogenetic and functional assessment of orthologs inference projects and methods.
+PLoS computational biology.
 2009;5: e1000262.
@@
 \leavevmode\hypertarget{ref-SalichosAndRokas2011}{}%
-29. Salichos L, Rokas A. Evaluating ortholog prediction algorithms in a
-yeast model clade. PloS one. 2011;6: e18755.
+30.
+Salichos L, Rokas A.
+Evaluating ortholog prediction algorithms in a yeast model clade.
+PloS one.
+2011;6: e18755.
 doi:\href{https://doi.org/10.1371/journal.pone.0018755}{10.1371/journal.pone.0018755}
@@
 \leavevmode\hypertarget{ref-uniprot}{}%
-30. Consortium TU. UniProt: the universal protein knowledgebase. Nucleic
-Acids Research. 2017;45: D158--D169.
-doi:\href{https://doi.org/10.1093/nar/gkw1099}{10.1093/nar/gkw1099}
+31. Consortium TU. UniProt: the universal protein knowledgebase. Nucleic Acids Research. 2017;45: D158--D169. doi:\href{https://doi.org/10.1093/nar/gkw1099}{10.1093/nar/gkw1099}
 
 \leavevmode\hypertarget{ref-gVolante}{}%
-31. Nishimura O, Hara Y, Kuraku S. gVolante for standardizing
-completeness assessment of genome and transcriptome assemblies.
-Bioinformatics. 2017;33: 3635--3637.
+32.
+Nishimura O, Hara Y, Kuraku S.
+gVolante for standardizing completeness assessment of genome and transcriptome assemblies.
+Bioinformatics.
+2017;33: 3635--3637.
 doi:\href{https://doi.org/10.1093/bioinformatics/btx445}{10.1093/bioinformatics/btx445}
@@
 \leavevmode\hypertarget{ref-CEGMA}{}%
-32. Parra G, Bradnam K, Ning Z, Keane T, Korf I. Assessing the gene
-space in draft genomes. Nucleic Acids Research. 2008;37: 289--297.
-doi:\href{https://doi.org/10.1093/nar/gkn916}{10.1093/nar/gkn916}
+33. Parra G, Bradnam K, Ning Z, Keane T, Korf I. Assessing the gene space in draft genomes. Nucleic Acids Research. 2008;37: 289--297. doi:\href{https://doi.org/10.1093/nar/gkn916}{10.1093/nar/gkn916}
 
 \leavevmode\hypertarget{ref-HISAT}{}%
-33. Kim D, Langmead B, Salzberg SL. HISAT: a fast spliced aligner with
-low memory requirements. Nature Methods. 2015;12: 357--360.
-doi:\href{https://doi.org/10.1038/nmeth.3317}{10.1038/nmeth.3317}
+34. Kim D, Langmead B, Salzberg SL. HISAT: a fast spliced aligner with low memory requirements. Nature Methods. 2015;12: 357--360. doi:\href{https://doi.org/10.1038/nmeth.3317}{10.1038/nmeth.3317}
 
 \leavevmode\hypertarget{ref-StringTie}{}%
-34. Pertea M, Pertea GM, Antonescu CM, Chang T-C, Mendell JT, Salzberg
-SL. StringTie enables improved reconstruction of a transcriptome from
-RNA-seq reads. Nature Biotechnology. 2015;33: 290--295.
+35.
+Pertea M, Pertea GM, Antonescu CM, Chang T-C, Mendell JT, Salzberg SL.
+StringTie enables improved reconstruction of a transcriptome from RNA-seq reads.
+Nature Biotechnology.
+2015;33: 290--295.
 doi:\href{https://doi.org/10.1038/nbt.3122}{10.1038/nbt.3122}
@@
 \leavevmode\hypertarget{ref-Tuxedo}{}%
-35. Pertea M, Kim D, Pertea GM, Leek JT, Salzberg SL. Transcript-level
-expression analysis of RNA-seq experiments with HISAT, StringTie and
-Ballgown. Nature Protocols. 2016;11: 1650--1667.
+36.
+Pertea M, Kim D, Pertea GM, Leek JT, Salzberg SL.
+Transcript-level expression analysis of RNA-seq experiments with HISAT, StringTie and Ballgown.
+Nature Protocols.
+2016;11: 1650--1667.
 doi:\href{https://doi.org/10.1038/nprot.2016.095}{10.1038/nprot.2016.095}
 
+\leavevmode\hypertarget{ref-IQTREE}{}%
+37.
+Minh BQ, Schmidt HA, Chernomor O, Schrempf D, Woodhams MD, Haeseler A von, et al.
+IQ-TREE 2: New Models and Efficient Methods for Phylogenetic Inference in the Genomic Era.
+Molecular Biology and Evolution.
+2020;37: 1530--1534.
+doi:\href{https://doi.org/10.1093/molbev/msaa015}{10.1093/molbev/msaa015}
+
+\leavevmode\hypertarget{ref-UFBoot2}{}%
+38.
+Hoang DT, Chernomor O, Haeseler A von, Minh BQ, Vinh LS.
+UFBoot2: Improving the Ultrafast Bootstrap Approximation.
+Molecular Biology and Evolution.
+2017;35: 518--522.
+doi:\href{https://doi.org/10.1093/molbev/msx281}{10.1093/molbev/msx281}
+
+\leavevmode\hypertarget{ref-ModelFinder}{}%
+39.
+Kalyaanamoorthy S, Minh BQ, Wong TKF, Haeseler A von, Jermiin LS.
+ModelFinder: fast model selection for accurate phylogenetic estimates.
+Nature Methods.
+2017;14: 587--589.
+doi:\href{https://doi.org/10.1038/nmeth.4285}{10.1038/nmeth.4285}
+
+\leavevmode\hypertarget{ref-PMSF}{}%
+40.
+Wang H-C, Minh BQ, Susko E, Roger AJ.
+Modeling Site Heterogeneity with Posterior Mean Site Frequency Profiles Accelerates Accurate Phylogenomic Estimation.
+Systematic Biology.
+2017;67: 216--235.
+doi:\href{https://doi.org/10.1093/sysbio/syx068}{10.1093/sysbio/syx068}
+
+\leavevmode\hypertarget{ref-Schrempf2019}{}%
+41.
+Schrempf D, Minh BQ, Haeseler A von, Kosiol C.
+Polymorphism-Aware Species Trees with Advanced Mutation Models, Bootstrap, and Rate Heterogeneity.
+Molecular Biology and Evolution.
+2019;36: 1294--1301.
+doi:\href{https://doi.org/10.1093/molbev/msz043}{10.1093/molbev/msz043}
+
+\leavevmode\hypertarget{ref-Soubrier2012}{}%
+42.
+Soubrier J, Steel M, Lee MSY, Sarkissian CD, Guindon S, Ho SYW, et al.
+The Influence of Rate Heterogeneity among Sites on the Time Dependence of Molecular Rates.
+Molecular Biology and Evolution.
+2012;29: 3345--3358.
+doi:\href{https://doi.org/10.1093/molbev/mss140}{10.1093/molbev/mss140}
+
+\leavevmode\hypertarget{ref-Yang1995}{}%
+43. Yang Z, Kumar S, Nei M. A new method of inference of ancestral nucleotide and amino acid sequences. Genetics. 1995;141: 1641--50.
+
 \leavevmode\hypertarget{ref-WebGestalt2019}{}%
-36. Liao Y, Wang J, Jaehnig EJ, Shi Z, Zhang B. WebGestalt 2019: gene
-set analysis toolkit with revamped UIs and APIs. Nucleic Acids Research.
+44.
+Liao Y, Wang J, Jaehnig EJ, Shi Z, Zhang B.
+WebGestalt 2019: gene set analysis toolkit with revamped UIs and APIs.
+Nucleic Acids Research.
 2019;47: W199--W205.
@@
 \leavevmode\hypertarget{ref-Reactome}{}%
-37. Jassal B, Matthews L, Viteri G, Gong C, Lorente P, Fabregat A, et
-al. The reactome pathway knowledgebase. Nucleic acids research. 2020;48:
-D498--D503.
+45.
+Jassal B, Matthews L, Viteri G, Gong C, Lorente P, Fabregat A, et al.
+The reactome pathway knowledgebase.
+Nucleic acids research.
+2020;48: D498--D503.
 doi:\href{https://doi.org/10.1093/nar/gkz1031}{10.1093/nar/gkz1031}
@@
 \leavevmode\hypertarget{ref-Wikipathways}{}%
-38. Slenter DN, Kutmon M, Hanspers K, Riutta A, Windsor J, Nunes N, et
-al. WikiPathways: a multifaceted pathway database bridging metabolomics
-to other omics research. Nucleic Acids Research. 2017;46: D661--D667.
+46.
+Slenter DN, Kutmon M, Hanspers K, Riutta A, Windsor J, Nunes N, et al.
+WikiPathways: a multifaceted pathway database bridging metabolomics to other omics research.
+Nucleic Acids Research.
+2017;46: D661--D667.
 doi:\href{https://doi.org/10.1093/nar/gkx1064}{10.1093/nar/gkx1064}
@@
 \leavevmode\hypertarget{ref-KEGG}{}%
-39. Kanehisa M, Goto S. KEGG: Kyoto Encyclopedia of Genes and Genomes.
-Nucleic Acids Research. 2000;28: 27--30.
-doi:\href{https://doi.org/10.1093/nar/28.1.27}{10.1093/nar/28.1.27}
+47. Kanehisa M, Goto S. KEGG: Kyoto Encyclopedia of Genes and Genomes. Nucleic Acids Research. 2000;28: 27--30. doi:\href{https://doi.org/10.1093/nar/28.1.27}{10.1093/nar/28.1.27}
+
+\leavevmode\hypertarget{ref-Peto:20151c2}{}%
+48.
+Peto R.
+Quantitative implications of the approximate irrelevance of mammalian body size and lifespan to lifelong cancer risk.
+Phil Trans R Soc B.
+2015;370: 20150198.
+doi:\href{https://doi.org/10.1098/rstb.2015.0198}{10.1098/rstb.2015.0198}
 
 \leavevmode\hypertarget{ref-Felsenstein1985}{}%
-40. Felsenstein J. Phylogenies and the Comparative Method. The American
-Naturalist. 1985;125: 1--15.
-doi:\href{https://doi.org/10.1086/284325}{10.1086/284325}
+49. Felsenstein J. Phylogenies and the Comparative Method. The American Naturalist. 1985;125: 1--15. doi:\href{https://doi.org/10.1086/284325}{10.1086/284325}
 
 \leavevmode\hypertarget{ref-MartinsAndHansen1997}{}%
-41. Martins EP, Hansen TF. Phylogenies and the Comparative Method: A
-General Approach to Incorporating Phylogenetic Information into the
-Analysis of Interspecific Data. The American Naturalist. 1997;149:
-646--667. doi:\href{https://doi.org/10.1086/286013}{10.1086/286013}
+50.
+Martins EP, Hansen TF.
+Phylogenies and the Comparative Method: A General Approach to Incorporating Phylogenetic Information into the Analysis of Interspecific Data.
+The American Naturalist.
+1997;149: 646--667.
+doi:\href{https://doi.org/10.1086/286013}{10.1086/286013}
 
 \leavevmode\hypertarget{ref-Armitage:19851c2}{}%
-42. Armitage P. Multistage models of carcinogenesis. Environmental
-health perspectives. 1985;63: 195--201.
-doi:\href{https://doi.org/10.1289/ehp.8563195}{10.1289/ehp.8563195}
+51. Armitage P. Multistage models of carcinogenesis. Environmental health perspectives. 1985;63: 195--201. doi:\href{https://doi.org/10.1289/ehp.8563195}{10.1289/ehp.8563195}
 
 \leavevmode\hypertarget{ref-Armitage:20041c2}{}%
-43. Armitage P, Doll R. The age distribution of cancer and a multi-stage
-theory of carcinogenesis. British Journal of Cancer. 2004;91: 6602297.
+52.
+Armitage P, Doll R.
+The age distribution of cancer and a multi-stage theory of carcinogenesis.
+British Journal of Cancer.
+2004;91: 6602297.
 doi:\href{https://doi.org/10.1038/sj.bjc.6602297}{10.1038/sj.bjc.6602297}
 
-\leavevmode\hypertarget{ref-Peto:20151c2}{}%
-44. Peto R. Quantitative implications of the approximate irrelevance of
-mammalian body size and lifespan to lifelong cancer risk. Phil Trans R
-Soc B. 2015;370: 20150198.
-doi:\href{https://doi.org/10.1098/rstb.2015.0198}{10.1098/rstb.2015.0198}
+\leavevmode\hypertarget{ref-Lynch2000}{}%
+53.
+Lynch M, Conery JS.
+The Evolutionary Fate and Consequences of Duplicate Genes.
+Science.
+2000;290: 1151--1155.
+doi:\href{https://doi.org/10.1126/science.290.5494.1151}{10.1126/science.290.5494.1151}
+
+\leavevmode\hypertarget{ref-Goodman2009}{}%
+54.
+Goodman M, Sterner KN, Islam M, Uddin M, Sherwood CC, Hof PR, et al.
+Phylogenomic analyses reveal convergent patterns of adaptive evolution in elephant and human ancestries.
+Proceedings of the National Academy of Sciences.
+2009;106: 20824--20829.
+doi:\href{https://doi.org/10.1073/pnas.0911239106}{10.1073/pnas.0911239106}
+
+\leavevmode\hypertarget{ref-Launonen2005}{}%
+55. Launonen V. Mutations in the human LKB1/STK11 gene. Human Mutation. 2005;26: 291--297. doi:\href{https://doi.org/10.1002/humu.20222}{10.1002/humu.20222}
+
+\leavevmode\hypertarget{ref-CooperAndPurvis2010}{}%
+56. Cooper N, Purvis A. Body Size Evolution in Mammals: Complexity in Tempo and Mode. The American Naturalist. 2010;175: 727--738. doi:\href{https://doi.org/10.1086/652466}{10.1086/652466}
+
+\leavevmode\hypertarget{ref-HoneAndBenton2004}{}%
+57.
+HONE D, BENTON M.
+The evolution of large size: how does Cope's Rule work?
+Trends in Ecology \& Evolution.
+2005;20: 4--6.
+doi:\href{https://doi.org/10.1016/j.tree.2004.10.012}{10.1016/j.tree.2004.10.012}
+
+\leavevmode\hypertarget{ref-HuntAndRoy2006}{}%
+58.
+Hunt G, Roy K.
+Climate change, body size evolution, and Cope's Rule in deep-sea ostracodes.
+Proceedings of the National Academy of Sciences.
+2006;103: 1347--1352.
+doi:\href{https://doi.org/10.1073/pnas.0510550103}{10.1073/pnas.0510550103}
+
+\leavevmode\hypertarget{ref-Jablonski1997}{}%
+59. Jablonski D. Body-size evolution in Cretaceous molluscs and the status of Cope's rule. Nature. 1997;385: 250--252. doi:\href{https://doi.org/10.1038/385250a0}{10.1038/385250a0}
+
+\leavevmode\hypertarget{ref-Tyner:20021c2}{}%
+60.
+Tyner SD, Venkatachalam S, Choi J, Jones S, Ghebranious N, Igelmann H, et al.
+p53 mutant mice that display early ageing-associated phenotypes.
+Nature.
+2002;415: 45.
+doi:\href{https://doi.org/10.1038/415045a}{10.1038/415045a}
+
+\leavevmode\hypertarget{ref-GarciaCao:20021c2}{}%
+61.
+García?Cao I, García?Cao M, Martín?Caballero J, Criado LM, Klatt P, Flores JM, et al.
+'Super p53' mice exhibit enhanced DNA damage response, are tumor resistant and age normally.
+The EMBO Journal.
+2002;21: 6225--6235.
+doi:\href{https://doi.org/10.1093/emboj/cdf595}{10.1093/emboj/cdf595}
+
+\leavevmode\hypertarget{ref-BowheadGenome}{}%
+62.
+Keane M, Semeiks J, Webb AE, Li YI, Quesada V, Craig T, et al.
+Insights into the Evolution of Longevity from the Bowhead Whale Genome.
+Cell Reports.
+2015;10: 112--122.
+doi:\href{https://doi.org/10.1016/j.celrep.2014.12.008}{10.1016/j.celrep.2014.12.008}
+
+\leavevmode\hypertarget{ref-seim2013}{}%
+63.
+Seim I, Fang X, Xiong Z, Lobanov AV, Huang Z, Ma S, et al.
+Genome analysis reveals insights into physiology and longevity of the Brandt?s bat Myotis brandtii.
+Nature Communications.
+2013;4: 2212.
+doi:\href{https://doi.org/10.1038/ncomms3212}{10.1038/ncomms3212}
+
+\leavevmode\hypertarget{ref-myoDav1_pteAle1}{}%
+64.
+Zhang G, Cowled C, Shi Z, Huang Z, Bishop-Lilly KA, Fang X, et al.
+Comparative Analysis of Bat Genomes Provides Insight into the Evolution of Flight and Immunity.
+Science.
+2012;339: 456--460.
+doi:\href{https://doi.org/10.1126/science.1230835}{10.1126/science.1230835}
+
+\leavevmode\hypertarget{ref-NakedMoleRatGenome}{}%
+65.
+Kim EB, Fang X, Fushan AA, Huang Z, Lobanov AV, Han L, et al.
+Genome sequencing reveals insights into physiology and longevity of the naked mole rat.
+Nature.
+2011;479: 223--7.
+doi:\href{https://doi.org/10.1038/nature10533}{10.1038/nature10533}
+
+\leavevmode\hypertarget{ref-BlindMoleRatGenome}{}%
+66.
+Fang X, Nevo E, Han L, Levanon EY, Zhao J, Avivi A, et al.
+Genome-wide adaptive complexes to underground stresses in blind mole rats Spalax.
+Nature Communications.
+2014;5: 3966.
+doi:\href{https://doi.org/10.1038/ncomms4966}{10.1038/ncomms4966}
+
+\leavevmode\hypertarget{ref-Rastogi2005}{}%
+67.
+Rastogi S, Liberles DA.
+Subfunctionalization of duplicated genes as a transition state to neofunctionalization.
+BMC Evolutionary Biology.
+2005;5: 28.
+doi:\href{https://doi.org/10.1186/1471-2148-5-28}{10.1186/1471-2148-5-28}
+
+\leavevmode\hypertarget{ref-QianAndZhang2014}{}%
+68. Qian W, Zhang J. Genomic evidence for adaptation by gene duplication. Genome research. 2014;24: 1356--62. doi:\href{https://doi.org/10.1101/gr.172098.114}{10.1101/gr.172098.114}
+
+\leavevmode\hypertarget{ref-Stoltzfus1999}{}%
+69. Stoltzfus A. On the Possibility of Constructive Neutral Evolution. Journal of Molecular Evolution. 1999;49: 169--181. doi:\href{https://doi.org/10.1007/pl00006540}{10.1007/pl00006540}
 

### 2020-06-03 paper_PLOS/paper_PLOS_draft.Rmd
Fixed Figures and converted to bookdown-based knitting, need to fix HTML and WORD outputs

@@
   code: "SUNY Buffalo"
+always_allow_html: true
 output:
+  bookdown::html_document2: default
+  bookdown::pdf_document2: default
   bookdown::pdf_book:
@@
       in_header: preamble.tex
-  pdf_document: default
-  word_document: default
-  workflowr::wflow_html: default
+  bookdown::word_document2: default
 csl: plos.csl
@@
 ```{r Libraries and functions, include = F}
-knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, out.width="6in",
-                      fig.retina = 2, fig.align = "left", dpi = 300)
+knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, out.width="6in", fig.pos = "H", out.extra = "",
+                      fig.retina = 2, dpi = 300)
 
@@
   scale_color_viridis_c(guide=F) +
-  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold"))) +
-  geom_tiplab(aes(label=Spc.short), offset = 0.4, color="black", size=4, parse=T)+
+  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold", size=6))) +
+  geom_tiplab(aes(label=Spc.short), offset = 0.4, color="black", size=2, parse=T)+
   geom_tiplab(aes(label=not.me(lnSize, N.Genes, threshold = 7)), geom="label",  offset=-0.5, color="black", size=2, fontface="bold")+ 
@@
 
+#Prep work
+ml.tree.increasing.geneCounts.withSize.table <- ml.tree.increasing.geneCounts.withSize %>% 
+  as_tibble %>% 
+  mutate(joint.label = str_c(label, ": ", N.Genes))
+
+ml.tree.increasing.geneCounts.withSize.table[which(ml.tree.increasing.geneCounts.withSize.table$label %in% genomeTable.atlantogenata$label),]$joint.label <- ""
+
+ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts.withSize.table %>% as.treedata()
+
 p.ml.tree.increasing.geneCounts.withSize.withBL <- ml.tree.increasing.geneCounts.withSize %>% 
@@
               color="black")+
-  geom_tiplab(aes(label=not.me(lnSize, N.Genes)), geom="label",  
-              offset=0, size=3,
+  geom_label_repel(aes(
+    label=yes.me(lnSize, joint.label, threshold=7)
+  ),
+    size=3, direction = "x", nudge_x = -0.2, force=10, segment.color = "black",
+    color="grey"
+  ) + 
+  geom_label_repel(aes(
+    label=not.me(lnSize, joint.label, threshold=7)
+  ),
+    size=3, direction = "x",nudge_x = -0.05, force=10, 
+    color="black") + 
+  geom_tiplab(aes(label=not.me(lnSize, N.Genes, threshold=7)), geom = "label", size=3, color="black") + 
+  geom_tiplab(aes(label=yes.me(lnSize, N.Genes, threshold=7)), geom = "label", size=3, color="grey") + 
+  geom_nodelab2(aes(subset=(label=="Root"), label = "Atlantogenata", x=branch), geom="label",
+               size=2,
                color="black")+
-  geom_tiplab(aes(label=yes.me(lnSize, N.Genes)), geom="label",  
-              offset=0.0, size=3,
-              color="grey")+ 
-  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes)), geom="label", 
-               nudge_y = 0.2, nudge_x = 0, size=3, 
-               color="black")+ 
-  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes)), geom="label", 
-               nudge_y = -0.2, nudge_x = 0, size=3, 
-               color="grey")+ 
-  geom_nodelab(aes(x=branch, label=not.me(lnSize, label)), geom="label", 
-               nudge_y = 0.2, nudge_x = -0.031, size=3, 
-               color="black")+ 
-  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label)), geom="label", 
-               nudge_y = -0.2, nudge_x = -0.031, size=3, 
-               color="grey")+ 
-  geom_nodelab2(aes(subset=label=="Root", x=branch), geom="label", 
-               size=3, 
-               color="grey")+ 
   theme_tree2() + 
-  xlim(-0.01,0.45) +
-  theme(legend.position = "bottom") +
-  xlab("Rate of Change, Body Size")
+  xlim(-0.01,0.4) +
+  xlab("Rate of Change, Gene Copy Number") + 
+  theme(
+    legend.position="bottom",
+    # panel.background = element_rect(fill = "lightgrey"),
+    text = element_text(face = "plain", 
+                        colour = "black", lineheight = 0.9, 
+                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
+                        debug = FALSE), 
+    axis.text.x = element_text(size = 4,
+                               colour = "black", face = "bold"),
+    axis.title = element_text(size = 6, 
+                              colour = "black", face = "bold"),
+    plot.title = element_text(size = 6, 
+                              colour = "black", lineheight = 1, face = "bold"), 
+    legend.title = element_text(size = 6, 
+                                face = "bold", colour = "black"), 
+    legend.text = element_text(size = 6, 
+                               face = "plain", colour = "black")
+  )
+
 # p.ml.tree.increasing.geneCounts.withSize.withBL
@@
     x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: One Contiguous Gene, One Hit"
+    y = "Query Coverage"
   )
@@
     x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: Two Contiguous Genes, Two Hits"
+    y = "Query Coverage"
   )
@@
     x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: One Fragmented Gene, Multiple Hits"
+    y = "Query Coverage"
   )
@@
     x = "Query Length",
-    y = "Query Coverage"#,
-    #title = "ECNC: Two Genes, Multiple Hits"
+    y = "Query Coverage"
   )
@@
     y = "Query Coverage"
-    # title = "ECNC: One Real Gene, Multiple Hits on Conserved Domain"
   )
@@
     x = "Query Length",
-    y = "Query Coverage",
-    title = "ECNC: Two Real Genes, Unidentified Tails"
+    y = "Query Coverage"
   )
@@
     axis.text.x = element_text(
-      angle = -330, vjust = 0,
+      angle = -270, vjust = 0,
       size = highlightTextSize(cor.GCT.CEGMA$Var1, cor.toHighlight, sizes = c(6,9)), 
@@
 For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways_cancer databases using WebGestalt ORA.
-Going through the tree, at no FDR $\leq$ 0.5 is there any significant pathway representation for genes that increased in the branches leading to <!--`r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`--> _Afrosoricida_, _Pseudoungulata_, _Elephantina_, _Loxodontini_, _Elephas maximus_, _Mammuthus_, _Loxodonta cyclotis_, _Palaeoloxodon antiquus_, _Loxodona_, _Mammuthus columbi_, _Mammuthus primigenius_, _Procavia capensis_, _Elephantidae_, or _Proboscidea_; furthermore, there are no significant pathway enrichments at FDR<0.5 for genes whose copy number did not change between branches (copy-number-stable) for _Afrotheria__.
+Going through the tree, at no FDR $\leq$ 0.5 is there any significant pathway representation for genes that increased in the branches leading to <!--`r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`--> _Afrosoricida_, _Pseudoungulata_, _Elephantina_, _Loxodontini_, _Elephas maximus_, _Mammuthus_, _Loxodonta cyclotis_, _Palaeoloxodon antiquus_, _Loxodona_, _Mammuthus columbi_, _Mammuthus primigenius_, _Procavia capensis_, _Elephantidae_, or _Proboscidea_; furthermore, there are no significant pathway enrichments at FDR<0.5 for genes whose copy number did not change between branches (copy-number-stable) for _Afrotheria_.
 Note that because _Xenarthra_ was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.
@@
   xlim(0,15) +
+  ylim(0,17)+
   theme_tree()+
@@
 
-```{r Figure-intro, fig.cap="Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]",  out.width="6in", fig.asp=1.3}
+(ref:Figure-intro) _Atlantogenatans_ with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]
+
+```{r Figure-intro, fig.cap="(ref:Figure-intro)",  out.width="6in", fig.asp=1.2}
 atlantogenata.tree.size.lifespan
@@
 
-```{r Figure-CancerSuscept, fig.cap="Cancer succecptibility across Atlantogenata.
-Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.", out.width="6in", fig.asp=1}
+(ref:Figure-CancerSuscept) Cancer succecptibility across _Atlantogenata._ Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.
+
+```{r Figure-CancerSuscept, fig.cap="(ref:Figure-CancerSuscept)", out.width="6in", fig.asp=1}
 p.tree.cancerSucceptibility.Atlantogenata
@@
 
-```{r Figure-GeneDup-Cladogram, fig.cap="Gene duplications occur readily throughout Atlantogenata.
-Shown here is a tree of Atlantogenatan species with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.", out.height="7in", out.width="6in"}
+(ref:Figure-GeneDup-Cladogram) Gene duplications occur readily throughout _Atlantogenata_.
+Shown here is a tree of _Atlantogenatan_ species with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.
+
+```{r Figure-GeneDup-Cladogram, fig.cap="(ref:Figure-GeneDup-Cladogram)", out.height="6in", out.width="6in"}
 p.ml.tree.increasing.geneCounts.withSize
@@
 
-```{r Figure-GenomeCor, fig.cap="Correlations between genome quality metrics and ECNC metrics.
-Gene copy number metrics, and the genome quality metrics most strongly associated with them, are highlighted in red.", fig.asp=1, out.width="6in", out.height="6in", fig.align="center"}
+(ref:Figure-GenomeCor) Correlations between genome quality metrics and ECNC metrics.
+Gene copy number metrics, and the genome quality metrics most strongly associated with them, are highlighted in red.
+
+```{r Figure-GenomeCor, fig.cap="(ref:Figure-GenomeCor)", fig.asp=1, out.width="6in", out.height="6in", fig.align="center"}
 p.cor.GCT.CEGMA
@@
 
-```{r Figure-loxAfrORA, fig.cap="Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.", out.width="6in", out.height="4in"}
+(ref:Figure-loxAfrORA) Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.
+
+```{r Figure-loxAfrORA, fig.cap="(ref:Figure-loxAfrORA)", out.width="6in", out.height="4in"}
 fig.loxAfr.ORA <- ggarrange(
@@
 
-```{r Figure-loxAfrTP53, fig.cap="", out.width="6in", out.height="6in"}
+(ref:Figure-loxAfrTP53) TP53-related genes are also duplicated and functional in _Loxodonta africana_.
+__A)__ Cladogram of _Atlantogenata_ highlighting along each branch when duplications for each gene occured.
+__B)__ Gene expression data collected from pulically-available RNA-seq data for each duplicate in __A)__.
+
+```{r Figure-loxAfrTP53, fig.cap="(ref:Figure-loxAfrTP53)", out.width="6in"}
 fig.TP53 <- ggarrange(
@@
 # TABLES
+
+```{r Table-BodySize, eval=T}
+t.bodysize.atlantogenata.withNodes <- st.Atlantogenata %>% 
+  as_tibble() %>% 
+  filter(!is.na(label)) %>%
+  select(
+    `Node`=label, 
+    `Size (log(g))`=Median, 
+    `95% CI (Low)` = `95%CI_Low`, 
+    `95% CI (High)` = `95%CI_High`,  
+    `Rate (sqrt)`=sqrt.rate
+  ) %>%
+  mutate_at(vars(-`Node`, ), . %>% as.numeric %>% round(2)) %>% 
+  kable(
+    ., 
+    booktabs = TRUE, longtable = T,
+    caption = 'Body Size and Confidence Intervals in Atlantogenata estimated using StableTraits.'
+  ) %>% kable_styling(latex_options = "repeat_header")
+t.bodysize.atlantogenata.withNodes
+```
 
@@
   ) %>% 
-  column_spec(4, width="50em") %>% 
-  column_spec(5, width="50em") %>% 
+  column_spec(4, width="10em") %>% 
+  column_spec(5, width="10em") %>% 
   kable_styling(latex_options = c("scale_down"))
-```
-
-```{r Table-BodySize, eval=T}
-t.bodysize.atlantogenata.withNodes <- st.Atlantogenata %>% 
-  as_tibble() %>% 
-  filter(!is.na(label)) %>%
-  select(
-    `Node`=label, 
-    `Size (log(g))`=Median, 
-    `95% CI (Low)` = `95%CI_Low`, 
-    `95% CI (High)` = `95%CI_High`,  
-    `Rate (sqrt)`=sqrt.rate
-  ) %>%
-  mutate_at(vars(-`Node`, ), . %>% as.numeric %>% round(2)) %>% 
-  kable(
-    ., 
-    booktabs = TRUE, longtable = T,
-    caption = 'Body Size and Confidence Intervals in Atlantogenata estimated using StableTraits.'
-  ) %>% kable_styling(latex_options = "repeat_header")
-t.bodysize.atlantogenata.withNodes
 ```
@@
 
-```{r Figure-ECNC, fig.cap="Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
+(ref:Figure-ECNC) Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
 __A)__ A single, contiguous gene homolog in a target genome with 100% query length coverage has an ECNC of 1.0.
@@
 Only 75% of the query sequence was covered in total between the hits, but for that 75%, each nucleotide has two hits.
-As such, ECNC is equal to 2.0 for this gene.", eval=T}
+As such, ECNC is equal to 2.0 for this gene.
+
+```{r Figure-ECNC, fig.cap="(ref:Figure-ECNC)", eval=T}
 fig.ECNC <- list(
@@
 ```{r Figure-RBHBStrategy, fig.cap="Full version of RecBlat strategy, but low quality", warning=T, eval=T}
-warning("TODO: Make HQ version of this figure")
 image_read("../output/figs/RBHBDiagram.png") %>% 
-  image_rotate(180)%>% 
+  # image_rotate(180)%>% 
+  image_flip() %>% 
   image_ggplot()

### 2020-06-03 paper_PLOS/paper_PLOS_draft.tex
Fixed Figures and converted to bookdown-based knitting, need to fix HTML and WORD outputs

@@
 \usepackage{pdflscape}
+
+\usepackage{float}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[H]
+} {
+    \endorigfigure
+}
 \usepackage{booktabs}
@@
 For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways\_cancer databases using WebGestalt ORA.
-Going through the tree, at no FDR \(\leq\) 0.5 is there any significant pathway representation for genes that increased in the branches leading to \emph{Afrosoricida}, \emph{Pseudoungulata}, \emph{Elephantina}, \emph{Loxodontini}, \emph{Elephas maximus}, \emph{Mammuthus}, \emph{Loxodonta cyclotis}, \emph{Palaeoloxodon antiquus}, \emph{Loxodona}, \emph{Mammuthus columbi}, \emph{Mammuthus primigenius}, \emph{Procavia capensis}, \emph{Elephantidae}, or \emph{Proboscidea}; furthermore, there are no significant pathway enrichments at FDR\textless{}0.5 for genes whose copy number did not change between branches (copy-number-stable) for \_Afrotheria\_\_.
+Going through the tree, at no FDR \(\leq\) 0.5 is there any significant pathway representation for genes that increased in the branches leading to \emph{Afrosoricida}, \emph{Pseudoungulata}, \emph{Elephantina}, \emph{Loxodontini}, \emph{Elephas maximus}, \emph{Mammuthus}, \emph{Loxodonta cyclotis}, \emph{Palaeoloxodon antiquus}, \emph{Loxodona}, \emph{Mammuthus columbi}, \emph{Mammuthus primigenius}, \emph{Procavia capensis}, \emph{Elephantidae}, or \emph{Proboscidea}; furthermore, there are no significant pathway enrichments at FDR\textless{}0.5 for genes whose copy number did not change between branches (copy-number-stable) for \emph{Afrotheria}.
 Note that because \emph{Xenarthra} was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.
@@
 
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-intro-1} \hfill{}
-
-\caption{Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]}\label{fig:Figure-intro}
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-intro-1} \caption{\emph{Atlantogenatans} with sequenced genomes, body sizes, and known lifespans {[}14,21{]}}\label{fig:Figure-intro}
 \end{figure}
 
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-BodySize-1} \hfill{}
-
-\caption{Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-BodySize-1} \caption{Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
 **A)** Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch.
@@
 
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-CancerSuscept-1} \hfill{}
-
-\caption{Cancer succecptibility across Atlantogenata.
-Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.}\label{fig:Figure-CancerSuscept}
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-CancerSuscept-1} \caption{Cancer succecptibility across \emph{Atlantogenata.} Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.}\label{fig:Figure-CancerSuscept}
 \end{figure}
 
-\begin{figure}
-
-\includegraphics[width=6in,height=7in]{paper_PLOS_draft_files/figure-latex/Figure-GeneDup-Cladogram-1} \hfill{}
-
-\caption{Gene duplications occur readily throughout Atlantogenata.
-Shown here is a tree of Atlantogenatan species with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.}\label{fig:Figure-GeneDup-Cladogram}
+\begin{figure}[H]
+\includegraphics[width=6in,height=6in,]{paper_PLOS_draft_files/figure-latex/Figure-GeneDup-Cladogram-1} \caption{Gene duplications occur readily throughout \emph{Atlantogenata}.
+Shown here is a tree of \emph{Atlantogenatan} species with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.}\label{fig:Figure-GeneDup-Cladogram}
 \end{figure}
 
-\begin{figure}
-
-{\centering \includegraphics[width=6in,height=6in]{paper_PLOS_draft_files/figure-latex/Figure-GenomeCor-1}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=6in,height=6in,]{paper_PLOS_draft_files/figure-latex/Figure-GenomeCor-1}
 
@@
 
-\begin{figure}
-
-\includegraphics[width=6in,height=4in]{paper_PLOS_draft_files/figure-latex/Figure-loxAfrORA-1} \hfill{}
-
-\caption{Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.}\label{fig:Figure-loxAfrORA}
+\begin{figure}[H]
+\includegraphics[width=6in,height=4in,]{paper_PLOS_draft_files/figure-latex/Figure-loxAfrORA-1} \caption{Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.}\label{fig:Figure-loxAfrORA}
 \end{figure}
 
-\begin{figure}
-
-\includegraphics[width=6in,height=6in]{paper_PLOS_draft_files/figure-latex/Figure-loxAfrTP53-1} \hfill{}
-
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-loxAfrTP53-1} \caption{TP53-related genes are also duplicated and functional in \emph{Loxodonta africana}.
+\textbf{A)} Cladogram of \emph{Atlantogenata} highlighting along each branch when duplications for each gene occured.
+\textbf{B)} Gene expression data collected from pulically-available RNA-seq data for each duplicate in \textbf{A)}.}\label{fig:Figure-loxAfrTP53}
 \end{figure}
@@
 \section{TABLES}\label{tables}}
-
-\begin{table}
-
-\caption{\label{tab:Table-CancerSuscept}Estimated Cancer Succeptibilty for nodes in Atlantogenata}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lrlllr}
-\toprule
-Node & Est. Lifespan & K1 & K2 & Change in K & log2 Change\\
-\midrule
-Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
-Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 & 4.07\\
-Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
-Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-\addlinespace
-Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 & 27.66\\
-Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
-Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 & 4.14\\
-Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
-\addlinespace
-Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 & -23.05\\
-Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 & 29.97\\
-Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 & 6.92\\
-Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 & 8.62\\
-\addlinespace
-Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 & 0.00\\
-Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 & -5.01\\
-Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 & 9.22\\
-Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 & 5.38\\
-Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
-\addlinespace
-Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
-Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 & 0.00\\
-Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 & 5.22\\
-Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 & 9.22\\
-Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 & 20.75\\
-\addlinespace
-Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 & -3.76\\
-Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 & 1.51\\
-Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 & 0.00\\
-Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 & 0.00\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:Table-GeneDup}Summary of duplications in Atlantogenata}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lllrrllr}
-\toprule
-Species & Common Name & Size (g) & \#Hits & \#Duplicated & \% Genes Found & \% Hits Duplicated & Mean ECNC/Hit\\
-\midrule
-Choloepus hoffmanni & Hoffmans Two-Toed Sloth & 4.3e+03 & 14082 & 3204 & 78.19\% & 22.75\% & 0.98\\
-Chrysochloris asiatica & Cape Golden Mole & 49 & 13547 & 2716 & 75.22\% & 20.05\% & 0.99\\
-Dasypus novemcinctus & Nine-Banded Armadillo & 4.8e+03 & 13819 & 2605 & 76.73\% & 18.85\% & 0.98\\
-Echinops telfairi & Lesser Hedgehog Tenrec & 87 & 12903 & 1670 & 71.64\% & 12.94\% & 0.99\\
-Elephantulus edwardii & Cape Elephant Shrew & 49 & 12884 & 3048 & 71.53\% & 23.66\% & 0.99\\
-\addlinespace
-Elephas maximus & Asian Elephant & 3.3e+06 & 14073 & 907 & 78.14\% & 6.44\% & 1.00\\
-Loxodonta africana & African Savanna Elephant & 4.6e+06 & 14051 & 940 & 78.01\% & 6.69\% & 1.00\\
-Loxodonta cyclotis & African Forest Elephant & 4.7e+06 & 14065 & 900 & 78.09\% & 6.40\% & 1.00\\
-Mammut americanum & American Mastodon & 6e+06 & 13840 & 737 & 76.84\% & 5.33\% & 1.00\\
-Mammuthus columbi & Columbian Mammoth & 6.6e+06 & 13059 & 426 & 72.51\% & 3.26\% & 1.00\\
-\addlinespace
-Mammuthus primigenius & Woolly Mammoth & 4.3e+06 & 13935 & 723 & 77.37\% & 5.19\% & 1.00\\
-Orycteropus afer & Aardvark & 5.3e+04 & 13880 & 1083 & 77.06\% & 7.80\% & 0.99\\
-Palaeoloxodon antiquus & Straight Tusked Elephant & 1e+07 & 13969 & 745 & 77.56\% & 5.33\% & 1.00\\
-Procavia capensis & Rock Hyrax & 3e+03 & 13672 & 788 & 75.91\% & 5.76\% & 1.00\\
-Trichechus manatus & Manatee & 6.9e+05 & 14092 & 1046 & 78.24\% & 7.42\% & 1.00\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:Table-ORAFreqDatabase}Number of pathways overrepresented among duplicated genes at different FDRs.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{llrrrr}
-\toprule
-Ancestor & Node & Pathways at FDR<0.1 & Pathways at FDR<0.2 & Pathways at FDR<0.3 & Pathways at FDR<0.5\\
-\midrule
-Afroinsectivora & Elephantulus edwardii & 252 & 37 & 30 & 87\\
-Afrosoricida & Chrysochloris asiatica & 90 & 48 & 43 & 105\\
-Afrosoricida & Echinops telfairi & 0 & 2 & 0 & 31\\
-Afrotheria & Afroinsectivora & 0 & 0 & 0 & 33\\
-Loxodontini & Loxodonta africana & 6 & 0 & 1 & 0\\
-\addlinespace
-Paenungulata & Tethytheria & 0 & 0 & 0 & 2\\
-Proboscidea & Mammut americanum & 0 & 0 & 0 & 6\\
-Pseudoungulata & Orycteropus afer & 27 & 67 & 29 & 67\\
-Tethytheria & Proboscidea & 0 & 3 & 0 & 3\\
-Tethytheria & Trichechus manatus & 4 & 0 & 0 & 2\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:Table-SRAData}NCBI SRA datasets used in this study, along with key biological and genome information.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lll>{\raggedright\arraybackslash}p{50em}>{\raggedright\arraybackslash}p{50em}}
-\toprule
-Organism & Common Name & Genome & SRA Acc. & Tissues\\
-\midrule
-Dasypus novemcinctus & Nine-banded armadillo & dasNov3 & SRR494779, SRR494767, SRR494780, SRR494770, SRR309130, SRR494771, SRR4043756, SRR494776, SRR494778, SRR4043762, SRR4043755, SRR6206923, SRR4043761, SRR4043760, SRR6206913, SRR4043763, SRR494772, SRR494781, SRR494774, SRR494777, SRR494775, SRR4043754, SRR1289524, SRR4043758, SRR6206903, SRR1289523, SRR4043759, SRR3222425, SRR494768, SRR494769, SRR6206908, SRR4043757, SRR494766, SRR6206918, SRR494773 & Kidney, Spleen, Cerebellum W/ Brainstem, Rt.
-Quadricep, Mid-Stage Pregnant Endometrium, Cervix, Lung, Liver, Skeletal Muscle, Ascending Colon, Pregnant Armadillo Endometrium, Heart, Placenta\\
-Loxodonta africana & African savanna elephant & loxAfr3, loxAfrC, loxAfr4 & SRR6307198, SRR1041765, SRR6307199, SRR6307201, SRR6307196, SRR6307202, SRR6307200, SRR6307195, SRR975188, SRR6307194, SRR6307204, SRR3222430, SRR6307205, SRR975189, SRR6307197, SRR6307203 & Blood, Fibroblast, Placenta\\
-Trichechus manatus latirostris & Manatee & triMan1, triManLat2 & SRR4228542, SRR4228545, SRR4228544, SRR4228539, SRR4228541, SRR4228538, SRR4228546, SRR4228537, SRR4228540, SRR4228543, SRR4228547 & Buffy Coat\\
-\bottomrule
-\end{tabular}}
-\end{table}
 
@@
 
+\caption{\label{tab:Table-CancerSuscept}Estimated Cancer Succeptibilty for nodes in Atlantogenata}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrlllr}
+\toprule
+Node & Est. Lifespan & K1 & K2 & Change in K & log2 Change\\
+\midrule
+Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
+Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 & 4.07\\
+Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
+Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+\addlinespace
+Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 & 27.66\\
+Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
+Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 & 4.14\\
+Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
+\addlinespace
+Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
+Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 & -23.05\\
+Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 & 29.97\\
+Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 & 6.92\\
+Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 & 8.62\\
+\addlinespace
+Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 & 0.00\\
+Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 & -5.01\\
+Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 & 9.22\\
+Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 & 5.38\\
+Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
+\addlinespace
+Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
+Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 & 0.00\\
+Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 & 5.22\\
+Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 & 9.22\\
+Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 & 20.75\\
+\addlinespace
+Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 & -3.76\\
+Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 & 1.51\\
+Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 & 0.00\\
+Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 & 0.00\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-GeneDup}Summary of duplications in Atlantogenata}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lllrrllr}
+\toprule
+Species & Common Name & Size (g) & \#Hits & \#Duplicated & \% Genes Found & \% Hits Duplicated & Mean ECNC/Hit\\
+\midrule
+Choloepus hoffmanni & Hoffmans Two-Toed Sloth & 4.3e+03 & 14082 & 3204 & 78.19\% & 22.75\% & 0.98\\
+Chrysochloris asiatica & Cape Golden Mole & 49 & 13547 & 2716 & 75.22\% & 20.05\% & 0.99\\
+Dasypus novemcinctus & Nine-Banded Armadillo & 4.8e+03 & 13819 & 2605 & 76.73\% & 18.85\% & 0.98\\
+Echinops telfairi & Lesser Hedgehog Tenrec & 87 & 12903 & 1670 & 71.64\% & 12.94\% & 0.99\\
+Elephantulus edwardii & Cape Elephant Shrew & 49 & 12884 & 3048 & 71.53\% & 23.66\% & 0.99\\
+\addlinespace
+Elephas maximus & Asian Elephant & 3.3e+06 & 14073 & 907 & 78.14\% & 6.44\% & 1.00\\
+Loxodonta africana & African Savanna Elephant & 4.6e+06 & 14051 & 940 & 78.01\% & 6.69\% & 1.00\\
+Loxodonta cyclotis & African Forest Elephant & 4.7e+06 & 14065 & 900 & 78.09\% & 6.40\% & 1.00\\
+Mammut americanum & American Mastodon & 6e+06 & 13840 & 737 & 76.84\% & 5.33\% & 1.00\\
+Mammuthus columbi & Columbian Mammoth & 6.6e+06 & 13059 & 426 & 72.51\% & 3.26\% & 1.00\\
+\addlinespace
+Mammuthus primigenius & Woolly Mammoth & 4.3e+06 & 13935 & 723 & 77.37\% & 5.19\% & 1.00\\
+Orycteropus afer & Aardvark & 5.3e+04 & 13880 & 1083 & 77.06\% & 7.80\% & 0.99\\
+Palaeoloxodon antiquus & Straight Tusked Elephant & 1e+07 & 13969 & 745 & 77.56\% & 5.33\% & 1.00\\
+Procavia capensis & Rock Hyrax & 3e+03 & 13672 & 788 & 75.91\% & 5.76\% & 1.00\\
+Trichechus manatus & Manatee & 6.9e+05 & 14092 & 1046 & 78.24\% & 7.42\% & 1.00\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-ORAFreqDatabase}Number of pathways overrepresented among duplicated genes at different FDRs.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{llrrrr}
+\toprule
+Ancestor & Node & Pathways at FDR<0.1 & Pathways at FDR<0.2 & Pathways at FDR<0.3 & Pathways at FDR<0.5\\
+\midrule
+Afroinsectivora & Elephantulus edwardii & 252 & 37 & 30 & 87\\
+Afrosoricida & Chrysochloris asiatica & 90 & 48 & 43 & 105\\
+Afrosoricida & Echinops telfairi & 0 & 2 & 0 & 31\\
+Afrotheria & Afroinsectivora & 0 & 0 & 0 & 33\\
+Loxodontini & Loxodonta africana & 6 & 0 & 1 & 0\\
+\addlinespace
+Paenungulata & Tethytheria & 0 & 0 & 0 & 2\\
+Proboscidea & Mammut americanum & 0 & 0 & 0 & 6\\
+Pseudoungulata & Orycteropus afer & 27 & 67 & 29 & 67\\
+Tethytheria & Proboscidea & 0 & 3 & 0 & 3\\
+Tethytheria & Trichechus manatus & 4 & 0 & 0 & 2\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-SRAData}NCBI SRA datasets used in this study, along with key biological and genome information.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lll>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}}
+\toprule
+Organism & Common Name & Genome & SRA Acc. & Tissues\\
+\midrule
+Dasypus novemcinctus & Nine-banded armadillo & dasNov3 & SRR494779, SRR494767, SRR494780, SRR494770, SRR309130, SRR494771, SRR4043756, SRR494776, SRR494778, SRR4043762, SRR4043755, SRR6206923, SRR4043761, SRR4043760, SRR6206913, SRR4043763, SRR494772, SRR494781, SRR494774, SRR494777, SRR494775, SRR4043754, SRR1289524, SRR4043758, SRR6206903, SRR1289523, SRR4043759, SRR3222425, SRR494768, SRR494769, SRR6206908, SRR4043757, SRR494766, SRR6206918, SRR494773 & Kidney, Spleen, Cerebellum W/ Brainstem, Rt.
+Quadricep, Mid-Stage Pregnant Endometrium, Cervix, Lung, Liver, Skeletal Muscle, Ascending Colon, Pregnant Armadillo Endometrium, Heart, Placenta\\
+Loxodonta africana & African savanna elephant & loxAfr3, loxAfrC, loxAfr4 & SRR6307198, SRR1041765, SRR6307199, SRR6307201, SRR6307196, SRR6307202, SRR6307200, SRR6307195, SRR975188, SRR6307194, SRR6307204, SRR3222430, SRR6307205, SRR975189, SRR6307197, SRR6307203 & Blood, Fibroblast, Placenta\\
+Trichechus manatus latirostris & Manatee & triMan1, triManLat2 & SRR4228542, SRR4228545, SRR4228544, SRR4228539, SRR4228541, SRR4228538, SRR4228546, SRR4228537, SRR4228540, SRR4228543, SRR4228547 & Buffy Coat\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
 \caption{\label{tab:Table-Genomes}Genomes used in this study.}
@@
 
-\textbackslash{}begin\{figure\}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-ECNC-1} \hfill{}
-
-\textbackslash{}caption\{Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-ECNC-1} \caption{Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
 \textbf{A)} A single, contiguous gene homolog in a target genome with 100\% query length coverage has an ECNC of 1.0.
@@
 Only 75\% of the query sequence was covered in total between the hits, but for that 75\%, each nucleotide has two hits.
-As such, ECNC is equal to 2.0 for this gene.\}\label{fig:Figure-ECNC}
-\textbackslash{}end\{figure\}
-
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-GeneDup-BLScaled-1} \hfill{}
-
-\caption{Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number}\label{fig:Figure-GeneDup-BLScaled}
+As such, ECNC is equal to 2.0 for this gene.}\label{fig:Figure-ECNC}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-GeneDup-BLScaled-1} \caption{Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number}\label{fig:Figure-GeneDup-BLScaled}
 \end{figure}
@@
 
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-RBHBStrategy-1} \hfill{}
-
-\caption{Full version of RecBlat strategy, but low quality}\label{fig:Figure-RBHBStrategy}
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-RBHBStrategy-1} \caption{Full version of RecBlat strategy, but low quality}\label{fig:Figure-RBHBStrategy}
 \end{figure}
 
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure-TimeTree-Eutheria-1} \hfill{}
-
-\caption{The time-calibrated Eutherian tree used as input for StableTraits, from Bininda-Emonds et al and Puttick and Thomas}\label{fig:Figure-TimeTree-Eutheria}
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-TimeTree-Eutheria-1} \caption{The time-calibrated Eutherian tree used as input for StableTraits, from Bininda-Emonds et al and Puttick and Thomas}\label{fig:Figure-TimeTree-Eutheria}
 \end{figure}

### 2020-06-04 paper_PLOS/paper_PLOS_draft.Rmd
New double-tree figures

@@
 suppressMessages(suppressWarnings(require(stargazer)))
+suppressMessages(suppressWarnings(require(ggnewscale)))
 source("../code/generalFunctions.R")
@@
 
+```{r Figure-DoubleTree-Bodysize-CancerSuscept-Genomes}
+
+### Body Size ###
+bsize.Atlantogenata.table <- bsize.Atlantogenata %>% as_tibble
+bsize.Atlantogenata.table[which(bsize.Atlantogenata.table$node == MRCA(bsize.Atlantogenata, "Pseudoungulata", "Elephantulus.edwardii")),]$label <- "Afroinsectivora"
+bsize.Atlantogenata <- bsize.Atlantogenata.table %>% as.treedata()
+
+lnsize.bsize.Atlantogenata <- bsize.Atlantogenata %>% as_tibble() %>% select(label, Species.strict, lnSize)
+bsize.Atlantogenata.genomes <-  bsize.Atlantogenata@phylo %>% 
+  treeio::drop.tip(bsize.Atlantogenata@phylo$tip.label[!bsize.Atlantogenata@phylo$tip.label %in% genomeTable.atlantogenata$label]) %>% as.tibble
+bsize.Atlantogenata.genomes <- left_join(bsize.Atlantogenata.genomes, lnsize.bsize.Atlantogenata) %>% as.treedata
+
+p.tree.atlantogenata.genomes <- bsize.Atlantogenata.genomes %>% 
+  ggtree(
+    aes(color=as.numeric(lnSize))
+  )
+
+### Cancer Suscept ###
+tree.cancerSucceptibility.Atlantogenata.table <- tree.cancerSucceptibility.Atlantogenata %>% as_tibble
+tree.cancerSucceptibility.Atlantogenata.table[which(tree.cancerSucceptibility.Atlantogenata.table$node == MRCA(tree.cancerSucceptibility.Atlantogenata, "Pseudoungulata", "Elephantulus edwardii")),]$label <- "Afroinsectivora"
+tree.cancerSucceptibility.Atlantogenata <- tree.cancerSucceptibility.Atlantogenata.table %>% as.treedata()
+
+cancerSucceptibility.Atlantogenata.genomes <- tree.cancerSucceptibility.Atlantogenata %>% 
+  as_tibble() %>%  
+  mutate(abslog2CancerSucceptabilityChange = log2CancerSucceptabilityChange %>% abs(), 
+         branch.length=abslog2CancerSucceptabilityChange) %>% 
+  select(label, abslog2CancerSucceptabilityChange, log2CancerSucceptabilityChange, lnK1)
+tree.cancerSucceptibility.Atlantogenata.genomes <-  tree.cancerSucceptibility.Atlantogenata %>% 
+  as_tibble %>% 
+  filter(!is.na(node)) %>%  
+  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% 
+  as.treedata %>% 
+  as.phylo() %>% 
+  treeio::drop.tip(tree.cancerSucceptibility.Atlantogenata@phylo$tip.label[!tree.cancerSucceptibility.Atlantogenata@phylo$tip.label %in% genomeTable.atlantogenata$Species.strict])  %>% as.tibble
+tree.cancerSucceptibility.Atlantogenata.genomes <- left_join(tree.cancerSucceptibility.Atlantogenata.genomes, cancerSucceptibility.Atlantogenata.genomes) %>% as.treedata
+
+p.tree.cancerSucceptibility.Atlantogenata.genomes <- tree.cancerSucceptibility.Atlantogenata.genomes %>% 
+  ggtree(
+    aes(color=as.numeric(abslog2CancerSucceptabilityChange))
+  )
+
+### Double Tree ###
+
+d.dbltree.1 <- p.tree.atlantogenata.genomes$data
+d.dbltree.2 <- p.tree.cancerSucceptibility.Atlantogenata.genomes$data
+
+d.dbltree.2$x <- max(d.dbltree.2$x) - d.dbltree.2$x + max(d.dbltree.1$x) + 500
+
+p.dbltree <-
+  bsize.Atlantogenata.genomes %>% 
+  ggtree(., aes(color=lnSize), size=2) +
+  scale_color_viridis_c(option = "A") + 
+  geom_nodepoint(size=2) + 
+  geom_tippoint(size=2) +
+  geom_tiplab(x=(min(d.dbltree.2$x)+max(d.dbltree.1$x))/2, color="black", hjust = 0.5) +
+  new_scale_color() + 
+  geom_tree(data=d.dbltree.2, aes(color=log2CancerSucceptabilityChange), inherit.aes = F, size=2) + 
+  geom_nodepoint(data=d.dbltree.2, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) + 
+  geom_tippoint(data=d.dbltree.2, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) +
+  # scale_color_viridis_c(option = "D") + 
+    scale_color_distiller(palette = "YlGnBu")+
+    theme(
+      plot.background = element_rect(fill="grey"),
+      panel.background = element_rect(fill="grey"),
+      legend.background = element_rect(fill="grey")
+    )
+
+  p.dbltree
+```
+
+```{r Figure-DoubleTree-Bodysize-CancerSuscept-All}
+
+### Body Size ###
+
+p.tree.atlantogenata.genomes <- bsize.Atlantogenata %>% 
+  ggtree(
+    aes(color=as.numeric(lnSize))
+  )
+
+### Cancer Suscept ###
+
+tree.cancerSucceptibility.Atlantogenata <-  tree.cancerSucceptibility.Atlantogenata %>% 
+  as_tibble %>% 
+  filter(!is.na(node)) %>%  
+  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% 
+  as.treedata
+
+p.tree.cancerSucceptibility.Atlantogenata.2 <- tree.cancerSucceptibility.Atlantogenata %>% 
+  ggtree(
+    aes(color=as.numeric(log2CancerSucceptabilityChange))
+  )
+
+### Double Tree ###
+
+d.dbltree.3 <- p.tree.atlantogenata.genomes$data
+d.dbltree.4 <- p.tree.cancerSucceptibility.Atlantogenata.2$data
+
+d.dbltree.4$x <- max(d.dbltree.4$x) - d.dbltree.4$x + max(d.dbltree.3$x) + 500
+
+p.dbltree.all <-
+  bsize.Atlantogenata %>% 
+  ggtree(., aes(color=lnSize), size=1) +
+  scale_color_viridis_c(option = "A") + 
+  geom_nodepoint(size=2) + 
+  geom_tippoint(size=2) +
+  geom_tiplab(x=(min(d.dbltree.4$x)+max(d.dbltree.3$x))/2, color="black", hjust = 0.5) +
+  new_scale_color() + 
+  geom_tree(data=d.dbltree.4, aes(color=log2CancerSucceptabilityChange), inherit.aes = F, size=1) + 
+  geom_nodepoint(data=d.dbltree.4, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) + 
+  geom_tippoint(data=d.dbltree.4, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) +
+  # scale_color_viridis_c(option = "D") + 
+    scale_color_distiller(palette = "YlGnBu")+
+    theme(
+      plot.background = element_rect(fill="grey"),
+      panel.background = element_rect(fill="grey"),
+      legend.background = element_rect(fill="grey")
+    )
+
+  p.dbltree.all
+```
+
 # References {#references .unnumbered}

### 2020-09-08 paper_PLOS/paper_PLOS_draft.tex
Penultimate draft

--- 2020-05-31 paper_PLOS/paper_PLOS_draft.tex
@@
             pdfauthor={},
-            pdftitle={Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans},
+            pdftitle={Pervasive duplication of tumor suppressors in Afrotherians preceded evolution of large bodies and reduced cancer risk in elephants},
             colorlinks=false,
@@
 \usepackage{pdflscape}
+\usepackage{float}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[H]
+} {
+    \endorigfigure
+}
 \usepackage{booktabs}
@@
 
-  \title{Pervasive duplication of tumor suppressor genes preceded parallel
-evolution of large bodied Atlantogenatans}
+  \title{Pervasive duplication of tumor suppressors in Afrotherians preceded evolution of large bodies and reduced cancer risk in elephants}
     \author[University of Chicago]{Juan Manuel Vazquez}
@@
   \begin{abstract}
-  Cancer is an intrinsic disease of multicellular organisms. Within a
-  species, the size of an animal, - correlated with the individual's
-  number of cells - and its lifespan - correlated with increasing cellular
-  damage over time - positively correlate with the risk any individual has
-  to form tumors. Between species, however, we do not observe any
-  correlation between size, lifespan, and cancer, a phenomenon that
-  referred to as \emph{Peto's Paradox}. Elephants are a particularly
-  interesting member of this class of paradoxical animals, since they are
-  a set of large species deeply nested in a clade of smaller species,
-  indicating a recent gain of size. Recent work has identified several
-  individual cases of gene duplicates contributing to the increased cancer
-  resistance of elephants, which suggests that duplication of tumor
-  suppressor genes may play a more general role in mediating Peto's
-  Paradox by increasing cancer resistance in large, long-lived species. By
-  using a Reciprocal Best-Hit BLAT search approach, we investigated copy
-  numbers of all protein-coding genes in \emph{Atlantogenatan} genomes to
-  see if there is any correlation between the copy number of duplicates
-  and changes body size along the phylogenetic tree. From an initial set
-  of 18,011 protein-coding genes in hg38, we identified a median of 13,880
-  genes in \emph{Atlantogenatan} genomes, of which a median of 940 genes
-  are duplicated. We find that, just as body size fluctuates throughout
-  \emph{Atlantogenata}, genes involved in tumor suppressor pathways are
-  also duplicated throughout the phylogenetic tree. Extant species of
-  elephants, however, show active transcription of both canonical and
-  duplicated copies of tummor suppressors that duplicated prior to and
-  during their sudden increase in body size, suggesting that the
-  duplication of tumor suppressor genes facilitates the evolution of
-  increased body size by compensating for the increased cancer risk.
+  The risk of developing cancer is correlated with body size and lifespan within
+  species. Between species, there is no correlation between body size or
+  lifespan, and cancer, suggesting that large, long-lived species have evolved
+  enhanced cancer protection mechanisms. Elephants and their relatives
+  (Proboscideans) are a particularly interesting lineage in which to explore
+  the mechanisms underlying the evolution of augmented cancer resistance because
+  they evolved large bodies recently within a clade of smaller bodied species
+  (Atlantogenatans). Here we explore the contribution of tumor suppressor
+  duplication to body size and cancer risk in Atlantogenatans. Unexpectedly, we
+  found that tumor suppressor duplication was pervasive in Atlantogenatan
+  genomes rather than restricted to Proboscideans. Proboscideans, however, have
+  duplicates in unique pathways that may underlie some aspects of their
+  remarkable anti-cancer cell biology. These data suggest that duplication of
+  tumor suppressor genes facilitated the evolution of increased body size by
+  compensating for decreasing intrinsic cancer risk.
   \end{abstract}
@@
 
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure0-1} \hfill{}
-
-\caption{Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]}\label{fig:Figure0}
-\end{figure}
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S.\thechapter.\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S.\thechapter.\arabic{figure}}%
+     }
 
@@
 
-One of the major constraints on the evolution of large body sizes in
-animals is an increased risk of developing cancer. If all cells in all
-organisms have a similar risk of malignant transformation and equivalent
-cancer suppression mechanisms, organism with many cells should have a
-higher prevalence of cancer than organisms with fewer cells. Consistent
-with this expectation there is a strong positive correlation between
-body size and cancer incidence within species, for example, human cancer
-incidence increases with increasing adult height {[}1,2{]} and cancer
-incidence is positively correlated with body size in dogs {[}3,4{]}.
-There is no correlation, however, between body size and cancer risk
-between species. This lack of correlation is often referred to as
-`Peto's Paradox' {[}5--7{]}. While it is clear that a resolution to
-Peto's Paradox must involve the evolution of enhanced cancer protection
-alongside increases in body size and lifespan, the specific genetic,
-molecular, and cellular mechanisms that underlie this resistance have
-proven elusive. {[}8--12{]}.
-
-Among the challenges for discovering how animals evolved enhanced cancer
-protection mechanisms is identifying lineages in which large bodied
-species are nested within species with small body sizes. Afrotherian
-mammals are generally small-bodied, similarly to the prediced common
-ancestor of Eutherian mammals. For example, maximum adult weights are
-\textasciitilde{}70g in golden moles, \textasciitilde{}120g in tenrecs,
-\textasciitilde{}170g in elephant shrews, \textasciitilde{}3kg in
-hyraxes, and 60kg in aardvarks {[}13{]}. However, while these extant
-species are relatively small, the fossil evidence demonstrates that
-their ancestral lineages reached enormous sizes. For example, while
-extant hyraxes are relatively small, the extinct Titanohyrax is
-estimated to have weighted up to \textasciitilde{}1300kg {[}14{]}. The
-largest members of Afrotheria, too, are dwarfed by the size of their
-recent ancestors: extant cows manatees are large bodied
-(\textasciitilde{}322-480kg) but are relatively small compared to the
-extinct Stellar's sea cow which is estimated to have weight 8000-10000kg
-{[}15{]}. Similarly African (4,800kg) and Asian elephants (3,200kg) are
-the largest living elephant species, but are dwarfed by the truly
-gigantic extinct Proboscideans such as Deinotherium
-(\textasciitilde{}132,000kg), Mammut borsoni (110,000kg), and the Asian
-straight-tusked elephant (\textasciitilde{}220,000kg), the largest known
-land mammal {[}16{]}. These large-bodied Afrotherian lineages are nested
-within small bodied species (Fig. 1) {[}17--20{]}, indicating that
-gigantism independently evolved in hyraxes, sea cows, and elephants
-(Paenungulates). Thus, Paenungulates are an excellent model system in
-which to explore the mechanisms that underlie the evolution of large
-body sizes and augmented cancer resistance.
-
-Although many mechanisms can potentially resolve Peto's paradox, among
-the most parsimonious routes to enhanced cancer resistance is through an
-increased copy number of tumor suppressors. Such an example has been
-seen in the case of candidate genes such as \emph{TP53} and \emph{LIF}
-{[}12,21,22{]} as well as in studies involving a limited set of
-candidate genes {[}23,24{]}. As these studies focus on \emph{a priori}
-gene sets, however, it remains unknown whether this is a general,
-genome-wide trend in Afrotherian genomes; and whether such a general
-trend is associated with the recent increases in body size -- and
-therefore expected cancer risk -- in these species.
-
-Here, we trace the evolution of body mass and gene copy number variation
-in Afrotherians in order to investigate whether gene duplications are
-enriched in large, long-lived species for genes involved in known tumor
-suppression pathways. Our estimates of the evolution of body mass,
-similarly to previous studies {[}17--20{]}, show that large body masses
-evolved in a step-wise manner, with major increases in body mass in the
-Pseudoungulata (17kg), Paenungulata (25kg), Tethytheria (296kg), and
-Proboscidea (4,100kg) stem-lineages. Furthermore, we see that the
-ancestral body size increases in Hydracoidia and Sirenia were
-independent events. To study the evolution of gene copy number, we used
-a genome-wide Reciprocal Best BLAT Hit (RBBH) method to identify gene
-duplications in Afrotherian genomes, and used maximum likelihood
-(treating copy number as a discrete trait) to infer the lineages in
-which those duplications occurred. We found gene duplications in
-lineages with increased body mass were enriched in functions related to
-tumor suppression, including regulation of the cell cycle, DNA damage
-repair, and regulation of apoptosis. These data suggest that duplication
-of tumor suppressors played a role in the evolution of large, long-lived
-in Afrotherians.
+Among the constraints on the evolution of large bodies and long life-spans in animals is an increased risk of developing cancer.
+If all cells in all organisms have a similar risk of malignant transformation and equivalent cancer suppression mechanisms, then organisms with many cells should have a higher prevalence of cancer than organisms with fewer cells, particularly because large and small animals have similar cell sizes (\emph{1}).
+Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, cancer incidence increases with increasing adult height in humans (\emph{2}, \emph{3}) and overall body size in dogs, cats, and cattle (\emph{4}--\emph{6}).
+There is no correlation, however, between body size and cancer risk between species; this lack of correlation is often referred to as `Peto's Paradox' (\emph{7}--\emph{9}).
+Indeed, cancer prevalence is relatively stable at \textasciitilde5\% across species with diverse body sizes ranging from the minuscule 51g grass mouse to the gargantuan 4800kg African elephant (\emph{10}--\emph{12}).
+While the ultimate resolution to Peto's Paradox is obvious, large bodied and long-lived species evolved enhanced cancer protection mechanisms, identifying and characterizing the proximate genetic, molecular, and cellular mechanisms that underlie the evolution of augmented cancer protection has proven difficult (\emph{13}--\emph{17}).
+
+One of the challenges for discovering how animals evolved enhanced cancer protection mechanisms is identifying lineages in which large bodied species are nested within species with small body sizes.
+\emph{Afrotherian} mammals are generally small-bodied, but also include the largest extant land mammals.
+For example, maximum adult weights are \textasciitilde70g in golden moles, \textasciitilde120g in tenrecs, \textasciitilde170g in elephant shrews, \textasciitilde3kg in hyraxes, and \textasciitilde60kg in aardvarks (\emph{18}).
+In contrast, while extant hyraxes are relatively small, the extinct \emph{Titanohyrax} is estimated to have weighed \textasciitilde1300kg (\emph{19}).
+The largest members of \emph{Afrotheria}, too, are dwarfed by the size of their recent ancestors: extant sea cows such as manatees are large bodied (\textasciitilde322-480kg) but are relatively small compared to the extinct Stellar's sea cow which is estimated to have weight \textasciitilde8000-10000kg (\emph{20}).
+Similarly African (4,800kg) and Asian elephants (3,200kg) are the large, but are dwarfed by the truly gigantic extinct Proboscideans such as \emph{Deinotherium} (\textasciitilde132,000kg), \emph{Mammut borsoni} (110,000kg), and the straight-tusked elephant (\textasciitilde220,000kg), the largest known land mammal (\emph{21}).
+Remarkably, these large-bodied \emph{Afrotherian} lineages are nested deeply within small bodied species (\textbf{Figure \ref{fig:Figure-intro}}) (\emph{22}--\emph{25}), indicating that gigantism independently evolved in hyraxes, sea cows, and elephants (\emph{Paenungulata}).
+Thus, \emph{Paenungulates} are an excellent model system in which to explore the mechanisms that underlie the evolution of large body sizes and augmented cancer resistance.
+
+Many mechanisms have been suggested to resolve Peto's paradox, including a decrease in the copy number of oncogenes, an increase in the copy number of tumor suppressor genes (\emph{7}, \emph{8}, \emph{26}), reduced metabolic rates, reduced retroviral activity and load (\emph{27}), and selection for `cheater' tumors that parasitize the growth of other tumors (\emph{28}), among many others.
+Among the most parsimonious routes to enhanced cancer resistance may be through an increased copy number of tumor suppressors.
+For example, transgenic mice with additional copies of TP53 have reduced cancer rates and extended lifespans but not increased cancer (\emph{29}), respectively, suggesting that changes in the copy number of tumor suppressors can effect cancer rates.
+Indeed, candidate genes studies have found that the elephant genome encodes duplicate tumor suppressors such as TP53 and LIF (\emph{10}, \emph{17}, \emph{30}) as well as other genes with putative tumor suppressive functions (\emph{31}, \emph{32}).
+These studies, however, focused on a priori candidate genes, thus it is unclear whether duplication of tumor suppressor genes is a general phenomenon in the elephant lineage.
+
+Here we trace the evolution of body mass, cancer risk, and gene copy number variation across Afrotherian genomes, including multiple living and extinct Proboscideans, to investigate whether duplication of tumor suppressors occurred coincident with the evolution of large body sizes.
+Our estimates of the evolution of body mass across Afrotheria show that large body masses evolved in a step-wise manner, similar to previous studies (\emph{22}--\emph{25}) and coincident with dramatic reductions in intrinsic cancer risk.
+To explore whether duplication of tumor suppressors occurred coincident with the evolution of large body sizes, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) strategy to identify gene duplications, and used maximum likelihood to infer the lineages in which those duplications occurred.
+Unexpectedly, we found that duplication of tumor suppressor genes was common in Afrotherians, both large and small, but that gene duplications in the Proboscidean lineage were enriched in pathways that may explain some of the unique cancer protection mechanisms observed iin elephant cells.
+These data suggest that duplication of tumor suppressor genes is pervasive in Afrotherians and proceeded the evolution of species with very large body sizes.
 
@@
 \hypertarget{ancestral-body-size-reconstruction}{%
-\subsection{Ancestral Body Size
-Reconstruction}\label{ancestral-body-size-reconstruction}}
-
-We built a time-calibrated supertree of Eutherian mammals by combining
-the time-calibrated molecular phylogeny of Bininda-Emonds \emph{et al.}
-{[}25{]} with the time-calibrated total evidence Afrotherian phylogeny
-from Puttick and Thomas {[}20{]}. While the Bininda-Emonds \emph{et al.}
-{[}25{]} phylogeny includes 1,679 species, only 34 are Afrotherian, and
-no fossil data are included. The inclusion of fossil data from extinct
-species is essential to ensure that ancestral state reconstructions of
-body mass are not biased by only including extant species. This can lead
-to inaccurate reconstructions, for example, if lineages convergently
-evolved large body masses from a small bodied ancestor. In contrast, the
-total evidence Afrotherian phylogeny of Puttick and Thomas {[}20{]}
-includes 77 extant species and fossil data from 39 extinct species.
-Therefore we replaced the Afrotherian clade in the Bininda-Emonds
-\emph{et al.} {[}25{]} phylogeny with the Afrotherian phylogeny of
-Puttick and Thomas {[}20{]} using Mesquite. Next, we jointly estimated
-rates of body mass evolution and reconstructed ancestral states using a
-generalization of the Brownian motion model that relaxes assumptions of
-neutrality and gradualism by considering increments to evolving
-characters to be drawn from a heavy-tailed stable distribution (the
-``Stable Model'') {[}26{]}. The stable model allows for occasional large
-jumps in traits and has previously been shown to out-perform other
-models of body mass evolution, including standard Brownian motion
-models, Ornstein--Uhlenbeck models, early burst maximum likelihood
-models, and heterogeneous multi-rate models {[}26{]}.
+\subsection{Ancestral Body Size Reconstruction}\label{ancestral-body-size-reconstruction}}
+
+We first built a time-calibrated supertree of Eutherian mammals by combining the time-calibrated molecular phylogeny of Bininda-Emonds et al.~(\emph{33}) with the time-calibrated total evidence Afrotherian phylogeny from Puttick and Thomas (\emph{25}).
+While the Bininda-Emonds et al.~(\emph{33}) phylogeny includes 1,679 species, only 34 are Afrotherian, and no fossil data are included.
+The inclusion of fossil data from extinct species is essential to ensure that ancestral state reconstructions of body mass are not biased by only including extant species.
+This can lead to inaccurate reconstructions, for example, if lineages convergently evolved large body masses from a small bodied ancestor.
+In contrast, the total evidence Afrotherian phylogeny of Puttick and Thomas (\emph{25}) includes 77 extant species and fossil data from 39 extinct species.
+Therefore we replaced the Afrotherian clade in the Bininda-Emonds et al.~(\emph{33}) phylogeny with the Afrotherian phylogeny of Puttick and Thomas (\emph{25}) using Mesquite.
+Next, we jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of the Brownian motion model that relaxes assumptions of neutrality and gradualism by considering increments to evolving characters to be drawn from a heavy-tailed stable distribution (the ``Stable Model'') implemented in StableTraits (\emph{34}).
+The stable model allows for large jumps in traits and has previously been shown to out-perform other models of body mass evolution, including standard Brownian motion models, Ornstein--Uhlenbeck models, early burst maximum likelihood models, and heterogeneous multi-rate models (\emph{34}).
 
 \hypertarget{identification-of-duplicate-genes}{%
-\subsection{Identification of Duplicate
-Genes}\label{identification-of-duplicate-genes}}
-
-\textbf{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit
-BLAT (RBHB) pipeline to identify putative homologs and estimate gene
-copy numbers across species (\textbf{Figure 1A}). The Reciprocal Best
-Hit (RBH) search strategy is conceptually straightforward: 1) Given a
-gene of interest \(G_A\) in a query genome \(A\), one searches a target
-genome \(B\) for all possible matches to \(G_A\); 2) For each of these
-hits, one then performs the reciprocal search in the original query
-genome to identify the highest-scoring hit; 3) A hit in genome \(B\) is
-defined as a homolog of gene \(G_A\) if and only if the original gene
-\(G_A\) is the top reciprocal search hit in genome \(A\). We selected
-BLAT {[}27{]} as our algorithm of choice, as this algorithm is sensitive
-to highly simliar (\textgreater{}90\% identity) sequences, thus
-identifying the highest-confidence homologs while minimizing many-to-one
-mapping problems when searching for multiple genes. RBH performs similar
-to other more complex methods of orthology prediction, and is
-particularly good at identifying incomplete genes that may be fragmented
-in low quality/poor assembled regions of the genome {[}28,29{]}.
-
-\textbf{Effective Copy Number By Coverage:} In lower-quality genomes,
-many genes are fragmented across multiple scaffolds, which results in
-BLAT calling multiple hits when in reality there is only one gene. To
-compensate for this, we came up with a novel statistic, Estimated Copy
-Number by Coverage (ECNC), which averages the number of times we see
-each nucleotides of a query sequence in a target genome over the total
-number of nucleotides of the query sequence found overall in each target
-genome (Supplementary Figure 1). This allows us to correct for genes
-that have been fragmented across incomplete genomes, while also taking
-into account missing sequences from the human query in the target
-genome. Mathematically, this can be written as:\\
-\[ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}\] where
-\(n\) is a given nucleotide in the query, \(l\) is the total length of
-the query, \(C_n\) is the number of instances that \(n\) is present
-within a reciprocal best hit, and \(bool(C_n)\) is 1 if \(C_n > 0\) or 0
-if \(C_n = 0\).
-
-\textbf{RecSearch Pipeline:} We created a custom Python pipeline for
-automating RBHB searches between a single reference genome and multiple
-target genomes using a list of query sequences from the reference
-genome. For the query sequences in our search, we used the hg38 Proteome
-provided by UniProt {[}30{]}, which is a comprehensive set of protein
-sequences curated from a combination of predicted and validated protein
-sequences generated by the UniProt Consortium. In order to refine our
-search, we omitted protein sequences originating from long, noncoding
-RNA loci (e.g.~LINC genes); poorly-studied genes from predicted open
-reading frames (C-ORFs); and sequences with highly repetitive sequences
-such as zinc fingers, protocadherins, and transposon-containing genes,
-as these were prone to high levels of false positive hits.
-
-After filtering out problematic protein queries (see ``Query gene
-inclusion criteria''), we then used our pipeline (Figure 1A) to search
-for all copies of our 18011 query genes in publicly available
-Afrotherian genomes, including African savannah elephant
-(\emph{Loxodonta africana}: loxAfr3, loxAfr4, loxAfrC), African forest
-elephant (\emph{Loxodonta cyclotis}: loxCycF), Asian Elephant
-(\emph{Elephas maximus}: eleMaxD), Woolly Mammoth (\emph{Mammuthus
-primigenius}: mamPriV), Colombian mammoth (\emph{Mammuthus columbi}:
-mamColU), American mastodon (\emph{Mammut americanum}: mamAmeI), Rock
-Hyrax (\emph{Procavia capensis}: proCap1, proCap2, proCap2\_HiC), West
-Indian Manatee (\emph{Trichechus manatus latirostris}: triManLat1,
-triManLat1\_HiC), Aardvark (\emph{Orycteropus afer}: oryAfe1,
-oryAfe1\_HiC), Lesser Hedgehog Tenrec (\emph{Echinops telfairi}:
-echTel2), Nine-banded armadillo (\emph{Dasypus novemcinctus}: dasNov3),
-Hoffman's two-toed sloth (\emph{Choloepus hoffmannii}: choHof1, choHof2,
-choHof2\_HiC), Cape golden mole (\emph{Chrysochloris asiatica}:
-chrAsi1), and Cape elephant shrew (\emph{Elephantulus edwardii}:
-eleEdw1). For many of these species, we covered multiple assemblies in
-order to test the effects of assembly size and quality on our hits.
-
-\textbf{Query gene inclusion criteria:} To assemble our query list, we
-first removed all unnamed genes from UP000005640. Next, we excluded
-genes from downstream analyses for which assignment of homology was
-uncertain, including uncharacterized ORFs (991), LOC (63), HLA genes
-(402), replication dependent histones (72), odorant receptors (499),
-ribosomal proteins (410), zinc finger transcription factors (1983),
-viral and repetitive-element-associated proteins (82) and any protein
-described as either ``Uncharacterized,'' ``Putative,'' or ``Fragment''
-by UniProt in UP000005640 (30724), leaving us with a final set of 37582
-query protein sequences, corresponding to 18011 genes.
-
-\textbf{Duplication gene inclusion criteria:} In order to condense
-transcript-level hits into single gene loci, and to resolve many-to-one
-genome mappings, we removed exons where transcripts from different genes
-overlapped, and merged overlapping transcripts of the same gene into a
-single gene locus call. The resulting gene-level copy number table was
-then combined with the maximum ECNC values observed for each gene in
-order to call gene duplications. We called a gene duplicated if its copy
-number was two or more, and if the maximum ECNC value of all the gene
-transcripts searched was 1.5 or greater; previous studies have shown
-that incomplete duplications can encode functional genes {[}12,22{]},
-therefore partial gene duplications were included provided they passed
-additional inclusion criteria. The ECNC cut off of 1.5 was selected
-empirically, as this value minimized the number of false positives seen
-in a test set of genes and genomes. The results of our initial search
-are summarized in Figure 1B. Overall, we identified 13880 genes across
-all species, or 77.1\% of our starting query genes.
-
-\textbf{Genome Quality Assessment using CEGMA:} In order to determine
-the effect of genome quality on our results, we used the gVolante
-webserver and CEGMA to assess the quality and completeness of the genome
-{[}31,32{]}. CEGMA was run using the default settings for mammals
-(``Cut-off length for sequence statistics and composition'' = 1;``CEGMA
-max intron length'' = 100000; ``CEGMA gene flanks'' = 10000, ``Selected
-reference gene set'' = CVG). For each genome, we generated a correlation
-matrix using the aforementioned genome quality scores, and either the
-mean Copy Number or mean ECNC for all hits in the genome.
-
-\hypertarget{evidence-for-functionality-of-identified-genes}{%
-\subsection{Evidence for Functionality of Identified
-Genes}\label{evidence-for-functionality-of-identified-genes}}
-
-To validate and filter out duplicate gene calls, we intersected our
-results with either gene prediction or transcriptomic evidence as a
-proxy for functionality.
-
-\textbf{Transcriptome Assembly:} For the African Savana Elephant, Asian
-Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated
-\emph{de novo} transcriptomes using publically-available RNA-sequencing
-data from NCBI SRA (Supplementary Table 1). We mapped reads to all
-genomes available for each species, and assembled transcripts using
-HISAT2 and StringTie, respectively {[}33--35{]}. RNA-sequencing data was
-not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax,
-Aardvark, or the Lesser Hedgehog Tenrec.
-
-\textbf{Gene Prediction:} We obtained tracks for genes predicted using
-GenScan for all the genomes available via UCSC Genome Browser: African
-savannah elephant (loxAfr3), Rock Hyrax (proCap1), West Indian Manatee
-(triManLat1), Aardvark (oryAfe1), Lesser Hedgehog Tenrec (echTel2),
-Nine-banded armadillo (dasNov3), Hoffman's Two-Toed Sloth (choHof1),
-Cape golden mole (chrAsi1), and Cape Elephant Shrew (eleEdw1); gene
-prediction tracks for higher-quality assemblies were not available.
-
-\textbf{Evidenced Duplicate Criteria:} We intersected our records of
-duplicate hits identified in each genome with the gene prediction tracks
-and/or transcriptome assemblies using \texttt{bedtools}. When multiple
-lines of evidence for functionality were present for a genome, we used
-the union of all intersections as the final output for evidenced
-duplicates. When analyzing the highest-quality assemblies available for
-each species, if a species had neither gene prediction tracks nor
-RNA-seq data for the highest-quality genome available, we conservatively
-included all hits for the genome in the final set of evidenced
-duplicates.
+\subsection{Identification of Duplicate Genes}\label{identification-of-duplicate-genes}}
+
+\emph{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit BLAT (RBHB) pipeline to identify putative homologs and estimate gene copy number across species.
+The Reciprocal Best Hit (RBH) search strategy is conceptually straightforward: 1) Given a gene of interest \(G_A\) in a query genome \(A\), one searches a target genome \(B\) for all possible matches to \(G_A\); 2) For each of these hits, one then performs the reciprocal search in the original query genome to identify the highest-scoring hit; 3) A hit in genome \(B\) is defined as a homolog of gene \(G_A\) if and only if the original gene \(G_A\) is the top reciprocal search hit in genome \(A\).
+We selected BLAT (\emph{35}) as our algorithm of choice, as this algorithm is sensitive to highly simliar (\textgreater90\% identity) sequences, thus identifying the highest-confidence homologs while minimizing many-to-one mapping problems when searching for multiple genes.
+RBH performs similar to other more complex methods of orthology prediction, and is particularly good at identifying incomplete genes that may be fragmented in low quality/poor assembled regions of the genome (\emph{36}, \emph{37}).
+
+\emph{Effective Copy Number By Coverage:} In low-quality genomes, many genes are fragmented across multiple scaffolds, which results in BLA(S)T-like methods calling multiple hits when in reality there is only one gene.
+To compensate for this, developed a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we hit each nucleotide of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (\textbf{Supplemental Figure \ref{fig:Figure-ECNC}}).
+This allows us to correct for genes that have been fragmented across incomplete genomes, while also taking into account missing sequences from the human query in the target genome.
+Mathematically, this can be written as:
+
+\[ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}\]
+
+where \(n\) is a given nucleotide in the query, \(l\) is the total length of the query, \(C_n\) is the number of instances that \(n\) is present within a reciprocal best hit, and \(bool(C_n)\) is 1 if \(C_n > 0\) or 0 if \(C_n = 0\).
+
+\emph{RecSearch Pipeline:} We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome.
+For the query sequences in our search, we used the hg38 UniProt proteome (\emph{38}), which is a comprehensive set of protein sequences curated from a combination of predicted and validated protein sequences generated by the UniProt Consortium.
+In order to refine our search, we omitted protein sequences originating from long, noncoding RNA loci (e.g.~LINC genes); poorly-studied genes from predicted open reading frames (C-ORFs); and sequences with highly repetitive sequences such as zinc fingers, protocadherins, and transposon-containing genes, as these were prone to high levels of false positive hits.
+After filtering out problematic protein queries, we then used our pipeline to search for all copies of 18011 query genes in publicly available \emph{Afrotherian} genomes (\ref{tab:Table-Genomes}), including African savannah elephant (\emph{Loxodonta africana}: loxAfr3, loxAfr4, loxAfrC), African forest elephant (\emph{Loxodonta cyclotis}: loxCycF), Asian Elephant (\emph{Elephas maximus}: eleMaxD), Woolly Mammoth (\emph{Mammuthus primigenius}: mamPriV), Colombian mammoth (\emph{Mammuthus columbi}: mamColU), American mastodon (\emph{Mammut americanum}: mamAmeI), Rock Hyrax (\emph{Procavia capensis}: proCap1, proCap2, proCap2\emph{HiC), West Indian Manatee (}Trichechus manatus latirostris\emph{: triManLat1, triManLat1}HiC), Aardvark (\emph{Orycteropus afer}: oryAfe1, oryAfe1\emph{HiC), Lesser Hedgehog Tenrec (}Echinops telfairi\emph{: echTel2), Nine-banded armadillo (}Dasypus novemcinctus\emph{: dasNov3), Hoffman's two-toed sloth (}Choloepus hoffmannii\emph{: choHof1, choHof2, choHof2}HiC), Cape golden mole (\emph{Chrysochloris asiatica}: chrAsi1), and Cape elephant shrew (\emph{Elephantulus edwardii}: eleEdw1).
+
+\emph{Query gene inclusion criteria:} To assemble our query list, we first removed all unnamed genes from UP000005640.
+Next, we excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (991 genes), LOC (63 genes), HLA genes (402 genes), replication dependent histones (72 genes), odorant receptors (499 genes), ribosomal proteins (410 genes), zinc finger transcription factors (1983 genes), viral and repetitive-element-associated proteins (82 genes) and any protein described as either ``Uncharacterized,'' ``Putative,'' or ``Fragment'' by UniProt in UP000005640 (30724 genes), leaving us with a final set of 37,582 query protein isoforms, corresponding to 18,011 genes.
+
+\emph{Duplication gene inclusion criteria:} In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call.
+The resulting gene-level copy number table was then combined with the maximum ECNC values observed for each gene in order to call gene duplications.
+We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes (\emph{17}, \emph{30}), therefore partial gene duplications were included provided they passed additional inclusion criteria (see below).
+The ECNC cut-off of 1.5 was selected empirically, as this value minimized the number of false positives seen in a test set of genes and genomes.
+The results of our initial search are summarized in \textbf{Figure @ref(fig: Figure-ML-UpSet-WC)A}.
+Overall, we identified 13880 genes across all species, or 77.1\% of our starting query genes.
+
+\emph{Genome Quality Assessment using CEGMA:} In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome (\emph{39}, \emph{40}).
+CEGMA was run using the default settings for mammals (``Cut-off length for sequence statistics and composition'' = 1;``CEGMA max intron length'' = 100000; ``CEGMA gene flanks'' = 10000, ``Selected reference gene set'' = CVG).
+For each genome, we generated a correlation matrix using the aforementioned genome quality scores, and either the mean Copy Number or mean ECNC for all hits in the genome.
+We observed that the percentage of duplicated genes in non-\emph{Pseudoungulatan} genomes was higher (12.94\% to 23.66\%) than \emph{Pseudoungulatan} genomes (3.26\% to 7.80\%).
+Mean Copy Number, mean ECNC, and mean CN (the lesser of Copy Number and ECNC per gene) moderately or strongly correlated with genomic quality, such as LD50, the number of scaffolds, and contigs with a length above either 100K or 1M (\textbf{Supplemental Figure \ref{fig:Figure-GenomeCor}}).
+The Afrosoricidians had the greatest correlation between poor genome quality and high gene duplication rates including larger numbers of private duplications.
+The correlations between genome quality metric and number of gene duplications was particularly high for Cape golden mole (\emph{Chrysochloris asiatica}: chrAsi1) and Cape elephant shrew (\emph{Elephantulus edwardii}: eleEdw1), therefore we excluded these species from downstream pathway enrichment analyses.
+
+\hypertarget{evidence-for-functionality-of-gene-duplicates}{%
+\subsection{Evidence for Functionality of Gene Duplicates}\label{evidence-for-functionality-of-gene-duplicates}}
+
+To validate and filter out duplicate gene calls, we intersected our results with either gene prediction or transcriptomic evidence as a proxy for functionality.
+
+\emph{Transcriptome Assembly}: For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated de novo transcriptomes using publically-available RNA-sequencing data from NCBI SRA (Supplemental Table \ref{tab:Table-SRAData}).
+We mapped reads to all genomes available for each species, and assembled transcripts using HISAT2 and StringTie, respectively (\emph{41}--\emph{43}).
+RNA-sequencing data was not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax, Aardvark, or the Lesser Hedgehog Tenrec.
+
+\emph{Gene Prediction}: We downloaded tracks for genes predicted using GenScan for all the genomes available via UCSC Genome Browser: African savannah elephant (loxAfr3), Rock Hyrax (proCap1), West Indian Manatee (triManLat1), Aardvark (oryAfe1), Lesser Hedgehog Tenrec (echTel2), Nine-banded armadillo (dasNov3), Hoffman's Two-Toed Sloth (choHof1), Cape golden mole (chrAsi1), and Cape Elephant Shrew (eleEdw1); gene prediction tracks for higher-quality assemblies were not available.
+
+\emph{Evidenced Duplicate Criteria}: We intersected our records of duplicate hits identified in each genome with the gene prediction tracks and/or transcriptome assemblies using bedtools.
+When multiple lines of evidence for functionality were present for a genome, we used the union of all intersections as the final output for evidenced duplicates.
+When analyzing the highest-quality assemblies available for each species, if a species had neither gene prediction tracks nor RNA-seq data for the highest-quality genome available, we conservatively included all hits for the genome in the final set of evidenced duplicates.
 
 \hypertarget{reconstruction-of-ancestral-copy-numbers}{%
-\subsection{Reconstruction of Ancestral Copy
-Numbers}\label{reconstruction-of-ancestral-copy-numbers}}
-
-We implemented a maximum likelihood method for determining the ancestral
-copy numbers of genes in \emph{Atlantogenata} using IQ-Tree. For this
-analysis, we used an unrooted species tree including only the
-aforementioned \emph{Atlantogenata} species. We generated PHYLIP files
-containing the copy number of each gene in the highest quality genome
-for each species, encoding genes on a scale from 1-31+ copies as 1-9,
-A-V; and encoding a gene's copy number as uncetain (``?'') when we did
-not identify it in the genome. We used the included tree-searching and
-model-testing functionality in IQ-Tree to determine the most likely
-topology for the species tree, and to obtain the most likely model for
-copy number changes in the genome. The most likely model for the
-evolution of copy number was inferred to be a Jukes-Cantor type model
-for morphological data with equal state frequencies, allowing for a
-proportion of invariable sites, and using a discrete Gamma model with 4
-rate categories. (``MK+FQ+I+G4''). We defined the ancestral state of a
-node if it had greater than an 80\% posterior probability.
+\subsection{Reconstruction of Ancestral Copy Numbers}\label{reconstruction-of-ancestral-copy-numbers}}
+
+We encoded the copy number of each gene for each species as a discrete trait ranging from 0 (one gene copy) to 31 (for 32+ gene copies) and used IQ-TREE to select the best-fitting model of character evolution (\emph{44}--\emph{48}), which was inferred to be a Jukes-Cantor type model for morphological data (MK) with equal character state frequencies (FQ) and rate heterogeneity across sites approximated by including a class of invariable sites (I) plus a discrete Gamma model with four rate categories (G4).
+Next we inferred gene duplication and loss events with the empirical Bayesian ancestral state reconstruction (ASR) method implemented in IQ-TREE (\emph{44}--\emph{48}), the best fitting model of character evolution (MK+FQ+GR+I) (\emph{49}, \emph{50}), and the unrooted species tree for \emph{Atlantogenata}.
+We considered ancestral state reconstructions to be reliable if they had Bayeisan Posterior Probability (BPP) \(\geq\) 0.80; less reliable reconstructions were excluded from pathway analyses.
 
 \hypertarget{pathway-enrichment-analysis}{%
-\subsection{Pathway Enrichment
-Analysis}\label{pathway-enrichment-analysis}}
-
-To determine which pathways were associated with duplicated genes in
-each species and lineage, we used WEBGESTALT to perform
-overrepresentation analysis (ORA) of the duplicated gene lists relative
-to our initial query gene list {[}36{]}. For the database of pathways
-used in the analysis, we used Reactome {[}37{]}, Wikipathways, {[}38{]},
-and KEGG {[}39{]}. For the ORA, we used FDR for determining
-significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and
-FDR=0.5.
-
-\hypertarget{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}{%
-\subsection{Lifespan Phylogenetic Least-Square Regression and
-Calculating Estimated Cancer Risk Thoughout
-Atlantogenata}\label{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}}
-
-In order to determine the cancer risk \(K\) at each node, we first
-needed to calcultate ancestral lifespans at each node. To do so, we used
-a Phylogenetic Generalized Least-Square Regression (PGLS) {[}40,41{]} to
-calculate estimated ancestral lifespans across \emph{Atlantogenata}
-using our estimates for body size at each node.
-
-Next, we used a simplified multistage cancer risk model for body size
-\(D\) and lifespan \(t\): \(K \approx Dt^6\) {[}7,42--44{]}. To
-determine the change in \(K\) between nodes, we obtained the ratio
-between the cancer risk \(K_{1}\) at any given node, and the cancer risk
-\(K_{2}\) at its ancestral node, using the equation
-\(\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}\).
-Finally, to simplify comparisons, we calculated the fold change cancer
-risk between a node and its ancestor as
-\(\log_{2}(\frac{K_{2}}{K_{1}})\).
+\subsection{Pathway Enrichment Analysis}\label{pathway-enrichment-analysis}}
+
+To determine if gene duplications were enriched in particular biological pathways, we used the WEB-based Gene SeT AnaLysis Toolkit (WebGestalt)(\emph{51}) to perform Over-Representation Analysis (ORA) using the Reactome database (\emph{52}).
+Gene duplicates in each lineage were used as the foreground gene set, and the initial query set was used as the background gene set.
+WebGestalt uses a hypergeometric test for statistical significance of pathway over-representation, which we refined using two methods: an False Discovery Rate (FDR)-based approach, and an empirical p-value approach (\emph{53}).
+The Benjamini-Hochberg FDR multiple-testing correction was generated by WebGestalt.
+In order to correct P-values based on an empirical distribution, we modified the approach used by Chen \emph{et al.} in Enrichr (\emph{53}) to generate a ``combined score'' for each pathway based on the hypergeometric P-value from WebGestalt, and a correction for expected rank for each pathway.
+In order to generate the table of expected ranks and variances for this approach, we randomly sampled foreground sets of 10-5,000 genes from our background set 5000 times, and used WebGestalt ORA to obtain a list of enriched terms and P-values for each run; we then compiled a table of Reactome terms with their expected frequencies and standard deviation.
+This data was used to calculate a Z-score for terms in an ORA run, and the combined score was calculated using the formula \(C = \log{(p)}\cdot{}z\).
+
+\hypertarget{estimating-the-evolution-of-cancer-risk}{%
+\subsection{Estimating the Evolution of Cancer Risk}\label{estimating-the-evolution-of-cancer-risk}}
+
+The dramatic increase in body mass and lifespan in some Afrotherian lineages, and the relatively constant rate of cancer across species of diverse body sizes (\emph{10}), indicates that those lineages must have also evolved reduced cancer risk.
+To infer the magnitude of these reductions we estimated differences in intrinsic cancer risk across extant and ancestral Afrotherians.
+Following Peto (\emph{54}), we estimate the intrinsic cancer risk (\(K\)) as the product of risk associated with body mass and lifespan.
+In order to determine (\(K\)) across species and at ancestral nodes (see below), we first estimated ancestral lifespans at each node.
+We used Phylogenetic Generalized Least-Square Regression (PGLS) (\emph{55}, \emph{56}), using a Brownian covariance matrix as implemented in the R package \texttt{ape} (\emph{57}), to calculate estimated ancestral lifespans across \emph{Atlantogenata} using our estimates for body size at each node.
+In order to estimate the intrinsic cancer risk of a species, we first inferred lifespans at ancestral nodes using PGLS and the model \(ln(lifespan) = \beta_{1}corBrowninan + \beta_{2}ln(size) + \epsilon\).
+Next, we calculated \(K_{1}\) at all nodes, and then estimated the fold-change in cancer susceptibility between ancestral and descendant nodes (\textbf{Figure @ref(fig:Figure-DoubleTree-Bodysize-Lifespan}).
+Next, in order to calculate \(K_{1}\) at all nodes, we used a simplified multistage cancer risk model for body size \(D\) and lifespan \(t\): \(K \approx Dt^6\) (\emph{9}, \emph{54}, \emph{58}, \emph{59}).
+The fold change in cancer risk between a node and its ancestor was then defined as \(\log_{2}(\frac{K_{2}}{K_{1}})\).
+
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
+
+All data analysis was performed using R version 4.0.2 (2020-06-22), and the complete reproducible manuscript along with code and data generation pipeline can be found on the author's GitHub page at \url{www.github.com/docmanny/smRecSearch/tree/publication} (\emph{57}, \emph{60}--\emph{104})
 
@@
 
-\hypertarget{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}{%
-\subsection{\texorpdfstring{Body size frequently and independently
-expands and contracts throughout
-\emph{Atlantogenata}}{Body size frequently and independently expands and contracts throughout Atlantogenata}}\label{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}}
-
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure1-1} \hfill{}
-
-\caption{Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
-**A)** Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch.
-Atlantogenata is highlighted at the bottom.
-**B)** Zoom-in of (**A**) on Atlantogenata.
-Silhuetes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size.
-**C)** Confidence interval plot for representative species and ancestral nodes.}\label{fig:Figure1}
+\hypertarget{step-wise-evolution-of-body-size-in-afrotherians}{%
+\subsection{Step-wise evolution of body size in Afrotherians}\label{step-wise-evolution-of-body-size-in-afrotherians}}
+
+Similar to previous studies of Afrotherian body size (\emph{25}, \emph{34}), we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95\% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a 67.36x increase in body mass in the stem-lineage of \emph{Pseudoungulata} (17.33kg), a 1.45x increase in body mass in the stem-lineage of \emph{Paenungulata} (25.08kg), a 11.82x increase in body mass in the stem-lineage of \emph{Tehthytheria} (296.56kg), and a 2.69x increase in body mass in the stem-lineage of \emph{Proboscidea} (4114.39kg) (\textbf{Figure \ref{fig:Figure-DoubleTree-Bodysize-Lifespan}A,B}).
+The ancestral \emph{Hyracoidea} was inferred to be relatively small (2.86kg-118.18kg), and rate accelerations were coincident with independent body mass increases in large hyraxes such as \emph{Titanohyrax andrewsi} (429.34kg, 67.36x increase).
+While the body mass of the ancestral \emph{Sirenian} was inferred to be large (61.7kg-955.51kg), a rate acceleration occurred coincident with a 10.59x increase in body mass in Stellar's sea cow.
+Rate accelerations also occurred coincident with 36.6x decrease in body mass in the stem-lineage of the dwarf elephants \emph{Elephas (Palaeoloxodon) antiquus falconeri} and \emph{Elephas cypriotes}.
+These data suggest that gigantism in \emph{Afrotherians} evolved step-wise, from small to medium bodies in the \emph{Pseudoungulata} stem-lineage, medium to large bodies in the \emph{Tehthytherian} stem-lineage and extinct hyraxes, and from large to exceptionally large bodies independently in the \emph{Proboscidean} stem-lineage and Stellar's sea cow (\textbf{Figure \ref{fig:Figure-DoubleTree-Bodysize-Lifespan}A,B}).
+
+\hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
+\subsection{Step-wise reduction of intrinsic cancer risk in large, long-lived Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}
+
+In order to account for a relatively stable cancer rate across species (\emph{10}--\emph{12}), intrinsic cancer risk must also evolve with changes body size and lifespan across species.
+As expected, intrinsic cancer susceptibility in \emph{Afrotheria} also varies with changes in body size and longevity (\textbf{Figure \ref{fig:Figure-DoubleTree-Bodysize-Lifespan}A}), with initial 0-log\_\{2\}-fold and 6.41-log\_\{2\}-fold decreases in the stem-lineage of Afrotheria and Xenarthra, respectively, followed by a 13.37-log\_\{2\}-fold decrease in Pseudoungulata, and a 1.49-log\_\{2\}-fold decrease in Aardvarks (\textbf{Figure \ref{fig:Figure-DoubleTree-Bodysize-Lifespan}A}).
+In contrast to the Paenungulate stem-lineage, there is a 7.84-log\_\{2\}-fold decrease in cancer risk in Tethytheria, a 0.67-log\_\{2\}-fold decrease in Manatee, and dramatic increases within Proboscidea including a 0.83-log\_\{2\}-fold decrease in Elephantidae and a 0.57-log\_\{2\}-fold decrease in the American Mastodon.
+Within the Elephantidae, Elephantina and Loxodontini have a 0.06-log\_\{2\}-fold decrease in cancer susceptibility, while susceptibility is relatively stable in Mammoths.
+The three extant Proboscideans, Asian Elephant, African Savana Elephant, and the African Forest Elephant, meanwhile, have similar decreases in body size, with increases in cancer susceptibility (\textbf{Figure \ref{fig:Figure-DoubleTree-Bodysize-Lifespan}A}).
+
+\hypertarget{pervasive-duplication-of-tumor-suppressor-genes-in-afrotheria}{%
+\subsection{Pervasive duplication of tumor suppressor genes in Afrotheria}\label{pervasive-duplication-of-tumor-suppressor-genes-in-afrotheria}}
+
+Our hypothesis was that genes which duplicated coincident with the evolution of increased body sizes (IBM) and reduced intrinsic cancer risk (RICR) would be uniquely enriched in tumor suppressor pathways compared to genes that duplicated in other lineages.
+Therefore, we tested if duplicated genes in each lineage were enriched in Reactome pathways using ORA implemented in WEBGESTALT.
+No pathways obviously related to cancer biology or tumor suppression were enriched in either the stem-lineage of \emph{Pseudoungulata} (67.36-fold IBM, 13.37-log\_\{2\}-fold RICR) or \emph{Paenungulata} (1.45-fold IBM, 1.17-log\_\{2\}-log\_\{2\}-fold RICR) (\textbf{Figure \ref{fig:Figure-DoubleTree-Bodysize-Lifespan}A}).
+Consistent with our hypothesis, 55.8\% (29/52) of the pathways that were enriched in the Tethytherian stem-lineage (11.82-fold IBM, 7.84-log\_\{2\}-fold RICR), 27.8\% (20/72) of the pathways that were enriched in the Proboscidean stem-lineage (1.06-fold IBM, 3.14-log\_\{2\}-fold RICR), and 28\% (33/118) of the pathways that were enriched within the Proboscidean clade were related to tumor suppression (\textbf{Figure \ref{fig:Figure-DoubleTree-Bodysize-Lifespan}C}).
+Similarly, 17.8\% (10/56) and 30\% (30/100) of the pathways that were enriched in manatee (1.11-fold IBM, 0.25, 0.22, 0.32, 2.29, 0.98-log\_\{2\}-fold RICR) and aardvark (67.36-fold IBM, 1.49-log\_\{2\}-fold RICR), respectively, were related to tumor suppression.
+In contrast, only 4.9\% (2/41) of the pathways that were enriched in hyrax (1.6-fold IBM, 1.49-log\_\{2\}-fold RICR) were related to tumor suppression (\textbf{Figure \ref{fig:Figure-ML-UpSet-WC}}).
+
+Unexpectedly, however, lineages without major increases in body size, lifespan, or intrinsic cancer risk were also enriched for tumor suppressor pathways.
+For example, 13.2\% (9/68), 36.1\% (13/36), and 20\% (20/100) of the pathways that were enriched in the stem-lineages of Afroinsectivoa and Afrosoricida, and in E.
+telfairi, respectively, were related to tumor suppression (\textbf{Figure \ref{fig:Figure-ML-UpSet-WC}}).
+These data suggest that either duplication of genes in tumor suppressor pathways is common in Afrotherians, or that there is a systemic bias in our pathways enrichment analyses.
+To explore this possibility further we generated 5000 randomly sampled gene sets of sizes between 10 and 5000 genes from the query gene set, and tested for enriched pathways using ORA implemented in WebGestalt.
+After filtering out pathways with a median p-value greater than 0.05, we saw no enrichments of cancer pathways above 157 genes; below that number, between 12.2\% and 18\% of enriched pathways were classified as cancer pathways by our criteria.
+Without considering p-values, the percentage of cancer pathways approaches 15.4\% (213/1381) in simulated sets as the number of genes in a set increases.
+When directly comparing our simulated results with the \emph{Atlantogenatan} results by gene set size, we see that only cancer pathway enrichments in \emph{Mammuthus columbi}, \emph{Paenungulata}, \emph{Elephantidae}, \emph{Loxodonta cyclotis}, \emph{Afrosoricida}, \emph{Tethytheria}, \emph{Elephas maximus}, \emph{Loxodonta africana}, \emph{Proboscidea}, \emph{Trichechus manatus}, \emph{Orycteropus afer}, \emph{Echinops telfairi}, \emph{Chrysochloris asiatica}, and \emph{Elephantulus edwardii} are above their simulated cancer pathway percentages.
+
+\hypertarget{tumor-suppressor-pathways-enriched-exclusively-within-proboscideans}{%
+\subsection{Tumor suppressor pathways enriched exclusively within Proboscideans}\label{tumor-suppressor-pathways-enriched-exclusively-within-proboscideans}}
+
+While duplication of tumor suppressor genes is common in Afrotheria, we identified 12 pathways related to tumor suppression that were enriched exclusively in the Proboscidean stem-lineage (\textbf{Figure \ref{fig:Figure-ML-UpSet-WC}C}).
+Among these were pathways related to the cell cycle, including ``G0 and Early G1'', ``G2/M Checkpoints'' and ``Phosphorylation of the APC/C'', pathways related to DNA damage repair including ``Global Genome Nucleotide Excision Repair (GG-NER)'', ``HDR through Single Strand Annealing (SSA)'', ``Gap-filling DNA repair synthesis and ligation in GG-NER'', ``Recognition of DNA damage by PCNA-containing replication complex'', and ``DNA Damage Recognition in GG-NER'', pathways related to telomere biology including ``Extension of Telomeres'' and ``Telomere Maintenance'', pathways related to the apoptosome including ``Activation of caspases through apoptosome-mediated cleavage'', pathways related to ``mTORC1-mediated signaling'' and ``mTOR signaling''.
+Thus we conclude that duplication of genes in tumor suppressor pathways is pervasive in Afrotherians, but that genes in some pathways related to cancer biology and tumor suppression are uniquely duplicated in large-bodied (or long-lived) Proboscideans.
+
+Among the genes uniquely duplicated within Proboscideans are TP53, COX20, LAMTOR5, PRDX1, STK11, BRD7, MAD2L1, BUB3, UBE2D1, SOD1, LIF, MAPRE1, CNOT11, CASP9, CD14, HMGB2.
+Two of these, TP53 and LIF, have been previously described (\emph{10}, \emph{17}, \emph{30}).
+These genes are significantly enriched in pathways involved in apoptosis, cell cycle regulation, and both upstream and downstream pathways involving TP53.
+The majority of these genes are actively expressed in extant African Elephants, suggesting that they maintain functionality (\textbf{Figure \ref{fig:Figure-LoxAfrExpressedDups}D}).
+
+\hypertarget{coordinated-duplication-of-tp53-related-genes-in-proboscidea}{%
+\subsection{Coordinated duplication of TP53-related genes in Proboscidea}\label{coordinated-duplication-of-tp53-related-genes-in-proboscidea}}
+
+Prior studies found that the ``master'' tumor suppressor TP53 duplicated multiple times in elephants (\emph{10}, \emph{17}), motivating us to further study duplication of genes involved in TP53-related pathways Proboscidea.
+We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, which has the most complete genome and for which several RNA-seq data sets are available.
+We found that the initial duplication of TP53 in Tethytheria, where body size expanded, was preceded by the duplication of GTF2F1 and STK11 in Paenungulata; and was coincident with the duplication of BRD7.
+These three genes are involved in regulating the transcription of TP53 (\emph{105}--\emph{108}), and their duplication prior to that of TP53 may have facilitated re-functionalization of TP53 retroduplicates.
+Interestingly, STK11 is a tumor suppressor gene in its own right, and plays additional roles in mediating tumor suppression via p21-induced senescence (\emph{106}).
+The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either coincident with TP53, in the case of SIAH1, or subsequently in Probodiscea, Elephantidae, or extant elephants (\textbf{Figure \ref{fig:Figure-LoxAfrExpressedDups}}).
+These genes are all expressed in RNA-seq data, suggesting that they encode functional genes (\textbf{Figure \ref{fig:Figure-LoxAfrExpressedDups}}).
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+Among the evolutionary, developmental, and life history constraints on the evolution of large bodies and long lifespans is an increased risk of developing cancer.
+While body size and lifespan are correlated with cancer risk within species, there is no correlation between species because large and long-lived organisms have evolved enhanced cancer suppression mechanisms.
+While this ultimate evolutionary explanation is straightforward (\emph{54}), determining the mechanisms that underlie the evolution of enhanced cancer protection is challenging because many mechanisms likely contribute to evolution of reduced cancer risk.
+Previous candidate gene studies in elephants have identified duplications of tumor suppressors such as TP53 and LIF, among others, suggesting that an increased copy number of tumor suppressors may contribute to the evolution of large body sizes in the elephant lineage (\emph{10}, \emph{17}, \emph{30}--\emph{32}).
+Here we: 1) Trace the evolution of body sizes in Eutherian mammals, with particular reference to Afrotherians; 2) Infer changes in cancer susceptibility across Afrotherian lineages; and 3) Use a genome-wide screen to determine gene duplications in Afrotherian genomes, including multiple living and extinct Proboscideans, are enriched in tumor suppressor pathways.
+
+\hypertarget{correlated-evolution-of-large-bodies-and-reduced-cancer-risk}{%
+\subsection{Correlated evolution of large bodies and reduced cancer risk}\label{correlated-evolution-of-large-bodies-and-reduced-cancer-risk}}
+
+One of the challenges for determining the mechanisms that underlie the evolution of large bodies, long lifespans, and reduced cancer risk is identifying lineages that are large and/or long-lived and closely related to species with normal body sizes and lifespans.
+One such lineage appears to be elephants, which are deeply nested within Afrotherians with smaller bodies and that are generally shorter-lived.
+To reconstruct the evolution of body size across Afrotherians, and determine in which lineages exceptional body sizes evolved generated a time-calibrated super tree of Eutherian mammals that combined the X and Y trees.
+Similar to previous studies, we found that the evolution of gigantism in Afrotheria was not limited to the living and extinct Proboscideans, but independently evolved in multiple lineages including hyraxes and sea cows.
+These data indicate that the Paenungulata is a particularly good model in which to explore mechanisms that underlie the evolution of exceptionally large bodies.
+
+The absence of a correlation between body size or lifespan and cancer risk indicates that large, long-lived species evolved reduced cancer risk.
+Previous studies, for example, have shown that while cancer incidence varies by species, the rate is relatively stable at \textasciitilde5\% for both large and small mammals; rates can increase significantly for species with transmissible cancers such as Tasmanian devils or that live in particularly polluted environments such as Minke whales that inhabit the St.~Lawrence estuary.
+Therefore, we used our ancestral body size and lifespan estimates to infer the magnitude of changes in intrinsic cancer risk across species (following Peto (2015)) that would be required to keep cancer rates stable across species.
+As expected, we identified significant reductions in intrinsic cancer risk in large bodied lineages, most notably in Pseudoungulata (13 log\_\{2\}-fold RICR), Tethytheria (7.84 log\_\{2\}-fold RICR), and Proboscidea (3.14 log\_\{2\}-fold RICR).
+
+\hypertarget{all-afrotheria-are-equal-but-some-afrotheria-are-more-equal-than-others}{%
+\subsection{All Afrotheria are equal, but some Afrotheria are more equal than others}\label{all-afrotheria-are-equal-but-some-afrotheria-are-more-equal-than-others}}
+
+The hundred- to hundred-million-fold reductions in intrinsic cancer risk associated with the evolution of large body sizes in some Afrotherian lineages, in particular Paenungulates such as elephants and mastodons, suggests that these lineages must have also evolved remarkable mechanisms to suppress cancer.
+While our initial hypothesis was that large bodied lineages would be uniquely enriched in duplicate tumor suppressor genes compared to other smaller bodied lineages, we unexpectedly found that duplication of genes in tumor suppressor pathways was common in Afrotheria regardless of body size.
+These data suggest that this abundance of tumor suppressors may have contributed to the evolution of large bodies and reduced cancer risk, but that these processes were not necessarily coincident.
+Interestingly, pervasive duplication of tumor suppressors may also have contributed to the repeated evolution of large bodies in hyraxes and sea cows, because at least some of the genetic changes that underlie the evolution of reduced cancer risk was common in this group.
+It remains to be determined whether our observation of pervasive duplication of tumor suppressors also occurs in other lineages.
+Using a similar reciprocal best BLAST/BLAT approach that focused on estimating copy number of known tumor suppressors in mammalian genomes, for example, Caulin et al.~(2015) found no evidence between copy number or tumor suppressor genes with either body mass or longevity, whilst Tollis et al.~(2020) found a correlation between copy number and longevity (but not body size) (\emph{12}, \emph{31}).
+These opposing conclusions may be results from differences in the number of genes (81 vs 548) and genomes (8 vs 63) analyzed, highlighting the need for genome-wide analyses of many species that vary in body size and longevity.
+
+While we found that duplication of tumor suppressor genes is common in Afrotheria, genes that duplicated in the Proboscidean stem-lineage (\textbf{Figure \ref{fig:Figure-ML-UpSet-WC}C}) were uniquely enriched in functions and pathways that may be related to the evolution of unique anti-cancer cellular phenotypes in the elephant lineage.
+Elephant cells, for example, cannot be experimentally immortalized (\emph{109}, \emph{110}), rapidly repair DNA damage (\emph{17}, \emph{111}, \emph{112}), are extremely resistant to oxidative stress (\emph{110}), yet are also extremely sensitive to DNA damage (\emph{10}, \emph{17}, \emph{30}).
+Several pathways related to DNA damage repair, nucleotide excision repair (NER) in particular, were uniquely enriched among genes that duplicated within Proboscideans suggesting a connection between duplication of genes involved in NER and rapid DNA damage repair (\emph{111}, \emph{112}).
+Similarly, the duplicate SOD1 gene in elephants may confer the resistance of elephant cells to oxidative stress (\emph{110}).
+Pathways related to the cell cycle were also uniquely enriched among genes that duplicated within Proboscideans, and cell cycle dynamics are different in elephants compared to other species; population doubling (PD) times for African and Asian elephant cells were 13-16 days, while it 21-28 days for other Afrotherians (\emph{110}).
+Finally, the role of ``mTOR signaling'' in the biology of aging is well-known.
+Collectively these data suggest that gene duplications within Proboscideans may underlie some of their cellular phenotypes that contribute to cancer resistance.
+
+\hypertarget{theres-no-such-thing-as-a-free-lunch-trade-offs-and-constraints-on-tumor-suppressor-copy-number}{%
+\subsection{There's no such thing as a free lunch: Trade-offs and constraints on tumor suppressor copy number}\label{theres-no-such-thing-as-a-free-lunch-trade-offs-and-constraints-on-tumor-suppressor-copy-number}}
+
+The lack of a preponderance of duplicate tumor suppressor in the stem-lineage of Proboscideans, coincident with the evolution of very large body sizes, may suggest that Afrotherians have all the additional tumor suppressors needed to evolve large body sizes, that duplication of tumor suppressors is not a general phenomena related to the evolution of cancer resistance or only a minor one, or reflect a trade-off between the protective effects of increased tumor suppressor number on cancer risk and potentially deleterious consequences of augmented tumor suppressor copy number.
+Indeed, increased copy number of tumor suppressors can have deleterious consequences.
+Overexpression of TP53 in mice, while protective against cancer, is associated with progeria, premature reproductive senescence, and early death; however, transgenic mice with a duplication of the TP53 locus that includes native regulatory elements are healthy and experience normal aging, while also demonstrating an enhanced response to cellular stress and lower rates of cancer (\emph{29}, \emph{113}).
+These data suggest duplication of tumor suppressors can contribute to augmented cancer resistance, if the duplication includes sufficient regulatory architecture to direct spatially and temporally appropriate gene expression.
+Thus it is fascinating that duplication of genes that regulate TP53 function, such as STK11, SIAH1, and BRD7, preceded the retroduplication TP53 in the Proboscidean stem-lineage, which may have mitigated toxicity arising from dosage imbalances.
+Similar co-duplication events may have alleviated the negative pleiotropy of tumor suppressor gene duplications to enable their persistence and allow for subsequent co-option during the evolution of cancer resistance.
+
+\hypertarget{conclusions-caveats-and-limitations}{%
+\subsection{Conclusions, caveats, and limitations}\label{conclusions-caveats-and-limitations}}
+
+Our genome-wide results suggest that duplication of tumor suppressors is pervasive in Afrotherians and may have enabled the evolution of larger body sizes by lowering intrinsic cancer risk either prior to (preadaptation or exaptation) or coincident with increasing body size.
+However, our study has several inherent limitations, for example, we have shown that genome quality plays an important role in our ability to identify duplicate genes and several species have poor quality genomes.
+Conversely, without comprehensive gene expression data we cannot be certain that duplicate genes are actually expressed.
+Duplication of tumor suppressor genes is also unlikely to be the only mechanism responsible for the evolution of large body sizes, long lifespans, and reduced cancer risk.
+The evolution of regulatory elements and coding genes with other non-canonical tumor suppressor functions are also important for mediating the cancer risk.
+We also assume that duplicate genes preserve their original functions and increase overall gene dosage.
+Many processes, however, such as developmental systems drift, neofunctionalization, and subfunctionalization can cause divergence in gene functions (\emph{114}--\emph{116}), leading to inaccurate inferences of dosage effects and pathway functions.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+We would like to thank Olga Duchenko and Eren Aiden at Baylor College of Medicine for the Hi-C scaffolded Procavia capensis, Trichechus manatus, Orycteropus afer, and Choloepus hoffmannii genomes.
+We would also like to thank
+D.H. Vazquez for his indispensable support.
+
+\hypertarget{conflicts-of-interest}{%
+\section{Conflicts of Interest}\label{conflicts-of-interest}}
+
+The Authors have no conflicts of interest to report.
+
+\hypertarget{funding-source}{%
+\section{Funding Source}\label{funding-source}}
+
+We would like to thank the Department of Human Genetics at the University of Chicago for supporting this project.
+
+\hypertarget{figures}{%
+\section{FIGURES}\label{figures}}
+
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-intro-1} \caption{\emph{Atlantogenatans} span a diverse range of body sizes (\emph{18}, \emph{25}).}\label{fig:Figure-intro}
 \end{figure}
 
-To trace the evolutionary history of body mass and lifespan in
-Afrotherians, we built a time-calibrated supertree of Eutherian mammals
-combining 1,679 species from Bininda-Emonds et al {[}25{]} with a total
-evidence Afrotherian phylogeny including 77 extant and fossil data from
-39 extinct species {[}20{]}. Fossil data from extinct species were
-included to ensure that ancestral state reconstructions of body mass in
-Afrotherians were not biased by only including extant species, which can
-lead to inaccurate reconstructions, for example, if lineages multiple
-lineages evolved large body masses from a small bodied ancestor. We
-jointly estimated rates of body mass evolution and reconstructed
-ancestral states using a generalization of a Brownian model of character
-evolution, which allows for occasional large jumps in traits (stable
-model) and out performs standard Brownian motion and Ornstein-Uhlenbeck
-models of character evolution {[}26{]}.
-
-Similar to previous studies of Afrotherian body size {[}20,26{]}, we
-found that the body mass of the Afrotherian ancestor was inferred to be
-small (0.26kg, 95\% CI: 0.31-3.01kg) and that substantial accelerations
-in the rate of body mass evolution occurred coincident with a 67.36x
-increase in body mass in the stem-lineage of \emph{Pseudoungulata}
-(17.33kg), a 1.45x increase in body mass in the stem-lineage of
-\emph{Paenungulata} (25.08kg), a 11.82x increase in body mass in the
-stem-lineage of \emph{Tehthytheria} (296.56kg), and a 2.69x increase in
-body mass in the stem-lineage of \emph{Proboscidea} (4114.39kg) (Figure
-1B,C). The ancestral \emph{Hyracoidea} was inferred to be relatively
-small 2.86kg-118.18kg, and rate accelerations were coincident with
-independent body mass increases in large hyraxes such as
-\emph{Titanohyrax andrewsi} 429.34kg; 67.36x increase. While the body
-mass of the ancestral \emph{Sirenian} was inferred to be large
-61.7kg-955.51kg, a rate acceleration occurred coincident with a 10.59x
-increase in body mass in Stellar's sea cow. Rate accelerations also
-occurred coincident with 36.6x decrease in body mass in the stem-lineage
-of the dwarf elephants \emph{Elephas (Palaeoloxodon) antiquus falconeri}
-and \emph{Elephas cypriotes}. These data suggest that gigantism in
-\emph{Afrotherians} evolved step-wise, from small to medium bodies in
-the \emph{Pseudoungulata} stem-lineage, medium to large bodies in the
-\emph{Tehthytherian} stem-lineage and extinct hyraxes, and from large to
-exceptionally large bodies independently in the \emph{Proboscidean}
-stem-lineage and Stellar's sea cow (Figure 1).
-
-\hypertarget{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}{%
-\subsection{Step-wise reduction of intrinsic cancer risk in large,
-long-lived
-Afrotherians}\label{step-wise-reduction-of-intrinsic-cancer-risk-in-large-long-lived-afrotherians}}
-
-\begin{figure}
-
-\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure2-1} \hfill{}
-
-\caption{Cancer succecptibility across Atlantogenata}\label{fig:Figure2}
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-DoubleTree-Bodysize-Lifespan-1} \caption{Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata.
+\textbf{A)} Back-to-back phylogeny for \emph{Atlantogenata}, scaled by log-fold increases in body mass (IBM, left) and log\_\{2\}-fold reduction in intrinsic cancer risk (RICR, right).
+\textbf{B)} Extant body sizes, plus estimated body sizes, for highlighted species and their ancestral nodes.
+\textbf{C)} Estimated Lifespans for body sizes in \textbf{B)}.
+Intervals are based on predicted lifespans of the 95\% CIs for body size.}\label{fig:Figure-DoubleTree-Bodysize-Lifespan}
 \end{figure}
 
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-ML-UpSet-WC-1} \caption{While duplications of genes in tumor supressor pathways are pervasive in \emph{Atlantogenata}, \emph{Proboscideans} have a greater number of cancer-related pathways enriched, in addition to duplications in various unique pathways.
+\textbf{A)} A cladogram of \emph{Afrotheria} highlighting the estimated number of genes duplicated at each node using maximum likelihood.
+\textbf{B)} Upset plot of cancer-related Reactome pathways enriched in each clade in \textbf{A}.
+\textbf{C)} Wordcloud of enriched pathways among Proboscideans that are exclusive to Proboscidea (purple), shared between Proboscidea and Tethytheria (blue), or shared between Proboscidea and one or more other lineages (green).}\label{fig:Figure-ML-UpSet-WC}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/Figure-LoxAfrExpressedDups-1} \caption{Duplications that occured in the African Savannah Elephant, \emph{Loxodonta africana}, are enriched for TP53-related and other tumor suppressor processes.
+\textbf{A)} Upset plot of cancer-related Reactome pathways in \emph{Loxodonta africana}, highlighting shared genes in each set, and the pathway class represented by the combinations.
+\textbf{B)} Inverted Upset plot from \textbf{A)}, showing the pathways shared by genes highlighted by WEBGESTALT in each pathway.
+\textbf{C)} Cladogram of \emph{Atlantogenata}, colored by the rate of gene duplication estimated by maximum likelihood.
+Dots represent a duplication event of the color-coded genes.
+\textbf{D)} Gene expression levels of genes in \textbf{C)} which have two or more expressed duplicates.}\label{fig:Figure-LoxAfrExpressedDups}
+\end{figure}
+
+\hypertarget{tables}{%
+\section{TABLES}\label{tables}}
+
 \begin{table}
 
-\caption{\label{tab:Fig2 Prep Table 1}Estimated Cancer Succeptibilty for nodes in Atlantogenata}
+\caption{\label{tab:Table-GeneDup}Summary of duplications in Atlantogenata}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lrlllr}
+\begin{tabular}[t]{l|l|l|r|r|l|l|r}
+\hline
+Species & Common Name & Size (g) & \#Hits & \#Duplicated & \% Genes Found & \% Hits Duplicated & Mean ECNC/Hit\\
+\hline
+Choloepus hoffmanni & Hoffmans Two-Toed Sloth & 4.8e+03 & 14082 & 3204 & 78.19\% & 22.75\% & 0.98\\
+\hline
+Chrysochloris asiatica & Cape Golden Mole & 49 & 13547 & 2716 & 75.22\% & 20.05\% & 0.99\\
+\hline
+Dasypus novemcinctus & Nine-Banded Armadillo & 4.3e+03 & 13819 & 2605 & 76.73\% & 18.85\% & 0.98\\
+\hline
+Echinops telfairi & Lesser Hedgehog Tenrec & 87 & 12903 & 1670 & 71.64\% & 12.94\% & 0.99\\
+\hline
+Elephantulus edwardii & Cape Elephant Shrew & 49 & 12884 & 3048 & 71.53\% & 23.66\% & 0.99\\
+\hline
+Elephas maximus & Asian Elephant & 3.3e+06 & 14073 & 907 & 78.14\% & 6.44\% & 1.00\\
+\hline
+Loxodonta africana & African Savanna Elephant & 4.6e+06 & 14051 & 940 & 78.01\% & 6.69\% & 1.00\\
+\hline
+Loxodonta cyclotis & African Forest Elephant & 4.7e+06 & 14065 & 900 & 78.09\% & 6.40\% & 1.00\\
+\hline
+Mammut americanum & American Mastodon & 6e+06 & 13840 & 737 & 76.84\% & 5.33\% & 1.00\\
+\hline
+Mammuthus columbi & Columbian Mammoth & 6.6e+06 & 13059 & 426 & 72.51\% & 3.26\% & 1.00\\
+\hline
+Mammuthus primigenius & Woolly Mammoth & 4.3e+06 & 13935 & 723 & 77.37\% & 5.19\% & 1.00\\
+\hline
+Orycteropus afer & Aardvark & 5.3e+04 & 13880 & 1083 & 77.06\% & 7.80\% & 0.99\\
+\hline
+Palaeoloxodon antiquus & Straight Tusked Elephant & 1e+07 & 13969 & 745 & 77.56\% & 5.33\% & 1.00\\
+\hline
+Procavia capensis & Rock Hyrax & 3e+03 & 13672 & 788 & 75.91\% & 5.76\% & 1.00\\
+\hline
+Trichechus manatus & Manatee & 6.9e+05 & 14092 & 1046 & 78.24\% & 7.42\% & 1.00\\
+\hline
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-ORA-pcCancer-withSim}Summary of Reactome Pathways in Atlantogenata}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{l|r|r|r|r|l}
+\hline
+Node & N. Genes & N. Pathways & \% Cancer Pathways & Simulated \% Cancer Pathways & p<0.05 in Simulation\\
+\hline
+Afroinsectivora & 36 & 65 & 0.1384615 & 0.1542361 & Yes\\
+\hline
+Afrosoricida & 79 & 27 & 0.3703704 & 0.1542361 & Yes\\
+\hline
+Chrysochloris.asiatica & 1591 & 100 & 0.2700000 & 0.1542361 & No\\
+\hline
+Echinops.telfairi & 587 & 100 & 0.2200000 & 0.1542361 & No\\
+\hline
+Elephantidae & 61 & 25 & 0.2000000 & 0.1302682 & Yes\\
+\hline
+Elephantulus.edwardii & 2103 & 100 & 0.2200000 & 0.1542361 & No\\
+\hline
+Elephas.maximus & 94 & 32 & 0.4062500 & 0.1772575 & Yes\\
+\hline
+Loxodona & 12 & 60 & 0.1000000 & 0.1453344 & Yes\\
+\hline
+Loxodonta.africana & 100 & 47 & 0.5319149 & 0.1542361 & Yes\\
+\hline
+Loxodonta.cyclotis & 76 & 35 & 0.3428571 & 0.1611374 & Yes\\
+\hline
+Loxodontini & 15 & 12 & 0.0000000 & 0.1381974 & Yes\\
+\hline
+Mammut.americanum & 52 & 16 & 0.0000000 & 0.1290878 & Yes\\
+\hline
+Mammuthus & 5 & 62 & 0.0483871 & 0.1529145 & Yes\\
+\hline
+Mammuthus.columbi & 28 & 26 & 0.2692308 & 0.1287958 & Yes\\
+\hline
+Mammuthus.primigenius & 35 & 16 & 0.0000000 & 0.1228288 & Yes\\
+\hline
+Orycteropus.afer & 504 & 100 & 0.3800000 & 0.1542361 & No\\
+\hline
+Paenungulata & 28 & 22 & 0.1818182 & 0.1287958 & Yes\\
+\hline
+Palaeoloxodon.antiquus & 35 & 8 & 0.0000000 & 0.1228288 & Yes\\
+\hline
+Proboscidea & 157 & 67 & 0.3880597 & 0.0952381 & Yes\\
+\hline
+Procavia.capensis & 383 & 35 & 0.0285714 & 0.1542361 & No\\
+\hline
+Pseudoungulata & 9 & 10 & 0.0000000 & 0.1489860 & Yes\\
+\hline
+Tethytheria & 83 & 46 & 0.6304348 & 0.1851852 & Yes\\
+\hline
+Trichechus.manatus & 484 & 47 & 0.2127660 & 0.1542361 & No\\
+\hline
+\end{tabular}}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:Table-SRAData}NCBI SRA datasets used in this study, along with key biological and genome information.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lll>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}}
 \toprule
-Node & Est. Lifespan & K1 & K2 & Change in K & log2 Change\\
+Organism & Common Name & Genome & SRA Acc. & Tissues\\
 \midrule
-Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
-Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 & 4.07\\
-Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
-Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-\addlinespace
-Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 & 27.66\\
-Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
-Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 & 4.14\\
-Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
-\addlinespace
-Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
-Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 & -23.05\\
-Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 & 29.97\\
-Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 & 6.92\\
-Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 & 8.62\\
-\addlinespace
-Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 & 0.00\\
-Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 & -5.01\\
-Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 & 9.22\\
-Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 & 5.38\\
-Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
-\addlinespace
-Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
-Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 & 0.00\\
-Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 & 5.22\\
-Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 & 9.22\\
-Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 & 20.75\\
-\addlinespace
-Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 & -3.76\\
-Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 & 1.51\\
-Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 & 0.00\\
-Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 & 0.00\\
+Dasypus novemcinctus & Nine-banded armadillo & dasNov3 & SRR494779, SRR494767, SRR494780, SRR494770, SRR309130, SRR494771, SRR4043756, SRR494776, SRR494778, SRR4043762, SRR4043755, SRR6206923, SRR4043761, SRR4043760, SRR6206913, SRR4043763, SRR494772, SRR494781, SRR494774, SRR494777, SRR494775, SRR4043754, SRR1289524, SRR4043758, SRR6206903, SRR1289523, SRR4043759, SRR3222425, SRR494768, SRR494769, SRR6206908, SRR4043757, SRR494766, SRR6206918, SRR494773 & Kidney, Spleen, Cerebellum W/ Brainstem, Rt.
+Quadricep, Mid-Stage Pregnant Endometrium, Cervix, Lung, Liver, Skeletal Muscle, Ascending Colon, Pregnant Armadillo Endometrium, Heart, Placenta\\
+Loxodonta africana & African savanna elephant & loxAfr3, loxAfrC, loxAfr4 & SRR6307198, SRR1041765, SRR6307199, SRR6307201, SRR6307196, SRR6307202, SRR6307200, SRR6307195, SRR975188, SRR6307194, SRR6307204, SRR3222430, SRR6307205, SRR975189, SRR6307197, SRR6307203 & Blood, Fibroblast, Placenta\\
+Trichechus manatus latirostris & Manatee & triMan1, triManLat2 & SRR4228542, SRR4228545, SRR4228544, SRR4228539, SRR4228541, SRR4228538, SRR4228546, SRR4228537, SRR4228540, SRR4228543, SRR4228547 & Buffy Coat\\
 \bottomrule
@@
 
-The dramatic increase in body mass and lifespan in some Afrotherian
-lineages implies those lineages evolved reduced cancer risk. To infer
-the magnitude of these reductions we estimated differences in cancer
-risk between small bodied, short-lived species and large bodied,
-long-lived species as well as for reconstructed ancestral Afrotherians.
-Following {[}44{]} we estimate the intrinsic cancer risk as the product
-of risk associated with body mass and lifespan. Differences in cancer
-susceptibility \(K\) due to body mass differences between species can be
-approximated simply as the fold difference in body mass (\(D\)) between
-species {[}44{]}. The risk of developing cancer also increases in
-proportion to the sixth power of age and is approximated by the formula
-\(Ct^6\), in which the proportionality constant C that determines
-susceptibility to cancer induction is multiplied by the sixth power of
-the age in years, \(t\) {[}44--46{]}. Thus we can estimate the intrinsic
-cancer risk for a species as \(K \approx Dt^6\).
-
-In order to estimate the intrinsic cancer risk of a species, we first
-obtained estimates for lifespans at ancestral nodes using PGLS and the
-model
-\(ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon\)
-(Table 2). With this information in hand, we calculated \(K_{1}\) at all
-nodes, and then estimated the fold change in cancer susceptibility
-between an ancestral node and a given node as \(\frac{K_{2}}{K_{1}}\)
-(\textbf{Table 4}). As body size and phylogeny are the strongest
-predictors of lifespan, we find that this regression is sufficiently
-robust without further metabolic or other covariates for our purposes
-herein.
-
-As shown in \textbf{Table 2}, cancer susceptibility skyrocketed at the
-initial divergence of Atlantogenata, followed by a generally upwards
-trend. At the common ancestor of Afrotheria there is an initial
-9.22-fold increase in cancer risk. In parallel to Afrotheria, cancer
-susceptibility increases 20.75-fold in Xenarthra. However, cancer risk
-slowly deflates as size decreases as one moves along the tree towards
-extant species, such as in Hoffman's Two Toed Sloth (1.51-fold change)
-and in the Nine-banded Armadillo (-3.76-fold change).
-
-Within Afrotheria, cancer susceptibility drops in Afrosoricida as
-species shrink (-4.61-fold, then stagnates for the Cape Golden Mole) -
-but then rises 5.22-fold towards the Lesser Hedgehog Tenrec. In
-parallel, Afroinsectivora does not increase in cancer susceptibility,
-and decreases once more at the Cape Elephant Shrew (-4.61-fold). The
-emergence of Pseudoungulata sees the next big leap in cancer
-susceptibility with a 9.22-fold increase. The Aardvark further increases
-5.38-fold, while we don't observe an increase at the common ancestor of
-Paenungulates. While the Rock Hyrax decreases in cancer susceptibility
-as expected (-5.01-fold), Tethytheria sees a sharp increase in cancer
-risk (6.92-fold). Within Tethytheria, the Manatee's cancer risk
-increases once more 8.62-fold, while Proboscidea's cancer risk drops
-precipitously along with its body size (-23.05-fold). Yet, within
-Proboscidea we see the biggest increases: right off the bat, we see that
-the cancer susceptibility of Elephantidae and the American Mastodon
-skyrocket by 27.66-fold and 29.97-fold, respectively. Both Elephantina
-and Loxodontini in Elephantidae have a 2.31-fold increase in cancer
-susceptibility. Within Elephantina, cancer susceptibility stays stable
-at Mammuthus and in the Colombian Mammoth, and slightly decreases in the
-Wooly Mammoth (-2.31-fold). The three extant elephants - Asian Elephant
-in Elephantina, the African Savana Elephant in Loxodontini, and the
-African Forest Elephant in Loxodona, meanwhile, have parallel and
-similar decreases in both size and cancer susceptibility (4.14-, 4.07-,
-and -2.31-fold, respectively). Neither the common ancestor of Loxodonta,
-nor the Straight-Tusked Mammoth see any further changes in cancer
-susceptibility.
-
-\hypertarget{identification-and-evolutionary-history-of-gene-duplications}{%
-\subsection{Identification and evolutionary history of gene
-duplications}\label{identification-and-evolutionary-history-of-gene-duplications}}
+\begin{table}[!htbp] \centering 
+  \caption{PGLS: ln(Lifespan) ~ ln(Size)} 
+  \label{} 
+\begin{tabular}{@{\extracolsep{5pt}}lc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+\\[-1.8ex] & ln(Lifespan) \\ 
+\hline \\[-1.8ex] 
+ ln(Size) & 0.200$^{***}$ \\ 
+  & (0.027) \\ 
+  & \\ 
+ Constant & 1.327$^{***}$ \\ 
+  & (0.237) \\ 
+  & \\ 
+\hline \\[-1.8ex] 
+Observations & 28 \\ 
+Log Likelihood & $-$22.160 \\ 
+Akaike Inf. Crit. & 50.320 \\ 
+Bayesian Inf. Crit. & 54.095 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
 
@@
 
-\caption{\label{tab:Figure3 Prep Genome Table}Summary of duplications in Atlantogenata}
+\caption{\label{tab:Table-Genomes}Genomes used in this study.}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lllrrllr}
+\begin{tabular}[t]{lllll}
 \toprule
-Species & Common Name & Size (g) & \#Hits & \#Duplicated & \% Genes Found & \% Hits Duplicated & Mean ECNC/Hit\\
+Species & Common Name & Genomes & Highest Quality Genome & Citation\\
 \midrule
-Choloepus hoffmanni & Hoffmans Two-Toed Sloth & 4.3e+03 & 14082 & 3204 & 78.19\% & 22.75\% & 0.98\\
-Chrysochloris asiatica & Cape Golden Mole & 49 & 13547 & 2716 & 75.22\% & 20.05\% & 0.99\\
-Dasypus novemcinctus & Nine-Banded Armadillo & 4.8e+03 & 13819 & 2605 & 76.73\% & 18.85\% & 0.98\\
-Echinops telfairi & Lesser Hedgehog Tenrec & 87 & 12903 & 1670 & 71.64\% & 12.94\% & 0.99\\
-Elephantulus edwardii & Cape Elephant Shrew & 49 & 12884 & 3048 & 71.53\% & 23.66\% & 0.99\\
+Choloepus hoffmanni & Hoffmans two-toed sloth & choHof1, choHof2, choHof-C\_hoffmanni-2.0.1\_HiC & choHof-C\_hoffmanni-2.0.1\_HiC & NA\\
+Chrysochloris asiatica & Cape golden mole & chrAsi1m & chrAsi1m & NA\\
+Dasypus novemcinctus & Nine-banded armadillo & dasNov3 & dasNov3 & NA\\
+Echinops telfairi & Lesser Hedgehog Tenrec & echTel2 & echTel2 & NA\\
+Elephantulus edwardii & Cape elephant shrew & eleEdw1m & eleEdw1m & NA\\
 \addlinespace
-Elephas maximus & Asian Elephant & 3.3e+06 & 14073 & 907 & 78.14\% & 6.44\% & 1.00\\
-Loxodonta africana & African Savanna Elephant & 4.6e+06 & 14051 & 940 & 78.01\% & 6.69\% & 1.00\\
-Loxodonta cyclotis & African Forest Elephant & 4.7e+06 & 14065 & 900 & 78.09\% & 6.40\% & 1.00\\
-Mammut americanum & American Mastodon & 6e+06 & 13840 & 737 & 76.84\% & 5.33\% & 1.00\\
-Mammuthus columbi & Columbian Mammoth & 6.6e+06 & 13059 & 426 & 72.51\% & 3.26\% & 1.00\\
+Elephas maximus & Asian elephant & eleMaxD & eleMaxD & NA\\
+Loxodonta africana & African savanna elephant & loxAfr3, loxAfrC, loxAfr4 & loxAfr4 & NA\\
+Loxodonta cyclotis & African forest elephant & loxCycF & loxCycF & NA\\
+Mammut americanum & American mastodon & mamAmeI & mamAmeI & NA\\
+Mammuthus columbi & Columbian mammoth & mamColU & mamColU & NA\\
 \addlinespace
-Mammuthus primigenius & Woolly Mammoth & 4.3e+06 & 13935 & 723 & 77.37\% & 5.19\% & 1.00\\
-Orycteropus afer & Aardvark & 5.3e+04 & 13880 & 1083 & 77.06\% & 7.80\% & 0.99\\
-Palaeoloxodon antiquus & Straight Tusked Elephant & 1e+07 & 13969 & 745 & 77.56\% & 5.33\% & 1.00\\
-Procavia capensis & Rock Hyrax & 3e+03 & 13672 & 788 & 75.91\% & 5.76\% & 1.00\\
-Trichechus manatus & Manatee & 6.9e+05 & 14092 & 1046 & 78.24\% & 7.42\% & 1.00\\
+Mammuthus primigenius & Woolly mammoth & mamPriV & mamPriV & NA\\
+Orycteropus afer & Aardvark & oryAfe1, oryAfe2 & oryAfe2 & NA\\
+Palaeoloxodon antiquus & Straight tusked elephant & palAntN & palAntN & NA\\
+Procavia capensis & Rock hyrax & proCap1, proCap2, proCap-Pcap\_2.0\_HiC & proCap-Pcap\_2.0\_HiC & NA\\
+Trichechus manatus latirostris & Manatee & triMan1, triManLat2 & triManLat2 & NA\\
 \bottomrule
@@
 
-\begin{figure}
-
-\includegraphics[width=6in,height=7in]{paper_PLOS_draft_files/figure-latex/Figure3-1} \hfill{}
-
-\caption{Gene duplications occur readily throughout Atlantogenata.
-A) A table summarizing duplication events in Atlantogenata.
-B) A tree of the species in Atlantogenata with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.}\label{fig:Figure3}
+\hypertarget{supplementary-figures}{%
+\section{SUPPLEMENTARY FIGURES}\label{supplementary-figures}}
+
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/SFigure-ECNC-1} \caption{Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs.
+\textbf{A)} A single, contiguous gene homolog in a target genome with 100\% query length coverage has an ECNC of 1.0.
+\textbf{B)} Two contiguous gene homologs, each with 100\% query length coverage have an ECNC of 2.0.
+\textbf{C)} A single gene homolog, split across multiple scaffolds and contigs in a fragmented target genome; BLAT identifies each fragment as a single hit.
+Per nucleotide of query sequence, there is only one corresponding nucleotide over all the hits, thus the ECNC is 1.0.
+\textbf{D)} Two gene homologs, one fragmented and one contiguous.
+100\% of nucleotides in the query sequence are represented between all hits; however, every nucleotide in the query has two matching nucleotides in the target genome, thus the ECNC is 2.0.
+\textbf{E)} One true gene homolog in the target genome, plus multiple hits of a conserved domain that span 20\% of the query sequence.
+While 100\% of the query sequence is represented in total, 20\% of the nucleotides have 4 hits.
+Thus, the ECNC for this gene is 1.45.
+\textbf{F)} Two real gene homologs; one hit is contiguous, one hit is fragmented in two, and the tail end of both sequences was not identified by BLAT due to sequence divergence.
+Only 75\% of the query sequence was covered in total between the hits, but for that 75\%, each nucleotide has two hits.
+As such, ECNC is equal to 2.0 for this gene.}\label{fig:SFigure-ECNC}
 \end{figure}
 
-Enhanced cancer suppression may have evolved through many mechanisms;
-among the most parsimonious is an increase in the copy number of genes
-with tumor suppressor functions. Previous studies focusing on candidate
-gene studies, for example, have identified increased copy number of the
-tumor suppressors \emph{TP53} and \emph{LIF} in elephants
-{[}12,21--24{]}. Therefore, in order to test whether this was pervasive
-genome-wide in \emph{Afrotherians}, we used a Reciprocal Best Hit BLAT
-(RBHB) approach to infer gene copy number in \emph{Afrotherian} and
-\emph{Atlantogenatan} genomes (Fig. 2A, Supplementary Figure 1). Because
-RBHB-like approaches can over-estimate copy number when genes are
-fragmented or incorrectly assembled across multiple scaffolds, we also
-inferred copy number using a complementary method that quantifies the
-ratio between observed and expected gene coverage per nucleotide (ECNC)
-(Supplementary Figure 1). By only including nucleotides from the query
-sequence that were observed in the target genome, we also correct for
-partial hits where some or all of the homologs of a gene have diverged
-from the human homolog.
-
-As our sequence database included various protein transcripts for each
-gene, in order to obtain gene-level copy number information and
-eliminate any many-to-one mappings of hits, we labeled each exon of
-every reciprocal best hit (RBH) with the gene corresponding to the query
-transcript and merged all overlapping exons; next, we eliminated any
-many-to-one exons that resulted from the previous step. Finally, we
-reassembled the gene loci based on the original transcript starts and
-ends, and the collapsed exon data, obtaining the full sequence of each
-RBH locus. Genes were considered to be duplicated if its copy number via
-RBHB was greater than or equal to 2, and the maximum ECNC among all
-transcripts prior to filtering was greater than or equal to 1.50 in
-order to account for partial hits that may be due to incomplete
-coverage.
-
-Once we applied these criteria to the results, we obtained the results
-shown in Figure 2A. Our approach positively identified an average of NA
-of the queries we searched. We observed that the percentage of
-duplicated genes in non-Pseudoungulatan genomes was significantly
-higher: while these genomes had duplicates percentages ranging anywhere
-from 3.26\% to 7.80\%, outgroup species' duplication rates ranged from
-12.94\% to 23.66\%.
-
-\begin{figure}
-
-{\centering \includegraphics[width=5in,height=5in]{paper_PLOS_draft_files/figure-latex/Figure4-1}
+\begin{figure}[H]
+\includegraphics[width=6in,]{paper_PLOS_draft_files/figure-latex/SFigure-TimeTree-Eutheria-1} \caption{The time-calibrated Eutherian tree used as input for StableTraits, from Bininda-Emonds \emph{et al.} (\emph{117}) and Puttick and Thomas (\emph{25}).}\label{fig:SFigure-TimeTree-Eutheria}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=6in,height=6in,]{paper_PLOS_draft_files/figure-latex/SFigure-GenomeCor-1}
 
@@
 
-\caption{Correlations between genome quality metrics and ECNC metrics}\label{fig:Figure4}
+\caption{Correlations between genome quality metrics and ECNC metrics.
+Gene copy number metrics, and the genome quality metrics most strongly associated with them, are highlighted in red.}\label{fig:SFigure-GenomeCor}
 \end{figure}
 
-Given the preliminary status of the genomes for \emph{Chrysochloris
-asiatica} and \emph{Elephantulus edwardii}; and the fact that more
-revisions have been performed for Paenungulatan genomes; we hypothesized
-that genome quality may be confounding the results we've observed. We
-used \texttt{CEGMA} and the \texttt{gVolante} server {[}31,32{]} to
-assess the quality of all our genomes, to see if any of these metrics
-correlated with our metrics of gene counts and copy number. As shown in
-Figure 3, mean ECNC, mean Copy Number, and mean CN (the lesser of Copy
-Number and ECNC per gene) correlate moderately strongly with genomic
-quality metrics related to the length and assembly quality, like LD50
-and the number of scaffolds and contigs with a length above either 100K
-or 1M, supporting our hypothesis that increases in copy number correlate
-.
-
-In order to determine when in the phylogeny our identified duplications
-occurred, we used maximum likelihood to reconstruct ancestral gene copy
-numbers across the phylogeny. We encoded the copy number of each gene as
-a discrete trait ranging from 0 (``?'') to 31+, and then ran the
-ancestral gene copy number reconstruction using the phylogenetics suite
-\texttt{IQTREE}, which rapidly tests and runs the best evolutionary
-model for a given trait of interest. Importantly, \texttt{IQTREE} can be
-set to only output states where the likelihood of the state is greater
-than a certain threshold. Setting the threshold to 0.8, we obtained the
-results shown in Figure 2B.
-
-Gene duplications events occured readily thoughout \emph{Atlantogenata}
-(Figure 2B). Among the genes that increased in copy number in the
-elephant lineage are TP53 and LIF, as previously described; however, we
-find that these two genes represent a fraction of the 940 genes that are
-duplicated in the African Elephant overall, which accumulated over
-various steps through their evolution. While the extinct elephantids
-have acceptable genome quality metrics according to CEGMA, they are
-nonetheless missing a significant number of sequences; this may
-contribute to the low number of duplicated genes that occured in
-internal nodes. The number of duplicates that occur at each branch is
-also proportional to the density of the sampling of the clade overall,
-as would be expected. In branches, such as \emph{Afrosoricida}, where
-the number of species is relatively minuscule compared to the size of
-the clade, we see many significantly larger numbers of duplications
-private to these species.
-
-\hypertarget{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}{%
-\subsection{Duplications that occurred recently in Probodiscea are
-enriched for tumor suppressor
-pathways}\label{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}}
-
-\begin{table}
-
-\caption{\label{tab:Branches which experienced an increase in size}Number of pathways overrepresented among duplicated genes at different FDRs.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{llrrrr}
-\toprule
-Ancestor & Node & Pathways at FDR<0.1 & Pathways at FDR<0.2 & Pathways at FDR<0.3 & Pathways at FDR<0.5\\
-\midrule
-Afroinsectivora & Elephantulus edwardii & 252 & 37 & 30 & 87\\
-Afrosoricida & Chrysochloris asiatica & 90 & 48 & 43 & 105\\
-Afrosoricida & Echinops telfairi & 0 & 2 & 0 & 31\\
-Afrotheria & Afroinsectivora & 0 & 0 & 0 & 33\\
-Loxodontini & Loxodonta africana & 6 & 0 & 1 & 0\\
-\addlinespace
-Paenungulata & Tethytheria & 0 & 0 & 0 & 2\\
-Proboscidea & Mammut americanum & 0 & 0 & 0 & 6\\
-Pseudoungulata & Orycteropus afer & 27 & 67 & 29 & 67\\
-Tethytheria & Proboscidea & 0 & 3 & 0 & 3\\
-Tethytheria & Trichechus manatus & 4 & 0 & 0 & 2\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{figure}
-
-\includegraphics[width=6in,height=4in]{paper_PLOS_draft_files/figure-latex/Figure5-1} \hfill{}
-
-\caption{Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.}\label{fig:Figure5}
-\end{figure}
-
-Our initial hypothesis was that genes which duplicated in lineages that
-experienced a growth in size would be enriched for membership in tumor
-suppressor pathways. Thus, we used WebGestalt and its Overrepresentation
-Analysis (ORA) functionality to determine what pathways were enriched in
-our duplicated gene sets in each branch relative to our initial query
-set. For our database, we used Reactome for our primary analysis, but
-additionally used the KEGG, Panther, Wikipathways, and
-Wikipathways\_cancer databases using WebGestalt ORA. Going through the
-tree, at no FDR\textless{}0.5 is there any significant pathway
-representation for genes that increased in the branches leading to , or
-\emph{Paenungulata}; furthermore, there are no significant pathway
-enrichments at FDR\textless{}0.5 for genes whose copy number did not
-change between branches (copy-number-stable) for , or \emph{Afrotheria}.
-Note that because \emph{Xenarthra} was selected as the outgroup, it is
-not possible to polarize the changes in their gene copy numbers along
-the tree.
-
-For the other branches, the number of pathways that came up as
-significantly enriched at each FDR is shown in Figure 3A (Supplementary
-Figure 3). For the species with high duplication rates and
-lower-quality, highly-fragmented genomes, such as with
-\emph{Chrysochloris asiatica} (20.05\% duplicated hits) and
-\emph{Elephantulus edwardii} (23.66\% duplicated hits), it is
-unsurprising that there is a proportionally large number of pathway
-enrichments. In the case of these two species, their many pathway
-enrichments also span an incredible range of processes at every level of
-biology; this, in combination with the high number of copy numbers
-identified for these genes, further suggests a need for improvement and
-refinement in these genomes. In the cell cycle pathways called as
-significant in the genomes of these two species plus \emph{Orycteropus
-afer} and \emph{Echinops telfairi}, the duplicated genes included in
-these sets are from the same gene families, such as the APC subunit
-family; the proteosome subunit families; and the protein phosphatase 2
-family, among others. It is highly possible that these results reflect
-true expansions of these gene families, especially in the higher-quality
-\emph{Orycteropus afer} genome; however, it is also possible that it
-simply reflects artificial duplication events, and so require further
-study.
-
-The pathway enrichments for genes whose copy number either did not
-change, or whose copy number increased, between \emph{Loxodona} and the
-African Elephant are shown in Figure 3B and 3C, respectively. Among the
-few enriched pathways in the African Elephant, we see that that two
-tumor suppression pathways - APC Complex-related pathways, and ``TP53
-Regulates Metabolic Genes''- appear not only in the case of stable
-genes, but also in the set of newly duplicated genes. The other pathways
-we see in the set of recently-duplicated genes include
-``Functionalization of Compounds'' and its daughter pathway
-``Xenobiotics''. Genes in these pathways serve to add functional groups
-to lipophylic compounds which would otherwise not be reactive in the
-cell, and are types of metabolic pathways. In the stable set we see
-enrichment of pathways such as ``Neuronal Systems'' and ``Axon
-Guidance,'' which fit in well with what is known about elephant biology
-and evolution {[}47{]}. Overall in elephants, we see enrichments within
-duplicated genes for pathways involved in what makes an elephant an
-elephant - including tumor suppressor pathways.
-
-\hypertarget{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}{%
-\subsection*{\texorpdfstring{Concerted duplication of TP53 and
-TP53-related genes towards
-\emph{Probodiscea}}{Concerted duplication of TP53 and TP53-related genes towards Probodiscea}}\label{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}}
-\addcontentsline{toc}{subsection}{Concerted duplication of TP53 and
-TP53-related genes towards \emph{Probodiscea}}
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Savage2007}{}%
+1.
+V.
+M.
+Savage, A.
+P.
+Allen, J.
+H.
+Brown, J.
+F.
+Gillooly, A.
+B.
+Herman, W.
+H.
+Woodruff, G.
+B.
+West, Scaling of number, size, and metabolic rate of cells with body size in mammals.
+\emph{Proceedings of the National Academy of Sciences}.
+\textbf{104}, 4718--4723 (2007).
+
 \leavevmode\hypertarget{ref-Green2011}{}%
-1. Green J, Cairns BJ, Casabonne D, Wright FL, Reeves G, Beral V, et al.
-Height and cancer incidence in the Million Women Study: prospective
-cohort, and meta-analysis of prospective studies of height and total
-cancer risk. The Lancet Oncology. 2011;12: 785--794.
-doi:\href{https://doi.org/10.1016/s1470-2045(11)70154-1}{10.1016/s1470-2045(11)70154-1}
+2.
+J.
+Green, B.
+J.
+Cairns, D.
+Casabonne, F.
+L.
+Wright, G.
+Reeves, V.
+Beral, for the M.
+W.
+S.
+collaborators, Height and cancer incidence in the Million Women Study: prospective cohort, and meta-analysis of prospective studies of height and total cancer risk.
+\emph{The Lancet Oncology}.
+\textbf{12}, 785--794 (2011).
 
 \leavevmode\hypertarget{ref-Nunney:20181c2}{}%
-2. Nunney L. Size matters: height, cell number and a person's risk of
-cancer. Proc R Soc B. 2018;285: 20181743.
-doi:\href{https://doi.org/10.1098/rspb.2018.1743}{10.1098/rspb.2018.1743}
+3. L. Nunney, Size matters: height, cell number and a person's risk of cancer. \emph{Proc. R. Soc. B}. \textbf{285}, 20181743 (2018).
 
 \leavevmode\hypertarget{ref-Dobson2013}{}%
-3. Dobson JM. Breed-predispositions to cancer in pedigree dogs. ISRN
-veterinary science. 2013;2013: 941275.
-doi:\href{https://doi.org/10.1155/2013/941275}{10.1155/2013/941275}
+4. J. M. Dobson, Breed-predispositions to cancer in pedigree dogs. \emph{ISRN veterinary science}. \textbf{2013}, 941275 (2013).
 
 \leavevmode\hypertarget{ref-Dorn1968}{}%
-4. Dorn CR, Taylor DON, Schneider R, Hibbard HH, Klauber MR. Survey of
-Animal Neoplasms in Alameda and Contra Costa Counties, California. II.
-Cancer Morbidity in Dogs and Cats From Alameda County\textless{}xref
-ref-type="fn" rid="FN2"\textgreater{}2\textless{}/xref\textgreater{}.
-JNCI: Journal of the National Cancer Institute. 1968;40: 307--318.
-doi:\href{https://doi.org/10.1093/jnci/40.2.307}{10.1093/jnci/40.2.307}
+5.
+C.
+R.
+Dorn, D.
+O.
+N.
+Taylor, R.
+Schneider, H.
+H.
+Hibbard, M.
+R.
+Klauber, Survey of Animal Neoplasms in Alameda and Contra Costa Counties, California.
+II.
+Cancer Morbidity in Dogs and Cats From Alameda County\textless xref ref-type="fn" rid="FN2"\textgreater2\textless/xref\textgreater{}.
+\emph{JNCI: Journal of the National Cancer Institute}.
+\textbf{40}, 307--318 (1968).
+
+\leavevmode\hypertarget{ref-Lucena2011}{}%
+6.
+R.
+B.
+Lucena, D.
+R.
+Rissi, G.
+D.
+Kommers, F.
+Pierezan, J.
+C.
+Oliveira-Filho, J.
+T.
+S.
+A.
+Macêdo, M.
+M.
+Flores, C.
+S.
+L.
+Barros, A Retrospective Study of 586 Tumours in Brazilian Cattle.
+\emph{Journal of Comparative Pathology}.
+\textbf{145}, 20--24 (2011).
 
 \leavevmode\hypertarget{ref-Caulin2011}{}%
-5. Caulin AF, Maley CC. Peto's Paradox: evolution's prescription for
-cancer prevention. Trends in ecology \& evolution. 2011;26: 175--82.
-doi:\href{https://doi.org/10.1016/j.tree.2011.01.002}{10.1016/j.tree.2011.01.002}
+7. A. F. Caulin, C. C. Maley, Peto's Paradox: evolution's prescription for cancer prevention. \emph{Trends in ecology \& evolution}. \textbf{26}, 175--82 (2011).
 
 \leavevmode\hypertarget{ref-Leroi2003}{}%
-6. Leroi AM, Koufopanou V, Burt A. Cancer selection. Nature Reviews
-Cancer. 2003;3: 226--231.
-doi:\href{https://doi.org/10.1038/nrc1016}{10.1038/nrc1016}
+8. A. M. Leroi, V. Koufopanou, A. Burt, Cancer selection. \emph{Nature Reviews Cancer}. \textbf{3}, 226--231 (2003).
 
 \leavevmode\hypertarget{ref-Peto1975}{}%
-7. Peto R, Roe F, Lee P, Levy L, Clack J. Cancer and ageing in mice and
-men. British Journal of Cancer. 1975;32: 411--426.
-doi:\href{https://doi.org/10.1038/bjc.1975.242}{10.1038/bjc.1975.242}
+9. R. Peto, F. Roe, P. Lee, L. Levy, J. Clack, Cancer and ageing in mice and men. \emph{British Journal of Cancer}. \textbf{32}, 411--426 (1975).
+
+\leavevmode\hypertarget{ref-Abegglen:JAMA2015}{}%
+10.
+L.
+M.
+Abegglen, A.
+F.
+Caulin, A.
+Chan, K.
+Lee, R.
+Robinson, M.
+S.
+Campbell, W.
+K.
+Kiso, D.
+L.
+Schmitt, P.
+J.
+Waddell, S.
+Bhaskara, S.
+T.
+Jensen, C.
+C.
+Maley, J.
+D.
+Schiffman, Potential Mechanisms for Cancer Resistance in Elephants and Comparative Cellular Response to DNA Damage in Humans.
+\emph{JAMA}.
+\textbf{314}, 1850--1860 (2015).
+
+\leavevmode\hypertarget{ref-Boddy2020}{}%
+11.
+A.
+M.
+Boddy, L.
+M.
+Abegglen, A.
+P.
+Pessier, J.
+D.
+Schiffman, C.
+C.
+Maley, C.
+Witte, Lifetime cancer prevalence and life history traits in mammals.
+\emph{Evolution, Medicine, and Public Health}, eoaa015 (2020).
+
+\leavevmode\hypertarget{ref-Tollis2020}{}%
+12.
+M.
+Tollis, E.
+Ferris, M.
+S.
+Campbell, V.
+K.
+Harris, S.
+M.
+Rupp, T.
+M.
+Harrison, W.
+K.
+Kiso, D.
+L.
+Schmitt, M.
+M.
+Garner, C.
+A.
+Aktipis, C.
+C.
+Maley, A.
+M.
+Boddy, M.
+Yandell, C.
+Gregg, J.
+D.
+Schiffman, L.
+M.
+Abegglen, Elephant Genomes Reveal Insights into Differences in Disease Defense Mechanisms between Species.
+\emph{bioRxiv}, 2020.05.29.124396 (2020).
 
 \leavevmode\hypertarget{ref-Ashur-Fabian2004}{}%
-8. Ashur-Fabian O, Avivi A, Trakhtenbrot L, Adamsky K, Cohen M, Kajakaro
-G, et al. Evolution of p53 in hypoxia-stressed Spalax mimics human tumor
-mutation. Proceedings of the National Academy of Sciences. 2004;101:
-12236--12241.
-doi:\href{https://doi.org/10.1073/pnas.0404998101}{10.1073/pnas.0404998101}
+13.
+O.
+Ashur-Fabian, A.
+Avivi, L.
+Trakhtenbrot, K.
+Adamsky, M.
+Cohen, G.
+Kajakaro, A.
+Joel, N.
+Amariglio, E.
+Nevo, G.
+Rechavi, Evolution of p53 in hypoxia-stressed Spalax mimics human tumor mutation.
+\emph{Proceedings of the National Academy of Sciences}.
+\textbf{101}, 12236--12241 (2004).
 
 \leavevmode\hypertarget{ref-Seluanov2008}{}%
-9. Seluanov A, Hine C, Bozzella M, Hall A, Sasahara THC, Ribeiro AACM,
-et al. Distinct tumor suppressor mechanisms evolve in rodent species
-that differ in size and lifespan. Aging cell. 2008;7: 813--23.
-doi:\href{https://doi.org/10.1111/j.1474-9726.2008.00431.x}{10.1111/j.1474-9726.2008.00431.x}
+14.
+A.
+Seluanov, C.
+Hine, M.
+Bozzella, A.
+Hall, T.
+H.
+C.
+Sasahara, A.
+A.
+C.
+M.
+Ribeiro, K.
+C.
+Catania, D.
+C.
+Presgraves, V.
+Gorbunova, Distinct tumor suppressor mechanisms evolve in rodent species that differ in size and lifespan.
+\emph{Aging cell}.
+\textbf{7}, 813--23 (2008).
 
 \leavevmode\hypertarget{ref-Gorbunova2012}{}%
-10. Gorbunova V, Hine C, Tian X, Ablaeva J, Gudkov AV, Nevo E, et al.
-Cancer resistance in the blind mole rat is mediated by concerted
-necrotic cell death mechanism. Proceedings of the National Academy of
-Sciences of the United States of America. 2012;109: 19392--6.
-doi:\href{https://doi.org/10.1073/pnas.1217211109}{10.1073/pnas.1217211109}
+15.
+V.
+Gorbunova, C.
+Hine, X.
+Tian, J.
+Ablaeva, A.
+V.
+Gudkov, E.
+Nevo, A.
+Seluanov, Cancer resistance in the blind mole rat is mediated by concerted necrotic cell death mechanism.
+\emph{Proceedings of the National Academy of Sciences of the United States of America}.
+\textbf{109}, 19392--6 (2012).
 
 \leavevmode\hypertarget{ref-Tian2013}{}%
-11. Tian X, Azpurua J, Hine C, Vaidya A, Myakishev-Rempel M, Ablaeva J,
-et al. High molecular weight hyaluronan mediates the cancer resistance
-of the naked mole-rat. 2013;499.
-doi:\href{https://doi.org/10.1038/nature12234}{10.1038/nature12234}
+16.
+X.
+Tian, J.
+Azpurua, C.
+Hine, A.
+Vaidya, M.
+Myakishev-Rempel, J.
+Ablaeva, Z.
+Mao, E.
+Nevo, V.
+Gorbunova, A.
+Seluanov, High molecular weight hyaluronan mediates the cancer resistance of the naked mole-rat.
+\textbf{499} (2013), doi:\href{https://doi.org/10.1038/nature12234}{10.1038/nature12234}.
 
 \leavevmode\hypertarget{ref-Sulak2015}{}%
-12. Sulak M, Fong L, Mika K, Chigurupati S, Yon L, Mongan NP, et al.
-TP53 copy number expansion is associated with the evolution of increased
-body size and an enhanced DNA damage response in elephants. eLife.
-2016;5: e11994.
-doi:\href{https://doi.org/10.7554/elife.11994}{10.7554/elife.11994}
+17.
+M.
+Sulak, L.
+Fong, K.
+Mika, S.
+Chigurupati, L.
+Yon, N.
+P.
+Mongan, R.
+D.
+Emes, V.
+J.
+Lynch, TP53 copy number expansion is associated with the evolution of increased body size and an enhanced DNA damage response in elephants.
+\emph{eLife}.
+\textbf{5}, e11994 (2016).
 
 \leavevmode\hypertarget{ref-HAGR}{}%
-13. Tacutu R, Craig T, Budovsky A, Wuttke D, Lehmann G, Taranukha D, et
-al. Human Ageing Genomic Resources: Integrated databases and tools for
-the biology and genetics of ageing. Nucleic Acids Research. 2013;41:
-D1027--D1033.
-doi:\href{https://doi.org/10.1093/nar/gks1155}{10.1093/nar/gks1155}
+18.
+R.
+Tacutu, T.
+Craig, A.
+Budovsky, D.
+Wuttke, G.
+Lehmann, D.
+Taranukha, J.
+Costa, V.
+E.
+Fraifeld, J.
+de Magalhães, Human Ageing Genomic Resources: Integrated databases and tools for the biology and genetics of ageing.
+\emph{Nucleic Acids Research}.
+\textbf{41}, D1027--D1033 (2013).
 
 \leavevmode\hypertarget{ref-Schwartz1995}{}%
-14. Schwartz GT, Rasmussen DT, Smith RJ. Body-Size Diversity and
-Community Structure of Fossil Hyracoids. Journal of Mammalogy. 1995;76:
-1088--1099. doi:\href{https://doi.org/10.2307/1382601}{10.2307/1382601}
+19. G. T. Schwartz, D. T. Rasmussen, R. J. Smith, Body-Size Diversity and Community Structure of Fossil Hyracoids. \emph{Journal of Mammalogy}. \textbf{76}, 1088--1099 (1995).
 
 \leavevmode\hypertarget{ref-Scheffer1972}{}%
-15. Scheffer VB. The Weight of the Steller Sea Cow. Journal of
-Mammalogy. 1972;53: 912--914.
-doi:\href{https://doi.org/10.2307/1379236}{10.2307/1379236}
+20. V. B. Scheffer, The Weight of the Steller Sea Cow. \emph{Journal of Mammalogy}. \textbf{53}, 912--914 (1972).
 
 \leavevmode\hypertarget{ref-Larramendi:20151c2}{}%
-16. Larramendi A. Shoulder Height, Body Mass, and Shape of
-Proboscideans. Acta Palaeontologica Polonica. 2015;61.
-doi:\href{https://doi.org/10.4202/app.00136.2014}{10.4202/app.00136.2014}
+21.
+A.
+Larramendi, Shoulder Height, Body Mass, and Shape of Proboscideans.
+\emph{Acta Palaeontologica Polonica}.
+\textbf{61} (2015), doi:\href{https://doi.org/10.4202/app.00136.2014}{10.4202/app.00136.2014}.
 
 \leavevmode\hypertarget{ref-OLeary2013a}{}%
-17. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
-NP, et al. The placental mammal ancestor and the post-K-Pg radiation of
-placentals. Science (New York, NY). 2013;339: 662--7.
-doi:\href{https://doi.org/10.1126/science.1229237}{10.1126/science.1229237}
+22.
+M.
+A.
+O'Leary, J.
+I.
+Bloch, J.
+J.
+Flynn, T.
+J.
+Gaudin, A.
+Giallombardo, N.
+P.
+Giannini, S.
+L.
+Goldberg, B.
+P.
+Kraatz, Z.-X.
+Luo, J.
+Meng, X.
+Ni, M.
+J.
+Novacek, F.
+A.
+Perini, Z.
+S.
+Randall, G.
+W.
+Rougier, E.
+J.
+Sargis, M.
+T.
+Silcox, N.
+B.
+Simmons, M.
+Spaulding, P.
+M.
+Velazco, M.
+Weksler, J.
+R.
+Wible, A.
+L.
+Cirranello, The placental mammal ancestor and the post-K-Pg radiation of placentals.
+\emph{Science (New York, N.Y.)}.
+\textbf{339}, 662--7 (2013).
 
 \leavevmode\hypertarget{ref-Springer2013}{}%
-18. Springer MS, Meredith RW, Teeling EC, Murphy WJ. Technical comment
-on "The placental mammal ancestor and the post-K-Pg radiation of
-placentals". Science (New York, NY). 2013;341: 613.
-doi:\href{https://doi.org/10.1126/science.1238025}{10.1126/science.1238025}
+23.
+M.
+S.
+Springer, R.
+W.
+Meredith, E.
+C.
+Teeling, W.
+J.
+Murphy, Technical comment on "The placental mammal ancestor and the post-K-Pg radiation of placentals".
+\emph{Science (New York, N.Y.)}.
+\textbf{341}, 613 (2013).
 
 \leavevmode\hypertarget{ref-OLeary2013b}{}%
-19. O'Leary MA, Bloch JI, Flynn JJ, Gaudin TJ, Giallombardo A, Giannini
-NP, et al. Response to comment on "The placental mammal ancestor and the
-post-K-Pg radiation of placentals". Science (New York, NY). 2013;341:
-613.
-doi:\href{https://doi.org/10.1126/science.1238162}{10.1126/science.1238162}
+24.
+M.
+A.
+O'Leary, J.
+I.
+Bloch, J.
+J.
+Flynn, T.
+J.
+Gaudin, A.
+Giallombardo, N.
+P.
+Giannini, S.
+L.
+Goldberg, B.
+P.
+Kraatz, Z.-X.
+Luo, J.
+Meng, X.
+Ni, M.
+J.
+Novacek, F.
+A.
+Perini, Z.
+Randall, G.
+W.
+Rougier, E.
+J.
+Sargis, M.
+T.
+Silcox, N.
+B.
+Simmons, M.
+Spaulding, P.
+M.
+Velazco, M.
+Weksler, J.
+R.
+Wible, A.
+L.
+Cirranello, Response to comment on "The placental mammal ancestor and the post-K-Pg radiation of placentals".
+\emph{Science (New York, N.Y.)}.
+\textbf{341}, 613 (2013).
 
 \leavevmode\hypertarget{ref-PuttickAndThomas2015}{}%
-20. Puttick MN, Thomas GH. Fossils and living taxa agree on patterns of
-body mass evolution: a case study with Afrotheria. Proceedings
-Biological sciences / The Royal Society. 2015;282: 20152023.
-doi:\href{https://doi.org/10.1098/rspb.2015.2023}{10.1098/rspb.2015.2023}
-
-\leavevmode\hypertarget{ref-Abegglen:JAMA2015}{}%
-21. Abegglen LM, Caulin AF, Chan A, Lee K, Robinson R, Campbell MS, et
-al. Potential Mechanisms for Cancer Resistance in Elephants and
-Comparative Cellular Response to DNA Damage in Humans. JAMA. 2015;314:
-1850--1860.
-doi:\href{https://doi.org/10.1001/jama.2015.13134}{10.1001/jama.2015.13134}
+25.
+M.
+N.
+Puttick, G.
+H.
+Thomas, Fossils and living taxa agree on patterns of body mass evolution: a case study with Afrotheria.
+\emph{Proceedings.
+Biological sciences / The Royal Society}.
+\textbf{282}, 20152023 (2015).
+
+\leavevmode\hypertarget{ref-Nunney1999}{}%
+26. L. Nunney, Lineage selection and the evolution of multistage carcinogenesis. \emph{Proceedings of the Royal Society of London. Series B: Biological Sciences}. \textbf{266}, 493--498 (1999).
+
+\leavevmode\hypertarget{ref-Katzourakis2014}{}%
+27. A. Katzourakis, G. Magiorkinis, A. G. Lim, S. Gupta, R. Belshaw, R. Gifford, Larger Mammalian Body Size Leads to Lower Retroviral Activity. \emph{PLoS Pathogens}. \textbf{10}, e1004214 (2014).
+
+\leavevmode\hypertarget{ref-Nagy2007}{}%
+28. J. D. Nagy, E. M. Victor, J. H. Cropper, Why don't all whales have cancer? A novel hypothesis resolving Peto's paradox. \emph{Integrative and Comparative Biology}. \textbf{47}, 317--328 (2007).
+
+\leavevmode\hypertarget{ref-GarciaCao:20021c2}{}%
+29.
+I.
+García‐Cao, M.
+García‐Cao, J.
+Martín‐Caballero, L.
+M.
+Criado, P.
+Klatt, J.
+M.
+Flores, J.
+Weill, M.
+A.
+Blasco, M.
+Serrano, 'Super p53' mice exhibit enhanced DNA damage response, are tumor resistant and age normally.
+\emph{The EMBO Journal}.
+\textbf{21}, 6225--6235 (2002).
 
 \leavevmode\hypertarget{ref-Vazquez2018}{}%
-22. Vazquez JM, Sulak M, Chigurupati S, Lynch VJ. A Zombie LIF Gene in
-Elephants Is Upregulated by TP53 to Induce Apoptosis in Response to DNA
-Damage. Cell Reports. 2018;24: 1765--1776.
-doi:\href{https://doi.org/10.1016/j.celrep.2018.07.042}{10.1016/j.celrep.2018.07.042}
+30.
+J.
+M.
+Vazquez, M.
+Sulak, S.
+Chigurupati, V.
+J.
+Lynch, A Zombie LIF Gene in Elephants Is Upregulated by TP53 to Induce Apoptosis in Response to DNA Damage.
+\emph{Cell Reports}.
+\textbf{24}, 1765--1776 (2018).
 
 \leavevmode\hypertarget{ref-Caulin2015}{}%
-23. Caulin AF, Graham TA, Wang L-S, Maley CC. Solutions to Peto's
-paradox revealed by mathematical modelling and cross-species cancer gene
-analysis. Philosophical transactions of the Royal Society of London
-Series B, Biological sciences. 2015;370: 20140222.
-doi:\href{https://doi.org/10.1098/rstb.2014.0222}{10.1098/rstb.2014.0222}
+31.
+A.
+F.
+Caulin, T.
+A.
+Graham, L.-S.
+Wang, C.
+C.
+Maley, Solutions to Peto's paradox revealed by mathematical modelling and cross-species cancer gene analysis.
+\emph{Philosophical transactions of the Royal Society of London.
+Series B, Biological sciences}.
+\textbf{370}, 20140222 (2015).
 
 \leavevmode\hypertarget{ref-Doherty2016}{}%
-24. Doherty A, Magalhães J de. Has gene duplication impacted the
-evolution of Eutherian longevity? Aging Cell. 2016;15: 978--980.
-doi:\href{https://doi.org/10.1111/acel.12503}{10.1111/acel.12503}
+32. A. Doherty, J. de Magalhães, Has gene duplication impacted the evolution of Eutherian longevity? \emph{Aging Cell}. \textbf{15}, 978--980 (2016).
 
 \leavevmode\hypertarget{ref-Bininda-Emonds2008}{}%
-25. Bininda-Emonds ORP, Cardillo M, Jones KE, MacPhee RDE, Beck RMD,
-Grenyer R, et al. Erratum: The delayed rise of present-day mammals.
-Nature. 2008;456: 274--274.
-doi:\href{https://doi.org/10.1038/nature07347}{10.1038/nature07347}
+33.
+O.
+R.
+P.
+Bininda-Emonds, M.
+Cardillo, K.
+E.
+Jones, R.
+D.
+E.
+MacPhee, R.
+M.
+D.
+Beck, R.
+Grenyer, S.
+A.
+Price, R.
+A.
+Vos, J.
+L.
+Gittleman, A.
+Purvis, Erratum: The delayed rise of present-day mammals.
+\emph{Nature}.
+\textbf{456}, 274--274 (2008).
 
 \leavevmode\hypertarget{ref-ElliotAndMooers2014}{}%
-26. Elliot MG, Mooers AØ. Inferring ancestral states without assuming
-neutrality or gradualism using a stable model of continuous character
-evolution. BMC evolutionary biology. 2014;14: 226.
-doi:\href{https://doi.org/10.1186/s12862-014-0226-8}{10.1186/s12862-014-0226-8}
+34.
+M.
+G.
+Elliot, A.
+Ø.
+Mooers, Inferring ancestral states without assuming neutrality or gradualism using a stable model of continuous character evolution.
+\emph{BMC evolutionary biology}.
+\textbf{14}, 226 (2014).
 
 \leavevmode\hypertarget{ref-blat}{}%
-27. Kent JW. BLAT?The BLAST-Like Alignment Tool. Genome Research.
-2002;12: 656--664.
-doi:\href{https://doi.org/10.1101/gr.229202}{10.1101/gr.229202}
+35. J. W. Kent, BLAT---The BLAST-Like Alignment Tool. \emph{Genome Research}. \textbf{12}, 656--664 (2002).
 
 \leavevmode\hypertarget{ref-AltenhoffAndDessimoz2009}{}%
-28. Altenhoff AM, Dessimoz C. Phylogenetic and functional assessment of
-orthologs inference projects and methods. PLoS computational biology.
-2009;5: e1000262.
-doi:\href{https://doi.org/10.1371/journal.pcbi.1000262}{10.1371/journal.pcbi.1000262}
+36. A. M. Altenhoff, C. Dessimoz, Phylogenetic and functional assessment of orthologs inference projects and methods. \emph{PLoS computational biology}. \textbf{5}, e1000262 (2009).
 
 \leavevmode\hypertarget{ref-SalichosAndRokas2011}{}%
-29. Salichos L, Rokas A. Evaluating ortholog prediction algorithms in a
-yeast model clade. PloS one. 2011;6: e18755.
-doi:\href{https://doi.org/10.1371/journal.pone.0018755}{10.1371/journal.pone.0018755}
+37. L. Salichos, A. Rokas, Evaluating ortholog prediction algorithms in a yeast model clade. \emph{PloS one}. \textbf{6}, e18755 (2011).
 
 \leavevmode\hypertarget{ref-uniprot}{}%
-30. Consortium TU. UniProt: the universal protein knowledgebase. Nucleic
-Acids Research. 2017;45: D158--D169.
-doi:\href{https://doi.org/10.1093/nar/gkw1099}{10.1093/nar/gkw1099}
+38. T. U. Consortium, UniProt: the universal protein knowledgebase. \emph{Nucleic Acids Research}. \textbf{45}, D158--D169 (2017).
 
 \leavevmode\hypertarget{ref-gVolante}{}%
-31. Nishimura O, Hara Y, Kuraku S. gVolante for standardizing
-completeness assessment of genome and transcriptome assemblies.
-Bioinformatics. 2017;33: 3635--3637.
-doi:\href{https://doi.org/10.1093/bioinformatics/btx445}{10.1093/bioinformatics/btx445}
+39. O. Nishimura, Y. Hara, S. Kuraku, gVolante for standardizing completeness assessment of genome and transcriptome assemblies. \emph{Bioinformatics}. \textbf{33}, 3635--3637 (2017).
 
 \leavevmode\hypertarget{ref-CEGMA}{}%
-32. Parra G, Bradnam K, Ning Z, Keane T, Korf I. Assessing the gene
-space in draft genomes. Nucleic Acids Research. 2008;37: 289--297.
-doi:\href{https://doi.org/10.1093/nar/gkn916}{10.1093/nar/gkn916}
+40. G. Parra, K. Bradnam, Z. Ning, T. Keane, I. Korf, Assessing the gene space in draft genomes. \emph{Nucleic Acids Research}. \textbf{37}, 289--297 (2008).
 
 \leavevmode\hypertarget{ref-HISAT}{}%
-33. Kim D, Langmead B, Salzberg SL. HISAT: a fast spliced aligner with
-low memory requirements. Nature Methods. 2015;12: 357--360.
-doi:\href{https://doi.org/10.1038/nmeth.3317}{10.1038/nmeth.3317}
+41. D. Kim, B. Langmead, S. L. Salzberg, HISAT: a fast spliced aligner with low memory requirements. \emph{Nature Methods}. \textbf{12}, 357--360 (2015).
 
 \leavevmode\hypertarget{ref-StringTie}{}%
-34. Pertea M, Pertea GM, Antonescu CM, Chang T-C, Mendell JT, Salzberg
-SL. StringTie enables improved reconstruction of a transcriptome from
-RNA-seq reads. Nature Biotechnology. 2015;33: 290--295.
-doi:\href{https://doi.org/10.1038/nbt.3122}{10.1038/nbt.3122}
+42.
+M.
+Pertea, G.
+M.
+Pertea, C.
+M.
+Antonescu, T.-C.
+Chang, J.
+T.
+Mendell, S.
+L.
+Salzberg, StringTie enables improved reconstruction of a transcriptome from RNA-seq reads.
+\emph{Nature Biotechnology}.
+\textbf{33}, 290--295 (2015).
 
 \leavevmode\hypertarget{ref-Tuxedo}{}%
-35. Pertea M, Kim D, Pertea GM, Leek JT, Salzberg SL. Transcript-level
-expression analysis of RNA-seq experiments with HISAT, StringTie and
-Ballgown. Nature Protocols. 2016;11: 1650--1667.
-doi:\href{https://doi.org/10.1038/nprot.2016.095}{10.1038/nprot.2016.095}
+43.
+M.
+Pertea, D.
+Kim, G.
+M.
+Pertea, J.
+T.
+Leek, S.
+L.
+Salzberg, Transcript-level expression analysis of RNA-seq experiments with HISAT, StringTie and Ballgown.
+\emph{Nature Protocols}.
+\textbf{11}, 1650--1667 (2016).
+
+\leavevmode\hypertarget{ref-IQTREE}{}%
+44.
+B.
+Q.
+Minh, H.
+A.
+Schmidt, O.
+Chernomor, D.
+Schrempf, M.
+D.
+Woodhams, A.
+von Haeseler, R.
+Lanfear, IQ-TREE 2: New Models and Efficient Methods for Phylogenetic Inference in the Genomic Era.
+\emph{Molecular Biology and Evolution}.
+\textbf{37}, 1530--1534 (2020).
+
+\leavevmode\hypertarget{ref-UFBoot2}{}%
+45. D. T. Hoang, O. Chernomor, A. von Haeseler, B. Q. Minh, L. S. Vinh, UFBoot2: Improving the Ultrafast Bootstrap Approximation. \emph{Molecular Biology and Evolution}. \textbf{35}, 518--522 (2017).
+
+\leavevmode\hypertarget{ref-ModelFinder}{}%
+46.
+S.
+Kalyaanamoorthy, B.
+Q.
+Minh, T.
+K.
+F.
+Wong, A.
+von Haeseler, L.
+S.
+Jermiin, ModelFinder: fast model selection for accurate phylogenetic estimates.
+\emph{Nature Methods}.
+\textbf{14}, 587--589 (2017).
+
+\leavevmode\hypertarget{ref-PMSF}{}%
+47.
+H.-C.
+Wang, B.
+Q.
+Minh, E.
+Susko, A.
+J.
+Roger, Modeling Site Heterogeneity with Posterior Mean Site Frequency Profiles Accelerates Accurate Phylogenomic Estimation.
+\emph{Systematic Biology}.
+\textbf{67}, 216--235 (2017).
+
+\leavevmode\hypertarget{ref-Schrempf2019}{}%
+48.
+D.
+Schrempf, B.
+Q.
+Minh, A.
+von Haeseler, C.
+Kosiol, Polymorphism-Aware Species Trees with Advanced Mutation Models, Bootstrap, and Rate Heterogeneity.
+\emph{Molecular Biology and Evolution}.
+\textbf{36}, 1294--1301 (2019).
+
+\leavevmode\hypertarget{ref-Soubrier2012}{}%
+49.
+J.
+Soubrier, M.
+Steel, M.
+S.
+Y.
+Lee, C.
+D.
+Sarkissian, S.
+Guindon, S.
+Y.
+W.
+Ho, A.
+Cooper, The Influence of Rate Heterogeneity among Sites on the Time Dependence of Molecular Rates.
+\emph{Molecular Biology and Evolution}.
+\textbf{29}, 3345--3358 (2012).
+
+\leavevmode\hypertarget{ref-Yang1995}{}%
+50. Z. Yang, S. Kumar, M. Nei, A new method of inference of ancestral nucleotide and amino acid sequences. \emph{Genetics}. \textbf{141}, 1641--50 (1995).
 
 \leavevmode\hypertarget{ref-WebGestalt2019}{}%
-36. Liao Y, Wang J, Jaehnig EJ, Shi Z, Zhang B. WebGestalt 2019: gene
-set analysis toolkit with revamped UIs and APIs. Nucleic Acids Research.
-2019;47: W199--W205.
-doi:\href{https://doi.org/10.1093/nar/gkz401}{10.1093/nar/gkz401}
+51. Y. Liao, J. Wang, E. J. Jaehnig, Z. Shi, B. Zhang, WebGestalt 2019: gene set analysis toolkit with revamped UIs and APIs. \emph{Nucleic Acids Research}. \textbf{47}, W199--W205 (2019).
 
 \leavevmode\hypertarget{ref-Reactome}{}%
-37. Jassal B, Matthews L, Viteri G, Gong C, Lorente P, Fabregat A, et
-al. The reactome pathway knowledgebase. Nucleic acids research. 2020;48:
-D498--D503.
-doi:\href{https://doi.org/10.1093/nar/gkz1031}{10.1093/nar/gkz1031}
-
-\leavevmode\hypertarget{ref-Wikipathways}{}%
-38. Slenter DN, Kutmon M, Hanspers K, Riutta A, Windsor J, Nunes N, et
-al. WikiPathways: a multifaceted pathway database bridging metabolomics
-to other omics research. Nucleic Acids Research. 2017;46: D661--D667.
-doi:\href{https://doi.org/10.1093/nar/gkx1064}{10.1093/nar/gkx1064}
-
-\leavevmode\hypertarget{ref-KEGG}{}%
-39. Kanehisa M, Goto S. KEGG: Kyoto Encyclopedia of Genes and Genomes.
-Nucleic Acids Research. 2000;28: 27--30.
-doi:\href{https://doi.org/10.1093/nar/28.1.27}{10.1093/nar/28.1.27}
+52.
+B.
+Jassal, L.
+Matthews, G.
+Viteri, C.
+Gong, P.
+Lorente, A.
+Fabregat, K.
+Sidiropoulos, J.
+Cook, M.
+Gillespie, R.
+Haw, F.
+Loney, B.
+May, M.
+Milacic, K.
+Rothfels, C.
+Sevilla, V.
+Shamovsky, S.
+Shorser, T.
+Varusai, J.
+Weiser, G.
+Wu, L.
+Stein, H.
+Hermjakob, P.
+D'Eustachio, The reactome pathway knowledgebase.
+\emph{Nucleic acids research}.
+\textbf{48}, D498--D503 (2020).
+
+\leavevmode\hypertarget{ref-Chen2013}{}%
+53.
+E.
+Y.
+Chen, C.
+M.
+Tan, Y.
+Kou, Q.
+Duan, Z.
+Wang, G.
+V.
+Meirelles, N.
+R.
+Clark, A.
+Ma'ayan, Enrichr: interactive and collaborative HTML5 gene list enrichment analysis tool.
+\emph{BMC bioinformatics}.
+\textbf{14}, 128 (2013).
+
+\leavevmode\hypertarget{ref-Peto:20151c2}{}%
+54. R. Peto, Quantitative implications of the approximate irrelevance of mammalian body size and lifespan to lifelong cancer risk. \emph{Phil. Trans. R. Soc. B}. \textbf{370}, 20150198 (2015).
 
 \leavevmode\hypertarget{ref-Felsenstein1985}{}%
-40. Felsenstein J. Phylogenies and the Comparative Method. The American
-Naturalist. 1985;125: 1--15.
-doi:\href{https://doi.org/10.1086/284325}{10.1086/284325}
+55. J. Felsenstein, Phylogenies and the Comparative Method. \emph{The American Naturalist}. \textbf{125}, 1--15 (1985).
 
 \leavevmode\hypertarget{ref-MartinsAndHansen1997}{}%
-41. Martins EP, Hansen TF. Phylogenies and the Comparative Method: A
-General Approach to Incorporating Phylogenetic Information into the
-Analysis of Interspecific Data. The American Naturalist. 1997;149:
-646--667. doi:\href{https://doi.org/10.1086/286013}{10.1086/286013}
+56.
+E.
+P.
+Martins, T.
+F.
+Hansen, Phylogenies and the Comparative Method: A General Approach to Incorporating Phylogenetic Information into the Analysis of Interspecific Data.
+\emph{The American Naturalist}.
+\textbf{149}, 646--667 (1997).
+
+\leavevmode\hypertarget{ref-ape2019}{}%
+57. E. Paradis, K. Schliep, Ape 5.0: An environment for modern phylogenetics and evolutionary analyses in R. \emph{Bioinformatics}. \textbf{35}, 526--528 (2019).
 
 \leavevmode\hypertarget{ref-Armitage:19851c2}{}%
-42. Armitage P. Multistage models of carcinogenesis. Environmental
-health perspectives. 1985;63: 195--201.
-doi:\href{https://doi.org/10.1289/ehp.8563195}{10.1289/ehp.8563195}
+58. P. Armitage, Multistage models of carcinogenesis. \emph{Environmental health perspectives}. \textbf{63}, 195--201 (1985).
 
 \leavevmode\hypertarget{ref-Armitage:20041c2}{}%
-43. Armitage P, Doll R. The age distribution of cancer and a multi-stage
-theory of carcinogenesis. British Journal of Cancer. 2004;91: 6602297.
-doi:\href{https://doi.org/10.1038/sj.bjc.6602297}{10.1038/sj.bjc.6602297}
-
-\leavevmode\hypertarget{ref-Peto:20151c2}{}%
-44. Peto R. Quantitative implications of the approximate irrelevance of
-mammalian body size and lifespan to lifelong cancer risk. Phil Trans R
-Soc B. 2015;370: 20150198.
-doi:\href{https://doi.org/10.1098/rstb.2015.0198}{10.1098/rstb.2015.0198}
-
-\leavevmode\hypertarget{ref-ArmitageAndDoll1954}{}%
-45. Armitage P, Doll R. The Age Distribution of Cancer and a Multi-stage
-Theory of Carcinogenesis. British Journal of Cancer. 1954;8: 1--12.
-doi:\href{https://doi.org/10.1038/bjc.1954.1}{10.1038/bjc.1954.1}
-
-\leavevmode\hypertarget{ref-Nordling1953}{}%
-46. Nordling CO. A New Theory on the Cancer-inducing Mechanism. British
-Journal of Cancer. 1953;7: 68--72.
-doi:\href{https://doi.org/10.1038/bjc.1953.8}{10.1038/bjc.1953.8}
-
-\leavevmode\hypertarget{ref-Goodman2009}{}%
-47. Goodman M, Sterner KN, Islam M, Uddin M, Sherwood CC, Hof PR, et al.
-Phylogenomic analyses reveal convergent patterns of adaptive evolution
-in elephant and human ancestries. Proceedings of the National Academy of
-Sciences. 2009;106: 20824--20829.
-doi:\href{https://doi.org/10.1073/pnas.0911239106}{10.1073/pnas.0911239106}
+59. P. Armitage, R. Doll, The age distribution of cancer and a multi-stage theory of carcinogenesis. \emph{British Journal of Cancer}. \textbf{91}, 6602297 (2004).
+
+\leavevmode\hypertarget{ref-R-ape}{}%
+60.
+E.
+Paradis, S.
+Blomberg, B.
+Bolker, J.
+Brown, S.
+Claramunt, J.
+Claude, H.
+S.
+Cuong, R.
+Desper, G.
+Didier, B.
+Durand, J.
+Dutheil, R.
+Ewing, O.
+Gascuel, T.
+Guillerme, C.
+Heibl, A.
+Ives, B.
+Jones, F.
+Krah, D.
+Lawson, V.
+Lefort, P.
+Legendre, J.
+Lemon, G.
+Louvel, E.
+Marcon, R.
+McCloskey, J.
+Nylander, R.
+Opgen-Rhein, A.-A.
+Popescu, M.
+Royer-Carenzi, K.
+Schliep, K.
+Strimmer, D.
+de Vienne, \emph{Ape: Analyses of phylogenetics and evolution} (2020; \url{https://CRAN.R-project.org/package=ape}).
+
+\leavevmode\hypertarget{ref-R-base}{}%
+61. R Core Team, \emph{R: A language and environment for statistical computing} (R Foundation for Statistical Computing, Vienna, Austria, 2019; \url{https://www.R-project.org/}).
+
+\leavevmode\hypertarget{ref-R-bookdown}{}%
+62. Y. Xie, \emph{Bookdown: Authoring books and technical documents with r markdown} (2020; \url{https://CRAN.R-project.org/package=bookdown}).
+
+\leavevmode\hypertarget{ref-R-broom.mixed}{}%
+63. B. Bolker, D. Robinson, \emph{Broom.mixed: Tidying methods for mixed models} (2020; \url{https://CRAN.R-project.org/package=broom.mixed}).
+
+\leavevmode\hypertarget{ref-R-data.table}{}%
+64. M. Dowle, A. Srinivasan, \emph{Data.table: Extension of `data.frame`} (2019; \url{https://CRAN.R-project.org/package=data.table}).
+
+\leavevmode\hypertarget{ref-R-dplyr}{}%
+65. H. Wickham, R. François, L. Henry, K. Müller, \emph{Dplyr: A grammar of data manipulation} (2020; \url{https://CRAN.R-project.org/package=dplyr}).
+
+\leavevmode\hypertarget{ref-R-forcats}{}%
+66. H. Wickham, \emph{Forcats: Tools for working with categorical variables (factors)} (2020; \url{https://CRAN.R-project.org/package=forcats}).
+
+\leavevmode\hypertarget{ref-R-geiger}{}%
+67.
+L.
+Harmon, M.
+Pennell, C.
+Brock, J.
+Brown, W.
+Challenger, J.
+Eastman, R.
+FitzJohn, R.
+Glor, G.
+Hunt, L.
+Revell, G.
+Slater, J.
+Uyeda, J.
+Weir, \emph{Geiger: Analysis of evolutionary diversification} (2020; \url{https://CRAN.R-project.org/package=geiger}).
+
+\leavevmode\hypertarget{ref-R-gghighlight}{}%
+68. H. Yutani, \emph{Gghighlight: Highlight lines and points in 'ggplot2'} (2020; \url{https://CRAN.R-project.org/package=gghighlight}).
+
+\leavevmode\hypertarget{ref-R-ggimage}{}%
+69. G. Yu, \emph{Ggimage: Use image in 'ggplot2'} (2020; \url{https://CRAN.R-project.org/package=ggimage}).
+
+\leavevmode\hypertarget{ref-R-ggnewscale}{}%
+70. E. Campitelli, \emph{Ggnewscale: Multiple fill and colour scales in 'ggplot2'} (2020; \url{https://CRAN.R-project.org/package=ggnewscale}).
+
+\leavevmode\hypertarget{ref-R-ggplot2}{}%
+71.
+H.
+Wickham, W.
+Chang, L.
+Henry, T.
+L.
+Pedersen, K.
+Takahashi, C.
+Wilke, K.
+Woo, H.
+Yutani, D.
+Dunnington, \emph{Ggplot2: Create elegant data visualisations using the grammar of graphics} (2020; \url{https://CRAN.R-project.org/package=ggplot2}).
+
+\leavevmode\hypertarget{ref-R-ggplotify}{}%
+72. G. Yu, \emph{Ggplotify: Convert plot to 'grob' or 'ggplot' object} (2020; \url{https://CRAN.R-project.org/package=ggplotify}).
+
+\leavevmode\hypertarget{ref-R-ggpubr}{}%
+73. A. Kassambara, \emph{Ggpubr: 'Ggplot2' based publication ready plots} (2020; \url{https://CRAN.R-project.org/package=ggpubr}).
+
+\leavevmode\hypertarget{ref-R-ggrepel}{}%
+74. K. Slowikowski, \emph{Ggrepel: Automatically position non-overlapping text labels with 'ggplot2'} (2020; \url{https://CRAN.R-project.org/package=ggrepel}).
+
+\leavevmode\hypertarget{ref-R-ggsci}{}%
+75. N. Xiao, \emph{Ggsci: Scientific journal and sci-fi themed color palettes for 'ggplot2'} (2018; \url{https://CRAN.R-project.org/package=ggsci}).
+
+\leavevmode\hypertarget{ref-R-ggtree}{}%
+76. G. Yu, T. T.-Y. Lam, \emph{Ggtree: An r package for visualization of tree and annotation data} (2020; \url{https://yulab-smu.github.io/treedata-book/}).
+
+\leavevmode\hypertarget{ref-R-kableExtra}{}%
+77. H. Zhu, \emph{KableExtra: Construct complex table with 'kable' and pipe syntax} (2019; \url{https://CRAN.R-project.org/package=kableExtra}).
+
+\leavevmode\hypertarget{ref-R-magick}{}%
+78. J. Ooms, \emph{Magick: Advanced graphics and image-processing in r} (2020; \url{https://CRAN.R-project.org/package=magick}).
+
+\leavevmode\hypertarget{ref-R-magrittr}{}%
+79. S. M. Bache, H. Wickham, \emph{Magrittr: A forward-pipe operator for r} (2014; \url{https://CRAN.R-project.org/package=magrittr}).
+
+\leavevmode\hypertarget{ref-R-nlme}{}%
+80. J. Pinheiro, D. Bates, R-core, \emph{Nlme: Linear and nonlinear mixed effects models} (2020; \url{https://CRAN.R-project.org/package=nlme}).
+
+\leavevmode\hypertarget{ref-R-plotly}{}%
+81.
+C.
+Sievert, C.
+Parmer, T.
+Hocking, S.
+Chamberlain, K.
+Ram, M.
+Corvellec, P.
+Despouy, \emph{Plotly: Create interactive web graphics via 'plotly.js'} (2020; \url{https://CRAN.R-project.org/package=plotly}).
+
+\leavevmode\hypertarget{ref-R-purrr}{}%
+82. L. Henry, H. Wickham, \emph{Purrr: Functional programming tools} (2020; \url{https://CRAN.R-project.org/package=purrr}).
+
+\leavevmode\hypertarget{ref-R-readr}{}%
+83. H. Wickham, J. Hester, R. Francois, \emph{Readr: Read rectangular text data} (2018; \url{https://CRAN.R-project.org/package=readr}).
+
+\leavevmode\hypertarget{ref-R-stargazer}{}%
+84. M. Hlavac, \emph{Stargazer: Well-formatted regression and summary statistics tables} (2018; \url{https://CRAN.R-project.org/package=stargazer}).
+
+\leavevmode\hypertarget{ref-R-stringr}{}%
+85. H. Wickham, \emph{Stringr: Simple, consistent wrappers for common string operations} (2019; \url{https://CRAN.R-project.org/package=stringr}).
+
+\leavevmode\hypertarget{ref-R-tibble}{}%
+86. K. Müller, H. Wickham, \emph{Tibble: Simple data frames} (2020; \url{https://CRAN.R-project.org/package=tibble}).
+
+\leavevmode\hypertarget{ref-R-tidyr}{}%
+87. H. Wickham, L. Henry, \emph{Tidyr: Tidy messy data} (2020; \url{https://CRAN.R-project.org/package=tidyr}).
+
+\leavevmode\hypertarget{ref-R-tidytree}{}%
+88. G. Yu, \emph{Tidytree: A tidy tool for phylogenetic tree data manipulation} (2020; \url{https://yulab-smu.github.io/treedata-book/}).
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+89. H. Wickham, \emph{Tidyverse: Easily install and load the 'tidyverse'} (2019; \url{https://CRAN.R-project.org/package=tidyverse}).
+
+\leavevmode\hypertarget{ref-R-treeio}{}%
+90. G. Yu, \emph{Treeio: Base classes and functions for phylogenetic tree input and output} (2020).
+
+\leavevmode\hypertarget{ref-R-UpSetR}{}%
+91. N. Gehlenborg, \emph{UpSetR: A more scalable alternative to venn and euler diagrams for visualizing intersecting sets} (2019; \url{https://CRAN.R-project.org/package=UpSetR}).
+
+\leavevmode\hypertarget{ref-bookdown2016}{}%
+92. Y. Xie, \emph{Bookdown: Authoring books and technical documents with R markdown} (Chapman; Hall/CRC, Boca Raton, Florida, 2016; \url{https://github.com/rstudio/bookdown}).
+
+\leavevmode\hypertarget{ref-geiger2009}{}%
+93.
+M.
+Alfaro, F.
+Santini, C.
+Brock, H.
+Alamillo, A.
+Dornburg, D.
+Rabosky, G.
+Carnevale, L.
+Harmon, Nine exceptional radiations plus high turnover explain species diversity in jawed vertebrates.
+\emph{Proceedings of the National Academy of Sciences of the United States of America}.
+\textbf{106}, 13410--13414 (2009).
+
+\leavevmode\hypertarget{ref-geiger2011}{}%
+94.
+J.
+Eastman, M.
+Alfaro, P.
+Joyce, A.
+Hipp, L.
+Harmon, A novel comparative method for identifying shifts in the rate of character evolution on trees.
+\emph{Evolution}.
+\textbf{65}, 3578--3589 (2011).
+
+\leavevmode\hypertarget{ref-geiger2012}{}%
+95.
+G.
+Slater, L.
+Harmon, D.
+Wegmann, P.
+Joyce, L.
+Revell, M.
+Alfaro, Fitting models of continuous trait evolution to incompletely sampled comparative data using approximate bayesian computation.
+\emph{Evolution}.
+\textbf{66}, 752--762 (2012).
+
+\leavevmode\hypertarget{ref-geiger2008}{}%
+96. L. Harmon, J. Weir, C. Brock, R. Glor, W. Challenger, GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics}. \textbf{24}, 129--131 (2008).
+
+\leavevmode\hypertarget{ref-geiger2014}{}%
+97.
+M.
+Pennell, J.
+Eastman, G.
+Slater, J.
+Brown, J.
+Uyeda, R.
+Fitzjohn, M.
+Alfaro, L.
+Harmon, Geiger v2.0: An expanded suite of methods for fitting macroevolutionary models to phylogenetic trees.
+\emph{Bioinformatics}.
+\textbf{30}, 2216--2218 (2014).
+
+\leavevmode\hypertarget{ref-ggplot22016}{}%
+98. H. Wickham, \emph{Ggplot2: Elegant graphics for data analysis} (Springer-Verlag New York, 2016; \url{https://ggplot2.tidyverse.org}).
+
+\leavevmode\hypertarget{ref-ggtree2020}{}%
+99. G. Yu, Using ggtree to visualize data on tree-like structures. \emph{Current Protocols in Bioinformatics}. \textbf{69}, e96 (2020).
+
+\leavevmode\hypertarget{ref-ggtree2018}{}%
+100. G. Yu, T. T.-Y. Lam, H. Zhu, Y. Guan, Two methods for mapping and visualizing associated data on phylogeny using ggtree. \emph{Molecular Biology and Evolution}. \textbf{35}, 3041--3043 (2018).
+
+\leavevmode\hypertarget{ref-ggtree2017}{}%
+101.
+G.
+Yu, D.
+Smith, H.
+Zhu, Y.
+Guan, T.
+T.-Y.
+Lam, Ggtree: An r package for visualization and annotation of phylogenetic trees with their covariates and other associated data.
+\emph{Methods in Ecology and Evolution}.
+\textbf{8}, 28--36 (2017).
+
+\leavevmode\hypertarget{ref-plotly2020}{}%
+102. C. Sievert, \emph{Interactive web-based data visualization with r, plotly, and shiny} (Chapman; Hall/CRC, 2020; \url{https://plotly-r.com}).
+
+\leavevmode\hypertarget{ref-tidyverse2019}{}%
+103.
+H.
+Wickham, M.
+Averick, J.
+Bryan, W.
+Chang, L.
+D.
+McGowan, R.
+François, G.
+Grolemund, A.
+Hayes, L.
+Henry, J.
+Hester, M.
+Kuhn, T.
+L.
+Pedersen, E.
+Miller, S.
+M.
+Bache, K.
+Müller, J.
+Ooms, D.
+Robinson, D.
+P.
+Seidel, V.
+Spinu, K.
+Takahashi, D.
+Vaughan, C.
+Wilke, K.
+Woo, H.
+Yutani, Welcome to the tidyverse.
+\emph{Journal of Open Source Software}.
+\textbf{4}, 1686 (2019).
+
+\leavevmode\hypertarget{ref-treeio2020}{}%
+104.
+L.-G.
+Wang, T.
+T.-Y.
+Lam, S.
+Xu, Z.
+Dai, L.
+Zhou, T.
+Feng, P.
+Guo, C.
+W.
+Dunn, B.
+R.
+Jones, T.
+Bradley, H.
+Zhu, Y.
+Guan, Y.
+Jiang, G.
+Yu, Treeio: An r package for phylogenetic tree input and output with richly annotated and associated data.
+\emph{Molecular Biology and Evolution}.
+\textbf{37}, 599--603 (2020).
+
+\leavevmode\hypertarget{ref-Liang2013}{}%
+105. J. Liang, G. B. Mills, AMPK: a contextual oncogene or tumor suppressor? \emph{Cancer research}. \textbf{73}, 2929--35 (2013).
+
+\leavevmode\hypertarget{ref-Launonen2005}{}%
+106. V. Launonen, Mutations in the human LKB1/STK11 gene. \emph{Human Mutation}. \textbf{26}, 291--297 (2005).
+
+\leavevmode\hypertarget{ref-Drost2010}{}%
+107.
+J.
+Drost, F.
+Mantovani, F.
+Tocco, R.
+Elkon, A.
+Comel, H.
+Holstege, R.
+Kerkhoven, J.
+Jonkers, P.
+M.
+Voorhoeve, R.
+Agami, G.
+D.
+Sal, BRD7 is a candidate tumour suppressor gene required for p53 function.
+\emph{Nature Cell Biology}.
+\textbf{12}, 380--389 (2010).
+
+\leavevmode\hypertarget{ref-Burrows2010}{}%
+108.
+A.
+E.
+Burrows, A.
+Smogorzewska, S.
+J.
+Elledge, Polybromo-associated BRG1-associated factor components BRD7 and BAF180 are critical regulators of p53 required for induction of replicative senescence.
+\emph{Proceedings of the National Academy of Sciences of the United States of America}.
+\textbf{107}, 14280--5 (2010).
+
+\leavevmode\hypertarget{ref-Fukuda:20161c2}{}%
+109.
+T.
+Fukuda, Y.
+Iino, M.
+Onuma, B.
+Gen, M.
+Inoue-Murayama, T.
+Kiyono, Expression of human cell cycle regulators in the primary cell line of the African savannah elephant (loxodonta africana) increases proliferation until senescence, but does not induce immortalization.
+\emph{In Vitro Cellular \& Developmental Biology - Animal}.
+\textbf{52}, 20--26 (2016).
+
+\leavevmode\hypertarget{ref-Gomes2011}{}%
+110.
+N.
+M.
+V.
+Gomes, O.
+A.
+Ryder, M.
+L.
+Houck, S.
+J.
+Charter, W.
+Walker, N.
+R.
+Forsyth, S.
+N.
+Austad, C.
+Venditti, M.
+Pagel, J.
+W.
+Shay, W.
+E.
+Wright, Comparative biology of mammalian telomeres: hypotheses on ancestral states and the roles of telomeres in longevity determination.
+\emph{Aging Cell}.
+\textbf{10}, 761--768 (2011).
+
+\leavevmode\hypertarget{ref-Hart1974}{}%
+111.
+R.
+W.
+Hart, R.
+B.
+Setlow, Correlation Between Deoxyribonucleic Acid Excision-Repair and Life-Span in a Number of Mammalian Species.
+\emph{Proceedings of the National Academy of Sciences}.
+\textbf{71}, 2169--2173 (1974).
+
+\leavevmode\hypertarget{ref-Francis1981}{}%
+112.
+A.
+A.
+Francis, W.
+H.
+Lee, J.
+D.
+Regan, The relationship of DNA excision repair of ultraviolet-induced lesions to the maximum life span of mammals.
+\emph{Mechanisms of Ageing and Development}.
+\textbf{16}, 181--189 (1981).
+
+\leavevmode\hypertarget{ref-Tyner:20021c2}{}%
+113.
+S.
+D.
+Tyner, S.
+Venkatachalam, J.
+Choi, S.
+Jones, N.
+Ghebranious, H.
+Igelmann, X.
+Lu, G.
+Soron, B.
+Cooper, C.
+Brayton, S.
+H.
+Park, T.
+Thompson, G.
+Karsenty, A.
+Bradley, L.
+A.
+Donehower, p53 mutant mice that display early ageing-associated phenotypes.
+\emph{Nature}.
+\textbf{415}, 45 (2002).
+
+\leavevmode\hypertarget{ref-Rastogi2005}{}%
+114. S. Rastogi, D. A. Liberles, Subfunctionalization of duplicated genes as a transition state to neofunctionalization. \emph{BMC Evolutionary Biology}. \textbf{5}, 28 (2005).
+
+\leavevmode\hypertarget{ref-QianAndZhang2014}{}%
+115. W. Qian, J. Zhang, Genomic evidence for adaptation by gene duplication. \emph{Genome research}. \textbf{24}, 1356--62 (2014).
+
+\leavevmode\hypertarget{ref-Stoltzfus1999}{}%
+116. A. Stoltzfus, On the Possibility of Constructive Neutral Evolution. \emph{Journal of Molecular Evolution}. \textbf{49}, 169--181 (1999).
+
+\leavevmode\hypertarget{ref-Bininda-Emonds2007}{}%
+117.
+O.
+R.
+P.
+Bininda-Emonds, M.
+Cardillo, K.
+E.
+Jones, R.
+D.
+E.
+MacPhee, R.
+M.
+D.
+Beck, R.
+Grenyer, S.
+A.
+Price, R.
+A.
+Vos, J.
+L.
+Gittleman, A.
+Purvis, The delayed rise of present-day mammals.
+\emph{Nature}.
+\textbf{446}, 507--512 (2007).
 

### 2020-09-08 paper_PLOS/test.Rmd
Penultimate draft

---
title: "test"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{R, child="test.tex"}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-09-08 code/empiricalFDRWebGestaltR.R

@@ -1,0 +1,239 @@
+library(tidyverse)
+options(readr.num_columns = 0)
+library(data.table)
+library(magrittr)
+library(WebGestaltR)
+library(parallel)
+library(ggpubr)
+source("../code/generalFunctions.R")
+source("../code/R_rainclouds.R")
+source("../code/summarySE.R")
+
+dir.create("../output/empiricalORA", recursive = T)
+
+no_cores <- detectCores() - 2 # I want to still use my laptop...
+cl <- makeCluster(no_cores, outfile="")
+clusterExport(cl, "WebGestaltR")
+
+set.seed(1234)
+
+reactome.ORA <- function(geneList, refList, is_output=F, fdr.thres=1){
+  suppressWarnings(require(WebGestaltR, quietly = T))
+  suppressMessages(suppressWarnings(WebGestaltR(
+    enrichMethod = "ORA",
+    enrichDatabase = "pathway_Reactome",
+    organism = "hsapiens",
+    interestGene = geneList,
+    interestGeneType = "genesymbol",
+    referenceGene = refList,
+    referenceGeneType = "genesymbol",
+    maxNum=1000,
+    fdrMethod="BH",
+    sigMethod="fdr",
+    fdrThr = fdr.thres,
+    reportNum=100,
+    isOutput = is_output,
+    # outputDirectory = outputDirectory,
+    hostName="http://www.webgestalt.org/"
+  )))
+}
+
+gene_list <- read_lines("../output/recBlastDBPrep/AvA_geneList_filtered.txt")
+geneCopyTable <- read_csv("../output/geneCopyTable/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata-GeneCopyTable_RBB_filtered_long.csv")
+genomeTable <- read_csv("../output/other/genomeTable.csv")
+
+geneCopyTable.stats <- geneCopyTable %>% 
+  group_by(Genome) %>% 
+  summarize(N.Hits = sum(!is.na(ECNC) & !is.na(Copy)),
+            N.Dup = sum(ECNC>=1.5 & Copy>1, na.rm = T))
+set.sizes <-c(1,3,5,9,12,15,28,35,52,61,76,83,94,157,383,484,504,587,1000,1591, 2103, 3000, 4000, 5000) %>% 
+  set_names(., as.character(.))
+
[…]

### 2020-09-08 code/generalFunctions.R

@@ -434,6 +434,21 @@ binTableFromList <- function (input) {
   data[, Gene:=elements][]
 }
 
+fromList <- function (input) {
+  # Same as original fromList()...
+  elements <- unique(unlist(input))
+  data <- unlist(lapply(input, function(x) {
+    x <- as.vector(match(elements, x))
+  }))
+  data[is.na(data)] <- as.integer(0)
+  data[data != 0] <- as.integer(1)
+  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
+  data <- data[which(rowSums(data) != 0), ]
+  names(data) <- names(input)
+  # ... Except now it conserves your original value names!
+  row.names(data) <- elements
+  return(data)
+}
 
 phylopic_uid_item_safer <- function (name) {
   x <- gsub("[^a-zA-Z]+", "+", name)
@@ -472,24 +487,30 @@ lineage.pathway.fdr.table <- function(path="output/ORA"){
       Database = ORAs %>% sapply(str_extract, "(?<=pathway_)[A-Za-z]+(?=-)"),
       FDR.Threshold = ORAs %>% sapply(str_extract, "(?<=FDR_)[0-9]\\.[0-9]+(?=/)"),
       Direction.lineage = ORAs %>% sapply(str_extract, "(?<=Project_).+?(?=/)"),
-      Pathway.EnrichRatio.FDR.Genes.Size = ORAs %>% sapply(
+      Pathway.EnrichRatio.FDR.pval.Genes.Size = ORAs %>% sapply(
         ., . %>% 
           read_tsv %>% 
-          select(description, enrichmentRatio, FDR, userId, size) %>% 
+          select(description, enrichmentRatio, FDR, pValue, userId, size) %>% 
           mutate_all(. %>% str_replace_all(",|;", ":")) %>% 
-          unite("PEF", description, enrichmentRatio, FDR, userId, size, sep = ",") %>% 
+          unite("PEF", description, enrichmentRatio, FDR, pValue, userId, size, sep = ",") %>% 
           pull(PEF) %>% 
           paste0(collapse=";")),
       fname = ORAs
     ) %>% 
     separate(Direction.lineage, c("Direction", "Ancestor", "to", "Node"), sep="_", extra = "merge") %>% 
-    separate_rows(Pathway.EnrichRatio.FDR.Genes.Size, sep=";") %>% 
[…]

### 2020-09-08 code/pub_ORATable.R

@@ -1,0 +1,39 @@
+#!/usr/bin/env Rscript
+
+library(data.table)
+library(stringr)
+library(magrittr)
+source("code/generalFunctions.R")
+
+
+# ORA.all.f <- "../output/pubFiles/ORA_table_pub.csv"
+# ORA.root.dir <- "../output/ORA/maxLikelihood_model_MK+FQ+I+G4-dataType_MORPH-asrMin_0.8/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata_RBB_filtered_dyn/pathway_Reactome-top100/"
+ORA.all.f <- snakemake@output[[1]]
+ORA.root.dir <- snakemake@input[[1]]
+
+dir.create(dirname(ORA.all.f), showWarnings = F, recursive = T)
+
+
+# ORA.root.dir <- "../output/ORA"
+# ORA.root.dir %>% 
+#   lineage.pathway.fdr.table %>% 
+#   mutate(
+#     EnrichRatio = as.numeric(EnrichRatio), FDR=as.numeric(FDR), Size=as.numeric(Size),
+#     GeneList = Genes %>% str_replace_all(";", "\n") %>% str_trunc(100, "center", ellipsis = "...\n"),
+#     Class = class.pathway(Pathway)
+#   ) %>% 
+#   fwrite(path=ORA.all.f)
+
+
+a <- dir(ORA.root.dir, pattern="Project_*", full.names = T) %>% 
+  (function(x){x[str_detect(x, "_stable_", negate = T)]}) %>% 
+  dir(pattern="enrichment_results", full.names=T) %>% 
+  set_names(., str_extract(., "(?<=enrichment_results_(increased|genesDuplicated)_).+(?=.txt)"))
+b <- lapply(
+  names(a), 
+  function(n, df=a){fread(df[[n]])[, branch:=n][]}
+) %>% 
+  rbindlist
+# b %>% split(., .$branch) %>% write.xlsx(ORA.all.f)
+b %>% 
+  fwrite(ORA.all.f)

### 2020-09-13 publication/manuscript.Rmd
Ready for BioRxiv

---
title: "Pervasive duplication of tumor suppressors in Afrotherians preceded evolution of large bodies and reduced cancer risk in elephants"
author:
- name: Juan Manuel Vazquez
  affiliation: "University of Chicago"
- affiliation: "SUNY Buffalo"
  name: Vincent J Lynch
address:
- address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
  code: "University of Chicago"
- address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
  code: "SUNY Buffalo"
always_allow_html: true
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    number_sections: yes
    includes:
      in_header: preamble.tex
  bookdown::html_document2: default
  bookdown::pdf_document2: default
  bookdown::word_document2: default
csl: science.csl
editor_options:
  chunk_output_type: console
bibliography: ["rbhb_paenungulates.bib", "packages.bib"]
abstract: |
  The risk of developing cancer is correlated with body size and lifespan within
  species. Between species, there is no correlation between body size or 
  lifespan, and cancer, suggesting that large, long-lived species have evolved 
  enhanced cancer protection mechanisms. Elephants and their relatives 
  (Proboscideans) are a particularly interesting lineage in which to explore 
  the mechanisms underlying the evolution of augmented cancer resistance because
  they evolved large bodies recently within a clade of smaller bodied species 
  (Atlantogenatans). Here we explore the contribution of tumor suppressor 
  duplication to body size and cancer risk in Atlantogenatans. Unexpectedly, we 
  found that tumor suppressor duplication was pervasive in Atlantogenatan 
  genomes rather than restricted to Proboscideans. Proboscideans, however, have 
  duplicates in unique pathways that may underlie some aspects of their 
  remarkable anti-cancer cell biology. These data suggest that duplication of 
  tumor suppressor genes facilitated the evolution of increased body size by 
  compensating for decreasing intrinsic cancer risk.

---

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S.\thechapter.\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S.\thechapter.\arabic{figure}}%
     }

<!-- \newcommand{\terminatesupplement}{% -->
<!--         \setcounter{table}{0} -->
<!--         \renewcommand{\thetable}{\thechapter.\arabic{table}}% -->
<!--         \setcounter{figure}{0} -->
<!--         \renewcommand{\thefigure}{\thechapter.\arabic{figure}}% -->
<!--      } -->

<!--_Text based on plos sample manuscript, see [http://journals.plos.org/ploscompbiol/s/latex](http://journals.plos.org/ploscompbiol/s/latex)_-->

```{r Libraries and functions, include = F}
knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, out.width="6in", fig.pos = "H", out.extra = "",
                      fig.retina = 2, dpi = 300, dev=c('cairo_pdf', 'svg', 'png','cairo_ps'))

if(basename(getwd()) != "paper_PLOS") setwd("paper_PLOS")

set.seed(1234)
library(data.table)
library(tidyverse)
library(magrittr)
library(viridis)
options(readr.num_columns = 0)
library(ape)
library(geiger)
library(nlme)
library(treeio)
library(tidytree)
library(ggtree)
library(ggimage)
library(ggpubr)
library(ggsci)
library(ggstance)
library(plotly)
library(ggplotify)
library(UpSetR)
library(ggrepel)
library(magick)
library(ggrepel)
library(gghighlight)
library(broom.mixed)
library(purrr)
library(kableExtra)
library(stargazer)
library(ggnewscale)
library(ComplexHeatmap)
library(rms)
library(patchwork)
library(ggwordcloud)
source("../code/generalFunctions.R")
```
```{r Data Files, include = F}
UP000005640 <- read_tsv("../data/input/UP000005640.withheader.tsv")
gene_list <- read_lines("../output/recBlastDBPrep/AvA_geneList.txt")
gene_list_filtered <- read_lines("../output/recBlastDBPrep/AvA_geneList_filtered.txt")
genomeTable <- read_csv("../output/other/genomeTable.csv")

stabletraits.progress <- read_tsv("../data/stableTraits/eutheria.progress")

time.tree <- read.tree("../data/stableTraits/eutheria.tree")
time.tree.atlantogenata <- tree_subset(time.tree, MRCA(time.tree, "Elephas.maximus", "Dasypus.novemcinctus"), levels_back = 0)

#################################################################################

tree.Atlantogenata.table <- read_csv("../output/pubFiles/cancerRisk-Atlantogenata.csv") %>% 
  mutate(label=str_replace_all(label, " ", "."))

class(tree.Atlantogenata.table) %<>% str_replace("spec_tbl_df", "tbl_tree")

tree.Atlantogenata <- tree.Atlantogenata.table %>% as.treedata()

#################################################################################

ml.treedata <- read.beast("../output/geneLists/maxLikelihood_model_MK+FQ+I+G4-dataType_MORPH-asrMin_0.8/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata_RBB_filtered_dyn/atlantogenata_RBB_filtered_dyn.nexus")

#################################################################################

geneCopyTable <- read_csv("../output/geneCopyTable/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata-GeneCopyTable_RBB_filtered_long.csv")

#################################################################################

ORA.all.f <- "../output/pubFiles/ORA_table_pub.csv"
ORA.all <- fread(ORA.all.f)[,
                              branch:=str_replace_all(branch, "_", ".")][, 
                            c("parent", "node") := tstrsplit(branch, ".to.")
                            ][, Class:=class.pathway(description)][pValue<=0.05]
ORA.all.Proboscidea <- ORA.all[is.na(node)][, nlPval := -log(pValue)][, node:=parent][]

#################################################################################

loxAf4.gene_transcript <- fread(
  "../output/RBHBID-to-StringTieTranscriptID/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/loxAfr4_RBB_ID-TranscriptID.tsv", 
  sep = "\t",
  header=F,
  col.names = c("Locus", "details")
  )[, 
    geneID := tstrsplit(details, ";", fixed=TRUE),
    by=Locus][,
      .(geneID = str_extract(geneID, '(?<=gene_id ")[A-Za-z0-9.]+(?=")')),
    by=Locus] %>% 
  unique

triManLat2.gene_transcript <- fread(
  "../output/RBHBID-to-StringTieTranscriptID/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/triManLat2_RBB_ID-TranscriptID.tsv", 
  sep = "\t",
  header=F,
  col.names = c("Locus", "details")
  )[, 
    geneID := tstrsplit(details, ";", fixed=TRUE),
    by=Locus][,
      .(geneID = str_extract(geneID, '(?<=gene_id ")[A-Za-z0-9.]+(?=")')),
    by=Locus] %>% 
  unique

#################################################################################

loxAfr4.t.f <- "../output/pubFiles/loxafr4_transcripts.csv"
if (file.exists(loxAfr4.t.f)){
  loxAfr4.transcripts <- fread(loxAfr4.t.f)
} else{
  loxAfr4.transcripts <- getTPM.dt("loxAfr4")
  dir.create(dirname(loxAfr4.t.f), showWarnings = F)
  loxAfr4.transcripts %>% fwrite(file = loxAfr4.t.f)
}

loxAfr4.geneTPM <- merge(
  loxAf4.gene_transcript,
  loxAfr4.transcripts,
  by="geneID"
  )[, Gene :=  str_remove(Locus, "_\\d+$")][]

setkey(loxAfr4.geneTPM, Gene)
setcolorder(loxAfr4.geneTPM)

triManLat2.t.f <- "../output/pubFiles/triManLat2_transcripts.csv"
if (file.exists(triManLat2.t.f)){
  triManLat2.transcripts <- fread(triManLat2.t.f)
} else{
  triManLat2.transcripts <- getTPM.dt("triManLat2")
  dir.create(dirname(triManLat2.t.f), showWarnings = F)
  triManLat2.transcripts %>% fwrite(file = triManLat2.t.f)
}

triManLat2.geneTPM <- merge(
  triManLat2.gene_transcript,
  triManLat2.transcripts,
  by="geneID"
  )[, Gene :=  str_remove(Locus, "_\\d+$")][]

setkey(triManLat2.geneTPM, Gene)
setcolorder(triManLat2.geneTPM)

loxAfr4.expressedDups <- loxAfr4.geneTPM %>% 
  group_by(Gene) %>% 
  dplyr::summarize(n = n_distinct(Locus)) %>% 
  filter(n>1) %>% 
  pull(Gene)

###########################################################################################################

genomeTable.atlantogenata <- genomeTable %>% filter(label %in% ml.treedata@phylo$tip.label)

###########################################################################################################

CEGMA.qc <- read_csv("../output/CEGMA/CEGMA_QC.csv") %>%
    separate_rows(Value, sep = ",") %>% 
  mutate(
    Genome = as.factor(Genome), 
    Stat = as.factor(Stat), 
    Value = Value %>% str_trim(side="both") %>% str_split(" ") %>% sapply(., "[", 1),
    Number = Value %>% str_extract("\\d+(\\.\\d+)?$") %>% as.numeric(),
    Type = Value %>% str_extract("^[A-Za-z\\.]+") %>% str_c(" (", ., ")") %>% if_else(is.na(.), "", .),
    NewStat = str_c(Stat, Type)
    ) %>% 
  select(-Value, -Type, -Stat, Genome, Stat=NewStat, Value=Number)
CEGMA.qc[is.na(CEGMA.qc$Value),]$Value <- 0
CEGMA.qc %<>% pivot_wider(id_cols=Genome, names_from = Stat, values_from=Value)
CEGMA.completeness <- read_csv("../output/CEGMA/CEGMA_Completeness_clean.csv")

###########################################################################################################

anage_full <- read_tsv("../data/AnAge/anage_build14.txt", col_types = list("Weaning (days)"=col_number(), "Weaning weight (g)"=col_number(), "References" = col_character())) %>% unite("label", Genus, Species, sep=".")
anage <- anage_full %>% 
  select(label,Family, "Maximum longevity (yrs)", "Adult weight (g)") %>% 
  mutate(lnSize.anage = sapply(
    `Adult weight (g)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  ),
  Lifespan = sapply(
    `Maximum longevity (yrs)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  )
  ) %>% 
  select(label, Family, lnSize.anage, Lifespan)

###########################################################################################################

ORA.term.stats <- fread("../output/empiricalORA/pValueStats_empiricalORA_seed1234_FDR1.csv")
stats.empirical.ds <- fread("../output/empiricalORA/stats_empiricalORA_seed1234_FDR100.csv")

```

```{r scales}
size.scale <- tree.Atlantogenata.table %>% pull(lnSize) %>% unique %>% as.numeric() %>% sort
size.scale.min <- size.scale[1]
size.scale.max <- size.scale[length(size.scale)]
get.size.color <- function(n, scale = size.scale){
  scale.colors = viridis::viridis(n=length(scale)) %>% 
    set_names(., as.character(scale))
  scale.colors[as.character(n)]
}

lifespan.scale <- tree.Atlantogenata.table %>% pull(Lifespan) %>% unique %>% as.numeric() %>% sort
lifespan.scale.min <- lifespan.scale[1]
lifespan.scale.max <- lifespan.scale[length(lifespan.scale)]
get.lifespan.color <- function(n, scale = lifespan.scale){
  scale.colors = viridis::viridis(n=length(scale), option = "C") %>% 
    set_names(., as.character(scale))
  scale.colors[as.character(n)]
}

scale.log2CancerSucceptabilityChange <- tree.Atlantogenata.table %>% 
  pull(log2CancerSucceptabilityChange) %>% 
  unique %>% 
  as.numeric() %>% 
  sort %>% 
  set_names(.,.)
scale.log2CancerSucceptabilityChange.min <- min(scale.log2CancerSucceptabilityChange)
scale.log2CancerSucceptabilityChange.max <- scale.log2CancerSucceptabilityChange[length(scale.log2CancerSucceptabilityChange)]

get.log2CancerSucceptabilityChange.color <- function(n, scale = scale.log2CancerSucceptabilityChange){
  scale.colors = viridis::viridis(n=length(scale)) %>% 
    set_names(., as.character(scale))
  scale.colors[as.character(n)]
}

clade.highlight <- c("Loxodontini", "Loxodonta africana", "Loxodona", 
                     "Loxodonta cyclotis", "Palaeoloxodon antiquus", 
                     "Elephantidae", "Elephantina", "Elephas maximus",
                     "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", 
                     "Proboscidea", "Mammut americanum", 
                     "Tethytheria", "Trichechus manatus", 
                     "Paenungulata", "Procavia capensis", 
                     "Pseudoungulata", "Orycteropus afer", 
                     "Macroscelidae", "Elephantulus edwardii", 
                     "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", 
                     "Afrotheria", 
                     "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", 
                     "Atlantogenata") %>% str_replace_all(" ", ".")

tree.Atlantogenata.table.RICR <- tree.Atlantogenata.table %>%
  mutate(
    K1 = (Lifespan)^6 * exp(lnSize),
    lnK1 = log(K1),
    K1.low = (Lifespan.low)^6 * exp(lnSize.low),
    lnK1.low = log(K1.low),
    K1.high = (Lifespan.high)^6 * exp(lnSize.high),
    lnK1.high = log(K1.high),
  )
tree.Atlantogenata.table.RICR2 <- tree.Atlantogenata.table.RICR %>% mutate(
  K2 = parent %>% 
      sapply(
        .,
        function(n, d=tree.Atlantogenata.table.RICR){
          d %>% filter(node==n) %>% pull(K1) %>% unique
        }
      ),
    lnK2 = parent %>% 
      sapply(
        .,
        function(n, d=tree.Atlantogenata.table.RICR){
          d %>% filter(node==n) %>% pull(lnK1) %>% unique
        }
      ),
    CancerSucceptabilityChange = K2/K1,
    log2CancerSucceptabilityChange = log2(K2)-log2(K1),
    K2.low = parent %>% 
      sapply(
        .,
        function(n, d=tree.Atlantogenata.table.RICR){
          d %>% filter(node==n) %>% pull(K1.low) %>% unique
        }
      ),
    lnK2.low = parent %>% 
      sapply(
        .,
        function(n, d=tree.Atlantogenata.table.RICR){
          d %>% filter(node==n) %>% pull(lnK1.low) %>% unique
        }
      ),
    CancerSucceptabilityChange.low = K2.low/K1.low,
    log2CancerSucceptabilityChange.low = log2(K2.low)-log2(K1.low),
    K2.high = parent %>% 
      sapply(
        .,
        function(n, d=tree.Atlantogenata.table.RICR){
          d %>% filter(node==n) %>% pull(K1.high) %>% unique
        }
      ),
    lnK2.high = parent %>% 
      sapply(
        .,
        function(n, d=tree.Atlantogenata.table.RICR){
          d %>% filter(node==n) %>% pull(lnK1.high) %>% unique
        }
      ),
    CancerSucceptabilityChange.high = K2.high/K1.high,
    log2CancerSucceptabilityChange.high = log2(K2.high)-log2(K1.high),
  )

dat.atlantogenata <- tree.Atlantogenata.table.RICR2 %>% 
  filter(label %in% clade.highlight) %>% 
  mutate(label = label %>% factor(., levels = clade.highlight)) %>% 
  dplyr::select(-parent, -node, -branch.length)

size.scale.expanded <- c(size.scale, dat.atlantogenata$lnSize.low, dat.atlantogenata$lnSize.high) %>% unique %>% as.numeric() %>% sort
size.scale.expanded.min <- size.scale.expanded[1]
size.scale.expanded.max <- size.scale.expanded[length(size.scale.expanded)]

lifespan.scale.expanded <- c(lifespan.scale, tree.Atlantogenata.table$Lifespan.low, tree.Atlantogenata.table$Lifespan.high) %>% unique %>% as.numeric() %>% sort
lifespan.scale.expanded.min <- lifespan.scale.expanded[1]
lifespan.scale.expanded.max <- lifespan.scale.expanded[length(lifespan.scale.expanded)]
```

```{r Query List, include=F}
UP000005640.full <- UP000005640 %>% 
  filter(!is.na(Name))
UP000005640.genenames <- unique(UP000005640.full$Name)
query.binTable <- list(AvA = gene_list, UP000005640 = UP000005640.genenames) %>% 
  binTableFromList() %>% 
  left_join(., UP000005640, by=c("Gene"="Name")) %>% 
  mutate(
    "C.ORF" = str_detect(Gene, "^C[0-9XY]{1,2}orf\\d+"),  # No C-Orfs
    "LOC" = str_detect(Gene, "^LOC\\d+"),   # No uncharacterized loci
    "LINC" = str_detect(Gene, "^LINC\\d+"),   # No Long "Non-Coding" RNA
    "HLA" = str_detect(Gene, "^HLA|^HCP5|^CD74"), # No HLAs
    "Histone" = str_detect(Gene, "^H[1-4]|^HIST\\d"), # No Histones
    "ZincFinger" = str_detect(Gene, "^ZNF\\d+"),   # No Zinc Fingers
    "Riboprotein" = str_detect(Gene, "^RP[LS]\\d+"),   # No Riboproteins
    "OlfactoryReceptor" = str_detect(Gene, "^OR\\d+"),   # No Olfactory Receptors
    "Uncharacterized" = str_detect(header, "[uU]ncharacterized"),  # No Uncharacterized proteins
    "Putative" = str_detect(header, "[pP]utative"),  # No putative proteins
    "Viral" = str_detect(header, "[pP]olyprotein|ERVK|HERV"),  # No viral proteins
    "Fragment" = str_detect(header, "[Ff]ragment")  # No fragment proteins
  )
```

```{r Gene Copy Trees, include=F}
ml.tree <- ml.treedata@phylo 
ml.treedata.table <- ml.treedata %>% 
  as_tibble() %>% 
  mutate_at(
    vars(-parent, -node, -branch.length, -label),
    . %>% as.character %>% if_else(is.na(.), "?", .) %>% if_else(.=="32", "32+", .)
  )

ml.tree %<>% full_join(., genomeTable.atlantogenata, by="label")
ml.treedata.table <- ml.treedata %>% 
  as_tibble()

bodysize.tree.atlantogenata <- tree.Atlantogenata.table %>% 
  filter(label %in% ml.treedata.table$label) %>% 
  select(label, lnSize)

ml.tree.nodes <- list(
  "Loxodonta.africana"= ml.treedata.table %>% filter(label=="Loxodonta.africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= ml.treedata.table %>% filter(label=="Trichechus.manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= ml.treedata.table %>% filter(label=="Procavia.capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= ml.treedata.table %>% filter(label=="Orycteropus.afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= ml.treedata.table %>% filter(label=="Choloepus.hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = ml.treedata.table %>% filter(label=="Dasypus.novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = ml.treedata.table %>% filter(label=="Chrysochloris.asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = ml.treedata.table %>% filter(label=="Echinops.telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = ml.treedata.table %>% filter(label=="Elephantulus.edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = ml.treedata.table %>% filter(label=="Loxodonta.cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = ml.treedata.table %>% filter(label=="Palaeoloxodon.antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = ml.treedata.table %>% filter(label=="Elephas.maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = ml.treedata.table %>% filter(label=="Mammuthus.columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = ml.treedata.table %>% filter(label=="Mammuthus.primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = ml.treedata.table %>% filter(label=="Mammut.americanum") %>% pull(node) %>% unique,
  "Afrotheria" =  ml.treedata %>% MRCA("Loxodonta.africana", "Chrysochloris.asiatica"),
  "Afroinsectivora" =  ml.treedata %>% MRCA("Elephantulus.edwardii", "Chrysochloris.asiatica"),
  "Afrosoricida" =  ml.treedata %>% MRCA("Echinops.telfairi", "Chrysochloris.asiatica"),
  "Pseudoungulata" =  ml.treedata %>% MRCA("Loxodonta.africana", "Orycteropus.afer"),
  "Paenungulata" =  ml.treedata %>% MRCA("Loxodonta.africana", "Procavia.capensis"),
  "Tethytheria" =  ml.treedata %>% MRCA("Loxodonta.africana", "Trichechus.manatus"),
  "Proboscidea" =  ml.treedata %>% MRCA("Loxodonta.africana", "Mammut.americanum"),
  "Elephantidae" =  ml.treedata %>% MRCA("Loxodonta.africana", "Elephas.maximus"),
  "Loxodontini" =  ml.treedata %>% MRCA("Loxodonta.africana", "Loxodonta.cyclotis"),
  "Loxodona" =  ml.treedata %>% MRCA("Loxodonta.cyclotis", "Palaeoloxodon.antiquus"),
  "Elephantina" =  ml.treedata %>% MRCA("Elephas.maximus", "Mammuthus.columbi"),
  "Mammuthus" =  ml.treedata %>% MRCA("Mammuthus.columbi", "Mammuthus.primigenius"),
  "Root" = ml.treedata %>% MRCA("Loxodonta.africana", "Dasypus.novemcinctus")
)

comparisons <- list(
  c("Proboscidea", "Elephantidae"),
  c("Elephantidae", "Elephantina"),
  c("Elephantina", "Elephas.maximus"),
  c("Elephantina", "Mammuthus"),
  c("Mammuthus", "Mammuthus.columbi"),
  c("Mammuthus", "Mammuthus.primigenius"),
  c("Elephantidae", "Loxodontini"),
  c("Loxodontini", "Loxodonta.africana"),
  c("Loxodontini", "Loxodona"),
  c("Loxodona", "Loxodonta.cyclotis"),
  c("Loxodona", "Palaeoloxodon.antiquus"),
  c("Proboscidea", "Mammut.americanum"),
  c("Tethytheria", "Proboscidea"),
  c("Tethytheria", "Trichechus.manatus"),
  c("Paenungulata", "Tethytheria"),
  c("Paenungulata", "Procavia.capensis"),
  c("Pseudoungulata", "Paenungulata"),
  c("Pseudoungulata", "Orycteropus.afer"),
  c("Afrotheria", "Pseudoungulata"),
  c("Afrotheria", "Afroinsectivora"),
  c("Afroinsectivora", "Elephantulus.edwardii"),
  c("Afroinsectivora", "Afrosoricida"),
  c("Afrosoricida", "Echinops.telfairi"),
  c("Afrosoricida", "Chrysochloris.asiatica"),
  c("Root", "Afrotheria"),
  c("Root", "Choloepus.hoffmanni"),
  c("Root", "Dasypus.novemcinctus")
) %>% 
  set_names(., sapply(.,"[[", 2)) %>% 
  sapply(.,. %>% paste0(., collapse="-to-") %>% paste0("increased-", ., collapse = ""))

ml.tree.increasing.geneLists <- comparisons %>% 
  lapply(
    .,
    function(f){
      dir(
        path = "../output/geneLists/maxLikelihood_model_MK+FQ+I+G4-dataType_MORPH-asrMin_0.8/AvA-pcScore0.1_pcIdent0.8_pcQuerySpan0.5_reverse-hg38_maskRep_noVarChr_fragWithGenes/atlantogenata_RBB_filtered_dyn/", 
        pattern=paste0(f, ".*"), 
        full.names = T
        ) %>% 
        read_lines() %>% 
        paste0(collapse = ",")
      }
  )

ml.tree.increasing.geneCounts <- ml.tree.increasing.geneLists %>% 
  lapply(., . %>% str_split(",") %>% unlist %>% stringi::stri_remove_empty(.) %>% length) %>% 
  as_tibble() %>% 
  pivot_longer(everything(), names_to = c("label"), values_to = "N.Genes")

ml.tree.increasing.geneCounts <- full_join(ml.tree, ml.tree.increasing.geneCounts, by="label")

ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts %>% 
  as_tibble %>% 
  full_join(., bodysize.tree.atlantogenata, by="label") %>%
  mutate(delta.GeneCN=log(branch.length)) %>% 
  as.treedata()

ml.tree.increasing.geneCounts.withSize@phylo$node.label[ml.tree.increasing.geneCounts@phylo$node.label == "Xenarthra"] <- NA

ml.tree.increasing.geneCounts.withSize.afrotheria.table <- ml.tree.increasing.geneCounts.withSize %>%
  drop.tip(c("Dasypus.novemcinctus", "Choloepus.hoffmanni")) %>% 
  as_tibble %>% mutate(Spc.short.bold = Spc.short %>% str_c('bold("', ., '")'),
                       joint.label = str_c(label, ": ", N.Genes)) 
# ml.tree.increasing.geneCounts.withSize.afrotheria.table$lnSize[which(ml.tree.increasing.geneCounts.withSize.afrotheria.table$label=="Root")] <- bsize.tree %>% as_tibble() %>% filter(node == MRCA(bsize.tree, "Loxodonta.africana", "Dasypus.novemcinctus")) %>% pull(lnSize) %>% as.numeric
ml.tree.increasing.geneCounts.withSize.afrotheria <- ml.tree.increasing.geneCounts.withSize.afrotheria.table%>% 
  as.treedata()
#Prep work
ml.tree.increasing.geneCounts.withSize.afrotheria.table[which(ml.tree.increasing.geneCounts.withSize.afrotheria.table$label %in% genomeTable.atlantogenata$label),]$joint.label <- NA
ml.tree.increasing.geneCounts.withSize.afrotheria <- ml.tree.increasing.geneCounts.withSize.afrotheria.table %>% as.treedata()
```

```{r Figure1Prep}

groups.atlantogenata <- list(
  # "Atlantogenata",
  # "Elephantidae" = c("Loxodonta.africana", "Loxodonta.cyclotis", "Palaeoloxodon.antiquus", "Elephas.maximus", "Mammuthus.columbi", "Mammuthus.primigenius"),
  "Proboscidea" = c("Loxodonta.africana", "Loxodonta.cyclotis", "Palaeoloxodon.antiquus", "Elephas.maximus", "Mammuthus.columbi", "Mammuthus.primigenius", "Mammut.americanum"),
  "Sirenia" = c("Trichechus.manatus", "Hydrodamalis.gigas", "Dugong.dugon"),
  "Hyracoidea" = c("Procavia.capensis", "Titanohyrax.andrewsi")#,
  # "Pseudoungulata" = c("Orycteropus.afer"),
  # "Xenarthra" = c("Dasypus.novemcinctus", "Choloepus.hoffmanni"),
  # "Afrosoricida" = c("Chrysochloris.asiatica"), 
  # "Macroscelidea" = c("Elephantulus.edwardii", "Echinops.telfairi")
)

anage.genomeSpecies <- anage %>% filter(label %in% genomeTable.atlantogenata$label)
anage.genomeSpecies <- anage.genomeSpecies %>% select(label, Lifespan) %>% 
  full_join(tree.Atlantogenata.table %>% 
              filter(label %in% genomeTable.atlantogenata$label) %>% 
              select(label, Size=Median, phyloPicUID), 
            by = "label") %>% 
  mutate(kgSize = Size %>% exp %>% divide_by(1000))

anage.genomeSpecies.plus <- anage.genomeSpecies %>% 
  bind_rows(
    tree.Atlantogenata.table %>% 
      filter(label %in% c("Hydrodamalis.gigas", "Titanohyrax.andrewsi", "Dugong.dugon")) %>% 
      select(label, Size=Median, phyloPicUID) %>% 
      mutate(kgSize = Size %>% exp %>% divide_by(1000))
  ) %>% 
  mutate(
    label = label %>% str_replace(" ", "."),
    clade=sapply(
      label, 
      function(x, y=split(rep(names(groups.atlantogenata), lengths(groups.atlantogenata)), unlist(groups.atlantogenata))){
        y[[x]] %>% ifelse(is.null(.), "", .)
      }
    ),
    Size.norm=Size %>% divide_by(max(Size)) %>% multiply_by(0.09),
    Size.log = Size
    )

# A few additional phyloPics:
anage.genomeSpecies.plus$phyloPicUID[which(anage.genomeSpecies.plus$label == "Hydrodamalis.gigas")] <- "23389deb-fe10-4c97-87c4-a606ebd1fec3"
anage.genomeSpecies.plus$phyloPicUID[which(anage.genomeSpecies.plus$label == "Titanohyrax.andrewsi")] <- "89845940-5587-425c-a7db-7f18bf1f7df7"
anage.genomeSpecies.plus$phyloPicUID[which(anage.genomeSpecies.plus$label == "Dugong.dugon")] <- "23389deb-fe10-4c97-87c4-a606ebd1fec3"

time.tree.atlantogenata.genomeSpecies.table <- time.tree.atlantogenata %>% 
  drop.tip(time.tree.atlantogenata$tip.label[!time.tree.atlantogenata$tip.label %in% anage.genomeSpecies.plus$label]) %>% 
  full_join(anage.genomeSpecies.plus %>% select(label, phyloPicUID, Size.norm, Size.log), by="label") %>% 
  as_tibble %>%
  mutate(
    prettyLabel = label %>% str_replace("\\.", " "),
    group=sapply(
    label, 
    function(x, y=split(rep(names(groups.atlantogenata), lengths(groups.atlantogenata)), unlist(groups.atlantogenata))){
      y[[x]] %>% ifelse(is.null(.), "", .)
    }
    )
    )

# Denote extinct species with cross

ded <- "^†~"
extinct.spc <- c("Titanohyrax.andrewsi", "Hydrodamalis.gigas", "Palaeoloxodon.antiquus", "Mammuthus.columbi", "Mammuthus.primigenius", "Mammut.americanum")

time.tree.atlantogenata.genomeSpecies.table$prettyLabel[which(time.tree.atlantogenata.genomeSpecies.table$label %in% extinct.spc)] %<>% str_c(., "'^'", "\u2020") #%>% str_replace(" ", "~") #str_c(., ded)
time.tree.atlantogenata.genomeSpecies.table$prettyLabel %<>% 
  # str_replace_all(" ", "~") %>% 
  str_c("'", ., "'")
# time.tree.atlantogenata.genomeSpecies.table$label[which(time.tree.atlantogenata.genomeSpecies.table$label %in% extinct.spc)] %<>% str_c(., ded)
# anage.genomeSpecies.plus$label[which(anage.genomeSpecies.plus$label %in% extinct.spc)] %<>% str_c(., ded)

time.tree.atlantogenata.genomeSpecies <- time.tree.atlantogenata.genomeSpecies.table %>% 
  as.treedata()

plot.label = data.frame(
  .panel = c('Tree', 'Body Size (kg)'), 
  lab = c("MYA", "Body Size (kg)"), 
  x=c(-25,5000), y=-1
  )

atlantogenata.tree.size.lifespan <-
  time.tree.atlantogenata.genomeSpecies %>% 
  ggtree(
  ) %>% 
  flip(tree_view = ., node1 = nodeid(time.tree.atlantogenata.genomeSpecies, "Loxodonta.africana"), node2 = MRCA(time.tree.atlantogenata.genomeSpecies, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis")) %>% 
  revts() +
  geom_tiplab(aes(color=group, image=phyloPicUID, size=I(Size.norm)), geom = "phylopic", align = T, linesize = 0, hjust = 0.5, offset = 10) +
  geom_tiplab(aes(color = group, label=prettyLabel), size=2, align = T, offset = 45, fontface="bold", linesize = 0, parse=T, family="Helvetica", hjust = 0.5
              ) +
  geom_tiplab(aes(subset = (group == ""), image=phyloPicUID, size=I(0.09)), geom = "phylopic", align = T, linesize = 0, color="black", hjust = 0.5, offset=10
              ) +
  geom_tiplab(aes(subset = (group == ""), label=prettyLabel), size=2, align = T, offset = 45, fontface="bold", linesize = 0, color="black", parse=T, family="Helvetica", hjust = 0.5
              ) +
  # geom_nodelab(aes(label=label %>% str_pad(width=14, side="both")), geom = "label", nudge_y = 0.25, nudge_x = -13, size=4) +
  # geom_nodepoint(aes(subset=!is.na(label)), size=2) +
  # xlab("Divergence time (MYA)") +
  geom_facet(
    data = anage.genomeSpecies.plus,
    panel = "Body Size (kg)",
    mapping = aes(x=kgSize, fill=clade),
    geom = ggstance::geom_barh,
    stat = "identity", 
    color = "black"
  ) + 
  geom_facet(
    data = anage.genomeSpecies.plus,
    panel = "Body Size (kg)",
    mapping = aes(x=kgSize %>% multiply_by(0.5) %>% if_else(.<1000, 2000, .),
                  label=kgSize %>% round(digits = 2) %>% format(scientific=F) %>% str_c(" kg")),
    geom = geom_text,
    fontface="bold", 
    family="Helvetica",

    size=2,
    color = "black"
  ) + 
  theme_tree2() +
  scale_color_manual(values = c("Proboscidea" = "#1B9E77", "Sirenia" = "#D95F02", "Hyracoidea" = "#7570B3"), limits = c("Proboscidea", "Sirenia", "Hyracoidea"), 
                     breaks= c("Proboscidea", "Sirenia", "Hyracoidea", "")) +
  scale_fill_manual(values = c("Proboscidea" = "#1B9E77", "Sirenia" = "#D95F02", "Hyracoidea" = "#7570B3"), limits = c("Proboscidea", "Sirenia", "Hyracoidea"), 
                     breaks= c("Proboscidea", "Sirenia", "Hyracoidea", ""), guide=guide_none()) +
  xlim_tree(c(-100, 65)) +
  scale_x_continuous(labels=abs, breaks=scale_breaks_tree)+
  theme(
    legend.position = "bottom",
    text = element_text(face = "plain", family="Helvetica", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text = element_text(size = 8, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 6, 
                              colour = "black", face = "bold"),
    # plot.title = element_text(size = 6, 
    #                           colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 8, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black"),
    strip.text = element_blank(), #element_text(size=16, face="bold"),
    strip.background = element_blank()
  )  + 
  guides(color=guide_legend(title="Clade", title.position = "left")) + 
  scale_y_continuous(limits=c(0, 20),
                        expand=c(0,0),
                        oob=function(x, ...) x) +
    geom_text(aes(label=lab), data=plot.label, fontface="bold", size=3, family="Helvetica") + 
    coord_cartesian(clip='off')  + 
    theme(plot.margin=margin(6, 6, 40, 6), rect=element_blank())

atlantogenata.tree.size.lifespan %<>% facet_widths(c(2,1))

```
```{r RBHB Stats}
n.TranscriptsSearched <- query.binTable %>% filter(!C.ORF, !LOC, !LINC, !HLA, !Histone, !ZincFinger,!Riboprotein, !OlfactoryReceptor, !Uncharacterized, !Putative, !Viral, !Fragment) %>% pull(ID) %>% unique %>% length()
n.GenesSearched <- query.binTable %>% filter(!C.ORF, !LOC, !LINC, !HLA, !Histone, !ZincFinger,!Riboprotein, !OlfactoryReceptor, !Uncharacterized, !Putative, !Viral, !Fragment) %>% pull(Gene) %>% unique %>% length()
genesPerGenome <- geneCopyTable %>% filter(Copy>=1, !is.na(ECNC)) %>% group_by(Genome) %>% summarise(
  Hits=n(), pc.Found=Hits/n.GenesSearched, 
  mean.ECNC = mean(ECNC), q1.ECNC = quantile(ECNC, 1/4), median.ECNC = median(ECNC), q3.ECNC = quantile(ECNC, 3/4), max.ECNC = max(ECNC), sd.ECNC = sd(ECNC), 
  mean.Copy = mean(Copy), q1.Copy = quantile(Copy, 1/4), median.Copy = median(Copy), q3.Copy = quantile(Copy, 3/4), max.Copy = max(Copy), sd.Copy = sd(Copy), 
  mean.ECNCPerHit = mean(ECNC/Copy), q1.ECNCPerHit = quantile(ECNC/Copy, 1/4), median.ECNCPerHit = median(ECNC/Copy), q3.ECNCPerHit = quantile(ECNC/Copy, 3/4), max.ECNCPerHit = max(ECNC/Copy), min.ECNCPerHit = min(ECNC/Copy), sd.ECNCPerHit = sd(ECNC/Copy))
duplicatedGenesPerGenome <- geneCopyTable %>% filter(Copy>=1, ECNC>=1.5) %>% group_by(Genome) %>% summarise(Duplicated=n())

genomeStatsRBHB <- full_join(genesPerGenome, duplicatedGenesPerGenome, by="Genome") %>% mutate(pc.Hits.Duplicated = Duplicated %>% divide_by(Hits))

median.genesPerGenome <- genesPerGenome %>% pull(Hits) %>% median
median.DuplicatedGenesPerGenome <- duplicatedGenesPerGenome %>% pull(Duplicated) %>% median
```
```{r Genome Table Prep}
genome.table.forFig <- genomeTable.atlantogenata %>% 
  left_join(anage, by="label") %>% 
  left_join(tree.Atlantogenata.table, by="label", suffix=c(".anage", ".tree")) %>% 
  mutate(
    Median = exp(Median) %>% formatC(format = "g", digits = 2),
    Lifespan = Lifespan.anage %>% exp %>% formatC(format = "g", digits = 2) %>% as.character() %>% str_replace_na(replacement = "(Unknown)")
  ) %>% 
  select(
    `Common Name` = `Common Name.tree`,
    Species = Species.strict.tree,
    # Clade = Family, 
    `Size (g)` = Median,
    # `Lifespan (y)`= Lifespan, 
    Genome = BestGenome.tree
  ) %>% 
  left_join(
    genomeStatsRBHB %>% 
      select(Genome, `#Hits`=Hits, `% Genes Found`=pc.Found, `#Duplicated` = Duplicated,`% Hits Duplicated` = pc.Hits.Duplicated, `Max Copy#`=max.Copy, `Max ECNC`=max.ECNC, `Mean ECNC/Hit`=mean.ECNCPerHit), 
    by="Genome"
  ) %>% 
  mutate(
    `Common Name` = str_to_title(`Common Name`),
    `% Genes Found` = `% Genes Found` %>% as.percentage,
    `% Hits Duplicated` = `% Hits Duplicated` %>% as.percentage,
    `Mean ECNC/Hit` = round(`Mean ECNC/Hit`, 2)
  )

```
```{r ORA-ranking}

ORA.cancer.pc.simulated <- stats.empirical.ds[, class:=class.pathway(description)][pValue.mean <= 0.05, .(N.cancer = sum(class != "Other"), N.Path = .N), by="N.set"][,pc := N.cancer/N.Path][]

ORA.ranking <- merge(
  stats.empirical.ds[!is.na(pValue.mean)], 
  ORA.term.stats, #[, .(perRun.mean, perRun.sd, pval.mean, pval.sd), by="N.set"], 
  by="N.set"
  )[,
    .(Z.freq=(PerRun.freq.mean-perRun.mean)/perRun.sd,
      Z.pval=(pValue.mean-pval.mean)/pval.sd),
    by=c("description", "N.set")
    ][]

# A Good summary table:
# merge(
#     ORA.all, 
#     ml.tree.increasing.geneCounts %>% as_tibble() %>% select(label, N.Genes), 
#     by.x="node", by.y="label", all.y=T
# )[,.(N.pathways = .N-sum(is.na(description))), by=c("node", "N.Genes")] %>% unique

ORA.all <- merge(  # Note this is a right join, so none of the (any/not/only)Proboscidea sets, but all the nodes are represented
    ORA.all, 
    ml.tree.increasing.geneCounts %>% as_tibble() %>% select(label, N.Genes), 
    by.x="node", by.y="label", all.y=T
  )

ORA.all.ranked <- merge(
  ORA.all,
  ORA.ranking[,.(N.Genes=N.set, description, Z.pval, Z.freq)], 
  by=c("description", "N.Genes")
)[,
  `:=`(pValue = as.numeric(pValue),
       nlPval = -log(pValue),
       FDR = as.numeric(FDR),
       FDR.sig.10 = FDR<=0.1,
       p.z.freq = pnorm(-abs(Z.freq)/2), 
       CombinedScore.freq=-log(pValue)*Z.freq)
][]

ORA.pcCancer.withSim <- merge(ORA.all[pValue <= 0.05, .(N.cancer = sum(Class != "Other"), N.Path = .N), by=c("node", "N.Genes")][,pc := N.cancer/N.Path, by=c("node", "N.Genes")][], ORA.cancer.pc.simulated[,N.Genes:=N.set][,.(pc.sim=pc, N.Genes)],all.x = T, all.y = F, by="N.Genes")

ORA.pcCancer.withSim[is.na(pc.sim)]$pc.sim <- stats.empirical.ds[, class:=class.pathway(description)][, .(N.cancer = sum(class != "Other"), N.Path = .N), by="N.set"][,pc := N.cancer/N.Path][,median(pc)]

ORA.pcCancer.aboveSim <- ORA.pcCancer.withSim[pc>pc.sim]

```
```{r LoxAfr Genes to Highlight}
genesToHighlight.plus <- list(
  "CellCycle" = c(
    "RBX1", #Cell cycle, ubiqu., Probo
    "LIN9", # Cell cycle, Probo
    "MND1", # Cell Cycle, Elephantidea,
    "E2F2", #Cell cycle, Tethytheria
    "MAX", # Cell Cycle, Tethytheria
    "SYCP3" #Cell cycle, unique to LoxAfr
  ),
  "APC" = c(
    "CDK1", # APC and Cell Cycle, Atlantogenata
    "RNF168", # ATM, cell cycle, Paen or Pseudo
    "RBBP8", # ATM Signaling, Cell Cycle, DSB repair, Tethy
    "MAD2L1", # APC, Tethytheria
    "CSNK1A1", #APC? Tethytheria
    "RPA2", # ATR, cell cycle, and stress response, Probo
    "PSMB3", # APC/C, ubiq, Probo
    "BUB3" # APC, 2 copies until 3 in LoxAfr
  ),
  "TP53" = c(
    "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata
    "STK11", #REgulation of TP53, Paenungulata
    "COX5A", #TP53 transcriptional regulation, PRobodiscea
    "BRD7", # TP53 regulation, Tethytheria
    "TP53", #TP53 duplicates in Tethytheria
    "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria
    "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants
    "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri
    "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants
    "COX20" # TP53 Regulates Metabolic Genes, LoxAfr
  ),
  "Stress" = c(
    "SOD1",  # Oxidative Stress, Probodiscea
    "SCMH1", # DNA damage response and OS-induced senescence, Probo
    "RNF2", # OSISenecense, Tethytheria
    "ORC3", # Replication Stress, Loxondontini, Elephas
    "POLR2K", # TC-NER, Elephantidae
    "HSBP1", # Heat Shock, Elephantidae
    "ERCC8", #Ubiquitin mediated proteolysis, Elephantidea
    "DTX3L", # Ubiquitination & Proteasome degradation, Loxodontini
    "DNAJC2", # Heat Shock, Probo
    "DET1" # Ubiq-mediate proteolysis, Probo
  ),
  "Death" = c(
    "RACK1", # Death Receptor Signaling, Tethy or Paen
    "PCBP2", # Ferroptosis, Loxodontini, Trichechus
    "NSMAF", # Death Receptor Signaling, Tethytheria
    "PCBP1", # Ferroptosis, Probodiscea
    "SERBP1", # Apoptosis, Probodiscea
    "CASP12", # Apoptosis, Probodiscea
    "NET1", # Apoptosis in cancers, Elephantidae
    "CSF2RB" #Apoptosis, Elephantidae
  ),
  "RAS" = c(
    "RASA1", # RAS signaling Pathway, Paen
    "RAB5B", # RAS signaling, Paen
    "NRAS", #Literally just RAS, Tethytheria
    "GSK3A" #RAS, Tethy
  ),
  "CancerRelatedGenes" = c(
    "MAPK3", # Many cancers, Afrotheria
    "GABARAP", #Senescence and autophagy in cancer, Paen or Tethy
    "HNRNPK", #LncRNA involvement in Wnt and Cancer, Probodiscea
    "CTTN", # Proteoglycans in Cancer, Loxodontini
    "CACYBP", # Gastric Cancer Network 2, Probodiscea
    "CHPT1", # Choline Metabolism in Cancer, Probodiscea
    "PYCR2", # Metabolic reprogramming in colon cancer, Elephantidae
    "PLPP2", # Choline Metabolism in Cancer, Loxodontini, Elephas, Trichechus
    "PLAC8" # Gastric Cancer Network 2, continuously increases towards Loxodona
  )
)

ml.treedata.loxAfrHighlight.plus <- genesToHighlight.plus %>% 
  unlist %>% 
  set_names(.,.) %>% 
  lapply(., function(gene){
    ml.treedata.table %>% 
      select(parent, node, branch.length, label, all_of(gene)) %>% 
      rename("Count"=all_of(gene))%>% 
      as.treedata()
    })

loxAfr4.geneTPM.genesToHighlight.plus <- merge(
  loxAfr4.geneTPM,
  geneCopyTable %>% 
    filter(str_detect(Genome, "loxAfr|triMan")) %>% 
    mutate(ECNC = str_c("(", ECNC, ")")) %>% 
    unite("CopyECNC", Copy, ECNC, sep=" ") %>% 
    spread(key="Genome", value=c("CopyECNC")),
  by="Gene"
)
```
```{r ECNC DIAGRAMS, include=F}
ecnc.df1 <- tibble(
  hit = as.character(1) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0),
  x2=c(100),
  y1=c(0),
  y2 = c(1)
)
ecnc.df2 <- tibble(
  hit = as.character(1:2) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0, 0),
  x2=c(100, 100),
  y1=c(0, 1),
  y2 = c(1, 2)
)
ecnc.df3 <- tibble(
  hit = as.character(1:3) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0,25,75),
  x2=c(25,75,100),
  y1=c(0, 0, 0),
  y2 = c(1, 1, 1)
)
ecnc.df4 <- tibble(
  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0,25,75, 0),
  x2=c(25,75,100, 100),
  y1=c(0, 0, 0, 1),
  y2 = c(1, 1, 1, 2)
)

ecnc.df5 <- tibble(
  hit = as.character(1:4) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  x1=c(0,0,0, 0),
  x2=c(100,15,15, 15),
  y1=c(0, 1, 2, 3),
  y2 = c(1, 2, 3, 4)
)

ecnc.df6 <- tibble(
  hit = as.character(c(1,2,1,3, 3)) %>% paste("Hit", ., sep=" ") %>% as.factor(),
  label = factor(c("Hit 1", "Hit 2", "Unidentified", "Unidentified", "Hit 3")),
  x1=c(0, 0, 75, 75, 50),
  x2=c(75, 50, 100, 100, 75),
  y1=c(0, 1, 0, 1, 1),
  y2 = c(1, 2, 1, 2, 2),
  transparent = c(0,0,1,1,0) %>% as.factor
)

ecnc.onegene.onehit <- 
  ecnc.df1 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.twogene.twohit <- 
  ecnc.df2 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.onegene.threehits <-
  ecnc.df3 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.twogene.fourhits <-
  ecnc.df4 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.onegene.manypartialhits <-
  ecnc.df5 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1.45, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.twogene.partialhits <-
  ecnc.df6 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit,
      alpha = transparent,
      linetype = transparent
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=label), color="black", inherit.aes = F)+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_alpha_manual(values = c("0"=1,"1"=0.5), guide=F) + 
  scale_linetype_manual(values = c("0"="solid","1"="dotted"), guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

```
```{r Correlations between genome quality scores}
CEGMA.completeness.clean <-
  CEGMA.completeness %>%
  filter(str_detect(Group, "_all")) %>%
  select(-`Missing Proteins`) %>%
  rename(
    n.Prots = `#Prots`,
    pc.Completeness = `%Completeness`,
    n.Total = `#Total`,
    pc.Ortho = `%Ortho`
  ) %>%
  reshape2::melt(., id = c("Genome", "Group")) %>%
  mutate(Group = Group %>% str_remove("_all")) %>%
  unite("variable", variable, Group) %>%
  pivot_wider(id_cols = Genome, names_from = variable, values_from = value)

CEGMA <- full_join(CEGMA.qc, CEGMA.completeness.clean, by="Genome")

GCT.CEGMA <- genomeStatsRBHB %>%
  select(-starts_with("q"), -starts_with("Median")) %>% 
  inner_join(., CEGMA, by="Genome") %>%
  mutate_if(is.character, as.factor)

cor.GCT.CEGMA <- GCT.CEGMA %>%
  mutate_if(is.factor, as.numeric) %>%
  as.matrix %>%
  cor %>%
  reorder_cormat %>%
  get_lower_tri %>%
  reshape2::melt(., na.rm=T)

cor.toHighlight <- c("mean.ECNC", "mean.Copy", "L50 sequence count", "# of sequences >  10M (nt)", "# of sequences >   1M (nt)", "# of sequences >  10K (nt)", "# of sequences >   1K (nt)", "# of sequences > 100K (nt)", "# of sequences", "Sum length of sequences > 1M  (nt)", "Total length (nt)", "Base composition (%) (N)", "mean.ECNCPerHit", "n.Prots_Partial","pc.Completeness_Partial", "Median sequence length (nt)", "Mean sequence length (nt)", "Sum length of sequences > 10M  (nt)")

```
```{r Phylogenetic Correlations}
to_drop <-tree.Atlantogenata.table %>% filter(!BestGenome %in% (GCT.CEGMA %>% pull(Genome))) %>% pull(label)

tree.GCT.CEGMA <- full_join(
  tree.Atlantogenata %>% 
    drop.tip(to_drop) %>% 
    as_tibble %>% 
    mutate(Genome = BestGenome),
  GCT.CEGMA, 
  by="Genome"
) %>% 
  as.treedata()
GCT.CEGMA.wGenome <- tree.GCT.CEGMA %>% as_tibble

```

# Introduction

Among the constraints on the evolution of large bodies and long life-spans in animals is an increased risk of developing cancer. If all cells in all organisms have a similar risk of malignant transformation and equivalent cancer suppression mechanisms, then organisms with many cells should have a higher prevalence of cancer than organisms with fewer cells, particularly because large and small animals have similar cell sizes [@Savage2007]. Consistent with this expectation there is a strong positive correlation between body size and cancer incidence within species, for example, cancer incidence increases with increasing adult height in humans [@Green2011; @Nunney:20181c2] and overall body size in dogs, cats, and cattle [@Dobson2013; @Dorn1968; @Lucena2011]. There is no correlation, however, between body size and cancer risk between species; this lack of correlation is often referred to as ‘Peto’s Paradox’ [@Caulin2011; @Leroi2003; @Peto1975]. Indeed, cancer prevalence is relatively stable at ~5% across species with diverse body sizes ranging from the minuscule 51g grass mouse to the gargantuan 4800kg African elephant [@Abegglen:JAMA2015; @Boddy2020; @Tollis2020]. While the ultimate resolution to Peto’s Paradox is obvious, large bodied and long-lived species evolved enhanced cancer protection mechanisms, identifying and characterizing the proximate genetic, molecular, and cellular mechanisms that underlie the evolution of augmented cancer protection has proven difficult [@Ashur-Fabian2004; @Seluanov2008; @Gorbunova2012; @Tian2013; @Sulak2015].

One of the challenges for discovering how animals evolved enhanced cancer protection mechanisms is identifying lineages in which large bodied species are nested within species with small body sizes. *Afrotherian* mammals are generally small-bodied, but also include the largest extant land mammals. For example, maximum adult weights are ~70g in golden moles, ~120g in tenrecs, ~170g in elephant shrews, ~3kg in hyraxes, and ~60kg in aardvarks [@HAGR]. In contrast, while extant hyraxes are relatively small, the extinct *Titanohyrax* is estimated to have weighed ~1300kg [@Schwartz1995]. The largest members of *Afrotheria*, too, are dwarfed by the size of their recent ancestors: extant sea cows such as manatees are large bodied (~322-480kg) but are relatively small compared to the extinct Stellar’s sea cow which is estimated to have weight ~8000-10000kg [@Scheffer1972]. Similarly African (4,800kg) and Asian elephants (3,200kg) are the large, but are dwarfed by the truly gigantic extinct Proboscideans such as *Deinotherium* (~132,000kg), *Mammut borsoni* (110,000kg), and the straight-tusked elephant (~220,000kg), the largest known land mammal [@Larramendi:20151c2]. Remarkably, these large-bodied *Afrotherian* lineages are nested deeply within small bodied species (**Figure \@ref(fig:Figure-intro)**) [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015], indicating that gigantism independently evolved in hyraxes, sea cows, and elephants (*Paenungulata*). Thus, *Paenungulates* are an excellent model system in which to explore the mechanisms that underlie the evolution of large body sizes and augmented cancer resistance.

Many mechanisms have been suggested to resolve Peto’s paradox, including a decrease in the copy number of oncogenes, an increase in the copy number of tumor suppressor genes [@Caulin2011; @Leroi2003; @Nunney1999], reduced metabolic rates, reduced retroviral activity and load [@Katzourakis2014], and selection for 'cheater' tumors that parasitize the growth of other tumors [@Nagy2007], among many others. Among the most parsimonious routes to enhanced cancer resistance may be through an increased copy number of tumor suppressors. For example, transgenic mice with additional copies of TP53 have reduced cancer rates and extended lifespans but not increased cancer [@GarciaCao:20021c2], respectively, suggesting that changes in the copy number of tumor suppressors can effect cancer rates. Indeed, candidate genes studies have found that the elephant genome encodes duplicate tumor suppressors such as TP53 and LIF [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018] as well as other genes with putative tumor suppressive functions [@Caulin2015; @Doherty2016]. These studies, however, focused on a priori candidate genes, thus it is unclear whether duplication of tumor suppressor genes is a general phenomenon in the elephant lineage.

Here we trace the evolution of body mass, cancer risk, and gene copy number variation across Afrotherian genomes, including multiple living and extinct Proboscideans, to investigate whether duplication of tumor suppressors occurred coincident with the evolution of large body sizes. Our estimates of the evolution of body mass across Afrotheria show that large body masses evolved in a step-wise manner, similar to previous studies [@OLeary2013a; @Springer2013; @OLeary2013b; @PuttickAndThomas2015] and coincident with dramatic reductions in intrinsic cancer risk. To explore whether duplication of tumor suppressors occurred coincident with the evolution of large body sizes, we used a genome-wide Reciprocal Best BLAT Hit (RBBH) strategy to identify gene duplications, and used maximum likelihood to infer the lineages in which those duplications occurred. Unexpectedly, we found that duplication of tumor suppressor genes was common in Afrotherians, both large and small, but that gene duplications in the Proboscidean lineage were enriched in pathways that may explain some of the unique cancer protection mechanisms observed iin elephant cells. These data suggest that duplication of tumor suppressor genes is pervasive in Afrotherians and proceeded the evolution of species with very large body sizes.

# Methods

## Ancestral Body Size Reconstruction

We first built a time-calibrated supertree of Eutherian mammals by combining the time-calibrated molecular phylogeny of Bininda-Emonds et al. [@Bininda-Emonds2008] with the time-calibrated total evidence Afrotherian phylogeny from Puttick and Thomas [@PuttickAndThomas2015]. While the Bininda-Emonds et al. [@Bininda-Emonds2008] phylogeny includes 1,679 species, only 34 are Afrotherian, and no fossil data are included. The inclusion of fossil data from extinct species is essential to ensure that ancestral state reconstructions of body mass are not biased by only including extant species. This can lead to inaccurate reconstructions, for example, if lineages convergently evolved large body masses from a small bodied ancestor. In contrast, the total evidence Afrotherian phylogeny of Puttick and Thomas [@PuttickAndThomas2015] includes 77 extant species and fossil data from 39 extinct species. Therefore we replaced the Afrotherian clade in the Bininda-Emonds et al. [@Bininda-Emonds2008] phylogeny with the Afrotherian phylogeny of Puttick and Thomas [@PuttickAndThomas2015] using Mesquite. Next, we jointly estimated rates of body mass evolution and reconstructed ancestral states using a generalization of the Brownian motion model that relaxes assumptions of neutrality and gradualism by considering increments to evolving characters to be drawn from a heavy-tailed stable distribution (the “Stable Model”) implemented in StableTraits [@ElliotAndMooers2014]. The stable model allows for large jumps in traits and has previously been shown to out-perform other models of body mass evolution, including standard Brownian motion models, Ornstein–Uhlenbeck models, early burst maximum likelihood models, and heterogeneous multi-rate models [@ElliotAndMooers2014].

## Identification of Duplicate Genes

_Reciprocal Best-Hit BLAT:_ We developed a reciprocal best hit BLAT (RBHB) pipeline to identify putative homologs and estimate gene copy number across species. The Reciprocal Best Hit (RBH) search strategy is conceptually straightforward: 1) Given a gene of interest $G_A$ in a query genome $A$, one searches a target genome $B$ for all possible matches to $G_A$; 2) For each of these hits, one then performs the reciprocal search in the original query genome to identify the highest-scoring hit; 3) A hit in genome $B$ is defined as a homolog of gene $G_A$ if and only if the original gene $G_A$ is the top reciprocal search hit in genome $A$. We selected BLAT [@blat] as our algorithm of choice, as this algorithm is sensitive to highly simliar (>90% identity) sequences, thus identifying the highest-confidence homologs while minimizing many-to-one mapping problems when searching for multiple genes. RBH performs similar to other more complex methods of orthology prediction, and is particularly good at identifying incomplete genes that may be fragmented in low quality/poor assembled regions of the genome [@AltenhoffAndDessimoz2009; @SalichosAndRokas2011].

_Effective Copy Number By Coverage:_ In low-quality genomes, many genes are fragmented across multiple scaffolds, which results in BLA(S)T-like methods calling multiple hits when in reality there is only one gene. To compensate for this, developed a novel statistic, Estimated Copy Number by Coverage (ECNC), which averages the number of times we hit each nucleotide of a query sequence in a target genome over the total number of nucleotides of the query sequence found overall in each target genome (**Supplemental Figure \@ref(fig:Figure-ECNC)**). This allows us to correct for genes that have been fragmented across incomplete genomes, while also taking into account missing sequences from the human query in the target genome. Mathematically, this can be written as:

$$ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}(1)$$

where $n$ is a given nucleotide in the query, $l$ is the total length of the query, $C_n$ is the number of instances that $n$ is present within a reciprocal best hit, and $bool(C_n)$ is 1 if $C_n > 0$ or 0 if $C_n = 0$.

_RecSearch Pipeline:_ We created a custom Python pipeline for automating RBHB searches between a single reference genome and multiple target genomes using a list of query sequences from the reference genome. For the query sequences in our search, we used the hg38 UniProt proteome [Accession #UP000005640; @uniprot], which is a comprehensive set of protein sequences curated from a combination of predicted and validated protein sequences generated by the UniProt Consortium. In order to refine our search, we omitted protein sequences originating from long, noncoding RNA loci (e.g. LINC genes); poorly-studied genes from predicted open reading frames (C-ORFs); and sequences with highly repetitive sequences such as zinc fingers, protocadherins, and transposon-containing genes, as these were prone to high levels of false positive hits.
After filtering out problematic protein queries, we then used our pipeline to search for all copies of 18011 query genes in publicly available *Afrotherian* genomes (\@ref(tab:Table-Genomes)), including African savannah elephant (*Loxodonta africana*: loxAfr3, loxAfr4, loxAfrC), African forest elephant (*Loxodonta cyclotis*: loxCycF), Asian Elephant (*Elephas maximus*: eleMaxD), Woolly Mammoth (*Mammuthus primigenius*: mamPriV), Colombian mammoth (*Mammuthus columbi*: mamColU), American mastodon (*Mammut americanum*: mamAmeI), Rock Hyrax (*Procavia capensis*: proCap1, proCap2, proCap2*HiC), West Indian Manatee (*Trichechus manatus latirostris*: triManLat1, triManLat1*HiC), Aardvark (*Orycteropus afer*: oryAfe1, oryAfe1*HiC), Lesser Hedgehog Tenrec (*Echinops telfairi*: echTel2), Nine-banded armadillo (*Dasypus novemcinctus*: dasNov3), Hoffman's two-toed sloth (*Choloepus hoffmannii*: choHof1, choHof2, choHof2*HiC), Cape golden mole (*Chrysochloris asiatica*: chrAsi1), and Cape elephant shrew (*Elephantulus edwardii*: eleEdw1).

_Query gene inclusion criteria:_ To assemble our query list, we first removed all unnamed genes from UP000005640. Next, we excluded genes from downstream analyses for which assignment of homology was uncertain, including uncharacterized ORFs (991 genes), LOC (63 genes), HLA genes (402 genes), replication dependent histones (72 genes), odorant receptors (499 genes), ribosomal proteins (410 genes), zinc finger transcription factors (1983 genes), viral and repetitive-element-associated proteins (82 genes) and any protein described as either “Uncharacterized,” “Putative,” or “Fragment” by UniProt in UP000005640 (30724 genes), leaving us with a final set of 37,582 query protein isoforms, corresponding to 18,011 genes.

_Duplication gene inclusion criteria:_ In order to condense transcript-level hits into single gene loci, and to resolve many-to-one genome mappings, we removed exons where transcripts from different genes overlapped, and merged overlapping transcripts of the same gene into a single gene locus call. The resulting gene-level copy number table was then combined with the maximum ECNC values observed for each gene in order to call gene duplications. We called a gene duplicated if its copy number was two or more, and if the maximum ECNC value of all the gene transcripts searched was 1.5 or greater; previous studies have shown that incomplete duplications can encode functional genes [@Sulak2015; @Vazquez2018], therefore partial gene duplications were included provided they passed additional inclusion criteria (see below). The ECNC cut-off of 1.5 was selected empirically, as this value minimized the number of false positives seen in a test set of genes and genomes. The results of our initial search are summarized in **Figure \@ref(fig: Figure-ML-UpSet-WC)A**. Overall, we identified 13880 genes across all species, or 77.1% of our starting query genes.

_Genome Quality Assessment using CEGMA:_ In order to determine the effect of genome quality on our results, we used the gVolante webserver and CEGMA to assess the quality and completeness of the genome [@gVolante; @CEGMA]. CEGMA was run using the default settings for mammals (“Cut-off length for sequence statistics and composition” = 1;“CEGMA max intron length” = 100000; “CEGMA gene flanks” = 10000, “Selected reference gene set” = CVG). For each genome, we generated a correlation matrix using the aforementioned genome quality scores, and either the mean Copy Number or mean ECNC for all hits in the genome. We observed that the percentage of duplicated genes in non-_Pseudoungulatan_ genomes was higher (12.94% to 23.66%) than *Pseudoungulatan* genomes (3.26% to 7.80%). Mean Copy Number, mean ECNC, and mean CN (the lesser of Copy Number and ECNC per gene) moderately or strongly correlated with genomic quality, such as LD50, the number of scaffolds, and contigs with a length above either 100K or 1M (**Supplemental Figure \@ref(fig:Figure-GenomeCor)**). The Afrosoricidians had the greatest correlation between poor genome quality and high gene duplication rates including larger numbers of private duplications. The correlations between genome quality metric and number of gene duplications was particularly high for Cape golden mole (*Chrysochloris asiatica*: chrAsi1) and Cape elephant shrew (*Elephantulus edwardii*: eleEdw1), therefore we excluded these species from downstream pathway enrichment analyses.

## Evidence for Functionality of Gene Duplicates

To validate and filter out duplicate gene calls, we intersected our results with either gene prediction or transcriptomic evidence as a proxy for functionality.

_Transcriptome Assembly_: For the African Savana Elephant, Asian Elephant, West Indian Manatee, and Nine-Banded Armadillo, we generated de novo transcriptomes using publically-available RNA-sequencing data from NCBI SRA (Supplemental Table \@ref(tab:Table-SRAData)). We mapped reads to all genomes available for each species, and assembled transcripts using HISAT2 and StringTie, respectively [@HISAT; @StringTie; @Tuxedo]. RNA-sequencing data was not available for Cape Golden Mole, Cape Elephant Shrew, Rock Hyrax, Aardvark, or the Lesser Hedgehog Tenrec.

_Gene Prediction_: We downloaded tracks for genes predicted using GenScan for all the genomes available via UCSC Genome Browser: African savannah elephant (loxAfr3), Rock Hyrax (proCap1), West Indian Manatee (triManLat1), Aardvark (oryAfe1), Lesser Hedgehog Tenrec (echTel2), Nine-banded armadillo (dasNov3), Hoffman’s Two-Toed Sloth (choHof1), Cape golden mole (chrAsi1), and Cape Elephant Shrew (eleEdw1); gene prediction tracks for higher-quality assemblies were not available.

_Evidenced Duplicate Criteria_: We intersected our records of duplicate hits identified in each genome with the gene prediction tracks and/or transcriptome assemblies using bedtools. When multiple lines of evidence for functionality were present for a genome, we used the union of all intersections as the final output for evidenced duplicates. When analyzing the highest-quality assemblies available for each species, if a species had neither gene prediction tracks nor RNA-seq data for the highest-quality genome available, we conservatively included all hits for the genome in the final set of evidenced duplicates.

## Reconstruction of Ancestral Copy Numbers

We encoded the copy number of each gene for each species as a discrete trait ranging from 0 (one gene copy) to 31 (for 32+ gene copies) and used IQ-TREE to select the best-fitting model of character evolution [@IQTREE; @UFBoot2; @ModelFinder; @PMSF; @Schrempf2019], which was inferred to be a Jukes-Cantor type model for morphological data (MK) with equal character state frequencies (FQ) and rate heterogeneity across sites approximated by including a class of invariable sites (I) plus a discrete Gamma model with four rate categories (G4). Next we inferred gene duplication and loss events with the empirical Bayesian ancestral state reconstruction (ASR) method implemented in IQ-TREE [@IQTREE; @UFBoot2; @ModelFinder; @PMSF; @Schrempf2019], the best fitting model of character evolution (MK+FQ+GR+I) [@Soubrier2012; @Yang1995], and the unrooted species tree for *Atlantogenata*. We considered ancestral state reconstructions to be reliable if they had Bayeisan Posterior Probability (BPP) $\geq$ 0.80; less reliable reconstructions were excluded from pathway analyses.

## Pathway Enrichment Analysis

To determine if gene duplications were enriched in particular biological pathways, we used the WEB-based Gene SeT AnaLysis Toolkit (WebGestalt)[@WebGestalt2019] to perform Over-Representation Analysis (ORA) using the Reactome database [@Reactome]. Gene duplicates in each lineage were used as the foreground gene set, and the initial query set was used as the background gene set. WebGestalt uses a hypergeometric test for statistical significance of pathway over-representation, which we refined using two methods: an False Discovery Rate (FDR)-based approach, and an empirical p-value approach [@Chen2013]. The Benjamini-Hochberg FDR multiple-testing correction was generated by WebGestalt. In order to correct P-values based on an empirical distribution, we modified the approach used by Chen *et al.* in Enrichr [@Chen2013] to generate a “combined score” for each pathway based on the hypergeometric P-value from WebGestalt, and a correction for expected rank for each pathway. In order to generate the table of expected ranks and variances for this approach, we randomly sampled foreground sets of 10-5,000 genes from our background set 5000 times, and used WebGestalt ORA to obtain a list of enriched terms and P-values for each run; we then compiled a table of Reactome terms with their expected frequencies and standard deviation. This data was used to calculate a Z-score for terms in an ORA run, and the combined score was calculated using the formula $C = \log{(p)}\cdot{}z$.

## Estimating the Evolution of Cancer Risk

The dramatic increase in body mass and lifespan in some Afrotherian lineages, and the relatively constant rate of cancer across species of diverse body sizes [@Abegglen:JAMA2015], indicates that those lineages must have also evolved reduced cancer risk. To infer the magnitude of these reductions we estimated differences in intrinsic cancer risk across extant and ancestral Afrotherians. Following Peto [@Peto:20151c2], we estimate the intrinsic cancer risk ($K$) as the product of risk associated with body mass and lifespan. In order to determine ($K$) across species and at ancestral nodes (see below), we first estimated ancestral lifespans at each node. We used Phylogenetic Generalized Least-Square Regression (PGLS) [@Felsenstein1985; @MartinsAndHansen1997], using a Brownian covariance matrix as implemented in the R package `ape` [version `r packageVersion("ape")`; @ape2019],  to calculate estimated ancestral lifespans across *Atlantogenata* using our estimates for body size at each node. In order to estimate the intrinsic cancer risk of a species, we first inferred lifespans at ancestral nodes using PGLS and the model $ln(lifespan) = \beta_{1}corBrowninan + \beta_{2}ln(size) + \epsilon$. Next, we calculated $K_{1}$ at all nodes, and then estimated the fold-change in cancer susceptibility between ancestral and descendant nodes (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan**). Next, in order to calculate $K_{1}$ at all nodes, we used a simplified multistage cancer risk model for body size $D$ and lifespan $t$: $K \approx Dt^6$ [@Armitage:19851c2; @Armitage:20041c2; @Peto:20151c2; @Peto1975]. The fold change in cancer risk between a node and its ancestor was then defined as $\log_{2}(\frac{K_{2}}{K_{1}})$.

## Data Analysis

All data analysis was performed using `r version$version.string`, and the complete reproducible manuscript along with code and data generation pipeline can be found on the author's GitHub page at [www.github.com/docmanny/smRecSearch/tree/publication](www.github.com/docmanny/smRecSearch/tree/publication) [@R-ape; @R-base; @R-bookdown; @R-broom.mixed; @R-data.table; @R-dplyr; @R-forcats; @R-geiger; @R-gghighlight; @R-ggimage; @R-ggnewscale; @R-ggplot2; @R-ggplotify; @R-ggpubr; @R-ggrepel; @R-ggsci; @R-ggtree; @R-kableExtra; @R-magick; @R-magrittr; @R-nlme; @R-plotly; @R-purrr; @R-readr; @R-stargazer; @R-stringr; @R-tibble; @R-tidyr; @R-tidytree; @R-tidyverse; @R-treeio; @R-UpSetR; @ape2019; @bookdown2016; @geiger2009; @geiger2011; @geiger2012; @geiger2008; @geiger2014; @ggplot22016; @ggtree2020; @ggtree2018; @ggtree2017; @plotly2020; @tidyverse2019; @treeio2020]

# Results

## Step-wise evolution of body size in Afrotherians

Similar to previous studies of Afrotherian body size [@ElliotAndMooers2014; @PuttickAndThomas2015], we found that the body mass of the Afrotherian ancestor was inferred to be small (0.26kg, 95% CI: 0.31-3.01kg) and that substantial accelerations in the rate of body mass evolution occurred coincident with a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Orycteropus.afer")` in body mass in the stem-lineage of *Pseudoungulata* (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Orycteropus.afer")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Procavia.capensis")` in body mass in the stem-lineage of *Paenungulata* (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Procavia.capensis")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Trichechus.manatus")` in body mass in the stem-lineage of *Tehthytheria* (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Trichechus.manatus")`), a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Mammut.americanum")` in body mass in the stem-lineage of *Elephantimorpha* (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Mammut.americanum")`), and  a `r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Numidotherium.koholense")` in body mass in the stem-lineage of *Proboscidea* (`r text.bodysize.ancNode.byMRCA("Loxodonta.africana", "Numidotherium.koholense")`) (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan)A,B**). The ancestral *Hyracoidea* was inferred to be relatively small (`r text.bodysize.ancNode.byMRCA.CI("Microhyrax.lavocati", "Procavia.capensis")`), and rate accelerations were coincident with independent body mass increases in large hyraxes such as *Titanohyrax andrewsi* (`r text.bodysize.byLabel("Titanohyrax.andrewsi")`, `r text.ratio.nodeByMRCA.ancNode(nodeA="Titanohyrax.andrewsi","Orycteropus.afer")`). While the body mass of the ancestral *Sirenian* was inferred to be large (`r text.bodysize.ancNode.byMRCA.CI("Trichechus.manatus", "Prorastomus.sirenoides")`), a rate acceleration occurred coincident with a `r text.ratio.node.ancnode.atLabel("Hydrodamalis.gigas")` in body mass in Stellar's sea cow. Rate accelerations also occurred coincident with `r text.ratio.nodeByMRCA.ancNode("Elephas.antiquus.falconeri", "Elephas.cypriotes")` in body mass in the stem-lineage of the dwarf elephants *Elephas (Palaeoloxodon) antiquus falconeri* and *Elephas cypriotes*. These data suggest that gigantism in *Afrotherians* evolved step-wise, from small to medium bodies in the *Pseudoungulata* stem-lineage, medium to large bodies in the *Tehthytherian* stem-lineage and extinct hyraxes, and from large to exceptionally large bodies independently in the *Proboscidean* stem-lineage and Stellar's sea cow (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan)A,B**).

## Step-wise reduction of intrinsic cancer risk in large, long-lived Afrotherians

In order to account for a relatively stable cancer rate across species [@Abegglen:JAMA2015; @Boddy2020; @Tollis2020], intrinsic cancer risk must also evolve with changes body size and lifespan across species. As expected, intrinsic cancer susceptibility in *Afrotheria* also varies with changes in body size and longevity (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan)A**), with initial `r tree.Atlantogenata.table %>% filter(label == "Afrotheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold and `r tree.Atlantogenata.table %>% filter(label == "Xenarthra") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decreases in the stem-lineage of Afrotheria and Xenarthra, respectively, followed by a `r tree.Atlantogenata.table %>% filter(label == "Pseudoungulata") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decrease in Pseudoungulata, and a `r tree.Atlantogenata.table %>% filter(label == "Orycteropus.afer") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decrease in Aardvarks (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan)A**). In contrast to the Paenungulate stem-lineage, there is a `r tree.Atlantogenata.table %>% filter(label == "Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decrease in cancer risk in Tethytheria, a `r tree.Atlantogenata.table %>% filter(label == "Trichechus.manatus") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decrease in Manatee, and dramatic increases within Proboscidea including a `r tree.Atlantogenata.table %>% filter(label == "Elephantidae") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decrease in Elephantidae and a `r tree.Atlantogenata.table %>% filter(label == "Mammut.americanum") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decrease in the American Mastodon. Within the Elephantidae, Elephantina and Loxodontini have a `r tree.Atlantogenata.table %>% filter(label == "Elephantina") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold decrease in cancer susceptibility, while susceptibility is relatively stable in Mammoths. The three extant Proboscideans, Asian Elephant, African Savana Elephant, and the African Forest Elephant, meanwhile, have similar decreases in body size, with increases in cancer susceptibility (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan)A**).

## Pervasive duplication of tumor suppressor genes in Afrotheria

Our hypothesis was that genes which duplicated coincident with the evolution of increased body sizes (IBM) and reduced intrinsic cancer risk (RICR) would be uniquely enriched in tumor suppressor pathways compared to genes that duplicated in other lineages. Therefore, we tested if duplicated genes in each lineage were enriched in Reactome pathways using ORA implemented in WEBGESTALT. No pathways obviously related to cancer biology or tumor suppression were enriched in either the stem-lineage of *Pseudoungulata* (`r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Orycteropus.afer") %>% str_remove("x increase")`-fold IBM, `r tree.Atlantogenata.table %>% filter(label == "Pseudoungulata") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold RICR) or *Paenungulata* (`r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Procavia.capensis") %>% str_remove("x increase")`-fold IBM, `r tree.Atlantogenata.table %>% filter(label == "Paenungulata") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-log_{2}-fold RICR) (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan)A**). Consistent with our hypothesis, 55.8% (29/52) of the pathways that were enriched in the Tethytherian stem-lineage (`r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Trichechus.manatus") %>% str_remove("x increase")`-fold IBM, `r tree.Atlantogenata.table %>% filter(label == "Tethytheria") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold RICR), 27.8% (20/72) of the pathways that were enriched in the Proboscidean stem-lineage (`r text.ratio.nodeByMRCA.ancNode("Loxodonta.africana", "Palaeoloxodon.antiquus") %>% str_remove("x increase")`-fold IBM, `r tree.Atlantogenata.table %>% filter(label == "Proboscidea") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold RICR), and 28% (33/118) of the pathways that were enriched within the Proboscidean clade were related to tumor suppression (**Figure \@ref(fig:Figure-DoubleTree-Bodysize-Lifespan)C**).  Similarly, 17.8% (10/56) and 30% (30/100) of the pathways that were enriched in manatee (`r text.ratio.nodeByMRCA.ancNode("Trichechus.manatus", "Prorastomus.sirenoides") %>% str_remove("x decrease")`-fold IBM, `r tree.Atlantogenata.table %>% filter(node == MRCA(tree.Atlantogenata.table, "Trichechus.manatus", "Prorastomus.sirenoides")) %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold RICR) and aardvark (`r text.ratio.nodeByMRCA.ancNode("Titanohyrax.andrewsi","Orycteropus.afer") %>% str_remove("x increase")`-fold IBM, `r tree.Atlantogenata.table %>% filter(label == "Orycteropus.afer") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold RICR), respectively, were related to tumor suppression. In contrast, only 4.9% (2/41) of the pathways that were enriched in hyrax (`r text.ratio.nodeByMRCA.ancNode("Microhyrax.lavocati", "Procavia.capensis") %>% str_remove("x decrease")`-fold IBM, `r MRCA(tree.Atlantogenata.table, "Microhyrax.lavocati", "Procavia.capensis") %>% pull(log2CancerSucceptabilityChange) %>% round(2) %>% as.numeric %>% abs`-log_{2}-fold RICR) were related to tumor suppression (**Figure \@ref(fig:Figure-ML-UpSet-WC)**).

Unexpectedly, however, lineages without major increases in body size, lifespan, or intrinsic cancer risk were also enriched for tumor suppressor pathways. For example, 13.2% (9/68), 36.1% (13/36), and 20% (20/100) of the pathways that were enriched in the stem-lineages of Afroinsectivoa and Afrosoricida, and in E. telfairi, respectively, were related to tumor suppression (**Figure \@ref(fig:Figure-ML-UpSet-WC)**). These data suggest that either duplication of genes in tumor suppressor pathways is common in Afrotherians, or that there is a systemic bias in our pathways enrichment analyses. To explore this possibility further we generated 5000 randomly sampled gene sets of sizes between 10 and 5000 genes from the query gene set, and tested for enriched pathways using ORA implemented in WebGestalt. After filtering out pathways with a median p-value greater than 0.05, we saw no enrichments of cancer pathways above 157 genes; below that number, between 12.2% and 18% of enriched pathways were classified as cancer pathways by our criteria. Without considering p-values, the percentage of cancer pathways approaches 15.4% (213/1381) in simulated sets as the number of genes in a set increases. When directly comparing our simulated results with the *Atlantogenatan* results by gene set size, we see that only cancer pathway enrichments in `r ORA.pcCancer.aboveSim$node %>% str_replace_all("\\.", " ") %>% sprintf("_%s_", .) %>% paste0(collapse=", ") %>% str_replace(", (?=_\\w+ ?\\w*_$)", ", and ")` are above their simulated cancer pathway percentages.

## Tumor suppressor pathways enriched exclusively within Proboscideans

While duplication of tumor suppressor genes is common in Afrotheria, we identified 12 pathways related to tumor suppression that were enriched exclusively in the Proboscidean stem-lineage (**Figure \@ref(fig:Figure-ML-UpSet-WC)C**). Among these were pathways related to the cell cycle, including “G0 and Early G1”, “G2/M Checkpoints” and “Phosphorylation of the APC/C”, pathways related to DNA damage repair  including “Global Genome Nucleotide Excision Repair (GG-NER)”, “HDR through Single Strand Annealing (SSA)”, “Gap-filling DNA repair synthesis and ligation in GG-NER”, “Recognition of DNA damage by PCNA-containing replication complex”, and “DNA Damage Recognition in GG-NER”, pathways related to telomere biology including “Extension of Telomeres” and “Telomere Maintenance”, pathways related to the apoptosome including “Activation of caspases through apoptosome-mediated cleavage”, pathways related to “mTORC1-mediated signaling” and “mTOR signaling”. Thus we conclude that duplication of genes in tumor suppressor pathways is pervasive in Afrotherians, but that genes in some pathways related to cancer biology and tumor suppression are uniquely duplicated in large-bodied (or long-lived) Proboscideans.

Among the genes uniquely duplicated within Proboscideans are TP53, COX20, LAMTOR5, PRDX1, STK11, BRD7, MAD2L1, BUB3, UBE2D1, SOD1, LIF, MAPRE1, CNOT11, CASP9, CD14, HMGB2. Two of these, TP53 and LIF, have been previously described [@Abegglen:JAMA2015; @Sulak2015; @Vazquez2018]. These genes are significantly enriched in pathways involved in apoptosis, cell cycle regulation, and both upstream and downstream pathways involving TP53. The majority of these genes are actively expressed in extant African Elephants, suggesting that they maintain functionality (**Figure \@ref(fig:Figure-LoxAfrExpressedDups)D**).

## Coordinated duplication of TP53-related genes in Proboscidea

Prior studies found that the “master” tumor suppressor TP53 duplicated multiple times in elephants [@Abegglen:JAMA2015; @Sulak2015], motivating us to further study duplication of genes involved in TP53-related pathways Proboscidea. We traced the evolution of all genes involved in this TP53-duplicated pathway that were duplicated in the African Elephant, which has the most complete genome and for which several RNA-seq data sets are available. We found that the initial duplication of TP53 in Tethytheria, where body size expanded, was preceded by the duplication of GTF2F1 and STK11 in Paenungulata; and was coincident with the duplication of BRD7. These three genes are involved in regulating the transcription of TP53 [@Liang2013; @Launonen2005; @Drost2010; @Burrows2010], and their duplication prior to that of TP53 may have facilitated re-functionalization of TP53 retroduplicates. Interestingly, STK11 is a tumor suppressor gene in its own right, and plays additional roles in mediating tumor suppression via p21-induced senescence [@Launonen2005]. The other genes that are duplicated in the pathway are all downstream of TP53; these genes duplicated either coincident with TP53, in the case of SIAH1, or subsequently in Probodiscea, Elephantidae, or extant elephants (**Figure \@ref(fig:Figure-LoxAfrExpressedDups)**). These genes are all expressed in RNA-seq data, suggesting that they encode functional genes (**Figure \@ref(fig:Figure-LoxAfrExpressedDups)**).

# Discussion

Among the evolutionary, developmental, and life history constraints on the evolution of large bodies and long lifespans is an increased risk of developing cancer. While body size and lifespan are correlated with cancer risk within species, there is no correlation between species because large and long-lived organisms have evolved enhanced cancer suppression mechanisms. While this ultimate evolutionary explanation is straightforward [@Peto:20151c2], determining the mechanisms that underlie the evolution of enhanced cancer protection is challenging because many mechanisms likely contribute to evolution of reduced cancer risk. Previous candidate gene studies in elephants have identified duplications of tumor suppressors such as TP53 and LIF, among others, suggesting that an increased copy number of tumor suppressors may contribute to the evolution of large body sizes in the elephant lineage [@Abegglen:JAMA2015; @Caulin2015; @Sulak2015; @Doherty2016; @Vazquez2018]. Here we: 1) Trace the evolution of body sizes in Eutherian mammals, with particular reference to Afrotherians; 2) Infer changes in cancer susceptibility across Afrotherian lineages; and 3) Use a genome-wide screen to determine gene duplications in Afrotherian genomes, including multiple living and extinct Proboscideans, are enriched in tumor suppressor pathways.

## Correlated evolution of large bodies and reduced cancer risk

One of the challenges for determining the mechanisms that underlie the evolution of large bodies, long lifespans, and reduced cancer risk is identifying lineages that are large and/or long-lived and closely related to species with normal body sizes and lifespans. One such lineage appears to be elephants, which are deeply nested within Afrotherians with smaller bodies and that are generally shorter-lived. To reconstruct the evolution of body size across Afrotherians, and determine in which lineages exceptional body sizes evolved generated a time-calibrated super tree of Eutherian mammals that combined the X and Y trees. Similar to previous studies, we found that the evolution of gigantism in Afrotheria was not limited to the living and extinct Proboscideans, but independently evolved in multiple lineages including hyraxes and sea cows. These data indicate that the Paenungulata is a particularly good model in which to explore mechanisms that underlie the evolution of exceptionally large bodies.

The absence of a correlation between body size or lifespan and cancer risk indicates that large, long-lived species evolved reduced cancer risk. Previous studies, for example, have shown that while cancer incidence varies by species, the rate is relatively stable at ~5% for both large and small mammals; rates can increase significantly for species with transmissible cancers such as Tasmanian devils or that live in particularly polluted environments such as Minke whales that inhabit the St. Lawrence estuary. Therefore, we used our ancestral body size and lifespan estimates to infer the magnitude of changes in intrinsic cancer risk across species (following Peto (2015)) that would be required to keep cancer rates stable across species. As expected, we identified significant reductions in intrinsic cancer risk in large bodied lineages, most notably in Pseudoungulata (13 log_{2}-fold RICR), Tethytheria (7.84 log_{2}-fold RICR), and Proboscidea (3.14 log_{2}-fold RICR).

## All Afrotheria are equal, but some Afrotheria are more equal than others

The hundred- to hundred-million-fold reductions in intrinsic cancer risk associated with the evolution of large body sizes in some Afrotherian lineages, in particular Paenungulates such as elephants and mastodons, suggests that these lineages must have also evolved remarkable mechanisms to suppress cancer. While our initial hypothesis was that large bodied lineages would be uniquely enriched in duplicate tumor suppressor genes compared to other smaller bodied lineages, we unexpectedly found that duplication of genes in tumor suppressor pathways was common in Afrotheria regardless of body size. These data suggest that this abundance of tumor suppressors may have contributed to the evolution of large bodies and reduced cancer risk, but that these processes were not necessarily coincident. Interestingly, pervasive duplication of tumor suppressors may also have contributed to the repeated evolution of large bodies in hyraxes and sea cows, because at least some of the genetic changes that underlie the evolution of reduced cancer risk was common in this group. It remains to be determined whether our observation of pervasive duplication of tumor suppressors also occurs in other lineages. Using a similar reciprocal best BLAST/BLAT approach that focused on estimating copy number of known tumor suppressors in mammalian genomes, for example, Caulin et al. (2015) found no evidence between copy number or tumor suppressor genes with either body mass or longevity, whilst Tollis et al. (2020) found a correlation between copy number and longevity (but not body size) [@Caulin2015; @Tollis2020]. These opposing conclusions may be results from differences in the number of genes (81 vs 548) and genomes (8 vs 63) analyzed, highlighting the need for genome-wide analyses of many species that vary in body size and longevity.

While we found that duplication of tumor suppressor genes is common in Afrotheria, genes that duplicated in the Proboscidean stem-lineage (**Figure \@ref(fig:Figure-ML-UpSet-WC)C**) were uniquely enriched in functions and pathways that may be related to the evolution of unique anti-cancer cellular phenotypes in the elephant lineage. Elephant cells, for example, cannot be experimentally immortalized [@Fukuda:20161c2; @Gomes2011], rapidly repair DNA damage [@Hart1974; @Francis1981; @Sulak2015], are extremely resistant to oxidative stress [@Gomes2011], yet are also extremely sensitive to DNA damage [@Sulak2015; @Vazquez2018; @Abegglen:JAMA2015]. Several pathways related to DNA damage repair, nucleotide excision repair (NER) in particular, were uniquely enriched among genes that duplicated within Proboscideans suggesting a connection between duplication of genes involved in NER and rapid DNA damage repair [@Hart1974; @Francis1981]. Similarly, the duplicate SOD1 gene in elephants may confer the resistance of elephant cells to oxidative stress [@Gomes2011]. Pathways related to the cell cycle were also uniquely enriched among genes that duplicated within Proboscideans, and cell cycle dynamics are different in elephants compared to other species; population doubling (PD) times for African and Asian elephant cells were 13-16 days, while it 21-28 days for other Afrotherians [@Gomes2011]. Finally, the role of “mTOR signaling” in the biology of aging is well-known. Collectively these data suggest that gene duplications within Proboscideans may underlie some of their cellular phenotypes that contribute to cancer resistance.

## There’s no such thing as a free lunch: Trade-offs and constraints on tumor suppressor copy number

The lack of a preponderance of duplicate tumor suppressor in the stem-lineage of Proboscideans, coincident with the evolution of very large body sizes, may suggest that Afrotherians have all the additional tumor suppressors needed to evolve large body sizes, that duplication of tumor suppressors is not a general phenomena related to the evolution of cancer resistance or only a minor one, or reflect a trade-off between the protective effects of increased tumor suppressor number on cancer risk and potentially deleterious consequences of augmented tumor suppressor copy number. Indeed, increased copy number of tumor suppressors can have deleterious consequences. Overexpression of TP53 in mice, while protective against cancer, is associated with progeria, premature reproductive senescence, and early death; however, transgenic mice with a duplication of the TP53 locus that includes native regulatory elements are healthy and experience normal aging, while also demonstrating an enhanced response to cellular stress and lower rates of cancer [@Tyner:20021c2; @GarciaCao:20021c2]. These data suggest duplication of tumor suppressors can contribute to augmented cancer resistance, if the duplication includes sufficient regulatory architecture to direct spatially and temporally appropriate gene expression. Thus it is fascinating that duplication of genes that regulate TP53 function, such as STK11, SIAH1, and BRD7, preceded the retroduplication TP53 in the Proboscidean stem-lineage, which may have mitigated toxicity arising from dosage imbalances. Similar co-duplication events may have alleviated the negative pleiotropy of tumor suppressor gene duplications to enable their persistence and allow for subsequent co-option during the evolution of cancer resistance.

## Conclusions, caveats, and limitations

Our genome-wide results suggest that duplication of tumor suppressors is pervasive in Afrotherians and may have enabled the evolution of larger body sizes by lowering intrinsic cancer risk either prior to (preadaptation or exaptation) or coincident with increasing body size. However, our study has several inherent limitations, for example, we have shown that genome quality plays an important role in our ability to identify duplicate genes and several species have poor quality genomes. Conversely, without comprehensive gene expression data we cannot be certain that duplicate genes are actually expressed. Duplication of tumor suppressor genes is also unlikely to be the only mechanism responsible for the evolution of large body sizes, long lifespans, and reduced cancer risk. The evolution of regulatory elements and coding genes with other non-canonical tumor suppressor functions are also important for mediating the cancer risk. We also assume that duplicate genes preserve their original functions and increase overall gene dosage. Many processes, however, such as developmental systems drift, neofunctionalization, and subfunctionalization can cause divergence in gene functions [@Rastogi2005; @QianAndZhang2014; @Stoltzfus1999], leading to inaccurate inferences of dosage effects and pathway functions.

# Acknowledgements

We would like to thank Olga Duchenko and Eren Aiden at Baylor College of Medicine for the Hi-C scaffolded Procavia capensis, Trichechus manatus, Orycteropus afer, and Choloepus hoffmannii genomes. We would also like to thank 
D.H. Vazquez for his indispensable support.

# Conflicts of Interest

The Authors have no conflicts of interest to report.

# Funding Source

We would like to thank the Department of Human Genetics at the University of Chicago for supporting this project.

# FIGURES

(ref:Figure-intro) *Atlantogenatans* span a diverse range of body sizes [@PuttickAndThomas2015; @HAGR].

```{r Figure-intro, fig.cap="(ref:Figure-intro)",  out.width="6in", fig.asp=1.2}
atlantogenata.tree.size.lifespan

ggsave("../output/pubFiles/Figures/Figure1.eps", plot=atlantogenata.tree.size.lifespan, width=5, height = 6, units="in", dpi=1200, device = grDevices::cairo_ps)
ggsave("../output/pubFiles/Figures/Figure1.png", plot=atlantogenata.tree.size.lifespan, width=5, height = 6, units="in", dpi=1200)
```

```{r Plot-Bodysize, include=F}
p.bodysize.atlantogenata <- dat.atlantogenata %>%
  ggplot(
    aes(
      y=as.factor(label),
      x=Median, 
      xmin=lnSize.low,
      xmax=lnSize.high
      )
  ) +
  geom_pointrange(size=0.5) +
  geom_point(aes(color=Median), size=1) +
  geom_point(aes(x=lnSize.low, color=lnSize.low), size=1.5, stroke=2, pch="[") +
  geom_point(aes(x=lnSize.high, color=lnSize.high), size=1.5, stroke=2, pch="]") +
  scale_color_distiller(limits=c(size.scale.expanded.min, size.scale.expanded.max), 
                        palette = "Blues", 
                        direction = 1,
                        guide = guide_colorbar(
                          title="Body size (log)",
                          title.theme =  element_text(size = 6, face = "bold", colour = "black"),
                          title.position = "top",
                          title.hjust = 0.5
                          )
                        ) +
  labs(
    x="Body size (95%CI)",
    y="Node"
  ) + 
  theme_pubclean()+
  labs_pubr() + 
  theme(
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text = element_text(size = 6, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 8, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 8, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black"),
    legend.position = "bottom"
  )
```

```{r Plot-Lifespan-Atlantogenata}
p.lifespan.atlantogenata <- dat.atlantogenata %>%
  ggplot(
    aes(
      y=as.factor(label),
      x=Lifespan, 
      xmin=Lifespan.low,
      xmax=Lifespan.high
      )
  ) +
  geom_pointrange(size=0.5) +
  geom_point(aes(color=Lifespan), size=1) +
  geom_point(aes(x=Lifespan.low, color=Lifespan.low), size=1.5, stroke=2, pch="[") +
  geom_point(aes(x=Lifespan.high, color=Lifespan.high), size=1.5, stroke=2, pch="]") +
  # scale_color_viridis_c(
  #   limits=c(lifespan.scale.expanded.min, lifespan.scale.expanded.max), 
  #   option = "C", 
  #   guide = guide_colorbar(
  #     title="Lifespan (yrs)",
  #     title.theme =  element_text(size = 6, face = "bold", colour = "black"),
  #     title.position = "top",
  #     title.hjust = 0.5
  #   )
  # ) +
  scale_color_distiller(
    palette = "Reds", direction = 1,
    guide = guide_colorbar(
      title="Lifespan (yrs)",
      title.theme =  element_text(size = 6, face = "bold", colour = "black"),
      title.position = "top",
      title.hjust = 0.5
    )
  ) + 
  labs(
    x="Estimated Lifespan from Body Sizes and Body Size 95% CI",
    y="Node"
  ) + 
  theme_pubclean()+
  labs_pubr() + 
  theme(
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text = element_text(size = 6, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 8, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 8, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black"),
    legend.position = "bottom"
  )
```

```{r Plot-RICR-Atlantogenata}
p.RICR.atlantogenata <- dat.atlantogenata %>%
  ggplot(
    aes(
      y=as.factor(label),
      x=log2CancerSucceptabilityChange, 
      # xmin=log2CancerSucceptabilityChange.low,
      # xmax=log2CancerSucceptabilityChange.high
      )
  ) +
  scale_color_viridis_c(
    option = "C",
    guide = guide_colorbar(
      title="Cancer Risk (log2)",
      title.theme =  element_text(size = 6, face = "bold", colour = "black"),
      title.position = "left",
      title.hjust = 0.5, 
      title.vjust = 0.75,
      reverse = T,
      order = 2
    )
  ) +
  # geom_pointrange(size=0.5) +
  geom_point(aes(color=Lifespan), size=1) +
  # geom_point(aes(x=Lifespan.low, color=log2CancerSucceptabilityChange.low), size=1.5, stroke=2, pch="[") +
  # geom_point(aes(x=Lifespan.high, color=log2CancerSucceptabilityChange.high), size=1.5, stroke=2, pch="]") +
  labs(
    x="Estimated RICR from Estimated Body Sizes and Lifespans",
    y="Node"
  ) + 
  theme_pubclean()+
  labs_pubr() + 
  theme(
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text = element_text(size = 6, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 8, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 8, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black"),
    legend.position = "bottom"
  )
ggsave("../output/pubFiles/Figures/Figure2-RICR.eps", plot=p.RICR.atlantogenata, width=9, height = 9, units="in", dpi=1200, device=grDevices::cairo_ps)
```

```{r Plot-DoubleTree-Bodysize-CancerSuscept, include=F}
### Body Size ###

p.tree.atlantogenata.genomes <- tree.Atlantogenata %>% 
  ggtree(
    aes(color=as.numeric(lnSize)),
    branch.length = "sqrt.rate"
  )

### Cancer Suscept ###
tree.Atlantogenata.table <-  tree.Atlantogenata.table %>%  
  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% 
  as.treedata

p.tree.Atlantogenata.table <- tree.Atlantogenata.table %>% 
  ggtree(
    aes(color=as.numeric(log2CancerSucceptabilityChange))
  )

### Double Tree ###

d.dbltree.3 <- p.tree.atlantogenata.genomes$data
d.dbltree.4 <- p.tree.Atlantogenata.table$data

d.dbltree.4$x <- max(d.dbltree.4$x) - d.dbltree.4$x + max(d.dbltree.3$x) + 500

p.dbltree.all <-
  tree.Atlantogenata %>% 
  ggtree(., branch.length = "sqrt.rate", aes(color=lnSize, fill=lnSize), size=1) +
  # scale_color_distiller("Body size (log)", palette = "Blues", direction = 1) + 
  scale_color_distiller(
    limits=c(size.scale.expanded.min, size.scale.expanded.max), 
    palette = "Blues", 
    direction = 1,
    guide = guide_colorbar(
      title="Body size (log)",
      title.theme =  element_text(size = 6, face = "bold", colour = "black"),
      title.position = "left",
      title.hjust = 0.5, 
      title.vjust = 0.75,
      order = 1
    )
    # guide=guide_none()
  ) +
  scale_fill_distiller(
    limits=c(lifespan.scale.expanded.min, lifespan.scale.expanded.max), 
    palette = "Reds", 
    direction = 1,
    guide = guide_colorbar(
      title="Lifespan",
      title.theme =  element_text(size = 6, face = "bold", colour = "black"),
      title.position = "left",
      title.hjust = 0.5, 
      title.vjust = 0.75,
      order = 3
    )
    # guide=guide_none()
  ) +
  geom_nodepoint(size=2) + 
  geom_tippoint(size=2) +
  # geom_tiplab(aes(label = if_else(is.na(Genome), "", label)), 
  #             x=(min(d.dbltree.4$x)+max(d.dbltree.3$x))/2, color="black", hjust = 0.5) +
  geom_text_repel(aes(label = if_else(is.na(Genome), "", label %>% str_replace_all("\\.", " "))), direction = "y", force = 0.5, segment.size = NA, size=2.5,
                  x=(min(d.dbltree.4$x)+max(d.dbltree.3$x))/2, color="black", hjust = 0.5) +
  new_scale_color() + 
  geom_tree(data=d.dbltree.4, aes(color=log2CancerSucceptabilityChange), inherit.aes = F, size=1) + 
  geom_nodepoint(data=d.dbltree.4, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) + 
  geom_tippoint(data=d.dbltree.4, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) +
  # scale_color_viridis_c(option = "D") + 
  # scale_color_distiller(
  #   palette = "Reds", direction = 1,
  #   guide = guide_colorbar(
  #     title="Cancer Risk (log2)",
  #     title.theme =  element_text(size = 6, face = "bold", colour = "black"),
  #     title.position = "top",
  #     title.hjust = 0.5
  #   )
  # ) + 
  scale_color_viridis_c(
    option = "C",
    guide = guide_colorbar(
      title="Cancer Risk (log2)",
      title.theme =  element_text(size = 6, face = "bold", colour = "black"),
      title.position = "left",
      title.hjust = 0.5, 
      title.vjust = 0.75,
      reverse = T,
      order = 2
    )
  ) +
  theme(
    plot.background = element_rect(fill="grey", color="grey"),
    panel.background = element_rect(fill="grey", color="grey"),
  legend.text = element_text(
    size = 6, 
    face = "plain", colour = "black"
  ),
    legend.background = element_blank(),
  legend.position = c(0.5, 0.15),#"bottom",
  legend.box = "vertical",
  legend.direction = "horizontal",
  legend.box.just = "right",
  legend.spacing.y = unit(0.5,"mm")
  # legend.direction = "vertical"
  )

p.dbltree.all
```

(ref:Figure-DoubleTree-Bodysize-Lifespan) Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata. **A)** Back-to-back phylogeny for *Atlantogenata*, scaled by log-fold increases in body mass (IBM, left) and log_{2}-fold reduction in intrinsic cancer risk (RICR, right). **B)** Extant body sizes, plus estimated body sizes, for highlighted species and their ancestral nodes. **C)** Estimated Lifespans for body sizes in **B)**. Intervals are based on predicted lifespans of the 95% CIs for body size.

```{r Figure-DoubleTree-Bodysize-Lifespan, fig.cap="(ref:Figure-DoubleTree-Bodysize-Lifespan)", fig.width=12, fig.height = 9}
fig.dtbl.a <- p.dbltree.all
fig.dtbl.b <- p.bodysize.atlantogenata + theme(
  legend.position = "none"
)
fig.dtbl.c <- p.lifespan.atlantogenata + theme(
  axis.title.y = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks.y = element_blank(),
  legend.position = "none"
)

fig.dtbl.bc <- ggarrange(
  fig.dtbl.b, fig.dtbl.c, 
  nrow=1, ncol = 2, 
  labels=c("B", "C"),
  font.label = list(size=6, face="bold")#,
  # common.legend = T,
  # legend = "bottom"
)

fig.dtbl <-
  ggarrange(
    fig.dtbl.a,
    ggplotify::as.ggplot(fig.dtbl.bc),
    nrow=2, ncol=1,
    labels = c("A", ""),
    font.label = list(size=6, face="bold")#, 
    # common.legend = T,
    # legend = "bottom"
  )
fig.dtbl
ggsave("../output/pubFiles/Figures/Figure2.eps", plot=fig.dtbl, width=12, height = 9, units="in", dpi=1200, device=grDevices::cairo_ps)
ggsave("../output/pubFiles/Figures/Figure2.png", plot=fig.dtbl, width=12, height = 9, units="in", dpi=1200)
# fig.dtbl.patch <- 
#   fig.dtbl.a / (fig.dtbl.b + fig.dtbl.c) + 
#   patchwork::plot_layout(guides="auto")

```

```{r Plot-GeneDup-Cladogram, fig.cap="(ref:Figure-GeneDup-Cladogram)", out.height="6in", out.width="6in"}
p.ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts.withSize.afrotheria  %>% 
  ggtree(
    .,
    branch.length = "none",
    # continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=1
    ) %>% 
  flip(tree_view = ., node1 =nodeid(., "Loxodonta.africana"), node2 = MRCA(., "Palaeoloxodon.antiquus", "Loxodonta.cyclotis"))  +
  scale_color_distiller(
    limits=c(size.scale.min, size.scale.max), 
    palette = "Blues", 
    direction = 1,
    guide=guide_none()
  ) +
  scale_fill_distiller(
    limits=c(size.scale.min, size.scale.max), 
    palette = "Blues", 
    direction = 1,
    guide = guide_colorbar(
      title="Body size\n(log kg)",
      title.theme =  element_text(size = 6, face = "bold", colour = "black"),
      title.position = "left", 
      title.hjust = 0.5,
      barheight = 0.5,
      barwidth = 3, 
      label.position = "bottom"
    )
  ) +
  #Standard
  geom_tiplab(aes(label=Spc.short.bold), family="Helvetica",
              offset = 0.03, color="black", size=2.5, parse=T)+
  geom_tiplab(aes(label=N.Genes),  family="Helvetica",
              hjust=1, color="black", size=2, geom="label", fontface="bold")+  
  geom_nodelab(aes(subset = !isTip & !label %in% c("Loxodontini", "Afrotheria","Loxodona", "Elephantina", "Mammuthus"), x=branch, 
                   label=N.Genes),  family="Helvetica",
               geom="label", nudge_x = -0.5, 
               color="black", size=2, fontface="bold")+ 
  geom_nodelab(aes(subset = !isTip & !label %in% c("Loxodontini","Loxodona", "Elephantina", "Proboscidea", "Afrosoricida", "Afrotheria", "Pseudoungulata", "Mammuthus"), x=branch, 
                   label=label),  family="Helvetica",
               geom="text", nudge_y = 0.52, nudge_x = -0.5, hjust=0.55,
               color="black", size=2, fontface="bold")+ 
  #Up shifted
  geom_nodelab(aes(subset = !isTip & label %in% c("Loxodontini", "Mammuthus"), x=branch,
                   label=N.Genes), family="Helvetica",
               geom="label", nudge_x = -0.5, nudge_y = 0.17,
               color="black", size=2, fontface="bold")+
  geom_nodelab(aes(subset = !isTip & label %in% c("Loxodontini", "Mammuthus"), x=branch,
                   label=label), family="Helvetica",
               geom="text", nudge_y = 0.67, nudge_x = -0.5, hjust=0.5,
               color="black", size=2, fontface="bold")+
  #Down shifted
  geom_nodelab(aes(subset = !isTip & label %in% c("Loxodona", "Elephantina"), x=branch,
                   label=N.Genes), family="Helvetica",
               geom="label", nudge_x = -0.5, nudge_y = -0.16,
               color="black", size=2, fontface="bold")+
  geom_nodelab(aes(subset = !isTip & label %in% c("Loxodona", "Elephantina"), x=branch,
                   label=label), family="Helvetica",
               geom="text", nudge_y = 0.37, nudge_x = -0.5, hjust=0.5,
               color="black", size=2, fontface="bold")+
  #Left shifted
  geom_nodelab(aes(subset = !isTip & label %in% c("Proboscidea", "Afrosoricida"), x=branch,
                   label=label), family="Helvetica",
               geom="text", nudge_y = 0.52, nudge_x = -0.5, hjust=0.75,
               color="black", size=2, fontface="bold")+
  #Right Shifted
  geom_nodelab(aes(subset = !isTip & label %in% c("Pseudoungulata"), #x=branch,
                   label=label), family="Helvetica",
               geom="text", nudge_y = 0.52, nudge_x = -0.5, hjust=0.5,
               color="black", size=2, fontface="bold")+
  # geom_text(label="Rate of Change\nBody Size", x=0, y=12.5, color = "black", size = 2, hjust=0.5, inherit.aes = F) +
  theme_tree() + 
  # geom_treescale(
  #   x=1.5, y=12.5,
  #  offset = -0.6, width = 1, linesize=1, fontsize = 3
  # )+
  # geom_treescale(
  #   x=3.5, y=1,
  #  offset = -0.6, width = 1, linesize=1, fontsize = 2.5,  family="Helvetica"
  # )+
  xlim(-1,10.5) +
  coord_cartesian(clip="off")+
  # xlab("\nRate of Change: Size") +
  theme(
    panel.background = element_blank(), 
    plot.background = element_blank(), 
    legend.background = element_blank(),
    legend.direction = "horizontal",
    legend.position=c(0.19,0.90),
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                       family="Helvetica",
                        debug = FALSE), 
    axis.text.x = element_blank(),
    axis.title.x = element_text(hjust=0.08, vjust=24),
    axis.title = element_text(size = 6, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 6, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 6, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black")
  )
# p.ml.tree.increasing.geneCounts.withSize2
```

```{r Plot-Upset, include=F}
ORA.all.increasing.pcCancer <- ORA.all %>% 
  mutate(node = str_replace_all(node, "_", ".")) %>% 
  # filter(Database=="Reactome", Direction == "increased", RBHB=="RBB", FDR.Threshold<=1, pValue<=0.05) %>%
  filter(!node %in% c("Chrysochloris.asiatica","Elephantulus.edwardii", "Afroinsectivora")) %>% # Remove problem nodes
  distinct(node, description, Class) %>% 
  group_by(node) %>% 
  summarise(N.Pathway = n(), N.Cancer = sum(Class != "Other")) %>% 
  mutate(node = str_replace_all(node, "\\.", " "))
  # mutate(node = str_replace_all(node, " ", "."))

ORA.all.increasing.pcCancer[is.na(ORA.all.increasing.pcCancer)] <- 0

phylo.levels <- p.ml.tree.increasing.geneCounts.withSize$data %>% 
  arrange(desc(y)) %>% 
  filter(!label %in% c("Chrysochloris.asiatica","Elephantulus.edwardii", "Afroinsectivora")) %>% # Remove problem nodes
  pull(label) %>%
  str_replace_all(., "\\.", " ") %>%
  # str_replace_all(., " ", ".") %>% 
  stringi::stri_remove_na()

list.ORA.increased.reactome <- ORA.all %>% 
  mutate(node = str_replace_all(node, "_", ".")) %>% 
  # filter(Database=="Reactome", Direction == "increased", RBHB=="RBB", FDR.Threshold<=1, pValue<=0.05, str_detect(node, "^Any$|^Any.*", negate = T)) %>% 
  filter(!node %in% c("Chrysochloris.asiatica","Elephantulus.edwardii", "Afroinsectivora")) %>% # Remove problem nodes
  distinct(node, description) %>% 
  mutate(node = str_replace_all(node, "\\.", " ")) %>%
  # mutate(node = str_replace_all(node, " ", ".")) %>% 
  split(., .$node) %>% 
  lapply(pull, "description")

list.ORA.increased.reactome <- list.ORA.increased.reactome[!is.na(list.ORA.increased.reactome)]

missing.sets <- setdiff(phylo.levels, names(list.ORA.increased.reactome))

combMat.ORA.increased.reactome<- list.ORA.increased.reactome %>%
  c(as.list(rep(list(character(0)), times=length(missing.sets)) %>% set_names(missing.sets))) %>% 
  (function(l, ord=phylo.levels){l[ord]}) %>% 
  make_comb_mat()

list.ORA.increased.reactome.cancer <- ORA.all %>% 
  mutate(node = str_replace_all(node, "_", ".")) %>% 
  # filter(Database=="Reactome", Direction == "increased", RBHB=="RBB", FDR.Threshold<=1, pValue<=0.05, str_detect(node, "^Any$|^Any.*", negate = T)) %>% 
  filter(!node %in% c("Chrysochloris.asiatica","Elephantulus.edwardii", "Afroinsectivora")) %>% # Remove problem nodes
  filter(Class != "Other") %>% 
  distinct(node, description) %>% 
  mutate(node = str_replace_all(node, "\\.", " ")) %>%
  split(., .$node) %>% 
  lapply(pull, "description")

missing.sets.cancer <- setdiff(phylo.levels, names(list.ORA.increased.reactome.cancer))

combMat.ORA.increased.reactome.cancer <- list.ORA.increased.reactome.cancer %>% 
  c(as.list(rep(list(character(0)), times=length(missing.sets.cancer)) %>% set_names(missing.sets.cancer))) %>% 
  (function(l, ord=phylo.levels){l[ord]}) %>% 
  make_comb_mat()

set.sizes.ORA <- combMat.ORA.increased.reactome %>% set_size() %>% as.data.frame() %>% t %>% as.data.frame %>% mutate(Class = "All")
set.sizes.ORA.cancer <- combMat.ORA.increased.reactome.cancer %>% set_size() %>% as.data.frame() %>% t %>% as.data.frame() %>% mutate(Class = "Cancer")

ORA.set.cancer.mat <- 
  bind_rows(
  set.sizes.ORA,
  set.sizes.ORA.cancer
) %>% 
  pivot_longer(-Class) %>% 
    pivot_wider(id_col=name, names_from=Class) %>% 
    mutate(Not.cancer = All-Cancer) %>% 
  dplyr::select(-All) %>% 
  mutate(name = factor(name, levels=set_name(combMat.ORA.increased.reactome.cancer))) %>%
  arrange(name)  %>%
    column_to_rownames("name")

ORA.comb.cancer.mat <- comb_name(combMat.ORA.increased.reactome.cancer) %>% 
  set_names(.,.) %>% 
  lapply(extract_comb, m=combMat.ORA.increased.reactome) %>% 
  lapply(. %>% class.pathway %>% unname %>% sapply(function(x){ifelse(x !="Other", 1, 0)})) %>% 
  lapply(function(x){c(All=length(as.numeric(x)), Cancer=sum(as.numeric(x)))}) %>% 
  as.data.frame %>% 
  rownames_to_column("Class") %>% 
  rename_all(. %>% str_remove("X")) %>% 
  data.table::transpose(make.names = "Class", keep.names = "ID") %>% 
  mutate(Not.cancer = All-Cancer) %>% 
  dplyr::select(-All) %>% 
  # mutate(ID = factor(ID, levels=comb_name(combMat.ORA.increased.reactome.cancer))) %>%
  # arrange(ID)  %>%
  column_to_rownames("ID")

annotate.comb.ORA <- ORA.comb.cancer.mat %>% 
  rownames_to_column("ID") %>% 
  transmute(ID=ID, pc.Cancer = divide_by(Cancer,sum(Cancer + Not.cancer)) %>% scales::label_percent(accuracy = 0.1)(.) %>% str_remove("^0.0%$")) %>% 
  # mutate(ID = factor(ID, levels=phylo.levels)) %>% 
  # arrange(ID)  %>% 
  column_to_rownames("ID") %>% t

color.comb.ORA <- ORA.comb.cancer.mat %>% 
  rownames_to_column("ID") %>% 
  transmute(ID=ID, pc.Cancer = Cancer %>% divide_by((Cancer + Not.cancer)) %>% 
              sapply(function(x){ifelse(x >= 0.25, 
                                        # "#F2AA4CFF",
                                        # "#00539CFF",
                                        "#2BAE66FF",
                                        "#101820FF")})
            ) %>% 
  # mutate(ID = factor(ID, levels=phylo.levels)) %>% 
  # arrange(ID) %>% 
  column_to_rownames("ID") %>% 
  t
  # split(., .$ID) %>%
  # lapply(pull, "pc.Cancer")

color.set.ORA <- ORA.set.cancer.mat %>% 
    rownames_to_column("ID") %>% 
    transmute(ID=ID, pc.Cancer = Cancer %>% divide_by((Cancer + Not.cancer)) %>% str_replace_na(replacement=0) %>% 
                  sapply(function(x){ifelse(x >= 0.25, 
                                            # "#F2AA4CFF",
                                            # "#00539CFF",
                                            "#2BAE66FF",
                                            "#101820FF")})
    ) %>% 
  mutate(ID = factor(ID, levels=phylo.levels)) %>% 
  arrange(ID)

color.set.ORA <- set_names(color.set.ORA$pc.Cancer, color.set.ORA$ID)

ordered.phylo <- combMat.ORA.increased.reactome.cancer %>% 
  ComplexHeatmap::set_name() %>% 
  factor(., levels=phylo.levels) %>% 
  sort %>% 
  as.character

# Desired Figure Size: 102mm tall, 152mm wide
## UpSet_Width_Total: 152mm
### Ratio of Histogram:Label:Sets: should be like 1:1:3
upset.total.width = 200
upset.main.width = 100
upset.label.width = 30
upset.sethist.width = 25
upset.padding.width = (upset.total.width - upset.main.width - upset.label.width - upset.sethist.width)

upset.total.height = 125
upset.main.height = 50
upset.combhist.height = 50
upset.padding.height = (upset.total.height - upset.main.height - upset.combhist.height)

upset.ORA.raw <-
  UpSet(
    combMat.ORA.increased.reactome.cancer,
    set_order = ordered.phylo,
    comb_order = order(comb_size(combMat.ORA.increased.reactome.cancer), decreasing = T),
    comb_col = color.comb.ORA,
    pt_size = unit(2, "mm"),
    border = FALSE,
    height = unit(upset.main.height, "mm"),
    width = unit(upset.main.width, "mm"),
    top_annotation = HeatmapAnnotation(
      "Cancer Pathway Intersections" = anno_barplot(
        # ORA.comb.cancer.mat,
        ORA.comb.cancer.mat[,1],
        # ylim = c(0, max(ORA.comb.cancer.mat %>% rowSums())*1.2),
        ylim = c(0, max(ORA.comb.cancer.mat[,1])*1.2),
        border = FALSE,
        gp = gpar(fill = c("#ADEFD1FF", "#00203FFF"),
                  fontsize = 4),
        axis_param = list(
          gp = gpar(fontsize= 6)
        ),
        height = unit(upset.combhist.height, "mm")
      ),
      annotation_name_side = "left",
      annotation_name_rot = 90,
      annotation_name_gp = gpar(fontsize=8)
    ),
    left_annotation = rowAnnotation(
      "Number of Reactome Pathways" = anno_barplot(
        ORA.set.cancer.mat,
        baseline = 0,
        xlim = c(0, max(ORA.set.cancer.mat %>% rowSums)*1.2),
        axis_param = list(
          direction="reverse",
          labels_rot = 0,
          gp = gpar(fontsize= 6)
        ),
        border = FALSE,
        gp = gpar(
          fill = c("#ADEFD1FF", "#00203FFF"),
          fontsize = 6
        ),
        width = unit(upset.sethist.width, "mm")
      ),
      set_name = anno_text(
        set_name(combMat.ORA.increased.reactome.cancer),
        # ordered.phylo,
        location = 0.5,
        just = "center",
        # width = max_text_width(
        #   set_name(combMat.ORA.increased.reactome),
        #   gp = gpar(
        #     fontsize = 4
        #   )
        # ) +
        #   unit(1, "mm"),
        width = unit(upset.label.width, "mm"),
        gp = gpar(
          # color=color.set.ORA,  #doesn't work
          fontsize = 7
        )
      ),
      # group = color.set.ORA %>% unname,
      annotation_name_gp = gpar(fontsize=8)
    ),
    right_annotation = NULL,
    show_row_names=F
  ) #%>%
  # draw(., padding = unit(c(upset.padding.height, upset.padding.width, upset.padding.height, upset.padding.width), "mm"))

upset.ORA <- grid.grabExpr(
  draw(
    upset.ORA.raw, 
    padding = unit(c(0, upset.padding.width, upset.padding.height, 2), "mm")
  ) #+
  #   decorate_annotation(
  #     "Cancer Pathway Intersections", 
  #     {
  #       grid.text(
  #         annotate.comb.ORA[column_order(upset.ORA.raw)], 
  #         x = seq_along(annotate.comb.ORA), 
  #         y = unit(comb_size(combMat.ORA.increased.reactome.cancer)[column_order(upset.ORA.raw)], "native") + unit(2, "pt"), 
  #         default.units = "native", 
  #         just = c("left", "bottom"), 
  #         gp = gpar(fontsize = 7, col = "black"), 
  #         rot = 45
  #       )
  #     }
  #   )
)

```

```{r Plot-Wordcloud}
# The WordCloud are enriched terms that are unique to duplications in the Proboscidea (purple), Tethytheria (blue), or in common with Proboscidea and at least one other lineage (green).
# wc.color.scheme <- c("ProboscideaAndTethytheria"="blue", "OnlyProboscidea" = "purple", "ProboscideaOther"="green", "other"="black")
proboscideans <- c("Loxodontini", "Loxodonta.africana", "Loxodona", 
                     "Loxodonta.cyclotis", "Palaeoloxodon.antiquus", 
                     "Elephantidae", "Elephantina", "Elephas.maximus",
                     "Mammuthus", "Mammuthus.primigenius", "Mammuthus.columbi", 
                     "Proboscidea", "Mammut.americanum")
ORA.all.cancer <- ORA.all[Class != "Other"]
# ORA.all.Proboscidea.cancer <- ORA.all.Proboscidea[Class != "Other"]
ORA.all.Proboscidea.cancer <- ORA.all.cancer[node %in% proboscideans][, .(nlPval = mean(-log(pValue))), by="description"][]

ORA.pathways.tethy <- ORA.all.cancer[node=="Tethytheria", description] %>% unique
ORA.pathways.notTethy <- ORA.all.cancer[node %in% setdiff(clade.highlight, c(proboscideans, "Tethytheria")), description] %>% unique
ORA.pathways.probscideans <- ORA.all.Proboscidea.cancer$description %>% unique

ORA.pathways.onlyProboscidea <- setdiff(ORA.pathways.probscideans, union(ORA.pathways.tethy, ORA.pathways.notTethy))
ORA.pathways.ProboscideaAndTethy <- setdiff(intersect(ORA.pathways.probscideans, ORA.pathways.tethy), ORA.pathways.notTethy)
ORA.pathways.ProboscideaAndOther <- setdiff(intersect(ORA.pathways.probscideans, ORA.pathways.notTethy), ORA.pathways.tethy)

ORA.ProboscideaAndTethytheria <- ORA.all.Proboscidea.cancer[description %in% ORA.pathways.ProboscideaAndTethy,
                                                            .(description = description %>% str_trunc(100, side="center"), nlPval, color="blue")]
ORA.ProboscideaOther <- ORA.all.Proboscidea.cancer[description %in% ORA.pathways.ProboscideaAndOther,
                                           .(description = description %>% str_trunc(100, side="center"), nlPval, color="green")]
ORA.onlyProboscidea <- ORA.all.Proboscidea.cancer[description %in% ORA.pathways.onlyProboscidea,
                                          .(description = description %>% str_trunc(100, side="center"), nlPval, color="purple")]

ORA.legend <- data.table(description="p=0.01", nlPval=-log(0.01), color="black")

ORA.wc <- list(
    ORA.onlyProboscidea,
    ORA.ProboscideaAndTethytheria,
    ORA.ProboscideaOther,
    ORA.legend
  ) %>% 
  rbindlist()

p.ORA.wc <- ORA.wc %>% 
  ggplot(aes(label=description, size=nlPval, color=color)) + 
  # geom_text_wordcloud_area(shape = "square") +
  geom_text_wordcloud() +
  scale_size_area(max_size = 1.8) +
  scale_color_identity() + 
  theme_void()
```

(ref:Figure-ML-UpSet-WC) While duplications of genes in tumor supressor pathways are pervasive in *Atlantogenata*, *Proboscideans* have a greater number of cancer-related pathways enriched, in addition to duplications in various unique pathways. **A)** A cladogram of *Afrotheria* highlighting the estimated number of genes duplicated at each node using maximum likelihood. **B)** Upset plot of cancer-related Reactome pathways enriched in each clade in **A**. **C)** Wordcloud of enriched pathways among Proboscideans that are exclusive to Proboscidea (purple), shared between Proboscidea and Tethytheria (blue), or shared between Proboscidea and one or more other lineages (green).

```{r Figure-ML-UpSet-WC, fig.cap="(ref:Figure-ML-UpSet-WC)", fig.width = 8, fig.height=6}
fig.muw.a <- p.ml.tree.increasing.geneCounts.withSize
fig.muw.b <- upset.ORA
fig.muw.c <- p.ORA.wc

fig.muw <- cowplot::ggdraw(fig.muw.b) + 
  cowplot::draw_plot(fig.muw.a, x = 0.08, y=0.42, width = .38, height = .5) + 
  cowplot::draw_plot(fig.muw.c, x = 0.52, y=0.4, width = .5, height = .6)
fig.muw
ggsave("../output/pubFiles/Figures/Figure3.eps", plot = fig.muw, width = 8, height=6, units = "in", device=grDevices::cairo_ps)
ggsave("../output/pubFiles/Figures/Figure3.png", plot = fig.muw, width = 8, height=6, units = "in")
```

```{r Plot-Upset-LoxAfrORA, include=F}

genesToHighlight.fig <- c("TP53", "COX20", "LAMTOR5", "PRDX1", "STK11", "BRD7", "MAD2L1", "BUB3", "UBE2D1", "SOD1", "LIF")

list.loxAfr.ORA.increased <- ORA.all.cancer %>% 
  rename(Genes = userId) %>% 
  filter(node == "Loxodonta.africana") %>% 
  select(description, Genes)%>% 
  mutate(description = description %>% 
           str_trunc(width = 60, side="right")) %>% 
  separate_rows(Genes, sep=";") %>% 
  split(.,.$description) %>% 
  lapply(pull, "Genes")

combMat.loxAfr.ORA.increased <- list.loxAfr.ORA.increased %>% 
  make_comb_mat()

combMat.loxAfr.ORA.increased.invert <- 
  list.loxAfr.ORA.increased %>% 
  unlist() %>% 
  set_names(names(.) %>% str_remove("\\d+$"), .) %>% 
  split(., names(.)) %>% 
  lapply(unname) %>% .[c("BUB3", "MAD2L1", "MAPRE1", "UBE2D1", "TP53", "CNOT11", "CASP9", "CD14", "HMGB2", "LAMTOR5", "COX20", "PRDX1")] %>% 
  make_comb_mat()

genesToHighlight.fig <- c(genesToHighlight.fig, combMat.loxAfr.ORA.increased.invert %>% set_name) %>% unique

highlight.gene.color <- ggsci::pal_d3("category20")(length(genesToHighlight.fig)) %>% 
  set_names(genesToHighlight.fig)

counts.combMat.loxAfr.ORA.increased <- list.loxAfr.ORA.increased %>% 
  fromList %>% t %>% 
  as.data.frame() %>% 
  rownames_to_column("description") %>% 
  mutate(description=factor(description, levels=set_name(combMat.loxAfr.ORA.increased))) %>% 
  column_to_rownames("description") %>% 
  as.matrix()

color.genes <- highlight.gene.color[dimnames(counts.combMat.loxAfr.ORA.increased)[[2]]]

pathway.class <- c(
  '1111100111101111011001000' = "Cell Cycle",
  '1111100111101110011011000' = "Cell Cycle",
  '0000010010010001100100111' = "TP53-Related",
  '0000000010001110000011000' = "Cell Cycle",
  '0000011000010000100000000' = "Apoptosis",
  '0000011000000000100000000' = "Apoptosis",
  '0000010000000000100000000' = "Apoptosis",
  '0000000000000000000100100' = "TOR Pathway",
  '0000000000000000000000011' = "TP53-Related",
  '0000000000000000000000100' = "TP53-Related"
)

pathways.color <- c(
"TP53-Related" = "#E7298A",
"Cell Cycle" = "#D95F02",
"Apoptosis" = "#7570B3",
"TOR Pathway" = "#1B9E77",
"Oxidative Stress" = "#66A61E"
)

gene.color.table <- comb_name(combMat.loxAfr.ORA.increased) %>%
  set_names(.,.) %>%
  lapply(extract_comb, m=combMat.loxAfr.ORA.increased) %>% 
  tibble(comb=names(.), genes = .) %>% 
  unnest(genes) %>% 
  mutate(class = pathway.class[comb]
  ) %>% 
  bind_rows(
    tibble(comb=c(NA, NA, NA, NA),
           genes = c("STK11", "BRD7", "SOD1","LIF"),
           class = c("Cell Cycle", "Cell Cycle", "Oxidative Stress", "Apoptosis")
    )
  ) %>% 
  mutate(color = pathways.color[class])

color.sets.Pathways <- gene.color.table %>% 
  filter(!is.na(comb), comb %in% comb_name(combMat.loxAfr.ORA.increased)) %>% 
  distinct(comb, color) %>% 
  mutate(comb = comb %>% factor(levels=comb_name(combMat.loxAfr.ORA.increased))) %>% 
  arrange(comb) %>% 
  split(., .$comb) %>% 
  sapply(. %>% pull("color") %>% unname)

UpSet.loxAfr.ORA.Pathway <- grid.grabExpr(
  UpSet(
    combMat.loxAfr.ORA.increased,
    comb_order = order(comb_size(combMat.loxAfr.ORA.increased), decreasing = T),
    # set_order = set_name(combMat.loxAfr.ORA.increased),
    set_order = order(set_size(combMat.loxAfr.ORA.increased), decreasing = T),
    height = unit(5, "cm"),
    width = unit(3.5, "cm"),
    top_annotation = HeatmapAnnotation(
      "Genes per Set" = anno_barplot(comb_size(combMat.loxAfr.ORA.increased),
                                             ylim = c(0, max(comb_size(combMat.loxAfr.ORA.increased))*1.1),
                                             border = FALSE,
                                             gp = gpar(fill = color.sets.Pathways,
                                                       fontsize = 7
                                                       ),
                                             axis_param = list(
                                               gp = gpar(fontsize= 7)
                                             ),
                                             height = unit(2, "cm")
      ),
      annotation_name_side = "left",
      annotation_name_rot = 90,
      annotation_name_gp = gpar(fontsize=8)
    ),
    left_annotation = rowAnnotation(
      "# Genes" = anno_barplot(counts.combMat.loxAfr.ORA.increased, 
                               border = FALSE,
                               axis_param = list(
                                 direction="reverse",
                                 labels_rot = 0,
                                 gp = gpar(fontsize= 7)
                               ),
                               gp=gpar(fill=color.genes,
                                       fontsize= 7),
                               width = unit(4, "cm")
      ),
      set_name = anno_text(set_name(combMat.loxAfr.ORA.increased),
                           location = 0.5,
                           just = "center",
                           width =  max_text_width(
                             set_name(combMat.loxAfr.ORA.increased),
                             gp = gpar(
                               fontsize = 6
                             )
                           ) +
                             unit(0.5, "cm"),
                           gp = gpar(
                             fontsize = 6
                           )
      ),
      annotation_name_gp = gpar(fontsize=8)
    ), 
    right_annotation = NULL,
    show_row_names=F
  ) %>% 
  draw(
  )
)

color.genes2 <- highlight.gene.color[combMat.loxAfr.ORA.increased.invert %>% set_name]

count.sets <- list.loxAfr.ORA.increased %>% 
  unlist() %>% 
  set_names(names(.) %>% str_remove("\\d+$"), .) %>% 
  split(., names(.)) %>% 
  lapply(unname) %>% 
  fromList %>% 
  t %>% 
  as.data.frame() %>% 
  rownames_to_column("description") %>% 
  mutate(description=factor(description, levels=set_name(combMat.loxAfr.ORA.increased.invert))) %>% 
  arrange(description) %>% 
  column_to_rownames("description") %>% 
  as.matrix()

sets.col <- c(
  'Activation of APC/C and APC/C:Cdc20 mediated degradation ...' = '#D95F02', 
'APC-Cdc20 mediated degradation of Nek2A' = '#D95F02', 
'APC:Cdc20 mediated degradation of cell cycle proteins pri...' = '#D95F02', 
'APC/C-mediated degradation of cell cycle proteins' = '#D95F02', 
'APC/C:Cdc20 mediated degradation of mitotic proteins' = '#D95F02', 
'Cdc20:Phospho-APC/C mediated degradation of Cyclin A' = '#D95F02', 
'Cell Cycle Checkpoints' = '#D95F02', 
'Inactivation of APC/C via direct inhibition of the APC/C ...' = '#D95F02', 
'Inhibition of the proteolytic activity of APC/C required ...' = '#D95F02', 
'Mitotic Anaphase' = '#D95F02', 
'Mitotic Metaphase and Anaphase' = '#D95F02', 
'Mitotic Spindle Checkpoint' = '#D95F02', 
'Regulation of APC/C activators between G1/S and early ana...' = '#D95F02', 
'Regulation of mitotic cell cycle' = '#D95F02', 
'Resolution of Sister Chromatid Cohesion' = '#D95F02', 
'Separation of Sister Chromatids' = '#D95F02', 
'Apoptosis' = '#7570B3', 
'Caspase activation via extrinsic apoptotic signalling pat...' = '#7570B3', 
'Intrinsic Pathway for Apoptosis' = '#7570B3', 
'Programmed Cell Death' = '#7570B3', 
'TP53 regulates transcription of additional cell cycle gen...' = '#E7298A', 
'TP53 Regulates Transcription of Cell Cycle Genes' = '#E7298A', 
'TP53 Regulates Metabolic Genes' = '#E7298A', 
'Regulation of PTEN gene transcription' = '#E7298A', 
'Ovarian tumor domain proteases' = '#E7298A' 
)[dimnames(count.sets)[[2]]]

UpSet.loxAfr.ORA.Genes <- grid.grabExpr(
  combMat.loxAfr.ORA.increased.invert %>% 
  UpSet(
    # Future manny note: it is vital that all the associated data/colors have the same row order as the sets (row) and combos (col) of the combMat
    comb_order = order(comb_size(combMat.loxAfr.ORA.increased.invert), decreasing=T),
    set_order = order(set_size(combMat.loxAfr.ORA.increased.invert), decreasing=T),
    # comb_col = color.genes.sets,
    height = unit(5, "cm"),
    width = unit(3.5, "cm"),
    top_annotation = HeatmapAnnotation(
      "Pathways shared\nbetween genes" = anno_barplot(comb_size(combMat.loxAfr.ORA.increased.invert),
                                                      ylim = c(0, max(comb_size(combMat.loxAfr.ORA.increased.invert))*1.1),
                                                      border = FALSE,
                                                      gp = gpar(fill="black", #fill = color.genes.sets,
                                                                fontsize = 7
                                                      ),
                                                      axis_param = list(
                                                        gp = gpar(fontsize= 7)
                                                      ),
                                                      height = unit(2, "cm")
      ),
      annotation_name_side = "left",
      annotation_name_rot = 90,
      annotation_name_gp = gpar(fontsize=8)
    ),
    left_annotation = rowAnnotation(
      "# Pathways" = anno_barplot(count.sets, 
                                  border = FALSE,
                                  axis_param = list(
                                    direction="reverse",
                                    labels_rot = 0,
                                    gp = gpar(fontsize= 7)
                                  ),
                                  gp=gpar(fill=sets.col %>% unname,
                                          fontsize= 7),
                                  width = unit(4, "cm")
      ),
      set_name = anno_text(set_name(combMat.loxAfr.ORA.increased.invert),
                           location = 0.5,
                           just = "center",
                           width = max_text_width(
                             set_name(combMat.loxAfr.ORA.increased.invert),
                             gp = gpar(
                               fontsize = 7
                             )
                           ) +
                             unit(1, "mm"),
                           gp = gpar(
                             fontsize = 7
                           )
      ),
      annotation_name_gp = gpar(fontsize=8)
    ), 
    right_annotation = NULL,
    show_row_names=F
  ) %>% 
  draw(
  )
  )

```

```{r Plot-loxAfrGene prep, include=F}
# loxAfr4.geneTPM.genesToHighlight.plus
    # "GTF2F1", #TP53 Transcriptional Regulation, Paenungulata    #Seems like a general TF...
    # "COX5A", #TP53 transcriptional regulation, PRobodiscea      #CytoC OXidase    
    # "STK11", #REgulation of TP53, Paenungulata                  #STK11 is a TSG
    # "BRD7", # TP53 regulation, Tethytheria                      #Required for p53-dependent oncogene-induced senescence
    # "TP53", #TP53 duplicates in Tethytheria
    # "SIAH1", # p53 pathway, Ubiquitin mediated proteolysis, Tethytheria   #E3 Ligase, induces apoptosis
    # "CASP9", # miRNA regulation of TP53, Cell Death probably, Extant Elephants  #Induces apoptosis, duh
    # "PRDX1", # TP53 regulates metabol, LoxAfr & MamPri          #Decomposes H2O2
    # "LAMTOR5", # Metabolism and TP53, 2x in Atlantogenata but 5x in oryAfe, 4-6 in modern elephants   #Antiapoptotic?
    # "COX20" # TP53 Regulates Metabolic Genes, LoxAfr            #CytoC OXidase

# Best to actually manually define categories

gene.pathway <- gene.color.table %>% 
  split(.,.$genes) %>% 
  sapply(. %>% pull("class") %>% unname)

gene.pathway.color <- gene.color.table %>% 
  distinct(class, color) %>% 
  split(.,.$class) %>% 
  lapply(. %>% pull("color") %>% unname)

loxAfr4.genes.fig <- loxAfr4.geneTPM[Gene %in% genesToHighlight.fig,
                                     ][,
                                       .(
                                         TPM=mean(TPM), 
                                         Copy = str_extract(Locus, "(?<=_)\\d+$") %>% as.numeric,
                                         Class=gene.pathway[Gene]
                                       ), 
                                       by=c("Gene","Locus")
                                       ][,
                                         NumberCopiesExpressed := .N,
                                         by="Gene"
                                         ][,
                                         Copy:= factor(Copy, levels=Copy %>% unique %>% sort)
                                       ][]

treedata.tp53.table <- ml.treedata.table %>% 
  select(parent, node, branch.length, label, all_of(genesToHighlight.fig)) %>% 
  full_join(., bodysize.tree.atlantogenata, by="label") %>%
  mutate_at(vars(all_of(genesToHighlight.fig)), . %>% sapply(. %>% ifelse(is.na(.), 1, .))) %>% 
  gene.increase.tree()

# # Look at colors with names using:
# tibble(a=names(highlight.gene.color), b=highlight.gene.color, x=rep(1:4, each=4), y=rep(1:4, times=4)) %>%
#   ggplot(aes(x=x,y=y, fill=a))+
#   scale_y_reverse()+
#   geom_tile()+
#   geom_text(aes(y=y+0.25, label=a))+
#   geom_text(aes(y=y-0.25, label=b))+
#   scale_fill_manual(values=ggsci::pal_d3("category20")(length(genesToHighlight.fig)) %>% set_names(genesToHighlight.fig))

p.treedata.loxAfr.dups.legendonly <- treedata.tp53.table %>% 
  mutate(label=str_replace_all(label, "\\.", " "), `Branch Length` = branch.length) %>% 
  as.treedata() %>% 
  ggtree(
    .,
    branch.length = "none",
    # continuous=T,
    aes(
      color=1/sqrt(`Branch Length`)
    ),
    size=1
  ) +  
  guides(
    color=guide_colorbar(title="Gene Duplication Rate", barheight = .5, barwidth = 5, direction = "horizontal",
                         title.position = "left", title.vjust = 0.5, title.hjust = 1.1,  title.theme = element_text(size = 8, face="bold", family="Helvetica"), label.hjust = 1)
  )+
  theme(
    legend.position="bottom",
    rect = element_blank(),
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", family="Helvetica", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    plot.title = element_text(size = 6, family="Helvetica", 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.box.background  = element_blank(),
    legend.title = element_text(size = 8, family="Helvetica", 
                                face = "bold", colour = "black"),
    legend.text = element_text(size = 6, family="Helvetica", 
                               face = "bold", colour = "black")
  )

p.treedata.loxAfr.dups <- treedata.tp53.table %>% 
  mutate(label=str_replace_all(label, "\\.", " "), `Branch Length` = branch.length) %>% 
  as.treedata() %>% 
  ggtree(
    .,
    branch.length = "none",
    # continuous=T,
    aes(
      color=1/sqrt(`Branch Length`)
    ),
    size=1
  ) +  
  # Row 1
  geom_point2(aes(x=branch, subset=(TP53>0), fill="TP53"),       pch=21, color="black", size=1.5, position = position_nudge(x=-0.3, y= 0.34)) +
  geom_point2(aes(x=branch, subset=(STK11>0), fill="STK11"),     pch=21, color="black", size=1.5, position = position_nudge(x=-0.1, y= 0.34)) +
  geom_point2(aes(x=branch, subset=(UBE2D1>0), fill="UBE2D1"),   pch=21, color="black", size=1.5, position = position_nudge(x= 0.1, y= 0.34)) +
  geom_point2(aes(x=branch, subset=(CNOT11>0), fill="CNOT11"),   pch=21, color="black", size=1.5, position = position_nudge(x= 0.3, y= 0.34)) +
  #  Row 2
  geom_point2(aes(x=branch, subset=(COX20>0), fill="COX20"),     pch=21, color="black", size=1.5, position = position_nudge(x=-0.3, y= 0.12)) +
  geom_point2(aes(x=branch, subset=(BRD7>0), fill="BRD7"),       pch=21, color="black", size=1.5, position = position_nudge(x=-0.1, y= 0.12)) +
  geom_point2(aes(x=branch, subset=(SOD1>0), fill="SOD1"),       pch=21, color="black", size=1.5, position = position_nudge(x= 0.1, y= 0.12)) +
  geom_point2(aes(x=branch, subset=(CASP9>0), fill="CASP9"),     pch=21, color="black", size=1.5, position = position_nudge(x= 0.3, y= 0.12)) +
  #  Row 3
  geom_point2(aes(x=branch, subset=(LAMTOR5>0), fill="LAMTOR5"), pch=21, color="black", size=1.5, position = position_nudge(x=-0.3, y= -0.12)) +
  geom_point2(aes(x=branch, subset=(MAD2L1>0), fill="MAD2L1"),   pch=21, color="black", size=1.5, position = position_nudge(x=-0.1, y= -0.12)) +
  geom_point2(aes(x=branch, subset=(LIF>0), fill="LIF"),         pch=21, color="black", size=1.5, position = position_nudge(x= 0.1, y= -0.12)) +
  geom_point2(aes(x=branch, subset=(CD14>0), fill="CD14"),       pch=21, color="black", size=1.5, position = position_nudge(x= 0.3, y= -0.12)) +
  #  Row 4
  geom_point2(aes(x=branch, subset=(PRDX1>0), fill="PRDX1"),     pch=21, color="black", size=1.5, position = position_nudge(x=-0.3, y= -0.34)) +
  geom_point2(aes(x=branch, subset=(BUB3>0), fill="BUB3"),       pch=21, color="black", size=1.5, position = position_nudge(x=-0.1, y= -0.34)) +
  geom_point2(aes(x=branch, subset=(MAPRE1>0), fill="MAPRE1"),   pch=21, color="black", size=1.5, position = position_nudge(x= 0.1, y= -0.34)) +
  geom_point2(aes(x=branch, subset=(HMGB2>0), fill="HMGB2"),     pch=21, color="black", size=1.5, position = position_nudge(x= 0.3, y= -0.34)) +
  geom_tiplab(color="black", fontface="bold", offset = 0.25, size=3, family="Helvetica") +
  scale_fill_manual(values = highlight.gene.color) + 
  guides(
    fill=guide_legend(title="Gene", 
                      title.position = "top", title.hjust = 0.5, title.vjust = 1, title.theme = element_text(size = 8, face="bold", family="Helvetica"), 
                      label.position = "bottom", nrow = 4, ncol=4, label.hjust = 0.5),
    color=guide_none()
  )+
  xlim(0,15) +
  ylim(0,17)+
  theme_tree()+
  theme(
    legend.position="bottom",
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", family="Helvetica", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    plot.title = element_text(size = 6, family="Helvetica", 
                              colour = "black", lineheight = 1, face = "bold"), 
    rect = element_blank(),
    legend.title = element_text(size = 8, family="Helvetica", 
                                face = "bold", colour = "black"),
    legend.text = element_text(size = 6, family="Helvetica", 
                               face = "bold", colour = "black"),
    legend.spacing.x = unit(-0.5, "mm"),
    legend.spacing.y = unit(-1, "mm")
  )

p.dups.genes.expression <- 
  loxAfr4.genes.fig[NumberCopiesExpressed>1] %>%
  ggplot(
    aes(
      x=TPM, 
      y=Copy, 
      color=Class
    )
  ) +
  geom_point(
    size=1.5
  )+
  geom_segment(
    aes(
      y=Copy,
      yend=Copy,
      x=0,
      xend=TPM
    )
  )+
  scale_x_sqrt()+
  scale_color_manual(values=gene.pathway.color) + 
  guides(color=guide_legend(ncol=3, nrow=2,#nrow = 2, ncol=3,
                            title= "Pathway", title.position = "left", vjust=0.5))+
  facet_grid(rows = vars(Gene), scales = "free_y", switch = "y") + 
  theme_pubclean()+
  labs(
    y="Copy",
    x="TPM (sqrt)"
  ) + 
  theme(
    legend.position="right",
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.y.left = element_text(face = "bold", size=7, angle = 0),
    legend.title = element_text(size = 8, face="bold"), 
    legend.text = element_text(size = 6),
    axis.title = element_text(size=8, face="bold"),
    axis.text.x = element_text(size=6, face="bold"),
    axis.text.y = element_text(size=8, face = "bold")
  )
```

(ref:Figure-LoxAfrExpressedDups) Duplications that occured in the African Savannah Elephant, *Loxodonta africana*, are enriched for TP53-related and other tumor suppressor processes. **A)** Upset plot of cancer-related Reactome pathways in *Loxodonta africana*, highlighting shared genes in each set, and the pathway class represented by the combinations. **B)** Inverted Upset plot from **A)**, showing the pathways shared by genes highlighted by WEBGESTALT in each pathway. **C)** Cladogram of *Atlantogenata*, colored by the rate of gene duplication estimated by maximum likelihood. Dots represent a duplication event of the color-coded genes. **D)** Gene expression levels of genes in **C)** which have two or more expressed duplicates.

```{r Figure-LoxAfrExpressedDups, fig.cap="(ref:Figure-LoxAfrExpressedDups)"}
areas <- c(area(1, 1, 2, 4),
           area(1, 5, 2, 7),
           area(3, 1, 5, 3),
           area(3, 4, 5, 5),
           area(3,6,4,7))

f4a <- UpSet.loxAfr.ORA.Pathway %>% wrap_elements()
f4b <- UpSet.loxAfr.ORA.Genes  %>% wrap_elements()
f4c <- p.treedata.loxAfr.dups + guides(fill=guide_none())
f4d <- facet_height(p.dups.genes.expression + guides(color=guide_none()), c(1,1,1, 1,1,1, 1,3,1, 1,1,2))
a <- cowplot::plot_grid(grid.grabExpr(cowplot::get_legend(p.treedata.loxAfr.dups) %>% grid.draw()), grid.grabExpr(cowplot::get_legend(p.treedata.loxAfr.dups.legendonly) %>% grid.draw()), grid.grabExpr(cowplot::get_legend(p.dups.genes.expression) %>% grid.draw()), nrow = 3, ncol=1)
f4 <-
  (f4a + f4b + f4c + f4d + a + patchwork::plot_layout(guides = "collect", design = areas)) #+ plot_annotation(tag_levels = 'A')

f4
ggsave("../output/pubFiles/Figures/Figure4.eps", plot=f4, width=12, height = 9, units="in", dpi=1200, device=grDevices::cairo_ps)
ggsave("../output/pubFiles/Figures/Figure4.png", plot=f4, width=12, height = 9, units="in", dpi=1200)

```

# TABLES

```{r Table-BodySize, eval=F}
t.bodysize.atlantogenata.withNodes <- st.Atlantogenata %>% 
  as_tibble() %>% 
  filter(!is.na(label)) %>%
  select(
    `Node`=label, 
    `Size (log(g))`=Median, 
    `95% CI (Low)` = `95%CI_Low`, 
    `95% CI (High)` = `95%CI_High`,  
    `Rate (sqrt)`=sqrt.rate
  ) %>%
  mutate_at(vars(-`Node`), . %>% as.numeric %>% round(2)) %>% 
  kable(
    ., 
    booktabs = TRUE, longtable = T,
    caption = 'Body Size and Confidence Intervals in Atlantogenata estimated using StableTraits.'
  ) %>% kable_styling(latex_options = "repeat_header")
t.bodysize.atlantogenata.withNodes
```

```{r Table-CancerSuscept, eval=F}
t.cancersucc.atl <- tree.Atlantogenata.table %>% 
  # filter(label %in% ml.treedata.table$label) %>%  
  # select(-parent, -node, -branch.length) %>%
  select(
    Node = label, 
    lnSize,
    `Est. Lifespan`=Lifespan,
    `K1` = K1,
    `K2` = K2,
    `Change in K` = CancerSucceptabilityChange,
    `log2 Change` = log2CancerSucceptabilityChange
  ) %>%
  mutate_at(
    vars(`K1`, `K2`, `Change in K`), 
    . %>% formatC(., format = "e", digits = 2)
            ) %>%
  mutate(
    `Est. Lifespan` = `Est. Lifespan` %>% round(2), 
    # `Change in K` = `Change in K` %>% round(2), 
    `log2 Change` = `log2 Change` %>% round(2)
  ) %>% 
  mutate(
    Ancestor = Node %>% 
    sapply(
        .,
        function(n, d=tree.Atlantogenata.table){
            p.node <- d %>% filter(label==n) %>% pull(parent) %>% unique
            d %>% filter(node == p.node) %>% pull(label) %>% unique %>% return
        }
    ) %>% str_replace_all("\\.", " ") %>% str_replace("Root", "Atlantogenata") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora")),
        Node = Node %>% str_replace_all("\\.", " ") %>% factor(., levels = c("Loxodontini", "Loxodonta africana","Loxodona", "Loxodonta cyclotis", "Palaeoloxodon antiquus", "Elephantidae", "Elephantina", "Elephas maximus", "Mammuthus", "Mammuthus primigenius", "Mammuthus columbi", "Proboscidea", "Mammut americanum", "Tethytheria", "Trichechus manatus", "Paenungulata", "Procavia capensis", "Pseudoungulata", "Orycteropus afer", "Macroscelidae", "Elephantulus edwardii", "Afrosoricida", "Chrysochloris asiatica", "Echinops telfairi", "Afrotheria", "Xenarthra", "Dasypus novemcinctus", "Choloepus hoffmanni", "Atlantogenata", "Afroinsectivora"))
  ) %>%
    arrange(Node) %>% 
  select(Node, `Est. Lifespan`, K1, K2, `Change in K`, `log2 Change`) %>% 
  kable(
    ., 
    booktabs = TRUE,
    caption = 'Estimated Cancer Succeptibilty for nodes in Atlantogenata'
  ) %>% 
  kable_styling(latex_options = c("scale_down"))

t.cancersucc.atl
```

```{r Table-GeneDup}
ggt.GenomeTable.withNote <- genome.table.forFig %>% 
  select(Species, `Common Name`, starts_with("Size"), starts_with("#"), starts_with("%"), starts_with("Mean")) %>% 
  arrange(Species) %T>% write_csv("../output/pubFiles/Table1.csv") %>% 
  kable(
    ., 
    # booktabs = TRUE,
    caption = 'Summary of duplications in Atlantogenata'
  ) %>% 
  kable_styling(latex_options = c("scale_down"))
ggt.GenomeTable.withNote
```

```{r Table-Top10Pathways-CombinedScore, eval=F}
ORA.top10 <- ORA.all %>% 
    filter(Direction=="increased", pValue<=0.05, FDR.Threshold=='1.00') %>% 
    arrange(desc(CombinedScore.freq)) %>% 
    select(-Model, -Database, -Direction, -EnrichRatio, -Genes, -RBHB, -is.Dynamic, -GeneList, -FDR.Threshold) %>% 
    distinct %>% 
    group_by(Ancestor, Node, Pathway, Class, FDR.sig.10, FDR, pValue) %>% 
    summarize_all(mean) %>% 
    ungroup() %>% 
    distinct %>% 
    group_by(Ancestor, Node) %>% 
    top_n(10) %>% 
    ungroup %>% 
    arrange(Node)

```

```{r Table-ORA-pcCancer-withSim}
t.ORA.sim <- ORA.pcCancer.withSim %>% 
  mutate(`p<0.05 in Simulation` = (`N.Genes` <= 157) %>% ifelse(., "Yes", "No")) %>% 
  select(Node=node, `N. Genes`= N.Genes, `N. Pathways` = N.Path, `% Cancer Pathways`= pc, `Simulated % Cancer Pathways` = pc.sim, `p<0.05 in Simulation`) %>% 
  arrange(Node) %T>% write_csv("../output/pubFiles/Table2.csv") %>% 
  kable(
    ., 
    # booktabs = TRUE,
    caption = 'Summary of Reactome Pathways in Atlantogenata'
  ) %>% 
  kable_styling(latex_options = c("scale_down"))
t.ORA.sim
```

```{r Table-SRAData, eval=T}
sra.data <- read_csv("../data/SraRunTable/SraRunTable.csv") %>% 
  select(Organism, Run, source_name, sample_name) %>% 
  filter(Organism %in% genomeTable.atlantogenata$Species) %>% 
     mutate(sample_name = sample_name %>% str_remove("Nine-banded Armadillo_ ID 09-86_")) %>% 
     mutate_all(. %>% str_replace_na("")) %>% 
     unite("Tissue", source_name, sample_name, sep="") %>% 
     filter(Tissue != "") %>% 
     group_by(Organism) %>% 
     summarise(Run = Run %>% unique %>% paste0(collapse=", "), 
               Tissue = Tissue %>% str_to_title() %>% unique %>% paste0(collapse=", ")) %>% 
     left_join(genomeTable.atlantogenata, by=c("Organism" = "Species")) %>% 
  select(Organism, `Common Name`, Genome, `SRA Acc.`=Run, Tissues=Tissue) #%>% 
  # mutate_all(. %>% str_wrap(width=50))
t.sra.data <- sra.data %T>% write_csv("../output/pubFiles/Table3.csv") %>%  
  kable(
    ., 
    booktabs = TRUE,
    caption = 'NCBI SRA datasets used in this study, along with key biological and genome information.'
  ) %>% 
  column_spec(4, width="10em") %>% 
  column_spec(5, width="10em") %>% 
  kable_styling(latex_options = c("scale_down"))

t.sra.data
```

```{r Table-PGLS-Lifespan, eval=T, results="asis"}
if(knitr::is_latex_output()){
  read_lines("../output/pubFiles/lifespan-PGLS.tex") %>% cat(sep = "\n")
} else if(knitr::is_html_output()){
  read_lines("../output/pubFiles/lifespan-PGLS.html") %>% cat(sep = "\n")
} else{
  read_lines("../output/pubFiles/lifespan-PGLS.html") %>% cat(sep = "\n")
}
```

```{r Table-Genomes, eval=T}
t.genomeTable.atlantogenata <- genomeTable.atlantogenata %>% 
    full_join(tribble(
        ~`Species`, ~`Citation`,
        "Choloepus hoffmanni", "\\citealp{DNAZoo}",
        "Chrysochloris asiatica", "GCA_000296735.1",
        "Dasypus novemcinctus", "GCA_000208655.2",
        "Echinops telfairi", "GCA_000313985.1",
        "Elephantulus edwardii", "GCA_000299155.1",
        "Elephas maximus", "\\citealp{Palkopoulou2018}",
        "Loxodonta africana", "ftp://ftp.broadinstitute.org/pub/assemblies/mammals/elephant/loxAfr4",
        "Loxodonta cyclotis", "\\citealp{Palkopoulou2018}",
        "Mammut americanum", "\\citealp{Palkopoulou2018}",
        "Mammuthus columbi", "\\citealp{Palkopoulou2018}",
        "Mammuthus primigenius", "\\citealp{Palkopoulou2015}",
        "Orycteropus afer", "\\citealp{DNAZoo}",
        "Palaeoloxodon antiquus", "\\citealp{Palkopoulou2018}",
        "Procavia capensis", "\\citealp{DNAZoo, 29mammals}",
        "Trichechus manatus latirostris", "\\citealp{DNAZoo, Foote2015}"
    ), by="Species") %>% 
  select(Species, `Common Name`, Genomes = Genome, `Highest Quality Genome` = BestGenome, Citation) %T>% write_csv("../output/pubFiles/Table5.csv") %>% 
  kable(
    ., 
    booktabs = TRUE,
    caption = 'Genomes used in this study.'
  ) %>% 
  kable_styling(latex_options = c("scale_down"))
t.genomeTable.atlantogenata
```

```{r Table-PathwayClasses}
t.reactome.class <- ORA.all[, .(Pathway=description, Class)] %>% 
  distinct %T>% write_csv("../output/pubFiles/Table5.csv") %>%  
  kable(
    ., 
    booktabs = TRUE,
    caption = 'Reactome Pathways Identified and Their Classifications'
  ) %>% 
  kable_styling(latex_options = c("scale_down"))
t.reactome.class
```

# SUPPLEMENTARY FIGURES

<!-- \beginsupplement -->

(ref:SFigure-ECNC) Estimated Copy Number by Coverage (ECNC) consolidates fragmented genes while accounting for missing domains in homologs. **A)** A single, contiguous gene homolog in a target genome with 100% query length coverage has an ECNC of 1.0. **B)** Two contiguous gene homologs, each with 100% query length coverage have an ECNC of 2.0. **C)** A single gene homolog, split across multiple scaffolds and contigs in a fragmented target genome; BLAT identifies each fragment as a single hit. Per nucleotide of query sequence, there is only one corresponding nucleotide over all the hits, thus the ECNC is 1.0. **D)** Two gene homologs, one fragmented and one contiguous. 100% of nucleotides in the query sequence are represented between all hits; however, every nucleotide in the query has two matching nucleotides in the target genome, thus the ECNC is 2.0. **E)** One true gene homolog in the target genome, plus multiple hits of a conserved domain that span 20% of the query sequence. While 100% of the query sequence is represented in total, 20% of the nucleotides have 4 hits. Thus, the ECNC for this gene is 1.45. **F)** Two real gene homologs; one hit is contiguous, one hit is fragmented in two, and the tail end of both sequences was not identified by BLAT due to sequence divergence. Only 75% of the query sequence was covered in total between the hits, but for that 75%, each nucleotide has two hits. As such, ECNC is equal to 2.0 for this gene.

```{r SFigure-ECNC, fig.cap="(ref:SFigure-ECNC)", eval=T}
fig.ECNC <- list(
    ecnc.onegene.onehit, #A
    ecnc.twogene.twohit, #B
    ecnc.onegene.threehits, #C 
    ecnc.twogene.fourhits, #D
    ecnc.onegene.manypartialhits, #E
    ecnc.twogene.partialhits #F
    ) %>% 
  lapply(
    function(p){
      p+theme_pubclean()
    }
  ) %>% 
  ggarrange(
    plotlist = .,
    ncol=3,
    nrow=2,
    labels = "AUTO"
    )
#A:"ECNC: One Contiguous Gene, One Hit"
#B: "ECNC: Two Contiguous Genes, Two Hits"
#C: "ECNC: One Fragmented Gene, Multiple Hits"
#D: "ECNC: Two Genes, Multiple Hits"
#E: "ECNC: One Real Gene, Multiple Hits on Conserved Domain"
#F: "ECNC: Two Real Genes, Unidentified Tails"
fig.ECNC
ggsave("../output/pubFiles/Figures/Supplement/FigureS1.eps", plot=fig.ECNC, width=6, height = 4, units="in", dpi=1200, device=grDevices::cairo_ps)
ggsave("../output/pubFiles/Figures/Supplement/FigureS1.png", plot=fig.ECNC, width=6, height = 4, units="in", dpi=1200)
```

```{r SFigure-GeneDup-BLScaled, fig.cap="Gene copy increases polarized along Atlantogenata, colored by ln(Body Size), with branch lengths equal to the change in gene copy number", eval=F}
p.ml.tree.increasing.geneCounts.withSize.withBL <- ml.tree.increasing.geneCounts.withSize.afrotheria %>% 
  ggtree(
    .,
    # branch.length = "none",
    #continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=1
    ) +
  scale_color_distiller(
    limits=c(size.scale.min, size.scale.max), 
    palette = "Blues", 
    direction = 1,
    guide=guide_none()
  ) +
  scale_fill_distiller(
    limits=c(size.scale.min, size.scale.max), 
    palette = "Blues", 
    direction = 1,
    guide = guide_colorbar(
      title="Body size (log kg)",
      title.theme =  element_text(size = 6, face = "bold", colour = "black"),
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  geom_tiplab(
    aes(label=Spc.short), 
    offset = 0.005, size=3,
    color="black", fontface="bold"
  )+
  # geom_label_repel(
  #   aes(label=yes.me(lnSize, joint.label, threshold=4)),
  #   size=3, direction = "x", nudge_x = -0.2, force=10, segment.color = "black",
  #   color="grey"
  # ) +
  # geom_label_repel(
  #   aes(label=not.me(lnSize, joint.label, threshold=4)),
  #   size=3, direction = "x",nudge_x = -0.05, force=10,
  #   color="black"
  # ) +
    geom_label_repel(
    aes(label=joint.label),
    size=3, force=1, direction="x", nudge_x=-0.1, segment.color = "black",
    color="black"
  ) +
  # geom_nodelab(
  #   aes(label=joint.label, x=-0.1), hjust = 0, geom="label", color="black", align=T, size=3
  # )+
  geom_tiplab(aes(label=N.Genes), geom = "label", size=3, color="black", hjust = 1) + 
  # geom_tiplab(aes(label=yes.me(lnSize, N.Genes, threshold=4)), geom = "label", size=3, color="grey") + 
  theme_tree() +  
  geom_treescale(
    #x=0, y=14, width = 1, 
   offset = -0.6, linesize=2
  )+
  xlim(-0.15,0.35) +
  xlab("Rate of Change, Body Size") +
  theme(
    legend.position="bottom",
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text.x = element_blank(),
    axis.title = element_text(size = 8, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 8, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 10, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 8, 
                               face = "plain", colour = "black")
  )
p.ml.tree.increasing.geneCounts.withSize.withBL
```

```{r Plot Body Size, include=F, eval=F}
p.tree <- tree.Atlantogenata %>% 
  ggtree(aes(color=lnSize), continuous=T, branch.length = "sqrt.rate") +
  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))")) +
  geom_highlight(node = bsize.ancAtlantogenata, fill="#EFD28D") +
  xlim(-100, NA) +
  theme_tree2() +
  theme(legend.position="right") + 
  xlab("Rate of Change, Body Size") + 
  theme(
    legend.position="right",
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text.x = element_text(size = 4, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 6, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 6, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 6, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black")
  ) + 
  guides(color=guide_none())

p.tree.atlantogenata <-  
  tree.Atlantogenata %>% 
  ggtree(
    aes(color=lnSize),
    branch.length = "sqrt.rate",
    continuous=T
  ) + 
  geom_tiplab(aes(image=highlight, size=I(0.1)), geom="phylopic",
              show_guide=F, align = F, linesize = 0, offset= 100
              ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Proboscidea, label="Proboscidea", 
                  fontsize = 2, align = T, offset = 50, angle=0, barsize = 1,
                  offset.text = 20,
                  color = bsize.Atlantogenata.table %>% filter(label=="Proboscidea") %>% pull(lnSize) %>% get.size.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Tethytheria, label="\nTethytheria", 
                  fontsize = 2, align = T, offset = 25, angle=0, barsize = 1, 
                  offset.text = 20, color = bsize.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(lnSize) %>% get.size.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Paenungulata, label="\n\nPaenungulata", 
                  fontsize = 2, align = T, angle=0, barsize = 1, offset.text = 20,
                  color = bsize.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(lnSize) %>% get.size.color
                  ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Afrosoricida, label="Afrosoricida",
                  fontsize = 2, align = T, angle=0, barsize = 1,
                  color=bsize.Atlantogenata.table %>% filter(label=="Afrosoricida") %>% pull(lnSize) %>% get.size.color
                  ) +
    geom_cladelabel(node=bsize.Atlantogenata.nodes$Macroscelidae, 
                    label="Macroscelidae", 
                    fontsize = 2, align = T, angle=0, barsize = 1,
                    color=bsize.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(lnSize) %>% get.size.color
                    ) +
  geom_cladelabel(node=bsize.Atlantogenata.nodes$Xenarthra, label="Xenarthra", 
                  fontsize = 2, align = T, angle=0, barsize = 1,
                  color=bsize.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(lnSize) %>% get.size.color
                  ) +
  scale_color_viridis_c(limits=c(size.scale.min, size.scale.max), 
                        guide = guide_colourbar(title="Body size (ln(g))"))+
  xlim(0,1750) + 
  theme_tree2() +
  xlab("Rate of Change, (Body Size, sqrt)") + 
  theme(
    legend.position="bottom",
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text.x = element_text(size = 4, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 6, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 6, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 6, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black")
  ) + 
  guides(color=guide_none())

```

(ref:SFigure-Bodysize) Body sizes rapidly and frequently expand in Eutherians, especially in Atlantogenata. **A)** Tree of Eutherian species, colored by ln(Body Size) and with branch lengths set to the rate of change in body sizes, normalized by the square root of the root branch. Atlantogenata is highlighted at the bottom. **B)** Zoom-in of (**A**) on *Atlantogenata*. Silhouettes for the African Elephant, West Indian Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole, Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their extant body sizes, while clade labels are colored based on the common ancestor's estimated body size. **C)** Confidence interval plot for representative species and ancestral nodes.

```{r SFigure-BodySize, fig.cap="(ref:SFigure-Bodysize)", out.width="6in", fig.asp=1, eval=F}
fig.bodysize.a <- p.tree
fig.bodysize.b <- p.tree.atlantogenata
fig.bodysize.c <- p.bodysize.atlantogenata

fig.bodysize.ab <- ggarrange(
  fig.bodysize.a, fig.bodysize.b, 
  nrow=1, ncol = 2, 
  labels="AUTO",
  widths = c(1,2),
  font.label = list(size=6, face="bold")
  )

fig.bodysize <-
  ggarrange(
  ggplotify::as.ggplot(fig.bodysize.ab),
  fig.bodysize.c,
  nrow=2, ncol=1,
  labels = c("", "C"),
  font.label = list(size=6, face="bold")
)
fig.bodysize
```

(ref:SFigure-CancerSuscept) Cancer succecptibility across *Atlantogenata.* Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.

```{r SFigure-CancerSuscept, fig.cap="(ref:SFigure-CancerSuscept)", out.width="6in", fig.asp=1, eval=F}
p.tree.Atlantogenata.table
```

(ref:SFigure-TimeTree-Eutheria) The time-calibrated Eutherian tree used as input for StableTraits, from Bininda-Emonds *et al.* [@Bininda-Emonds2007] and Puttick and Thomas [@PuttickAndThomas2015].

```{r SFigure-TimeTree-Eutheria, fig.cap="(ref:SFigure-TimeTree-Eutheria)", warning=F, eval=T}
p.time.tree.Eutheria <-
  time.tree %>%
  ggtree(
    size=0.2,
    layout="circular"
  ) %>%
  ggtree::revts() +
    geom_highlight(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), fill="goldenrod") + geom_cladelabel(node = MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), label="Atlantogenata", align = F, offset = 2, offset.text = 2, geom = "label", fill="goldenrod", fontsize = 2) +
  geom_highlight(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill="aquamarine") + geom_cladelabel(node = MRCA(time.tree, "Myotis.lucifugus", "Pteropus.vampyrus"), fill = "aquamarine", label="Chiroptera", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label", fontsize = 2) +
    geom_highlight(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1") + geom_cladelabel(node = MRCA(time.tree, "Cavia.porcellus", "Mus.musculus"), fill="firebrick1", label="Rodentia", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label", fontsize = 2) +
  geom_highlight(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue") + geom_cladelabel(node = MRCA(time.tree, "Gorilla.gorilla", "Otolemur.crassicaudatus"), fill="deepskyblue", label="Primata", align = F, offset = 2, offset.text = 3, hjust = 0, geom = "label", fontsize = 2) +
  geom_highlight(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1") + geom_cladelabel(node = MRCA(time.tree, "Balaena.mysticetus", "Tursiops.truncatus"), fill="darkseagreen1", label="Cetacea", align = F, offset = 2, offset.text = 2, hjust = 0, geom = "label", fontsize = 2) +
  theme_tree() +
  theme(plot.margin = margin(5,15,5,5))
```

```{r SFigure-TimeTree-Atlantogenata, fig.cap="The time-calibrated Atlantogenatan tree used as input for StableTraits, from Puttick and Thomas", warning=F, eval=T}
time.tree.atlantogenata <- time.tree %>%
  tree_subset(., MRCA(time.tree, "Loxodonta.africana", "Dasypus.novemcinctus"), levels_back = 0)

p.time.tree.Atlantogenata <-
  time.tree.atlantogenata %>% 
  # as_tibble %>%
  # mutate(time = branch.length) %>%
  # select(-parent, -node, -branch.length) %>%
  # full_join(., bsize.tree %>% as_tibble, by="label") %>%
  # as.treedata %>%
  ggtree(
    size=0.5,
    layout="circular"
  ) %>%
  ggtree::revts() +
  geom_highlight(node = MRCA(time.tree.atlantogenata, "Loxodonta.africana", "Procavia.capensis"), fill="goldenrod") + 
    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Loxodonta.africana", "Procavia.capensis"), label="Paenungulata", align = F, offset = 2, offset.text = 10, geom = "label", fill="goldenrod", fontsize = 2) +
  geom_highlight(node = MRCA(time.tree.atlantogenata, "Elephantulus.edwardii", "Chambius.kasserinensis"), fill="firebrick1") + 
    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Elephantulus.edwardii", "Chambius.kasserinensis"), fill="firebrick1", label="Macroscelidae", align = F, offset = 2, offset.text = 3, hjust = 1, geom = "label", fontsize = 2) +
  geom_highlight(node = MRCA(time.tree.atlantogenata, "Chrysochloris.asiatica", "Echinops.telfairi"), fill="deepskyblue") + 
    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Chrysochloris.asiatica", "Echinops.telfairi"), fill="deepskyblue", label="Afrosoricida", align = F, offset = 2, offset.text = 15, hjust = 0, geom = "label", fontsize = 2) +
  geom_highlight(node = MRCA(time.tree.atlantogenata, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), fill="darkseagreen1") + 
    geom_cladelabel(node = MRCA(time.tree.atlantogenata, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), fill="darkseagreen1", label="Xenarthra", align = F, offset = 2, offset.text = 7, hjust = 0, geom = "label", fontsize = 2) +
      geom_point2(aes(subset=(label %in% c("Choloepus.hoffmanni", "Chrysochloris.asiatica", "Dasypus.novemcinctus", "Echinops.telfairi", "Elephantulus.edwardii", "Elephas.maximus", "Loxodonta.africana", "Loxodonta.cyclotis", "Mammut.americanum", "Mammuthus.columbi", "Mammuthus.primigenius", "Orycteropus.afer", "Palaeoloxodon.antiquus", "Procavia.capensis", "Trichechus.manatus"))), size=1, pch = 16, color="purple")+
  theme_tree()+
  theme(plot.margin = margin(5,5,5,5))

p.time.tree <- ggarrange(p.time.tree.Eutheria, p.time.tree.Atlantogenata, 
                         nrow=1, ncol=2, labels = "AUTO")
ggsave("../output/pubFiles/Figures/Supplement/FigureS2.eps", plot=p.time.tree, width=6, height = 6, units="in", dpi=1200, device=grDevices::cairo_ps)
ggsave("../output/pubFiles/Figures/Supplement/FigureS2.png", plot=p.time.tree, width=6, height = 6, units="in", dpi=1200)
```

(ref:SFigure-DoubleTree-Bodysize-CancerSuscept-Genomes) Back-to-back phylogeny for extant *Atlantogenatan* species, scaled by log-fold increases in body mass (IBM, left) and log_{2}-fold reduction in intrinsic cancer risk (RICR, right).

```{r SFigure-DoubleTree-Bodysize-CancerSuscept-Genomes, fig.cap="(ref:SFigure-DoubleTree-Bodysize-CancerSuscept-Genomes)", eval=F}

### Body Size ###
bsize.Atlantogenata.table <- bsize.Atlantogenata %>% as_tibble
bsize.Atlantogenata.table[which(bsize.Atlantogenata.table$node == MRCA(bsize.Atlantogenata, "Pseudoungulata", "Elephantulus.edwardii")),]$label <- "Afroinsectivora"
bsize.Atlantogenata <- bsize.Atlantogenata.table %>% as.treedata()

lnsize.bsize.Atlantogenata <- bsize.Atlantogenata %>% as_tibble() %>% select(label, Species.strict, lnSize)
bsize.Atlantogenata.genomes <-  bsize.Atlantogenata@phylo %>% 
  treeio::drop.tip(bsize.Atlantogenata@phylo$tip.label[!bsize.Atlantogenata@phylo$tip.label %in% genomeTable.atlantogenata$label]) %>% as_tibble
bsize.Atlantogenata.genomes <- left_join(bsize.Atlantogenata.genomes, lnsize.bsize.Atlantogenata) %>% as.treedata

p.tree.atlantogenata.genomes <- bsize.Atlantogenata.genomes %>% 
  ggtree(
    aes(color=as.numeric(lnSize))
  )

### Cancer Suscept ###
tree.Atlantogenata.table.table <- tree.Atlantogenata.table %>% as_tibble
tree.Atlantogenata.table.table[which(tree.Atlantogenata.table.table$node == MRCA(tree.Atlantogenata.table, "Pseudoungulata", "Elephantulus edwardii")),]$label <- "Afroinsectivora"
tree.Atlantogenata.table <- tree.Atlantogenata.table.table %>% as.treedata()

cancerSucceptibility.Atlantogenata.genomes <- tree.Atlantogenata.table %>% 
  as_tibble() %>%  
  mutate(abslog2CancerSucceptabilityChange = log2CancerSucceptabilityChange %>% abs(), 
         branch.length=abslog2CancerSucceptabilityChange) %>% 
  select(label, abslog2CancerSucceptabilityChange, log2CancerSucceptabilityChange, lnK1)
tree.Atlantogenata.table.genomes <-  tree.Atlantogenata.table %>% 
  as_tibble %>% 
  filter(!is.na(node)) %>%  
  mutate(branch.length = log2CancerSucceptabilityChange %>% abs()) %>% 
  as.treedata %>% 
  as.phylo() %>% 
  treeio::drop.tip(tree.Atlantogenata.table@phylo$tip.label[!tree.Atlantogenata.table@phylo$tip.label %in% genomeTable.atlantogenata$Species.strict])  %>% as_tibble
tree.Atlantogenata.table.genomes <- left_join(tree.Atlantogenata.table.genomes, cancerSucceptibility.Atlantogenata.genomes) %>% as.treedata

p.tree.Atlantogenata.table.genomes <- tree.Atlantogenata.table.genomes %>% 
  ggtree(
    aes(color=as.numeric(abslog2CancerSucceptabilityChange))
  )

### Double Tree ###

d.dbltree.1 <- p.tree.atlantogenata.genomes$data
d.dbltree.2 <- p.tree.Atlantogenata.table.genomes$data

d.dbltree.2$x <- max(d.dbltree.2$x) - d.dbltree.2$x + max(d.dbltree.1$x) + 500

p.dbltree <-
  bsize.Atlantogenata.genomes %>% 
  ggtree(., aes(color=lnSize), size=1) +
  scale_color_distiller("Body size (log)", palette = "Blues", direction = -1) + 
  geom_nodepoint(size=2) + 
  geom_tippoint(size=2) +
  geom_text_repel(aes(label = label %>% str_replace_all("\\.", " ")), direction = "y", force = 0.2, segment.size = NA,
              x=(min(d.dbltree.2$x)+max(d.dbltree.1$x))/2, color="black", hjust = 0.5) +
  new_scale_color() + 
  geom_tree(data=d.dbltree.2, aes(color=log2CancerSucceptabilityChange), inherit.aes = F, size=1) + 
  geom_nodepoint(data=d.dbltree.2, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) + 
  geom_tippoint(data=d.dbltree.2, aes(color=log2CancerSucceptabilityChange, x=x, y=y), inherit.aes = F, size=2) +
  # scale_color_viridis_c(option = "D") + 
  scale_color_distiller("Cancer Risk (log2)", palette = "Reds", direction = 1) + 
    theme(
      plot.background = element_rect(fill="grey"),
      panel.background = element_rect(fill="grey"),
      legend.background = element_rect(fill="grey")
    )

p.dbltree
```

(ref:SFigure-GenomeCor) Correlations between genome quality metrics and ECNC metrics. Gene copy number metrics, and the genome quality metrics most strongly associated with them, are highlighted in red.

```{r SFigure-GenomeCor, fig.cap="(ref:SFigure-GenomeCor)", fig.asp=1, out.width="6in", out.height="6in", fig.align="center"}
p.cor.GCT.CEGMA <-
  cor.GCT.CEGMA %>%
  ggplot(
    aes(x=Var1, y=Var2, fill = value)
  )+
  geom_tile(color = "white")+
  theme_minimal()+
  coord_fixed()+
  scale_x_discrete(
    position = "top"
  ) +
  scale_y_discrete(position = "right") +
  scale_fill_viridis_c(breaks = c(-1,-0.5, 0, 0.5, 1),labels = c(-1,-0.5, 0, 0.5, 1)) +
  theme_pubclean() +
  theme(
    axis.text.x = element_text(
      angle = -270, vjust = 0,
      size = highlightTextSize(cor.GCT.CEGMA$Var1, cor.toHighlight, sizes = c(3,3)), 
      color = highlightTextColor(cor.GCT.CEGMA$Var1, cor.toHighlight, color = c("black", "red")), 
      hjust = 0
    ),
    axis.text.y = element_text(
      size = highlightTextSize(cor.GCT.CEGMA$Var2, cor.toHighlight, sizes = c(3,3)), 
      color = highlightTextColor(cor.GCT.CEGMA$Var2, cor.toHighlight, color = c("black", "red"))
    ),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 1),
    legend.position = c(0.4,0.7),
    legend.direction = "horizontal",
    legend.text = element_text(size=3),
    legend.title =  element_text(size=5)
  ) +
  guides(
    fill = guide_colorbar(
      title="Pearson\nCorrelation",
      barwidth = 2, barheight = 0.5,
      title.position = "top", title.hjust = 0.5,
      draw.llim = T
    )
  )

p.cor.GCT.CEGMA

ggsave("../output/pubFiles/Figures/Supplement/FigureS3.eps", plot=p.cor.GCT.CEGMA, width=6, height = 4, units="in", dpi=1200, device=grDevices::cairo_ps)
ggsave("../output/pubFiles/Figures/Supplement/FigureS3.png", plot=p.cor.GCT.CEGMA, width=6, height = 4, units="in", dpi=1200)
```

# References {#references .unnumbered}

### 2020-09-13 code/generalFunctions.R

@@ -434,21 +434,6 @@ binTableFromList <- function (input) {
   data[, Gene:=elements][]
 }
 
-fromList <- function (input) {
-  # Same as original fromList()...
-  elements <- unique(unlist(input))
-  data <- unlist(lapply(input, function(x) {
-    x <- as.vector(match(elements, x))
-  }))
-  data[is.na(data)] <- as.integer(0)
-  data[data != 0] <- as.integer(1)
-  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
-  data <- data[which(rowSums(data) != 0), ]
-  names(data) <- names(input)
-  # ... Except now it conserves your original value names!
-  row.names(data) <- elements
-  return(data)
-}
 
 phylopic_uid_item_safer <- function (name) {
   x <- gsub("[^a-zA-Z]+", "+", name)
@@ -487,30 +472,24 @@ lineage.pathway.fdr.table <- function(path="output/ORA"){
       Database = ORAs %>% sapply(str_extract, "(?<=pathway_)[A-Za-z]+(?=-)"),
       FDR.Threshold = ORAs %>% sapply(str_extract, "(?<=FDR_)[0-9]\\.[0-9]+(?=/)"),
       Direction.lineage = ORAs %>% sapply(str_extract, "(?<=Project_).+?(?=/)"),
-      Pathway.EnrichRatio.FDR.pval.Genes.Size = ORAs %>% sapply(
+      Pathway.EnrichRatio.FDR.Genes.Size = ORAs %>% sapply(
         ., . %>% 
           read_tsv %>% 
-          select(description, enrichmentRatio, FDR, pValue, userId, size) %>% 
+          select(description, enrichmentRatio, FDR, userId, size) %>% 
           mutate_all(. %>% str_replace_all(",|;", ":")) %>% 
-          unite("PEF", description, enrichmentRatio, FDR, pValue, userId, size, sep = ",") %>% 
+          unite("PEF", description, enrichmentRatio, FDR, userId, size, sep = ",") %>% 
           pull(PEF) %>% 
           paste0(collapse=";")),
       fname = ORAs
     ) %>% 
     separate(Direction.lineage, c("Direction", "Ancestor", "to", "Node"), sep="_", extra = "merge") %>% 
[…]
